### 2020-01-27 wuhan.tex
putting all in one directory so that jonathan can MAKE

\documentclass[12pt]{article}
\usepackage[top=1in,left=1in, right = 1in, footskip=1in]{geometry}

\usepackage{graphicx}
%\usepackage{adjustbox}

\newcommand{\eref}[1]{(\ref{eq:#1})}
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\Fref}[1]{Fig.~\ref{fig:#1}}
\newcommand{\sref}[1]{Sec.~\ref{#1}}
\newcommand{\frange}[2]{Fig.~\ref{fig:#1}--\ref{fig:#2}}
\newcommand{\tref}[1]{Table~\ref{tab:#1}}
\newcommand{\tlab}[1]{\label{tab:#1}}
\newcommand{\seminar}{SE\mbox{$^m$}I\mbox{$^n$}R}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

%\usepackage{lineno}
%\linenumbers

\usepackage[pdfencoding=auto, psdextra]{hyperref}

\usepackage{natbib}
\bibliographystyle{chicago}
\date{\today}

\usepackage{xspace}
\newcommand*{\ie}{i.e.\@\xspace}

\usepackage{color}

\newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}} 
\newcommand{\Ro}{\Rx{0}}
\newcommand{\RR}{\ensuremath{{\mathcal R}}}
\newcommand{\Rhat}{\ensuremath{{\hat\RR}}}
\newcommand{\tsub}[2]{#1_{{\textrm{\tiny #2}}}}

\newcommand{\comment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}

%% \newcommand{\rev}[1]{\comment{red}{REV}{#1}}
\newcommand{\rev}[1]{}

\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\new}[1]{\textcolor{blue}{#1}}

\begin{document}

\begin{flushleft}{
	\Large
	\textbf\newline{
		Assessing uncertainties associated with early estimates of the basic reproduction number during the novel coronavirus (2019-nCoV) outbreak
	}
}
\newline
\\
Sang Woo Park\textsuperscript{1}
Jonathan Dushoff\textsuperscript{2,3}
\\

\bigskip
\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, USA
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, Ontario, Canada
\\
\textbf{3} Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Ontario, Canada
\\
\bigskip

% *swp2@princeton.edu
\end{flushleft}

\section*{Abstract}

\pagebreak

As the novel coronavirus (2019-nCoV) continues to spread,
researchers are rushing to publish their estimates of the basic 
reproductive number $\mathcal R_0$.
The basic reproductive number (i.e., the 
average number of secondary cases generated 
by a primary case in a fully susceptible population) is of particular interest 
because it allows prediction about the final size of an epidemic.
While their efforts are valuable, their analyses rely on several
assumptions that could immediately affect their estimates of $\mathcal R_0$ and
the associated uncertainties in their estimates.
Here, we illustrate the degree to which their assumptions can affect
their estimates and stress few principles that needs to be taken into
consideration.

\section*{Early estimates of $\mathcal R_0$}

Early in an outbreak, $\mathcal R_0$ cannot be estimated directly,
whereas the exponential growth rate $r$ can be estimated reliably.
Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of 
generation intervals (i.e., the time between when a person become 
infected and that person infects another person), the basic reproduction
number can be estimated via the Euler-Lotka equation:
\begin{equation}
1/\mathcal R_0 = \int \exp(-r\tau) g(\tau) d\tau.
\end{equation}
Therefore, early estimates of $\mathcal R_0$ must 
necessarily depend on the assumptions about $r$ and $g(\tau)$.

\begin{figure}[t]
\includegraphics[width=\textwidth]{../fig/comparison.pdf}
\caption{
Early estimates of $\mathcal R_0$ and associated assumptions about $r$ and $g(\tau)$.
}
\end{figure}

We use the gamma approximation framework to demonstrate
how assumptions about $r$ and $g(\tau)$
affects $\mathcal R_0$.
Assuming that generation intervals follow a gamma distribution 
with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
we have
\begin{equation}
\mathcal R_0 = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
\end{equation}
This equation demonstrates that generation-interval distributions
with a larger mean (higher $\bar{G}$) or less variability (lower $\kappa$)
will give a higher estimate of $\mathcal R_0$.
Likewise, strong assumptions about $\bar{G}$ and $\kappa$ will give
estimates of $\mathcal R_0$ with narrow confidence intervals.

Figure 1 clearly demonstrates the lack of uncertainties in the underlying parameters
of the current $\mathcal R_0$ estimates.
In particular, no studies properly propagate uncertainties associated with the
amount of variability in generation intervals, which can have 
large effects on the estimates of $\mathcal R_0$
For example, when the relative mean generation interval is long ($r\bar{G}\approx 2$ based on the estimates by \cite{zhaoncov}),
varying $\kappa$ from 0 to 1 reduces the estimate of $\mathcal R_0$ by more than a two fold.

To assess the amount of uncertainties in the estimate of $\mathcal R_0$,
we come up with distributions that span over the assumptions made by other people:
\begin{equation}
\begin{aligned}
r &\sim \mathrm{Gamma}(\alpha=5, \beta=5/0.15)\\
\bar{G} &\sim \mathrm{Gamma}(\alpha=10, \beta=10/8.5)\\
\kappa &\sim \mathrm{Gamma}(\alpha=10, \beta=10/0.35)\\
\end{aligned}
\end{equation}
These give median estimate of $\mathcal R_0$ of 2.37 with 95\% qunatile of 1.26--9.01.

\begin{table}[!th]
\begin{center}
\begin{tabular}{l|c|c|c|c}
Study & $\mathcal R_0$ & $r$ ($\textrm{days}^{-1}$) & $\bar G$ (days) & $\kappa$ \\
\hline
\cite{imaincov} & 2.5 (1.5--3.5) & 0.14 (0.05--0.21) & 8.4 & unspecified$^\ast$ \\
\hline
\cite{majumderncov} & 2.0--3.3 & 0.07--0.20 & 6--10 & 0 \\
\hline
\cite{readncov} & 3.8 (3.6--4.0) & 0.25 (0.25--0.26)$^\dagger$ & 7.6 & 0.5 \\
\hline
\cite{riouncov} & 2.2 (1.4--3.8) & 0.09 (0.03--0.27) & 7--14 & 0.5\\
\hline
\cite{zhaoncov} & 5.47 (4.16--7.1)$^\ddagger$ & 0.25 (0.21--0.30) & 7.6--8.4 & 0.2
\end{tabular}
\end{center}
\caption{
\textbf{Parameter estimates and assumptions}
Exponential growth rate $r$ is calculated by using the following formula: $r = (\mathcal R_0^\kappa - 1)/(\kappa \bar{G})$.
$^\ast$We assume $\kappa = 0.5$ to calculate $r$.
$^\dagger$Exponential growth rate $r$ is calculated by using the exact formula for the SEIR model \citep{ma2014estimating}.
$^\ddagger$Assuming no increase in the reporting rate.
}
\end{table}

\bibliography{wuhan}

\end{document}

### 2020-01-27 analysis/sample.R

@@ -1,23 +1,0 @@
-sample_imai <- function(n=1) {
-  # qgamma(0.025, 20, 20/2.5)
-  # qgamma(0.975, 20, 20/2.5)
-  dplyr::data_frame(
-    gbar=8.4,
-    kappa=0.5,
-    R=rgamma(n, 20, rate=20/2.5),
-    r=(R^kappa-1)/(kappa*gbar),
-    rho=r*gbar
-  )
-}
-
-sample_majumder <- function(n=1) {
-  # qgamma(0.025, 20, 20/2.5)
-  # qgamma(0.975, 20, 20/2.5)
-  dplyr::data_frame(
-    gbar=runif(n, min=6,max=10),
-    kappa=0,
-    r=log(2)/gbar,
-    R=exp(r*gbar),
-    rho=r*gbar
-  )
-}

### 2020-01-27 calculate_r.R

@@ -1,0 +1,1 @@
+source("sample.R")

### 2020-01-27 comparison.R

@@ -5,11 +5,10 @@ source("../data/estimate.R")
 g0 <- ggplot(Rdata) +
   geom_point(aes(est, study)) +
   geom_segment(aes(lwr, study, xend=upr, yend=study)) +
-  scale_x_continuous("Exponential growth rate") +
+  scale_x_continuous("Basic reproductive number") +
   theme(
     axis.title.y = element_blank()
   )
-
 
 g1 <- ggplot(rdata) +
   geom_point(aes(est, study)) +

### 2020-01-27 data/estimate.R

@@ -1,41 +1,0 @@
-Rdata <- data.frame(
-  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder et al. (2020)"),
-  est=c(2.5, 2.2, 3.8, 5.47, NA ),
-  lwr=c(1.5, 1.4, 3.6, 4.16, 2),
-  upr=c(3.5, 3.8, 4, 7.1, 3.3)
-)
-
-rdata <- data.frame(
-  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder et al. (2020)"),
-  est=c((2.5^0.5 - 1)/(0.5 * 8.4), 
-        (2.2^0.5 - 1)/(0.5 * 10.5),
-        1/2*(-(1/4+1/3.6)+sqrt((1/4-1/3.6)^2+4*1.07*1/4)),
-        (5.47^0.2 - 1)/(0.2 * 8),
-        NA),
-  lwr=c((1.5^0.5 - 1)/(0.5 * 8.4),
-        (1.4^0.5 - 1)/(0.5 * 14),
-        1/2*(-(1/4+1/3.6)+sqrt((1/4-1/3.6)^2+4*1.06*1/4)),
-        (4.16^0.2 - 1)/(0.2 * 8),
-        log(2)/10),
-  upr=c((3.5^0.5 - 1)/(0.5 * 8.4),
-        (3.8^0.5 - 1)/(0.5 * 7),
-        1/2*(-(1/4+1/3.6)+sqrt((1/4-1/3.6)^2+4*1.09*1/4)),
-        (7.1^0.2 - 1)/(0.2 * 8),
-        log(3.3)/6)
-)
-
-gdata <- data.frame(
-  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder et al. (2020)"),
-  est=c(8.4, NA, 7.6, NA, NA ),
-  lwr=c(8.4, 7, 7.6, 7.6, 6),
-  upr=c(8.4, 14, 7.6, 8.4, 10)
-)
-
-kappadata <- data.frame(
-  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
[…]

### 2020-01-27 doc/seir.R

@@ -1,10 +1,0 @@
-gamma <- 1/4
-mu <- 1/3.6
-
-f <- function(x) {
-  (gamma * mu/(gamma-mu) * (exp(-mu * x) - exp(-gamma * x))) * (x - 7.6)^2
-}
-
-integrate(f, 0, 1000)
-
-28.96/7.6^2

### 2020-01-27 published_estimate.R

@@ -1,0 +1,21 @@
+Rdata <- data.frame(
+  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
+          "Zhao et al. (2020)", "Majumder et al. (2020)", "Liu et al. (2020)"),
+  est=c(2.5, 2.2, 3.8, 5.47, NA, 2.92),
+  lwr=c(1.5, 1.4, 3.6, 4.16, 2, 2.28),
+  upr=c(3.5, 3.8, 4, 7.1, 3.1, 3.67)
+)
+
+gdata <- data.frame(
+  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
+          "Zhao et al. (2020)", "Majumder et al. (2020)", "Liu et al. (2020)"),
+  est=c(8.4, NA, 7.6, NA, NA, 8.4),
+  lwr=c(8.4, 7, 7.6, 7.6, 6, 8.4),
+  upr=c(8.4, 14, 7.6, 8.4, 10, 8.4)
+)
+
+kappadata <- data.frame(
+  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
+          "Zhao et al. (2020)", "Majumder et al. (2020)", "Liu et al. (2020)"),
+  est=c(NA, 0.5, 0.5, 0.2, 0, 0.2)
+)

### 2020-01-27 sample.R

@@ -1,0 +1,74 @@
+sample_imai <- function(n=1) {
+  # qgamma(0.025, 20, 20/2.5)
+  # qgamma(0.975, 20, 20/2.5)
+  dplyr::data_frame(
+    gbar=8.4,
+    kappa=0.5,
+    R=rgamma(n, 20, rate=20/2.5),
+    r=(R^kappa-1)/(kappa*gbar),
+    rho=r*gbar
+  )
+}
+
+sample_majumder <- function(n=1) {
+  dplyr::data_frame(
+    gbar=runif(n, min=6,max=10),
+    kappa=0,
+    r=runif(n, max=log(2)/6, min=log(3.1)/10),
+    R=exp(r*gbar),
+    rho=r*gbar
+  )
+}
+
+sample_read <- function(n=1) {
+  # qgamma(0.025, 1400, 1400/3.8)
+  # qgamma(0.975, 1400, 1400/3.8)
+  
+  dplyr::data_frame(
+    gbar=7.6,
+    kappa=0.5,
+    R=rgamma(n, 1400, rate=1400/3.8),
+    r=(R^kappa-1)/(kappa*gbar),
+    rho=r*gbar
+  )
+}
+
+sample_riou <- function(n=1) {
+  # qgamma(0.05, 12, 12/2.5)
+  # qgamma(0.95, 12, 12/2.5)
+  # qgamma(0.5, 12, 12/2.5)
+  
+  dplyr::data_frame(
+    gbar=runif(n, 7, 14),
+    kappa=0.5,
+    R=rgamma(n, 12, rate=12/2.5),
+    r=(R^kappa-1)/(kappa*gbar),
+    rho=r*gbar
+  )
+}
+
+sample_zhao <- function(n=1) {
+  # qgamma(0.025, 50, 50/5.47)
+  # qgamma(0.975, 50, 50/5.47)
+  
+  dplyr::data_frame(
+    gbar=runif(n, 7.6, 8.4),
+    kappa=0.2,
+    R=rgamma(n, 50, rate=50/5.47),
+    r=(R^kappa-1)/(kappa*gbar),
+    rho=r*gbar
+  )
+}
+
+sample_liu <- function(n=1) {
+  # qgamma(0.025, 65, 65/2.92)
+  # qgamma(0.975, 65, 65/2.92)
+  
+  dplyr::data_frame(
+    gbar=8.4,
+    kappa=(3.8/8.4)^2,
[…]

### 2020-01-27 wuhan.tex
more updates

@@
 \begin{figure}[t]
-\includegraphics[width=\textwidth]{../fig/comparison.pdf}
+\includegraphics[width=\textwidth]{comparison.pdf}
 \caption{
@@
 
-To assess the amount of uncertainties in the estimate of $\mathcal R_0$,
-we come up with distributions that span over the assumptions made by other people:
 \begin{equation}
 \begin{aligned}
-r &\sim \mathrm{Gamma}(\alpha=5, \beta=5/0.15)\\
-\bar{G} &\sim \mathrm{Gamma}(\alpha=10, \beta=10/8.5)\\
-\kappa &\sim \mathrm{Gamma}(\alpha=10, \beta=10/0.35)\\
+\mathcal R_i &\sim \textrm{distrib}(\theta_i)\\
+\bar G_i &\sim \textrm{distrib}(\theta_i)\\
+\kappa &\sim \textrm{distrib}(\theta_i)\\
 \end{aligned}
 \end{equation}
-These give median estimate of $\mathcal R_0$ of 2.37 with 95\% qunatile of 1.26--9.01.
-
-\begin{table}[!th]
-\begin{center}
-\begin{tabular}{l|c|c|c|c}
-Study & $\mathcal R_0$ & $r$ ($\textrm{days}^{-1}$) & $\bar G$ (days) & $\kappa$ \\
-\hline
-\cite{imaincov} & 2.5 (1.5--3.5) & 0.14 (0.05--0.21) & 8.4 & unspecified$^\ast$ \\
-\hline
-\cite{majumderncov} & 2.0--3.3 & 0.07--0.20 & 6--10 & 0 \\
-\hline
-\cite{readncov} & 3.8 (3.6--4.0) & 0.25 (0.25--0.26)$^\dagger$ & 7.6 & 0.5 \\
-\hline
-\cite{riouncov} & 2.2 (1.4--3.8) & 0.09 (0.03--0.27) & 7--14 & 0.5\\
-\hline
-\cite{zhaoncov} & 5.47 (4.16--7.1)$^\ddagger$ & 0.25 (0.21--0.30) & 7.6--8.4 & 0.2
-\end{tabular}
-\end{center}
-\caption{
-\textbf{Parameter estimates and assumptions}
-Exponential growth rate $r$ is calculated by using the following formula: $r = (\mathcal R_0^\kappa - 1)/(\kappa \bar{G})$.
-$^\ast$We assume $\kappa = 0.5$ to calculate $r$.
-$^\dagger$Exponential growth rate $r$ is calculated by using the exact formula for the SEIR model \citep{ma2014estimating}.
-$^\ddagger$Assuming no increase in the reporting rate.
-}
-\end{table}
 

### 2020-01-27 MCMC_all.R

@@ -1,0 +1,57 @@
+library(MCMCpack)
+load("sample_data.rda")
+
+nsample <- 10000
+
+priorfun <- function(theta) {
+  sum(dgamma(theta, 0.1, 0.1, log=TRUE))
+}
+
+postfun <- function(theta) {
+  mean_r <- theta[1]
+  shape_r <- theta[2]
+  rate_r <- shape_r/mean_r
+  
+  mean_gbar <- theta[3]
+  shape_gbar <- theta[4]
+  rate_gbar <- shape_gbar/mean_gbar
+  
+  rate_kappa <- 1/theta[5]
+  
+  ss <- sample(nsample, 1)
+  
+  r <- sapply(sample_data, function(x) x$r[ss])
+  gbar <- sapply(sample_data, function(x) x$gbar[ss])
+  kappa <- sapply(sample_data, function(x) x$kappa[ss])
+  
+  ll <- sum(dgamma(r, shape=shape_r, rate=rate_r, log=TRUE)) +
+    sum(dgamma(gbar, shape=shape_gbar, rate=rate_gbar, log=TRUE)) +
+    sum(dexp(kappa, rate=rate_kappa, log=TRUE)) +
+    priorfun(theta)
+  
+  if (is.na(ll) || is.nan(ll)) {
+    return(-Inf)
+  }
+  
+  ll
+}
+
+V <- matrix(0, 5, 5)
+diag(V) <- c(1e-3, 40, 0.2, 50, 0.03)
+
+reslist <- vector('list', 4)
+
+for (i in 1:4) {
+  reslist[[i]] <- MCMCmetrop1R(postfun,
+                               theta.init=c(0.2, 10, 8, 70, 0.2),
+                               thin=200,
+                               burnin=100000,
+                               mcmc=100000,
+                               verbose=50000,
+                               seed=i,
+                               V=V)
+}
+
+MCMC_all <- as.mcmc.list(reslist)
+
+save("MCMC_all", file="MCMC_all.rda")

### 2020-01-27 R0_all.R

@@ -1,0 +1,22 @@
+load("sample_data.rda")
+load("MCMC_all.rda")
+
+r <- unlist(unlist(MCMC_all[,1]))
+gbar <- unlist(unlist(MCMC_all[,3]))
+kappa <- unlist(unlist(MCMC_all[,5]))
+
+R0_all <- lapply(sample_data, function(x) {
+  R <- (1+kappa*gbar*r)^(1/kappa)
+  
+  data.frame(
+    est=median(R),
+    lwr=quantile(R, 0.025),
+    upr=quantile(R, 0.975)
+  )
+}) %>%
+  bind_rows(.id="study") %>%
+  mutate(
+    type="all"
+  )
+
+save("R0_all", file="R0_all.rda")

### 2020-01-27 R0_all_uncorrelated.R

@@ -1,0 +1,22 @@
+load("sample_data.rda")
+load("MCMC_all.rda")
+
+r <- sample(unlist(unlist(MCMC_all[,1])), replace=FALSE)
+gbar <- sample(unlist(unlist(MCMC_all[,3])), replace=FALSE)
+kappa <- sample(unlist(unlist(MCMC_all[,5])), replace=FALSE)
+
+R0_all_uncorrelated <- lapply(sample_data, function(x) {
+  R <- (1+kappa*gbar*r)^(1/kappa)
+  
+  data.frame(
+    est=median(R),
+    lwr=quantile(R, 0.025),
+    upr=quantile(R, 0.975)
+  )
+}) %>%
+  bind_rows(.id="study") %>%
+  mutate(
+    type="allu"
+  )
+
+save("R0_all_uncorrelated", file="R0_all_uncorrelated.rda")

### 2020-01-27 R0_gbar.R

@@ -1,0 +1,23 @@
+load("sample_data.rda")
+load("MCMC_all.rda")
+
+gbar <- unlist(unlist(MCMC_all[,3]))
+
+R0_gbar <- lapply(sample_data, function(x) {
+  R <- (1+x$kappa*gbar*x$r)^(1/x$kappa)
+  
+  if (all(x$kappa == 0))
+    R <- exp(gbar*x$r)
+  
+  data.frame(
+    est=median(R),
+    lwr=quantile(R, 0.025),
+    upr=quantile(R, 0.975)
+  )
+}) %>%
+  bind_rows(.id="study") %>%
+  mutate(
+    type="gbar"
+  )
+
+save("R0_gbar", file="R0_gbar.rda")

### 2020-01-27 R0_kappa.R

@@ -1,0 +1,20 @@
+load("sample_data.rda")
+load("MCMC_all.rda")
+
+kappa <- unlist(unlist(MCMC_all[,5]))
+
+R0_kappa <- lapply(sample_data, function(x) {
+  R <- (1+kappa*x$rho)^(1/kappa)
+  
+  data.frame(
+    est=median(R),
+    lwr=quantile(R, 0.025),
+    upr=quantile(R, 0.975)
+  )
+}) %>%
+  bind_rows(.id="study") %>%
+  mutate(
+    type="kappa"
+  )
+
+save("R0_kappa", file="R0_kappa.rda")

### 2020-01-27 R0_r.R

@@ -1,0 +1,23 @@
+load("sample_data.rda")
+load("MCMC_all.rda")
+
+r <- unlist(unlist(MCMC_all[,1]))
+
+R0_r <- lapply(sample_data, function(x) {
+  R <- (1+x$kappa*x$gbar*r)^(1/x$kappa)
+  
+  if (all(x$kappa == 0))
+    R <- exp(x$gbar*r)
+  
+  data.frame(
+    est=median(R),
+    lwr=quantile(R, 0.025),
+    upr=quantile(R, 0.975)
+  )
+}) %>%
+  bind_rows(.id="study") %>%
+  mutate(
+    type="r"
+  )
+
+save("R0_r", file="R0_r.rda")

### 2020-01-27 R0_uncertainty.R

@@ -1,16 +1,0 @@
-set.seed(101)
-
-n <- 10000
-
-r <- rgamma(n, shape=5, rate=5/0.15)
-g <- rgamma(n, shape=10, rate=10/8.5)
-kappa <- rgamma(n, shape=10, rate=10/0.35)
-
-quantile(r, c(0.025, 0.975))
-quantile(g, c(0.025, 0.975))
-quantile(kappa, c(0.025, 0.975))
-
-R0 <- (1+g*r*kappa)^(1/kappa)
-
-median(R0)
-quantile(R0, c(0.025, 0.975))

### 2020-01-27 calculate_r.R

@@ -1,1 +1,0 @@
-source("sample.R")

### 2020-01-27 compare_R0.R

@@ -1,0 +1,33 @@
+library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
+source("published_estimate.R")
+
+load("R0_r.rda")
+load("R0_gbar.rda")
+load("R0_kappa.rda")
+load("R0_all.rda")
+
+R0all <- Rdata %>%
+  mutate(
+    type="base"
+  ) %>%
+  bind_rows(
+    R0_r, R0_gbar, R0_kappa, R0_all
+  ) %>%
+  mutate(
+    type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
+                labels=c("base", "growth rate", "GI mean", "GI variation", "all"))
+  )
+
+g1 <- ggplot(R0all) +
+  geom_point(aes(study, est, col=type), position=position_dodge(0.5)) +
+  geom_errorbar(aes(study, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),
+                width=0) +
+  scale_y_continuous("Basic reproductive number") +
+  theme(
+    axis.title.x = element_blank(),
+    legend.position = "top",
+    legend.title = element_blank()
+  )
+
+ggsave("compare_R0.pdf", g1, width=12, height=4)

### 2020-01-27 compare_assumption.R

@@ -1,0 +1,71 @@
+library(ggplot2); theme_set(theme_bw())
+library(gridExtra)
+source("published_estimate.R")
+
+load("sample_data.rda")
+
+load("MCMC_all.rda")
+
+rdata2 <- sample_data %>%
+  lapply(function(x) {
+    data.frame(
+      est=median(x$r),
+      lwr=quantile(x$r, 0.025),
+      upr=quantile(x$r, 0.975)
+    )
+  }) %>%
+  bind_rows(.id="study")
+
+avg_r <- data.frame(
+  median=median(unlist(unlist(MCMC_all[,1]))),
+  lwr=quantile(unlist(unlist(MCMC_all[,1])), 0.025),
+  upr=quantile(unlist(unlist(MCMC_all[,1])), 0.975),
+  study="Average"
+)
+
+avg_gen <- data.frame(
+  median=median(unlist(unlist(MCMC_all[,3]))),
+  lwr=quantile(unlist(unlist(MCMC_all[,3])), 0.025),
+  upr=quantile(unlist(unlist(MCMC_all[,3])), 0.975),
+  study="Average"
+)
+
+avg_kappa <- data.frame(
+  median=median(unlist(unlist(MCMC_all[,5]))),
+  lwr=quantile(unlist(unlist(MCMC_all[,5])), 0.025),
+  upr=quantile(unlist(unlist(MCMC_all[,5])), 0.975),
+  study="Average"
+)
+
+g1 <- ggplot(rdata2) +
+  geom_point(aes(est, study)) +
+  geom_point(data=avg_r, aes(median, study), col=2) +
+  geom_segment(aes(lwr, study, xend=upr, yend=study)) +
+  geom_segment(data=avg_r, aes(lwr, study, xend=upr, yend=study), col=2) +
+  scale_x_continuous("Exponential growth rate") +
+  theme(
+    axis.title.y = element_blank()
+  )
+
+g2 <- ggplot(gdata) +
+  geom_point(aes(est, study)) +
+  geom_point(data=avg_gen, aes(median, study), col=2) +
+  geom_segment(aes(lwr, study, xend=upr, yend=study)) +
+  geom_segment(data=avg_gen, aes(lwr, study, xend=upr, yend=study), col=2) +
+  scale_x_continuous("Mean generation time") +
+  theme(
+    axis.title.y = element_blank()
+  )
+
+g3 <- ggplot(kappadata) +
+  geom_point(aes(est, study)) +
[…]

### 2020-01-27 comparison.R

@@ -1,39 +1,0 @@
-library(ggplot2); theme_set(theme_bw())
-library(gridExtra)
-source("../data/estimate.R")
-
-g0 <- ggplot(Rdata) +
-  geom_point(aes(est, study)) +
-  geom_segment(aes(lwr, study, xend=upr, yend=study)) +
-  scale_x_continuous("Basic reproductive number") +
-  theme(
-    axis.title.y = element_blank()
-  )
-
-g1 <- ggplot(rdata) +
-  geom_point(aes(est, study)) +
-  geom_segment(aes(lwr, study, xend=upr, yend=study)) +
-  scale_x_continuous("Exponential growth rate") +
-  theme(
-    axis.title.y = element_blank()
-  )
-
-g2 <- ggplot(gdata) +
-  geom_point(aes(est, study)) +
-  geom_segment(aes(lwr, study, xend=upr, yend=study)) +
-  scale_x_continuous("Mean generation time") +
-  theme(
-    axis.title.y = element_blank()
-  )
-
-g3 <- ggplot(kappadata) +
-  geom_point(aes(est, study)) +
-  # geom_segment(aes(lwr, study, xend=upr, yend=study)) +
-  scale_x_continuous("Squared coefficient of variation") +
-  theme(
-    axis.title.y = element_blank()
-  )
-
-gtot <- arrangeGrob(g0, g1, g2, g3, nrow=2)
-
-ggsave("comparison.pdf", gtot, width=8, height=4)

### 2020-01-27 sample_data.R

@@ -1,0 +1,15 @@
+source("sample_author.R")
+
+nsample <- 10000
+
+set.seed(101)
+sample_data <- list(
+  `Imai et al. (2020)`=sample_imai(nsample),
+  `Liu et al. (2020)`=sample_liu(nsample),
+  `Majumder et al. (2020)`=sample_majumder(nsample),
+  `Read et al. (2020)`=sample_read(nsample),
+  `Riou and Althaus (2020)`=sample_riou(nsample),
+  `Zhao et al. (2020)`=sample_zhao(nsample)
+)
+
+save(sample_data, file="sample_data.rda")

### 2020-01-28 wuhan.tex
stuff

@@
 	\textbf\newline{
-		Assessing uncertainties associated with early estimates of the basic reproduction number during the novel coronavirus (2019-nCoV) outbreak
+		Evaluating uncertainties associated with early estimates of the basic reproduction number during the novel coronavirus (2019-nCoV) outbreak
 	}
@@
 
-As the novel coronavirus (2019-nCoV) continues to spread,
-researchers are rushing to publish their estimates of the basic 
-reproductive number $\mathcal R_0$.
-The basic reproductive number (i.e., the 
+\section{Introduction}
+
+Since the end of December 2019, a novel coronavirus (2019-nCoV) continues to
+spread in China and in other parts of the world.
+As of January 27th, 2020, the World Health Organization has
+confirmed 2798 cases, including 37 confirmed cases in 11 different
+countries, outside China \citep{who27report}.
+Although the virus is suspected to have originated
+from animal reservoirs, a recent case from Viet Nam
+demonstrated its ability to transmit between
+humans \citep{who26report},
+posing a greater threat for a wider spread.
+
+Many researchers have been rushing to publish their 
+analysis of the outbreak \citep{imaincov, riouncov,readncov,zhaoncov,majumderncov,liuncov} 
+and, in particular, their
+estimates of the basic 
+reproductive number $\mathcal R_0$ (i.e., the 
 average number of secondary cases generated 
-by a primary case in a fully susceptible population) is of particular interest 
+by a primary case in a fully susceptible population).
+The basic reproductive number is of particular interest 
 because it allows prediction about the final size of an epidemic.
@@
 assumptions that could immediately affect their estimates of $\mathcal R_0$ and
-the associated uncertainties in their estimates.
-Here, we illustrate the degree to which their assumptions can affect
-their estimates and stress few principles that needs to be taken into
+the associated uncertainties.
+
+Here, we present a simple framework for evaluating uncertainties 
+associated with parameter estimates
+across a wide range of models.
+Our results indicate that most published estimates of $\mathcal R_0$
+are likely to be overly confident. 
+We also lay down several principles that needs to be taken into
 consideration.
@@
 
-Early in an outbreak, $\mathcal R_0$ cannot be estimated directly,
-whereas the exponential growth rate $r$ can be estimated reliably.
+Early in an outbreak, $\mathcal R_0$ cannot be estimated directly.
+Instead, estimates of $\mathcal R_0$ are often inferred from
+the exponential growth rate $r$ can be estimated reliably from incidence data.
 Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of 
@@
 \end{equation}
-Therefore, early estimates of $\mathcal R_0$ must 
-necessarily depend on the assumptions about $r$ and $g(\tau)$.
+In other words, estimates of $\mathcal R_0$
+necessarily depends on the assumptions about 
+growth rate $r$ and the shape of generation-interval distribution $g(\tau)$.
 
 \begin{figure}[t]
-\includegraphics[width=\textwidth]{comparison.pdf}
+\includegraphics[width=\textwidth]{compare_assumption.pdf}
 \caption{

### 2020-01-28 wuhan.tex
stuff

@@
 
-\section*{Early estimates of $\mathcal R_0$}
+\section{Methods and results}
 
@@
 Instead, estimates of $\mathcal R_0$ are often inferred from
-the exponential growth rate $r$ can be estimated reliably from incidence data.
+the exponential growth rate $r$, which can be estimated reliably from incidence data.
 Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of
@@
 In other words, estimates of $\mathcal R_0$
-necessarily depends on the assumptions about 
-growth rate $r$ and the shape of generation-interval distribution $g(\tau)$.
+will depend on the assumptions about the
+exponential growth rate $r$ and the shape of generation-interval distribution $g(\tau)$.
 
@@
 
-We use the gamma approximation framework to demonstrate
-how assumptions about $r$ and $g(\tau)$
-affects $\mathcal R_0$.
+We use the gamma approximation framework to first characterize the
+amount of uncertainty present in the underlying parameters and then 
+to assess the 
+degree to which these uncertainties
+affect the estimate of $\mathcal R_0$.
 Assuming that generation intervals follow a gamma distribution 
@@
 \end{equation}
-This equation demonstrates that generation-interval distributions
-with a larger mean (higher $\bar{G}$) or less variability (lower $\kappa$)
-will give a higher estimate of $\mathcal R_0$.
-Likewise, strong assumptions about $\bar{G}$ and $\kappa$ will give
-estimates of $\mathcal R_0$ with narrow confidence intervals.
+This equation demonstrates that a generation-interval distribution
+that assumes a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
+will give a higher estimate of $\mathcal R_0$ for the same value of $r$.
 
-Figure 1 clearly demonstrates the lack of uncertainties in the underlying parameters
-of the current $\mathcal R_0$ estimates.
-In particular, no studies properly propagate uncertainties associated with the
-amount of variability in generation intervals, which can have 
-large effects on the estimates of $\mathcal R_0$
-For example, when the relative mean generation interval is long ($r\bar{G}\approx 2$ based on the estimates by \cite{zhaoncov}),
-varying $\kappa$ from 0 to 1 reduces the estimate of $\mathcal R_0$ by more than a two fold.
+First, we gathered information on estimates of $\mathcal R_0$ and their
+assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN).
+Then, we model each parameter ($\mathcal R_0$, $\bar G$, and $\kappa$) with an appropriate probability distribution (either uniform or gamma) that matches the reported estimates or assumptions.
+For example, \cite{imaincov} estimated $\mathcal R_0 = 2.5$
 
-\begin{equation}
-\begin{aligned}
-\mathcal R_i &\sim \textrm{distrib}(\theta_i)\\
-\bar G_i &\sim \textrm{distrib}(\theta_i)\\
-\kappa &\sim \textrm{distrib}(\theta_i)\\
-\end{aligned}
-\end{equation}
+\pagebreak
 

### 2020-01-28 sample_author.R

@@ -1,10 +1,11 @@
 sample_imai <- function(n=1) {
-  # qgamma(0.025, 20, 20/2.5)
-  # qgamma(0.975, 20, 20/2.5)
+  # qgamma(0.025, 20, 20/2.6)
+  # qgamma(0.975, 20, 20/2.6)
+  # qgamma(0.5, 20, 20/2.6)
   dplyr::data_frame(
     gbar=8.4,
     kappa=0.5,
-    R=rgamma(n, 20, rate=20/2.5),
+    R=rgamma(n, 20, rate=20/2.6),
     r=(R^kappa-1)/(kappa*gbar),
     rho=r*gbar
   )

### 2020-01-28 wuhan.tex
big update

@@
 \mathcal R_0 = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
+\label{eq:gamma}
 \end{equation}
@@
 
-First, we gathered information on estimates of $\mathcal R_0$ and their
+First, we gather information on estimates of $\mathcal R_0$ and their
 assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN).
-Then, we model each parameter ($\mathcal R_0$, $\bar G$, and $\kappa$) with an appropriate probability distribution (either uniform or gamma) that matches the reported estimates or assumptions.
-For example, \cite{imaincov} estimated $\mathcal R_0 = 2.5$
+For each study, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and uncertainty ranges.
+For example, \cite{imaincov} estimated $\mathcal R_0 = 2.6$ (95\% CI: 1.5--3.5);
+we model this parameter as a gamma distribution with mean of 2.6 and a shape parameter of 28, which has a 95.2\% probability of containing a value between 1.5 and 3.5.
+Then, we draw 100,000 random samples from these distributions and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
+\begin{equation}
+r = \frac{\mathcal R_0^\kappa - 1}{\kappa \bar{G}}
+\end{equation}
+
+To define ranges of appropriate uncertainties for each parameter ($r$, $\bar G$, and $\kappa$),
+we use a Bayesian multilevel modeling approach by assuming that all \emph{observed} (i.e., estimated or assumed) parameter values come from the same distribution:
+\begin{equation}
+\begin{aligned}
+r_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
+\bar{G}_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
+\kappa_i &\sim \mathrm{Exponential}(\mathrm{mean}=\mu_\kappa),\\
+(\mu_r, \alpha_r, \mu_G, \alpha_G, \mu_\kappa) &\sim \mathrm{Gamma}(\mathrm{shape}=0.1, \mathrm{rate}=0.1),
+\end{aligned}
+\end{equation}
+where $i$ represents the index across 6 studies that we consider.
+We use an exponential distribution to model $\kappa_i$ to account for the fact that the IDEA model, which \cite{majumderncov} uses, assumes that the generation-interval distribution is fixed ($\kappa = 0$).
+Hyperparameters $\mu_r$, $\alpha_r$, $\mu_G$, $\alpha_G$, and $\mu_\kappa$ are estimated with a Markov Chain Monte Carlo (MCMC) by drawing a random set for $r_i$, $\bar{G}_i$, and $\kappa_i$ at each MCMC step from the 100,000 samples that we calculated previously.
+We run 4 parallel chains that consist of 100,000 burnin steps and 100,000 sampling steps.
+Posterior samples are thinned every 200 steps.
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference}.
 

### 2020-01-28 MCMC_all.R

@@ -1,7 +1,7 @@
 library(MCMCpack)
 load("sample_data.rda")
 
-nsample <- 10000
+nsample <- 100000
 
 priorfun <- function(theta) {
   sum(dgamma(theta, 0.1, 0.1, log=TRUE))
@@ -37,7 +37,7 @@ postfun <- function(theta) {
 }
 
 V <- matrix(0, 5, 5)
-diag(V) <- c(1e-3, 40, 0.2, 50, 0.03)
+diag(V) <- c(5e-3, 40, 0.5, 50, 0.03)
 
 reslist <- vector('list', 4)
 

### 2020-01-28 sample_author.R

@@ -1,11 +1,10 @@
 sample_imai <- function(n=1) {
-  # qgamma(0.025, 20, 20/2.6)
-  # qgamma(0.975, 20, 20/2.6)
-  # qgamma(0.5, 20, 20/2.6)
+  # pgamma(3.5, 28, 28/2.6)-pgamma(1.5, 28, 28/2.6)
+  
   dplyr::data_frame(
     gbar=8.4,
     kappa=0.5,
-    R=rgamma(n, 20, rate=20/2.6),
+    R=rgamma(n, 28, rate=28/2.6),
     r=(R^kappa-1)/(kappa*gbar),
     rho=r*gbar
   )
@@ -22,8 +21,7 @@ sample_majumder <- function(n=1) {
 }
 
 sample_read <- function(n=1) {
-  # qgamma(0.025, 1400, 1400/3.8)
-  # qgamma(0.975, 1400, 1400/3.8)
+  # pgamma(4, 1400, 1400/3.8) - pgamma(3.6, 1400, 1400/3.8)
   
   dplyr::data_frame(
     gbar=7.6,
@@ -35,40 +33,36 @@ sample_read <- function(n=1) {
 }
 
 sample_riou <- function(n=1) {
-  # qgamma(0.05, 12, 12/2.5)
-  # qgamma(0.95, 12, 12/2.5)
-  # qgamma(0.5, 12, 12/2.5)
-  
+  # pgamma(3.8, 12, 12/2.2) - pgamma(1.4, 12, 12/2.2)
+
   dplyr::data_frame(
     gbar=runif(n, 7, 14),
     kappa=0.5,
-    R=rgamma(n, 12, rate=12/2.5),
+    R=rgamma(n, 12, rate=12/2.2),
     r=(R^kappa-1)/(kappa*gbar),
     rho=r*gbar
   )
 }
 
 sample_zhao <- function(n=1) {
-  # qgamma(0.025, 50, 50/5.47)
-  # qgamma(0.975, 50, 50/5.47)
+  # pgamma(7.1, 54, 54/5.47) - pgamma(4.16, 54, 54/5.47)
   
   dplyr::data_frame(
     gbar=runif(n, 7.6, 8.4),
     kappa=0.2,
-    R=rgamma(n, 50, rate=50/5.47),
+    R=rgamma(n, 54, rate=54/5.47),
     r=(R^kappa-1)/(kappa*gbar),
     rho=r*gbar
   )
 }
 
[…]

### 2020-01-28 sample_data.R

@@ -1,6 +1,6 @@
 source("sample_author.R")
 
-nsample <- 10000
+nsample <- 100000
 
 set.seed(101)
 sample_data <- list(

### 2020-01-28 wuhan.tex
more more more

--- 2020-01-28 wuhan.tex
@@
 	\textbf\newline{
-		Evaluating uncertainties associated with early estimates of the basic reproduction number during the novel coronavirus (2019-nCoV) outbreak
+		Evaluating uncertainties associated with early estimates of the basic reproductive number during the novel coronavirus (2019-nCoV) outbreak
 	}
@@
 
-Since the end of December 2019, a novel coronavirus (2019-nCoV) continues to
-spread in China and in other parts of the world.
-As of January 27th, 2020, the World Health Organization has
+Since December 2019, a novel coronavirus (2019-nCoV) has been
+spreading in China and in other parts of the world.
+As of January 27th, 2020, the World Health Organization (WHO) has
 confirmed 2798 cases, including 37 confirmed cases in 11 different
 countries, outside China \citep{who27report}.
-Although the virus is suspected to have originated
+Although the virus is believed to have originated
 from animal reservoirs, a recent case from Viet Nam
-demonstrated its ability to transmit between
+demonstrated its ability to directly transmit between
 humans \citep{who26report},
-posing a greater threat for a wider spread.
+posing a greater threat for its spread.
 
@@
 The basic reproductive number is of particular interest 
-because it allows prediction about the final size of an epidemic.
+because it allows prediction about the final size of an epidemic \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note}.
 While their efforts are valuable, their analyses rely on several
@@
 are likely to be overly confident. 
-We also lay down several principles that needs to be taken into
-consideration.
+We also lay down several principles that need to be taken into
+consideration when fitting models to early incidence data.
 
-\section{Methods and results}
+\section{Methods}
 
-Early in an outbreak, $\mathcal R_0$ cannot be estimated directly.
-Instead, estimates of $\mathcal R_0$ are often inferred from
-the exponential growth rate $r$, which can be estimated reliably from incidence data.
+Early in an outbreak, $\mathcal R_0$ cannot be estimated directly;
+instead, estimates of $\mathcal R_0$ are often inferred from
+the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{chowell2003sars, mills2004transmissibility, nishiura2009transmission, nishiura2010pros, ma2014estimating}.
 Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of
 generation intervals (i.e., the time between when a person become 
-infected and that person infects another person), the basic reproduction
-number can be estimated via the Euler-Lotka equation:
+infected and that person infects another person \citep{svensson2007note}), the basic reproductive
+number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{equation}
@@
 In other words, estimates of $\mathcal R_0$
-will depend on the assumptions about the
-exponential growth rate $r$ and the shape of generation-interval distribution $g(\tau)$.
+must depend on the assumptions about the
+exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
 
-\begin{figure}[t]
-\includegraphics[width=\textwidth]{compare_assumption.pdf}
-\caption{
-Early estimates of $\mathcal R_0$ and associated assumptions about $r$ and $g(\tau)$.
-}
-\end{figure}
-
-We use the gamma approximation framework to first characterize the
-amount of uncertainty present in the underlying parameters and then 
-to assess the 
+We use the gamma approximation framework \citep{park2019practical} to first characterize the
+amount of uncertainty present in the underlying parameters and then assess the 
 degree to which these uncertainties
@@
 \mathcal R_0 = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
+\label{eq:gamma}
 \end{equation}
 This equation demonstrates that a generation-interval distribution
-that assumes a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
+that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
 will give a higher estimate of $\mathcal R_0$ for the same value of $r$.
 
-First, we gathered information on estimates of $\mathcal R_0$ and their
+First, we gather information on estimates of $\mathcal R_0$ and their
 assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN).
-Then, we model each parameter ($\mathcal R_0$, $\bar G$, and $\kappa$) with an appropriate probability distribution (either uniform or gamma) that matches the reported estimates or assumptions.
-For example, \cite{imaincov} estimated $\mathcal R_0 = 2.5$
+For each study $i$, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and their uncertainty ranges:
+\begin{equation}
+\begin{aligned}
+\mathcal R_{0i} &\sim D_{i, \mathcal R}(\theta_{i, \mathcal R})\\
+{\bar G}_i &\sim D_{i, G}(\theta_{i, G})\\
+\kappa_i &\sim D_{i, \kappa}(\theta_{i, \kappa})\\
+\end{aligned}
+\end{equation}
+where $D$ is a probability distribution and $\theta$ is a parameter set of the probability distribution $D$.
+For example, \cite{imaincov} estimated $\mathcal R_0 = 2.6$ (95\% CI: 1.5--3.5);
+we model this parameter as a gamma distribution with mean of 2.6 and a shape parameter of 28, which has a 95\% probability of containing a value between 1.5 and 3.5.
+Then, we draw 100,000 random samples from these distributions ($\mathcal R_{0i}^n$, ${\bar G}_i^n$, and $\kappa_i^n$ where $n = 1, \dots, 100,000$) and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
+\begin{equation}
+r_i^n = \frac{{\mathcal R_{0i}^n}^{\kappa_i^n} - 1}{\kappa_i^n \bar{G}_i^n}.
+\end{equation}
+This allows us to approximte the probability distribution of the exponential growth rate as a discrete probability distribution.
+
+To define appropriate ranges of uncertainties for each parameter ($r$, $\bar G$, and $\kappa$),
+we use a Bayesian multilevel modeling approach by assuming that all \emph{observed} (i.e., estimated or assumed) parameter values come from the same distribution:
+\begin{equation}
+\begin{aligned}
+r_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
+\bar{G}_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
+\kappa_i^n &\sim \mathrm{Exponential}(\mathrm{mean}=\mu_\kappa),\\
+(\mu_r, \alpha_r, \mu_G, \alpha_G, \mu_\kappa) &\sim \mathrm{Gamma}(\mathrm{shape}=0.1, \mathrm{rate}=0.1),
+\end{aligned}
+\end{equation}
+where $n$ is randomly selected between 1 and 100,000 at each Metropolis-Hastings step;
+this approach is analogous to Bayesian methods for analyzing phylogenetic data, which can rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
+We use an exponential distribution to model $\kappa_i$ to account for the fact that the IDEA model, which \cite{majumderncov} uses, assumes a fixed generation interval ($\kappa = 0$).
+We run 4 parallel chains that consist of 100,000 burnin steps and 100,000 sampling steps.
+Posterior samples are thinned every 200 steps.
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all parameters \citep{gelman1992inference}.
+
+\section{Results}
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{compare_assumption.pdf}
+\caption{
+\textbf{Comparisons of estimated and assumed parameter values with their average values calculated from the multilevel model.}
+I'm a caption
+}
+\label{fig:assumption}
+\end{figure}
+
+\fref{assumption} compares the estimated and assumed values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared coefficient of variation $\kappa$ of the generation-interval distribution with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+We find that there is a large uncertainty associated with the underlying parameters;
+many models rely on stronger assumptions that ignore these uncertainties.
+Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different models.
+% For example, assuming $r = 0.15\,\mathrm{days}^{-1}$ and $\bar{G} = 8\,\mathrm{days}$, 
+% varying $\kappa$ from 0 to 0.5 reduce the estimate of $\mathcal R_0$ frmo 3.3 to 2.6.
+
+To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
+we replace the estimated or assumed values of $r$, $\bar G$, and $\kappa$ with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (15 out 18).
+We estimate  narrower confidence intervals for \cite{liuncov} and \cite{majumderncov} when we include uncertainties in $\kappa$ because they assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
+when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ becomes less sensitive to the values of $r$ and $\bar G$.
+We estimate narrower confidence intervals for \citep{riouncov} because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
+
+We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimate of $\mathcal R_0$ and the width of the confidence interval (\fref{R0}).
+For example, recalculating $\mathcal R_0$ by \cite{majumderncov} using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+
+\begin{figure}[t]
+\includegraphics[width=\textwidth]{compare_R0.pdf}
+\caption{
+\textbf{Effect of uncertainties in underlying parameters on the estimate of the basic reproductive number.}
+I'm a caption
+}
+\label{fig:R0}
+\end{figure}
 

### 2020-01-28 compare_R0.R

@@ -19,6 +19,20 @@ R0all <- Rdata %>%
                 labels=c("base", "growth rate", "GI mean", "GI variation", "all"))
   )
 
+R0all %>%
+  filter(type!="all", type!="base") %>%
+  mutate(
+    width=upr-lwr
+  ) %>%
+  merge(summarize(group_by(Rdata, study), bwidth=upr-lwr)) %>%
+  mutate(
+    ww=width>bwidth
+  ) %>%
+  summarize(
+    sum(ww)
+  )
+
+
 g1 <- ggplot(R0all) +
   geom_point(aes(study, est, col=type), position=position_dodge(0.5)) +
   geom_errorbar(aes(study, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),

### 2020-01-28 compare_assumption.R

@@ -42,7 +42,7 @@ g1 <- ggplot(rdata2) +
   geom_point(data=avg_r, aes(median, study), col=2) +
   geom_segment(aes(lwr, study, xend=upr, yend=study)) +
   geom_segment(data=avg_r, aes(lwr, study, xend=upr, yend=study), col=2) +
-  scale_x_continuous("Exponential growth rate") +
+  scale_x_continuous(expression(Exponentia~ growth~rate~(days^{-1}))) +
   theme(
     axis.title.y = element_blank()
   )
@@ -52,7 +52,7 @@ g2 <- ggplot(gdata) +
   geom_point(data=avg_gen, aes(median, study), col=2) +
   geom_segment(aes(lwr, study, xend=upr, yend=study)) +
   geom_segment(data=avg_gen, aes(lwr, study, xend=upr, yend=study), col=2) +
-  scale_x_continuous("Mean generation time") +
+  scale_x_continuous("Mean generation time (days)") +
   theme(
     axis.title.y = element_blank()
   )
@@ -68,4 +68,4 @@ g3 <- ggplot(kappadata) +
 
 gtot <- arrangeGrob(g1, g2, g3, nrow=1)
 
-ggsave("compare_assumption.pdf", gtot, width=12, height=4)
+ggsave("compare_assumption.pdf", gtot, width=12, height=3)

### 2020-01-29 wuhan.tex
more writing

@@
 }
-\newline
-\\
-Sang Woo Park\textsuperscript{1}
-Jonathan Dushoff\textsuperscript{2,3}
-\\
 
-\bigskip
-\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, USA
-\\
-\textbf{2} Department of Biology, McMaster University, Hamilton, Ontario, Canada
-\\
-\textbf{3} Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Ontario, Canada
-\\
 \bigskip
@@
 
-First, we gather information on estimates of $\mathcal R_0$ and their
-assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN).
+First, we gathered information on estimates of $\mathcal R_0$ and their
+assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN) between January 24th, 2020 and January 26th, 2020.
 For each study $i$, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and their uncertainty ranges:
@@
 Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different models.
-% For example, assuming $r = 0.15\,\mathrm{days}^{-1}$ and $\bar{G} = 8\,\mathrm{days}$, 
-% varying $\kappa$ from 0 to 0.5 reduce the estimate of $\mathcal R_0$ frmo 3.3 to 2.6.
 
@@
 
-We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimate of $\mathcal R_0$ and the width of the confidence interval (\fref{R0}).
-For example, recalculating $\mathcal R_0$ by \cite{majumderncov} using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
+For example, recalculating $\mathcal R_0$ for \cite{majumderncov} by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+There are two explanations for this result.
+First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
+$r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ than with $\mu_G$.
+Second, assuming a fixed generation time makes their estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
 
@@
 
+Finally, we recalculate $\mathcal R_0$ by accounting for all uncertainties (i.e., using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$) and comapre it with the reported $\mathcal R_0$ estimates.
+We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
+While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
+This result does not imply that our assumptions are too weak;
+we believe that this confidence interval accurately reflects the level of uncertainties present in the information that were available when these models were fitted.
+In fact, our results show that incorporating uncertainties in all parameters can give narrower confidence intervals than doing so one at a time because nonlinear effects of parameters could cancel out and give appropriately wide confidence intervals.
+
+\section{Discussion}
+
 \pagebreak

### 2020-01-29 R0_all_uncorrelated.R

@@ -1,22 +1,0 @@
-load("sample_data.rda")
-load("MCMC_all.rda")
-
-r <- sample(unlist(unlist(MCMC_all[,1])), replace=FALSE)
-gbar <- sample(unlist(unlist(MCMC_all[,3])), replace=FALSE)
-kappa <- sample(unlist(unlist(MCMC_all[,5])), replace=FALSE)
-
-R0_all_uncorrelated <- lapply(sample_data, function(x) {
-  R <- (1+kappa*gbar*r)^(1/kappa)
-  
-  data.frame(
-    est=median(R),
-    lwr=quantile(R, 0.025),
-    upr=quantile(R, 0.975)
-  )
-}) %>%
-  bind_rows(.id="study") %>%
-  mutate(
-    type="allu"
-  )
-
-save("R0_all_uncorrelated", file="R0_all_uncorrelated.rda")

### 2020-01-29 compare_R0.R

@@ -32,6 +32,15 @@ R0all %>%
     sum(ww)
   )
 
+R0all %>%
+  filter(type=="all") %>%
+  mutate(
+    width=upr-lwr
+  ) %>%
+  merge(summarize(group_by(Rdata, study), bwidth=upr-lwr)) %>%
+  mutate(
+    ww=width>bwidth
+  )
 
 g1 <- ggplot(R0all) +
   geom_point(aes(study, est, col=type), position=position_dodge(0.5)) +

### 2020-01-29 wuhan.tex
anonymyze papers

@@
 	\textbf\newline{
-		Evaluating uncertainties associated with early estimates of the basic reproductive number during the novel coronavirus (2019-nCoV) outbreak
+		Evaluating uncertainties associated with preliminary estimates of the basic reproductive number during the novel coronavirus (2019-nCoV) outbreak
 	}
@@
 First, we gathered information on estimates of $\mathcal R_0$ and their
-assumptions about the underlying generation-interval distributions from preprint servers (e.g., bioRxiv, medRxiv, and SSRN) between January 24th, 2020 and January 26th, 2020.
+assumptions about the underlying generation-interval distributions from
+6 articles that were uploaded to
+preprint servers (e.g., bioRxiv, medRxiv, and SSRN) between January 24th, 2020 and January 26th, 2020.
 For each study $i$, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and their uncertainty ranges:
@@
 where $D$ is a probability distribution and $\theta$ is a parameter set of the probability distribution $D$.
-For example, \cite{imaincov} estimated $\mathcal R_0 = 2.6$ (95\% CI: 1.5--3.5);
+For example, one study estimated $\mathcal R_0 = 2.6$ (95\% CI: 1.5--3.5);
 we model this parameter as a gamma distribution with mean of 2.6 and a shape parameter of 28, which has a 95\% probability of containing a value between 1.5 and 3.5.
@@
 \end{equation}
-This allows us to approximte the probability distribution of the exponential growth rate as a discrete probability distribution.
+This allows us to approximate the probability distribution of the exponential growth rate as a discrete probability distribution.
 
@@
 this approach is analogous to Bayesian methods for analyzing phylogenetic data, which can rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
-We use an exponential distribution to model $\kappa_i$ to account for the fact that the IDEA model, which \cite{majumderncov} uses, assumes a fixed generation interval ($\kappa = 0$).
+We use an exponential distribution to model $\kappa_i$ since one model assumes a fixed generation interval ($\kappa = 0$).
 We run 4 parallel chains that consist of 100,000 burnin steps and 100,000 sampling steps.
@@
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (15 out 18).
-We estimate  narrower confidence intervals for \cite{liuncov} and \cite{majumderncov} when we include uncertainties in $\kappa$ because they assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
+We estimate narrower confidence intervals for Study 2 and Study 3 when we include uncertainties in $\kappa$ because they assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
 when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ becomes less sensitive to the values of $r$ and $\bar G$.
-We estimate narrower confidence intervals for \citep{riouncov} because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
+We estimate narrower confidence intervals when we use the average mean generation time for Study 5 because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
 
 We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
-For example, recalculating $\mathcal R_0$ for \cite{majumderncov} by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
 There are two explanations for this result.
@@
 
-Finally, we recalculate $\mathcal R_0$ by accounting for all uncertainties (i.e., using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$) and comapre it with the reported $\mathcal R_0$ estimates.
+Finally, we recalculate $\mathcal R_0$ by accounting for all uncertainties (i.e., using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$) and compare it with the reported $\mathcal R_0$ estimates.
 We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
@@
 
+Estimating the basic reproductive number is crucial for predicting the course of an outbreak and planning intervention strategies.
+When the basic reproductive numbers are estimated from a wide range of models that rely on different assumptions, 
+it can difficult to assess the statistical validity of these estimates.
+Here, we present a simple framework for evaluating estimates of the basic reproductive number and associated model assumptions using the gamma approximation of the generation-interval distribution \citep{park2019practical}.
+
 \pagebreak

### 2020-01-29 compare_R0.R

@@ -1,6 +1,7 @@
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
 source("published_estimate.R")
+source("study_number.R")
 
 load("R0_r.rda")
 load("R0_gbar.rda")
@@ -17,7 +18,8 @@ R0all <- Rdata %>%
   mutate(
     type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
                 labels=c("base", "growth rate", "GI mean", "GI variation", "all"))
-  )
+  ) %>%
+  merge(study_number)
 
 R0all %>%
   filter(type!="all", type!="base") %>%
@@ -43,8 +45,8 @@ R0all %>%
   )
 
 g1 <- ggplot(R0all) +
-  geom_point(aes(study, est, col=type), position=position_dodge(0.5)) +
-  geom_errorbar(aes(study, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),
+  geom_point(aes(anon, est, col=type), position=position_dodge(0.5)) +
+  geom_errorbar(aes(anon, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),
                 width=0) +
   scale_y_continuous("Basic reproductive number") +
   theme(
@@ -53,4 +55,4 @@ g1 <- ggplot(R0all) +
     legend.title = element_blank()
   )
 
-ggsave("compare_R0.pdf", g1, width=12, height=4)
+ggsave("compare_R0.pdf", g1, width=6, height=4)

### 2020-01-29 compare_assumption.R

@@ -1,6 +1,7 @@
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
 source("published_estimate.R")
+source("study_number.R")
 
 load("sample_data.rda")
 
@@ -37,28 +38,28 @@ avg_kappa <- data.frame(
   study="Average"
 )
 
-g1 <- ggplot(rdata2) +
-  geom_point(aes(est, study)) +
+g1 <- ggplot(merge(rdata2, study_number)) +
+  geom_point(aes(est, anon)) +
+  geom_segment(aes(lwr, anon, xend=upr, yend=anon)) +
   geom_point(data=avg_r, aes(median, study), col=2) +
-  geom_segment(aes(lwr, study, xend=upr, yend=study)) +
   geom_segment(data=avg_r, aes(lwr, study, xend=upr, yend=study), col=2) +
   scale_x_continuous(expression(Exponentia~ growth~rate~(days^{-1}))) +
   theme(
     axis.title.y = element_blank()
   )
 
-g2 <- ggplot(gdata) +
-  geom_point(aes(est, study)) +
+g2 <- ggplot(merge(gdata, study_number)) +
+  geom_point(aes(est, anon)) +
+  geom_segment(aes(lwr, anon, xend=upr, yend=anon)) +
   geom_point(data=avg_gen, aes(median, study), col=2) +
-  geom_segment(aes(lwr, study, xend=upr, yend=study)) +
   geom_segment(data=avg_gen, aes(lwr, study, xend=upr, yend=study), col=2) +
   scale_x_continuous("Mean generation time (days)") +
   theme(
     axis.title.y = element_blank()
   )
 
-g3 <- ggplot(kappadata) +
-  geom_point(aes(est, study)) +
+g3 <- ggplot(merge(kappadata, study_number)) +
+  geom_point(aes(est, anon)) +
   geom_point(data=avg_kappa, aes(median, study), col=2) +
   geom_segment(data=avg_kappa, aes(lwr, study, xend=upr, yend=study), col=2) +
   scale_x_continuous("Squared coefficient of variation") +

### 2020-01-29 study_number.R

@@ -1,0 +1,5 @@
+study_number <- data.frame(
+  anon=paste0("Study ", 1:6),
+  study=c("Imai et al. (2020)", "Liu et al. (2020)", "Majumder et al. (2020)", 
+          "Read et al. (2020)", "Riou and Althaus (2020)", "Zhao et al. (2020)")
+)

### 2020-01-29 wuhan.tex
fiddling

@@
 spreading in China and in other parts of the world.
-As of January 27th, 2020, the World Health Organization (WHO) has
-confirmed 2798 cases, including 37 confirmed cases in 11 different
-countries, outside China \citep{who27report}.
+As of January 28th, 2020, the World Health Organization (WHO) has
+confirmed 4593 cases, including 56 confirmed cases in 14 different
+countries, outside China \citep{who28report}.
 Although the virus is believed to have originated
@@
 Many researchers have been rushing to publish their 
-analysis of the outbreak \citep{imaincov, riouncov,readncov,zhaoncov,majumderncov,liuncov} 
-and, in particular, their
-estimates of the basic 
-reproductive number $\mathcal R_0$ (i.e., the 
+analysis of the outbreak and, in particular, their
+estimates of the basic reproductive number $\mathcal R_0$ (i.e., the 
 average number of secondary cases generated 
-by a primary case in a fully susceptible population).
+by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
 The basic reproductive number is of particular interest 
 because it allows prediction about the final size of an epidemic \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note}.
-While their efforts are valuable, their analyses rely on several
+While their efforts are valuable, their analyses rely on
 assumptions that could immediately affect their estimates of $\mathcal R_0$ and
@@
 
-Here, we present a simple framework for evaluating uncertainties 
+Here, we present a statistical framework for evaluating uncertainties 
 associated with parameter estimates
@@
 Early in an outbreak, $\mathcal R_0$ cannot be estimated directly;
-instead, estimates of $\mathcal R_0$ are often inferred from
+instead, $\mathcal R_0$ is often inferred from
 the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{chowell2003sars, mills2004transmissibility, nishiura2009transmission, nishiura2010pros, ma2014estimating}.
-Given estimates of the exponential growth rate $r$ and the distribution $g(\tau)$ of
+Given an estimate of the exponential growth rate $r$ and a distribution $g(\tau)$ of
 generation intervals (i.e., the time between when a person become 
@@
 In other words, estimates of $\mathcal R_0$
-must depend on the assumptions about the
+depend on the assumptions about the
 exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
 
-We use the gamma approximation framework \citep{park2019practical} to first characterize the
+Here, we use the gamma approximation framework \citep{park2019practical} to first characterize the
 amount of uncertainty present in the underlying parameters and then assess the 

### 2020-01-29 wuhan.tex
keep making changes

@@
 Assuming that generation intervals follow a gamma distribution 
-with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
+with the mean $\bar G$ and the squared coefficient of variation (CV$^2$) $\kappa$, 
 we have
@@
 
+\begin{table}[t]
+\begin{center}
+\footnotesize
+\begin{tabular}{l|l|l|l}
+ & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
+\hline
+Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ \\
+\hline
+Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 \\
+\hline
+Study 3 & 2.0--3.1 & 6--10 & 0 \\
+\hline
+Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 \\
+\hline
+Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5\\
+\hline
+Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2
+\end{tabular}
+\end{center}
+\caption{
+\textbf{Estimates of the basic reproductive number and their assumptions about the generation-interval distributions.}
+$^\ast$We treated these intervals as a 95\% confidence interval in our analysis.
+$^\dagger$We assumed $\kappa = 0.5$ in our analysis.
+$^\ddagger$The authors presented $\mathcal R_0$ estimates under different assumptions; we used their baseline scenario in our analysis.
+}
+\end{table}
+
 First, we gathered information on estimates of $\mathcal R_0$ and their
@@
 6 articles that were uploaded to
-preprint servers (e.g., bioRxiv, medRxiv, and SSRN) between January 24th, 2020 and January 26th, 2020.
+preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
 For each study $i$, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and their uncertainty ranges:
@@
 where $D$ is a probability distribution and $\theta$ is a parameter set of the probability distribution $D$.
-For example, one study estimated $\mathcal R_0 = 2.6$ (95\% CI: 1.5--3.5);
-we model this parameter as a gamma distribution with mean of 2.6 and a shape parameter of 28, which has a 95\% probability of containing a value between 1.5 and 3.5.
+For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
+we model this parameter as a gamma distribution with mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for the list of all distributions).
 Then, we draw 100,000 random samples from these distributions ($\mathcal R_{0i}^n$, ${\bar G}_i^n$, and $\kappa_i^n$ where $n = 1, \dots, 100,000$) and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
@@
 This allows us to approximate the probability distribution of the exponential growth rate as a discrete probability distribution.
+
+\begin{table}[t]
+\begin{center}
+\footnotesize
+\begin{tabular}{l|l|l|l}
+ & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
+\hline
+Study 1 & $\mathrm{Gamma}(\alpha=28, \beta=28/2.6)$ & constant & constant \\
+\hline
+Study 2 & $\mathrm{Gamma}(\alpha=67, \beta=67/2.92)$ & constant & constant \\
+\hline
+Study 3 & N.A. & $\mathrm{Uniform}(6, 10)$ & constant\\
+\hline
+Study 4 & $\mathrm{Gamma}(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
+\hline
+Study 5 & $\mathrm{Gamma}(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
+\hline
+Study 6 & $\mathrm{Gamma}(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant
+\end{tabular}
+\end{center}
+\caption{
+\textbf{Probability distributions for the estimates of $\mathcal R_0$ and for the assumed values of $\bar G$ and $\kappa$.}
+The probability distribution for the exponential growth rate $r$ for each study is obtained by drawing random samples from these distributions and taking the inverse of \eref{gamma}.
+The gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
+Constant distributions assume a fixed values according to Table 1.
+}
+\end{table}
 

### 2020-01-29 sample_author.R

@@ -14,7 +14,7 @@ sample_majumder <- function(n=1) {
   dplyr::data_frame(
     gbar=runif(n, min=6,max=10),
     kappa=0,
-    r=runif(n, max=log(2)/6, min=log(3.1)/10),
+    r=0.114,
     R=exp(r*gbar),
     rho=r*gbar
   )

### 2020-01-29 wuhan.tex
a bit more

@@
 
+As the novel coronavirus (2019-nCoV) continues to spread,
+
 \pagebreak
@@
 
-Since December 2019, a novel coronavirus (2019-nCoV) has been
+Since December 2019, the novel coronavirus (2019-nCoV) has been
 spreading in China and in other parts of the world.
@@
 across a wide range of models.
-Our results indicate that most published estimates of $\mathcal R_0$
+Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak
 are likely to be overly confident.
-We also lay down several principles that need to be taken into
-consideration when fitting models to early incidence data.
 
@@
 For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
-we model this parameter as a gamma distribution with mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for the list of all distributions).
+we model this parameter as a gamma distribution with mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for the list of distributions that we assume).
 Then, we draw 100,000 random samples from these distributions ($\mathcal R_{0i}^n$, ${\bar G}_i^n$, and $\kappa_i^n$ where $n = 1, \dots, 100,000$) and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
@@
 
-To define appropriate ranges of uncertainties for each parameter ($r$, $\bar G$, and $\kappa$),
+To capture the average ranges of uncertainties for each parameter ($r$, $\bar G$, and $\kappa$),
 we use a Bayesian multilevel modeling approach by assuming that all \emph{observed} (i.e., estimated or assumed) parameter values come from the same distribution:
@@
 
-\fref{assumption} compares the estimated and assumed values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared coefficient of variation $\kappa$ of the generation-interval distribution with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared CV $\kappa$ of the generation-interval distribution with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
 We find that there is a large uncertainty associated with the underlying parameters;
@@
 To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
-we replace the estimated or assumed values of $r$, $\bar G$, and $\kappa$ with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+we replace the estimated (or assumed) values of $r$, $\bar G$, and $\kappa$ with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (15 out 18).
 We estimate narrower confidence intervals for Study 2 and Study 3 when we include uncertainties in $\kappa$ because they assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
-when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ becomes less sensitive to the values of $r$ and $\bar G$.
-We estimate narrower confidence intervals when we use the average mean generation time for Study 5 because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
+when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$.
+We estimate narrower confidence intervals when we use the average mean generation time $\mu_G$ to recalculate $\mathcal R_0$ for Study 5 because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
 
@@
 First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
-$r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ than with $\mu_G$.
-Second, assuming a fixed generation time makes their estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
+$r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ (95\% CI: 0.12--0.28 days$^{-1}$) than with $\mu_G$ (95\% CI: 6.6--10.2 days).
+Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
 
@@
 
-Estimating the basic reproductive number is crucial for predicting the course of an outbreak and planning intervention strategies.
-When the basic reproductive numbers are estimated from a wide range of models that rely on different assumptions, 
-it can difficult to assess the statistical validity of these estimates.
-Here, we present a simple framework for evaluating estimates of the basic reproductive number and associated model assumptions using the gamma approximation of the generation-interval distribution \citep{park2019practical}.
+Estimating the basic reproductive number $\mathcal R_0$ is crucial for predicting the course of an outbreak and planning intervention strategies.
+When $\mathcal R_0$ is estimated from a wide range of models that rely on different assumptions, it can difficult to assess the statistical validity of these estimates.
+Here, we present a simple framework that allows us to evaluate the validity of the estimates of $\mathcal R_0$ and associated model assumptions using the gamma approximation of the generation-interval distribution \citep{park2019practical}.
+We then apply the framework to the preliminary estimates of $\mathcal R_0$ during the novel coronavirus outbreak.
+Our results demonstrates the importance of accounting for uncertainties associated with the underlying generation-interval distributions, especially with the amount of dispersion in the generation intervals:
+assuming a narrow generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.
+
+Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approches that the estimates rely on.
+Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
+Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, 
+fitting a model to cumulative number of cases can bias the parameter estimate and give overly narrow confidence intervals as it violates the ``independence of error'' assumption \citep{ma2014estimating, king2015avoidable}.
+For example, fitting the exponential growth model to the daily number of confirmed 2019-nCoV cases give $r=??$ (95\% CI: ??
+-- ??), whereas fitting the same model to the daily number of cumulative confirmed cases give $r=??$ (95\% CI: ??
+-- ??).
+We strongly emphasize that all models should be fitted to raw incidence data for any future analyses.
+
+Concluding paragraph with a message?
 

### 2020-01-29 wuhan.tex
stuff

@@
 
-As the novel coronavirus (2019-nCoV) continues to spread,
-
 \pagebreak
@@
 spreading in China and in other parts of the world.
-As of January 28th, 2020, the World Health Organization (WHO) has
+As of January 28th, 2020, the World Health Organization has
 confirmed 4593 cases, including 56 confirmed cases in 14 different
@@
 
-Many researchers have been rushing to publish their 
+As the disease continues to spread, 
+many disease modelers have been rushing to publish their 
 analysis of the outbreak and, in particular, their
@@
 because it allows prediction about the final size of an epidemic \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note}.
-While their efforts are valuable, their analyses rely on
-assumptions that could immediately affect their estimates of $\mathcal R_0$ and
-the associated uncertainties.
+While their efforts are valuable, it is difficult to
+compare their estimates of $\mathcal R_0$ because their methods 
+and assumptions, which can directly affect their estimates and
+the associated uncertainties, vary widely.
 
-Here, we present a statistical framework for evaluating uncertainties 
-associated with parameter estimates
-across a wide range of models.
+Here, we present a simple, statistical framework for evaluating uncertainties 
+associated with the estimates of $\mathcal R_0$
+across a wide range of models and compare estimates of $\mathcal R_0$ for the outbreak.
 Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak
@@
 
-Here, we use the gamma approximation framework \citep{park2019practical} to first characterize the
-amount of uncertainty present in the underlying parameters and then assess the 
+Here, we use the gamma approximation framework \citep{park2019practical} to characterize the
+amount of uncertainty present in the underlying parameters and assess the 
 degree to which these uncertainties
@@
 
-To capture the average ranges of uncertainties for each parameter ($r$, $\bar G$, and $\kappa$),
+To capture the average ranges of uncertainties in each parameter ($r$, $\bar G$, and $\kappa$),
 we use a Bayesian multilevel modeling approach by assuming that all \emph{observed} (i.e., estimated or assumed) parameter values come from the same distribution:
@@
 
-\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared CV $\kappa$ of the generation-interval distribution with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared CV $\kappa$ of the generation-interval distribution with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
 We find that there is a large uncertainty associated with the underlying parameters;
 many models rely on stronger assumptions that ignore these uncertainties.
-Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different models.
+Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
 
 To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
-we replace the estimated (or assumed) values of $r$, $\bar G$, and $\kappa$ with the average values ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+we replace the estimated (or assumed) values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (15 out 18).
-We estimate narrower confidence intervals for Study 2 and Study 3 when we include uncertainties in $\kappa$ because they assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
+We estimate narrower confidence intervals for Study 2 and Study 3 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because both studies assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
 when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$.
-We estimate narrower confidence intervals when we use the average mean generation time $\mu_G$ to recalculate $\mathcal R_0$ for Study 5 because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
+We estimate narrower confidence intervals when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ for Study 5 because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
 
-We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
+We find that accounting for uncertainties in the estimate of $r$ (i.e., replacing the estimated $r$ values with $\mu_r$) has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
 For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
@@
 
-Finally, we recalculate $\mathcal R_0$ by accounting for all uncertainties (i.e., using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$) and compare it with the reported $\mathcal R_0$ estimates.
+Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
 We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
@@
 Estimating the basic reproductive number $\mathcal R_0$ is crucial for predicting the course of an outbreak and planning intervention strategies.
-When $\mathcal R_0$ is estimated from a wide range of models that rely on different assumptions, it can difficult to assess the statistical validity of these estimates.
-Here, we present a simple framework that allows us to evaluate the validity of the estimates of $\mathcal R_0$ and associated model assumptions using the gamma approximation of the generation-interval distribution \citep{park2019practical}.
-We then apply the framework to the preliminary estimates of $\mathcal R_0$ during the novel coronavirus outbreak.
-Our results demonstrates the importance of accounting for uncertainties associated with the underlying generation-interval distributions, especially with the amount of dispersion in the generation intervals:
+Here, we evaluated the statistical validity of the estimates of $\mathcal R_0$ for the novel coronavirus outbreak using a novel statistical framework, based on the gamma approximation of the generation-interval distribution \citep{park2019practical}.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, especially with the amount of dispersion in the generation intervals:
 assuming a narrow generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.

### 2020-01-29 wuhan.tex

@@
 Since December 2019, the novel coronavirus (2019-nCoV) has been
-spreading in China and in other parts of the world.
+spreading in China and other parts of the world.
 As of January 28th, 2020, the World Health Organization has
@@
 As the disease continues to spread, 
-many disease modelers have been rushing to publish their 
+many researchers have been rushing to publish their 
 analysis of the outbreak and, in particular, their
@@
 by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-The basic reproductive number is of particular interest 
-because it allows prediction about the final size of an epidemic \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note}.
-While their efforts are valuable, it is difficult to
-compare their estimates of $\mathcal R_0$ because their methods 
-and assumptions, which can directly affect their estimates and
-the associated uncertainties, vary widely.
+Estimating the basic reproductive number is of particular interest during an outbreak
+because it allows prediction about the final size of the outbreak \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note} and the required level of intervention \citep{anderson1991infectious}.
+However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty)
+when the estimation methods and their underlying assumptions vary widely.
 
@@
 associated with the estimates of $\mathcal R_0$
-across a wide range of models and compare estimates of $\mathcal R_0$ for the outbreak.
+across a wide range of models using a generation-interval based approach and compare early estimates of $\mathcal R_0$ of the 2019-nCoV outbreak.
 Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak

### 2020-01-29 wuhan.tex
some stuff

@@
 As the disease continues to spread, 
-many researchers have been rushing to publish their 
+many researchers are rushing to publish their 
 analysis of the outbreak and, in particular, their
@@
 by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-Estimating the basic reproductive number is of particular interest during an outbreak
+Estimating the basic reproductive number is of interest during an outbreak
 because it allows prediction about the final size of the outbreak \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note} and the required level of intervention \citep{anderson1991infectious}.
 However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty)
-when the estimation methods and their underlying assumptions vary widely.
+when the estimation methods and their underlying assumptions vary widely,
+especially since these assumptions can directly affect the estimates.
 
-Here, we present a simple, statistical framework for evaluating uncertainties 
+Here, we present a simple framework for evaluating uncertainties 
 associated with the estimates of $\mathcal R_0$
@@
 Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak
-are likely to be overly confident.
+are likely to have been overly confident.
 
@@
 \end{equation}
-In other words, estimates of $\mathcal R_0$
+In other words, estimates of $\mathcal R_0$ must
 depend on the assumptions about the
@@
 
-Here, we use the gamma approximation framework \citep{park2019practical} to characterize the
-amount of uncertainty present in the underlying parameters and assess the 
+Here, we use the gamma approximation framework \citep{park2019practical} to (1) 
+characterize the
+amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution
+and (2) assess the 
 degree to which these uncertainties
@@
 preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
-For each study $i$, we model $\mathcal R_0$, $\bar G$, and $\kappa$ with an appropriate probability distribution (either uniform or gamma) that roughly matches the estimated or assumed values and their uncertainty ranges:
+For each study $i$, we model $\mathcal R_{0i}$, $\bar G_i$, and $\kappa_i$ with appropriate probability distributions (either uniform or gamma) that roughly captures the reported values and their uncertainty ranges.
+For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
+we model this parameter as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for the list of distributions that we assume).
+If a study assumed a fixed value for a parameter, we hold that parameters constant as well, using the same assumed value.
+
+We draw 100,000 random samples from the probability distributions that represent the estimates of $\mathcal R_0$ and the assumed values of $\bar G$ and $\kappa$
+and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
 \begin{equation}
-\begin{aligned}
-\mathcal R_{0i} &\sim D_{i, \mathcal R}(\theta_{i, \mathcal R})\\
-{\bar G}_i &\sim D_{i, G}(\theta_{i, G})\\
-\kappa_i &\sim D_{i, \kappa}(\theta_{i, \kappa})\\
-\end{aligned}
+r_i = \frac{{\mathcal R_{0i}}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
 \end{equation}
-where $D$ is a probability distribution and $\theta$ is a parameter set of the probability distribution $D$.
-For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
-we model this parameter as a gamma distribution with mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for the list of distributions that we assume).
-Then, we draw 100,000 random samples from these distributions ($\mathcal R_{0i}^n$, ${\bar G}_i^n$, and $\kappa_i^n$ where $n = 1, \dots, 100,000$) and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
-\begin{equation}
-r_i^n = \frac{{\mathcal R_{0i}^n}^{\kappa_i^n} - 1}{\kappa_i^n \bar{G}_i^n}.
-\end{equation}
-This allows us to approximate the probability distribution of the exponential growth rate as a discrete probability distribution.
+This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.
 
@@
 
-To capture the average ranges of uncertainties in each parameter ($r$, $\bar G$, and $\kappa$),
-we use a Bayesian multilevel modeling approach by assuming that all \emph{observed} (i.e., estimated or assumed) parameter values come from the same distribution:
+We calculate the pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) with a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same Gamma distribution:
 \begin{equation}
 \begin{aligned}
-r_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
-\bar{G}_i^n &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
-\kappa_i^n &\sim \mathrm{Exponential}(\mathrm{mean}=\mu_\kappa),\\
-(\mu_r, \alpha_r, \mu_G, \alpha_G, \mu_\kappa) &\sim \mathrm{Gamma}(\mathrm{shape}=0.1, \mathrm{rate}=0.1),
+r_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
+\bar{G}_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
+\kappa_i &\sim \mathrm{Exponential}(\mathrm{mean}=\mu_\kappa),\\
 \end{aligned}
 \end{equation}
-where $n$ is randomly selected between 1 and 100,000 at each Metropolis-Hastings step;
-this approach is analogous to Bayesian methods for analyzing phylogenetic data, which can rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
-We use an exponential distribution to model $\kappa_i$ since one model assumes a fixed generation interval ($\kappa = 0$).
-We run 4 parallel chains that consist of 100,000 burnin steps and 100,000 sampling steps.
-Posterior samples are thinned every 200 steps.
+We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing random samples at each Metropolis-Hastings step from 100,000 samples that we calculated earlier;
+this approach is analogous to Bayesian methods for analyzing phylogenetic data, which pften rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
+Weak priors are assumed on the hyperparameters:
+\begin{equation}
+(\mu_r, \alpha_r, \mu_G, \alpha_G, \mu_\kappa) \sim \mathrm{Gamma}(\mathrm{shape}=0.1, \mathrm{rate}=0.1).
+\end{equation}
+We run 4 parallel Markov Chain Monte Carlo (MCMC) chains that consist of 200,000 burnin steps and 200,000 sampling steps.
+Posterior samples are thinned every 400 steps.
 Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all parameters \citep{gelman1992inference}.
@@
 Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure2.pdf}
+\caption{
+\textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of $\mathcal R_0$.}
+I'm a caption
+}
+\label{fig:eff}
+\end{figure}
+
+To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
+we
 
@@
 fitting a model to cumulative number of cases can bias the parameter estimate and give overly narrow confidence intervals as it violates the ``independence of error'' assumption \citep{ma2014estimating, king2015avoidable}.
-For example, fitting the exponential growth model to the daily number of confirmed 2019-nCoV cases give $r=??$ (95\% CI: ??
--- ??), whereas fitting the same model to the daily number of cumulative confirmed cases give $r=??$ (95\% CI: ??
--- ??).
-We strongly emphasize that all models should be fitted to raw incidence data for any future analyses.
+We emphasize that all models should be fitted to raw incidence data for any future analyses.
 

### 2020-01-29 MCMC_all.R

@@ -7,6 +7,8 @@ priorfun <- function(theta) {
   sum(dgamma(theta, 0.1, 0.1, log=TRUE))
 }
 
+sample_data$`Majumder et al. (2020)`$kappa <- 0.02
+
 postfun <- function(theta) {
   mean_r <- theta[1]
   shape_r <- theta[2]
@@ -16,7 +18,9 @@ postfun <- function(theta) {
   shape_gbar <- theta[4]
   rate_gbar <- shape_gbar/mean_gbar
   
-  rate_kappa <- 1/theta[5]
+  mean_kappa <- theta[5]
+  shape_kappa <- theta[6]
+  rate_kappa <- shape_kappa/mean_kappa
   
   ss <- sample(nsample, 1)
   
@@ -26,7 +30,7 @@ postfun <- function(theta) {
   
   ll <- sum(dgamma(r, shape=shape_r, rate=rate_r, log=TRUE)) +
     sum(dgamma(gbar, shape=shape_gbar, rate=rate_gbar, log=TRUE)) +
-    sum(dexp(kappa, rate=rate_kappa, log=TRUE)) +
+    sum(dgamma(kappa, shape=shape_kappa, rate=rate_kappa, log=TRUE)) +
     priorfun(theta)
   
   if (is.na(ll) || is.nan(ll)) {
@@ -36,17 +40,17 @@ postfun <- function(theta) {
   ll
 }
 
-V <- matrix(0, 5, 5)
-diag(V) <- c(5e-3, 40, 0.5, 50, 0.03)
+V <- matrix(0, 6, 6)
+diag(V) <- c(5e-3, 50, 0.5, 60, 0.04, 1)
 
 reslist <- vector('list', 4)
 
 for (i in 1:4) {
   reslist[[i]] <- MCMCmetrop1R(postfun,
-                               theta.init=c(0.2, 10, 8, 70, 0.2),
-                               thin=200,
-                               burnin=100000,
-                               mcmc=100000,
+                               theta.init=c(0.2, 10, 8, 70, 0.2, 1),
+                               thin=400,
+                               burnin=200000,
+                               mcmc=200000,
                                verbose=50000,
                                seed=i,
                                V=V)

### 2020-01-29 MCMC_gen.R

@@ -1,0 +1,51 @@
+library(MCMCpack)
+load("sample_data.rda")
+
+rpar <- sapply(sample_data, function(x) {
+  mm <- mean(x$r)
+  shape <- mm^2/var(x$r)
+  rate <- shape/mm
+  
+  c(shape=shape, rate=rate, mm=mm)
+})
+
+priorfun <- function(theta) {
+  
+  
+  sum(dgamma(theta[1:nstudy], shape=rpar[1,], rate=rpar[2,], log=TRUE)) +
+    dgamma(theta[nstudy+1], 0.1, 0.1, log=TRUE) +
+    dgamma(theta[nstudy+2], 0.1, 0.1, log=TRUE)
+}
+
+postfun <- function(theta,
+                    nstudy=6) {
+  mean <- theta[nstudy+1]
+  shape <- theta[nstudy+2]
+  rate <- shape/mean
+  
+  ll <- sum(dgamma(theta[1:nstudy], shape=shape, rate=rate, log=TRUE)) +
+    priorfun(theta, nstudy=nstudy)
+  
+  if (is.na(ll) || is.nan(ll)) {
+    return(-Inf)
+  }  
+  
+  ll
+}
+
+reslist <- vector('list', 4)
+
+for (i in 1:4) {
+  reslist[[i]] <- MCMCmetrop1R(postfun,
+                               theta.init=c(rpar[3,], mean(rpar[3,]), 10),
+                               thin=200,
+                               burnin=40000,
+                               mcmc=40000,
+                               verbose=20000,
+                               seed=i)
+  
+}
+
+MCMC_r <- as.mcmc.list(reslist)
+
+save("MCMC_r", file="MCMC_r.rda")

### 2020-01-29 R0_all.R

@@ -1,3 +1,5 @@
+library(coda)
+
 load("sample_data.rda")
 load("MCMC_all.rda")
 

### 2020-01-29 compare_assumption.R

@@ -1,4 +1,5 @@
 library(ggplot2); theme_set(theme_bw())
+library(dplyr)
 library(gridExtra)
 source("published_estimate.R")
 source("study_number.R")

### 2020-01-29 figure2.R

@@ -1,0 +1,75 @@
+library(coda)
+library(ggplot2); theme_set(theme_bw())
+
+load("sample_data.rda")
+load("MCMC_all.rda")
+
+r <- unlist(unlist(MCMC_all[,1]))
+gbar <- unlist(unlist(MCMC_all[,3]))
+kappa <- unlist(unlist(MCMC_all[,5]))
+
+r_narrow <- (r + median(r))/2
+
+compfun <- function(x) {
+  dd <- data.frame(
+    est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
+    lwr=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
+    upr=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
+    type="base"
+  )
+  
+  dd_r <- data.frame(
+    est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
+    lwr=quantile((1+median(kappa)*median(gbar)*x)^(1/median(kappa)), 0.025),
+    upr=quantile((1+median(kappa)*median(gbar)*x)^(1/median(kappa)), 0.975),
+    type="growth rate"
+  )
+  
+  dd_gbar <- data.frame(
+    est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
+    lwr=quantile((1+median(kappa)*gbar*median(x))^(1/median(kappa)), 0.025),
+    upr=quantile((1+median(kappa)*gbar*median(x))^(1/median(kappa)), 0.975),
+    type="GI mean"
+  )
+  
+  dd_both <- data.frame(
+    est=median((1+median(kappa)*gbar*x)^(1/median(kappa))),
+    lwr=quantile((1+median(kappa)*gbar*x)^(1/median(kappa)), 0.025),
+    upr=quantile((1+median(kappa)*gbar*x)^(1/median(kappa)), 0.975),
+    type="growth rate\n+\nGI mean"
+  )
+  
+  dd_all <- data.frame(
+    est=median((1+kappa*gbar*x)^(1/kappa)),
+    lwr=quantile((1+kappa*gbar*x)^(1/kappa), 0.025),
+    upr=quantile((1+kappa*gbar*x)^(1/kappa), 0.975),
+    type="all"
+  )
+  
+  dd_combine <- rbind(dd, dd_r, dd_gbar, dd_both, dd_all)
+  
+  dd_combine
+}
+
[…]

### 2020-01-30 wuhan.tex
fiddle

@@
 \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of $\mathcal R_0$.}
-I'm a caption
+(base): $\mathcal R_0$ estimate based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+(growth rate): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_r$ while assuming median estimates of $\mu_G$ and $\mu_\kappa$.
+(GI mean): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_G$ while assuming median estimates of $\mu_r$ and $\mu_\kappa$.
+(growth rate + GI mean): $\mathcal R_0$ estimate based on the the full posterior distributions of $\mu_r$ and $\mu_G$ while assuming a median estimate of $\mu_\kappa$.
+(all): $\mathcal R_0$ estimate based on the the full posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
 }
@@
 
-To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
-we
+\fref{eff} demonstrates the degree to which how the uncertainty in each parameter propogates to the uncertainties in $\mathcal R_0$.
+\swp{JD: can you explain figure 2?}
 
-To evaluate how the uncertainty in each parameter affects the estimate of $\mathcal R_0$,
-we replace the estimated (or assumed) values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate the basic reproductive number $\mathcal R_0$ (\fref{R0}).
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (15 out 18).
-We estimate narrower confidence intervals for Study 2 and Study 3 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because both studies assume narrow generation-interval distributions ($\kappa = 0.2$ and $\kappa=0$, respectively);
+Finally, we assess the validity of estimates of $\mathcal R_0$ across different studies by 
+replacing the their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (16 out 18).
+We estimate narrower confidence intervals for Study 2 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0$);
 when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$.
-We estimate narrower confidence intervals when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ for Study 5 because the range of uncertainty in the mean generation time $\bar G$ that they consider is much wider than ours (\fref{assumption}).
+We estimate narrower confidence intervals for Study 5 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
 
-We find that accounting for uncertainties in the estimate of $r$ (i.e., replacing the estimated $r$ values with $\mu_r$) has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
-For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
-There are two explanations for this result.
-First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
-$r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ (95\% CI: 0.12--0.28 days$^{-1}$) than with $\mu_G$ (95\% CI: 6.6--10.2 days).
-Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
+% We find that accounting for uncertainties in the estimate of $r$ (i.e., replacing the estimated $r$ values with $\mu_r$) has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
+% For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+% There are two explanations for this result.
+% First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
+% $r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ (95\% CI: 0.12--0.28 days$^{-1}$) than with $\mu_G$ (95\% CI: 6.6--10.2 days).
+% Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
 
@@
 
-Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
+When we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
 We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).

### 2020-01-30 compare_R0.R

@@ -8,12 +8,12 @@ load("R0_gbar.rda")
 load("R0_kappa.rda")
 load("R0_all.rda")
 
-R0all <- Rdata %>%
+R0all <- Rdata_adj %>%
   mutate(
     type="base"
   ) %>%
   bind_rows(
-    R0_r, R0_gbar, R0_kappa, R0_all
+    R0_r, R0_gbar, R0_kappa
   ) %>%
   mutate(
     type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
@@ -26,7 +26,7 @@ R0all %>%
   mutate(
     width=upr-lwr
   ) %>%
-  merge(summarize(group_by(Rdata, study), bwidth=upr-lwr)) %>%
+  merge(summarize(group_by(Rdata_adj, study), bwidth=upr-lwr)) %>%
   mutate(
     ww=width>bwidth
   ) %>%
@@ -45,6 +45,8 @@ R0all %>%
   )
 
 g1 <- ggplot(R0all) +
+  geom_hline(data=R0_all, aes(yintercept=lwr), lty=2) +
+  geom_hline(data=R0_all, aes(yintercept=upr), lty=2) +
   geom_point(aes(anon, est, col=type), position=position_dodge(0.5)) +
   geom_errorbar(aes(anon, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),
                 width=0) +

### 2020-01-30 figure2.R

@@ -57,12 +57,7 @@ g1 <- ggplot(dd_combine) +
   geom_point(aes(type, est)) +
   geom_errorbar(aes(type, min=lwr, max=upr), width=0) +
   xlab("Uncertainty type") +
-  scale_y_continuous("Basic reproductive number", limits=c(2, 6)) +
-  theme(
-    panel.grid = element_blank(),
-    panel.border = element_blank(),
-    axis.line = element_line()
-  )
+  scale_y_continuous("Basic reproductive number", limits=c(2, 6))
 
 dd_combine2 <- compfun(r_narrow)
 

### 2020-01-30 published_estimate.R

@@ -6,6 +6,14 @@ Rdata <- data.frame(
   upr=c(3.5, 3.8, 4, 7.1, 3.1, 3.67)
 )
 
+Rdata_adj <- data.frame(
+  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
+          "Zhao et al. (2020)", "Majumder et al. (2020)", "Liu et al. (2020)"),
+  est=c(2.5, 2.2, 3.8, 5.47, NA, 2.92),
+  lwr=c(1.5, qgamma(0.025, 12, 12/2.2), 3.6, 4.16, exp(0.114 * qunif(0.025, 6, 10)), 2.28),
+  upr=c(3.5, qgamma(0.975, 12, 12/2.2), 4, 7.1, exp(0.114 * qunif(0.975, 6, 10)), 3.67)
+)
+
 gdata <- data.frame(
   study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
           "Zhao et al. (2020)", "Majumder et al. (2020)", "Liu et al. (2020)"),

### 2020-01-30 wuhan.tex
Have to call my Mom

@@
 \newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
+\newcommand{\jd}[1]{\comment{magenta}{JD}{#1}}
 \newcommand{\new}[1]{\textcolor{blue}{#1}}
@@
 posing a greater threat for its spread.
+\jd{It is spread primarily from person to person; that's been obvious for a while and is now confirmed. We should find a better cite and leave Viet Nam out of it.}
 
+\jd{What's up with the line breaks??}
 As the disease continues to spread, 
-many researchers are rushing to publish their 
-analysis of the outbreak and, in particular, their
+many researchers have published
+analyses of the outbreak, many focusing in particular on
 estimates of the basic reproductive number $\mathcal R_0$ (i.e., the 
@@
 Estimating the basic reproductive number is of interest during an outbreak
-because it allows prediction about the final size of the outbreak \citep{anderson1991infectious, ma2006generality, arino2007final, andreasen2011final, miller2012note} and the required level of intervention \citep{anderson1991infectious}.
+because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+\jd{We really only need one cite for this whole s., but I think Ma will make David happy.}
 However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty)
@@
 instead, $\mathcal R_0$ is often inferred from
-the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{chowell2003sars, mills2004transmissibility, nishiura2009transmission, nishiura2010pros, ma2014estimating}.
+the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
+\jd{Trimmed references; in particular, don't really want to point people to Chowell for stats.}
 Given an estimate of the exponential growth rate $r$ and a distribution $g(\tau)$ of
@@
 affect the estimate of $\mathcal R_0$.
+\jd{We should cite another group as well about this approximation.}
 Assuming that generation intervals follow a gamma distribution
@@
 
-We calculate the pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) with a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same Gamma distribution:
+\jd{Are you still using exponential?}
+We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same Gamma distribution:
 \begin{equation}
@@
 this approach is analogous to Bayesian methods for analyzing phylogenetic data, which pften rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
+\jd{use mean (not rate) and shape; specify units for the means.}
 Weak priors are assumed on the hyperparameters:

### 2020-01-30 wuhan.tex
conversation with Daniel

@@
 
-\jd{What's up with the line breaks??}
-As the disease continues to spread, 
-many researchers have published
-analyses of the outbreak, many focusing in particular on
-estimates of the basic reproductive number $\mathcal R_0$ (i.e., the 
-average number of secondary cases generated 
-by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-Estimating the basic reproductive number is of interest during an outbreak
-because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
-\jd{We really only need one cite for this whole s., but I think Ma will make David happy.}
-However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty)
-when the estimation methods and their underlying assumptions vary widely,
-especially since these assumptions can directly affect the estimates.
+As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number $\mathcal R_0$ (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
 
@@
 the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}. 
-\jd{Trimmed references; in particular, don't really want to point people to Chowell for stats.}
 Given an estimate of the exponential growth rate $r$ and a distribution $g(\tau)$ of
@@
 
-Here, we use the gamma approximation framework \citep{park2019practical} to (1) 
-characterize the
-amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution
-and (2) assess the 
-degree to which these uncertainties
-affect the estimate of $\mathcal R_0$.
+Here, we use the gamma approximation framework \citep{park2019practical} to (1) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of $\mathcal R_0$.
 \jd{We should cite another group as well about this approximation.}
@@
 
-\jd{Are you still using exponential?}
+\jd{Fix exponential?}
 We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same Gamma distribution:
@@
 this approach is analogous to Bayesian methods for analyzing phylogenetic data, which pften rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
-\jd{use mean (not rate) and shape; specify units for the means.}
+\jd{use mean (not rate) and shape; specify units for the means. Use this as an opportunity to question your choices maybe. Consider using 1 week and 1/week for the means.}
 Weak priors are assumed on the hyperparameters:
@@
 
+\jd{Change speculative to hypothetical.}
 \begin{figure}[!ht]

### 2020-01-30 wuhan.tex
after meeting with Daniel

@@
 
-We draw 100,000 random samples from the probability distributions that represent the estimates of $\mathcal R_0$ and the assumed values of $\bar G$ and $\kappa$
-and calculate the exponential growth rate $r$ via the inverse of \eref{gamma}:
+For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions that represent the estimates of $\mathcal R_0$ and the assumed values of $\bar G$ and $\kappa$
+and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
 \begin{equation}
@@
 \end{equation}
-We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing random samples at each Metropolis-Hastings step from 100,000 samples that we calculated earlier;
-this approach is analogous to Bayesian methods for analyzing phylogenetic data, which pften rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
+We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step;
+this approach is analogous to Bayesian methods for analyzing phylogenetic data, which often rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
 \jd{use mean (not rate) and shape; specify units for the means. Use this as an opportunity to question your choices maybe. Consider using 1 week and 1/week for the means.}

### 2020-01-30 wuhan.tex
push

--- 2020-01-30 wuhan.tex
@@
 
-Since December 2019, the novel coronavirus (2019-nCoV) has been
-spreading in China and other parts of the world.
-As of January 28th, 2020, the World Health Organization has
-confirmed 4593 cases, including 56 confirmed cases in 14 different
-countries, outside China \citep{who28report}.
-Although the virus is believed to have originated
-from animal reservoirs, a recent case from Viet Nam
-demonstrated its ability to directly transmit between
-humans \citep{who26report},
+Since December 2019, the novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
+As of January 28th, 2020, the World Health Organization has confirmed 4593 cases, including 56 confirmed cases in 14 different countries, outside China \citep{who28report}.
+Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, 
+there is clear evidence that it can be directly tranmitted between humans,
 posing a greater threat for its spread.

### 2020-01-30 wuhan.tex
pull on request

--- 2020-01-30 wuhan.tex
@@
 
-Here, we present a simple framework for evaluating uncertainties 
+Here, we show that a wide range of approaches to estimating \Ro\ present a simple framework for evaluating uncertainties 
 associated with the estimates of $\mathcal R_0$

### 2020-01-30 wuhan.tex
new figs

--- 2020-01-30 wuhan.tex
@@
 \bar{G}_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
-\kappa_i &\sim \mathrm{Exponential}(\mathrm{mean}=\mu_\kappa),\\
+\kappa_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\alpha_\kappa).\\
 \end{aligned}
@@
 \jd{use mean (not rate) and shape; specify units for the means. Use this as an opportunity to question your choices maybe. Consider using 1 week and 1/week for the means.}
-Weak priors are assumed on the hyperparameters:
+Weakly informative priors are assumed on the hyperparameters:
 \begin{equation}
-(\mu_r, \alpha_r, \mu_G, \alpha_G, \mu_\kappa) \sim \mathrm{Gamma}(\mathrm{shape}=0.1, \mathrm{rate}=0.1).
+\begin{aligned}
+\mu_r &\sim \mathrm{Gamma}(\mathrm{mean}=0.1\,\mathrm{day}^{-1},\,\mathrm{shape}=0.1)\\
+\mu_G &\sim \mathrm{Gamma}(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\
+(\mu_\kappa, \alpha_r, \alpha_G, \alpha_\kappa) &\sim \mathrm{Gamma}(\mathrm{mean}=1,\,\mathrm{shape}=0.1).
+\end{aligned}
 \end{equation}

### 2020-01-30 MCMC_all.R

@@ -4,7 +4,9 @@ load("sample_data.rda")
 nsample <- 100000
 
 priorfun <- function(theta) {
-  sum(dgamma(theta, 0.1, 0.1, log=TRUE))
+  sum(dgamma(theta[-c(1, 3)], 0.1, 0.1/1, log=TRUE)) +
+    dgamma(theta[1], 0.1, 0.1/0.1, log=TRUE) +
+    dgamma(theta[3], 0.1, 0.1/7, log=TRUE)
 }
 
 sample_data$`Majumder et al. (2020)`$kappa <- 0.02

### 2020-01-30 wuhan.tex
pullpush

--- 2020-01-30 wuhan.tex
@@
 
-Here, we show that a wide range of approaches to estimating \Ro\ present a simple framework for evaluating uncertainties 
-associated with the estimates of $\mathcal R_0$
-across a wide range of models using a generation-interval based approach and compare early estimates of $\mathcal R_0$ of the 2019-nCoV outbreak.
+Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+We argue that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimate uncertainty should propagate error from all three sources.
+
+We evaluate six disparate models posted online in daterange that estimated \Ro\ for the 2019-nCoV outbreak. <C-F3>
 Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak

### 2020-01-30 figure2.R
fig2

@@ -8,8 +8,6 @@ r <- unlist(unlist(MCMC_all[,1]))
 gbar <- unlist(unlist(MCMC_all[,3]))
 kappa <- unlist(unlist(MCMC_all[,5]))
 
-r_narrow <- (r + median(r))/2
-
 compfun <- function(x) {
   dd <- data.frame(
     est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
@@ -59,12 +57,4 @@ g1 <- ggplot(dd_combine) +
   xlab("Uncertainty type") +
   scale_y_continuous("Basic reproductive number", limits=c(2, 6))
 
-dd_combine2 <- compfun(r_narrow)
-
-g2 <- g1 %+% dd_combine2
-
-gtot <- arrangeGrob(g1 + ggtitle("A. Jan 24th - Jan 26th"),
-            g2 + ggtitle("B. Speculative scenario"),
-            nrow=1)
-
-ggsave("figure2.pdf", gtot, width=8, height=3)
+ggsave("figure2.pdf", g1, width=6, height=4)

### 2020-01-30 wuhan.tex
More ambitious intro

@@
 
-We evaluate six disparate models posted online in daterange that estimated \Ro\ for the 2019-nCoV outbreak. <C-F3>
-Our results indicate that most early estimates of $\mathcal R_0$ for the outbreak
-are likely to have been overly confident.
+We evaluate six disparate models posted online between January 24--26, 2020 that estimated \Ro\ for the 2019-nCoV outbreak.
+We use our framework to construct pooled estimates for the underlying parameters.
+We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
+We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
+
+Our results indicate that many early estimates of \Ro\ were likely overly confident.
 

### 2020-01-30 compare_assumption.R
updating figs

@@ -19,50 +19,69 @@ rdata2 <- sample_data %>%
   bind_rows(.id="study")
 
 avg_r <- data.frame(
-  median=median(unlist(unlist(MCMC_all[,1]))),
+  est=median(unlist(unlist(MCMC_all[,1]))),
   lwr=quantile(unlist(unlist(MCMC_all[,1])), 0.025),
   upr=quantile(unlist(unlist(MCMC_all[,1])), 0.975),
-  study="Average"
+  anon="Pooled estimate"
 )
 
 avg_gen <- data.frame(
-  median=median(unlist(unlist(MCMC_all[,3]))),
+  est=median(unlist(unlist(MCMC_all[,3]))),
   lwr=quantile(unlist(unlist(MCMC_all[,3])), 0.025),
   upr=quantile(unlist(unlist(MCMC_all[,3])), 0.975),
-  study="Average"
+  anon="Pooled estimate"
 )
 
 avg_kappa <- data.frame(
-  median=median(unlist(unlist(MCMC_all[,5]))),
+  est=median(unlist(unlist(MCMC_all[,5]))),
   lwr=quantile(unlist(unlist(MCMC_all[,5])), 0.025),
   upr=quantile(unlist(unlist(MCMC_all[,5])), 0.975),
-  study="Average"
+  anon="Pooled estimate"
 )
 
-g1 <- ggplot(merge(rdata2, study_number)) +
-  geom_point(aes(est, anon)) +
-  geom_segment(aes(lwr, anon, xend=upr, yend=anon)) +
-  geom_point(data=avg_r, aes(median, study), col=2) +
-  geom_segment(data=avg_r, aes(lwr, study, xend=upr, yend=study), col=2) +
+rdata3 <- mutate(merge(rdata2, study_number)) %>%
+  bind_rows(avg_r) %>%
+  mutate(
+    anon=factor(anon, levels=c("Pooled estimate", paste0("Study ", 6:1)))
+  )
+
+g1 <- ggplot(rdata3) +
+  geom_point(aes(est, anon), col=c(1, 1, 1, 1, 1, 1, 2),
+             size=c(1, 1, 1, 1, 1, 1, 4)) +
+  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, 1, 1, 1, 1, 2),
+               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 2)) +
[…]

### 2020-01-30 study_number.R

@@ -1,5 +1,5 @@
 study_number <- data.frame(
   anon=paste0("Study ", 1:6),
-  study=c("Imai et al. (2020)", "Liu et al. (2020)", "Majumder et al. (2020)", 
-          "Read et al. (2020)", "Riou and Althaus (2020)", "Zhao et al. (2020)")
+  study=c("Imai et al. (2020)", "Liu et al. (2020)", 
+          "Read et al. (2020)", "Riou and Althaus (2020)", "Zhao et al. (2020)", "Majumder et al. (2020)")
 )

### 2020-01-30 wuhan.tex
delete comments

@@
 posing a greater threat for its spread.
-\jd{It is spread primarily from person to person; that's been obvious for a while and is now confirmed. We should find a better cite and leave Viet Nam out of it.}
 
@@
 
-\jd{Fix exponential?}
 We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same Gamma distribution:
@@
 this approach is analogous to Bayesian methods for analyzing phylogenetic data, which often rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
-\jd{use mean (not rate) and shape; specify units for the means. Use this as an opportunity to question your choices maybe. Consider using 1 week and 1/week for the means.}
 Weakly informative priors are assumed on the hyperparameters:
@@
 \begin{aligned}
-\mu_r &\sim \mathrm{Gamma}(\mathrm{mean}=0.1\,\mathrm{day}^{-1},\,\mathrm{shape}=0.1)\\
+\mu_r &\sim \mathrm{Gamma}(\mathrm{mean}=0.1\,\mathrm{days}^{-1},\,\mathrm{shape}=0.1)\\
 \mu_G &\sim \mathrm{Gamma}(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\
@@
 
-\jd{Change speculative to hypothetical.}
 \begin{figure}[!ht]

### 2020-01-30 wuhan.tex
new citation

@@
 
-Here, we use the gamma approximation framework \citep{park2019practical} to (1) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of $\mathcal R_0$.
-\jd{We should cite another group as well about this approximation.}
+Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (1) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of $\mathcal R_0$.
 Assuming that generation intervals follow a gamma distribution

### 2020-01-30 compare_R0.R
new fig

@@ -13,14 +13,15 @@ R0all <- Rdata_adj %>%
     type="base"
   ) %>%
   bind_rows(
-    R0_r, R0_gbar, R0_kappa
+    R0_r, R0_gbar, R0_kappa,
+    filter(R0_all, study=="Imai et al. (2020)")
   ) %>%
   mutate(
     type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
                 labels=c("base", "growth rate", "GI mean", "GI variation", "all"))
   ) %>%
   merge(study_number)
-
+  
 R0all %>%
   filter(type!="all", type!="base") %>%
   mutate(
@@ -45,8 +46,8 @@ R0all %>%
   )
 
 g1 <- ggplot(R0all) +
-  geom_hline(data=R0_all, aes(yintercept=lwr), lty=2) +
-  geom_hline(data=R0_all, aes(yintercept=upr), lty=2) +
+  geom_hline(data=R0_all, aes(yintercept=lwr), lty=2, col="#C77CFF") +
+  geom_hline(data=R0_all, aes(yintercept=upr), lty=2, col="#C77CFF") +
   geom_point(aes(anon, est, col=type), position=position_dodge(0.5)) +
   geom_errorbar(aes(anon, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),
                 width=0) +

### 2020-01-30 wuhan.tex
First part of new Results

--- 2020-01-30 wuhan.tex
@@
 Our results indicate that many early estimates of \Ro\ were likely overly confident.
+
+\section{newResults}
+
+\begin{table}[t]
+\begin{center}
+\footnotesize
+\begin{tabular}{l|l|l|l}
+ & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
+\hline
+Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ \\
+\hline
+Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 \\
+\hline
+Study 3 & 2.0--3.1 & 6--10 & 0 \\
+\hline
+Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 \\
+\hline
+Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5\\
+\hline
+Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2
+\end{tabular}
+\end{center}
+\caption{
+\textbf{Estimates of the basic reproductive number and their assumptions about the generation-interval distributions.}
+$^\ast$We treated these intervals as a 95\% confidence interval in our analysis.
+$^\dagger$We assumed $\kappa = 0.5$ in our analysis.
+$^\ddagger$The authors presented $\mathcal R_0$ estimates under different assumptions; we used their baseline scenario in our analysis.
+}
+\end{table}
+
+We gathered information on estimates of $\mathcal R_0$ and their assumptions about the underlying generation-interval distributions from 6 articles that were uploaded to preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
+For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a gamma distribution, which matched the assumptions of all of the studies).
+We modeled values reported using gamma distributions or confidence intervals with gamma distributions, and values reported using uniform distributions with uniform distributions \jd{SWP pls check}.
+For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
+we model this parameter as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
+If a study assumed a fixed value for a parameter, we hold that parameters constant as well, using the same assumed value. \jd{I kind of want to drop this sentence.}
+
+\begin{table}[t]
+\begin{center}
+\footnotesize
+\begin{tabular}{l|l|l|l}
+ & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
+\hline
+Study 1 & $\mathrm{Gamma}(\alpha=28, \beta=28/2.6)$ & constant & constant \\
+\hline
+Study 2 & $\mathrm{Gamma}(\alpha=67, \beta=67/2.92)$ & constant & constant \\
+\hline
+Study 3 & N.A. & $\mathrm{Uniform}(6, 10)$ & constant\\
+\hline
+Study 4 & $\mathrm{Gamma}(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
+\hline
+Study 5 & $\mathrm{Gamma}(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
+\hline
+Study 6 & $\mathrm{Gamma}(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant
+\end{tabular}
+\end{center}
+\caption{
+\textbf{Probability distributions for the estimates of $\mathcal R_0$ and for the assumed values of $\bar G$ and $\kappa$.}
+The probability distribution for the exponential growth rate $r$ for each study is obtained by drawing random samples from these distributions and taking the inverse of \eref{gamma}.
+The gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
+Constant distributions assume a fixed values according to Table 1.
+}
+\end{table}
+
+We then constructed a family of parameter sets for each study, and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{compare_assumption.pdf}
+\caption{
+\textbf{Comparisons of estimated and assumed parameter values with their average values calculated from the multilevel model.}
+I'm a caption; \emph{no, you're not.}
+}
+\label{fig:assumption}
+\end{figure}
+
+\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared CV $\kappa$ of the generation-interval distribution with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+We find that there is a large uncertainty associated with the underlying parameters;
+many models rely on stronger assumptions that ignore these uncertainties.
+Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
+
+\begin{figure}[!ht]
+\includegraphics[width=\textwidth]{figure2.pdf}
+\caption{
+\textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of $\mathcal R_0$.}
+(base): $\mathcal R_0$ estimate based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+(growth rate): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_r$ while assuming median estimates of $\mu_G$ and $\mu_\kappa$.
+(GI mean): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_G$ while assuming median estimates of $\mu_r$ and $\mu_\kappa$.
+(growth rate + GI mean): $\mathcal R_0$ estimate based on the the full posterior distributions of $\mu_r$ and $\mu_G$ while assuming a median estimate of $\mu_\kappa$.
+(all): $\mathcal R_0$ estimate based on the the full posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+}
+\label{fig:eff}
+\end{figure}
+
+\fref{eff} demonstrates the degree to which how the uncertainty in each parameter propogates to the uncertainties in $\mathcal R_0$.
+\swp{JD: can you explain figure 2?}
+
+Finally, we assess the validity of estimates of $\mathcal R_0$ across different studies by 
+replacing the their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (16 out 18).
+We estimate narrower confidence intervals for Study 2 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0$);
+when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$.
+We estimate narrower confidence intervals for Study 5 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
+
+% We find that accounting for uncertainties in the estimate of $r$ (i.e., replacing the estimated $r$ values with $\mu_r$) has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
+% For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+% There are two explanations for this result.
+% First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
+% $r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ (95\% CI: 0.12--0.28 days$^{-1}$) than with $\mu_G$ (95\% CI: 6.6--10.2 days).
+% Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
+
+\begin{figure}[t]
+\includegraphics[width=\textwidth]{compare_R0.pdf}
+\caption{
+\textbf{Effect of uncertainties in underlying parameters on the estimate of the basic reproductive number.}
+I'm a caption
+}
+\label{fig:R0}
+\end{figure}
+
+When we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
+We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
+While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
+This result does not imply that our assumptions are too weak;
+we believe that this confidence interval accurately reflects the level of uncertainties present in the information that were available when these models were fitted.
+In fact, our results show that incorporating uncertainties in all parameters can give narrower confidence intervals than doing so one at a time because nonlinear effects of parameters could cancel out and give appropriately wide confidence intervals.
+
+\section{Discussion}
+
+Estimating the basic reproductive number $\mathcal R_0$ is crucial for predicting the course of an outbreak and planning intervention strategies.
+Here, we evaluated the statistical validity of the estimates of $\mathcal R_0$ for the novel coronavirus outbreak using a novel statistical framework, based on the gamma approximation of the generation-interval distribution \citep{park2019practical}.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, especially with the amount of dispersion in the generation intervals:
+assuming a narrow generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.
+
+Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approches that the estimates rely on.
+Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
+Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, 
+fitting a model to cumulative number of cases can bias the parameter estimate and give overly narrow confidence intervals as it violates the ``independence of error'' assumption \citep{ma2014estimating, king2015avoidable}.
+We emphasize that all models should be fitted to raw incidence data for any future analyses.
+
+Concluding paragraph with a message?
 
@@
 
-\section{Results}
-
-\jd{Change speculative to hypothetical.}
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{compare_assumption.pdf}
-\caption{
-\textbf{Comparisons of estimated and assumed parameter values with their average values calculated from the multilevel model.}
-I'm a caption
-}
-\label{fig:assumption}
-\end{figure}
-
-\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared CV $\kappa$ of the generation-interval distribution with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
-We find that there is a large uncertainty associated with the underlying parameters;
-many models rely on stronger assumptions that ignore these uncertainties.
-Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
-
-\begin{figure}[!ht]
-\includegraphics[width=\textwidth]{figure2.pdf}
-\caption{
-\textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of $\mathcal R_0$.}
-(base): $\mathcal R_0$ estimate based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-(growth rate): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_r$ while assuming median estimates of $\mu_G$ and $\mu_\kappa$.
-(GI mean): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_G$ while assuming median estimates of $\mu_r$ and $\mu_\kappa$.
-(growth rate + GI mean): $\mathcal R_0$ estimate based on the the full posterior distributions of $\mu_r$ and $\mu_G$ while assuming a median estimate of $\mu_\kappa$.
-(all): $\mathcal R_0$ estimate based on the the full posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-}
-\label{fig:eff}
-\end{figure}
-
-\fref{eff} demonstrates the degree to which how the uncertainty in each parameter propogates to the uncertainties in $\mathcal R_0$.
-\swp{JD: can you explain figure 2?}
-
-Finally, we assess the validity of estimates of $\mathcal R_0$ across different studies by 
-replacing the their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (16 out 18).
-We estimate narrower confidence intervals for Study 2 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0$);
-when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$.
-We estimate narrower confidence intervals for Study 5 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
-
-% We find that accounting for uncertainties in the estimate of $r$ (i.e., replacing the estimated $r$ values with $\mu_r$) has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
-% For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
-% There are two explanations for this result.
-% First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
-% $r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ (95\% CI: 0.12--0.28 days$^{-1}$) than with $\mu_G$ (95\% CI: 6.6--10.2 days).
-% Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
-
-\begin{figure}[t]
-\includegraphics[width=\textwidth]{compare_R0.pdf}
-\caption{
-\textbf{Effect of uncertainties in underlying parameters on the estimate of the basic reproductive number.}
-I'm a caption
-}
-\label{fig:R0}
-\end{figure}
-
-When we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
-We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
-While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
-This result does not imply that our assumptions are too weak;
-we believe that this confidence interval accurately reflects the level of uncertainties present in the information that were available when these models were fitted.
-In fact, our results show that incorporating uncertainties in all parameters can give narrower confidence intervals than doing so one at a time because nonlinear effects of parameters could cancel out and give appropriately wide confidence intervals.
-
-\section{Discussion}
-
-Estimating the basic reproductive number $\mathcal R_0$ is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we evaluated the statistical validity of the estimates of $\mathcal R_0$ for the novel coronavirus outbreak using a novel statistical framework, based on the gamma approximation of the generation-interval distribution \citep{park2019practical}.
-Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, especially with the amount of dispersion in the generation intervals:
-assuming a narrow generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.
-
-Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approches that the estimates rely on.
-Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, 
-fitting a model to cumulative number of cases can bias the parameter estimate and give overly narrow confidence intervals as it violates the ``independence of error'' assumption \citep{ma2014estimating, king2015avoidable}.
-We emphasize that all models should be fitted to raw incidence data for any future analyses.
-
-Concluding paragraph with a message?
-
 \pagebreak

### 2020-01-30 wuhan.tex
captions

--- 2020-01-30 wuhan.tex
@@
 \begin{tabular}{l|l|l|l}
- & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
+ & Basic reproductive number & Mean generation time & CV$^2$ generation time \\
 \hline
-Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ \\
+Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 days & unspecified$^\dagger$ \\
 \hline
-Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 \\
+Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 days & 0.2 \\
 \hline
-Study 3 & 2.0--3.1 & 6--10 & 0 \\
+Study 3 & 2.0--3.1 & 6--10 days & 0 \\
 \hline
-Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 \\
+Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 \\
 \hline
-Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5\\
+Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5\\
 \hline
-Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2
+Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2\\
+\hline
 \end{tabular}
@@
 \textbf{Estimates of the basic reproductive number and their assumptions about the generation-interval distributions.}
-$^\ast$We treated these intervals as a 95\% confidence interval in our analysis.
-$^\dagger$We assumed $\kappa = 0.5$ in our analysis.
-$^\ddagger$The authors presented $\mathcal R_0$ estimates under different assumptions; we used their baseline scenario in our analysis.
+Estimates of $\mathcal R_0$ and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
+$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
+$^\dagger$We assume $\kappa = 0.5$ in our analysis.
+$^\ddagger$The authors presented $\mathcal R_0$ estimates under different assumptions; we use their baseline scenario in our analysis.
 }

### 2020-01-30 wuhan.tex
working on text

--- 2020-01-30 wuhan.tex
@@
 \caption{
-\textbf{Probability distributions for the estimates of $\mathcal R_0$ and for the assumed values of $\bar G$ and $\kappa$.}
+\textbf{Probability distributions $\mathcal R_0$, $\bar G$, and $\kappa$.}
+We use these 
 The probability distribution for the exponential growth rate $r$ for each study is obtained by drawing random samples from these distributions and taking the inverse of \eref{gamma}.

### 2020-01-30 wuhan.tex
suddenly work on other stuff

--- 2020-01-30 wuhan.tex
@@
 
-\section{newResults}
+\section{Results}
 
@@
 \caption{
-\textbf{Comparisons of estimated and assumed parameter values with their average values calculated from the multilevel model.}
-I'm a caption; \emph{no, you're not.}
+\textbf{Parameter values.}
+We inferred point estimates, uniform distributions (shown in orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (see text).
 }
@@
 Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
+It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval.
 
@@
 
-\fref{eff} demonstrates the degree to which how the uncertainty in each parameter propogates to the uncertainties in $\mathcal R_0$.
-\swp{JD: can you explain figure 2?}
+\fref{eff} shows the effects of propagating uncertainty in different combinations would affect estimates and CIs for our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
+Propagating error from the growth rate (which all but one of the models did) is absolutely crucial. 
+Propagating error from the generation interval also has important effects, however.
+Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example.
 
-Finally, we assess the validity of estimates of $\mathcal R_0$ across different studies by 
+We also evaluate the estimates of $\mathcal R_0$ across different studies by 
 replacing the their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals in most cases (16 out 18).
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but two cases.
 We estimate narrower confidence intervals for Study 2 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0$);

### 2020-01-30 compare_R0.R
pushing figs

@@ -14,13 +14,16 @@ R0all <- Rdata_adj %>%
   ) %>%
   bind_rows(
     R0_r, R0_gbar, R0_kappa,
-    filter(R0_all, study=="Imai et al. (2020)")
+    mutate(filter(R0_all, study=="Imai et al. (2020)"))
   ) %>%
   mutate(
     type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
                 labels=c("base", "growth rate", "GI mean", "GI variation", "all"))
   ) %>%
-  merge(study_number)
+  merge(study_number) %>%
+  mutate(
+    anon=ifelse(type=="all", "Pooled estimate", anon)
+  )
   
 R0all %>%
   filter(type!="all", type!="base") %>%

### 2020-01-30 wuhan.tex
pushing figs
Next part of Results

@@
 when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$.
+\jd{don't know what this last part means.
+Doesn't lower kappa mean exactly the opposite?
+It's what's making the estimate volatile?}
 We estimate narrower confidence intervals for Study 5 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
@@
 \textbf{Effect of uncertainties in underlying parameters on the estimate of the basic reproductive number.}
-I'm a caption
+I'm a caption; \emph{No you're not.}
 }
@@
 While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
-This result does not imply that our assumptions are too weak;
+This result does not imply that assumptions based on the pooled estimate are too weak; \jd{better in Discussion?}
 we believe that this confidence interval accurately reflects the level of uncertainties present in the information that were available when these models were fitted.
 In fact, our results show that incorporating uncertainties in all parameters can give narrower confidence intervals than doing so one at a time because nonlinear effects of parameters could cancel out and give appropriately wide confidence intervals.
+\jd{Are you sure?
+I don't think we have a mechanism for cancelling; I think that the narrow GIs cause instability.}
 

### 2020-01-30 wuhan.tex
tweaks

--- 2020-01-30 wuhan.tex
@@
 We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
-
 Our results indicate that many early estimates of \Ro\ were likely overly confident.
@@
 For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a gamma distribution, which matched the assumptions of all of the studies).
-We modeled values reported using gamma distributions or confidence intervals with gamma distributions, and values reported using uniform distributions with uniform distributions \jd{SWP pls check}.
+We modeled values reported using confidence intervals with gamma distributions, and values reported using ranges with uniform distributions.
 For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
 we model this parameter as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
-If a study assumed a fixed value for a parameter, we hold that parameters constant as well, using the same assumed value. \jd{I kind of want to drop this sentence.}
+We then constructed a family of parameter sets for each study, and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
 
@@
 \hline
-Study 3 & N.A. & $\mathrm{Uniform}(6, 10)$ & constant\\
+Study 3 & N.A.$^\ast$ & $\mathrm{Uniform}(6, 10)$ & constant\\
 \hline
@@
 Constant distributions assume a fixed values according to Table 1.
+$^\ast$Study 3 effectively fits an exponential curve to the cumulative number of confirmed cases.
+\emph{All} uncertainty in their $\mathcal R_0$ estimate can be explained by the uncertainty range they consider for $\bar G$.
+Instead of modeling $\mathcal R_0$ with a probability distribution and recalculating $r$, we simply fix $r=0.114\,\mathrm{days}^{-1}$.
 }
 \end{table}
-
-We then constructed a family of parameter sets for each study, and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
 
@@
 \begin{aligned}
-\mu_r &\sim \mathrm{Gamma}(\mathrm{mean}=0.1\,\mathrm{day}^{-1},\,\mathrm{shape}=0.1)\\
+\mu_r &\sim \mathrm{Gamma}(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=0.1)\\
 \mu_G &\sim \mathrm{Gamma}(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\

### 2020-01-30 wuhan.tex
Rewrite Discussion; nothing added

--- 2020-01-30 wuhan.tex
@@
 Estimating the basic reproductive number $\mathcal R_0$ is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we evaluated the statistical validity of the estimates of $\mathcal R_0$ for the novel coronavirus outbreak using a novel statistical framework, based on the gamma approximation of the generation-interval distribution \citep{park2019practical}.
-Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, especially with the amount of dispersion in the generation intervals:
-assuming a narrow generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.
+Here, we used a simple framework \citep{park2019practical} to compare  estimates of $\mathcal R_0$ for the novel coronavirus outbreak.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including with the amount of dispersion in the generation intervals:
+although our pooled estimates are relatively insensitive to the estimated uncertainty, our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.
 
-Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approches that the estimates rely on.
+Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approches used to estimate the exponential growth rate $r$.
 Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, 
-fitting a model to cumulative number of cases can bias the parameter estimate and give overly narrow confidence intervals as it violates the ``independence of error'' assumption \citep{ma2014estimating, king2015avoidable}.
-We emphasize that all models should be fitted to raw incidence data for any future analyses.
-
-Concluding paragraph with a message?
+Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, fitting a model to cumulative number of without accounting for problems with the dependent error structure can both bias parameter narrow confidence intervals \citep{ma2014estimating, king2015avoidable}.
+Naive fits to cumulative incidence data should be avoided.
 

### 2020-01-30 wuhan.tex
a new Discussion ¶
tweak

--- 2020-01-30 wuhan.tex
@@
 
-\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared CV $\kappa$ of the generation-interval distribution with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared coefficient of variation (CV) $\kappa$ of the generation-interval distribution with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
 We find that there is a large uncertainty associated with the underlying parameters;
@@
 \mu_G &\sim \mathrm{Gamma}(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\
-(\mu_\kappa, \alpha_r, \alpha_G, \alpha_\kappa) &\sim \mathrm{Gamma}(\mathrm{mean}=1,\,\mathrm{shape}=0.1).
+\mu_G &\sim \mathrm{Gamma}(\mathrm{mean}=0.5,\,\mathrm{shape}=0.1)\\
+(\alpha_r, \alpha_G, \alpha_\kappa) &\sim \mathrm{Gamma}(\mathrm{mean}=1,\,\mathrm{shape}=0.1).
 \end{aligned}

### 2020-01-30 wuhan.tex

--- 2020-01-30 wuhan.tex
@@
 Naive fits to cumulative incidence data should be avoided.
+
+We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components.
+We are hopeful that this will provide a guide to understanding and reconciling different estimates early in an epidemic.
 

### 2020-01-30 wuhan.tex
Title

--- 2020-01-30 wuhan.tex
@@
 	\textbf\newline{
-		Evaluating uncertainties associated with preliminary estimates of the basic reproductive number during the novel coronavirus (2019-nCoV) outbreak
+		Reconciling early-outbreak preliminary estimates of the basic reproductive number: a new framework and applications to the novel coronavirus (2019-nCoV) outbreak
 	}
@@
 
+We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components.
+We are hopeful that this will provide a guide to understanding and reconciling different estimates early in an epidemic.
+
 \section{Methods}

### 2020-01-30 MCMC_all.R
push

@@ -4,9 +4,10 @@ load("sample_data.rda")
 nsample <- 100000
 
 priorfun <- function(theta) {
-  sum(dgamma(theta[-c(1, 3)], 0.1, 0.1/1, log=TRUE)) +
+  sum(dgamma(theta[-c(1, 3, 5)], 0.1, 0.1/1, log=TRUE)) +
     dgamma(theta[1], 0.1, 0.1*7, log=TRUE) +
-    dgamma(theta[3], 0.1, 0.1/7, log=TRUE)
+    dgamma(theta[3], 0.1, 0.1/7, log=TRUE) +
+    dgamma(theta[5], 0.1, 0.1/0.5, log=TRUE)
 }
 
 sample_data$`Majumder et al. (2020)`$kappa <- 0.02
@@ -43,7 +44,7 @@ postfun <- function(theta) {
 }
 
 V <- matrix(0, 6, 6)
-diag(V) <- c(5e-3, 50, 0.5, 100, 0.2, 1)
+diag(V) <- c(2e-3, 50, 1, 200, 0.1, 0.5)/10
 
 reslist <- vector('list', 4)
 

### 2020-01-30 figure2.R

@@ -55,6 +55,6 @@ g1 <- ggplot(dd_combine) +
   geom_point(aes(type, est)) +
   geom_errorbar(aes(type, min=lwr, max=upr), width=0) +
   xlab("Uncertainty type") +
-  scale_y_continuous("Basic reproductive number", limits=c(2, 5.6), expand=c(0, 0))
+  scale_y_continuous("Basic reproductive number", limits=c(2, 6), expand=c(0, 0))
 
 ggsave("figure2.pdf", g1, width=6, height=4)

### 2020-01-30 wuhan.tex
tweaking

@@
 there is clear evidence that it can be directly tranmitted between humans,
-posing a greater threat for its spread.
+posing a greater threat for its spread \citep{huang2020clinical}.
 
 As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number $\mathcal R_0$ (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+Estimating the basic reproductive number is of particular interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
@@
 Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We argue that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimate uncertainty should propagate error from all three sources.
+We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimate uncertainty should propagate error from all three sources.
 
@@
 We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
-Our results indicate that many early estimates of \Ro\ were likely overly confident.
 
@@
 \hline
-Study 3 & 2.0--3.1 & 6--10 days & 0 \\
+Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 \\
 \hline
-Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 \\
+Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5\\
 \hline
-Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5\\
+Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2\\
 \hline
-Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2\\
+Study 6 & 2.0--3.1 & 6--10 days & 0 \\
 \hline
@@
 \hline
-Study 3 & N.A.$^\ast$ & $\mathrm{Uniform}(6, 10)$ & constant\\
+Study 3 & $\mathrm{Gamma}(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
 \hline
-Study 4 & $\mathrm{Gamma}(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
+Study 4 & $\mathrm{Gamma}(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
 \hline
-Study 5 & $\mathrm{Gamma}(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
+Study 5 & $\mathrm{Gamma}(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant\\
 \hline
-Study 6 & $\mathrm{Gamma}(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant
+Study 6 & N.A.$^\ast$ & $\mathrm{Uniform}(6, 10)$ & constant\\
+\hline
 \end{tabular}
@@
 \caption{
-\textbf{Probability distributions $\mathcal R_0$, $\bar G$, and $\kappa$.}
+\textbf{Probability distributions for $\mathcal R_0$, $\bar G$, and $\kappa$.}
 We use these 
@@
 Constant distributions assume a fixed values according to Table 1.
-$^\ast$Study 3 effectively fits an exponential curve to the cumulative number of confirmed cases.
-\emph{All} uncertainty in their $\mathcal R_0$ estimate can be explained by the uncertainty range they consider for $\bar G$.
-Instead of modeling $\mathcal R_0$ with a probability distribution and recalculating $r$, we simply fix $r=0.114\,\mathrm{days}^{-1}$.
+$^\ast$Study 6 effectively fits an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+Instead of modeling $\mathcal R_0$ with a probability distribution and recalculating $r$, we simply fix $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in $\mathcal R_0$ they report, when combined with the range of $\bar G$ they consider.
 }
@@
 
-\begin{figure}[!ht]
+\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared coefficient of variation (CV) $\kappa$ of the generation-interval distribution from different studies with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+We find that there is a large uncertainty associated with the underlying parameters;
+many models rely on stronger assumptions that ignore these uncertainties.
+Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
+It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval.
+
+\begin{figure}[t]
 \includegraphics[width=\textwidth]{compare_assumption.pdf}
 \caption{
-\textbf{Parameter values.}
+\textbf{Comparisons of the reported parameter values with our pooled estimates.}
 We inferred point estimates, uniform distributions (shown in orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (see text).
+Study 1 did not 
 }
@@
 
-\fref{assumption} compares the estimated (or assumed) values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared coefficient of variation (CV) $\kappa$ of the generation-interval distribution with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
-We find that there is a large uncertainty associated with the underlying parameters;
-many models rely on stronger assumptions that ignore these uncertainties.
-Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
-It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval.
+\fref{eff} shows the effects of propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
+Propagating error from the growth rate (which all but one of the models did) is absolutely crucial. 
+Propagating error from the generation interval also has important effects, however.
+Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example.
 
@@
 
-\fref{eff} shows the effects of propagating uncertainty in different combinations would affect estimates and CIs for our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
-Propagating error from the growth rate (which all but one of the models did) is absolutely crucial. 
-Propagating error from the generation interval also has important effects, however.
-Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example.
-
 We also evaluate the estimates of $\mathcal R_0$ across different studies by 
 replacing the their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but two cases.
-We estimate narrower confidence intervals for Study 2 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0$);
-when lower values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$.
-\jd{don't know what this last part means.
-Doesn't lower kappa mean exactly the opposite?
-It's what's making the estimate volatile?}
-We estimate narrower confidence intervals for Study 5 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.
+We estimate narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0.2$ and $0$ respectively);
+when higher values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
+We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
 
@@
 
-When we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
+Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
 We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
@@
 
-\begin{table}[t]
-\begin{center}
-\footnotesize
-\begin{tabular}{l|l|l|l}
- & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
-\hline
-Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ \\
-\hline
-Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 \\
-\hline
-Study 3 & 2.0--3.1 & 6--10 & 0 \\
-\hline
-Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 \\
-\hline
-Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5\\
-\hline
-Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2
-\end{tabular}
-\end{center}
-\caption{
-\textbf{Estimates of the basic reproductive number and their assumptions about the generation-interval distributions.}
-$^\ast$We treated these intervals as a 95\% confidence interval in our analysis.
-$^\dagger$We assumed $\kappa = 0.5$ in our analysis.
-$^\ddagger$The authors presented $\mathcal R_0$ estimates under different assumptions; we used their baseline scenario in our analysis.
-}
-\end{table}
-
 First, we gathered information on estimates of $\mathcal R_0$ and their
@@
 This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.
-
-\begin{table}[t]
-\begin{center}
-\footnotesize
-\begin{tabular}{l|l|l|l}
- & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
-\hline
-Study 1 & $\mathrm{Gamma}(\alpha=28, \beta=28/2.6)$ & constant & constant \\
-\hline
-Study 2 & $\mathrm{Gamma}(\alpha=67, \beta=67/2.92)$ & constant & constant \\
-\hline
-Study 3 & N.A. & $\mathrm{Uniform}(6, 10)$ & constant\\
-\hline
-Study 4 & $\mathrm{Gamma}(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
-\hline
-Study 5 & $\mathrm{Gamma}(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
-\hline
-Study 6 & $\mathrm{Gamma}(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant
-\end{tabular}
-\end{center}
-\caption{
-\textbf{Probability distributions for the estimates of $\mathcal R_0$ and for the assumed values of $\bar G$ and $\kappa$.}
-The probability distribution for the exponential growth rate $r$ for each study is obtained by drawing random samples from these distributions and taking the inverse of \eref{gamma}.
-The gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
-Constant distributions assume a fixed values according to Table 1.
-}
-\end{table}
 

### 2020-01-30 compare_R0.R

@@ -22,7 +22,7 @@ R0all <- Rdata_adj %>%
   ) %>%
   merge(study_number) %>%
   mutate(
-    anon=ifelse(type=="all", "Pooled estimate", anon)
+    anon=ifelse(type=="all", "Pooled estimate", as.character(anon))
   )
   
 R0all %>%
@@ -33,9 +33,6 @@ R0all %>%
   merge(summarize(group_by(Rdata_adj, study), bwidth=upr-lwr)) %>%
   mutate(
     ww=width>bwidth
-  ) %>%
-  summarize(
-    sum(ww)
   )
 
 R0all %>%

### 2020-01-30 compare_assumption.R

@@ -47,9 +47,9 @@ rdata3 <- mutate(merge(rdata2, study_number)) %>%
 
 g1 <- ggplot(rdata3) +
   geom_point(aes(est, anon), col=c(1, 1, 1, 1, 1, 1, 2),
-             size=c(1, 1, 1, 1, 1, 1, 4)) +
+             size=c(2, 2, 2, 2, 2, 2, 4)) +
   geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, 1, 1, 1, 1, 2),
-               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 2)) +
+               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.5)) +
   scale_x_continuous(expression(Exponential~growth~rate~(days^{-1}))) +
   theme(
     axis.title.y = element_blank()
@@ -63,9 +63,9 @@ gdata2 <- mutate(merge(gdata, study_number)) %>%
 
 g2 <- ggplot(gdata2) +
   geom_point(aes(est, anon), col=c(1, 1, 1, 1, 1, 1, 2),
-             size=c(1, 1, 1, 1, 1, 1, 4)) +
+             size=c(2, 2, 2, 2, 2, 2, 4)) +
   geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, 1, 1, 1, 1, 2),
-               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 2)) +
+               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.5)) +
   scale_x_continuous("Mean generation time (days)") +
   theme(
     axis.title.y = element_blank()
@@ -81,9 +81,9 @@ kappadata2 <- mutate(merge(kappadata, study_number)) %>%
 g3 <- ggplot(kappadata2) +
   geom_point(aes(est, anon), shape=c(2, 16, 16, 16, 16, 16, 16),
              col=c(1, 1, 1, 1, 1, 1, 2),
-             size=c(1, 1, 1, 1, 1, 1, 4)) +
+             size=c(2, 2, 2, 2, 2, 2, 4)) +
   geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, 1, 1, 1, 1, 2),
-               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 2)) +
[…]

### 2020-01-30 wuhan.tex
tweaking stuff

@@
 there is clear evidence that it can be directly tranmitted between humans,
-posing a greater threat for its spread \citep{huang2020clinical}.
+posing a greater threat for its spread \citep{huang2020clinical,who26report}.
 
@@
 \caption{
-\textbf{Estimates of the basic reproductive number and their assumptions about the generation-interval distributions.}
+\textbf{Estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
 Estimates of $\mathcal R_0$ and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
@@
 $^\ast$Study 6 effectively fits an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
-Instead of modeling $\mathcal R_0$ with a probability distribution and recalculating $r$, we simply fix $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in $\mathcal R_0$ they report, when combined with the range of $\bar G$ they consider.
+Instead of modeling $\mathcal R_0$ with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in $\mathcal R_0$ they report, when combined with the range of $\bar G$ they consider.
 }
@@
 We inferred point estimates, uniform distributions (shown in orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (see text).
-Study 1 did not 
+Black points and lines: reported parameter values and ranges of uncertainties considered.
+Black triangle: we assumed $\kappa=0.5$ for Study 1 as they do not report the parameter values that they used.
+Red points and lines: median and 95\% confidence intervals of our pooled estimates.
 }
@@
 \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of $\mathcal R_0$.}
-(base): $\mathcal R_0$ estimate based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-(growth rate): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_r$ while assuming median estimates of $\mu_G$ and $\mu_\kappa$.
-(GI mean): $\mathcal R_0$ estimate based on the the full posterior distribution of $\mu_G$ while assuming median estimates of $\mu_r$ and $\mu_\kappa$.
-(growth rate + GI mean): $\mathcal R_0$ estimate based on the the full posterior distributions of $\mu_r$ and $\mu_G$ while assuming a median estimate of $\mu_\kappa$.
-(all): $\mathcal R_0$ estimate based on the the full posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+We compare estimates of $\mathcal R_0$ under five scenarios that propagate different combinations of uncertainties.
+(base): $\mathcal R_0$ estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+(growth rate): $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
+(GI mean): $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
+(growth rate + GI mean): $\mathcal R_0$ estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
+(all): $\mathcal R_0$ estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+Vertical lines represent the 95\% confidence intervals.
 }
@@
 
+Here, we considered studies that were published between January 24th, 2020 and Janury 26th, 2020.
+As the epidemic continues and more data becomes available, 
+some of these analyses
+
 We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components.
@@
 will give a higher estimate of $\mathcal R_0$ for the same value of $r$.
-
-First, we gathered information on estimates of $\mathcal R_0$ and their
-assumptions about the underlying generation-interval distributions from
-6 articles that were uploaded to
-preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
-For each study $i$, we model $\mathcal R_{0i}$, $\bar G_i$, and $\kappa_i$ with appropriate probability distributions (either uniform or gamma) that roughly captures the reported values and their uncertainty ranges.
-For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
-we model this parameter as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for the list of distributions that we assume).
-If a study assumed a fixed value for a parameter, we hold that parameters constant as well, using the same assumed value.
 
@@
 Posterior samples are thinned every 400 steps.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all parameters \citep{gelman1992inference}.
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference}.
 

### 2020-01-30 wuhan.tex
maybe I can send it...

@@
 
-% We find that accounting for uncertainties in the estimate of $r$ (i.e., replacing the estimated $r$ values with $\mu_r$) has the largest effect on the estimates of $\mathcal R_0$ and the width of associated confidence intervals (\fref{R0}).
-% For example, recalculating $\mathcal R_0$ for Study 3 by using the average $r$ values gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
-% There are two explanations for this result.
-% First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
-% $r$ has greater overall effect on $\mathcal R_0$ because there is more uncertainty associated with our estimate of $\mu_r$ (95\% CI: 0.12--0.28 days$^{-1}$) than with $\mu_G$ (95\% CI: 6.6--10.2 days).
-% Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$.
+Consistent with our previous observations (\fref{eff}),
+we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ (\fref{R0}).
+For example, recalculating $\mathcal R_0$ for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+There are two explanations for this result.
+First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
+$r$ has a greater overall effect on $\mathcal R_0$ because it is associated with more uncertainty;
+the range of the 95\% confidence interval on $\mu_r$ spaans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
+Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$ as explained previously.
 
@@
 \caption{
-\textbf{Effect of uncertainties in underlying parameters on the estimate of the basic reproductive number.}
-I'm a caption; \emph{No you're not.}
+\textbf{Sensitivity of the reported $\mathcal R_0$ estimates with respect to our pooled estimates of the underlying parameters.}
+We replace the reported paramaeter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our correponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate $\mathcal R_0$.
+The pooled estimate of $\mathcal R_0$ is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$;
+horizontal dashed lines represent the 95\% confidence intervals of our pooled estimate of $\mathcal R_0$.
+The reported $\mathcal R_0$ estimates (base) have been adjusted to show the approximate 95\% confidence interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
 }
@@
 While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
-This result does not imply that assumptions based on the pooled estimate are too weak; \jd{better in Discussion?}
-we believe that this confidence interval accurately reflects the level of uncertainties present in the information that were available when these models were fitted.
-In fact, our results show that incorporating uncertainties in all parameters can give narrower confidence intervals than doing so one at a time because nonlinear effects of parameters could cancel out and give appropriately wide confidence intervals.
-\jd{Are you sure?
-I don't think we have a mechanism for cancelling; I think that the narrow GIs cause instability.}
+This result does not imply that assumptions based on the pooled estimate are too weak;
+we believe that this confidence interval accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
 
@@
 Naive fits to cumulative incidence data should be avoided.
-
-Here, we considered studies that were published between January 24th, 2020 and Janury 26th, 2020.
-As the epidemic continues and more data becomes available, 
-some of these analyses
 
@@
 
-For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions that represent the estimates of $\mathcal R_0$ and the assumed values of $\bar G$ and $\kappa$
-and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
+For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of $\mathcal R_0$ and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
 \begin{equation}
@@
 
-We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same Gamma distribution:
+We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
 \begin{equation}
@@
 this approach is analogous to Bayesian methods for analyzing phylogenetic data, which often rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
+Since the gamma distribution does not allow zeroes, we use $\kappa =0.02$ instead for Study 6.
+
 Weakly informative priors are assumed on the hyperparameters:

### 2020-01-30 wuhan.tex
last tweak before i send

@@
 As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number $\mathcal R_0$ (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-Estimating the basic reproductive number is of particular interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
@@
 \caption{
-\textbf{Estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
+\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
 Estimates of $\mathcal R_0$ and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
@@
 We gathered information on estimates of $\mathcal R_0$ and their assumptions about the underlying generation-interval distributions from 6 articles that were uploaded to preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
-For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a gamma distribution, which matched the assumptions of all of the studies).
+For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a gamma distribution).
 We modeled values reported using confidence intervals with gamma distributions, and values reported using ranges with uniform distributions.
@@
 \textbf{Probability distributions for $\mathcal R_0$, $\bar G$, and $\kappa$.}
-We use these 
-The probability distribution for the exponential growth rate $r$ for each study is obtained by drawing random samples from these distributions and taking the inverse of \eref{gamma}.
+We use these probability distributions to obtain a probability distribution for the exponential growth rate $r$.
 The gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
@@
 
-\begin{figure}[t]
+\begin{figure}[!th]
 \includegraphics[width=\textwidth]{compare_R0.pdf}

### 2020-01-30 wuhan.tex
* wuhan.tex: tweaks/spelling

@@
 \usepackage{color}
-
+\usepackage{xspace}
 \newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}} 
-\newcommand{\Ro}{\Rx{0}}
+\newcommand{\Ro}{\Rx{0}\xspace}
 \newcommand{\RR}{\ensuremath{{\mathcal R}}}
@@
 Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, 
-there is clear evidence that it can be directly tranmitted between humans,
+there is clear evidence that it can be directly transmitted between humans,
 posing a greater threat for its spread \citep{huang2020clinical,who26report}.
 
-As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number $\mathcal R_0$ (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
 Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
-However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
+However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
 
 Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimate uncertainty should propagate error from all three sources.
+We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
 
@@
 \textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
-Estimates of $\mathcal R_0$ and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
+Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
 $^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
 $^\dagger$We assume $\kappa = 0.5$ in our analysis.
-$^\ddagger$The authors presented $\mathcal R_0$ estimates under different assumptions; we use their baseline scenario in our analysis.
+$^\ddagger$The authors presented \Ro estimates under different assumptions; we use their baseline scenario in our analysis.
 }
@@
 
-We gathered information on estimates of $\mathcal R_0$ and their assumptions about the underlying generation-interval distributions from 6 articles that were uploaded to preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
-For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a gamma distribution).
-We modeled values reported using confidence intervals with gamma distributions, and values reported using ranges with uniform distributions.
-For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
-we model this parameter as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
+We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were uploaded to preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
+For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a Gamma distribution).
+We used Gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
+For example, Study 2 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
+we model this parameter as a Gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
 We then constructed a family of parameter sets for each study, and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
 
+\newcommand{\gammdist}{\mathrm{Gamma}}
 \begin{table}[t]
@@
 \hline
-Study 1 & $\mathrm{Gamma}(\alpha=28, \beta=28/2.6)$ & constant & constant \\
+Study 1 & $\gammdist(\alpha=28, \beta=28/2.6)$ & constant & constant \\
 \hline
-Study 2 & $\mathrm{Gamma}(\alpha=67, \beta=67/2.92)$ & constant & constant \\
+Study 2 & $\gammdist(\alpha=67, \beta=67/2.92)$ & constant & constant \\
 \hline
-Study 3 & $\mathrm{Gamma}(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
+Study 3 & $\gammdist(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
 \hline
-Study 4 & $\mathrm{Gamma}(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
+Study 4 & $\gammdist(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
 \hline
-Study 5 & $\mathrm{Gamma}(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant\\
+Study 5 & $\gammdist(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant\\
 \hline
@@
 \caption{
-\textbf{Probability distributions for $\mathcal R_0$, $\bar G$, and $\kappa$.}
+\textbf{Probability distributions for \Ro, $\bar G$, and $\kappa$.}
 We use these probability distributions to obtain a probability distribution for the exponential growth rate $r$.
-The gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
-Constant distributions assume a fixed values according to Table 1.
+The Gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
+Constant distributions assume fixed values according to Table 1.
 $^\ast$Study 6 effectively fits an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
@@
 many models rely on stronger assumptions that ignore these uncertainties.
-Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
-It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval.
+Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of \Ro, even though the assumed values range from 0 to 0.5 across different studies.
+It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., CV$^2$ equal to 1).
 
@@
 $r$ has a greater overall effect on $\mathcal R_0$ because it is associated with more uncertainty;
-the range of the 95\% confidence interval on $\mu_r$ spaans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
+the range of the 95\% confidence interval on $\mu_r$ spans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
 Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$ as explained previously.
@@
 \textbf{Sensitivity of the reported $\mathcal R_0$ estimates with respect to our pooled estimates of the underlying parameters.}
-We replace the reported paramaeter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our correponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate $\mathcal R_0$.
+We replace the reported parameter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate $\mathcal R_0$.
 The pooled estimate of $\mathcal R_0$ is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$;
@@
 
-Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approches used to estimate the exponential growth rate $r$.
+Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approaches used to estimate the exponential growth rate $r$.
 Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
@@
 \begin{aligned}
-r_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
-\bar{G}_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
-\kappa_i &\sim \mathrm{Gamma}(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\alpha_\kappa).\\
+r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
+\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
+\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\alpha_\kappa).\\
 \end{aligned}
@@
 this approach is analogous to Bayesian methods for analyzing phylogenetic data, which often rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
-Since the gamma distribution does not allow zeroes, we use $\kappa =0.02$ instead for Study 6.
+Since the gamma distribution does not allow zeros, we use $\kappa =0.02$ instead for Study 6.
 
@@
 \begin{aligned}
-\mu_r &\sim \mathrm{Gamma}(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=0.1)\\
-\mu_G &\sim \mathrm{Gamma}(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\
-\mu_G &\sim \mathrm{Gamma}(\mathrm{mean}=0.5,\,\mathrm{shape}=0.1)\\
-(\alpha_r, \alpha_G, \alpha_\kappa) &\sim \mathrm{Gamma}(\mathrm{mean}=1,\,\mathrm{shape}=0.1).
+\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=0.1)\\
+\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\
+\mu_G &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=0.1)\\
+(\alpha_r, \alpha_G, \alpha_\kappa) &\sim \gammdist(\mathrm{mean}=1,\,\mathrm{shape}=0.1).
 \end{aligned}

### 2020-01-30 wuhan.tex
bmb comments

@@
 \newcommand{\jd}[1]{\comment{magenta}{JD}{#1}}
+\newcommand{\bmb}[1]{\comment{magenta}{BMB}{#1}}
 \newcommand{\new}[1]{\textcolor{blue}{#1}}
@@
 
+\bmb{general comments (Daniel may have taken notes on these already):
+  \begin{itemize}
+  \item try flatter-at-origin priors (e.g. half-Cauchy)?
+  \item Stan implementation?
+  \item comment on broader uncertainty context (King/Rohani on observation + process error, major reporting irregularities, non-stationarity \ldots)
+  \item cite Elderd et al on point estimates??
+  \item colorspace package ({\tt scale\_colour\_discrete\_qualitative} or other) for Figure 3?
+  \end{itemize}
+}
+
 \pagebreak

### 2020-01-30 R0_all.R

@@ -1,4 +1,5 @@
 library(coda)
+library(dplyr)
 
 load("sample_data.rda")
 load("MCMC_all.rda")

### 2020-01-30 wuhan_DC.tex

--- 2020-01-30 wuhan.tex
@@
 \usepackage{amsfonts}
+
+\usepackage[dvipsnames]{xcolor}
 
@@
 \newcommand{\jd}[1]{\comment{magenta}{JD}{#1}}
+\newcommand{\dc}[1]{\comment{Orange}{DC}{#1}}
 \newcommand{\new}[1]{\textcolor{blue}{#1}}
@@
 However, it can be difficult to assess the validity of the estimates of $\mathcal R_0$ (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
+\dc{Even if this is not directly addressed here, don't you also want to talk about observation error/delays, which are clearly present with this outbreak?}
 
 Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimate uncertainty should propagate error from all three sources.
+\dc{define what the generation interval is here.}
+We highlight the importance that clear and reasonable assumptions about these three quantities and that estimate uncertainty should propagate error from all three sources.
 
@@
 We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
+
+\dc{Maybe add more explicitly somewhere in the introduction that inevitably, early in an epidemic, we have poor observations, hence large uncertainty, around case reporting (affecting $r$) and contact tracing (affecting $g$).}
 
@@
 $^\ddagger$The authors presented $\mathcal R_0$ estimates under different assumptions; we use their baseline scenario in our analysis.
+\dc{You should include a column with the references of the studies and not keep them anonymous.}
 }
@@
 We modeled values reported using confidence intervals with gamma distributions, and values reported using ranges with uniform distributions.
+\dc{the ``uniform'' part relates to $\bar{G}$, right? If so, it's not clear neither here, nor later, only in Table 2}
+
 For example, Study 2 estimated $\mathcal R_0 = 2.92$ (95\% CI: 2.28--3.67);
@@
 Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of $\mathcal R_0$, even though the assumed values range from 0 to 0.5 across different studies.
+\dc{so what did the studies do?
+Especially the ones where you reported the studied considered a range for the mean GI?
+Maybe you want to clarify because I find it confusing you say they did not take into account GI but still report the range in Table 1 }
 It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval.
@@
 Black triangle: we assumed $\kappa=0.5$ for Study 1 as they do not report the parameter values that they used.
-Red points and lines: median and 95\% confidence intervals of our pooled estimates.
+Red points and lines: median and 95\% confidence intervals of our pooled estimates.\dc{shouldn't we be talking about ``credible'' intervals, given the method is Bayesian?}
 }
@@
 Propagating error from the growth rate (which all but one of the models did) is absolutely crucial.
+\dc{we should quantify here, this is the Results section and moreover it would strengthen the point.
+We should say something like ``uncertainty on $R_0$, measured with our 95\%CI is of the order of 3 (eyeballing 2.25--5.25)''}
 Propagating error from the generation interval also has important effects, however.
-Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example.
+\dc{Again, let's quantify, here it's about half as less, ie 1.5 (eyeballing 2.6--4.1)}
+Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example.\dc{why? this comes out of  nowhere...}
 
@@
 We also evaluate the estimates of $\mathcal R_0$ across different studies by 
-replacing the their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.
-We estimate narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0.2$ and $0$ respectively);
+replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+\dc{It just occurs to me now, but it's worth briefly describing the methods of those 6 studies... }
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but three cases.
+We estimate narrower confidence intervals for Study 2 and Study 6 \dc{really? I don't see that on the figure... I'm confused...}
+ when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0.2$ and $0$ respectively);
 when higher values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
@@
 we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ (\fref{R0}).
-For example, recalculating $\mathcal R_0$ for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+For example, recalculating $\mathcal R_0$ for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range reported (2.0--3.1).
 There are two explanations for this result.
@@
 We replace the reported paramaeter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our correponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate $\mathcal R_0$.
+\dc{I would like to see also the effect when all are included for each study (so adding a 5th error bar) and furthermore that would flow with the previous Figure 2}
 The pooled estimate of $\mathcal R_0$ is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$;
@@
 The reported $\mathcal R_0$ estimates (base) have been adjusted to show the approximate 95\% confidence interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
+\dc{please change the y-axis tick values to round ones, it will be easier to read.}
 }
@@
 We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
-While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
+While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.\dc{``credible'' intervals?}
 This result does not imply that assumptions based on the pooled estimate are too weak;
@@
 although our pooled estimates are relatively insensitive to the estimated uncertainty, our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.
-
+\dc{good point!}
 Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approches used to estimate the exponential growth rate $r$.
 Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, fitting a model to cumulative number of without accounting for problems with the dependent error structure can both bias parameter narrow confidence intervals \citep{ma2014estimating, king2015avoidable}.
+Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, fitting a model to cumulative number of without accounting for problems with the dependent error structure can both bias parameter narrow confidence intervals \citep{ma2014estimating, king2015avoidable}\dc{rephrase?}.
 Naive fits to cumulative incidence data should be avoided.
@@
 Given an estimate of the exponential growth rate $r$ and a distribution $g(\tau)$ of
-generation intervals (i.e., the time between when a person become 
+\dc{INTRISIC!
+sorry to be pedantic here, but taking backward GI for intrinsic ones is really a common mistake people do, especially early in the outbreak where censoring occurs.At  least coming from us-- it should be specified...} generation intervals (i.e., the time between when a person become
 infected and that person infects another person \citep{svensson2007note}), the basic reproductive
@@
 
-Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (1) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of $\mathcal R_0$.
-Assuming that generation intervals follow a gamma distribution
+Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (1) to characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of $\mathcal R_0$.
+Assuming that intrinsic generation intervals follow a gamma distribution
 with the mean $\bar G$ and the squared coefficient of variation (CV$^2$) $\kappa$, 
@@
 This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.
+
+\dc{I think I understand why you do that.
+But I'm worried the uncertainty for $r$ comes artificially from whatever methods the Studies used, not the actual observation process we would have by observing (imperfectly) directly the cases.
+I guess the studies have all more or less the same incidence data... why not estimating uncertainty for $r$ directly from incidence? }
 

### 2020-01-30 wuhan.tex
minor tweaks

@@
 \newcommand{\bmb}[1]{\comment{magenta}{BMB}{#1}}
+\newcommand{\dc}[1]{\comment{Orange}{DC}{#1}}
 \newcommand{\new}[1]{\textcolor{blue}{#1}}
@@
 However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
+\citep{Even if this is not directly addressed here, don’t you also want to talk about observation error/delays, which are clearly present with this outbreak?}
 
 Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical}.
 We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
@@
 We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
+\citep{Maybe add more explicitly somewhere in the introduction that inevitably, early in an epidemic, we have poor observations, hence large uncertainty, around case reporting (affecting r) and contact tracing (affecting g).}
 
@@
 \footnotesize
-\begin{tabular}{l|l|l|l}
- & Basic reproductive number & Mean generation time & CV$^2$ generation time \\
+\begin{tabular}{l|l|l|l|c}
+ & Basic reproductive number & Mean generation time & CV$^2$ generation time & Reference \\
 \hline
-Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 days & unspecified$^\dagger$ \\
+Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 days & unspecified$^\dagger$ & \cite{imaincov} \\
 \hline
-Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 days & 0.2 \\
+Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 days & 0.2 & \cite{liuncov} \\
 \hline
-Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 \\
+Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 & \cite{readncov} \\
 \hline
-Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5\\
+Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5 &  \\
 \hline

### 2020-01-30 wuhan.tex
piush

@@
 \usepackage{color}
-\usepackage{xspace}
+
 \newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}} 
@@
 \newcommand{\bmb}[1]{\comment{magenta}{BMB}{#1}}
-\newcommand{\dc}[1]{\comment{Orange}{DC}{#1}}
+\newcommand{\dc}[1]{\comment{magenta}{DC}{#1}}
 \newcommand{\new}[1]{\textcolor{blue}{#1}}
@@
 }
+\newline
+\\
+Sang Woo Park\textsuperscript{?}
+Benjamin M. Bolker\textsuperscript{?}
+David Champredon\textsuperscript{?}
+Bryan T. Grenfell\textsuperscript{?}
+Jonathan Dushoff\textsuperscript{?}
+\\
 
+\bigskip
+\textbf{1} Department of Mathematics \& Statistics, McMaster University, Hamilton, Ontario, Canada
+\\
+\textbf{2} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, USA
+\\
+\textbf{3} Department of Biology, McMaster University, Hamilton, Ontario, Canada
+\\
+\textbf{4} Agent-Based Modelling Laboratory, York University, Toronto, Ontario, Canada
+\\
+\textbf{5} Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Ontario, Canada
+\\
 \bigskip
@@
 However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
-\citep{Even if this is not directly addressed here, don’t you also want to talk about observation error/delays, which are clearly present with this outbreak?}
+\dc{Even if this is not directly addressed here, don’t you also want to talk about observation error/delays, which are clearly present with this outbreak?}
 
@@
 We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
-\citep{Maybe add more explicitly somewhere in the introduction that inevitably, early in an epidemic, we have poor observations, hence large uncertainty, around case reporting (affecting r) and contact tracing (affecting g).}
+\dc{Maybe add more explicitly somewhere in the introduction that inevitably, early in an epidemic, we have poor observations, hence large uncertainty, around case reporting (affecting r) and contact tracing (affecting g).}
 
@@
 \hline
-Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5 &  \\
+Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5 & \cite{riouncov} \\
 \hline
-Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2\\
+Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2 & \cite{zhaoncov} \\
 \hline
-Study 6 & 2.0--3.1 & 6--10 days & 0 \\
+Study 6 & 2.0--3.1 & 6--10 days & 0 & \cite{majumderncov} \\
 \hline
@@
 We then constructed a family of parameter sets for each study, and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
+\dc{the “uniform” part relates to G¯, right? If so, it’s not clear
+neither here, nor later, only in Table 2}
 

### 2020-01-30 wuhan.tex
author list
fiddling

@@
 
+\usepackage{xspace}
 \newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}} 
@@
 David Champredon\textsuperscript{?}
+Michale Li\textsuperscript{?}
 Bryan T. Grenfell\textsuperscript{?}
@@
 \\
-
 \bigskip
@@
 We then constructed a family of parameter sets for each study, and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
-\dc{the “uniform” part relates to G¯, right? If so, it’s not clear
+\dc{the “uniform” part relates to G, right? If so, it’s not clear
 neither here, nor later, only in Table 2}
@@
 many models rely on stronger assumptions that ignore these uncertainties.
-Surprisingly, no studies take into account how the variability in generation intervals (represented by $\kappa$) affects their estimates of \Ro, even though the assumed values range from 0 to 0.5 across different studies.
-It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., CV$^2$ equal to 1).
+Surprisingly, no studies take into account how the variation in generation intervals affects their estimates of \Ro:
+all studies assumed fixed values for $\kappa$, ranging from 0 to 0.5.
+It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$).
 
@@
 \fref{eff} shows the effects of propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
-Propagating error from the growth rate (which all but one of the models did) is absolutely crucial. 
-Propagating error from the generation interval also has important effects, however.
-Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example.
+Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
+uncertainty on $\mathcal R_0$, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
+Propagating error from the generation interval also has important effects, however (95\% CI: 2.59--4.12).
+Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example -- 
+propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
 
@@
 We compare estimates of $\mathcal R_0$ under five scenarios that propagate different combinations of uncertainties.
-(base): $\mathcal R_0$ estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-(growth rate): $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
-(GI mean): $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
-(growth rate + GI mean): $\mathcal R_0$ estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
-(all): $\mathcal R_0$ estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+\textbf{base}: $\mathcal R_0$ estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+\textbf{growth rate}: $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
+\textbf{GI mean}: $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
+\textbf{growth rate + GI mean}: $\mathcal R_0$ estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
+\textbf{all}: $\mathcal R_0$ estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
 Vertical lines represent the 95\% confidence intervals.
@@
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.
-We estimate narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution ($\kappa=0.2$ and $0$ respectively);
+We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
 when higher values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
+We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
 
@@
 \textbf{Sensitivity of the reported $\mathcal R_0$ estimates with respect to our pooled estimates of the underlying parameters.}
-We replace the reported parameter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate $\mathcal R_0$.
-The pooled estimate of $\mathcal R_0$ is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$;
-horizontal dashed lines represent the 95\% confidence intervals of our pooled estimate of $\mathcal R_0$.
-The reported $\mathcal R_0$ estimates (base) have been adjusted to show the approximate 95\% confidence interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
+We replace the reported parameter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate $\mathcal R_0$ (\textbf{growth rate}, \textbf{GI mean}, and \textbf{GI variation}).
+The pooled estimate of $\mathcal R_0$ is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$ (\textbf{all});
+this corresponds to replacing all reported parameter values with our pooled estimates, which gives identical results across all studies.
+Horizontal dashed lines represent the 95\% confidence intervals of our pooled estimate of $\mathcal R_0$.
+The reported $\mathcal R_0$ estimates (\textbf{base}) have been adjusted to show the approximate 95\% confidence interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
 }
@@
 Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, fitting a model to cumulative number of without accounting for problems with the dependent error structure can both bias parameter narrow confidence intervals \citep{ma2014estimating, king2015avoidable}.
+Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, fitting a model to cumulative incidence, instead of raw incidence, can both bias parameter and give narrow confidence intervals if non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
 Naive fits to cumulative incidence data should be avoided.
@@
 the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
-Given an estimate of the exponential growth rate $r$ and a distribution $g(\tau)$ of
-generation intervals (i.e., the time between when a person become 
-infected and that person infects another person \citep{svensson2007note}), the basic reproductive
+Given an estimate of the exponential growth rate $r$ and an \emph{instrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
 number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
@@
 
+\dc{I think I understand why you do that. But I’m worried the uncertainty for r comes
+artificially from whatever methods the Studies used, not the actual observation process we
+would have by observing (imperfectly) directly the cases. I guess the studies have all more or
+less the same incidence data... why not estimating uncertainty for r directly from incidence}
+
 We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
@@
 Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference}.
+95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the posterior distribution.
+
+\dc{It just occurs to me now, but it’s worth briefly
+describing the methods of those 6 studies...}
 

### 2020-01-30 compare_R0.R

@@ -1,5 +1,6 @@
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
+library(dplyr)
 source("published_estimate.R")
 source("study_number.R")
 
@@ -51,7 +52,7 @@ g1 <- ggplot(R0all) +
   geom_point(aes(anon, est, col=type), position=position_dodge(0.5)) +
   geom_errorbar(aes(anon, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),
                 width=0) +
-  scale_y_continuous("Basic reproductive number") +
+  scale_y_continuous("Basic reproductive number", breaks=c(2, 4, 6, 8, 10)) +
   theme(
     axis.title.x = element_blank(),
     legend.position = "top",

### 2020-01-30 wuhan.tex
update author

@@
 Sang Woo Park\textsuperscript{?}
-Benjamin M. Bolker\textsuperscript{?}
 David Champredon\textsuperscript{?}
-Michale Li\textsuperscript{?}
+Michael Li\textsuperscript{?}
 Bryan T. Grenfell\textsuperscript{?}
@@
 \bigskip
-\textbf{1} Department of Mathematics \& Statistics, McMaster University, Hamilton, Ontario, Canada
+\textbf{1} Department of ???
 \\
-\textbf{2} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, USA
+\textbf{2} Department of ???
 \\
-\textbf{3} Department of Biology, McMaster University, Hamilton, Ontario, Canada
+\textbf{3} Department of ???
 \\
-\textbf{4} Agent-Based Modelling Laboratory, York University, Toronto, Ontario, Canada
+\textbf{4} Department of ???
 \\
-\textbf{5} Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Ontario, Canada
+\textbf{5} Department of ???
 \\
@@
 Constant distributions assume fixed values according to Table 1.
-$^\ast$Study 6 effectively fits an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+$^\ast$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling $\mathcal R_0$ with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in $\mathcal R_0$ they report, when combined with the range of $\bar G$ they consider.
@@
 
+\section*{Acknowledgements}
+
+We thank Ben Bolker for providing useful comments on the manuscript.
+
 \pagebreak

### 2020-01-30 wuhan.tex
more update

@@
 
-As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+As the disease continues to spread, many researchers have published analyses of the outbreak, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
 Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
@@
 
-\fref{eff} shows the effects of propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
+\fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
 Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
@@
 We also evaluate the estimates of $\mathcal R_0$ across different studies by 
-replacing the their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.

### 2020-01-30 wuhan.tex
more tweaks

@@
 Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+We congratulate these researchers for their timely contribution and those who made the data publicly available.
 However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
@@
 The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical}.
-We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
+We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
 
@@
 \begin{center}
-\footnotesize
+\scriptsize
 \begin{tabular}{l|l|l|l|c}
- & Basic reproductive number & Mean generation time & CV$^2$ generation time & Reference \\
+ & Basic reproductive number & Mean generation time (days) & CV$^2$ generation time & Reference \\
 \hline
-Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 days & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
 \hline
-Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 days & 0.2 & \cite{liuncov} \\
+Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
 \hline
-Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 & \cite{readncov} \\
+Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
 \hline
-Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5 & \cite{riouncov} \\
+Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
 \hline
-Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2 & \cite{zhaoncov} \\
+Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
 \hline
-Study 6 & 2.0--3.1 & 6--10 days & 0 & \cite{majumderncov} \\
+Study 6 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
 \hline

### 2020-01-30 wuhan.tex
tired

@@
 Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical}.
+The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
+early in an epidemic, information is scarce and, inevitably, there is large uncertainty around contact tracing (affecting the estimates of generation intervals) and case reports (affecting the estimates of exponential growth rate).
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
@@
 We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
-\dc{Maybe add more explicitly somewhere in the introduction that inevitably, early in an epidemic, we have poor observations, hence large uncertainty, around case reporting (affecting r) and contact tracing (affecting g).}
 

### 2020-01-30 wuhan.tex
fig 3; tweaking

@@
 $^\ast$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
-Instead of modeling $\mathcal R_0$ with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in $\mathcal R_0$ they report, when combined with the range of $\bar G$ they consider.
+Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in \Ro they report, when combined with the range of $\bar G$ they consider.
 }
@@
 Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
-uncertainty on $\mathcal R_0$, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
+uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
 Propagating error from the generation interval also has important effects, however (95\% CI: 2.59--4.12).
@@
 \caption{
-\textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of $\mathcal R_0$.}
-We compare estimates of $\mathcal R_0$ under five scenarios that propagate different combinations of uncertainties.
-\textbf{base}: $\mathcal R_0$ estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-\textbf{growth rate}: $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
-\textbf{GI mean}: $\mathcal R_0$ estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
-\textbf{growth rate + GI mean}: $\mathcal R_0$ estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
-\textbf{all}: $\mathcal R_0$ estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+\textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
+We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties.
+\textbf{base}: \Ro estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+\textbf{growth rate}: \Ro estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
+\textbf{GI mean}: \Ro estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
+\textbf{growth rate + GI mean}: \Ro estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
+\textbf{all}: \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
 Vertical lines represent the 95\% confidence intervals.
@@
 
-We also evaluate the estimates of $\mathcal R_0$ across different studies by 
-replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number $\mathcal R_0$ (\fref{R0}).
+We also evaluate the estimates of \Ro across different studies by 
+replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.
-We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate $\mathcal R_0$ because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
-when higher values of $\kappa$ are used, their estimates of $\mathcal R_0$ become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate $\mathcal R_0$ (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
+We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
+We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
 
 Consistent with our previous observations (\fref{eff}),
-we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of $\mathcal R_0$ (\fref{R0}).
-For example, recalculating $\mathcal R_0$ for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
+For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
 There are two explanations for this result.
-First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on $\mathcal R_0$ under the gamma approximation framework \eref{gamma},
-$r$ has a greater overall effect on $\mathcal R_0$ because it is associated with more uncertainty;
+First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework \eref{gamma},
+$r$ has a greater overall effect on \Ro because it is associated with more uncertainty;
 the range of the 95\% confidence interval on $\mu_r$ spans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
-Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of $\mathcal R_0$ too sensitive to $r$ and $\bar G$ as explained previously.
+Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as explained previously.
 
@@
 \caption{
-\textbf{Sensitivity of the reported $\mathcal R_0$ estimates with respect to our pooled estimates of the underlying parameters.}
-We replace the reported parameter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate $\mathcal R_0$ (\textbf{growth rate}, \textbf{GI mean}, and \textbf{GI variation}).
-The pooled estimate of $\mathcal R_0$ is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$ (\textbf{all});
+\textbf{Sensitivity of the reported \Ro estimates with respect to our pooled estimates of the underlying parameters.}
+We replace the reported parameter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro (\textbf{growth rate}, \textbf{GI mean}, and \textbf{GI variation}).
+The pooled estimate of \Ro is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$ (\textbf{all});
 this corresponds to replacing all reported parameter values with our pooled estimates, which gives identical results across all studies.
-Horizontal dashed lines represent the 95\% confidence intervals of our pooled estimate of $\mathcal R_0$.
-The reported $\mathcal R_0$ estimates (\textbf{base}) have been adjusted to show the approximate 95\% confidence interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
+Horizontal dashed lines represent the 95\% confidence intervals of our pooled estimate of \Ro.
+The reported \Ro estimates (\textbf{base}) have been adjusted to show the approximate 95\% confidence interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
 }
@@
 
-Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate $\mathcal R_0$ and compare it with the reported $\mathcal R_0$ estimates.
-We estimate the average $\mathcal R_0$ to be 3.1 (95\% CI: 2.1--5.7).
-While our pooled estimate of $\mathcal R_0$ is similar to other reported values, our calculation gives wider confidence intervals than all of them.
+Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate \Ro and compare it with the reported \Ro estimates.
+We estimate the average \Ro to be 3.1 (95\% CI: 2.1--5.7).
+While our pooled estimate of \Ro is similar to other reported values, our calculation gives wider confidence intervals than all of them.
 This result does not imply that assumptions based on the pooled estimate are too weak;
@@
 
-Estimating the basic reproductive number $\mathcal R_0$ is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we used a simple framework \citep{park2019practical} to compare estimates of $\mathcal R_0$ for the novel coronavirus outbreak.
+Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
+Here, we used a simple framework \citep{park2019practical} to compare estimates of \Ro for the novel coronavirus outbreak.
 Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including with the amount of dispersion in the generation intervals:
-although our pooled estimates are relatively insensitive to the estimated uncertainty, our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of $\mathcal R_0$ too sensitive to the estimates of the exponential growth rate $r$.
+although our pooled estimates are relatively insensitive to the estimated uncertainty, our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro too sensitive to the estimates of the exponential growth rate $r$.
 
-Narrow confidence intervals associated with early estimates of $\mathcal R_0$ partly depend on the statistical approaches used to estimate the exponential growth rate $r$.
+Narrow confidence intervals associated with early estimates of \Ro partly depend on the statistical approaches used to estimate the exponential growth rate $r$.
 Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
@@
 
-Early in an outbreak, $\mathcal R_0$ cannot be estimated directly;
-instead, $\mathcal R_0$ is often inferred from
+Early in an outbreak, \Ro cannot be estimated directly;
+instead, \Ro is often inferred from
 the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
@@
 \end{equation}
-In other words, estimates of $\mathcal R_0$ must
+In other words, estimates of \Ro must
 depend on the assumptions about the
@@
 
-Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (1) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of $\mathcal R_0$.
+Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (1) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of \Ro.
 Assuming that generation intervals follow a gamma distribution
@@
 that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
-will give a higher estimate of $\mathcal R_0$ for the same value of $r$.
+will give a higher estimate of \Ro for the same value of $r$.
 
-For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of $\mathcal R_0$ and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
+For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Ro and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
 \begin{equation}
@@
 \mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\
-\mu_G &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=0.1)\\
+\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=0.1)\\
 (\alpha_r, \alpha_G, \alpha_\kappa) &\sim \gammdist(\mathrm{mean}=1,\,\mathrm{shape}=0.1).
@@
   \begin{itemize}
-  \item try flatter-at-origin priors (e.g. half-Cauchy)?
-  \item Stan implementation?
+  \item (\textbf{later on}) try flatter-at-origin priors (e.g. half-Cauchy)? Stan implementation?
   \item comment on broader uncertainty context (King/Rohani on observation + process error, major reporting irregularities, non-stationarity \ldots)
-  \item cite Elderd et al on point estimates??
-  \item colorspace package ({\tt scale\_colour\_discrete\_qualitative} or other) for Figure 3?
+  \item cite \cite{elderd2006uncertainty} on point estimates
   \end{itemize}

### 2020-01-30 compare_R0.R

@@ -1,5 +1,6 @@
 library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
+library(colorspace)
 library(dplyr)
 source("published_estimate.R")
 source("study_number.R")
@@ -46,17 +47,23 @@ R0all %>%
     ww=width>bwidth
   )
 
+all_col <- "#000000" ## black
 g1 <- ggplot(R0all) +
-  geom_hline(data=R0_all, aes(yintercept=lwr), lty=2, col="#C77CFF") +
-  geom_hline(data=R0_all, aes(yintercept=upr), lty=2, col="#C77CFF") +
+  ## tried to collapse these two lines to one using annotate or other clever
+  ## tricks, failed ...
+  geom_hline(data=R0_all, aes(yintercept=lwr), lty=2, col=all_col, alpha=0.5) +
+  geom_hline(data=R0_all, aes(yintercept=upr), lty=2, col=all_col, alpha=0.5) +
   geom_point(aes(anon, est, col=type), position=position_dodge(0.5)) +
   geom_errorbar(aes(anon, ymin=lwr, ymax=upr, col=type), position=position_dodge(0.5),
                 width=0) +
   scale_y_continuous("Basic reproductive number", breaks=c(2, 4, 6, 8, 10)) +
+  ## scale_colour_discrete_qualitative() +
+  scale_colour_manual(values=c(colorspace::qualitative_hcl(4),"#000000")) +
   theme(
     axis.title.x = element_blank(),
     legend.position = "top",
     legend.title = element_blank()
   )
+
 
 ggsave("compare_R0.pdf", g1, width=6, height=4)

### 2020-01-30 wuhan.tex
i sense merge conflicts

--- 2020-01-30 wuhan.tex
@@
 Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical}.
+The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
+early in an epidemic, information is scarce and, inevitably, there is large uncertainty around case reports (affecting the estimates of exponential growth rate) and contact tracing (affecting the estimates of generation intervals).
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
@@
 
-We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were uploaded to preprint servers between January 24th, 2020 and January 26th, 2020 (Table 1).
+We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
 For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a Gamma distribution).
@@
 neither here, nor later, only in Table 2}
+\swp{This is coincidence... Not sure how to explain it...}
 
@@
 all studies assumed fixed values for $\kappa$, ranging from 0 to 0.5.
+Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
 It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$).
@@
 
+\subsection{Gamma approximation framework for linking $r$ and $\mathcal R_0$}
+
 Early in an outbreak, $\mathcal R_0$ cannot be estimated directly;
@@
 
+\subsection{Description of the studies}
+
+We reviewed six modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.
+\cite{imaincov} and \cite{rioucov} estimated the basic reproduction number by simulating branching process models
+
+majumder 26th
+read 21st
+zhao 22nd
+liu 23rd
+Two based on 18th.
+
+\subsection{Statistical framework}
+
 For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of $\mathcal R_0$ and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
@@
 \end{equation}
-This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.
+This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
+uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.
 
@@
 less the same incidence data... why not estimating uncertainty for r directly from incidence}
+\swp{that is exactly the point... we want to be able to reflect their methods... added a sentence about it.}
 
@@
   \item comment on broader uncertainty context (King/Rohani on observation + process error, major reporting irregularities, non-stationarity \ldots)
-  \item cite Elderd et al on point estimates??
   \item colorspace package ({\tt scale\_colour\_discrete\_qualitative} or other) for Figure 3?

### 2020-01-30 wuhan.tex

--- 2020-01-30 wuhan.tex
@@
 The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
-early in an epidemic, information is scarce and, inevitably, there is large uncertainty around contact tracing (affecting the estimates of generation intervals) and case reports (affecting the estimates of exponential growth rate).
+early in an epidemic, information is scarce and, inevitably, there is large uncertainty around case reports (affecting the estimates of exponential growth rate) and contact tracing (affecting the estimates of generation intervals).
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
@@
 
+\subsection{Gamma approximation framework for linking $r$ and $\mathcal R_0$}
+
 Early in an outbreak, \Ro cannot be estimated directly;
-instead, \Ro is often inferred from
-the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
+instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
 Given an estimate of the exponential growth rate $r$ and an \emph{instrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
@@
 
+\subsection{Description of the studies}
+
+We reviewed six modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.
+\cite{imaincov} and \cite{rioucov} estimated the basic reproduction number by simulating branching process models
+
+majumder 26th
+read 21st
+zhao 22nd
+liu 23rd
+Two based on 18th.
+
+\subsection{Statistical framework}
+
 For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Ro and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
@@
   \item comment on broader uncertainty context (King/Rohani on observation + process error, major reporting irregularities, non-stationarity \ldots)
-  \item cite \cite{elderd2006uncertainty} on point estimates
   \end{itemize}

### 2020-01-30 wuhan.tex
one less comment

@@
 1/\mathcal R_0 = \int \exp(-r\tau) g(\tau) d\tau.
+\label{eq:euler}
 \end{equation}
@@
 Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.
-\cite{imaincov} and \cite{rioucov} estimated the basic reproduction number by simulating branching process models
-
-majumder 26th
-read 21st
-zhao 22nd
-liu 23rd
-Two based on 18th.
+There is a wide variation in their statistical methods and the amount of data they used to infer \Ro.
+\cite{imaincov} and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18th.
+\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1st and January 21st from major cities in China and other countries.
+\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22nd and January 23rd, respectively, and inferred \Ro\ via the Euler-Lotka equation \eref{euler}.
+\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26th, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
 
@@
 
-\dc{It just occurs to me now, but it’s worth briefly
-describing the methods of those 6 studies...}
-
 \bmb{general comments (Daniel may have taken notes on these already):

### 2020-01-30 wuhan.tex
need to get dinner

@@
 
+More broadly, disease modelers should always consider uncertainties in the data generating processes.
+Process vs obs noise.
+
+Our results indicate that the uncertainty in the estimates of the exponential growth rate $r$ matters the most.
+However, as the epidemic progresses and more data becomes available, inference on will become more precise, eventually decreasing.
+Uncertainty in the exponential growth rate will decrease, 
+and the effect of uncertainty in the mean generation time will become more important.
+
 We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components.

### 2020-01-30 wuhan.tex
stuff

--- 2020-01-30 wuhan.tex
@@
 
+More broadly, disease modelers should always consider uncertainties in the data generating processes.
+Process vs obs noise.
+
+Our results indicate that the uncertainty in the estimates of the exponential growth rate $r$ matters the most.
+However, as the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
+Uncertainty in the exponential growth rate will eventually decrease, 
+and the effect of uncertainty in the mean generation time will become more important.
+
 We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components.

### 2020-01-31 wuhan.tex
tweaking

@@
 	\textbf\newline{
-		Reconciling early-outbreak preliminary estimates of the basic reproductive number: a new framework and applications to the novel coronavirus (2019-nCoV) outbreak
+		Reconciling early-outbreak preliminary estimates of the basic reproductive number and its uncertainty: a new framework and applications to the novel coronavirus (2019-nCoV) outbreak
 	}
@@
 Michael Li\textsuperscript{?}
+Joshua S. Weitz\textsuperscript{?}
 Bryan T. Grenfell\textsuperscript{?}
@@
 However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
-\dc{Even if this is not directly addressed here, don’t you also want to talk about observation error/delays, which are clearly present with this outbreak?}
 
@@
 The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
-early in an epidemic, information is scarce and, inevitably, there is large uncertainty around case reports (affecting the estimates of exponential growth rate) and contact tracing (affecting the estimates of generation intervals).
+therefore, estimates of $\mathcal R_0$ from different models directly depend on their implicit assumptions about the shape of the generation-interval distribution and the exponential growth rate.
+Early in an epidemic, information is scarce and, inevitably, there is large uncertainty around case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
 
-We evaluate six disparate models posted online between January 24--26, 2020 that estimated \Ro\ for the 2019-nCoV outbreak.
-We use our framework to construct pooled estimates for the underlying parameters.
+We evaluate six disparate models published online between January 24--26, 2020 that estimated \Ro\ for the 2019-nCoV outbreak \citep{imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
+We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$. 
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. 
@@
 We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
-For each study $i$, we analyze explicitly or implicitly reported distributions of the reproductive number $\mathcal R_{0i}$, the mean generation interval $\bar G_i$, and the generation-interval dispersion parameter $\kappa_i$ (equivalent to the squared coefficient of variation for a Gamma distribution).
-We used Gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
+As none of these studies report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond their model assumptions.
+We do so by modeling explicitly or implicitly reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
+we used Gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
 For example, Study 2 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
-we model this parameter as a Gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
-We then constructed a family of parameter sets for each study, and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
-\dc{the “uniform” part relates to G, right? If so, it’s not clear
-neither here, nor later, only in Table 2}
-\swp{This is coincidence... Not sure how to explain it...}
+we model this estimate as a Gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
+We then constructed a family of parameter sets -- which include $r$, $\bar G$, and $\kappa$ -- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
 
@@
 \begin{tabular}{l|l|l|l}
- & Basic reproductive number & mean generation time (days) & CV$^2$ generation time \\
+ & Basic reproductive number & Mean generation time (days) & CV$^2$ generation time \\
 \hline

### 2020-01-31 wuhan.tex
want to brush my teeth

@@
 \scriptsize
-\begin{tabular}{l|l|l|l|c}
- & Basic reproductive number & Mean generation time (days) & CV$^2$ generation time & Reference \\
+\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.5cm}|l}
+ & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation interval\newline dispersion $\kappa$ \\
 \hline
@@
 \begin{center}
-\footnotesize
-\begin{tabular}{l|l|l|l}
- & Basic reproductive number & Mean generation time (days) & CV$^2$ generation time \\
+\scriptsize
+\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.5cm}}
+ & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation interval\newline dispersion $\kappa$  \\
 \hline
-Study 1 & $\gammdist(\alpha=28, \beta=28/2.6)$ & constant & constant \\
+Study 1 & $\gammdist(\alpha=28, \beta=28/2.6)$ & 8.4 & 0.5 \\
 \hline
-Study 2 & $\gammdist(\alpha=67, \beta=67/2.92)$ & constant & constant \\
+Study 2 & $\gammdist(\alpha=67, \beta=67/2.92)$ & 8.4 & 0.2 \\
 \hline
-Study 3 & $\gammdist(\alpha=1400, \beta=1400/3.8)$ & constant & constant \\
+Study 3 & $\gammdist(\alpha=1400, \beta=1400/3.8)$ & 7.6 & 0.5 \\
 \hline
-Study 4 & $\gammdist(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & constant\\
+Study 4 & $\gammdist(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
 \hline
-Study 5 & $\gammdist(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & constant\\
+Study 5 & $\gammdist(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & 0.2\\
 \hline
-Study 6 & N.A.$^\ast$ & $\mathrm{Uniform}(6, 10)$ & constant\\
+Study 6 & $\exp(r \bar G)^\ast$ & $\mathrm{Uniform}(6, 10)$ & 0\\
 \hline
@@
 The Gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
-Constant distributions assume fixed values according to Table 1.
+Constant values are fixed according to Table 1.
 $^\ast$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.

### 2020-01-31 wuhan.tex
stuff

@@
 \newcommand{\dc}[1]{\comment{magenta}{DC}{#1}}
+\newcommand{\jsw}[1]{\comment{magenta}{JSW}{#1}}
 \newcommand{\new}[1]{\textcolor{blue}{#1}}
@@
 Vertical lines represent the 95\% confidence intervals.
+\jsw{why not show all combinations}
 }
@@
 we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
-For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 4.0$ (95\% CI: 2.3--10.1), which is much wider than the range they reported (2.0--3.1).
+For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the range they reported (2.0--3.1).
 There are two explanations for this result.
@@
 Naive fits to cumulative incidence data should be avoided.
+\jsw{Ok... it's not strictly speaking a logistical model fit is it?  Also, this sentence is quite loose ('familiar to many people', which people).}
 
@@
 We are hopeful that this will provide a guide to understanding and reconciling different estimates early in an epidemic.
+\jsw{remind us what they are: 
+three simple components: the growth rate, mean generation interval, and generation interval dispersion.}
 

### 2020-01-31 wuhan.tex
tweak

@@
 
-Early in an outbreak, \Ro cannot be estimated directly;
+Early in an outbreak, \Ro is difficult to estimate directly;
 instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.

### 2020-01-31 wuhan.tex
k

@@
 three simple components: the growth rate, mean generation interval, and generation interval dispersion.}
+\jsw{Here is my take: you are doing a Nate Silver-esque approach, pooling estimates in a smart way without necessarily redoing all the models. Isn't that right? That's cool!  
+But, I have one nagging worry: the folks measured r - yet this method says that they measured the uncertainty incorrectly...
+My nagging worry has to do with the structure of the multi-level modeling scheme.
+Aren't they all using the same data?
+Or are they measuring different data types (because the multi-level scheme you use has the data in parallel, whereas I think the data used may be nested in ways I don't yet understand).}
 
@@
 Since the gamma distribution does not allow zeros, we use $\kappa =0.02$ instead for Study 6.
+\jsw{Just a question: was the data the same?
+I assume the models fit to different data?
+I like this idea, except for the fact that these are not different outbreaks.
+I feel like this kind of multilevel model would be right if we had the same fundameental disease dynamics taking palce in different places.
+But what happens if the models are just different, even if the data is the same.
+Again, I have a small philosophical doubt with the implication that the i-s denote the same disease mechanism...
+and why are all the hyper-parameters gammas?}
 

### 2020-01-31 wuhan.tex
tewak

@@
 %\usepackage{adjustbox}
+
+\usepackage{lineno}
+\linenumbers
 
@@
 \\
-Sang Woo Park\textsuperscript{?}
+Sang Woo Park\textsuperscript{1}
 David Champredon\textsuperscript{?}
+David Earn\textsuperscript{?}
 Michael Li\textsuperscript{?}
@@
 \bigskip
-\textbf{1} Department of ???
+\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, NJ, USA
 \\

### 2020-01-31 wuhan.tex
adding bryan's comments

@@
 we believe that this confidence interval accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+\swp{Bryan: any comment about why R0 isn’t at the lower end of previous estimates? (Guess someone might ask this)}
 
@@
 
-More broadly, disease modelers should always consider uncertainties in the data generating processes.
-Process vs obs noise.
+More broadly, disease modelers should always consider different sources of uncertainties in the data generating process.
+Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
 
@@
 and the effect of uncertainty in the mean generation time will become more important.
+Furthermore,
 

### 2020-01-31 compare_R0.R

@@ -65,5 +65,4 @@ g1 <- ggplot(R0all) +
     legend.title = element_blank()
   )
 
-
 ggsave("compare_R0.pdf", g1, width=6, height=4)

### 2020-01-31 wuhan.tex
One change to Intro
pushign some stuff

@@
 Sang Woo Park\textsuperscript{1}
-David Champredon\textsuperscript{?}
+David Champredon\textsuperscript{2}
 David Earn\textsuperscript{?}
@@
 \bigskip
-\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, NJ, USA
+\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, USA
 \\
-\textbf{2} Department of ???
+\textbf{2} Department of Pathology and Laboratory Medicine, University of Western Ontario, London, Ontario, Canada
 \\
@@
 
-More broadly, disease modelers should always consider different sources of uncertainties in the data generating process.
+More broadly, disease modelers should carefully consider different sources of uncertainties in the data generating process.
 Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
+When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible .
 
 Our results indicate that the uncertainty in the estimates of the exponential growth rate $r$ matters the most.
-However, as the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
+As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
 Uncertainty in the exponential growth rate will eventually decrease, 

### 2020-01-31 wuhan.tex
repaired line numbers and added my affiliations

--- 2020-01-31 wuhan.tex
@@
 
-\usepackage{lineno}
+\usepackage{xcolor}
+\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
 \linenumbers
@@
 David Champredon\textsuperscript{?}
-David Earn\textsuperscript{?}
+David J.\,D.\ Earn\textsuperscript{3}
 Michael Li\textsuperscript{?}
@@
 \\
-\textbf{3} Department of ???
+\textbf{3} Department of Mathematics and Statistics and M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
 \\
@@
 number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 In other words, estimates of \Ro must
@@
 we have
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 This equation demonstrates that a generation-interval distribution
@@
 For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Ro and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
@@
 We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step;
@@
 Weakly informative priors are assumed on the hyperparameters:
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 We run 4 parallel Markov Chain Monte Carlo (MCMC) chains that consist of 200,000 burnin steps and 200,000 sampling steps.

### 2020-01-31 wuhan.tex

@@
 Sang Woo Park\textsuperscript{1}
-David Champredon\textsuperscript{?}
+David Champredon\textsuperscript{2}
 David J.\,D.\ Earn\textsuperscript{3}
@@
 Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
-We congratulate these researchers for their timely contribution and those who made the data publicly available.
+We commend these researchers for their timely contribution and those who made the data publicly available.
 However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.

### 2020-01-31 wuhan.tex
please no merge conflicts

--- 2020-01-31 wuhan.tex
@@
 Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
-When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible .
+When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible answers.
 
-Our results indicate that the uncertainty in the estimates of the exponential growth rate $r$ matters the most.
+Here, we only focused on estimates of \Ro that were published within a very short frame of time.
+Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: uncertainties in generation-interval distribution is important.
 As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
@@
 and the effect of uncertainty in the mean generation time will become more important.
-Furthermore,
 

### 2020-01-31 wuhan.tex
aff

@@
 David Champredon\textsuperscript{2}
-David J.\,D.\ Earn\textsuperscript{3}
+David J.\,D.\ Earn\textsuperscript{3,4}
 Michael Li\textsuperscript{?}
@@
 Bryan T. Grenfell\textsuperscript{?}
-Jonathan Dushoff\textsuperscript{?}
+Jonathan Dushoff\textsuperscript{?,4}
 \\
@@
 \\
-\textbf{3} Department of Mathematics and Statistics and M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
+\textbf{3} Department of Mathematics and Statistics, McMaster University, Hamilton, ON, Canada
 \\
-\textbf{4} Department of ???
+\textbf{4} M.\,G.\,DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, ON, Canada
 \\

### 2020-01-31 wuhan.tex
A couple of substantive comments

--- 2020-01-31 wuhan.tex
@@
 Sang Woo Park\textsuperscript{1}
-David Champredon\textsuperscript{2}
+David Champredon\textsuperscript{?}
 David Earn\textsuperscript{?}
@@
 \bigskip
-\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, USA
+\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, NJ, USA
 \\
-\textbf{2} Department of Pathology and Laboratory Medicine, University of Western Ontario, London, Ontario, Canada
+\textbf{2} Department of ???
 \\
@@
 Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
-It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$).
+It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly.
 
@@
 
-\fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
+\jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
+But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
+\fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.
+For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
 Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
+\jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
+Who wanted this and why?
+I can write something about the comparison above instead?}.
 uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
@@
 
-More broadly, disease modelers should carefully consider different sources of uncertainties in the data generating process.
+More broadly, disease modelers should always consider different sources of uncertainties in the data generating process.
 Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
-When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible answers.
 
-Here, we only focused on estimates of \Ro that were published within a very short frame of time.
-Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: uncertainties in generation-interval distribution is important.
-As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
+Our results indicate that the uncertainty in the estimates of the exponential growth rate $r$ matters the most.
+However, as the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
 Uncertainty in the exponential growth rate will eventually decrease, 
 and the effect of uncertainty in the mean generation time will become more important.
+Furthermore,
 

### 2020-01-31 wuhan.tex
An apology of sorts ☺
let me pull

--- 2020-01-31 wuhan.tex
@@
 Here, we only focused on estimates of \Ro that were published within a very short frame of time.
-Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: uncertainties in generation-interval distribution is important.
+Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: understanding uncertainties in generation-interval distribution is important.
 As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
-Uncertainty in the exponential growth rate will eventually decrease, 
+Uncertainty in the exponential growth rate will inevitably decrease, 
 and the effect of uncertainty in the mean generation time will become more important.
+A recent epidemiological analysis of the patient-level data revealed large uncertainty in serial intervals, which is defined as the time between symptom onsets of an infector and an infectee, (mean: 7.5 days; 95\% CI: .3--19) \citep{nejmncov};
 

### 2020-01-31 wuhan.tex

--- 2020-01-31 wuhan.tex
@@
 But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
+\jd{I have no problem doing this in the next round, though. Time is important.}
 \fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.

### 2020-01-31 wuhan.tex
push?

@@
 Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
-When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible answers.
+When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible answers \citep{ma2014estimating}.
+In general, however, ignoring process error and fitting deterministic models can give overly narrow confidence intervals \citep{king2015avoidable}.
 
-Here, we only focused on estimates of \Ro that were published within a very short frame of time.
-Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: understanding uncertainties in generation-interval distribution is important.
+Here, we focused on estimates of \Ro that were published within a very short frame of time (January 24th--26th).
+Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: understanding uncertainties in generation-interval distribution is important.
 As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
@@
 and the effect of uncertainty in the mean generation time will become more important.
-A recent epidemiological analysis of the patient-level data revealed large uncertainty in serial intervals, which is defined as the time between symptom onsets of an infector and an infectee, (mean: 7.5 days; 95\% CI: .3--19) \citep{nejmncov};
+We strongly emphasize that more efforts should be put into tracing the chain of transmission and accurately characterizing the generation-interval distribution of the novel coronavirus.
+
+% A recent epidemiological analysis of 452 patients from Wuhan \citep{nejmncov} revealed large uncertainty in the serial-interval distributions (mean: 7.5 days; 95\% CI: 5.3--19).
+% The serial interval, which is defined as the time between symptom onset of an infector and an infectee, is closely linked to the generation-interval distribution \citep{svensson2007note}.
+% Their results suggest that there is likely even more uncertainty in the generation time of the novel coronavirus.
 

### 2020-01-31 wuhan.tex
discussion is hard

@@
 
-More broadly, disease modelers should carefully consider different sources of uncertainties in the data generating process.
+More broadly, disease modelers always need to consider all potential sources of uncertainties in the data generating process and weigh their importance.
 Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
@@
 
-Here, we focused on estimates of \Ro that were published within a very short frame of time (January 24th--26th).
+Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 24th--26th).
 Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: understanding uncertainties in generation-interval distribution is important.
@@
 Uncertainty in the exponential growth rate will inevitably decrease,
-and the effect of uncertainty in the mean generation time will become more important.
-We strongly emphasize that more efforts should be put into tracing the chain of transmission and accurately characterizing the generation-interval distribution of the novel coronavirus.
-
-% A recent epidemiological analysis of 452 patients from Wuhan \citep{nejmncov} revealed large uncertainty in the serial-interval distributions (mean: 7.5 days; 95\% CI: 5.3--19).
-% The serial interval, which is defined as the time between symptom onset of an infector and an infectee, is closely linked to the generation-interval distribution \citep{svensson2007note}.
-% Their results suggest that there is likely even more uncertainty in the generation time of the novel coronavirus.
+and the effect of uncertainty in the generation-interval distribution will become even more important.
+We strongly emphasize the need for accurate characterization of the transmission chains via contact tracing and better statistical framework for inferring generation-interval distributions from such data \citep{britton2019estimation}.
+Combined efforts is crucial for predicting the course of an epidemic and controlling the outbreak.
 

### 2020-01-31 wuhan.tex
fiddle
Trying to resolve and/or comment out comments.

--- 2020-01-31 wuhan.tex
@@
 
-\jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
+%% \jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
 But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
-\jd{I have no problem doing this in the next round, though. Time is important.}
+%% \jd{I have no problem doing this in the next round, though. Time is important.}
 \fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.
 For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
-Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
-\jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
+Comparing the models that include only some sources of uncertainty to the confidences intervals from the ``all'' model, we see that propagating error from the growth rate (which all but one of the models did) is absolutely crucial: the middle bar, which lacks growth-rate uncertainty is far too narrow.
+%% \jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
 Who wanted this and why?
 I can write something about the comparison above instead?}.
-uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
-Propagating error from the generation interval also has important effects, however (95\% CI: 2.59--4.12).
-Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example -- 
-propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
+%% uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
+Propagating error from the generation interval also has important effects. 
+%% however (95\% CI: 2.59--4.12).
+Once these two are included, the impact of leaving out uncertainty in the dispersion is small, though noticeable, in this particular example.
+%% -- propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
 
@@
 Vertical lines represent the 95\% confidence intervals.
-\jsw{why not show all combinations}
+%% \jsw{why not show all combinations}
+%% {We discussed this, but came down on the side of simplicity. There are no examples I can think of where people included uncertainty in κ but left out one of the two big ones.}
 }
@@
 Here, we only focused on estimates of \Ro that were published within a very short frame of time.
-Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: understanding uncertainties in generation-interval distribution is important.
+Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: uncertainties in generation-interval distribution is important.
 As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
-Uncertainty in the exponential growth rate will inevitably decrease, 
+Uncertainty in the exponential growth rate will eventually decrease, 
 and the effect of uncertainty in the mean generation time will become more important.
-A recent epidemiological analysis of the patient-level data revealed large uncertainty in serial intervals, which is defined as the time between symptom onsets of an infector and an infectee, (mean: 7.5 days; 95\% CI: .3--19) \citep{nejmncov};
 

### 2020-01-31 wuhan.tex

@@
 
-More broadly, disease modelers should carefully consider different sources of uncertainties in the data generating process.
+More broadly, disease modelers always need to consider all potential sources of uncertainties in the data generating process and weigh their importance.
 Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
-When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible answers.
+When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible answers \citep{ma2014estimating}.
+In general, however, ignoring process error and fitting deterministic models can give overly narrow confidence intervals \citep{king2015avoidable}.
 
-Here, we only focused on estimates of \Ro that were published within a very short frame of time.
-Although our analysis reflects a snapshot of a fast moving epidemic, our lessons will hold: uncertainties in generation-interval distribution is important.
+Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 24th--26th).
+Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: understanding uncertainties in generation-interval distribution is important.
 As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
-Uncertainty in the exponential growth rate will eventually decrease, 
-and the effect of uncertainty in the mean generation time will become more important.
+Uncertainty in the exponential growth rate will inevitably decrease,
+and the effect of uncertainty in the generation-interval distribution will become even more important.
+We strongly emphasize the need for accurate characterization of the transmission chains via contact tracing and better statistical framework for inferring generation-interval distributions from such data \citep{britton2019estimation}.
+A combined effort between healthcare workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
 

### 2020-01-31 wuhan.tex
pushing mli paragraph

@@
 \newcommand{\jsw}[1]{\comment{magenta}{JSW}{#1}}
+\newcommand{\mli}[1]{\comment{magenta}{MLi}{#1}}
 \newcommand{\new}[1]{\textcolor{blue}{#1}}
@@
 
+\mli{Nevertheless, this is our best approximation in estimating the potential growth rate, and R0 and its uncertainty from these studies.
+To confidently evaluate and compare these parameters across studies, we would need to know the assumptions made from each research and be able to reproduce their estimates.
+Initial estimates of R0 is essential to guide control; however, discrepancies in parameter estimates (and uncertainties) across studies using similar information of the same outbreak can be a potential step back in making control.
+(i.e., are these ranges in R0 estimates due to assumptions modelers make, or is R0 more uncertain when we think?).
+To aid policymakers to guide control, we suggest modelers to (1) propagate uncertainties in parameters and honestly report them, (2) reproducible methods.
+
+}
+
 \section{Methods}

### 2020-01-31 wuhan.tex
push

--- 2020-01-31 wuhan.tex
@@
 
-\newcommand{\eref}[1]{(\ref{eq:#1})}
+\newcommand{\eref}[1]{Eq.~\ref{eq:#1}}
 \newcommand{\fref}[1]{Fig.~\ref{fig:#1}}
@@
 
-Since December 2019, the novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
-As of January 28th, 2020, the World Health Organization has confirmed 4593 cases, including 56 confirmed cases in 14 different countries, outside China \citep{who28report}.
-Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, 
-there is clear evidence that it can be directly transmitted between humans,
-posing a greater threat for its spread \citep{huang2020clinical,who26report}.
+Since December 2019, the novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
+Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, its ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
+As of January 30th, 2020, the World Health Organization (WHO) has confirmed 7818 cases, including 82 confirmed cases in 18 different countries, outside China \citep{who28report}.
+WHO has now declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
 
@@
 
-%% \jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
+\jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
 But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
-%% \jd{I have no problem doing this in the next round, though. Time is important.}
+\jd{I have no problem doing this in the next round, though. Time is important.}
 \fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.
 For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
-Comparing the models that include only some sources of uncertainty to the confidences intervals from the ``all'' model, we see that propagating error from the growth rate (which all but one of the models did) is absolutely crucial: the middle bar, which lacks growth-rate uncertainty is far too narrow.
-%% \jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
+Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
+\jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
 Who wanted this and why?
 I can write something about the comparison above instead?}.
-%% uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
-Propagating error from the generation interval also has important effects. 
-%% however (95\% CI: 2.59--4.12).
-Once these two are included, the impact of leaving out uncertainty in the dispersion is small, though noticeable, in this particular example.
-%% -- propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
+uncertainty in \Ro, measured with our 95\% CI, is more than two-fold in this case (95\% CI: 2.22--5.16).
+Propagating error from the generation interval also has important effects, however (95\% CI: 2.59--4.12).
+Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example -- 
+propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
 
@@
 Vertical lines represent the 95\% confidence intervals.
-%% \jsw{why not show all combinations}
-%% {We discussed this, but came down on the side of simplicity. There are no examples I can think of where people included uncertainty in κ but left out one of the two big ones.}
+\jsw{why not show all combinations}
 }
@@
 There are two explanations for this result.
-First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework \eref{gamma},
+First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
 $r$ has a greater overall effect on \Ro because it is associated with more uncertainty;
@@
 \cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1st and January 21st from major cities in China and other countries.
-\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22nd and January 23rd, respectively, and inferred \Ro\ via the Euler-Lotka equation \eref{euler}.
+\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22nd and January 23rd, respectively, and inferred \Ro\ via the Euler-Lotka equation (\eref{euler}).
 \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26th, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.

### 2020-01-31 wuhan.tex
Results

--- 2020-01-31 wuhan.tex
@@
 For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
-Comparing the models that include only some sources of uncertainty to the confidences intervals from the ``all'' model, we see that propagating error from the growth rate (which all but one of the models did) is absolutely crucial: the middle bar, which lacks growth-rate uncertainty is far too narrow.
-%% \jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
-Who wanted this and why?
-I can write something about the comparison above instead?}.
+Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar, which lacks growth-rate uncertainty, is far too narrow.
 %% uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
@@
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than ours (\fref{assumption}).
+We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
 
@@
 First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework \eref{gamma},
-$r$ has a greater overall effect on \Ro because it is associated with more uncertainty;
-the range of the 95\% confidence interval on $\mu_r$ spans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
-Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as explained previously.
+$r$ has a greater overall effect on \Ro because it is associated with more uncertainty (\fref{assumption}).
+%% \jd{I am less against these words, although I don't love having them. But we would need a bit more precision (easiest would be to report >2.4, < 1.6 or something (I made the specific numbers up).
+%% the range of the 95\% confidence interval on $\mu_r$ spans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
+Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as discussed previously.
 
@@
 Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate \Ro and compare it with the reported \Ro estimates.
-We estimate the average \Ro to be 3.1 (95\% CI: 2.1--5.7).
-While our pooled estimate of \Ro is similar to other reported values, our calculation gives wider confidence intervals than all of them.
+Our estimated \Ro from the pooled distribution has a median of 3.1 (95\% CI: 2.1--5.7).
+While the point estimate of \Ro is similar to other reported values from this date range, the confidence intervals are wider than those of other studies.
 This result does not imply that assumptions based on the pooled estimate are too weak;
 we believe that this confidence interval accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
-\swp{Bryan: any comment about why R0 isn’t at the lower end of previous estimates? (Guess someone might ask this)}
+%% \swp{Bryan: any comment about why R0 isn’t at the lower end of previous estimates? (Guess someone might ask this)}
+%% \jd{Why should it be at lower end?}
 

### 2020-01-31 wuhan.tex

--- 2020-01-31 wuhan.tex
@@
 
-\jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
+%% \jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
 But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
-\jd{I have no problem doing this in the next round, though. Time is important.}
+%% \jd{I have no problem doing this in the next round, though. Time is important.}
 \fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.
 For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
-Propagating error from the growth rate (which all but one of the models did) is absolutely crucial:
-\jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
+Comparing the models that include only some sources of uncertainty to the confidences intervals from the ``all'' model, we see that propagating error from the growth rate (which all but one of the models did) is absolutely crucial: the middle bar, which lacks growth-rate uncertainty is far too narrow.
+%% \jd{I really dislike these in-text numbers, plus it misses the point that the cruciality of them depends primarily on something else (the comparison of things that leave them out with the last two bars).
 Who wanted this and why?
 I can write something about the comparison above instead?}.
-uncertainty in \Ro, measured with our 95\% CI, is more than two-fold in this case (95\% CI: 2.22--5.16).
-Propagating error from the generation interval also has important effects, however (95\% CI: 2.59--4.12).
-Once these two are included, the impact of leaving out uncertainty in the dispersion would be small, in this particular example -- 
-propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
+%% uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
+Propagating error from the generation interval also has important effects. 
+%% however (95\% CI: 2.59--4.12).
+Once these two are included, the impact of leaving out uncertainty in the dispersion is small, though noticeable, in this particular example.
+%% -- propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
 
@@
 Or are they measuring different data types (because the multi-level scheme you use has the data in parallel, whereas I think the data used may be nested in ways I don't yet understand).}
+
+\mli{Nevertheless, this is our best approximation in estimating the potential growth rate, and R0 and its uncertainty from these studies.
+To confidently evaluate and compare these parameters across studies, we would need to know the assumptions made from each research and be able to reproduce their estimates.
+Initial estimates of R0 is essential to guide control; however, discrepancies in parameter estimates (and uncertainties) across studies using similar information of the same outbreak can be a potential step back in making control.
+(i.e., are these ranges in R0 estimates due to assumptions modelers make, or is R0 more uncertain when we think?).
+To aid policymakers to guide control, we suggest modelers to (1) propagate uncertainties in parameters and honestly report them, (2) reproducible methods.
+
+}
 

### 2020-01-31 wuhan.tex

--- 2020-01-31 wuhan.tex
@@
 There are two explanations for this result.
-First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework \eref{gamma},
+First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
 $r$ has a greater overall effect on \Ro because it is associated with more uncertainty (\fref{assumption}).

### 2020-01-31 wuhan.tex
stuff

@@
 
-We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components.
+We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components: the exponential growth rate, mean generation interval, and generation interval dispersion.
 We are hopeful that this will provide a guide to understanding and reconciling different estimates early in an epidemic.
-\jsw{remind us what they are: 
-three simple components: the growth rate, mean generation interval, and generation interval dispersion.}
 \jsw{Here is my take: you are doing a Nate Silver-esque approach, pooling estimates in a smart way without necessarily redoing all the models. Isn't that right? That's cool!  
@@
 
-We reviewed six modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
+We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
 Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.

### 2020-01-31 wuhan.tex
minor

@@
 To aid policymakers to guide control, we suggest modelers to (1) propagate uncertainties in parameters and honestly report them, (2) reproducible methods.
-
 }
@@
 We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were upload on preprint servers (bioRxiv, medRxiv, and SSRN), and one report was published from Imperial College London \citep{imaincov}.
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to preprint servers (bioRxiv, medRxiv, and SSRN), and one report was posted on the website of Imperial College London \citep{imaincov}.
 There is a wide variation in their statistical methods and the amount of data they used to infer \Ro.

### 2020-01-31 wuhan.tex
more minor tweaks

@@
 \\
-Sang Woo Park\textsuperscript{1}
+Sang Woo Park\textsuperscript{1,*}
 David Champredon\textsuperscript{2}
 David J.\,D.\ Earn\textsuperscript{3,4}
-Michael Li\textsuperscript{?}
-Joshua S. Weitz\textsuperscript{?}
-Bryan T. Grenfell\textsuperscript{?}
-Jonathan Dushoff\textsuperscript{?,4}
+Michael Li\textsuperscript{5}
+Joshua S. Weitz\textsuperscript{6, 7}
+Bryan T. Grenfell\textsuperscript{1,8,9}
+Jonathan Dushoff\textsuperscript{3,4,5,*}
 \\
 \bigskip
-\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, New Jersey, USA
+\textbf{1} Department of Ecology and Evolutionary Biology, Princeton University, Princeton, NJ, USA
 \\
@@
 \\
-\textbf{5} Department of ???
+\textbf{5} Department of Biology, McMaster University, Hamilton, ON, Canada
+\\
+\textbf{6} School of Biological Sciences, Georgia Institute of Technology, Atlanta, GA, USA
+\\
+\textbf{7} School of Physics, Georgia Institute of Technology, Atlanta, GA, USA
+\\
+\textbf{8} Division of International Epidemiology and Population Studies, Fogarty International Center, National Institutes of Health, Bethesda, MD, USA
+\\
+\textbf{9} Woodrow Wilson School of Public and International Affairs, Princeton University, Princeton, NJ, USA
 \\
@@
 
-% *swp2@princeton.edu
+*Corresponding authors: swp2@princeton.edu and dushoff@mcmaster.ca
 \end{flushleft}
@@
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
+However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can affect the estimates.
 
@@
 We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
-As none of these studies report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond their model assumptions.
+As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond their model assumptions.
 We do so by modeling explicitly or implicitly reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
@@
 \hline
-Study 1 & $\gammdist(\alpha=28, \beta=28/2.6)$ & 8.4 & 0.5 \\
+Study 1 & $\gammdist(\mathrm{mean}=2.6, \alpha=28)$ & 8.4 & 0.5 \\
 \hline
-Study 2 & $\gammdist(\alpha=67, \beta=67/2.92)$ & 8.4 & 0.2 \\
+Study 2 & $\gammdist(\mathrm{mean}=2.92, \alpha=67)$ & 8.4 & 0.2 \\
 \hline
-Study 3 & $\gammdist(\alpha=1400, \beta=1400/3.8)$ & 7.6 & 0.5 \\
+Study 3 & $\gammdist(\mathrm{mean}=3.8, \alpha=1400)$ & 7.6 & 0.5 \\
 \hline
-Study 4 & $\gammdist(\alpha=12, \beta=12/2.2)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
+Study 4 & $\gammdist(\mathrm{mean}=2.2, \alpha=12)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
 \hline
-Study 5 & $\gammdist(\alpha=54, \beta=54/5.47)$ & $\mathrm{Uniform}(7.6, 8.4)$ & 0.2\\
+Study 5 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)$ & 0.2\\
 \hline
@@
 We use these probability distributions to obtain a probability distribution for the exponential growth rate $r$.
-The Gamma distribution is parameterized by shape $\alpha$ and rate $\beta$ parameters.
+The Gamma distribution is parameterized by its mean and shape $\alpha$.
 Constant values are fixed according to Table 1.

### 2020-01-31 wuhan.tex
stuff

@@
 
-%% \jd{The pictures with linear scale for Ro (and r) are fine, and the Ro ones look similar.
-But I would say that it's a good principle to use log scales for things like Ro and r, where it's more or less clear that the important distances are multiplicative.}
-%% \jd{I have no problem doing this in the next round, though. Time is important.}
 \fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.
 For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
-Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar, which lacks growth-rate uncertainty, is far too narrow.
-%% uncertainty on \Ro, measured with our 95\% CI, is more than two fold in this case (95\% CI: 2.22--5.16).
+Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is far too narrow.
 Propagating error from the generation interval also has important effects.
-%% however (95\% CI: 2.59--4.12).
 Once these two are included, the impact of leaving out uncertainty in the dispersion is small, though noticeable, in this particular example.
-%% -- propagating errors from $r$ and $\bar G$ gives 95\% CI of 2.16--5.58 whereas propagating all errors gives 95\% CI of 2.06--5.89 (compare \textbf{growth rate + GI mean} with \textbf{all} in \fref{eff}).
 
@@
 Vertical lines represent the 95\% confidence intervals.
-%% \jsw{why not show all combinations}
-%% {We discussed this, but came down on the side of simplicity. There are no examples I can think of where people included uncertainty in κ but left out one of the two big ones.}
 }
@@
 we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
-For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the range they reported (2.0--3.1).
+For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the uncertainty range they reported (2.0--3.1).
 There are two explanations for this result.
@@
 $r$ has a greater overall effect on \Ro because it is associated with more uncertainty (\fref{assumption}).
-%% \jd{I am less against these words, although I don't love having them. But we would need a bit more precision (easiest would be to report >2.4, < 1.6 or something (I made the specific numbers up).
-%% the range of the 95\% confidence interval on $\mu_r$ spans over more than a two-fold difference, whereas the range of the 95\% confidence intervals on $\mu_G$ does not.
 Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as discussed previously.
@@
 we believe that this confidence interval accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
-%% \swp{Bryan: any comment about why R0 isn’t at the lower end of previous estimates? (Guess someone might ask this)}
-%% \jd{Why should it be at lower end?}
 

### 2020-01-31 wuhan.tex
update figs

@@
 \textbf{Comparisons of the reported parameter values with our pooled estimates.}
-We inferred point estimates, uniform distributions (shown in orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (see text).
-Black points and lines: reported parameter values and ranges of uncertainties considered.
-Black triangle: we assumed $\kappa=0.5$ for Study 1 as they do not report the parameter values that they used.
-Red points and lines: median and 95\% confidence intervals of our pooled estimates.
+We inferred point estimates (black), uniform distributions (orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (red; see text).
+Open triangle: we assumed $\kappa=0.5$ for Study 1 as they do not report their generation interval dispersion.
 }

### 2020-01-31 compare_assumption.R

@@ -46,10 +46,10 @@ rdata3 <- mutate(merge(rdata2, study_number)) %>%
   )
 
 g1 <- ggplot(rdata3) +
-  geom_point(aes(est, anon), col=c(1, 1, 1, 1, 1, 1, 2),
+  geom_point(aes(est, anon), col=c("purple", "purple", 1, "purple", "purple", "purple", 2),
              size=c(2, 2, 2, 2, 2, 2, 4)) +
-  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, 1, 1, 1, 1, 2),
-               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.5)) +
+  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c("purple", "purple", 1, "purple", "purple", "purple", 2),
+               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.2)) +
   scale_x_continuous(expression(Exponential~growth~rate~(days^{-1}))) +
   theme(
     axis.title.y = element_blank()
@@ -62,10 +62,10 @@ gdata2 <- mutate(merge(gdata, study_number)) %>%
   )
 
 g2 <- ggplot(gdata2) +
-  geom_point(aes(est, anon), col=c(1, 1, 1, 1, 1, 1, 2),
+  geom_point(aes(est, anon), col=c(1, 1, "orange", 1, "orange", "orange", 2),
              size=c(2, 2, 2, 2, 2, 2, 4)) +
-  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, 1, 1, 1, 1, 2),
-               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.5)) +
+  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, "orange", 1, "orange", "orange", 2),
+               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.2)) +
   scale_x_continuous("Mean generation time (days)") +
   theme(
     axis.title.y = element_blank()
@@ -83,7 +83,7 @@ g3 <- ggplot(kappadata2) +
              col=c(1, 1, 1, 1, 1, 1, 2),
              size=c(2, 2, 2, 2, 2, 2, 4)) +
[…]

### 2020-01-31 wuhan.tex
something about two kinds of noise (revised what was there)

--- 2020-01-31 wuhan.tex
@@
 
-Narrow confidence intervals associated with early estimates of \Ro partly depend on the statistical approaches used to estimate the exponential growth rate $r$.
+In this study, we focused on propagating errors arising from implicit or explicit estimates of growth rate and generation intervals.
+Other key issues underlying early estimates of \Ro\ include statistical independence and types of noise.
+
 Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-Although the formula for the cumulative curve (e.g., logistic model) is much simpler and familiar to many people, fitting a model to cumulative incidence, instead of raw incidence, can both bias parameter and give narrow confidence intervals if non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
+This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the result non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
 Naive fits to cumulative incidence data should be avoided.
-\jsw{Ok... it's not strictly speaking a logistical model fit is it?  Also, this sentence is quite loose ('familiar to many people', which people).}
 
-More broadly, disease modelers always need to consider all potential sources of uncertainties in the data generating process and weigh their importance.
-Two main sources are process error, which refers to stochasticity in the process of disease spread, and observation error, which often refers to imperfect reporting.
-When the amount observation error dominates the noise in the data, ignoring process error and fitting phenomenological models can give sensible answers \citep{ma2014estimating}.
-In general, however, ignoring process error and fitting deterministic models can give overly narrow confidence intervals \citep{king2015avoidable}.
+There are many real-world sources of noise in real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).
+Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
+This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \citep{ma2014estimating}).
+Modelers should be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \citep{king2015avoidable}).
 

### 2020-01-31 wuhan.tex
starting on abstract

--- 2020-01-31 wuhan.tex
@@
 \section*{Abstract}
+A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
+As epidemic progresses, many disease modelers have been prioritized their attention to estimate the basic reproductive number, which is defined as average number of secondary cases caused by a primary case.
+While their efforts are extremely valuable, their modeling approaches and, inevitably, their estimates vary widely.
 
@@
 
-Since December 2019, the novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
+Since December 2019, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
 Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, its ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.

### 2020-01-31 wuhan.tex
A bit more discussion; up to Mike's part.

--- 2020-01-31 wuhan.tex
@@
 Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 24th--26th).
-Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: understanding uncertainties in generation-interval distribution is important.
-As the epidemic progresses and more data becomes available, inference on epidemic trajectories will become more precise.
-Uncertainty in the exponential growth rate will inevitably decrease,
-and the effect of uncertainty in the generation-interval distribution will become even more important.
+Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: confidence intervals must combine different sources of uncertainty. 
+In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.
+
 We strongly emphasize the need for accurate characterization of the transmission chains via contact tracing and better statistical framework for inferring generation-interval distributions from such data \citep{britton2019estimation}.
@@
 We are hopeful that this will provide a guide to understanding and reconciling different estimates early in an epidemic.
-\jsw{Here is my take: you are doing a Nate Silver-esque approach, pooling estimates in a smart way without necessarily redoing all the models. Isn't that right? That's cool!  
-But, I have one nagging worry: the folks measured r - yet this method says that they measured the uncertainty incorrectly...
-My nagging worry has to do with the structure of the multi-level modeling scheme.
-Aren't they all using the same data?
-Or are they measuring different data types (because the multi-level scheme you use has the data in parallel, whereas I think the data used may be nested in ways I don't yet understand).}
 

### 2020-01-31 wuhan.tex
code
merge-pull

@@
 
-We strongly emphasize the need for accurate characterization of the transmission chains via contact tracing and better statistical framework for inferring generation-interval distributions from such data \citep{britton2019estimation}.
-A combined effort between healthcare workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
+We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical framework for inferring generation-interval distributions from such data \citep{britton2019estimation}.
+A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
+%% \jd{Leave the medical folks out; we don't want to ad
 

### 2020-01-31 wuhan.tex
.
push again

@@
 A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
-%% \jd{Leave the medical folks out; we don't want to ad
+%% \jd{Leave the medical folks out; we don't want to add to the pressure on them during an outbreak.}
+We also emphasize the value of transparency from modelers. 
+Model estimates during an outbreak, even in pre-prints, should include code links and complete explanations. 
+Ideally, the code should not rely on closed-source programs.
 
@@
 
-\mli{Nevertheless, this is our best approximation in estimating the potential growth rate, and R0 and its uncertainty from these studies.
+%% \mli{Nevertheless, this is our best approximation in estimating the potential growth rate, and R0 and its uncertainty from these studies.
 To confidently evaluate and compare these parameters across studies, we would need to know the assumptions made from each research and be able to reproduce their estimates.

### 2020-01-31 wuhan.tex
piush
de-comment the methods, don't think I changed anything

@@
 To aid policymakers to guide control, we suggest modelers to (1) propagate uncertainties in parameters and honestly report them, (2) reproducible methods.
-}
 
@@
 
-\dc{I think I understand why you do that. But I’m worried the uncertainty for r comes
-artificially from whatever methods the Studies used, not the actual observation process we
-would have by observing (imperfectly) directly the cases. I guess the studies have all more or
-less the same incidence data... why not estimating uncertainty for r directly from incidence}
-\swp{that is exactly the point... we want to be able to reflect their methods... added a sentence about it.}
-
 We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
@@
 Since the gamma distribution does not allow zeros, we use $\kappa =0.02$ instead for Study 6.
-\jsw{Just a question: was the data the same?
-I assume the models fit to different data?
-I like this idea, except for the fact that these are not different outbreaks.
-I feel like this kind of multilevel model would be right if we had the same fundameental disease dynamics taking palce in different places.
-But what happens if the models are just different, even if the data is the same.
-Again, I have a small philosophical doubt with the implication that the i-s denote the same disease mechanism...
-and why are all the hyper-parameters gammas?}
 
@@
 95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the posterior distribution.
-
-\bmb{general comments (Daniel may have taken notes on these already):
-  \begin{itemize}
-  \item (\textbf{later on}) try flatter-at-origin priors (e.g. half-Cauchy)? Stan implementation?
-  \item comment on broader uncertainty context (King/Rohani on observation + process error, major reporting irregularities, non-stationarity \ldots)
-  \end{itemize}
-}
+\texttt{R} code is available in GitHub (https://github.com/parksw3/nCoV\_framework).
 

### 2020-01-31 wuhan.tex
push

--- 2020-01-31 wuhan.tex
@@
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
-As epidemic progresses, many disease modelers have been prioritized their attention to estimate the basic reproductive number, which is defined as average number of secondary cases caused by a primary case.
+As epidemic progresses, many disease modelers have prioritized their attention to estimating the basic reproductive number \Ro, which is defined as the average number of secondary cases caused by a primary case.
 While their efforts are extremely valuable, their modeling approaches and, inevitably, their estimates vary widely.
+Here, we present a novel framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to arly estimates of \Ro\ for the 2019-nCoV outbreak.
+Our results demonstrate the importance of propagating uncertainties in all three quantities, in particular in the mean generation interval.
+We recommend that more efforts should be
 
@@
 95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the posterior distribution.
+\texttt{R} code is available in GitHub (https://github.com/parksw3/nCoV\_framework).
 

### 2020-01-31 wuhan.tex
spelling

@@
 While their efforts are extremely valuable, their modeling approaches and, inevitably, their estimates vary widely.
-Here, we present a novel framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to arly estimates of \Ro\ for the 2019-nCoV outbreak.
+Here, we present a novel framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro\ for the 2019-nCoV outbreak.
 Our results demonstrate the importance of propagating uncertainties in all three quantities, in particular in the mean generation interval.
@@
 instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
-Given an estimate of the exponential growth rate $r$ and an \emph{instrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
+Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
 number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:

### 2020-01-31 wuhan.tex
can't think of a good concluding sentence for the abstract...

@@
 Here, we present a novel framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro\ for the 2019-nCoV outbreak.
-Our results demonstrate the importance of propagating uncertainties in all three quantities, in particular in the mean generation interval.
-We recommend that more efforts should be
+Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.
 

### 2020-01-31 wuhan.tex
deleting comments
Tighten abstract

@@
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
-As epidemic progresses, many disease modelers have prioritized their attention to estimating the basic reproductive number \Ro, which is defined as the average number of secondary cases caused by a primary case.
-While their efforts are extremely valuable, their modeling approaches and, inevitably, their estimates vary widely.
-Here, we present a novel framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro\ for the 2019-nCoV outbreak.
+As the epidemic progresses, many disease modelers have prioritized estimating the basic reproductive number \Ro, defined as the average number of secondary cases caused by a primary case.
+While these efforts are extremely valuable, their modeling approaches and the resulting estimates vary widely.
+Here, we present a framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro\ for the 2019-nCoV outbreak.
 Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.

### 2020-01-31 wuhan.tex
minor details

--- 2020-01-31 wuhan.tex
@@
 \begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.5cm}|l}
- & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation interval\newline dispersion $\kappa$ \\
+ & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
 \hline
@@
 \begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.5cm}}
- & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation interval\newline dispersion $\kappa$  \\
+ & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$  \\
 \hline
@@
 
-\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared coefficient of variation (CV) $\kappa$ of the generation-interval distribution from different studies with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
 We find that there is a large uncertainty associated with the underlying parameters;
@@
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.
-We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
@@
 A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
-%% \jd{Leave the medical folks out; we don't want to add to the pressure on them during an outbreak.}
 We also emphasize the value of transparency from modelers.
@@
 We are hopeful that this will provide a guide to understanding and reconciling different estimates early in an epidemic.
-
-%% \mli{Nevertheless, this is our best approximation in estimating the potential growth rate, and R0 and its uncertainty from these studies.
-To confidently evaluate and compare these parameters across studies, we would need to know the assumptions made from each research and be able to reproduce their estimates.
-Initial estimates of R0 is essential to guide control; however, discrepancies in parameter estimates (and uncertainties) across studies using similar information of the same outbreak can be a potential step back in making control.
-(i.e., are these ranges in R0 estimates due to assumptions modelers make, or is R0 more uncertain when we think?).
-To aid policymakers to guide control, we suggest modelers to (1) propagate uncertainties in parameters and honestly report them, (2) reproducible methods.
 
@@
 Assuming that generation intervals follow a gamma distribution
-with the mean $\bar G$ and the squared coefficient of variation (CV$^2$) $\kappa$, 
+with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
 we have

### 2020-01-31 wuhan.tex
stuff

@@
 \scriptsize
-\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.5cm}|l}
+\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
  & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
@@
 \scriptsize
-\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.5cm}}
+\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.7cm}}
  & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$  \\

### 2020-01-31 wuhan.tex
Last conference with Daniel

@@
 Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.
+While rapid response during an outbreak can be valuable, avoiding over-confidence is also important.
+Modelers should work with field-workers to develop better methods for characterizing generation intervals.
 
@@
 \scriptsize
-\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
- & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
+\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.5cm}|l}
+ & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation interval\newline dispersion $\kappa$ \\
 \hline
@@
 \scriptsize
-\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.7cm}}
- & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$  \\
+\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.5cm}}
+ & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation interval\newline dispersion $\kappa$  \\
 \hline
@@
 
-\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the squared coefficient of variation (CV) $\kappa$ of the generation-interval distribution from different studies with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
 We find that there is a large uncertainty associated with the underlying parameters;
@@
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.
-We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the squared CV in generation time $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
@@
 This result does not imply that assumptions based on the pooled estimate are too weak;
-we believe that this confidence interval accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+we believe that this confidence interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
+Here, our focus is on certainty, not on the reason for these discrepancies.
+%% JD: WE should add a citation to the model-ensemble literature and say we don't explore but there's some reason to think that the pooled version might be relatively good.
 
@@
 Assuming that generation intervals follow a gamma distribution
-with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
+with the mean $\bar G$ and the squared coefficient of variation (CV$^2$) $\kappa$, 
 we have

### 2020-01-31 wuhan.tex
remove line numbers
one last read

@@
 \usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
-\linenumbers
+% \linenumbers
 
@@
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can affect the estimates.
+However, it can be difficult to assess the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can affect the estimates.
 
@@
 Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
-This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \citep{ma2014estimating}).
-Modelers should be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \citep{king2015avoidable}).
+This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \cite{ma2014estimating}).
+Modelers should be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable}).
 

### 2020-01-31 MCMC_gen.R
delete old files

@@ -1,51 +1,0 @@
-library(MCMCpack)
-load("sample_data.rda")
-
-rpar <- sapply(sample_data, function(x) {
-  mm <- mean(x$r)
-  shape <- mm^2/var(x$r)
-  rate <- shape/mm
-  
-  c(shape=shape, rate=rate, mm=mm)
-})
-
-priorfun <- function(theta) {
-  
-  
-  sum(dgamma(theta[1:nstudy], shape=rpar[1,], rate=rpar[2,], log=TRUE)) +
-    dgamma(theta[nstudy+1], 0.1, 0.1, log=TRUE) +
-    dgamma(theta[nstudy+2], 0.1, 0.1, log=TRUE)
-}
-
-postfun <- function(theta,
-                    nstudy=6) {
-  mean <- theta[nstudy+1]
-  shape <- theta[nstudy+2]
-  rate <- shape/mean
-  
-  ll <- sum(dgamma(theta[1:nstudy], shape=shape, rate=rate, log=TRUE)) +
-    priorfun(theta, nstudy=nstudy)
-  
-  if (is.na(ll) || is.nan(ll)) {
-    return(-Inf)
-  }  
-  
-  ll
-}
-
-reslist <- vector('list', 4)
-
-for (i in 1:4) {
-  reslist[[i]] <- MCMCmetrop1R(postfun,
-                               theta.init=c(rpar[3,], mean(rpar[3,]), 10),
-                               thin=200,
-                               burnin=40000,
-                               mcmc=40000,
-                               verbose=20000,
-                               seed=i)
-  
-}
-
-MCMC_r <- as.mcmc.list(reslist)
-
-save("MCMC_r", file="MCMC_r.rda")

### 2020-01-31 ncov.tex
final change
Caveats

--- 2020-01-31 wuhan.tex
@@
 
-\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
 We find that there is a large uncertainty associated with the underlying parameters;
@@
 \fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.
-For illustrative purposes, we use our pooled estimates, which represent a reasonable proxy for the state of knowledge as of 26 January.
+For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of 26 January.
 Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is far too narrow.
@@
 we believe that this confidence interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
 Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
 Here, our focus is on certainty, not on the reason for these discrepancies.
-%% JD: WE should add a citation to the model-ensemble literature and say we don't explore but there's some reason to think that the pooled version might be relatively good.
 
@@
 Since the gamma distribution does not allow zeros, we use $\kappa =0.02$ instead for Study 6.
+We note that this approach does not account for non-independence between the parameter estimates made by different modelers. 
+As we add more models, we expect the pooled estimates to become sharper even when the models no longer add more information.
+Thus, the pooled estimator should be interpreted with care.
 
@@
 
-\bibliography{wuhan}
+\bibliography{ncov}
 

### 2020-01-31 ncov.tex
minor tweaks

@@
 
-For each study, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Ro and the assumed values of $\bar G$ and $\kappa$ and calculating the exponential growth rate $r$ via the inverse of \eref{gamma}:
+For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of $\mathcal R_{0i}$ and the assumed values of $\bar G_i$ and $\kappa_i$ and calculating the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
 \begin{linenomath*}

### 2020-01-31 ncov.tex
update

@@
 Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.
-While rapid response during an outbreak can be valuable, avoiding over-confidence is also important.
+While rapid response during an outbreak is important, careful consideration of methodology is also important.
 Modelers should work with field-workers to develop better methods for characterizing generation intervals.
@@
 
-There are many real-world sources of noise in real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).
+There are many sources of noise in real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).
 Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
@@
 We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to preprint servers (bioRxiv, medRxiv, and SSRN), and one report was posted on the website of Imperial College London \citep{imaincov}.
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN), and one report was posted on the website of Imperial College London \citep{imaincov}.
 There is a wide variation in their statistical methods and the amount of data they used to infer \Ro.

### 2020-01-31 ncov.tex
update
AUTO

@@
 Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.
-While rapid response during an outbreak is important, careful consideration of methodology is also important.
+While rapid response during an outbreak is important, careful consideration of methodology is also worthwhile.
 Modelers should work with field-workers to develop better methods for characterizing generation intervals.
@@
 
+There are other important phenomena not covered by our simple framework.
+Examples that seem relevant to this outbreak include: changing reporting rates, changing generation intervals, and reporting delays (including the effects of weekends and holidays.)
+\jd{This \P\ should be fleshed out and citated.}
+
 Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 24th--26th).

### 2020-02-01 ncov.tex
two unresolved comments from DC

@@
 As the epidemic progresses, many disease modelers have prioritized estimating the basic reproductive number \Ro, defined as the average number of secondary cases caused by a primary case.
-While these efforts are extremely valuable, their modeling approaches and the resulting estimates vary widely.
+These efforts are extremely valuable;
+however, their modeling approaches and the resulting estimates vary widely, despite relying on similar data source.
 Here, we present a framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro\ for the 2019-nCoV outbreak.
@@
 Modelers should work with field-workers to develop better methods for characterizing generation intervals.
+\dc{This implies it is not currently the case but we do not give any justification for this assertion in this paper...
+that may be unfair.
+I know people in the UK put a lot of effort into this for Ebola.
+Maybe you have the same people in China doing the same thing.
+My point is that we should say it is not currently shared publicly.}
+\swp{Modelers should continue to work closely with field-workers to develop better methods for characterizing generation intervals and make their analyses and data publicly available?}
 
@@
 We evaluate six disparate models published online between January 24--26, 2020 that estimated \Ro\ for the 2019-nCoV outbreak \citep{imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
-We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$. 
+We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
+the pooled estimates reflect the amount of uncertainties present in modeling approaches and their underlying assumptions.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. 
@@
 we model this estimate as a Gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
-We then constructed a family of parameter sets -- which include $r$, $\bar G$, and $\kappa$ -- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates (see Methods).
+We then constructed a family of parameter sets -- which include $r$, $\bar G$, and $\kappa$ -- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates: $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make (see Methods).
 
@@
 
-\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) that we calculate from our multilevel model.
+\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model.
 We find that there is a large uncertainty associated with the underlying parameters;
@@
 Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
-It is also interesting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly.
+It is also worth to noting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
 
@@
 
-\fref{eff} shows how propagating uncertainty ($\mu_r$, $\mu_G$, and $\mu_\kappa$) in different combinations would affect estimates and CIs for \Ro.
+\fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro.
 For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of 26 January.
@@
 We also evaluate the estimates of \Ro across different studies by 
-replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals all but three cases.
+replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
+This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions drive the sensitiivty.
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but three cases.
 We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
@@
 Ideally, the code should not rely on closed-source programs.
+\dc{I agree in principle but that's too harsh. People should be free to use whatever they want (Matlab...) as long as they publish the methods and code.}
 
@@
 
-Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (1) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (2) assess the degree to which these uncertainties affect the estimate of \Ro.
+Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
 Assuming that generation intervals follow a gamma distribution
@@
 \end{linenomath*}
-We run 4 parallel Markov Chain Monte Carlo (MCMC) chains that consist of 200,000 burnin steps and 200,000 sampling steps.
+We run 4 independent Markov Chain Monte Carlo chains that consist of 200,000 burnin steps and 200,000 sampling steps.
 Posterior samples are thinned every 400 steps.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference}.
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference} --
+trace plots and marginal posterior distribution plots are presented in Appendix.
 95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the posterior distribution.
@@
 
+\pagebreak
+\appendix
+\renewcommand\thefigure{A\arabic{figure}}
+\setcounter{figure}{0}    
+\section*{Appendix}
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{posterior_chain.pdf}
+\caption{
+\textbf{Trace plots of the multilevel model.}
+Each chain is represented by a different color.
+}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{posterior_dist.pdf}
+\caption{
+\textbf{Marginal posterior distributions of the multilevel model.}
+Each chain is represented by a different color.
+}
+\end{figure}
+
 \end{document}

### 2020-02-01 compare_R0.R

@@ -66,3 +66,5 @@ g1 <- ggplot(R0all) +
   )
 
 ggsave("compare_R0.pdf", g1, width=6, height=4)
+ggsave("compare_R0.png", g1, width=6, height=4)
+

### 2020-02-01 figure2.R

@@ -52,9 +52,16 @@ compfun <- function(x) {
 dd_combine <- compfun(r)
 
 g1 <- ggplot(dd_combine) +
-  geom_point(aes(type, est)) +
-  geom_errorbar(aes(type, min=lwr, max=upr), width=0) +
+  geom_point(aes(type, est), size=7) +
+  geom_errorbar(aes(type, min=lwr, max=upr), width=0, lwd=2) +
   xlab("Uncertainty type") +
-  scale_y_continuous("Basic reproductive number", limits=c(2, 6), expand=c(0, 0))
+  scale_y_continuous("Basic reproductive number", limits=c(2, 6), expand=c(0, 0)) +
+  theme(
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.border = element_blank(),
+    axis.line = element_line()
+  )
 
-ggsave("figure2.pdf", g1, width=6, height=4)
+ggsave("figure2.pdf", g1, width=8, height=4)

### 2020-02-01 posterior.R

@@ -1,0 +1,111 @@
+library(coda)
+library(dplyr)
+library(tidyr)
+library(ggplot2); theme_set(theme_bw())
+library(colorspace)
+library(gridExtra)
+
+load("MCMC_all.rda")
+
+parnames <- c("mu[r]~(days^{-1})", "alpha[r]", "mu[G]~(days)", "alpha[G]", "mu[kappa]", "alpha[kappa]")
+
+gelman.diag(MCMC_all)
+
+MCMC_data <- lapply(MCMC_all, function(x) {
+  dd <- as.data.frame(x)
+  dd <- setNames(dd, parnames)
+  dd$nn <- 1:nrow(dd)
+  dd
+}) %>%
+  bind_rows(.id="chain") 
+
+## there's probably a better way of doing this
+## e.g., gather and use facet_wrap
+## but I want to customize y axis titles...
+
+g1 <- ggplot(MCMC_data) +
+  geom_line(aes(nn, `mu[r]~(days^{-1})`, col=chain)) +
+  scale_x_continuous("MCMC steps (thinned every 400 steps)", expand=c(0, 0),
+                     breaks=c(0, 100, 200, 300, 400), limits=c(0, 500)) +
+  scale_y_continuous(expression(mu[r]~(days^{-1}))) +
+  scale_colour_manual(values=colorspace::qualitative_hcl(4)) +
+  theme(
+    panel.grid = element_blank(),
+    legend.position = "none"
+  )
+
+g2 <- ggplot(MCMC_data) +
+  geom_line(aes(nn, `alpha[r]`, col=chain)) +
+  scale_x_continuous("MCMC steps (thinned every 400 steps)", expand=c(0, 0),
+                     breaks=c(0, 100, 200, 300, 400), limits=c(0, 500)) +
+  scale_y_continuous(expression(alpha[r])) +
+  scale_colour_manual(values=colorspace::qualitative_hcl(4)) +
+  theme(
+    panel.grid = element_blank(),
+    legend.position = "none"
+  )
+
+g3 <- ggplot(MCMC_data) +
+  geom_line(aes(nn, `mu[G]~(days)`, col=chain)) +
+  scale_x_continuous("MCMC steps (thinned every 400 steps)", expand=c(0, 0),
+                     breaks=c(0, 100, 200, 300, 400), limits=c(0, 500)) +
+  scale_y_continuous(expression(mu[G]~(days))) +
+  scale_colour_manual(values=colorspace::qualitative_hcl(4)) +
+  theme(
[…]

### 2020-02-02 ncov.tex
miscellaneous edits

@@
 \usepackage{xspace}
-\newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}} 
-\newcommand{\Ro}{\Rx{0}\xspace}
+\newcommand{\Rx}[1]{\ensuremath{{\mathcal R}_{#1}}\xspace}
+% https://tex.stackexchange.com/questions/86565/drawbacks-of-xspace
+% avoid double-\xspace
+\newcommand{\Ro}{\ensuremath{{\mathcal R}_{0}}\xspace}
 \newcommand{\RR}{\ensuremath{{\mathcal R}}}
@@
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
-As the epidemic progresses, many disease modelers have prioritized estimating the basic reproductive number \Ro, defined as the average number of secondary cases caused by a primary case.
-These efforts are extremely valuable;
-however, their modeling approaches and the resulting estimates vary widely, despite relying on similar data source.
-Here, we present a framework for comparing different estimates of \Ro\ across a wide range of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro\ for the 2019-nCoV outbreak.
+As the epidemic progresses, many disease modelers have prioritized
+\bmb{focused on? (this is what we say in the body of the paper)} estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case.
+These efforts are valuable;
+however, their modeling approaches and the resulting estimates vary widely, despite relying on similar data sources.
+\bmb{why is this a problem/opportunity?
+``however'' suggests some sort of problem, but it isn't explicitly stated what it is \ldots}
+% don't need {\ } after \Ro because we use xspace
+Here, we present a framework for comparing and combining different estimates of \Ro across a wide of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
 Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.
+
 While rapid response during an outbreak is important, careful consideration of methodology is also worthwhile. 
@@
 \swp{Modelers should continue to work closely with field-workers to develop better methods for characterizing generation intervals and make their analyses and data publicly available?}
+\bmb{The last two sentences seem disconnected from the rest. What was the point you wanted to make? (I think the abstract would work without them.)}
 
@@
 Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, its ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
-As of January 30th, 2020, the World Health Organization (WHO) has confirmed 7818 cases, including 82 confirmed cases in 18 different countries, outside China \citep{who28report}.
-WHO has now declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
+As of January 30, 2020, the World Health Organization (WHO) had confirmed 7818 cases, including 82 confirmed cases in 18 different countries, outside China \citep{who28report}.
+Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
 
-As the disease continues to spread, many researchers have published analyses of the outbreak, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+As the disease continued to spread, many researchers published analyses of the outbreak, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, it can be difficult to assess the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can affect the estimates.
+However, it can be difficult to assess a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.
 
-Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
-therefore, estimates of $\mathcal R_0$ from different models directly depend on their implicit assumptions about the shape of the generation-interval distribution and the exponential growth rate.
-Early in an epidemic, information is scarce and, inevitably, there is large uncertainty around case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
-We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.
+Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
+therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the shape of the generation-interval distribution and the exponential growth rate.
+Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
+We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources \citep{elderd2006uncertainty}.
 
-We evaluate six disparate models published online between January 24--26, 2020 that estimated \Ro\ for the 2019-nCoV outbreak \citep{imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
+We evaluate six disparate models published online between January 24--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
 We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
-the pooled estimates reflect the amount of uncertainties present in modeling approaches and their underlying assumptions.
+the pooled estimates reflect the uncertainties present in modeling approaches and their underlying assumptions.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. 
-We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
+We also use our framework to unravel \bmb{tease apart?} which assumptions of these different models led to their different estimates and confidence intervals.
 
@@
 \begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
- & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
+ & Basic reproductive\newline number \Ro & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
 \hline
@@
 
-We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
+We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 24 and January 26, 2020 (Table 1).
 As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond their model assumptions.
@@
 we model this estimate as a Gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
-We then constructed a family of parameter sets -- which include $r$, $\bar G$, and $\kappa$ -- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates: $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+We then constructed a family of parameter sets --- which include $r$, $\bar G$, and $\kappa$ --- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates for each parameter ($\mu_r$, $\mu_G$, and $\mu_\kappa$).
 The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make (see Methods).
@@
 \begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.7cm}}
- & Basic reproductive\newline number \Ro\ & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$  \\
+ & Basic reproductive\newline number \Ro & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$  \\
 \hline
@@
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro.
-For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of 26 January.
+For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 26.
 Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is far too narrow.
@@
   \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
+  \bmb{May I suggest that you use {\tt theme\_set(theme\_bw(base\_size=18))} (or some large value) rather than (or in addition to) increasing the sizes of the points and lines?}
 We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties.
@@
 replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
-This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions drive the sensitiivty.
+This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions drive the sensitivity.
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but three cases.
@@
 we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
-For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\mathcal R_0 = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the uncertainty range they reported (2.0--3.1).
+For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the uncertainty range they reported (2.0--3.1).
 There are two explanations for this result.
@@
 In this study, we focused on propagating errors arising from implicit or explicit estimates of growth rate and generation intervals.
-Other key issues underlying early estimates of \Ro\ include statistical independence and types of noise.
+Other key issues underlying early estimates of \Ro include statistical independence and types of noise.
 
@@
 This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the result non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
-Naive fits to cumulative incidence data should be avoided.
+Naive fits to cumulative incidence data should therefore be avoided.
 
-There are many sources of noise in real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).
+Many sources of noise affect real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).
 Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
@@
 
-Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 24th--26th).
+Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 24--26).
 Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: confidence intervals must combine different sources of uncertainty. 
@@
 Model estimates during an outbreak, even in pre-prints, should include code links and complete explanations.
-Ideally, the code should not rely on closed-source programs.
+For maximal reproducibility, the code should not rely on closed-source software.
 \dc{I agree in principle but that's too harsh. People should be free to use whatever they want (Matlab...) as long as they publish the methods and code.}
+\bmb{We could leave this out, or phrase it positively (``Methods based on open-source tools allow for maximal reproducibility.'')}
 
-We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro\ in terms of three simple components: the exponential growth rate, mean generation interval, and generation interval dispersion.
-We are hopeful that this will provide a guide to understanding and reconciling different estimates early in an epidemic.
+We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro in terms of three simple components: the exponential growth rate, mean generation interval, and generation interval dispersion.
+We hope this will help researchers understand and reconcile disparate estimates early in an epidemic.
 
@@
 
-\subsection{Gamma approximation framework for linking $r$ and $\mathcal R_0$}
+\subsection{Gamma approximation framework for linking $r$ and $\Ro$}
 
@@
 \begin{equation}
-1/\mathcal R_0 = \int \exp(-r\tau) g(\tau) d\tau.
+1/\Ro = \int \exp(-r\tau) g(\tau) d\tau.
 \label{eq:euler}
@@
 \begin{equation}
-\mathcal R_0 = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
+\Ro = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
 \label{eq:gamma}
@@
 
-We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 24th, 2020 and January 26th, 2020 (Table 1).
+We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 24 and January 26, 2020 (Table 1).
 Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN), and one report was posted on the website of Imperial College London \citep{imaincov}.
-There is a wide variation in their statistical methods and the amount of data they used to infer \Ro.
-\cite{imaincov} and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18th.
-\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1st and January 21st from major cities in China and other countries.
-\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22nd and January 23rd, respectively, and inferred \Ro\ via the Euler-Lotka equation (\eref{euler}).
-\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26th, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
+They vary widely in their statistical methods and in the amount of data used to infer \Ro.
+\cite{imaincov} and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18.
+\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
+\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22 and January 23, respectively, and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
+\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
 
@@
 
-For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of $\mathcal R_{0i}$ and the assumed values of $\bar G_i$ and $\kappa_i$ and calculating the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
+For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculating the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
 \begin{linenomath*}
 \begin{equation}
-r_i = \frac{{\mathcal R_{0i}}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
+r_i = \frac{\Rx{0i}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
 \end{equation}
@@
 \end{linenomath*}
-We run 4 independent Markov Chain Monte Carlo chains that consist of 200,000 burnin steps and 200,000 sampling steps.
+We run 4 independent Markov Chain Monte Carlo chains each consisting of 200,000 burnin steps and 200,000 sampling steps.
 Posterior samples are thinned every 400 steps.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference} --
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
 trace plots and marginal posterior distribution plots are presented in Appendix.
-95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the posterior distribution.
+95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
 \texttt{R} code is available in GitHub (https://github.com/parksw3/nCoV\_framework).
@@
 Each chain is represented by a different color.
+\bmb{it seems a little weird that the densities of the $\alpha$ parameters are positive at zero. Is this a kernel-density artifact?}
 }

### 2020-02-03 ncov.tex
tweaks

@@
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
-As the epidemic progresses, many disease modelers have prioritized
-\bmb{focused on? (this is what we say in the body of the paper)} estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case.
+As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case.
 These efforts are valuable;
@@
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. 
-We also use our framework to unravel \bmb{tease apart?} which assumptions of these different models led to their different estimates and confidence intervals.
+We also use our framework to tease apart which assumptions of these different models led to their different estimates and confidence intervals.
 
@@
   \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
-  \bmb{May I suggest that you use {\tt theme\_set(theme\_bw(base\_size=18))} (or some large value) rather than (or in addition to) increasing the sizes of the points and lines?}
 We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties.

### 2020-02-03 figure2.R

@@ -1,5 +1,5 @@
 library(coda)
-library(ggplot2); theme_set(theme_bw())
+library(ggplot2); theme_set(theme_bw(base_size=18))
 
 load("sample_data.rda")
 load("MCMC_all.rda")
@@ -55,7 +55,7 @@ g1 <- ggplot(dd_combine) +
   geom_point(aes(type, est), size=7) +
   geom_errorbar(aes(type, min=lwr, max=upr), width=0, lwd=2) +
   xlab("Uncertainty type") +
-  scale_y_continuous("Basic reproductive number", limits=c(2, 6), expand=c(0, 0)) +
+  scale_y_continuous("Basic reproductive number", limits=c(2, 6.5), expand=c(0, 0)) +
   theme(
     panel.grid.major.x = element_blank(),
     panel.grid.minor.x = element_blank(),

### 2020-02-03 posterior.R

@@ -97,7 +97,7 @@ MCMC_summ <- gather(MCMC_data, key, value, -nn, -chain) %>%
   mutate(key=factor(key, levels=parnames))
 
 gpost <- ggplot(MCMC_summ) +
-  geom_density(aes(value, col=chain)) +
+  geom_histogram(aes(value), position="identity", fill=NA, bins=50) +
   facet_wrap(~key, scale="free", nrow=3,
              labeller = label_parsed)  +
   scale_x_continuous("Parameter values", limits=c(0, NA), expand=c(0, 0)) +

### 2020-02-03 sample_author.R

@@ -51,8 +51,8 @@ sample_zhao <- function(n=1) {
     gbar=runif(n, 7.6, 8.4),
     kappa=0.2,
     R=rgamma(n, 54, rate=54/5.47),
-    r=(R^kappa-1)/(kappa*gbar),
-    rho=r*gbar
+    r=(R^kappa-1)/(kappa*8), ## using average gbar to calculate r; this better reflects what the authors do.
+    rho=r*8
   )
 }
 

### 2020-02-03 ncov.tex
consistency

@@
 \begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
- & Basic reproductive\newline number \Ro & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
+ & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
 \hline
@@
 \begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.7cm}}
- & Basic reproductive\newline number \Ro & Mean generation\newline time $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$  \\
+ & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$  \\
 \hline
@@
 \hline
-Study 5 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)$ & 0.2\\
+Study 5 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)^\ast$ & 0.2\\
 \hline
-Study 6 & $\exp(r \bar G)^\ast$ & $\mathrm{Uniform}(6, 10)$ & 0\\
+Study 6 & $\exp(r \bar G)^\dagger$ & $\mathrm{Uniform}(6, 10)$ & 0\\
 \hline
@@
 Constant values are fixed according to Table 1.
-$^\ast$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+$^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
+We still account for this uncertainty in our pooled estimates ($\mu_G$).
+$^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in \Ro they report, when combined with the range of $\bar G$ they consider.
@@
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation time $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation time $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
+We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation interval $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
 
@@
 There are two explanations for this result.
-First, even though the exponential growth rate $r$ and the mean generation time $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
+First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
 $r$ has a greater overall effect on \Ro because it is associated with more uncertainty (\fref{assumption}).
-Second, assuming a fixed generation time ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as discussed previously.
+Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as discussed previously.
 

### 2020-02-03 compare_assumption.R

@@ -66,7 +66,7 @@ g2 <- ggplot(gdata2) +
              size=c(2, 2, 2, 2, 2, 2, 4)) +
   geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, "orange", 1, "orange", "orange", 2),
                lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.2)) +
-  scale_x_continuous("Mean generation time (days)") +
+  scale_x_continuous("Mean generation interval (days)") +
   theme(
     axis.title.y = element_blank()
   )

### 2020-02-03 ncov.tex
adding Ben

@@
 Sang Woo Park\textsuperscript{1,*}
-David Champredon\textsuperscript{2}
+Benjamin M.\ Bolker\textsuperscript{2,3,4}
+David Champredon\textsuperscript{5}
 David J.\,D.\ Earn\textsuperscript{3,4}
-Michael Li\textsuperscript{5}
-Joshua S. Weitz\textsuperscript{6, 7}
-Bryan T. Grenfell\textsuperscript{1,8,9}
-Jonathan Dushoff\textsuperscript{3,4,5,*}
+Michael Li\textsuperscript{2}
+Joshua S.\ Weitz\textsuperscript{6, 7}
+Bryan T.\ Grenfell\textsuperscript{1,8,9}
+Jonathan Dushoff\textsuperscript{2,3,4,*}
 \\
@@
 \\
-\textbf{2} Department of Pathology and Laboratory Medicine, University of Western Ontario, London, Ontario, Canada
+\textbf{2} Department of Biology, McMaster University, Hamilton, ON, Canada
 \\
@@
 \\
-\textbf{5} Department of Biology, McMaster University, Hamilton, ON, Canada
+\textbf{5} Department of Pathology and Laboratory Medicine, University of Western Ontario, London, Ontario, Canada
 \\

### 2020-02-03 ncov.tex
ok
update

@@
 
-We evaluate six disparate models published online between January 24--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
+We evaluate six disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
 We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
@@
 \hline
-Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 1 & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
 \hline
-Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
+Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
 \hline
-Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
+Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
 \hline
-Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
+Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
 \hline
-Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
+Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
 \hline
-Study 6 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
+Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
+\hline
+Study 7 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
 \hline
@@
 
-We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 24 and January 26, 2020 (Table 1).
+We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 23 and January 26, 2020 (Table 1).
 As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond their model assumptions.
@@
 we used Gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
-For example, Study 2 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
+For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
 we model this estimate as a Gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
@@
 \hline
-Study 1 & $\gammdist(\mathrm{mean}=2.6, \alpha=28)$ & 8.4 & 0.5 \\
+Study 1 & $\mathrm{Uniform}(1.5, 3.5)$ & 10 & 1 \\
 \hline
-Study 2 & $\gammdist(\mathrm{mean}=2.92, \alpha=67)$ & 8.4 & 0.2 \\
+Study 2 & $\gammdist(\mathrm{mean}=2.6, \alpha=28)$ & 8.4 & 0.5 \\
 \hline
-Study 3 & $\gammdist(\mathrm{mean}=3.8, \alpha=1400)$ & 7.6 & 0.5 \\
+Study 3 & $\gammdist(\mathrm{mean}=2.92, \alpha=67)$ & 8.4 & 0.2 \\
 \hline
-Study 4 & $\gammdist(\mathrm{mean}=2.2, \alpha=12)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
+Study 4 & $\gammdist(\mathrm{mean}=3.8, \alpha=1400)$ & 7.6 & 0.5 \\
 \hline
-Study 5 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)^\ast$ & 0.2\\
+Study 5 & $\gammdist(\mathrm{mean}=2.2, \alpha=12)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
 \hline
-Study 6 & $\exp(r \bar G)^\dagger$ & $\mathrm{Uniform}(6, 10)$ & 0\\
+Study 6 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)^\ast$ & 0.2\\
+\hline
+Study 7 & $\exp(r \bar G)^\dagger$ & $\mathrm{Uniform}(6, 10)$ & 0\\
 \hline
@@
 Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
-It is also worth to noting that none of the six studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
+It is also worth noting that none but one studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
 
@@
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro.
-For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 26.
+For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26.
 Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is far too narrow.
@@
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but three cases.
-We estimate slightly narrower confidence intervals for Study 2 and Study 6 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+We estimate slightly narrower confidence intervals for Study 3 and Study 7 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
@@
 we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
-For example, recalculating \Ro for Study 6 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the uncertainty range they reported (2.0--3.1).
+For example, recalculating \Ro for Study 7 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the uncertainty range they reported (2.0--3.1).
 There are two explanations for this result.
@@
 
-Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 24--26).
+Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 23--26).
 Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: confidence intervals must combine different sources of uncertainty. 
@@
 
-We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 24 and January 26, 2020 (Table 1).
+We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
 Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN), and one report was posted on the website of Imperial College London \citep{imaincov}.
@@
 
-\section*{Acknowledgements}
-
-We thank Ben Bolker for providing useful comments on the manuscript.
-
 \pagebreak

### 2020-02-03 MCMC_all.R

@@ -4,25 +4,29 @@ load("sample_data.rda")
 nsample <- 100000
 
 priorfun <- function(theta) {
-  sum(dgamma(theta[-c(1, 3, 5)], 0.1, 0.1/1, log=TRUE)) +
-    dgamma(theta[1], 0.1, 0.1*7, log=TRUE) +
-    dgamma(theta[3], 0.1, 0.1/7, log=TRUE) +
-    dgamma(theta[5], 0.1, 0.1/0.5, log=TRUE)
+  sum(dnorm(theta[-c(1, 3, 5)], 0, 10, log=TRUE)) +
+    dgamma(theta[1], 2, 2*7, log=TRUE) +
+    dgamma(theta[3], 2, 2/7, log=TRUE) +
+    dgamma(theta[5], 2, 2/0.5, log=TRUE)
 }
 
 sample_data$`Majumder et al. (2020)`$kappa <- 0.02
 
 postfun <- function(theta) {
+  if (any(theta < 0))
+    return(-Inf)
+  
+  ## theta[2], theta[4], theta[6] are sd
   mean_r <- theta[1]
-  shape_r <- theta[2]
+  shape_r <- mean_r^2/theta[2]^2
   rate_r <- shape_r/mean_r
   
   mean_gbar <- theta[3]
-  shape_gbar <- theta[4]
+  shape_gbar <- mean_gbar^2/theta[4]^2
   rate_gbar <- shape_gbar/mean_gbar
   
   mean_kappa <- theta[5]
-  shape_kappa <- theta[6]
+  shape_kappa <- mean_kappa^2/theta[6]^2
   rate_kappa <- shape_kappa/mean_kappa
   
   ss <- sample(nsample, 1)
@@ -44,17 +48,17 @@ postfun <- function(theta) {
 }
 
 V <- matrix(0, 6, 6)
-diag(V) <- c(2e-3, 50, 1, 200, 0.1, 0.5)/10
+diag(V) <- c(2e-3, 0.004, 0.3, 0.3, 0.1, 0.1)/10
 
 reslist <- vector('list', 4)
 
 for (i in 1:4) {
   reslist[[i]] <- MCMCmetrop1R(postfun,
-                               theta.init=c(0.2, 10, 8, 70, 0.2, 1),
-                               thin=400,
-                               burnin=200000,
-                               mcmc=200000,
-                               verbose=50000,
+                               theta.init=c(0.2, 0.1, 8, 1, 0.5, 0.5),
+                               thin=1000,
[…]

### 2020-02-03 compare_R0.R

@@ -67,4 +67,3 @@ g1 <- ggplot(R0all) +
 
 ggsave("compare_R0.pdf", g1, width=6, height=4)
 ggsave("compare_R0.png", g1, width=6, height=4)
-

### 2020-02-03 compare_assumption.R

@@ -42,14 +42,14 @@ avg_kappa <- data.frame(
 rdata3 <- mutate(merge(rdata2, study_number)) %>%
   bind_rows(avg_r) %>%
   mutate(
-    anon=factor(anon, levels=c("Pooled estimate", paste0("Study ", 6:1)))
+    anon=factor(anon, levels=c("Pooled estimate", paste0("Study ", 7:1)))
   )
 
 g1 <- ggplot(rdata3) +
-  geom_point(aes(est, anon), col=c("purple", "purple", 1, "purple", "purple", "purple", 2),
-             size=c(2, 2, 2, 2, 2, 2, 4)) +
-  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c("purple", "purple", 1, "purple", "purple", "purple", 2),
-               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.2)) +
+  geom_point(aes(est, anon), col=c("orange", "purple", "purple", 1, "purple", "purple", "purple", 2),
+             size=c(2, 2, 2, 2, 2, 2, 2, 4)) +
+  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c("orange", "purple", "purple", 1, "purple", "purple", "purple", 2),
+               lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.2)) +
   scale_x_continuous(expression(Exponential~growth~rate~(days^{-1}))) +
   theme(
     axis.title.y = element_blank()
@@ -58,14 +58,14 @@ g1 <- ggplot(rdata3) +
 gdata2 <- mutate(merge(gdata, study_number)) %>%
   bind_rows(avg_gen) %>%
   mutate(
-    anon=factor(anon, levels=c("Pooled estimate", paste0("Study ", 6:1)))
+    anon=factor(anon, levels=c("Pooled estimate", paste0("Study ", 7:1)))
   )
 
 g2 <- ggplot(gdata2) +
-  geom_point(aes(est, anon), col=c(1, 1, "orange", 1, "orange", "orange", 2),
-             size=c(2, 2, 2, 2, 2, 2, 4)) +
-  geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, "orange", 1, "orange", "orange", 2),
[…]

### 2020-02-03 figure2.R

@@ -55,7 +55,7 @@ g1 <- ggplot(dd_combine) +
   geom_point(aes(type, est), size=7) +
   geom_errorbar(aes(type, min=lwr, max=upr), width=0, lwd=2) +
   xlab("Uncertainty type") +
-  scale_y_continuous("Basic reproductive number", limits=c(2, 6.5), expand=c(0, 0)) +
+  scale_y_continuous("Basic reproductive number", limits=c(2, 4.6), expand=c(0, 0)) +
   theme(
     panel.grid.major.x = element_blank(),
     panel.grid.minor.x = element_blank(),

### 2020-02-03 published_estimate.R

@@ -1,29 +1,29 @@
 Rdata <- data.frame(
-  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder et al. (2020)", "Liu et al. (2020)"),
-  est=c(2.5, 2.2, 3.8, 5.47, NA, 2.92),
-  lwr=c(1.5, 1.4, 3.6, 4.16, 2, 2.28),
-  upr=c(3.5, 3.8, 4, 7.1, 3.1, 3.67)
+  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
+          "Zhao et al. (2020)", "Majumder and Mandl (2020)", "Liu et al. (2020)"),
+  est=c(NA, 2.5, 2.2, 3.8, 5.47, NA, 2.92),
+  lwr=c(1.5, 1.5, 1.4, 3.6, 4.16, 2, 2.28),
+  upr=c(3.5, 3.5, 3.8, 4, 7.1, 3.1, 3.67)
 )
 
 Rdata_adj <- data.frame(
-  study=c("Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder et al. (2020)", "Liu et al. (2020)"),
-  est=c(2.5, 2.2, 3.8, 5.47, NA, 2.92),
-  lwr=c(1.5, qgamma(0.025, 12, 12/2.2), 3.6, 4.16, exp(0.114 * qunif(0.025, 6, 10)), 2.28),
-  upr=c(3.5, qgamma(0.975, 12, 12/2.2), 4, 7.1, exp(0.114 * qunif(0.975, 6, 10)), 3.67)
+  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
+          "Zhao et al. (2020)", "Majumder and Mandl (2020)", "Liu et al. (2020)"),
+  est=c(NA, 2.5, 2.2, 3.8, 5.47, NA, 2.92),
+  lwr=c(qunif(0.025, 1.5, 3.6), 1.5, qgamma(0.025, 12, 12/2.2), 3.6, 4.16, exp(0.114 * qunif(0.025, 6, 10)), 2.28),
[…]

### 2020-02-03 sample_author.R

@@ -1,3 +1,25 @@
+sample_bedford <- function(n=1) {
+  dplyr::data_frame(
+    gbar=10,
+    kappa=1,
+    R=runif(n, 1.5, 3.5),
+    r=(R^kappa-1)/(kappa*gbar),
+    rho=r*gbar
+  )
+}
+
+sample_imai <- function(n=1) {
+  # pgamma(3.5, 28, 28/2.6)-pgamma(1.5, 28, 28/2.6)
+  
+  dplyr::data_frame(
+    gbar=8.4,
+    kappa=0.5,
+    R=rgamma(n, 28, rate=28/2.6),
+    r=(R^kappa-1)/(kappa*gbar),
+    rho=r*gbar
+  )
+}
+
 sample_imai <- function(n=1) {
   # pgamma(3.5, 28, 28/2.6)-pgamma(1.5, 28, 28/2.6)
   

### 2020-02-03 sample_data.R

@@ -4,9 +4,10 @@ nsample <- 100000
 
 set.seed(101)
 sample_data <- list(
+  `Bedford et al. (2020)`=sample_bedford(nsample),
   `Imai et al. (2020)`=sample_imai(nsample),
   `Liu et al. (2020)`=sample_liu(nsample),
-  `Majumder et al. (2020)`=sample_majumder(nsample),
+  `Majumder and Mandl (2020)`=sample_majumder(nsample),
   `Read et al. (2020)`=sample_read(nsample),
   `Riou and Althaus (2020)`=sample_riou(nsample),
   `Zhao et al. (2020)`=sample_zhao(nsample)

### 2020-02-03 study_number.R

@@ -1,5 +1,5 @@
 study_number <- data.frame(
-  anon=paste0("Study ", 1:6),
-  study=c("Imai et al. (2020)", "Liu et al. (2020)", 
-          "Read et al. (2020)", "Riou and Althaus (2020)", "Zhao et al. (2020)", "Majumder et al. (2020)")
+  anon=paste0("Study ", 1:7),
+  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Liu et al. (2020)", 
+          "Read et al. (2020)", "Riou and Althaus (2020)", "Zhao et al. (2020)", "Majumder and Mandl (2020)")
 )

### 2020-02-03 ncov.tex
tweaks here and there

@@
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, it can be difficult to assess a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.
+However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.
 
@@
 
-We evaluate six disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}. 
+We evaluate seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
@@
 
-We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 6 articles that were published online between January 23 and January 26, 2020 (Table 1).
+We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23 and January 26, 2020 (Table 1).
 As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond their model assumptions.
@@
 Surprisingly, no studies take into account how the variation in generation intervals affects their estimates of \Ro:
-all studies assumed fixed values for $\kappa$, ranging from 0 to 0.5.
+all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
 Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
@@
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro.
-For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26.
-Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is far too narrow.
-Propagating error from the generation interval also has important effects.
-Once these two are included, the impact of leaving out uncertainty in the dispersion is small, though noticeable, in this particular example.
+For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26 (\fref{eff}A).
+Comparing the models that include only some sources of uncertainty to the \textbf{all} model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is far too narrow.
+In this case, propagating error from the mean generation interval has negligible effect as the uncertainty in $r$ is much greater.
+Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare \textbf{growth rate + GI mean} with \textbf{all});
+reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential.
+
+As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important (\fref{eff}B).
+Only propagating error from the growth rate gives very narrow confidence intervals.
+Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow narrow confidence intervals.
+We expect this hypothetical example to better reflect the current situation, as increased data availability will allow researchers to estimate $r$ with more certainty.
 
@@
   \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
-We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties.
+We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties (A) based on our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) and (B) assuming a 4-fold reduction in uncertainty of our pooled estimate of the exponential growth rate (using $(\mu_r + 3\times\mathrm{median}(\mu_r))/4$, instead).
 \textbf{base}: \Ro estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
@@
 This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions drive the sensitivity.
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but three cases.
-We estimate slightly narrower confidence intervals for Study 3 and Study 7 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but 7 cases.
+We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 4 when we use our pooled estimate of the mean generation interval $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
+We estimate narrower confidence intervals for Study 5 and Study 7 when we use our pooled estimate of the mean generation interval $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
+We also estimate narrower confidence intervals for Study 1 when we replace their estimated $r$ or $\bar G$ with our pooled estimtates for similar reasons.
 
 Consistent with our previous observations (\fref{eff}),
-we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro (\fref{R0}).
-For example, recalculating \Ro for Study 7 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--9.8), which is much wider than the uncertainty range they reported (2.0--3.1).
+we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
+For example, recalculating \Ro for Study 7 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--8.6), which is much wider than the uncertainty range they reported (2.0--3.1).
 There are two explanations for this result.
@@
 Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as discussed previously.
+One exception is Study 1: we find that there estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
+Since they assume an exponentially distributed generation interval ($\kappa=1$), their estimate of \Ro is not very sensitive to $r$ and $\bar G$.
 
@@
 Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate \Ro and compare it with the reported \Ro estimates.
-Our estimated \Ro from the pooled distribution has a median of 3.1 (95\% CI: 2.1--5.7).
-While the point estimate of \Ro is similar to other reported values from this date range, the confidence intervals are wider than those of other studies.
+Our estimated \Ro from the pooled distribution has a median of 2.9 (95\% CI: 2.1--4.5).
+While the point estimate of \Ro is similar to other reported values from this date range, the confidence intervals are wider than all but one study.
 This result does not imply that assumptions based on the pooled estimate are too weak;
@@
 In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
-Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
+Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates;
+for example, Study 6 appears to be an outlier but the authors, in fact, explore several different scenarios with changes in reporting rate, under which their estimates of \Ro are similar to other estimates.
 Here, our focus is on certainty, not on the reason for these discrepancies.

### 2020-02-03 compare_R0.R

@@ -65,5 +65,5 @@ g1 <- ggplot(R0all) +
     legend.title = element_blank()
   )
 
-ggsave("compare_R0.pdf", g1, width=6, height=4)
-ggsave("compare_R0.png", g1, width=6, height=4)
+ggsave("compare_R0.pdf", g1, width=7, height=4)
+ggsave("compare_R0.png", g1, width=7, height=4)

### 2020-02-03 figure2.R

@@ -1,5 +1,6 @@
 library(coda)
-library(ggplot2); theme_set(theme_bw(base_size=18))
+library(ggplot2); theme_set(theme_bw(base_size=16))
+library(gridExtra)
 
 load("sample_data.rda")
 load("MCMC_all.rda")
@@ -51,10 +52,13 @@ compfun <- function(x) {
 
 dd_combine <- compfun(r)
 
+dd_combine2 <- compfun((r + median(r)*3)/4)
+
 g1 <- ggplot(dd_combine) +
   geom_point(aes(type, est), size=7) +
   geom_errorbar(aes(type, min=lwr, max=upr), width=0, lwd=2) +
   xlab("Uncertainty type") +
+  ggtitle("A. Baseline") +
   scale_y_continuous("Basic reproductive number", limits=c(2, 4.6), expand=c(0, 0)) +
   theme(
     panel.grid.major.x = element_blank(),
@@ -64,4 +68,20 @@ g1 <- ggplot(dd_combine) +
     axis.line = element_line()
   )
 
-ggsave("figure2.pdf", g1, width=8, height=4)
+g2 <- ggplot(dd_combine2) +
+  geom_point(aes(type, est), size=7) +
+  geom_errorbar(aes(type, min=lwr, max=upr), width=0, lwd=2) +
+  xlab("Uncertainty type") +
+  ggtitle("B. Reduced uncertainty in the growth rate") +
+  scale_y_continuous("Basic reproductive number", limits=c(2.3, 3.8), expand=c(0, 0)) +
+  theme(
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.border = element_blank(),
+    axis.line = element_line()
+  )
+
+gtot <- arrangeGrob(g1, g2, nrow=1)
+
+ggsave("figure2.pdf", gtot, width=12, height=6)

### 2020-02-04 ncov.tex
update

@@
 $^\dagger$We assume $\kappa = 0.5$ in our analysis.
-$^\ddagger$The authors presented \Ro estimates under different assumptions; we use their baseline scenario in our analysis.
+$^\ddagger$The authors presented \Ro estimates under different assumptions regarding the reporting rate; we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
 }
@@
 Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates;
-for example, Study 6 appears to be an outlier but the authors, in fact, explore several different scenarios with changes in reporting rate, under which their estimates of \Ro are similar to other estimates.
+for example, Study 6 may appears to be an outlier but, in fact, the authors explore several different scenarios that allows for changes in reporting rate over time, under which their estimates of \Ro are similar to other estimates.
 Here, our focus is on certainty, not on the reason for these discrepancies.
@@
 In this study, we focused on propagating errors arising from implicit or explicit estimates of growth rate and generation intervals.
+These estimates of the exponential growth rate depend on how incidence has been reported.
+For emerging pathogens such as 2019-nCoV, we expect the reporting rate to increase over time due to limited resources and uncertainty in case definition initially;
+for example, \cite{zhaoncov} (Study 6) found that that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
+
 Other key issues underlying early estimates of \Ro include statistical independence and types of noise.
@@
 Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the result non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
+This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
 Naive fits to cumulative incidence data should therefore be avoided.

### 2020-02-04 ncov.tex
more tweaks

@@
 In this study, we focused on propagating errors arising from implicit or explicit estimates of growth rate and generation intervals.
-These estimates of the exponential growth rate depend on how incidence has been reported.
-For emerging pathogens such as 2019-nCoV, we expect the reporting rate to increase over time due to limited resources and uncertainty in case definition initially;
-for example, \cite{zhaoncov} (Study 6) found that that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
-
 Other key issues underlying early estimates of \Ro include statistical independence and types of noise.
@@
 There are other important phenomena not covered by our simple framework. 
-Examples that seem relevant to this outbreak include: changing reporting rates, changing generation intervals, and reporting delays (including the effects of weekends and holidays.)
+Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
+For emerging pathogens such as 2019-nCoV, we expect the reporting rate to increase over time due to limited resources and uncertainty in case definition initially;
+for example, \cite{zhaoncov} (Study 6) found that that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
+
 \jd{This \P\ should be fleshed out and citated.}
@@
 
-We reviewed 6 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN), and one report was posted on the website of Imperial College London \citep{imaincov}.
+We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
 They vary widely in their statistical methods and in the amount of data used to infer \Ro.
-\cite{imaincov} and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18.
+\cite{bedfordncov}, \cite{imaincov}, and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
 \cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.

### 2020-02-04 ncov.tex
removing evaluate from the paper

@@
 
-We evaluate seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
@@
 
-We also evaluate the estimates of \Ro across different studies by 
+We also compare the estimates of \Ro across different studies by 
 replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
@@
 For emerging pathogens such as 2019-nCoV, we expect the reporting rate to increase over time due to limited resources and uncertainty in case definition initially;
-for example, \cite{zhaoncov} (Study 6) found that that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
-
-\jd{This \P\ should be fleshed out and citated.}
+in particular, \cite{zhaoncov} (Study 6) found that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
+Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{harris1990reporting, tsuzuki2017dynamics, tariq2019assessing}.
+Even if there is a constant delay between infection and detection, we would expect the associated confidence intervals to be affected (estimates of the growth rate will be unbiased).
+Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
+Disease modelers who want to make serious inference about an outbreak should consider all these factors.
 
@@
 Model estimates during an outbreak, even in pre-prints, should include code links and complete explanations.
-For maximal reproducibility, the code should not rely on closed-source software.
-\dc{I agree in principle but that's too harsh. People should be free to use whatever they want (Matlab...) as long as they publish the methods and code.}
-\bmb{We could leave this out, or phrase it positively (``Methods based on open-source tools allow for maximal reproducibility.'')}
+We suggest using methods based on open-source tools allow for maximal reproducibility.
 
-We have provided a basis for evaluating and comparing exponential-growth based estimates of \Ro in terms of three simple components: the exponential growth rate, mean generation interval, and generation interval dispersion.
+We have provided a basis for comparing exponential-growth based estimates of \Ro in terms of three simple components: the exponential growth rate, mean generation interval, and generation interval dispersion.
 We hope this will help researchers understand and reconcile disparate estimates early in an epidemic.

### 2020-02-04 ncov.tex
going to take a break for a bit

@@
 As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case.
-These efforts are valuable;
-however, their modeling approaches and the resulting estimates vary widely, despite relying on similar data sources.
-\bmb{why is this a problem/opportunity?
-``however'' suggests some sort of problem, but it isn't explicitly stated what it is \ldots}
-% don't need {\ } after \Ro because we use xspace
+While their efforts are valuable, it can be difficult to navigate through different analyses as their modeling approaches and the resulting estimates vary widely, despite relying on similar data sources.
 Here, we present a framework for comparing and combining different estimates of \Ro across a wide of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
 Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.
-
-While rapid response during an outbreak is important, careful consideration of methodology is also worthwhile. 
-Modelers should work with field-workers to develop better methods for characterizing generation intervals.
-\dc{This implies it is not currently the case but we do not give any justification for this assertion in this paper...
-that may be unfair.
-I know people in the UK put a lot of effort into this for Ebola.
-Maybe you have the same people in China doing the same thing.
-My point is that we should say it is not currently shared publicly.}
-\swp{Modelers should continue to work closely with field-workers to develop better methods for characterizing generation intervals and make their analyses and data publicly available?}
-\bmb{The last two sentences seem disconnected from the rest. What was the point you wanted to make? (I think the abstract would work without them.)}
 
@@
 Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, its ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
-As of January 30, 2020, the World Health Organization (WHO) had confirmed 7818 cases, including 82 confirmed cases in 18 different countries, outside China \citep{who28report}.
+As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \citep{who02report}.
 Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
 
-As the disease continued to spread, many researchers published analyses of the outbreak, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+As the disease continued to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
@@
 Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
-It is also worth noting that none but one studies explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
+It is also worth noting that none but one study explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
 
@@
 In this case, propagating error from the mean generation interval has negligible effect as the uncertainty in $r$ is much greater.
-Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare \textbf{growth rate + GI mean} with \textbf{all});
+Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare \textbf{growth rate + GI mean} with \textbf{all}):
 reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential.
@@
 As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important (\fref{eff}B).
-Only propagating error from the growth rate gives very narrow confidence intervals.
-Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow narrow confidence intervals.
+Propagating error only from the growth rate gives very narrow confidence intervals.
+Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow confidence intervals.
 We expect this hypothetical example to better reflect the current situation, as increased data availability will allow researchers to estimate $r$ with more certainty.
@@
 replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
-This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions drive the sensitivity.
+This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions affect their estimates.
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but 7 cases.
@@
 Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates;
-for example, Study 6 may appears to be an outlier but, in fact, the authors explore several different scenarios that allows for changes in reporting rate over time, under which their estimates of \Ro are similar to other estimates.
+for example, Study 6 may appears to be an outlier but, in fact, the authors explore several different scenarios that allows for changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
 Here, our focus is on certainty, not on the reason for these discrepancies.
@@
 Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including with the amount of dispersion in the generation intervals:
-although our pooled estimates are relatively insensitive to the estimated uncertainty, our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro too sensitive to the estimates of the exponential growth rate $r$.
+our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro too sensitive to the estimates of the exponential growth rate $r$.
 
@@
 \begin{aligned}
-r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\alpha_r),\\
-\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\alpha_G),\\
-\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\alpha_\kappa).\\
+r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
+\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
+\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
 \end{aligned}
@@
 \end{linenomath*}
+where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
 We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step;
 this approach is analogous to Bayesian methods for analyzing phylogenetic data, which often rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
-Since the gamma distribution does not allow zeros, we use $\kappa =0.02$ instead for Study 6.
+Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
@@
 
-Weakly informative priors are assumed on the hyperparameters:
+We use weakly informative priors on hyperparameters representing pooled means and uninformative priors on hyperparameters representing between-study standard deviations:
 \begin{linenomath*}
@@
 \begin{aligned}
-\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=0.1)\\
-\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=0.1)\\
-\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=0.1)\\
-(\alpha_r, \alpha_G, \alpha_\kappa) &\sim \gammdist(\mathrm{mean}=1,\,\mathrm{shape}=0.1).
+\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
+\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
+\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
+(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{Half-Normal}(0, 10).
 \end{aligned}
@@
 \end{linenomath*}
-We run 4 independent Markov Chain Monte Carlo chains each consisting of 200,000 burnin steps and 200,000 sampling steps.
-Posterior samples are thinned every 400 steps.
+
+We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
+Posterior samples are thinned every 1000 steps.
 Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};

### 2020-02-04 ncov.tex
prior done

@@
 
-We use weakly informative priors on hyperparameters representing pooled means and uninformative priors on hyperparameters representing between-study standard deviations:
+We use weakly informative priors on hyperparameters representing pooled means and noninformative priors on hyperparameters representing between-study standard deviations:
 \begin{linenomath*}
@@
 \mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
-(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{Half-Normal}(0, 10).
+(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
 \end{aligned}
@@
 \end{linenomath*}
+We initially imposed Gamma priors with low shape parameters ($< 1$) on between-study shape parameters (i.e., the inverse of the between-study squared coefficient of variation) of our hierarchical model but later found out that the resulting posterior distribution could be sensitive to our choices of shape parameters of the Gamma priors \citep{gelman2006prior}.
+Furthermore, imposing strong priors (e.g., half-$t$ with 4 degrees of freedom and a standard deviation of 1) on between-study shape parameters \textit{a priori} assumes that there is large between-study variance and has a strong effect on the resulting posterior.
+On the other hand, imposing weak priors (e.g., half-Cauchy with a scale of 5) on between-study shape parameters \textit{a priori} allows for very low between-study variance and leads to poor mixing.
+Therefore, we followed recommendations outlined in \citep{gelman2006prior} by reparameterizing the hierarchical model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations: $\textrm{half-normal}(0, 10)$.
 
-We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
+We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 b
+urnin steps and 500,000 sampling steps.
 Posterior samples are thinned every 1000 steps.

### 2020-02-04 posterior.R
new posterior figures

@@ -7,7 +7,7 @@ library(gridExtra)
 
 load("MCMC_all.rda")
 
-parnames <- c("mu[r]~(days^{-1})", "alpha[r]", "mu[G]~(days)", "alpha[G]", "mu[kappa]", "alpha[kappa]")
+parnames <- c("mu[r]~(days^{-1})", "sigma[r]", "mu[G]~(days)", "sigma[G]", "mu[kappa]", "sigma[kappa]")
 
 gelman.diag(MCMC_all)
 
@@ -35,10 +35,10 @@ g1 <- ggplot(MCMC_data) +
   )
 
 g2 <- ggplot(MCMC_data) +
-  geom_line(aes(nn, `alpha[r]`, col=chain)) +
+  geom_line(aes(nn, `sigma[r]`, col=chain)) +
   scale_x_continuous("MCMC steps (thinned every 400 steps)", expand=c(0, 0),
                      breaks=c(0, 100, 200, 300, 400), limits=c(0, 500)) +
-  scale_y_continuous(expression(alpha[r])) +
+  scale_y_continuous(expression(sigma[r])) +
   scale_colour_manual(values=colorspace::qualitative_hcl(4)) +
   theme(
     panel.grid = element_blank(),
@@ -57,10 +57,10 @@ g3 <- ggplot(MCMC_data) +
   )
 
 g4 <- ggplot(MCMC_data) +
-  geom_line(aes(nn, `alpha[G]`, col=chain)) +
+  geom_line(aes(nn, `sigma[G]`, col=chain)) +
   scale_x_continuous("MCMC steps (thinned every 400 steps)", expand=c(0, 0),
                      breaks=c(0, 100, 200, 300, 400), limits=c(0, 500)) +
-  scale_y_continuous(expression(alpha[G])) +
+  scale_y_continuous(expression(sigma[G])) +
   scale_colour_manual(values=colorspace::qualitative_hcl(4)) +
   theme(
     panel.grid = element_blank(),
@@ -79,10 +79,10 @@ g5 <- ggplot(MCMC_data) +
   )
 
 g6 <- ggplot(MCMC_data) +
-  geom_line(aes(nn, `alpha[kappa]`, col=chain)) +
+  geom_line(aes(nn, `sigma[kappa]`, col=chain)) +
   scale_x_continuous("MCMC steps (thinned every 400 steps)", expand=c(0, 0),
[…]

### 2020-02-04 ncov.tex
new posterior figures
tweaks

@@
 
-As the disease continued to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
@@
 Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare \textbf{growth rate + GI mean} with \textbf{all}):
-reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential.
+reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
 
 As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important (\fref{eff}B).
-Propagating error only from the growth rate gives very narrow confidence intervals.
+Only propagating error from the growth rate gives very narrow confidence intervals. 
 Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow confidence intervals.
@@
 For emerging pathogens such as 2019-nCoV, we expect the reporting rate to increase over time due to limited resources and uncertainty in case definition initially;
-in particular, \cite{zhaoncov} (Study 6) found that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
+for example, \cite{zhaoncov} (Study 6) found that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
 Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{harris1990reporting, tsuzuki2017dynamics, tariq2019assessing}.
-Even if there is a constant delay between infection and detection, we would expect the associated confidence intervals to be affected (estimates of the growth rate will be unbiased).
+Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
 Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
-Disease modelers who want to make serious inference about an outbreak should consider all these factors.
+Accounting for these factors is crucial for making accurate inference.
 
@@
 Each chain is represented by a different color.
-\bmb{it seems a little weird that the densities of the $\alpha$ parameters are positive at zero. Is this a kernel-density artifact?}
 }

### 2020-02-05 ncov.tex
need to wordsmith

@@
 	\textbf\newline{
-		Reconciling early-outbreak preliminary estimates of the basic reproductive number and its uncertainty: a new framework and applications to the novel coronavirus (2019-nCoV) outbreak
+		Reconciling early-outbreak preliminary estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (2019-nCoV) outbreak
 	}
@@
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
-As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case.
-While their efforts are valuable, it can be difficult to navigate through different analyses as their modeling approaches and the resulting estimates vary widely, despite relying on similar data sources.
-Here, we present a framework for comparing and combining different estimates of \Ro across a wide of models by decomposing it into three key quantities (the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$) and apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
-Our results emphasize the importance of propagating uncertainties in all three quantities, in particular in the shape of the generation-interval distribution.
+As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case in an otherwise susceptible population.
+The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
+Here, we present a framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
+Our results emphasize the importance of propagating uncertainties including the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
 
@@
 Since December 2019, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
-Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, its ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
+Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of 2019-nCoV ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
 As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \citep{who02report}.
@@
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
-We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources \citep{elderd2006uncertainty}.
+We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty of \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
 
@@
 \textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
-Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
+Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
 $^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
@@
 We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23 and January 26, 2020 (Table 1).
-As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond their model assumptions.
+As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond to their model assumptions.
 We do so by modeling explicitly or implicitly reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
@@
 $^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
-Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in \Ro they report, when combined with the range of $\bar G$ they consider.
+Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
 }
@@
 Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
-It is also worth noting that none but one study explicitly or implicitly assumed an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption which used to be extremely common, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
+It is also worth noting that all but one study explicitly or implicitly did not assume an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption often used in disease modeling, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
 
@@
 We inferred point estimates (black), uniform distributions (orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (red; see text).
-Open triangle: we assumed $\kappa=0.5$ for Study 1 as they do not report their generation interval dispersion.
+Open triangle: we assumed $\kappa=0.5$ for Study 1 which does not report generation-interval dispersion.
 }
@@
 For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26 (\fref{eff}A).
-Comparing the models that include only some sources of uncertainty to the \textbf{all} model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is far too narrow.
-In this case, propagating error from the mean generation interval has negligible effect as the uncertainty in $r$ is much greater.
-Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare \textbf{growth rate + GI mean} with \textbf{all}):
-reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
+Comparing the models that include only some sources of uncertainty to the \textbf{all} model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is relatively narrow.
+In this case, propagating error from the mean generation interval has negligible effect compared to propagating the uncertainty in $r$.
+Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare \textbf{growth rate + GI mean} with \textbf{all}).
+For example, reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
 
@@
 Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow confidence intervals.
-We expect this hypothetical example to better reflect the current situation, as increased data availability will allow researchers to estimate $r$ with more certainty.
+We expect this hypothetical example to better reflect more recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
+\jsw{Just pointing out that new data might reveal that the growth rate is no longer exponential which would make an estimate of r mean something altogether different.}
 
@@
 Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as discussed previously.
-One exception is Study 1: we find that there estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-Since they assume an exponentially distributed generation interval ($\kappa=1$), their estimate of \Ro is not very sensitive to $r$ and $\bar G$.
+One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
+Since Study 1 assumes an exponentially distributed generation interval ($\kappa=1$), its estimate of \Ro is relatively insensitive to $r$ and $\bar G$.
 
@@
 for example, Study 6 may appears to be an outlier but, in fact, the authors explore several different scenarios that allows for changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
-Here, our focus is on certainty, not on the reason for these discrepancies.
+Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
 
@@
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we used a simple framework \citep{park2019practical} to compare estimates of \Ro for the novel coronavirus outbreak.
-Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including with the amount of dispersion in the generation intervals:
-our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro too sensitive to the estimates of the exponential growth rate $r$.
+Here, we used a multi-level Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals:
+our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro overly sensitive to the estimates of the exponential growth rate $r$.
 
@@
 Other key issues underlying early estimates of \Ro include statistical independence and types of noise.
+\jsw{Why is this here? Fine sentence, but belongs 2 paragraphs down and is largely redundant.}
+\swp{I proprose we drop these two sentences.}
 
@@
 Naive fits to cumulative incidence data should therefore be avoided.
+\jsw{Continue to this paragraph which builds on the point of growth rate r}
+\swp{No idea what we should continue.}
 
@@
 This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \cite{ma2014estimating}).
-Modelers should be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable}).
+Modelers should be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable, taylor2016stochasticity}).
 
@@
 Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
-Accounting for these factors is crucial for making accurate inference.
+Accounting for these factors is crucial for making accurate inferences.
 
 Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 23--26).
-Although our analysis only reflects a snapshot of a fast-moving epidemic, our lessons will hold: confidence intervals must combine different sources of uncertainty. 
+Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: confidence intervals must combine different sources of uncertainty. 
 In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.
 
-We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical framework for inferring generation-interval distributions from such data \citep{britton2019estimation}.
+We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical frameworks for inferring generation-interval distributions from such data \citep{britton2019estimation}.
 A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
@@
 
-We have provided a basis for comparing exponential-growth based estimates of \Ro in terms of three simple components: the exponential growth rate, mean generation interval, and generation interval dispersion.
-We hope this will help researchers understand and reconcile disparate estimates early in an epidemic.
+In summary, we have provided a basis for comparing exponential-growth based estimates of \Ro and its associated uncertainty in terms of three components: the exponential growth rate, mean generation interval, and generation interval dispersion.
+We hope this framework will help researchers understand and reconcile disparate estimates of disease strength early in an epidemic.
 
@@
 On the other hand, imposing weak priors (e.g., half-Cauchy with a scale of 5) on between-study shape parameters \textit{a priori} allows for very low between-study variance and leads to poor mixing.
-Therefore, we followed recommendations outlined in \citep{gelman2006prior} by reparameterizing the hierarchical model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations: $\textrm{half-normal}(0, 10)$.
+Therefore, we followed recommendations outlined in \citep{gelman2006prior} by reparameterizing the hierarchical model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
 
-We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 b
-urnin steps and 500,000 sampling steps.
+We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
 Posterior samples are thinned every 1000 steps.

### 2020-02-05 ncov.tex
worthsmithed (is this a word? probably not)

@@
 We do so by modeling explicitly or implicitly reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
-we used Gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
+we used gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
 For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
-we model this estimate as a Gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
+we model this estimate as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
 We then constructed a family of parameter sets --- which include $r$, $\bar G$, and $\kappa$ --- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates for each parameter ($\mu_r$, $\mu_G$, and $\mu_\kappa$).
@@
 We use these probability distributions to obtain a probability distribution for the exponential growth rate $r$.
-The Gamma distribution is parameterized by its mean and shape $\alpha$.
+The gamma distribution is parameterized by its mean and shape $\alpha$.
 Constant values are fixed according to Table 1.
@@
 \end{linenomath*}
-We initially imposed Gamma priors with low shape parameters ($< 1$) on between-study shape parameters (i.e., the inverse of the between-study squared coefficient of variation) of our hierarchical model but later found out that the resulting posterior distribution could be sensitive to our choices of shape parameters of the Gamma priors \citep{gelman2006prior}.
-Furthermore, imposing strong priors (e.g., half-$t$ with 4 degrees of freedom and a standard deviation of 1) on between-study shape parameters \textit{a priori} assumes that there is large between-study variance and has a strong effect on the resulting posterior.
-On the other hand, imposing weak priors (e.g., half-Cauchy with a scale of 5) on between-study shape parameters \textit{a priori} allows for very low between-study variance and leads to poor mixing.
-Therefore, we followed recommendations outlined in \citep{gelman2006prior} by reparameterizing the hierarchical model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+Here, we followed recommendations outlined in \citep{gelman2006prior} by reparameterizing the hierarchical gamma model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We initially imposed gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) of our hierarchical model but such priors can over-penalize the resulting posterior distribution, depending on their shape parameters \citep{gelman2006prior}.
+We also found that imposing strong priors on between-study shape parameters can be problematic as they \textit{a priori} assume large between-study variance, whereas imposing weak priors can allow for very low between-study variance and lead to poor mixing.
 

### 2020-02-05 ncov.tex
more edits!!!!!!!!!!!!!!!!!!!

@@
 \newcommand{\jd}[1]{\comment{magenta}{JD}{#1}}
+\newcommand{\djde}[1]{\comment{magenta}{DJDE}{#1}}
 \newcommand{\bmb}[1]{\comment{magenta}{BMB}{#1}}
@@
 Here, we present a framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
-Our results emphasize the importance of propagating uncertainties including the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak, estimating the \Ro from the pooled distribution as 2.9 (95\% CI: 2.1-4.5).
+Our results emphasize the importance of propagating propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+
+\swp{JSW suggests including our pooled estimate of R in the abstract. Not sure if we want that.}
+\swp{DJDE suggests using overline G rather than bar G. I think overline looks ugly.}
 
@@
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
-We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty of \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
+We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
 
@@
 We estimate narrower confidence intervals for Study 5 and Study 7 when we use our pooled estimate of the mean generation interval $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
-We also estimate narrower confidence intervals for Study 1 when we replace their estimated $r$ or $\bar G$ with our pooled estimtates for similar reasons.
+We also estimate narrower confidence intervals for Study 1 when we replace their estimated $r$ or $\bar G$ with our pooled estimates for similar reasons.
 
@@
 
-Of the six studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
+Of the seven studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
 This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
@@
 Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
-For emerging pathogens such as 2019-nCoV, we expect the reporting rate to increase over time due to limited resources and uncertainty in case definition initially;
-for example, \cite{zhaoncov} (Study 6) found that estimates of \Ro could change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 16--20.
+For emerging pathogens such as 2019-nCoV, we expect the reporting rate likely to be very low until sufficient resources become available and case definitions become certain;
+for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic protocols were released \citep{who17protocol}, and January 20.
 Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{harris1990reporting, tsuzuki2017dynamics, tariq2019assessing}.
@@
 In summary, we have provided a basis for comparing exponential-growth based estimates of \Ro and its associated uncertainty in terms of three components: the exponential growth rate, mean generation interval, and generation interval dispersion.
-We hope this framework will help researchers understand and reconcile disparate estimates of disease strength early in an epidemic.
+We hope this framework will help researchers understand and reconcile disparate estimates of disease transmission early in an epidemic.
 

### 2020-02-05 ncov.tex
tweak

@@
 For emerging pathogens such as 2019-nCoV, we expect the reporting rate likely to be very low until sufficient resources become available and case definitions become certain;
-for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic protocols were released \citep{who17protocol}, and January 20.
+for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
 Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{harris1990reporting, tsuzuki2017dynamics, tariq2019assessing}.

### 2020-02-05 ncov.tex
typo
tweak again

@@
 Here, we present a framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak, estimating the \Ro from the pooled distribution as 2.9 (95\% CI: 2.1-4.5).
+We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak\jsw{, estimating the \Ro from the pooled distribution as 2.9 (95\% CI: 2.1-4.5)}.
 Our results emphasize the importance of propagating propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
 
-\swp{JSW suggests including our pooled estimate of R in the abstract. Not sure if we want that.}
+\swp{JSW suggests including our pooled estimate of R in the abstract.
+Not sure if we want that.
+The pooled estimates have some caveats and they're based on very early estimates.
+I think it's OK to leave them out.}
 \swp{DJDE suggests using overline G rather than bar G. I think overline looks ugly.}
@@
 \jsw{Why is this here? Fine sentence, but belongs 2 paragraphs down and is largely redundant.}
-\swp{I proprose we drop these two sentences.}
+\swp{I propose we drop these two sentences.}
 

### 2020-02-05 ncov.tex
one more comment!

@@
 \swp{DJDE suggests using overline G rather than bar G. I think overline looks ugly.}
+
+\swp{Some of the papers that we look at have been published now.
+Where do we want to address that?
+Discussion?
+Methods?
+I'm thinking of adding a sentence to the Methods section where we review 7 papers: ``Some of these studies have now been published in peer reviewed journals or have been updated with better quantification of uncertainties [CITE].'' I think this is a fair thing to say.}
 

### 2020-02-05 ncov.tex
final update before I go to bed

--- 2020-02-05 ncov.tex
@@
 
-As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
@@
 \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
+Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+\swp{I think it's fair to add the last sentence because we do not want to misrepresent these studies.}
 

### 2020-02-05 ncov.tex
tweak

@@
 \cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
-\cite{zhaoncov} and \cite{liuncov} fitted exponential growth models to incidence data up to January 22 and January 23, respectively, and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
-\cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
+\cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
+\cite{liuncov} modeled the observed incidence up to January 23 as a function of previous incidence, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
+Finally, \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
 Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.

### 2020-02-05 ncov.tex
final tweak

@@
 \cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
-\cite{liuncov} modeled the observed incidence up to January 23 as a function of previous incidence, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
+\cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
 Finally, \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.

### 2020-02-05 ncov.tex
Some edits

@@
 
-\swp{JSW suggests including our pooled estimate of R in the abstract.
-Not sure if we want that.
-The pooled estimates have some caveats and they're based on very early estimates.
-I think it's OK to leave them out.}
-\swp{DJDE suggests using overline G rather than bar G. I think overline looks ugly.}
+\jd{no overline, no estimate in abstract.}
 
@@
 The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
-therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the shape of the generation-interval distribution and the exponential growth rate.
+therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the generation-interval distribution and the exponential growth rate.
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
@@
 Constant values are fixed according to Table 1.
-$^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
+\\\mbox{}\quad $^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
 We still account for this uncertainty in our pooled estimates ($\mu_G$).
-$^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+\\\mbox{}\quad $^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
@@
 It is also worth noting that all but one study explicitly or implicitly did not assume an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption often used in disease modeling, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
+\jd{Needs to be reworded, but do we still think it's even worth noting?}
 
@@
 Open triangle: we assumed $\kappa=0.5$ for Study 1 which does not report generation-interval dispersion.
+\jd{Study number doesn't match triangle, also is this clear elsewhere?}
 }
@@
 As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important (\fref{eff}B).
-Only propagating error from the growth rate gives very narrow confidence intervals. 
+Propagating error only from the growth rate gives very narrow confidence intervals in this case. 
 Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow confidence intervals.
@@
 \jsw{Just pointing out that new data might reveal that the growth rate is no longer exponential which would make an estimate of r mean something altogether different.}
+\jd{We should address issues with exponentiality in the Discussion; nothing is needed here I think.}
 
@@
 There are two explanations for this result.
+\jd{Reword below.}
 First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
@@
 Since Study 1 assumes an exponentially distributed generation interval ($\kappa=1$), its estimate of \Ro is relatively insensitive to $r$ and $\bar G$.
+\jd{Are these really both the same Study 1? I'd like to take a minute to think about this \P.}
 
@@
 In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
-Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates;
-for example, Study 6 may appears to be an outlier but, in fact, the authors explore several different scenarios that allows for changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
+Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
+We not in particular that, while the baseline example we used from Study 6 appears to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
 Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
@@
 
+\jd{Should we lead with the Bayesian fit or the approximation framework? The fit is useful but crude. The approximation is the part I hope will stick.}
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we used a multi-level Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
+Here, we used a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
 Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals:
 our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro overly sensitive to the estimates of the exponential growth rate $r$.
-
-In this study, we focused on propagating errors arising from implicit or explicit estimates of growth rate and generation intervals.
-Other key issues underlying early estimates of \Ro include statistical independence and types of noise.
-\jsw{Why is this here? Fine sentence, but belongs 2 paragraphs down and is largely redundant.}
-\swp{I propose we drop these two sentences.}
 
@@
 Naive fits to cumulative incidence data should therefore be avoided.
-\jsw{Continue to this paragraph which builds on the point of growth rate r}
-\swp{No idea what we should continue.}
 
@@
 This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \cite{ma2014estimating}).
-Modelers should be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable, taylor2016stochasticity}).
+Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable, taylor2016stochasticity}).
 
@@
 Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
-For emerging pathogens such as 2019-nCoV, we expect the reporting rate likely to be very low until sufficient resources become available and case definitions become certain;
+For emerging pathogens such as 2019-nCoV, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
 for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
@@
 
-Here, we focused on the estimates of \Ro that were published within a very short frame of time (January 23--26).
+Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26).
 Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: confidence intervals must combine different sources of uncertainty. 
@@
 Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
-\swp{I think it's fair to add the last sentence because we do not want to misrepresent these studies.}
 
@@
 
+\jd{Do you really prefer $wk^{-1} to /wk? I never understood that idea.}
 We use weakly informative priors on hyperparameters representing pooled means and noninformative priors on hyperparameters representing between-study standard deviations:
@@
 \end{linenomath*}
-Here, we followed recommendations outlined in \citep{gelman2006prior} by reparameterizing the hierarchical gamma model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
-We initially imposed gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) of our hierarchical model but such priors can over-penalize the resulting posterior distribution, depending on their shape parameters \citep{gelman2006prior}.
+Here, we followed recommendations outlined in \cite{gelman2006prior} by reparameterizing the hierarchical gamma model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We initially tried gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) of our hierarchical model but such priors can over-penalize the resulting posterior distribution, depending on their shape parameters \citep{gelman2006prior}.
 We also found that imposing strong priors on between-study shape parameters can be problematic as they \textit{a priori} assume large between-study variance, whereas imposing weak priors can allow for very low between-study variance and lead to poor mixing.
+\jd{Talk?}
 

### 2020-02-05 ncov.tex
cleaning

@@
 Here, we present a framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak\jsw{, estimating the \Ro from the pooled distribution as 2.9 (95\% CI: 2.1-4.5)}.
+We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
 Our results emphasize the importance of propagating propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
-
-\jd{no overline, no estimate in abstract.}
 
@@
 Constant values are fixed according to Table 1.
-\\\mbox{}\quad $^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
+$^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
 We still account for this uncertainty in our pooled estimates ($\mu_G$).
-\\\mbox{}\quad $^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+$^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
@@
 Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
-It is also worth noting that all but one study explicitly or implicitly did not assume an exponentially distributed generation interval (i.e., $\kappa=1$), an assumption often used in disease modeling, particularly implicitly (e.g., \cite{may1987transmission, anderson1991infectious, pybus2001epidemic, ferguson2005strategies}).
-\jd{Needs to be reworded, but do we still think it's even worth noting?}
 
@@
 We inferred point estimates (black), uniform distributions (orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (red; see text).
-Open triangle: we assumed $\kappa=0.5$ for Study 1 which does not report generation-interval dispersion.
-\jd{Study number doesn't match triangle, also is this clear elsewhere?}
+Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval dispersion.
 }
@@
 We expect this hypothetical example to better reflect more recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
-\jsw{Just pointing out that new data might reveal that the growth rate is no longer exponential which would make an estimate of r mean something altogether different.}
-\jd{We should address issues with exponentiality in the Discussion; nothing is needed here I think.}
 
@@
 There are two explanations for this result.
-\jd{Reword below.}
+\jd{Reword below.} \swp{Not sure why you want to reword this. We went over this paragraph so many times.}
 First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
 $r$ has a greater overall effect on \Ro because it is associated with more uncertainty (\fref{assumption}).
-Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$ as discussed previously.
+Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$.
 One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-Since Study 1 assumes an exponentially distributed generation interval ($\kappa=1$), its estimate of \Ro is relatively insensitive to $r$ and $\bar G$.
+Since Study 1 assumes an exponentially distributed generation interval ($\kappa=1$), its estimate of \Ro becomes more sensitive to $r$ and $\bar G$ when a narrower generation-interval distribution is assumed ($\kappa < 1$), resulting in a wider confidence interval.
 \jd{Are these really both the same Study 1? I'd like to take a minute to think about this \P.}
+\swp{Clarified this.}
 
@@
 \jd{Should we lead with the Bayesian fit or the approximation framework? The fit is useful but crude. The approximation is the part I hope will stick.}
+\swp{How about this?}
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we used a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
+Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
 Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals:
@@
 Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26).
+\swp{During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
+However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of \Ro should no longer rely solely on $r$.}
 Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: confidence intervals must combine different sources of uncertainty. 
-In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.
+In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate \swp{(and other epidemiological parameters)} will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.
 
@@
 
-\jd{Do you really prefer $wk^{-1} to /wk? I never understood that idea.}
 We use weakly informative priors on hyperparameters representing pooled means and noninformative priors on hyperparameters representing between-study standard deviations:
@@
 \jd{Talk?}
+\swp{???}
 

### 2020-02-05 ncov.tex
> 7.3 minutes

@@
 There are two explanations for this result.
-\jd{Reword below.} \swp{Not sure why you want to reword this. We went over this paragraph so many times.}
-First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical effects on \Ro under the gamma approximation framework (\eref{gamma} in Methods),
-$r$ has a greater overall effect on \Ro because it is associated with more uncertainty (\fref{assumption}).
+First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical mathematical effects on \Ro in our framework (\eref{gamma} in Methods),
+$r$ is more influential in this case because it is associated with more uncertainty (\fref{assumption}).
 Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$.
 One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-Since Study 1 assumes an exponentially distributed generation interval ($\kappa=1$), its estimate of \Ro becomes more sensitive to $r$ and $\bar G$ when a narrower generation-interval distribution is assumed ($\kappa < 1$), resulting in a wider confidence interval.
-\jd{Are these really both the same Study 1? I'd like to take a minute to think about this \P.}
-\swp{Clarified this.}
+This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that include this assumption make \Ro relatively insensitive and thus tend have particularly narrow confidence intervals.
 
@@
 
-\jd{Should we lead with the Bayesian fit or the approximation framework? The fit is useful but crude. The approximation is the part I hope will stick.}
-\swp{How about this?}
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
@@
 Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26).
-\swp{During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
-However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of \Ro should no longer rely solely on $r$.}
+During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
+However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibilty that $r$ is decreasing through time.
 Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: confidence intervals must combine different sources of uncertainty. 
-In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate \swp{(and other epidemiological parameters)} will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.
+In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate (and other epidemiological parameters) will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.
 
@@
 Here, we followed recommendations outlined in \cite{gelman2006prior} by reparameterizing the hierarchical gamma model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
-We initially tried gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) of our hierarchical model but such priors can over-penalize the resulting posterior distribution, depending on their shape parameters \citep{gelman2006prior}.
+We had initially tried gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) of our hierarchical model but found this put too much on large between-study variances; this is in fact a known problem \citep{gelman2006prior}, and thus was a poor prior choice.
 We also found that imposing strong priors on between-study shape parameters can be problematic as they \textit{a priori} assume large between-study variance, whereas imposing weak priors can allow for very low between-study variance and lead to poor mixing.
-\jd{Talk?}
-\swp{???}
 

### 2020-02-05 ncov.tex
minor tweak

@@
 One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that include this assumption make \Ro relatively insensitive and thus tend have particularly narrow confidence intervals.
+This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive and thus tend to have particularly narrow confidence intervals.
 

### 2020-02-05 ncov.tex
super minor tweak

@@
 Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
-We not in particular that, while the baseline example we used from Study 6 appears to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
+We note in particular that, while the baseline example we used from Study 6 appears to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
 Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
@@
 
+\section*{Acknowledgements}
+
+We thank Daihai He for providing helpful comments on the manuscript.
+
 \pagebreak

### 2020-02-05 ncov.tex
another typo

@@
 During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
-However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibilty that $r$ is decreasing through time.
+However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
 Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: confidence intervals must combine different sources of uncertainty. 

### 2020-02-06 ncov.tex
super minor tweak

@@
 Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
-We note in particular that, while the baseline example we used from Study 6 appears to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
+We note in particular that, while the baseline example we used from Study 6 may appear to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
 Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.

### 2020-02-06 ncov.tex
prior discussion, some tiny tweaks

@@
 \usepackage[top=1in,left=1in, right = 1in, footskip=1in]{geometry}
-
+\usepackage{url}
 \usepackage{graphicx}
@@
 We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
 They vary widely in their statistical methods and in the amount of data used to infer \Ro.
@@
 
-We use weakly informative priors on hyperparameters representing pooled means and noninformative priors on hyperparameters representing between-study standard deviations:
+We use weakly informative priors on hyperparameters representing pooled means and non-informative priors on hyperparameters representing between-study standard deviations:
 \begin{linenomath*}
@@
 \end{linenomath*}
-Here, we followed recommendations outlined in \cite{gelman2006prior} by reparameterizing the hierarchical gamma model in terms of its mean and standard deviation and imposing a noninformative but proper prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
-We had initially tried gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) of our hierarchical model but found this put too much on large between-study variances; this is in fact a known problem \citep{gelman2006prior}, and thus was a poor prior choice.
-We also found that imposing strong priors on between-study shape parameters can be problematic as they \textit{a priori} assume large between-study variance, whereas imposing weak priors can allow for very low between-study variance and lead to poor mixing.
+We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
+This phenomenon is a known problem \citep{gelman2006prior}.
+Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
+half-Student-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
+half-Cauchy(0,5)) can lead to poor mixing.
 
@@
 95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
-\texttt{R} code is available in GitHub (https://github.com/parksw3/nCoV\_framework).
+\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
 

### 2020-02-06 ncov.tex
justifying sticking with older estimates
thinspace in integral
typo

@@
 	\textbf\newline{
-		Reconciling early-outbreak preliminary estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (2019-nCoV) outbreak
+		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (2019-nCoV) outbreak
 	}
@@
 We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
-Our results emphasize the importance of propagating propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+Our results emphasize the importance of propagating uncertainties in all components of \Ro, especially the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
 
@@
 We also use our framework to tease apart which assumptions of these different models led to their different estimates and confidence intervals.
+Despite the availability of more recent and/or updated estimates of \Ro, we restrict ourselves to the estimates above in order to focus on the resolution of uncertainty in the earliest stages of an epidemic.
 
@@
 \begin{equation}
-1/\Ro = \int \exp(-r\tau) g(\tau) d\tau.
+1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
 \label{eq:euler}

### 2020-02-06 figure2.R
autocommit from git.mk

@@ -82,6 +82,9 @@ g2 <- ggplot(dd_combine2) +
     axis.line = element_line()
   )
 
+print(g1)
+print(g2)
+
 gtot <- arrangeGrob(g1, g2, nrow=1)
 
 ggsave("figure2.pdf", gtot, width=12, height=6)

### 2020-02-06 compare_assumption.R
push

@@ -1,3 +1,4 @@
+library(coda)
 library(ggplot2); theme_set(theme_bw())
 library(dplyr)
 library(gridExtra)

### 2020-02-06 ncov_lancet_ID.tex
lancet id version

--- 2020-02-06 ncov.tex
@@
 \usepackage{natbib}
-\bibliographystyle{chicago}
+\setcitestyle{numbers} 
+\setcitestyle{square}
+\bibliographystyle{unsrt}
 \date{\today}
@@
 
+\pagebreak
+
 \section*{Abstract}
+
+\subsection*{Background}
+
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
@@
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
-Here, we present a framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
+We present a framework for comparing and combining different estimates of \Ro across a wide range of models.
+
+\subsection*{Methods}
+
+We decompose the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+We use a Bayesian multilevel model to assess uncertainties associated with these quantities.
+We apply our framework to compare early estimates of \Ro for the 2019-nCoV outbreak.
+
+\subsection*{Findings}
+
+We find that most estimates rely on strong assumptions, particularly about the generation-interval dispersion.
+
+\subsection*{Interpretation}
+
 Our results emphasize the importance of propagating uncertainties in all components of \Ro, especially the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+
+\subsection*{Funding}
+
+BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC). ML was supported by Canadian Institutes of Health Research (CIHR).
 
@@
 Since December 2019, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
-Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of 2019-nCoV ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
+Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of 2019-nCoV ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical}.
 As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \citep{who02report}.
-Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
+% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
 
-As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious}).
+Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \citep{anderson1991infectious}.
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.
+However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups when the estimation methods and their underlying assumptions vary widely.
 
 Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
+The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wallinga2007generation};
 therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the generation-interval distribution and the exponential growth rate.
-Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
+Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding disease transmission.
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
@@
 We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
-the pooled estimates reflect the uncertainties present in modeling approaches and their underlying assumptions.
+the pooled estimates reflect the uncertainties present in modeling approaches.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
@@
 \begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
- & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
+ & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & Sources \\
 \hline
@@
 
-We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23 and January 26, 2020 (Table 1).
-As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond to their model assumptions.
-We do so by modeling explicitly or implicitly reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
+We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
+As most studies do not report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond to their model assumptions.
+We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
 we used gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
@@
 all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
-Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
+% Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
 
@@
 replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
-This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions affect their estimates.
+This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but 7 cases.
-We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we account for proper uncertainties in the generation-interval dispersion because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 5 and Study 7 when we use our pooled estimate of the mean generation interval $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
-We also estimate narrower confidence intervals for Study 1 when we replace their estimated $r$ or $\bar G$ with our pooled estimates for similar reasons.
+We estimate narrower confidence intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
+We also estimate narrower confidence intervals for Study 1 when we substitute their estimated $r$ or $\bar G$ with our pooled estimates for similar reasons.
+Substituting the reported $r$ or $\bar G$ from Study 1 with our pooled estimates give narrower confidence intervals for similar reasons.
 
-Consistent with our previous observations (\fref{eff}),
-we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
+We find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
 For example, recalculating \Ro for Study 7 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--8.6), which is much wider than the uncertainty range they reported (2.0--3.1).
@@
 This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \cite{ma2014estimating}).
-Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable, taylor2016stochasticity}).
+Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable}).
 
@@
 For emerging pathogens such as 2019-nCoV, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
-for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
-Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{harris1990reporting, tsuzuki2017dynamics, tariq2019assessing}.
+for example, Zhao \textit{et al.} \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
+Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{tariq2019assessing}.
 Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
@@
 Early in an outbreak, \Ro is difficult to estimate directly;
-instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
+instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
 Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
@@
 They vary widely in their statistical methods and in the amount of data used to infer \Ro.
-\cite{bedfordncov}, \cite{imaincov}, and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
-\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
-\cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
-\cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
-Finally, \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
+Bedford \textit{et al.} \cite{bedfordncov}, Imai \textit{et al.} \cite{imaincov}, and Riou and Althaus \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
+Read \textit{et al.} \cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
+Zhao \textit{et al.} \cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
+Liu \textit{et al.} \cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
+Finally, Majumder and Mandl \textit{et al.} \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
 Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
@@
 \end{linenomath*}
-We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We followed recommendations outlined in Gelman \textit{et al.} \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
 We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.

### 2020-02-06 ncov_lancet_ID.tex
almost there

@@
 Since December 2019, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
-Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of 2019-nCoV ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical}.
+Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of 2019-nCoV to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical}.
 As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \citep{who02report}.
@@
 
-As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious}).
+As the disease continues to spread, many researchers have been publishing their preliminary analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \citep{anderson1991infectious}.
@@
 where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
-We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step;
-this approach is analogous to Bayesian methods for analyzing phylogenetic data, which often rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
+We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
 Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.

### 2020-02-06 ncov_lancet_ID.tex
cut 3 more

@@
 
-Since December 2019, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
-Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of 2019-nCoV to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical}.
+Since December 2019 \citep{pneumonia}, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
+Although the virus is believed to have originated from animal reservoirs, the ability of 2019-nCoV to directly transmit between humans has posed a greater threat for its spread \citep{chan2020familial}.
 As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \citep{who02report}.
@@
 
-As the disease continues to spread, many researchers have been publishing their preliminary analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious}).
+As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \citep{anderson1991infectious}.
@@
 Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
-This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \cite{ma2014estimating}).
-Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable}).
+This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \cite{ma2014estimating}.
+Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \cite{king2015avoidable}.
 

### 2020-02-06 ncov_lancet_ID.tex
30 citations

@@
 
-Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
 Assuming that generation intervals follow a gamma distribution

### 2020-02-06 ncov_lancet_ID.tex
research in context

@@
 
-We decompose the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We use a Bayesian multilevel model to assess uncertainties associated with these quantities.
-We apply our framework to compare early estimates of \Ro for the 2019-nCoV outbreak.
+We decompose \Ro into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+We use a Bayesian multilevel model tomeasure uncertainties associated with these quantities.
+We assess early estimates of \Ro for the 2019-nCoV outbreak across these ranges of uncertainties.
 
@@
 
-We find that most estimates rely on strong assumptions, particularly about the generation-interval dispersion.
+We find that most early estimates of \Ro rely on strong assumptions, especially about the generation-interval dispersion.
+In particular, estimates that rely on narrow generation-interval distributions are overly sensitive to estimates of the exponential growth rate.
 
@@
 BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC). ML was supported by Canadian Institutes of Health Research (CIHR).
+
+\pagebreak
+
+\section*{Research in context}
 
@@
 We estimate narrower confidence intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
-We also estimate narrower confidence intervals for Study 1 when we substitute their estimated $r$ or $\bar G$ with our pooled estimates for similar reasons.
 Substituting the reported $r$ or $\bar G$ from Study 1 with our pooled estimates give narrower confidence intervals for similar reasons.

### 2020-02-07 ncov_lancet_ID.tex
editing lancet ID version

@@
 We decompose \Ro into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We use a Bayesian multilevel model tomeasure uncertainties associated with these quantities.
-We assess early estimates of \Ro for the 2019-nCoV outbreak across these ranges of uncertainties.
+We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.
+We assess early estimates of \Ro for the 2019-nCoV outbreak using the pooled estimates.
 
@@
 \section*{Research in context}
+
+\subsection*{Evidence before this study}
+
+7 analyses of the 2019-nCoV outbreak were published online as pre-prints between January 23--26, 2020.
+Their modeling approaches varied widely and estimates of uncertainties in \Ro ranged from 1.5 to 7.1.
+
+This section should include a description of all the evidence that the authors considered before undertaking this study.
+Authors should briefly state: the sources (databases, journal or book reference lists, etc) searched; the criteria used to include or exclude studies (including the exact start and end dates of the search), which should not be limited to English language publications; the search terms used; the quality (risk of bias) of that evidence; and the pooled estimate derived from meta-analysis of the evidence, if appropriate
+
+\subsection*{Added value of this study}
+
+Authors should describe here how their findings add value to the existing evidence
+
+\subsection*{Implications of all the available evidence}
+
+Authors should state the implications for practice or policy and future research of their study combined with existing evidence.Research in context panels should not contain references; key studies mentioned here should be referenced in the main text.
 
@@
 
-\section{Results}
+\section{Methods}
+
+\subsection{Description of the studies}
 
@@
 We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
+They vary widely in their modeling approaches: a branching process model \citep{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \citep{readncov}, an exponential growth model \citep{zhaoncov}, a Poisson offspring distribution model \citep{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \citep{majumderncov}.
+Four studies estimated \Ro by directly fitting their models to incidence data \citep{readncov, zhaoncov, liuncov, majumderncov}.
+The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0})
+Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+
+\subsection{Gamma approximation framework for linking $r$ and $\Ro$}
+
+Early in an outbreak, \Ro is difficult to estimate directly;
+instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
+Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
+number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
+\begin{linenomath*}
+\begin{equation}
+1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
+\label{eq:euler}
+\end{equation}
+\end{linenomath*}
+In other words, estimates of \Ro must
+depend on the assumptions about the
+exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+
+Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+Assuming that generation intervals follow a gamma distribution
+with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
+we have
+\begin{linenomath*}
+\begin{equation}
+\Ro = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
+\label{eq:gamma}
+\end{equation}
+\end{linenomath*}
+This equation demonstrates that a generation-interval distribution
+that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
+will give a higher estimate of \Ro for the same value of $r$.
+
+\subsection{Statistical framework}
+
 As most studies do not report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond to their model assumptions.
@@
 we model this estimate as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
-We then constructed a family of parameter sets --- which include $r$, $\bar G$, and $\kappa$ --- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates for each parameter ($\mu_r$, $\mu_G$, and $\mu_\kappa$).
-The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make (see Methods).
+For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculate the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
+\begin{linenomath*}
+\begin{equation}
+r_i = \frac{\Rx{0i}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
+\end{equation}
+\end{linenomath*}
+This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
+uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.
 
@@
 \end{table}
+
+We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
+\begin{linenomath*}
+\begin{equation}
+\begin{aligned}
+r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
+\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
+\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
+\end{aligned}
+\end{equation}
+\end{linenomath*}
+where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
+We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
+Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
+We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
+As we add more models, the pooled estimates can become sharper even when the models no longer add more information.
+Thus, the pooled estimator should be interpreted with care.
+
+We use weakly informative priors on hyperparameters representing pooled means and non-informative priors on hyperparameters representing between-study standard deviations:
+\begin{linenomath*}
+\begin{equation}
+\begin{aligned}
+\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
+\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
+\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
+(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
+\end{aligned}
+\end{equation}
+\end{linenomath*}
+We followed recommendations outlined in Gelman \textit{et al.} \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
+This phenomenon is a known problem \citep{gelman2006prior}.
+Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
+half-Student-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
+half-Cauchy(0,5)) can lead to poor mixing.
+
+We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
+Posterior samples are thinned every 1000 steps.
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
+trace plots and marginal posterior distribution plots are presented in Appendix.
+95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
+\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
+
+\section{Results}
 
@@
 
-\section{Methods}
+\section*{Contributors}
 
-\subsection{Gamma approximation framework for linking $r$ and $\Ro$}
+SWP and JD developed the statistical framework. 
+SWP reviewed the published literature.
+SWP performed the analysis. 
+SWP, BMB, and JD created the figures. 
+SWP and JD wrote the first draft.
+All authors contributed to the writing and approval of the final report.
 
-Early in an outbreak, \Ro is difficult to estimate directly;
-instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
-Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
-number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
-\begin{linenomath*}
-\begin{equation}
-1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
-\label{eq:euler}
-\end{equation}
-\end{linenomath*}
-In other words, estimates of \Ro must
-depend on the assumptions about the
-exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+\section*{Declaration of interests}
 
-Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
-Assuming that generation intervals follow a gamma distribution
-with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
-we have
-\begin{linenomath*}
-\begin{equation}
-\Ro = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
-\label{eq:gamma}
-\end{equation}
-\end{linenomath*}
-This equation demonstrates that a generation-interval distribution
-that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
-will give a higher estimate of \Ro for the same value of $r$.
-
-\subsection{Description of the studies}
-
-We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
-They vary widely in their statistical methods and in the amount of data used to infer \Ro.
-Bedford \textit{et al.} \cite{bedfordncov}, Imai \textit{et al.} \cite{imaincov}, and Riou and Althaus \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
-Read \textit{et al.} \cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
-Zhao \textit{et al.} \cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
-Liu \textit{et al.} \cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
-Finally, Majumder and Mandl \textit{et al.} \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
-Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
-
-\subsection{Statistical framework}
-
-For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculating the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
-\begin{linenomath*}
-\begin{equation}
-r_i = \frac{\Rx{0i}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
-\end{equation}
-\end{linenomath*}
-This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
-uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.
-
-We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
-\begin{linenomath*}
-\begin{equation}
-\begin{aligned}
-r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
-\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
-\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
-\end{aligned}
-\end{equation}
-\end{linenomath*}
-where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
-We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
-Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
-We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-As we add more models, we expect the pooled estimates to become sharper even when the models no longer add more information.
-Thus, the pooled estimator should be interpreted with care.
-
-We use weakly informative priors on hyperparameters representing pooled means and non-informative priors on hyperparameters representing between-study standard deviations:
-\begin{linenomath*}
-\begin{equation}
-\begin{aligned}
-\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
-\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
-\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
-(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
-\end{aligned}
-\end{equation}
-\end{linenomath*}
-We followed recommendations outlined in Gelman \textit{et al.} \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
-We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
-This phenomenon is a known problem \citep{gelman2006prior}.
-Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
-half-Student-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
-half-Cauchy(0,5)) can lead to poor mixing.
-
-We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
-Posterior samples are thinned every 1000 steps.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
-trace plots and marginal posterior distribution plots are presented in Appendix.
-95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
-\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
+We declare no competing interests.
 

### 2020-02-07 ncov_lancet_ID.tex
crap

@@
 \usepackage{natbib}
+\usepackage[superscript,biblabel]{cite}
 \setcitestyle{numbers} 
@@
 \\
-Sang Woo Park\textsuperscript{1,*}
-Benjamin M.\ Bolker\textsuperscript{2,3,4}
-David Champredon\textsuperscript{5}
-David J.\,D.\ Earn\textsuperscript{3,4}
-Michael Li\textsuperscript{2}
-Joshua S.\ Weitz\textsuperscript{6, 7}
-Bryan T.\ Grenfell\textsuperscript{1,8,9}
-Jonathan Dushoff\textsuperscript{2,3,4,*}
+Sang Woo Park BSc\textsuperscript{1,*}
+Benjamin M.\ Bolker PhD\textsuperscript{2,3,4}
+David Champredon PhD\textsuperscript{5}
+David J.\,D.\ Earn PhD\textsuperscript{3,4}
+Michael Li PhD\textsuperscript{2}
+Joshua S.\ Weitz PhD\textsuperscript{6, 7}
+Bryan T.\ Grenfell PhD\textsuperscript{1,8,9}
+Jonathan Dushoff PhD\textsuperscript{2,3,4,*}
 \\
@@
 
-We decompose \Ro into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+We reviewed 7 studies that were published online between January 23--26, 2020.
+We decompose their \Ro estimates into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
 We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.
@@
 
-7 analyses of the 2019-nCoV outbreak were published online as pre-prints between January 23--26, 2020.
-Their modeling approaches varied widely and estimates of uncertainties in \Ro ranged from 1.5 to 7.1.
+% 7 analyses of the 2019-nCoV outbreak were published online as pre-prints between January 23--26, 2020.
+% Their modeling approaches varied widely and estimates of uncertainties in \Ro ranged from 1.5 to 7.1.
 
@@
 
-Since December 2019 \citep{pneumonia}, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
-Although the virus is believed to have originated from animal reservoirs, the ability of 2019-nCoV to directly transmit between humans has posed a greater threat for its spread \citep{chan2020familial}.
-As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \citep{who02report}.
-% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
+Since December 2019 \cite{pneumonia}, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
+Although the virus is believed to have originated from animal reservoirs, the ability of 2019-nCoV to directly transmit between humans has posed a greater threat for its spread \cite{chan2020familial}.
+As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \cite{who02report}.
+% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \cite{whoemer}.
 
-As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious}).
-Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \citep{anderson1991infectious}.
+As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
+Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.
 We commend these researchers for their timely contribution and those who made the data publicly available.
@@
 Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wallinga2007generation};
+The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \cite{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \cite{wallinga2007generation};
 therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the generation-interval distribution and the exponential growth rate.
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding disease transmission.
-We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
+We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \cite{elderd2006uncertainty}.
 
-We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
@@
 We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
-They vary widely in their modeling approaches: a branching process model \citep{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \citep{readncov}, an exponential growth model \citep{zhaoncov}, a Poisson offspring distribution model \citep{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \citep{majumderncov}.
-Four studies estimated \Ro by directly fitting their models to incidence data \citep{readncov, zhaoncov, liuncov, majumderncov}.
-The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0})
-Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+Five studies \cite{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \cite{imaincov}; and one report was posted on nextstrain.org \cite{bedfordncov}.
+Their modeling approaches vary widely: a branching process model \cite{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \cite{readncov}, an exponential growth model \cite{zhaoncov}, a Poisson offspring distribution model \cite{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \cite{majumderncov}.
+Four studies estimated \Ro by directly fitting their models to incidence data \cite{readncov, zhaoncov, liuncov, majumderncov}.
+The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \cite{imaincov0})
+Some of these studies have now been published in peer-reviewed journals \cite{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \cite{readncov2}.
 
@@
 Early in an outbreak, \Ro is difficult to estimate directly;
-instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
-Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
-number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
+instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \cite{ma2014estimating}.
+Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \cite{champredon2015intrinsic}, the basic reproductive
+number can be estimated via the Euler-Lotka equation \cite{wallinga2007generation}:
 \begin{linenomath*}
@@
 
-Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+Here, we use the gamma approximation framework \cite{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
 Assuming that generation intervals follow a gamma distribution
@@
 We still account for this uncertainty in our pooled estimates ($\mu_G$).
-$^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+$^\dagger$Study 6 uses the IDEA model \cite{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
@@
 \end{linenomath*}
-where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
+where $\mu_r$, $\mu_G$, $\mu_\kappa$ represent the pooled estimates, and $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations.
 We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
@@
 
-We use weakly informative priors on hyperparameters representing pooled means and non-informative priors on hyperparameters representing between-study standard deviations:
+We use weakly informative priors on hyperparameters:
 \begin{linenomath*}
@@
 We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
-This phenomenon is a known problem \citep{gelman2006prior}.
+This phenomenon is a known problem \cite{gelman2006prior}.
 Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
@@
 Posterior samples are thinned every 1000 steps.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \cite{gelman1992inference};
 trace plots and marginal posterior distribution plots are presented in Appendix.
@@
 all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
-% Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
+% Assuming fixed parameter values can lead to overly strong conclusions \cite{elderd2006uncertainty}.
 
@@
 For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26 (\fref{eff}A).
-Comparing the models that include only some sources of uncertainty to the \textbf{all} model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (\textbf{GI mean}), which lacks growth-rate uncertainty, is relatively narrow.
+Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (``GI mean''), which lacks growth-rate uncertainty, is relatively narrow.
 In this case, propagating error from the mean generation interval has negligible effect compared to propagating the uncertainty in $r$.
-Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare \textbf{growth rate + GI mean} with \textbf{all}).
+Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare ``growth rate + GI mean'' with ``all'').
 For example, reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
@@
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but 7 cases.
-We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we account for proper uncertainties in the generation-interval dispersion because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we account for proper uncertainties in the generation-interval dispersion because they assume a narrow generation-interval distribution (compare ``base'' with ``GI variation'');
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
+We estimate narrower confidence intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare ``base'' with ``GI mean'') because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
 Substituting the reported $r$ or $\bar G$ from Study 1 with our pooled estimates give narrower confidence intervals for similar reasons.
@@
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
+Here, we use a gamma approximation \cite{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
 Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals:
@@
 Of the seven studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
+This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \cite{ma2014estimating, king2015avoidable}.
 Naive fits to cumulative incidence data should therefore be avoided.
@@
 This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \cite{ma2014estimating}.
-Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \cite{king2015avoidable}.
+Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \cite{king2015avoidable,taylor2016stochasticity}.
 
@@
 For emerging pathogens such as 2019-nCoV, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
-for example, Zhao \textit{et al.} \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
-Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{tariq2019assessing}.
-Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
-Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
+for example, Zhao \textit{et al.} \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \cite{who17protocol}, and January 20.
+Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \cite{tariq2019assessing}.
+% Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
+Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \cite{hethcote2002effects}.
 Accounting for these factors is crucial for making accurate inferences.
@@
 Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26).
-During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
+During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \cite{anderson1991infectious}.
 However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
@@
 
-We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical frameworks for inferring generation-interval distributions from such data \citep{britton2019estimation}.
+We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical frameworks for inferring generation-interval distributions from such data \cite{britton2019estimation}.
 A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.

### 2020-02-07 ncov_lancet_ID.tex
stuff

@@
 
-\usepackage{natbib}
-\usepackage[superscript,biblabel]{cite}
-\setcitestyle{numbers} 
-\setcitestyle{square}
-\bibliographystyle{unsrt}
+\usepackage[numbers,sort&compress]{natbib}
+\bibliographystyle{unsrtnat}
 \date{\today}
@@
 \\
-Sang Woo Park BSc\textsuperscript{1,*}
-Benjamin M.\ Bolker PhD\textsuperscript{2,3,4}
-David Champredon PhD\textsuperscript{5}
-David J.\,D.\ Earn PhD\textsuperscript{3,4}
-Michael Li PhD\textsuperscript{2}
-Joshua S.\ Weitz PhD\textsuperscript{6, 7}
-Bryan T.\ Grenfell PhD\textsuperscript{1,8,9}
-Jonathan Dushoff PhD\textsuperscript{2,3,4,*}
+Sang Woo Park\textsuperscript{1,*}
+Benjamin M.\ Bolker\textsuperscript{2,3,4}
+David Champredon\textsuperscript{5}
+David J.\,D.\ Earn\textsuperscript{3,4}
+Michael Li\textsuperscript{2}
+Joshua S.\ Weitz\textsuperscript{6, 7}
+Bryan T.\ Grenfell\textsuperscript{1,8,9}
+Jonathan Dushoff\textsuperscript{2,3,4,*}
 \\
@@
 
-\pagebreak
-
 \section*{Abstract}
@@
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
-We present a framework for comparing and combining different estimates of \Ro across a wide range of models.
+
+\subsection*{Aim}
+
+We aimed to develop a framework for comparing and combining different estimates of \Ro across a wide range of models.
 
@@
 
-We reviewed 7 studies that were published online between January 23--26, 2020.
+We reviewed 7 model-based analyses of the 2019-nCoV outbreak that were published online between January 23--26, 2020.
 We decompose their \Ro estimates into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
 We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.
-We assess early estimates of \Ro for the 2019-nCoV outbreak using the pooled estimates.
 
-\subsection*{Findings}
+\subsection*{Results}
 
 We find that most early estimates of \Ro rely on strong assumptions, especially about the generation-interval dispersion.
-In particular, estimates that rely on narrow generation-interval distributions are overly sensitive to estimates of the exponential growth rate.
+Estimates that rely on narrow generation-interval distributions are overly sensitive to estimates of the exponential growth rate.
 
-\subsection*{Interpretation}
+\subsection*{Conclusion}
 
@@
 
-\subsection*{Funding}
+\section*{Keywords}
 
-BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC). ML was supported by Canadian Institutes of Health Research (CIHR).
+Basic reproductive number, 2019-nCoV
 
-\pagebreak
+\section*{Funding}
 
-\section*{Research in context}
+BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC).
+ML was supported by Canadian Institutes of Health Research (CIHR).
+The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
 
-\subsection*{Evidence before this study}
+\section*{Declaration of interests}
 
-% 7 analyses of the 2019-nCoV outbreak were published online as pre-prints between January 23--26, 2020.
-% Their modeling approaches varied widely and estimates of uncertainties in \Ro ranged from 1.5 to 7.1.
+We declare no competing interests.
 
-This section should include a description of all the evidence that the authors considered before undertaking this study.
-Authors should briefly state: the sources (databases, journal or book reference lists, etc) searched; the criteria used to include or exclude studies (including the exact start and end dates of the search), which should not be limited to English language publications; the search terms used; the quality (risk of bias) of that evidence; and the pooled estimate derived from meta-analysis of the evidence, if appropriate
+\section*{Acknowledgements}
 
-\subsection*{Added value of this study}
+We thank Daihai He for providing helpful comments on the manuscript.
 
-Authors should describe here how their findings add value to the existing evidence
+\section*{Contribution}
 
-\subsection*{Implications of all the available evidence}
-
-Authors should state the implications for practice or policy and future research of their study combined with existing evidence.Research in context panels should not contain references; key studies mentioned here should be referenced in the main text.
+SWP and JD developed the statistical framework. 
+SWP reviewed the published literature.
+SWP performed the analysis. 
+SWP, BMB, and JD created the figures. 
+SWP and JD wrote the first draft.
+All authors contributed to the writing and approval of the final report.
 
@@
 
-As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
+As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints\cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.
@@
 \begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
- & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & Sources \\
+ & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
 \hline
-Study 1 & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
+Study 1 & 1.5--3.5 & 10 & 1 & Bedford \textit{et al.} \cite{bedfordncov} \\
 \hline
-Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & Imai \textit{et al.} \cite{imaincov} \\
 \hline
-Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
+Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & Liu \textit{et al.} \cite{liuncov} \\
 \hline
-Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
+Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & Read \textit{et al.} \cite{readncov} \\
 \hline
-Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
+Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & Riou and Althaus \cite{riouncov} \\
 \hline
-Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
+Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & Zhao \textit{et al.} \cite{zhaoncov} \\
 \hline
-Study 7 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
+Study 7 & 2.0--3.1 & 6--10 & 0 & Majumder and Mandl \cite{majumderncov} \\
 \hline
@@
 
-\section*{Contributors}
-
-SWP and JD developed the statistical framework. 
-SWP reviewed the published literature.
-SWP performed the analysis. 
-SWP, BMB, and JD created the figures. 
-SWP and JD wrote the first draft.
-All authors contributed to the writing and approval of the final report.
-
-\section*{Declaration of interests}
-
-We declare no competing interests.
-
-\section*{Acknowledgements}
-
-We thank Daihai He for providing helpful comments on the manuscript.
-
 \pagebreak
 
-\bibliography{ncov}
+\bibliography{ncov_abbr}
 

### 2020-02-07 ncov_euro.tex
euro

--- 2020-02-07 ncov_lancet_ID.tex
@@
 
-As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints\cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
+As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.
@@
 
-We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
+We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution.
+The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make:
 \begin{linenomath*}
@@
 Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
-half-Student-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
+half-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
 half-Cauchy(0,5)) can lead to poor mixing.

### 2020-02-07 ncov.tex
minor modifications and funding
Tiny changes

@@
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
-As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case in an otherwise susceptible population.
+As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro -- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
@@
 We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
-Our results emphasize the importance of propagating uncertainties in all components of \Ro, especially the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.
 
@@
 
+\section*{Funding}
+
+BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC). ML was supported by Canadian Institutes of Health Research (CIHR).
+
 \pagebreak

### 2020-02-07 ncov_euro.tex
pdf
oops

@@
 
+\pagebreak
+
 \section*{Abstract}
@@
 A novel coronavirus (2019-nCoV) has recently emerged as a global threat.
-As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case in an otherwise susceptible population.
+As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro -- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
@@
 
-Our results emphasize the importance of propagating uncertainties in all components of \Ro, especially the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+
+\pagebreak
 
@@
 
-Basic reproductive number, 2019-nCoV
+Basic reproductive number, 2019-nCoV, novel coronavirus, Bayesian multilevel model
 
@@
 
-As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
+As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov, riouncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
 Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.

### 2020-02-07 ncov_euro.tex
add data av
Same tiny changes ☺

--- 2020-02-07 ncov_euro.tex
@@
 
-Our results emphasize the importance of propagating uncertainties in all components of \Ro, especially the shape of the generation-interval distribution in efforts to estimate \Ro at the outset of an epidemic.
+Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.
+
+\pagebreak
 

### 2020-02-07 ncov_euro.tex
final?
not sure why this changed

@@
 
+\section*{Data availability}
+
+\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
+
 \pagebreak
@@
 95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
-\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
 

### 2020-02-26 ncov.tex
update

@@
 	\textbf\newline{
-		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (2019-nCoV) outbreak
+		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (SARS-CoV-2) outbreak
 	}
@@
 \section*{Abstract}
-A novel coronavirus (2019-nCoV) has recently emerged as a global threat. 
+A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat. 
 As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro -- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
-Here, we present a framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We then apply our framework to early estimates of \Ro for the 2019-nCoV outbreak.
+Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+We then apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak.
 Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.
@@
 
-Since December 2019, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world \citep{pneumonia}.
-Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of 2019-nCoV ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
-As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \citep{who02report}.
-Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
+Since December 2019, a novel coronavirus (SARS-CoV-2) has been spreading in China and other parts of the world \citep{pneumonia}.
+Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of SARS-CoV-2 ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
+As of February 26, 2020, the World Health Organization (WHO) has confirmed 81109 cases, including 2918 confirmed cases in 37 different countries, outside China \citep{who37report}.
+% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
 
@@
 
-We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
-We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
+We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+We use a Bayesian multilevel model to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
 the pooled estimates reflect the uncertainties present in modeling approaches and their underlying assumptions.
@@
 \begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
- & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
+ & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
 \hline
-Study 1 & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
+Study 1 & 1.5--3.5 & 10 & 1 & Bedford \textit{et al.} \cite{bedfordncov} \\
 \hline
-Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & Imai \textit{et al.} \cite{imaincov} \\
 \hline
-Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
+Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & Liu \textit{et al.} \cite{liuncov} \\
 \hline
-Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
+Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & Read \textit{et al.} \cite{readncov} \\
 \hline
-Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
+Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & Riou and Althaus \cite{riouncov} \\
 \hline
-Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
+Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & Zhao \textit{et al.} \cite{zhaoncov} \\
 \hline
-Study 7 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
+Study 7 & 2.0--3.1 & 6--10 & 0 & Majumder and Mandl \cite{majumderncov} \\
 \hline
@@
 
-We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23 and January 26, 2020 (Table 1).
-As most studies do not report their estimates of the exponential growth rate or the associated confidence intervals, we first recalculate the exponential growth rate that correspond to their model assumptions.
-We do so by modeling explicitly or implicitly reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
+We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
+Their modeling approaches vary widely: a branching process model \citep{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \citep{readncov}, an exponential growth model \citep{zhaoncov}, a Poisson offspring distribution model \citep{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \citep{majumderncov}.
+Four studies estimated \Ro by directly fitting their models to incidence data \citep{readncov, zhaoncov, liuncov, majumderncov}.
+The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0})
+Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+
+\subsection{Gamma approximation framework for linking $r$ and $\Ro$}
+
+Early in an outbreak, \Ro is difficult to estimate directly;
+instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
+Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
+number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
+\begin{linenomath*}
+\begin{equation}
+1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
+\label{eq:euler}
+\end{equation}
+\end{linenomath*}
+In other words, estimates of \Ro must
+depend on the assumptions about the
+exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+
+Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+Assuming that generation intervals follow a gamma distribution with the mean $\bar G$ and the squared coefficient of variation $\kappa$, we have
+\begin{linenomath*}
+\begin{equation}
+\Ro = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
+\label{eq:gamma}
+\end{equation}
+\end{linenomath*}
+This equation demonstrates that a generation-interval distribution that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$) will give a higher estimate of \Ro for the same value of $r$ \citep{wallinga2007generation}.
+
+\subsection{Statistical framework}
+
+As most studies do not report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond to their model assumptions.
+We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
 we used gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
@@
 
+We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution.
+The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make:
+\begin{linenomath*}
+\begin{equation}
+\begin{aligned}
+r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
+\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
+\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
+\end{aligned}
+\end{equation}
+\end{linenomath*}
+where $\mu_r$, $\mu_G$, $\mu_\kappa$ represent the pooled estimates, and $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations.
+We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
+Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
+We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
+As we add more models, the pooled estimates can become sharper even when the models no longer add more information.
+Thus, the pooled estimator should be interpreted with care.
+
+We use weakly informative priors on hyperparameters:
+\begin{linenomath*}
+\begin{equation}
+\begin{aligned}
+\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
+\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
+\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
+(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
+\end{aligned}
+\end{equation}
+\end{linenomath*}
+We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
+This phenomenon is a known problem \citep{gelman2006prior}.
+Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
+half-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
+half-Cauchy(0,5)) can lead to poor mixing.
+
+We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
+Posterior samples are thinned every 1000 steps.
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
+trace plots and marginal posterior distribution plots are presented in Appendix.
+95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
+
+\section{Results}
+
 \fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model.
@@
 replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
-This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties and test how underlying assumptions affect their estimates.
+This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
 We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but 7 cases.
-We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we use our pooled estimate of the generation-interval dispersion $\mu_\kappa$ to recalculate \Ro because they assume a narrow generation-interval distribution (compare \textbf{base} with \textbf{GI variation} in \fref{R0});
+We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we account for proper uncertainties in the generation-interval dispersion because they assume a narrow generation-interval distribution (compare ``base'' with ``GI variation'');
 when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 5 and Study 7 when we use our pooled estimate of the mean generation interval $\mu_G$ to recalculate \Ro (compare \textbf{base} with \textbf{GI mean} in \fref{R0}) because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
-We also estimate narrower confidence intervals for Study 1 when we replace their estimated $r$ or $\bar G$ with our pooled estimates for similar reasons.
+We estimate narrower confidence intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare ``base'' with ``GI mean'') because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
+Substituting the reported $r$ or $\bar G$ from Study 1 with our pooled estimates give narrower confidence intervals for similar reasons.
 
-Consistent with our previous observations (\fref{eff}),
-we find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
+We find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
 For example, recalculating \Ro for Study 7 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--8.6), which is much wider than the uncertainty range they reported (2.0--3.1).
@@
 Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
-This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts (e.g., \cite{ma2014estimating}).
-Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals (e.g., \cite{king2015avoidable, taylor2016stochasticity}).
+This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \citep{ma2014estimating}.
+Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \citep{king2015avoidable,taylor2016stochasticity}.
 
@@
 Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
-For emerging pathogens such as 2019-nCoV, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
+For emerging pathogens such as SARS-CoV-2, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
 for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
-Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{harris1990reporting, tsuzuki2017dynamics, tariq2019assessing}.
+Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{tariq2019assessing}.
 Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
@@
 
-\section{Methods}
-
-\subsection{Gamma approximation framework for linking $r$ and $\Ro$}
-
-Early in an outbreak, \Ro is difficult to estimate directly;
-instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{mills2004transmissibility, nishiura2009transmission, ma2014estimating}.
-Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
-number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
-\begin{linenomath*}
-\begin{equation}
-1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
-\label{eq:euler}
-\end{equation}
-\end{linenomath*}
-In other words, estimates of \Ro must
-depend on the assumptions about the
-exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-
-Here, we use the gamma approximation framework \citep{mcbryde2009early, nishiura2009transmission, roberts2011early, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
-Assuming that generation intervals follow a gamma distribution
-with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
-we have
-\begin{linenomath*}
-\begin{equation}
-\Ro = \left(1 + \kappa r \bar{G}\right)^{1/\kappa}.
-\label{eq:gamma}
-\end{equation}
-\end{linenomath*}
-This equation demonstrates that a generation-interval distribution
-that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
-will give a higher estimate of \Ro for the same value of $r$.
-
-\subsection{Description of the studies}
-
-We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
-They vary widely in their statistical methods and in the amount of data used to infer \Ro.
-\cite{bedfordncov}, \cite{imaincov}, and \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
-\cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
-\cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
-\cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
-Finally, \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
-Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
-
-\subsection{Statistical framework}
-
-For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculating the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
-\begin{linenomath*}
-\begin{equation}
-r_i = \frac{\Rx{0i}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
-\end{equation}
-\end{linenomath*}
-This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
-uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.
-
-We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
-\begin{linenomath*}
-\begin{equation}
-\begin{aligned}
-r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
-\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
-\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
-\end{aligned}
-\end{equation}
-\end{linenomath*}
-where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
-We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step;
-this approach is analogous to Bayesian methods for analyzing phylogenetic data, which often rely on drawing random samples of phylogenetic trees from a discrete set to account for phylogenetic uncertainty \citep{pagel2004bayesian,bedford2014integrating}.
-Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
-We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-As we add more models, we expect the pooled estimates to become sharper even when the models no longer add more information.
-Thus, the pooled estimator should be interpreted with care.
-
-We use weakly informative priors on hyperparameters representing pooled means and non-informative priors on hyperparameters representing between-study standard deviations:
-\begin{linenomath*}
-\begin{equation}
-\begin{aligned}
-\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
-\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
-\mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
-(\sigma_r, \sigma_G, \sigma_\kappa) &\sim \textrm{half-normal}(0, 10).
-\end{aligned}
-\end{equation}
-\end{linenomath*}
-We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
-We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
-This phenomenon is a known problem \citep{gelman2006prior}.
-Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
-half-Student-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
-half-Cauchy(0,5)) can lead to poor mixing.
-
-We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
-Posterior samples are thinned every 1000 steps.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
-trace plots and marginal posterior distribution plots are presented in Appendix.
-95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
-\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
-
-\section*{Acknowledgements}
-
-We thank Daihai He for providing helpful comments on the manuscript.
-
-\section*{Funding}
-
-BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC). ML was supported by Canadian Institutes of Health Research (CIHR).
-
 \pagebreak
 
-\bibliography{ncov}
+\bibliography{ncov_abbr}
 

### 2020-02-26 ncov.tex
funding and stuff

@@
 
+\section*{Funding}
+
+BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC).
+ML was supported by Canadian Institutes of Health Research (CIHR).
+The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
+\section*{Declaration of interests}
+
+We declare no competing interests.
+
+\section*{Acknowledgements}
+
+We thank Daihai He for providing helpful comments on the manuscript.
+
+\section*{Contribution}
+
+SWP and JD developed the statistical framework. 
+SWP reviewed the published literature.
+SWP performed the analysis. 
+SWP, BMB, and JD created the figures. 
+SWP and JD wrote the first draft.
+All authors contributed to the writing and approval of the final report.
+
+\section*{Data availability}
+
+\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
+
+\pagebreak
+
 \bibliography{ncov_abbr}

### 2020-02-26 ncov.tex
ok

@@
 We then apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak.
+We show that many early \Ro estimates are overly confident.
 Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.
@@
 Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of SARS-CoV-2 ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
-As of February 26, 2020, the World Health Organization (WHO) has confirmed 81109 cases, including 2918 confirmed cases in 37 different countries, outside China \citep{who37report}.
+As of February 26, 2020, the World Health Organization (WHO) has confirmed 81,109 cases, including 2,918 confirmed cases in 37 different countries, outside China \citep{who37report}.
 % Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
@@
 Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
-Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals:
-our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro overly sensitive to the estimates of the exponential growth rate $r$.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals.
+Our analysis of individual studies shows that many early estimates of \Ro rely on overly strong assumptions.
 

### 2020-02-28 ncov.tex
..
update

@@
 
+\section*{Keywords}
+
+SARS-CoV-2, COVID-19, novel coronavirus, basic reproductive number, generation interval, Bayesian multilevel model
+
 \pagebreak
@@
 Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of SARS-CoV-2 ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
-As of February 26, 2020, the World Health Organization (WHO) has confirmed 81,109 cases, including 2,918 confirmed cases in 37 different countries, outside China \citep{who37report}.
+As of February 27, 2020, the World Health Organization (WHO) has confirmed 82,294 cases of the coronavirus disease (COVID-19), including 3,664 confirmed cases in 46 different countries, outside China \citep{who38report}.
 % Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
@@
 Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals.
-Our analysis of individual studies shows that many early estimates of \Ro rely on overly strong assumptions.
+Our analysis of individual studies shows that many early estimates of \Ro rely on strong assumptions.
 
@@
 
-\section*{Declaration of interests}
+\section*{Competing interests}
 

### 2020-04-24 ncov.tex
minor changes
editing

@@
 
+\pagebreak
+
 \section*{Abstract}
 A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat. 
-As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro -- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
+As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro\ --- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
@@
 SARS-CoV-2, COVID-19, novel coronavirus, basic reproductive number, generation interval, Bayesian multilevel model
-
-\pagebreak
 
@@
 \hline
-Study 1 & 1.5--3.5 & 10 & 1 & Bedford \textit{et al.} \cite{bedfordncov} \\
+Study 1 & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
 \hline
-Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & Imai \textit{et al.} \cite{imaincov} \\
+Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
 \hline
-Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & Liu \textit{et al.} \cite{liuncov} \\
+Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
 \hline
-Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & Read \textit{et al.} \cite{readncov} \\
+Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
 \hline
-Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & Riou and Althaus \cite{riouncov} \\
+Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
 \hline
-Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & Zhao \textit{et al.} \cite{zhaoncov} \\
+Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
 \hline
-Study 7 & 2.0--3.1 & 6--10 & 0 & Majumder and Mandl \cite{majumderncov} \\
+Study 7 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
 \hline
@@
 Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
-Assuming that generation intervals follow a gamma distribution with the mean $\bar G$ and the squared coefficient of variation $\kappa$, we have
+Assuming that generation intervals follow a gamma distribution with mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation of a gamma distribution, we have
 \begin{linenomath*}
@@
 As most studies do not report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond to their model assumptions.
-We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
+We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions;
 we used gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
@@
 trace plots and marginal posterior distribution plots are presented in Appendix.
-95\% confidence intervals are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
+95\% credible intervals (CI) are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
 
@@
 In this case, propagating error from the mean generation interval has negligible effect compared to propagating the uncertainty in $r$.
-Uncertainty in the generation-interval dispersion also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare ``growth rate + GI mean'' with ``all'').
-For example, reducing the dispersion parameter $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
+Uncertainty in the generation-interval dispersion $\kappa$ also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare ``growth rate + GI mean'' with ``all'').
+For example, reducing $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
 
@@
 \textbf{Sensitivity of the reported \Ro estimates with respect to our pooled estimates of the underlying parameters.}
-We replace the reported parameter values (growth rate $r$, GI mean $\bar G$, and GI variation $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro (\textbf{growth rate}, \textbf{GI mean}, and \textbf{GI variation}).
+We replace the reported parameter values (growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro (\textbf{growth rate}, \textbf{GI mean}, and \textbf{GI variation}).
 The pooled estimate of \Ro is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$ (\textbf{all});

### 2020-04-24 compare_assumption.R

@@ -85,7 +85,7 @@ g3 <- ggplot(kappadata2) +
              size=c(2, 2, 2, 2, 2, 2, 2, 4)) +
   geom_segment(aes(lwr, anon, xend=upr, yend=anon), col=c(1, 1, 1, 1, 1, 1, 1, 2),
                lwd=c(0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 1.2)) +
-  scale_x_continuous("Squared coefficient of variation",
+  scale_x_continuous("Generation-interval dispersion",
                      breaks=c(0, 0.25, 0.5, 0.75, 1)) +
   theme(
     axis.title.y = element_blank()

### 2020-04-24 ncov.tex
some minor changes

@@
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. 
-We also use our framework to tease apart which assumptions of these different models led to their different estimates and confidence intervals.
+We also use our framework to tease apart which assumptions of these different models led to their different estimates and credible intervals.
 Despite the availability of more recent and/or updated estimates of \Ro, we restrict ourselves to the estimates above in order to focus on the resolution of uncertainty in the earliest stages of an epidemic.
@@
 Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
-$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
+$^\ast$We treat these intervals as a 95\% credible interval in our analysis.
 $^\dagger$We assume $\kappa = 0.5$ in our analysis.
@@
 We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions;
-we used gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
+we used gamma distributions to model values reported with confidence/credible intervals and uniform distributions to model values reported with ranges.
 For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
@@
 \begin{aligned}
-\mu_r &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week}^{-1},\,\mathrm{shape}=2)\\
-\mu_G &\sim \gammdist(\mathrm{mean}=1\,\mathrm{week},\,\mathrm{shape}=2)\\
+\mu_r &\sim \gammdist(\mathrm{mean}=1/7\,\mathrm{days}^{-1},\,\mathrm{shape}=2)\\
+\mu_G &\sim \gammdist(\mathrm{mean}=7\,\mathrm{days},\,\mathrm{shape}=2)\\
 \mu_\kappa &\sim \gammdist(\mathrm{mean}=0.5,\,\mathrm{shape}=2)\\
@@
 many models rely on stronger assumptions that ignore these uncertainties.
-Surprisingly, no studies take into account how the variation in generation intervals affects their estimates of \Ro:
+Surprisingly, no studies take into account how generation-interval dispersion affects their estimates of \Ro:
 all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
@@
 \textbf{Comparisons of the reported parameter values with our pooled estimates.}
-We inferred point estimates (black), uniform distributions (orange) or confidence intervals (purple) for each parameter from each study, and combined them into pooled estimates (red; see text).
+We inferred point estimates (black), uniform distributions (orange) or confidence/credible intervals (purple) for each parameter from each study, and combined them into pooled estimates (red; see text).
 Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval dispersion.
@@
 As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important (\fref{eff}B).
-Propagating error only from the growth rate gives very narrow confidence intervals in this case. 
-Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow confidence intervals.
+Propagating error only from the growth rate gives very narrow credible intervals in this case. 
+Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow credible intervals.
 We expect this hypothetical example to better reflect more recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
@@
 \textbf{all}: \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-Vertical lines represent the 95\% confidence intervals.
+Vertical lines represent the 95\% credible intervals.
 }
@@
 This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
-We find that incorporating uncertainties one at a time increases the width of the confidence intervals in all but 7 cases.
-We estimate narrower confidence intervals for Study 3, Study 6, and Study 7 when we account for proper uncertainties in the generation-interval dispersion because they assume a narrow generation-interval distribution (compare ``base'' with ``GI variation'');
-when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower confidence intervals.
-We estimate narrower confidence intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare ``base'' with ``GI mean'') because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
-Substituting the reported $r$ or $\bar G$ from Study 1 with our pooled estimates give narrower confidence intervals for similar reasons.
+We find that incorporating uncertainties one at a time increases the width of the credible intervals in all but 7 cases.
+We estimate narrower credible intervals for Study 3, Study 6, and Study 7 when we account for proper uncertainties in the generation-interval dispersion because they assume a narrow generation-interval distribution (compare ``base'' with ``generation-interval dispersion'');
+when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower credible intervals.
+We estimate narrower credible intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare ``base'' with ``mean generation interval'') because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
+Substituting the reported $r$ or $\bar G$ from Study 1 with our pooled estimates give narrower credible intervals for similar reasons.
 
@@
 One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive and thus tend to have particularly narrow confidence intervals.
+This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive and thus tend to have particularly narrow credible intervals.
 
@@
 \textbf{Sensitivity of the reported \Ro estimates with respect to our pooled estimates of the underlying parameters.}
-We replace the reported parameter values (growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro (\textbf{growth rate}, \textbf{GI mean}, and \textbf{GI variation}).
+We replace the reported parameter values (growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro (\textbf{growth rate}, \textbf{mean generation interval}, and \textbf{generation-interval dispersion}).
 The pooled estimate of \Ro is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$ (\textbf{all});
 this corresponds to replacing all reported parameter values with our pooled estimates, which gives identical results across all studies.
-Horizontal dashed lines represent the 95\% confidence intervals of our pooled estimate of \Ro.
-The reported \Ro estimates (\textbf{base}) have been adjusted to show the approximate 95\% confidence interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
+Horizontal dashed lines represent the 95\% credible intervals of our pooled estimate of \Ro.
+The reported \Ro estimates (\textbf{base}) have been adjusted to show the approximate 95\% credible interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
 }
@@
 Our estimated \Ro from the pooled distribution has a median of 2.9 (95\% CI: 2.1--4.5).
-While the point estimate of \Ro is similar to other reported values from this date range, the confidence intervals are wider than all but one study.
+While the point estimate of \Ro is similar to other reported values from this date range, the credible intervals are wider than all but one study.
 This result does not imply that assumptions based on the pooled estimate are too weak;
-we believe that this confidence interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+we believe that this credible interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
 In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
@@
 Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{tariq2019assessing}.
-Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
+Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated credible intervals.
 Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
@@
 However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
-Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: confidence intervals must combine different sources of uncertainty. 
+Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: credible intervals must combine different sources of uncertainty. 
 In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate (and other epidemiological parameters) will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.

### 2020-04-24 compare_R0.R

@@ -20,7 +20,7 @@ R0all <- Rdata_adj %>%
   ) %>%
   mutate(
     type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
-                labels=c("base", "growth rate", "GI mean", "GI variation", "all"))
+                labels=c("base", "growth rate", "mean generation interval", "generation-interval dispersion", "all"))
   ) %>%
   merge(study_number) %>%
   mutate(

### 2020-04-25 ncov.tex
lots of edits..

@@
 % \linenumbers
+
+\usepackage{pdflscape}
+
+\usepackage{afterpage}
 
@@
 A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat. 
-As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro\ --- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
+As the epidemic progresses, disease modelers continue to focus on estimating the basic reproductive number \Ro\ --- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
@@
 
-Since December 2019, a novel coronavirus (SARS-CoV-2) has been spreading in China and other parts of the world \citep{pneumonia}.
-Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of SARS-CoV-2 ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
-As of February 27, 2020, the World Health Organization (WHO) has confirmed 82,294 cases of the coronavirus disease (COVID-19), including 3,664 confirmed cases in 46 different countries, outside China \citep{who38report}.
-% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
+Since December 2019, a novel coronavirus (SARS-CoV-2) has been spreading in many parts of the world \citep{pneumonia}.
+Although the virus is likely to have originated from animal hosts \citep{andersen2020proximal}, the ability of SARS-CoV-2 to directly transmit between humans, particularly without symptoms, has posed a greater threat for its spread \citep{he2020temporal}.
+As of April 23, 2020, more than 2 million cases of the coronavirus disease 2019 (COVID-19) have been confirmed internationally \citep{who94}.
 
-As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
-Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+As SARS-CoV-2 began to spread in other parts of China, outside Hubei province, as well as in other countries, many analyses of the outbreak were published as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}).
+These analyses particularly focused on estimating the basic reproductive number \Ro\ --- the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition} --- in order to assess the pandemic potential of SARS-CoV-2.
+Such rapid dissemination of the early analyses played an important role in shaping the response to the outbreak \citep{majumder2020early}.
+
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.
+However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) varied significantly even though most analyses relied on reports of confirmed cases from Hubei, China.
+Understanding the drivers of the differences between \Ro estimates is necessary as \Ro provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+It can be difficult to compare a disparate set of estimates of \Ro when the estimation methods and their underlying assumptions vary widely.
 
-Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
-therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the generation-interval distribution and the exponential growth rate.
+Here, we show that a wide range of approaches to estimating \Ro during the exponential growth period can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, characterizes the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
+therefore, estimates of \Ro from different models directly depend on their assumptions about the generation-interval distribution and the exponential growth rate.
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
+Ignoring these uncertainties inevitably lead to strong conclusions.
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
 
-We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+We compare seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 We use a Bayesian multilevel model to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
-the pooled estimates reflect the uncertainties present in modeling approaches and their underlying assumptions.
+the pooled estimates represent average of the estimated and/or assumed parameter values and reflect the uncertainties present in modeling approaches and their underlying assumptions.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
@@
 
-\begin{table}[t]
+\afterpage{%
+\clearpage
+    \begin{landscape}
+\begin{table}[!th]
 \begin{center}
 \scriptsize
-\begin{tabular}{l|p{3.5cm}|p{2.5cm}|p{2.7cm}|l}
- & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
+\begin{tabular}{l|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{2.4cm}|p{2.7cm}|l}
+ & Model & Data & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
 \hline
-Study 1 & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
+Study 1 & Deterministic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
 \hline
-Study 2 & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 2 & Stochastic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020)  & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
 \hline
-Study 3 & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
+Study 3 & Poisson offspring distribution model & Confirmed cases from China and other countries\newline (Dec 29, 2019--Jan 23, 2020) & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
 \hline
-Study 4 & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
+Study 4 & Metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model & Reported cases from China\newline (Jan 1--21, 2020) & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
 \hline
-Study 5 & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
+Study 5 & Stochastic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
 \hline
-Study 6 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
+Study 6 & Exponential growth model & Reported cases from China\newline (Jan 1--21, 2020) & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
 \hline
-Study 7 & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
+Study 7 & Incidence Decay and Exponential Adjustment (IDEA) model & Reported cases from Wuhan\newline (Dec 8, 2019--Jan 26, 2020) & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
 \hline
@@
 \caption{
-\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
-Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
+\textbf{Summary of the models used and reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
+}
+Model details, estimates of \Ro, and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
+Generation-interval dispersion represent the squared coefficient of variations in generation intervals.
 $^\ast$We treat these intervals as a 95\% credible interval in our analysis.
@@
 $^\ddagger$The authors presented \Ro estimates under different assumptions regarding the reporting rate; we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
+\end{table}
+\end{landscape}
+\clearpage
 }
-\end{table}
 
-We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
-Their modeling approaches vary widely: a branching process model \citep{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \citep{readncov}, an exponential growth model \citep{zhaoncov}, a Poisson offspring distribution model \citep{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \citep{majumderncov}.
-Four studies estimated \Ro by directly fitting their models to incidence data \citep{readncov, zhaoncov, liuncov, majumderncov}.
-The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0})
-Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+We gathered information on estimates of \Ro and their model assumptions from 7 articles that were published online between January 23--26, 2020.
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on \url{nextstrain.org} \citep{bedfordncov}.
+Model details are summarized in Table 1.
 
-\subsection{Gamma approximation framework for linking $r$ and $\Ro$}
+\subsection{Comparing model assumptions}
 
-Early in an outbreak, \Ro is difficult to estimate directly;
-instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
-Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
-number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
+Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic grows exponentially in the beginning.
+While the IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth \citep{fisman2013idea}, Study 7 estimates $d=0$ across all parameters they consider, in which case the IDEA model becomes equivalent to the exponential growth model.
+When the epidemic is growing exponentially, the estimated basic reproduction number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{wallinga2007generation}:
 \begin{linenomath*}
@@
 \end{linenomath*}
-In other words, estimates of \Ro must
-depend on the assumptions about the
-exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+In other words, all model assumptions come down to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence intervals of the model parameters will be necessarily narrow.
+Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and shape of the generation-interval distribution.
+
+Since only one out of seven studies report the exponential growth rate, we summarize the 
+In Table 1, we summarized the reported
+
+Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+
+\subsection{Gamma approximation framework for linking $r$ and $\Ro$}
 

### 2020-04-26 ncov.tex
too much

@@
 
-%% \newcommand{\rev}[1]{\comment{red}{REV}{#1}}
-\newcommand{\rev}[1]{}
+\newcommand{\rev}[1]{\comment{red}{REV}{#1}}
 
@@
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) varied significantly even though most analyses relied on reports of confirmed cases from Hubei, China.
-Understanding the drivers of the differences between \Ro estimates is necessary as \Ro provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary significantly even though most analyses relied on reports of confirmed cases from China, particularly from the Hubei province.
+Understanding the differences between \Ro estimates is necessary as \Ro provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 It can be difficult to compare a disparate set of estimates of \Ro when the estimation methods and their underlying assumptions vary widely.
@@
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
-Ignoring these uncertainties inevitably lead to strong conclusions.
+Ignoring these uncertainties inevitably leads to strong conclusions.
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
 
-We compare seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
-We use a Bayesian multilevel model to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
-the pooled estimates represent average of the estimated and/or assumed parameter values and reflect the uncertainties present in modeling approaches and their underlying assumptions.
+We present a statistical framework for averaging across estimated basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+Previous studies have directly calculated the average of reported \Ro values but such methods undermine differences in underlying model assumptions and statistical methods \citep{majumder2020early, liu2020reproductive}.
+Instead, we decompose estimated \Ro into three key quantities ($r$, $\bar G$, and $\kappa$) and calculate the average (pooled estimates) of these key parameters.
+Calculating \Ro based on the pooled estimates allow us to appropriately average across the uncertainties present in modeling approaches and their underlying assumptions.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
-We also use our framework to tease apart which assumptions of these different models led to their different estimates and credible intervals.
+We also use our framework to tease apart which assumptions of these different models led to their different estimates and the associated degrees of uncertainty.
 Despite the availability of more recent and/or updated estimates of \Ro, we restrict ourselves to the estimates above in order to focus on the resolution of uncertainty in the earliest stages of an epidemic.
@@
 \hline
-Study 1 & Deterministic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
+Study 1 & Deterministic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
 \hline
-Study 2 & Stochastic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020)  & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 2 & Stochastic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020)  & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
 \hline
@@
 \hline
-Study 5 & Stochastic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
+Study 5 & Stochastic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
 \hline
@@
 \hline
-Study 7 & Incidence Decay and Exponential Adjustment (IDEA) model & Reported cases from Wuhan\newline (Dec 8, 2019--Jan 26, 2020) & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
+Study 7 & Incidence Decay and Exponential Adjustment (IDEA) model & Reported cases from Wuhan City, China\newline (Dec 8, 2019--Jan 26, 2020) & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
 \hline
@@
 Generation-interval dispersion represent the squared coefficient of variations in generation intervals.
-$^\ast$We treat these intervals as a 95\% credible interval in our analysis.
+$^\ast$These intervals reflect \Ro values for best and worst scenarios. We treat these intervals as a 90\% credible interval in our analysis.
 $^\dagger$We assume $\kappa = 0.5$ in our analysis.
@@
 Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic grows exponentially in the beginning.
-While the IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth \citep{fisman2013idea}, Study 7 estimates $d=0$ across all parameters they consider, in which case the IDEA model becomes equivalent to the exponential growth model.
-When the epidemic is growing exponentially, the estimated basic reproduction number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{wallinga2007generation}:
+While the IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth when $d \neq 0$ \citep{fisman2013idea}, Study 7 estimates $d=0$ across all parameters they consider.
+When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{wallinga2007generation}:
 \begin{linenomath*}
 \begin{equation}
-1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
+\frac{1}{\Ro} = \int \exp(-r\tau) g(\tau) \, d\tau.
 \label{eq:euler}
@@
 In other words, all model assumptions come down to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence intervals of the model parameters will be necessarily narrow.
-Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and shape of the generation-interval distribution.
+For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence intervals associated with the exponential growth rates or parameters of the generation-interval distributions will be necessarily narrow.
+Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproduction number.
 
-Since only one out of seven studies report the exponential growth rate, we summarize the 
-In Table 1, we summarized the reported
+As most studies do not report their estimates of the exponential growth rate, we summarize model outcomes using reported (either estimated or assumed) values of the basic reproductive number \Ro, mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation (Table 1);
+we re-estimate the corresponding exponential growth rates later.
+Study 2 only reports their assumptions about the mean generation interval; for brevity, we assume $\kappa = 0.5$ in our analysis.
+Study 6 presents \Ro estimates under 12 different scenarios regarding reporting rates (0, 0.5, 1 or 2 fold increase in reporting rate) and the shapes of the generation-interval distributions (MERS-like, SARS-like, and average);
+we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
+While estimates of \Ro and associated confidence intervals for Study 6 in Table 1 represents estimates of \Ro based on $\bar G = 8\,\mathrm{days}$, we account for the uncertainty they consider for $\bar G$ in our formal analysis.
+\rev{Page 7, Line 20. Figure 1 - Caption: Did the authors test the sensitivity of the value of kappa of
+Study 2 (e.g., 0 or 1) on the pooled estimate?}
 
-Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+While most studies report confidence or credible intervals to quantify uncertainties associated with their estimates, some rely on different measures.
+In particular, study 2 reports a range of \Ro for the worst and best case scenarios, which correspond to the values of \Ro such that 95\% and 5\% of the simulated total number of cases by Janaury 18, 2020 is greater than or equal to 4000, respectively;
+for brevity, we treat these intervals as a 90\% credible interval in our analysis.
+Uncertainty ranges reported by Study 1 and Study 7 are assumed to be uniform ranges.
+
+%While some studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}, we strictly focus on estimates that were published between January 23--26, 2020.
+%The main purpose of this study is to illustrate that estimates of \Ro from different models can be understood by comparing their underlying assumptions about the exponential growth rates and generation-interval distributions, rather than to provide more precise or accurate estimate of \Ro.
 
@@
 
-Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+Here, we use the gamma approximation framework to the generation-interval distribution \citep{nishiura2009transmission, mcbryde2009early, roberts2011early, trichereau2012estimation, nishiura2015theoretical, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.
 Assuming that generation intervals follow a gamma distribution with mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation of a gamma distribution, we have
@@
 
-As most studies do not report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond to their model assumptions.
-We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions;
-we used gamma distributions to model values reported with confidence/credible intervals and uniform distributions to model values reported with ranges.
+\rev{The authors consider that all the 7 studies have the same weight in the estimation of basic
+reproduction weight even though their estimations and precision vary a lot. The authors might
+explain why they make this hypothesis and why they did not explore a weighted estimation
+based on the confidence of the estimation of R0.}
+
+\rev{In each chain, the authors use 500,000 sampling steps and remove every 1000th step.
+Therefore, the posterior distributions of each parameter consist of 2,000 samples (500 samples
+per chain with 4 independent chains). Is it sufficient for the effective sample size? Is it too much
+for the burn-in steps of 500,000 and the thinning steps of 1,000?}
+
+As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that correspond to their model assumptions:
+Since the estimates of the basic reproductive number \Ro entirely depends on the exponential growth rate and the shape of generation-interval distributions, we can calulate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions;
+we used gamma distributions to model values reported with confidence/credible intervals and uniform distributions to model values reported with ranges (Table 2).
 For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
 we model this estimate as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
+
 For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculate the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
@@
 Thus, the pooled estimator should be interpreted with care.
+\rev{Authors state that “We note that this approach does not account for non- independence between the parameter estimates made by different modelers.” Authors should discuss this point more in depth.
+This seems to me an important limitation.
+With this respect I wonder what data the studies focus on.
+Are all analyzing the same epidemic (in China) considering same or similar data?
+Again, more details on these studies (in this case the location, period of the epidemic, and data source) should be provided.}
 
+\rev{Page 6, Line 29. Are the priors for $\mu_r$, $\mu_G$, and $\mu_\kappa$ weakly informative? Did the
+authors check the sensibility with new priors?}
 We use weakly informative priors on hyperparameters:
@@
 half-Cauchy(0,5)) can lead to poor mixing.
+\rev{Page 6 Lines 44-47. Please clarify the sentence: “Alternative choices of prior … can lead to
+poor mixing”.}
 
@@
 \section{Results}
+
+\rev{In Figures 2 and 3, the authors replace the reported parameters (r, $\bar G$, and kappa) with the
+pooled estimates mu ($\mu_r$, $\mu_G$, and $\mu_\kappa$). The authors do not give information about
+the other sources of uncertainty linked with the variation of parameter sigma ($\sigma_r$, $\sigma_G$,
+$\sigma_\kappa$).}
+
+\rev{In figure 1, the posterior density distribution for the 3 parameters [exponential growth rate (r),
+Mean generation interval ($\bar G$), Generation interval dispersion (kappa))] must replace the
+segment representing the median and the CI. In figure 1, the legend must be self-sufficient. ‘See
+text’ it is not possible. A short explanation is necessary.}
+
+\rev{In the first paragraph of the Results page 7 the value of the median and the CI of the 3
+parameters must be presented.}
 
@@
 
+\rev{The estimation of R0 median and its credible interval “2.9 (95\% CI: 2.1-4.5)” must be presented
+in page 7 with the figure 2 in the second paragraph of Results and not at the end of the Results
+part (page 9 line 26).}
+
+\rev{The uncertainty related to kappa is not described in Figure 2. Even the authors mentioned in
+Page 7, Lines 44-50, the results should also be visualized in Figure 2 along with other
+parameters.}
+
+\rev{Page 7, Line 38. The authors might give some information about the incubation period of
+COVID-19. It could explain why using the data of 26 January is still acceptable even if it was the
+3rd day of the city lockdown.}
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro.
@@
 We expect this hypothetical example to better reflect more recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
+
+\rev{Please clearly explain the formula in the caption of Figure 2 (Page 8, line 33) in the Methods
+section.}
 
@@
 
+\rev{Page 8, Line 50. Please clarify the sentence: ”We find that incorporating uncertainties one at a
+time increases the width of the confidence intervals in all but 7 cases”. It is not clear particularly
+at “in all but 7 cases”.}
 We also compare the estimates of \Ro across different studies by 
@@
 
+\rev{In the Results section, the authors occasionally discuss the results, i.e., Page 8, line 50 to Page
+9, line 23. The opinions should be in the Discussion section.
+In the Discussion section, the authors should discuss the quality of the reported estimates from
+7 studies used in this subject. Some of them do not provide detailed methods and assumptions.
+Also, most of them are preprint articles. It could be another important source of uncertainty on
+R0.}
+
+\rev{Although the authors are in the case of the COVID-19 epidemic in Hubei before any
+confinement, the authors do not mention in the introduction or in the discussion the other
+sources of uncertainty linked to the data (such as the variations in diagnostic results related to
+the day of the week, a saturation of diagnostic test capacity, transparency of data,
+representativeness of samples and improvement of detection capacity as time goes by.}
+
+\rev{Page 4, Table 1. Please check Study 4, the article has been revised since 27 January 2020,
+and their estimates also changed.}
+
+\rev{Page 6, Table 2. Please recheck the parameters of Study 5, which is inconsistent with the 90%
+confidence interval in Table 1. The 90\% confidence interval for a Gamma distribution with mean
+2.2 and shape 12 is 1.27 to 3.34, while the interval in Table 1 is 1.4 to 3.8.}
+
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
@@
 Naive fits to cumulative incidence data should therefore be avoided.
+\rev{
+At the beginning of the Discussion section authors list a series of important issues and possible sources of errors that often limit the utility of early outbreak analyses (e.g.
+fitting the cumulative curve of cases, assuming a fixed generation interval when in fact it varies during the outbreak, changing reporting rate, etc.).
+Compared with these issues the main focus of the paper (not properly accounting for the uncertainty in the generation time and growth rate parameters) appears a small caveat.
+Maybe author could give a sense of the importance of the element they are focusing on with the respect to the other issues?
+}
 

### 2020-04-26 ncov.tex
edits

@@
 
-We present a statistical framework for averaging across estimated basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+We present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 Previous studies have directly calculated the average of reported \Ro values but such methods undermine differences in underlying model assumptions and statistical methods \citep{majumder2020early, liu2020reproductive}.
@@
 For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence intervals associated with the exponential growth rates or parameters of the generation-interval distributions will be necessarily narrow.
-Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproduction number.
+Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.
 
@@
 
-\subsection{Statistical framework}
+\subsection{Re-estimation of the exponential growth rate}
 
-\rev{The authors consider that all the 7 studies have the same weight in the estimation of basic
-reproduction weight even though their estimations and precision vary a lot. The authors might
-explain why they make this hypothesis and why they did not explore a weighted estimation
-based on the confidence of the estimation of R0.}
-
-\rev{In each chain, the authors use 500,000 sampling steps and remove every 1000th step.
-Therefore, the posterior distributions of each parameter consist of 2,000 samples (500 samples
-per chain with 4 independent chains). Is it sufficient for the effective sample size? Is it too much
-for the burn-in steps of 500,000 and the thinning steps of 1,000?}
-
-As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that correspond to their model assumptions:
+As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that correspond to their model assumptions.
 Since the estimates of the basic reproductive number \Ro entirely depends on the exponential growth rate and the shape of generation-interval distributions, we can calulate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions;
-we used gamma distributions to model values reported with confidence/credible intervals and uniform distributions to model values reported with ranges (Table 2).
+First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
+We used gamma distributions to model values reported with confidence/credible intervals and uniform distributions to model values reported with ranges;
+when confidence/credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimate value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence/credible limits is equal to the reported confidence/credible level.
+This probability does not need to be equi-tailed.
 For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
@@
 
-For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculate the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
+For each study $i$, we construct a family of parameter sets by drawing $10^5$ random samples from the probability distributions (Table 2) that represent the estimates of $(\Ro)_{i,m}$ and the assumed values of $\bar G_{i,m}$ and $\kappa_{i,m}$ and calculate the exponential growth rate $r_{i,m}$ via the inverse of \eref{gamma}:
 \begin{linenomath*}
 \begin{equation}
-r_i = \frac{\Rx{0i}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
+r_{i,m} = \frac{\left[(\Ro)_{i,m}\right]^{\kappa_{i,m}} - 1}{\kappa_{i,m} \bar{G}_{i,m}},
 \end{equation}
 \end{linenomath*}
-This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
-uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.
+where $m=1,\dots,10^5$.
+This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.
+Uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.
+Then, the set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ defines an empirical parameter distribution for each study without having to re-fit each model.
+
+For most cases, we model $(\Ro)_i$, $\bar G_i$, and $\kappa_i$ using independent distributions (as specified in Table 2) to calculate $r_i$.
+In practice, this assumption does not affect our analysis because we are re-calculating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
+in other words, we still obtain the same estimates and the associated uncertainty of $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.
+For study 6, we fix $\bar G=8\,\textrm{days}$ and use the gamma distribution (Table 2) that corresponds to $\mathcal R_0 = 5.47$ (95\% CI: 4.16--7.10) during the re-estimation step for $r$ to remain consistent with the original study that assumed $\bar G=8\,\textrm{days}$ to obtain this particular estimate;
+the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like) days.
+We account for uncertainties in $\bar G$ for Study 6 (Table 1) in all other steps in order to properly incorporate parameter uncertainties in the estimate of \Ro.
+Study 7 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$ in the original article.
 
@@
 We still account for this uncertainty in our pooled estimates ($\mu_G$).
-$^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+$^\dagger$Study 7 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
@@
 \end{table}
+
+\subsection{Pooled-estimates}
+
+\rev{The authors consider that all the 7 studies have the same weight in the estimation of basic
+reproduction weight even though their estimations and precision vary a lot. The authors might
+explain why they make this hypothesis and why they did not explore a weighted estimation
+based on the confidence of the estimation of R0.}
+
+\rev{In each chain, the authors use 500,000 sampling steps and remove every 1000th step.
+Therefore, the posterior distributions of each parameter consist of 2,000 samples (500 samples
+per chain with 4 independent chains). Is it sufficient for the effective sample size? Is it too much
+for the burn-in steps of 500,000 and the thinning steps of 1,000?}
 

### 2020-04-27 ncov.tex
what did I do today...

@@
 \begin{aligned}
-r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
-\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
-\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
+(r_1, \dots, r_7) &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
+(\bar{G}_1, \dots, \bar{G}_7) &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
+(\kappa_1, \dots, \kappa_7) &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
 \end{aligned}
@@
 Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
+
+Here, we assume that all 7 studies are equally weighted.
+During the initial phase of an outbreak of a novel virus such as SARS-CoV-2, uncertainties in parameter estimates are primarily driven by the underlying assumptions; narrow confidence intervals from a study are likely to reflect their strong assumptions rather than its precision of inference.
+Moreover, we cannot assess which assumptions are more realistic during the initial phase of an outbreak a novel virus due to the scarcity of information.
+Therefore, it is difficult to come up with a sensible way to weigh different estimates, and we do not explore weighted estimation.
+
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-As we add more models, the pooled estimates can become sharper even when the models no longer add more information.
+Typically, Bayesian hierarchical epidemic modeling seeks to leverage estimates across different outbreaks of a same disease.
+In this case, most estimates are based on reported number of cases from China, particularly from Wuhan province; thus, differences in estimates are likely to be driven by differences in the methods and assumptions, rather than epidemiological differences.
+In other words, the pooled estimates can become sharper as we add more models even when the models no longer add more information about the epidemic.
 Thus, the pooled estimator should be interpreted with care.

### 2020-04-27 ncov.tex
edits

@@
 
-\rev{The authors consider that all the 7 studies have the same weight in the estimation of basic
-reproduction weight even though their estimations and precision vary a lot. The authors might
-explain why they make this hypothesis and why they did not explore a weighted estimation
-based on the confidence of the estimation of R0.}
-
-\rev{In each chain, the authors use 500,000 sampling steps and remove every 1000th step.
-Therefore, the posterior distributions of each parameter consist of 2,000 samples (500 samples
-per chain with 4 independent chains). Is it sufficient for the effective sample size? Is it too much
-for the burn-in steps of 500,000 and the thinning steps of 1,000?}
-
 We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution.
@@
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-Typically, Bayesian hierarchical epidemic modeling seeks to leverage estimates across different outbreaks of a same disease.
-In this case, most estimates are based on reported number of cases from China, particularly from Wuhan province; thus, differences in estimates are likely to be driven by differences in the methods and assumptions, rather than epidemiological differences.
-In other words, the pooled estimates can become sharper as we add more models even when the models no longer add more information about the epidemic.
+Typically, Bayesian multilevel models seek to leverage estimates across spatially and/or temporally varying outbreaks of a same disease.
+In this case, most estimates are based on reported number of cases from China, particularly from Wuhan province.
+Differences among estimates are likely to reflect differences in estimation methods and underlying assumptions, rather than epidemiological differences.
+The pooled estimates can become sharper as we add more models even when the models no longer add more information about the epidemic.
 Thus, the pooled estimator should be interpreted with care.
+
 \rev{Authors state that “We note that this approach does not account for non- independence between the parameter estimates made by different modelers.” Authors should discuss this point more in depth.
@@
 
-\rev{Page 6, Line 29. Are the priors for $\mu_r$, $\mu_G$, and $\mu_\kappa$ weakly informative? Did the
-authors check the sensibility with new priors?}
+\subsection{Prior distributions}
+
 We use weakly informative priors on hyperparameters:
@@
 \end{linenomath*}
+These priors are chosen such that their 95\% quantiles are sufficiently wider than biologically realistic parameter ranges.
+Specifically, 95\% quantiles for $\mu_r$, $\mu_G$, and $\mu_\kappa$ are 0.02--0.40 $\mathrm{days}^{-1}$, 0.8--19.5 days, and 0.1--1.4, respectively.
+Therefore, we do not expect our results to be sensitive to these priors.
+
 We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
@@
 This phenomenon is a known problem \citep{gelman2006prior}.
-Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
-half-$t(\mu=0,\sigma=1,\nu=4)$  assumes \textit{a priori} that between-study variance is large,  while weak priors (e.g.
-half-Cauchy(0,5)) can lead to poor mixing.
-\rev{Page 6 Lines 44-47. Please clarify the sentence: “Alternative choices of prior … can lead to
-poor mixing”.}
+Alternative choices of prior for the between-study shape parameters are also suboptimal. 
+Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$ assumes \textit{a priori} that between-study variance is large (and therefore do not pool different estimates sufficiently).
+Weak priors (e.g. half-Cauchy(0,5)) can lead to inefficient sampling and poor convergence.
 

### 2020-04-28 compare_R0.tex
working on it...

% Created by tikzDevice version 0.12.3 on 2020-04-27 21:16:13
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 18.22) rectangle (572.66,247.13);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 18.22) rectangle (572.66,247.13);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 25.76) --
	(572.66, 25.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 79.98) --
	(572.66, 79.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,134.20) --
	(572.66,134.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,188.41) --
	(572.66,188.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,242.63) --
	(572.66,242.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 52.87) --
	(572.66, 52.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,107.09) --
	(572.66,107.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,161.31) --
	(572.66,161.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,215.52) --
	(572.66,215.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 18.22) --
	( 67.22,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.72, 18.22) --
	(133.72,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.23, 18.22) --
	(200.23,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (266.73, 18.22) --
	(266.73,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.24, 18.22) --
	(333.24,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.75, 18.22) --
	(399.75,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.25, 18.22) --
	(466.25,247.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.76, 18.22) --
	(532.76,247.13);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 55.23) -- (572.66, 55.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,122.16) -- (572.66,122.16);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25, 77.19) rectangle ( 69.18, 81.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25, 77.19) -- ( 69.18, 81.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25, 81.12) -- ( 69.18, 77.19);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.42, 78.87) -- (148.97, 78.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.19, 76.10) -- (146.19, 81.65);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (135.92, 58.38) --
	(139.84, 58.38) --
	(139.84, 62.30) --
	(135.92, 62.30) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (129.57, 75.55) --
	(132.21, 70.98) --
	(126.92, 70.98) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (121.25, 66.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.92, 66.49) -- (215.47, 66.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.70, 63.71) -- (212.70, 69.26);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (202.42, 66.60) --
	(206.35, 66.60) --
	(206.35, 70.53) --
	(202.42, 70.53) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (196.07, 82.29) --
	(198.71, 77.71) --
	(193.43, 77.71) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (187.76, 67.91) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.43, 66.66) -- (281.98, 66.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.20, 63.89) -- (279.20, 69.44);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (268.93, 76.35) --
	(272.85, 76.35) --
	(272.85, 80.27) --
	(268.93, 80.27) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (262.58, 98.08) --
	(265.22, 93.51) --
	(259.93, 93.51) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (254.26, 77.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.93,101.40) -- (348.48,101.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.71, 98.63) -- (345.71,104.18);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (335.43,112.03) --
	(339.36,112.03) --
	(339.36,115.96) --
	(335.43,115.96) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (329.08, 75.97) --
	(331.73, 71.40) --
	(326.44, 71.40) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (320.77,101.64) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.44, 55.76) -- (414.99, 55.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.22, 52.99) -- (412.22, 58.54);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (401.94, 47.95) --
	(405.86, 47.95) --
	(405.86, 51.88) --
	(401.94, 51.88) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (395.59, 99.37) --
	(398.23, 94.79) --
	(392.95, 94.79) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (387.28, 56.59) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.95,105.92) -- (481.50,105.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.72,103.15) -- (478.72,108.70);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (468.45,157.79) --
	(472.37,157.79) --
	(472.37,161.71) --
	(468.45,161.71) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (462.09, 93.47) --
	(464.74, 88.89) --
	(459.45, 88.89) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (453.78,145.76) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.45, 55.35) -- (548.00, 55.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.23, 52.58) -- (545.23, 58.13);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (534.95, 68.19) --
	(538.88, 68.19) --
	(538.88, 72.11) --
	(534.95, 72.11) --
	cycle;
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (528.60,108.72) --
	(531.24,104.14) --
	(525.96,104.14) --
	cycle;
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] (520.29, 66.12) circle (  1.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22,122.16) --
	( 67.22,122.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22,122.16) --
	( 67.22, 55.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 55.23) --
	( 67.22, 55.23);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.19,148.57) --
	(146.19,148.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.19,148.57) --
	(146.19, 42.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.19, 42.52) --
	(146.19, 42.52);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.88, 83.60) --
	(137.88, 83.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.88, 83.60) --
	(137.88, 38.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.88, 38.44) --
	(137.88, 38.44);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.57, 93.55) --
	(129.57, 93.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.57, 93.55) --
	(129.57, 57.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.57, 57.68) --
	(129.57, 57.68);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.25, 92.13) --
	(121.25, 92.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.25, 92.13) --
	(121.25, 40.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.25, 40.71) --
	(121.25, 40.71);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.70,109.43) --
	(212.70,109.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.70,109.43) --
	(212.70, 40.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.70, 40.25) --
	(212.70, 40.25);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (204.38,108.47) --
	(204.38,108.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (204.38,108.47) --
	(204.38, 40.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (204.38, 40.60) --
	(204.38, 40.60);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (196.07,112.61) --
	(196.07,112.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (196.07,112.61) --
	(196.07, 59.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (196.07, 59.21) --
	(196.07, 59.21);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.76,105.16) --
	(187.76,105.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.76,105.16) --
	(187.76, 40.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.76, 40.54) --
	(187.76, 40.54);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.20, 84.95) --
	(279.20, 84.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.20, 84.95) --
	(279.20, 52.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.20, 52.73) --
	(279.20, 52.73);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (270.89,103.19) --
	(270.89,103.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (270.89,103.19) --
	(270.89, 59.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (270.89, 59.79) --
	(270.89, 59.79);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (262.58,154.24) --
	(262.58,154.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (262.58,154.24) --
	(262.58, 65.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (262.58, 65.38) --
	(262.58, 65.38);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.26, 97.96) --
	(254.26, 97.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.26, 97.96) --
	(254.26, 60.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.26, 60.10) --
	(254.26, 60.10);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.71,126.97) --
	(345.71,126.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.71,126.97) --
	(345.71, 74.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.71, 74.24) --
	(345.71, 74.24);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.40,131.59) --
	(337.40,131.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.40,131.59) --
	(337.40,100.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.40,100.29) --
	(337.40,100.29);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (329.08,101.76) --
	(329.08,101.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (329.08,101.76) --
	(329.08, 55.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (329.08, 55.45) --
	(329.08, 55.45);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.77,107.13) --
	(320.77,107.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.77,107.13) --
	(320.77, 96.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.77, 96.34) --
	(320.77, 96.34);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.22, 98.29) --
	(412.22, 98.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.22, 98.29) --
	(412.22, 29.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.22, 29.35) --
	(412.22, 29.35);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.90, 91.28) --
	(403.90, 91.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.90, 91.28) --
	(403.90, 28.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.90, 28.63) --
	(403.90, 28.63);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (395.59,165.53) --
	(395.59,165.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (395.59,165.53) --
	(395.59, 61.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (395.59, 61.03) --
	(395.59, 61.03);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] (387.28, 96.20) --
	(387.28, 96.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (387.28, 96.20) --
	(387.28, 29.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (387.28, 29.36) --
	(387.28, 29.36);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.72,148.26) --
	(478.72,148.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.72,148.26) --
	(478.72, 74.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.72, 74.69) --
	(478.72, 74.69);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,223.85) --
	(470.41,223.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,223.85) --
	(470.41,114.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,114.90) --
	(470.41,114.90);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.09,146.48) --
	(462.09,146.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.09,146.48) --
	(462.09, 62.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.09, 62.15) --
	(462.09, 62.15);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] (453.78,191.49) --
	(453.78,191.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (453.78,191.49) --
	(453.78,109.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (453.78,109.22) --
	(453.78,109.22);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (545.23, 67.69) --
	(545.23, 67.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (545.23, 67.69) --
	(545.23, 46.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (545.23, 46.39) --
	(545.23, 46.39);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (536.91, 79.88) --
	(536.91, 79.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (536.91, 79.88) --
	(536.91, 63.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (536.91, 63.12) --
	(536.91, 63.12);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.60,236.72) --
	(528.60,236.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.60,236.72) --
	(528.60, 61.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.60, 61.88) --
	(528.60, 61.88);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] (520.29, 82.44) --
	(520.29, 82.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (520.29, 82.44) --
	(520.29, 52.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (520.29, 52.97) --
	(520.29, 52.97);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 18.22) rectangle (572.66,247.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 49.84) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,104.06) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,158.27) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,212.49) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 52.87) --
	( 27.31, 52.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,107.09) --
	( 27.31,107.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,161.31) --
	( 27.31,161.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,215.52) --
	( 27.31,215.52);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 15.47) --
	( 67.22, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.72, 15.47) --
	(133.72, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.23, 15.47) --
	(200.23, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (266.73, 15.47) --
	(266.73, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.24, 15.47) --
	(333.24, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.75, 15.47) --
	(399.75, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.25, 15.47) --
	(466.25, 18.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.76, 15.47) --
	(532.76, 18.22);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 67.22,  7.21) {Pooled estimate};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (133.72,  7.21) {Study 1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (200.23,  7.21) {Study 2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (266.73,  7.21) {Study 3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (333.24,  7.21) {Study 4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (399.75,  7.21) {Study 5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (466.25,  7.21) {Study 6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (532.76,  7.21) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,132.67) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 67.57,258.13) rectangle (532.41,283.58);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 78.57,263.63) rectangle ( 93.02,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{225,106,134}

\path[fill=fillColor] ( 85.79,270.85) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{225,106,134}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.01,270.85) -- ( 91.57,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (120.69,263.63) rectangle (135.14,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{144,152,0}

\path[fill=fillColor] (127.91,273.90) --
	(130.56,269.33) --
	(125.27,269.33) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{144,152,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.13,270.85) -- (133.69,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (238.08,263.63) rectangle (252.53,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{0,173,154}

\path[fill=fillColor] (243.34,268.89) --
	(247.27,268.89) --
	(247.27,272.82) --
	(243.34,272.82) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,173,154}

\path[draw=drawColor,line width= 0.6pt,line join=round] (239.52,270.85) -- (251.09,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (358.89,263.63) rectangle (373.35,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (363.34,270.85) -- (368.89,270.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.12,268.08) -- (366.12,273.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{145,131,230}

\path[draw=drawColor,line width= 0.6pt,line join=round] (360.34,270.85) -- (371.90,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (497.67,263.63) rectangle (512.12,278.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.93,268.89) rectangle (506.86,272.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.93,268.89) -- (506.86,272.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.93,272.82) -- (506.86,268.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.11,270.85) -- (510.68,270.85);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 98.52,267.82) {base};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.64,267.82) {exponential growth rate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (258.03,267.82) {mean generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (378.85,267.82) {generation-interval dispersion};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (517.62,267.82) {all};
\end{scope}
\end{tikzpicture}

\end{document}

### 2020-04-28 compare_assumption.tex
working on it...

% Created by tikzDevice version 0.12.3 on 2020-04-27 16:28:44
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{gray}{0.95}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (  0.00,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 38.04,  0.00) -- ( 38.04,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07,  0.00) -- ( 76.07,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (114.11,  0.00) -- (114.11,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.15,  0.00) -- (152.15,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.18,  0.00) -- (190.18,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.22,  0.00) -- (228.22,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.26,  0.00) -- (266.26,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.29,  0.00) -- (304.29,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.33,  0.00) -- (342.33,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (380.37,  0.00) -- (380.37,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.41,  0.00) -- (418.41,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.44,  0.00) -- (456.44,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.48,  0.00) -- (494.48,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (532.52,  0.00) -- (532.52,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.55,  0.00) -- (570.55,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (608.59,  0.00) -- (608.59,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (646.63,  0.00) -- (646.63,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (684.66,  0.00) -- (684.66,216.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (722.70,  0.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 11.41) -- (722.70, 11.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 22.82) -- (722.70, 22.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 34.23) -- (722.70, 34.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 45.64) -- (722.70, 45.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 57.06) -- (722.70, 57.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 68.47) -- (722.70, 68.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 79.88) -- (722.70, 79.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 91.29) -- (722.70, 91.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,102.70) -- (722.70,102.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,114.11) -- (722.70,114.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,125.52) -- (722.70,125.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,136.93) -- (722.70,136.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,148.34) -- (722.70,148.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,159.75) -- (722.70,159.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,171.17) -- (722.70,171.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,182.58) -- (722.70,182.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,193.99) -- (722.70,193.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,205.40) -- (722.70,205.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,216.81) -- (722.70,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (240.90,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (240.90,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 71.72, 37.19) rectangle (235.40,211.31);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 71.72, 37.19) rectangle (235.40,211.31);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 93.56, 37.19) --
	( 93.56,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (131.56, 37.19) --
	(131.56,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (169.55, 37.19) --
	(169.55,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (207.55, 37.19) --
	(207.55,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72, 49.93) --
	(235.40, 49.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72, 71.16) --
	(235.40, 71.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72, 92.40) --
	(235.40, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,113.63) --
	(235.40,113.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,134.87) --
	(235.40,134.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,156.10) --
	(235.40,156.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,177.34) --
	(235.40,177.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.72,198.57) --
	(235.40,198.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 74.56, 37.19) --
	( 74.56,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.56, 37.19) --
	(112.56,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.55, 37.19) --
	(150.55,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.55, 37.19) --
	(188.55,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.55, 37.19) --
	(226.55,211.31);
\definecolor{drawColor}{RGB}{255,165,0}
\definecolor{fillColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (131.47, 71.16) circle (  2.50);
\definecolor{drawColor}{RGB}{160,32,240}
\definecolor{fillColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (128.69, 92.40) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (128.47,113.63) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (117.88,198.57) circle (  2.50);
\definecolor{drawColor}{RGB}{160,32,240}
\definecolor{fillColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (169.46,134.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (108.06,156.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (170.22,177.34) circle (  2.50);
\definecolor{drawColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 95.51, 71.16) -- (167.58, 71.16);
\definecolor{drawColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 96.55, 92.40) -- (163.42, 92.40);

\path[draw=drawColor,line width= 0.8pt,line join=round] (114.86,113.63) -- (141.83,113.63);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) -- (117.88,198.57);
\definecolor{drawColor}{RGB}{160,32,240}

\path[draw=drawColor,line width= 0.8pt,line join=round] (164.38,134.87) -- (174.59,134.87);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 79.16,156.10) -- (149.69,156.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] (152.24,177.34) -- (187.73,177.34);
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (105.33, 49.85) --
	(105.57, 49.84) --
	(105.81, 49.83) --
	(106.05, 49.82) --
	(106.29, 49.81) --
	(106.53, 49.80) --
	(106.77, 49.79) --
	(107.01, 49.78) --
	(107.25, 49.77) --
	(107.49, 49.76) --
	(107.73, 49.74) --
	(107.97, 49.73) --
	(108.21, 49.71) --
	(108.45, 49.69) --
	(108.69, 49.68) --
	(108.93, 49.66) --
	(109.17, 49.64) --
	(109.41, 49.62) --
	(109.65, 49.60) --
	(109.89, 49.58) --
	(110.13, 49.56) --
	(110.37, 49.54) --
	(110.61, 49.51) --
	(110.85, 49.49) --
	(111.09, 49.47) --
	(111.33, 49.45) --
	(111.57, 49.43) --
	(111.81, 49.40) --
	(112.05, 49.38) --
	(112.29, 49.36) --
	(112.53, 49.34) --
	(112.77, 49.32) --
	(113.01, 49.30) --
	(113.25, 49.28) --
	(113.49, 49.26) --
	(113.73, 49.24) --
	(113.97, 49.22) --
	(114.21, 49.20) --
	(114.45, 49.18) --
	(114.69, 49.16) --
	(114.93, 49.13) --
	(115.17, 49.11) --
	(115.41, 49.09) --
	(115.65, 49.06) --
	(115.89, 49.03) --
	(116.13, 49.00) --
	(116.37, 48.96) --
	(116.61, 48.93) --
	(116.85, 48.89) --
	(117.09, 48.84) --
	(117.33, 48.79) --
	(117.57, 48.74) --
	(117.81, 48.69) --
	(118.05, 48.63) --
	(118.29, 48.57) --
	(118.53, 48.50) --
	(118.77, 48.44) --
	(119.01, 48.37) --
	(119.25, 48.30) --
	(119.49, 48.23) --
	(119.73, 48.16) --
	(119.97, 48.08) --
	(120.21, 48.01) --
	(120.44, 47.94) --
	(120.68, 47.87) --
	(120.92, 47.80) --
	(121.16, 47.72) --
	(121.40, 47.65) --
	(121.64, 47.58) --
	(121.88, 47.51) --
	(122.12, 47.43) --
	(122.36, 47.35) --
	(122.60, 47.27) --
	(122.84, 47.18) --
	(123.08, 47.09) --
	(123.32, 47.00) --
	(123.56, 46.89) --
	(123.80, 46.79) --
	(124.04, 46.67) --
	(124.28, 46.55) --
	(124.52, 46.43) --
	(124.76, 46.30) --
	(125.00, 46.16) --
	(125.24, 46.02) --
	(125.48, 45.88) --
	(125.72, 45.73) --
	(125.96, 45.58) --
	(126.20, 45.42) --
	(126.44, 45.27) --
	(126.68, 45.11) --
	(126.92, 44.96) --
	(127.16, 44.80) --
	(127.40, 44.64) --
	(127.64, 44.49) --
	(127.88, 44.33) --
	(128.12, 44.17) --
	(128.36, 44.02) --
	(128.60, 43.86) --
	(128.84, 43.70) --
	(129.08, 43.54) --
	(129.32, 43.38) --
	(129.56, 43.22) --
	(129.80, 43.06) --
	(130.04, 42.89) --
	(130.28, 42.73) --
	(130.52, 42.56) --
	(130.76, 42.40) --
	(131.00, 42.23) --
	(131.24, 42.07) --
	(131.48, 41.90) --
	(131.72, 41.74) --
	(131.96, 41.58) --
	(132.20, 41.43) --
	(132.44, 41.27) --
	(132.68, 41.13) --
	(132.92, 40.99) --
	(133.16, 40.85) --
	(133.40, 40.72) --
	(133.64, 40.60) --
	(133.88, 40.48) --
	(134.12, 40.37) --
	(134.36, 40.26) --
	(134.60, 40.16) --
	(134.84, 40.07) --
	(135.08, 39.98) --
	(135.32, 39.90) --
	(135.56, 39.82) --
	(135.80, 39.75) --
	(136.04, 39.68) --
	(136.28, 39.61) --
	(136.52, 39.55) --
	(136.76, 39.50) --
	(137.00, 39.45) --
	(137.24, 39.41) --
	(137.48, 39.38) --
	(137.72, 39.35) --
	(137.96, 39.33) --
	(138.20, 39.32) --
	(138.44, 39.31) --
	(138.68, 39.32) --
	(138.92, 39.32) --
	(139.16, 39.34) --
	(139.40, 39.37) --
	(139.64, 39.40) --
	(139.88, 39.44) --
	(140.12, 39.49) --
	(140.36, 39.54) --
	(140.60, 39.60) --
	(140.84, 39.67) --
	(141.08, 39.74) --
	(141.32, 39.82) --
	(141.56, 39.90) --
	(141.80, 39.99) --
	(142.04, 40.08) --
	(142.28, 40.18) --
	(142.52, 40.29) --
	(142.76, 40.40) --
	(143.00, 40.51) --
	(143.24, 40.62) --
	(143.48, 40.75) --
	(143.72, 40.87) --
	(143.96, 40.99) --
	(144.20, 41.12) --
	(144.44, 41.25) --
	(144.68, 41.38) --
	(144.92, 41.51) --
	(145.16, 41.65) --
	(145.40, 41.77) --
	(145.64, 41.90) --
	(145.88, 42.03) --
	(146.12, 42.15) --
	(146.36, 42.26) --
	(146.60, 42.38) --
	(146.84, 42.49) --
	(147.08, 42.59) --
	(147.32, 42.69) --
	(147.56, 42.79) --
	(147.80, 42.88) --
	(148.04, 42.97) --
	(148.28, 43.05) --
	(148.52, 43.13) --
	(148.76, 43.21) --
	(149.00, 43.29) --
	(149.24, 43.37) --
	(149.48, 43.45) --
	(149.72, 43.53) --
	(149.96, 43.61) --
	(150.20, 43.70) --
	(150.44, 43.78) --
	(150.68, 43.86) --
	(150.92, 43.95) --
	(151.16, 44.04) --
	(151.40, 44.13) --
	(151.64, 44.22) --
	(151.88, 44.32) --
	(152.12, 44.41) --
	(152.36, 44.52) --
	(152.60, 44.62) --
	(152.84, 44.72) --
	(153.08, 44.83) --
	(153.32, 44.94) --
	(153.56, 45.06) --
	(153.80, 45.18) --
	(154.04, 45.30) --
	(154.28, 45.42) --
	(154.52, 45.54) --
	(154.76, 45.67) --
	(155.00, 45.80) --
	(155.24, 45.92) --
	(155.48, 46.05) --
	(155.72, 46.18) --
	(155.96, 46.30) --
	(156.20, 46.42) --
	(156.44, 46.53) --
	(156.68, 46.65) --
	(156.92, 46.75) --
	(157.16, 46.85) --
	(157.40, 46.95) --
	(157.64, 47.04) --
	(157.88, 47.12) --
	(158.12, 47.20) --
	(158.36, 47.27) --
	(158.60, 47.34) --
	(158.84, 47.40) --
	(159.08, 47.45) --
	(159.32, 47.50) --
	(159.56, 47.55) --
	(159.80, 47.59) --
	(160.04, 47.62) --
	(160.28, 47.66) --
	(160.52, 47.69) --
	(160.76, 47.71) --
	(161.00, 47.74) --
	(161.24, 47.76) --
	(161.48, 47.78) --
	(161.72, 47.81) --
	(161.96, 47.83) --
	(162.20, 47.85) --
	(162.44, 47.88) --
	(162.68, 47.91) --
	(162.92, 47.94) --
	(163.16, 47.97) --
	(163.40, 48.01) --
	(163.64, 48.05) --
	(163.88, 48.09) --
	(164.12, 48.13) --
	(164.36, 48.18) --
	(164.60, 48.22) --
	(164.84, 48.27) --
	(165.08, 48.32) --
	(165.32, 48.36) --
	(165.56, 48.41) --
	(165.80, 48.46) --
	(166.04, 48.50) --
	(166.28, 48.54) --
	(166.52, 48.59) --
	(166.76, 48.63) --
	(167.00, 48.66) --
	(167.24, 48.70) --
	(167.48, 48.74) --
	(167.72, 48.77) --
	(167.96, 48.81) --
	(168.20, 48.84) --
	(168.44, 48.88) --
	(168.68, 48.91) --
	(168.92, 48.95) --
	(169.16, 48.98) --
	(169.40, 49.01) --
	(169.64, 49.05) --
	(169.88, 49.08) --
	(170.12, 49.11) --
	(170.36, 49.14) --
	(170.60, 49.17) --
	(170.84, 49.20) --
	(171.08, 49.22) --
	(171.32, 49.25) --
	(171.56, 49.27) --
	(171.80, 49.29) --
	(172.04, 49.31) --
	(172.28, 49.33) --
	(172.52, 49.34) --
	(172.76, 49.36) --
	(173.00, 49.37) --
	(173.24, 49.38) --
	(173.48, 49.39) --
	(173.72, 49.41) --
	(173.96, 49.42) --
	(174.20, 49.43) --
	(174.44, 49.44) --
	(174.68, 49.45) --
	(174.92, 49.46) --
	(175.16, 49.47) --
	(175.40, 49.48) --
	(175.64, 49.49) --
	(175.88, 49.50) --
	(176.12, 49.51) --
	(176.36, 49.52) --
	(176.60, 49.53) --
	(176.84, 49.53) --
	(177.08, 49.54) --
	(177.32, 49.55) --
	(177.56, 49.55) --
	(177.80, 49.56) --
	(178.04, 49.56) --
	(178.28, 49.57) --
	(178.52, 49.58) --
	(178.76, 49.58) --
	(179.00, 49.59) --
	(179.24, 49.59) --
	(179.48, 49.60) --
	(179.72, 49.60) --
	(179.96, 49.61) --
	(180.20, 49.62) --
	(180.44, 49.63) --
	(180.68, 49.64) --
	(180.92, 49.65) --
	(181.16, 49.66) --
	(181.40, 49.67) --
	(181.64, 49.68) --
	(181.88, 49.69) --
	(182.12, 49.70) --
	(182.36, 49.72) --
	(182.60, 49.73) --
	(182.84, 49.74) --
	(183.08, 49.75) --
	(183.32, 49.76) --
	(183.56, 49.77) --
	(183.80, 49.78) --
	(184.04, 49.79) --
	(184.28, 49.80) --
	(184.52, 49.81) --
	(184.76, 49.81) --
	(185.00, 49.82) --
	(185.24, 49.82) --
	(185.48, 49.83) --
	(185.72, 49.83) --
	(185.96, 49.83) --
	(186.20, 49.83) --
	(186.44, 49.83) --
	(186.68, 49.83) --
	(186.92, 49.83) --
	(187.16, 49.83) --
	(187.40, 49.83) --
	(187.64, 49.83) --
	(187.88, 49.82) --
	(188.12, 49.82) --
	(188.36, 49.82) --
	(188.60, 49.82) --
	(188.84, 49.81) --
	(189.08, 49.81) --
	(189.32, 49.81) --
	(189.56, 49.81) --
	(189.80, 49.81) --
	(190.04, 49.81) --
	(190.28, 49.81) --
	(190.52, 49.81) --
	(190.76, 49.81) --
	(191.00, 49.81) --
	(191.24, 49.81) --
	(191.48, 49.82) --
	(191.72, 49.82) --
	(191.96, 49.82) --
	(192.20, 49.83) --
	(192.44, 49.83) --
	(192.68, 49.84) --
	(192.92, 49.84) --
	(193.16, 49.85) --
	(193.40, 49.86) --
	(193.64, 49.86) --
	(193.88, 49.87) --
	(194.12, 49.87) --
	(194.36, 49.87) --
	(194.60, 49.88) --
	(194.84, 49.88) --
	(195.08, 49.88) --
	(195.32, 49.88) --
	(195.56, 49.88) --
	(195.80, 49.88) --
	(196.04, 49.88) --
	(196.28, 49.88) --
	(196.52, 49.88) --
	(196.76, 49.88) --
	(197.00, 49.88) --
	(197.24, 49.88) --
	(197.48, 49.88) --
	(197.72, 49.88) --
	(197.96, 49.88) --
	(198.20, 49.88) --
	(198.44, 49.88) --
	(198.68, 49.88) --
	(198.92, 49.88) --
	(199.16, 49.88) --
	(199.40, 49.89) --
	(199.64, 49.89) --
	(199.88, 49.89) --
	(200.12, 49.89) --
	(200.36, 49.90) --
	(200.60, 49.90) --
	(200.84, 49.90) --
	(201.08, 49.91) --
	(201.32, 49.91) --
	(201.56, 49.91) --
	(201.80, 49.91) --
	(202.04, 49.92) --
	(202.28, 49.92) --
	(202.52, 49.92) --
	(202.76, 49.92) --
	(203.00, 49.92) --
	(203.24, 49.93) --
	(203.48, 49.93) --
	(203.72, 49.93) --
	(203.96, 49.93) --
	(204.20, 49.93) --
	(204.44, 49.93) --
	(204.68, 49.93) --
	(204.92, 49.93) --
	(205.16, 49.93) --
	(205.40, 49.93) --
	(205.64, 49.93) --
	(205.88, 49.93) --
	(206.12, 49.93) --
	(206.36, 49.93) --
	(206.60, 49.93) --
	(206.84, 49.93) --
	(207.08, 49.93) --
	(207.32, 49.93) --
	(207.56, 49.93) --
	(207.80, 49.93) --
	(208.04, 49.93) --
	(208.28, 49.93) --
	(208.52, 49.93) --
	(208.76, 49.93) --
	(209.00, 49.93) --
	(209.24, 49.93) --
	(209.48, 49.93) --
	(209.72, 49.92) --
	(209.96, 49.92) --
	(210.20, 49.92) --
	(210.44, 49.92) --
	(210.68, 49.92) --
	(210.92, 49.92) --
	(211.16, 49.92) --
	(211.40, 49.91) --
	(211.64, 49.91) --
	(211.88, 49.91) --
	(212.12, 49.91) --
	(212.36, 49.91) --
	(212.60, 49.91) --
	(212.84, 49.91) --
	(213.08, 49.91) --
	(213.32, 49.91) --
	(213.56, 49.90) --
	(213.80, 49.90) --
	(214.04, 49.90) --
	(214.28, 49.91) --
	(214.52, 49.91) --
	(214.76, 49.91) --
	(215.00, 49.91) --
	(215.24, 49.91) --
	(215.48, 49.91) --
	(215.72, 49.91) --
	(215.96, 49.91) --
	(216.20, 49.92) --
	(216.44, 49.92) --
	(216.68, 49.92) --
	(216.92, 49.92) --
	(217.16, 49.92) --
	(217.40, 49.92) --
	(217.64, 49.92) --
	(217.88, 49.93) --
	(218.12, 49.93) --
	(218.36, 49.93) --
	(218.60, 49.93) --
	(218.84, 49.93) --
	(219.08, 49.93) --
	(219.32, 49.93) --
	(219.56, 49.93) --
	(219.80, 49.93) --
	(220.04, 49.93) --
	(220.28, 49.93) --
	(220.52, 49.93) --
	(220.76, 49.93) --
	(221.00, 49.93) --
	(221.24, 49.93) --
	(221.48, 49.93) --
	(221.72, 49.93) --
	(221.96, 49.93) --
	(222.20, 49.93) --
	(222.44, 49.93) --
	(222.68, 49.93) --
	(222.92, 49.93) --
	(223.16, 49.93) --
	(223.40, 49.93) --
	(223.64, 49.93) --
	(223.88, 49.93) --
	(224.12, 49.92) --
	(224.36, 49.92) --
	(224.60, 49.92) --
	(224.84, 49.92) --
	(225.08, 49.92) --
	(225.32, 49.92) --
	(225.56, 49.92) --
	(225.80, 49.91) --
	(226.04, 49.91) --
	(226.28, 49.91) --
	(226.52, 49.91) --
	(226.76, 49.91) --
	(227.00, 49.91) --
	(227.24, 49.91) --
	(227.48, 49.91) --
	(227.72, 49.90) --
	(227.96, 49.90) --
	(227.96, 49.96) --
	(227.72, 49.96) --
	(227.48, 49.96) --
	(227.24, 49.95) --
	(227.00, 49.95) --
	(226.76, 49.95) --
	(226.52, 49.95) --
	(226.28, 49.95) --
	(226.04, 49.95) --
	(225.80, 49.95) --
	(225.56, 49.95) --
	(225.32, 49.94) --
	(225.08, 49.94) --
	(224.84, 49.94) --
	(224.60, 49.94) --
	(224.36, 49.94) --
	(224.12, 49.94) --
	(223.88, 49.94) --
	(223.64, 49.93) --
	(223.40, 49.93) --
	(223.16, 49.93) --
	(222.92, 49.93) --
	(222.68, 49.93) --
	(222.44, 49.93) --
	(222.20, 49.93) --
	(221.96, 49.93) --
	(221.72, 49.93) --
	(221.48, 49.93) --
	(221.24, 49.93) --
	(221.00, 49.93) --
	(220.76, 49.93) --
	(220.52, 49.93) --
	(220.28, 49.93) --
	(220.04, 49.93) --
	(219.80, 49.93) --
	(219.56, 49.93) --
	(219.32, 49.93) --
	(219.08, 49.93) --
	(218.84, 49.93) --
	(218.60, 49.93) --
	(218.36, 49.93) --
	(218.12, 49.93) --
	(217.88, 49.94) --
	(217.64, 49.94) --
	(217.40, 49.94) --
	(217.16, 49.94) --
	(216.92, 49.94) --
	(216.68, 49.94) --
	(216.44, 49.94) --
	(216.20, 49.94) --
	(215.96, 49.95) --
	(215.72, 49.95) --
	(215.48, 49.95) --
	(215.24, 49.95) --
	(215.00, 49.95) --
	(214.76, 49.95) --
	(214.52, 49.95) --
	(214.28, 49.96) --
	(214.04, 49.96) --
	(213.80, 49.96) --
	(213.56, 49.96) --
	(213.32, 49.96) --
	(213.08, 49.96) --
	(212.84, 49.95) --
	(212.60, 49.95) --
	(212.36, 49.95) --
	(212.12, 49.95) --
	(211.88, 49.95) --
	(211.64, 49.95) --
	(211.40, 49.95) --
	(211.16, 49.94) --
	(210.92, 49.94) --
	(210.68, 49.94) --
	(210.44, 49.94) --
	(210.20, 49.94) --
	(209.96, 49.94) --
	(209.72, 49.94) --
	(209.48, 49.94) --
	(209.24, 49.93) --
	(209.00, 49.93) --
	(208.76, 49.93) --
	(208.52, 49.93) --
	(208.28, 49.93) --
	(208.04, 49.93) --
	(207.80, 49.93) --
	(207.56, 49.93) --
	(207.32, 49.93) --
	(207.08, 49.93) --
	(206.84, 49.93) --
	(206.60, 49.93) --
	(206.36, 49.93) --
	(206.12, 49.93) --
	(205.88, 49.93) --
	(205.64, 49.93) --
	(205.40, 49.93) --
	(205.16, 49.93) --
	(204.92, 49.93) --
	(204.68, 49.93) --
	(204.44, 49.93) --
	(204.20, 49.93) --
	(203.96, 49.93) --
	(203.72, 49.93) --
	(203.48, 49.93) --
	(203.24, 49.94) --
	(203.00, 49.94) --
	(202.76, 49.94) --
	(202.52, 49.94) --
	(202.28, 49.94) --
	(202.04, 49.94) --
	(201.80, 49.95) --
	(201.56, 49.95) --
	(201.32, 49.95) --
	(201.08, 49.95) --
	(200.84, 49.96) --
	(200.60, 49.96) --
	(200.36, 49.96) --
	(200.12, 49.97) --
	(199.88, 49.97) --
	(199.64, 49.97) --
	(199.40, 49.98) --
	(199.16, 49.98) --
	(198.92, 49.98) --
	(198.68, 49.98) --
	(198.44, 49.98) --
	(198.20, 49.98) --
	(197.96, 49.98) --
	(197.72, 49.98) --
	(197.48, 49.98) --
	(197.24, 49.98) --
	(197.00, 49.98) --
	(196.76, 49.98) --
	(196.52, 49.98) --
	(196.28, 49.98) --
	(196.04, 49.98) --
	(195.80, 49.98) --
	(195.56, 49.98) --
	(195.32, 49.98) --
	(195.08, 49.98) --
	(194.84, 49.98) --
	(194.60, 49.98) --
	(194.36, 49.99) --
	(194.12, 49.99) --
	(193.88, 50.00) --
	(193.64, 50.00) --
	(193.40, 50.01) --
	(193.16, 50.01) --
	(192.92, 50.02) --
	(192.68, 50.02) --
	(192.44, 50.03) --
	(192.20, 50.03) --
	(191.96, 50.04) --
	(191.72, 50.04) --
	(191.48, 50.05) --
	(191.24, 50.05) --
	(191.00, 50.05) --
	(190.76, 50.05) --
	(190.52, 50.05) --
	(190.28, 50.06) --
	(190.04, 50.06) --
	(189.80, 50.05) --
	(189.56, 50.05) --
	(189.32, 50.05) --
	(189.08, 50.05) --
	(188.84, 50.05) --
	(188.60, 50.04) --
	(188.36, 50.04) --
	(188.12, 50.04) --
	(187.88, 50.04) --
	(187.64, 50.03) --
	(187.40, 50.03) --
	(187.16, 50.03) --
	(186.92, 50.03) --
	(186.68, 50.03) --
	(186.44, 50.03) --
	(186.20, 50.03) --
	(185.96, 50.03) --
	(185.72, 50.03) --
	(185.48, 50.04) --
	(185.24, 50.04) --
	(185.00, 50.04) --
	(184.76, 50.05) --
	(184.52, 50.06) --
	(184.28, 50.06) --
	(184.04, 50.07) --
	(183.80, 50.08) --
	(183.56, 50.09) --
	(183.32, 50.10) --
	(183.08, 50.11) --
	(182.84, 50.12) --
	(182.60, 50.13) --
	(182.36, 50.14) --
	(182.12, 50.16) --
	(181.88, 50.17) --
	(181.64, 50.18) --
	(181.40, 50.19) --
	(181.16, 50.20) --
	(180.92, 50.21) --
	(180.68, 50.22) --
	(180.44, 50.23) --
	(180.20, 50.24) --
	(179.96, 50.25) --
	(179.72, 50.26) --
	(179.48, 50.26) --
	(179.24, 50.27) --
	(179.00, 50.28) --
	(178.76, 50.28) --
	(178.52, 50.29) --
	(178.28, 50.29) --
	(178.04, 50.30) --
	(177.80, 50.30) --
	(177.56, 50.31) --
	(177.32, 50.31) --
	(177.08, 50.32) --
	(176.84, 50.33) --
	(176.60, 50.33) --
	(176.36, 50.34) --
	(176.12, 50.35) --
	(175.88, 50.36) --
	(175.64, 50.37) --
	(175.40, 50.38) --
	(175.16, 50.39) --
	(174.92, 50.40) --
	(174.68, 50.41) --
	(174.44, 50.42) --
	(174.20, 50.43) --
	(173.96, 50.44) --
	(173.72, 50.45) --
	(173.48, 50.47) --
	(173.24, 50.48) --
	(173.00, 50.49) --
	(172.76, 50.50) --
	(172.52, 50.52) --
	(172.28, 50.53) --
	(172.04, 50.55) --
	(171.80, 50.57) --
	(171.56, 50.59) --
	(171.32, 50.61) --
	(171.08, 50.64) --
	(170.84, 50.66) --
	(170.60, 50.69) --
	(170.36, 50.72) --
	(170.12, 50.75) --
	(169.88, 50.78) --
	(169.64, 50.81) --
	(169.40, 50.85) --
	(169.16, 50.88) --
	(168.92, 50.91) --
	(168.68, 50.95) --
	(168.44, 50.98) --
	(168.20, 51.02) --
	(167.96, 51.05) --
	(167.72, 51.09) --
	(167.48, 51.12) --
	(167.24, 51.16) --
	(167.00, 51.20) --
	(166.76, 51.24) --
	(166.52, 51.28) --
	(166.28, 51.32) --
	(166.04, 51.36) --
	(165.80, 51.41) --
	(165.56, 51.45) --
	(165.32, 51.50) --
	(165.08, 51.55) --
	(164.84, 51.59) --
	(164.60, 51.64) --
	(164.36, 51.69) --
	(164.12, 51.73) --
	(163.88, 51.77) --
	(163.64, 51.81) --
	(163.40, 51.85) --
	(163.16, 51.89) --
	(162.92, 51.92) --
	(162.68, 51.95) --
	(162.44, 51.98) --
	(162.20, 52.01) --
	(161.96, 52.03) --
	(161.72, 52.05) --
	(161.48, 52.08) --
	(161.24, 52.10) --
	(161.00, 52.12) --
	(160.76, 52.15) --
	(160.52, 52.18) --
	(160.28, 52.21) --
	(160.04, 52.24) --
	(159.80, 52.27) --
	(159.56, 52.31) --
	(159.32, 52.36) --
	(159.08, 52.41) --
	(158.84, 52.46) --
	(158.60, 52.52) --
	(158.36, 52.59) --
	(158.12, 52.66) --
	(157.88, 52.74) --
	(157.64, 52.82) --
	(157.40, 52.91) --
	(157.16, 53.01) --
	(156.92, 53.11) --
	(156.68, 53.22) --
	(156.44, 53.33) --
	(156.20, 53.44) --
	(155.96, 53.56) --
	(155.72, 53.69) --
	(155.48, 53.81) --
	(155.24, 53.94) --
	(155.00, 54.06) --
	(154.76, 54.19) --
	(154.52, 54.32) --
	(154.28, 54.44) --
	(154.04, 54.56) --
	(153.80, 54.69) --
	(153.56, 54.80) --
	(153.32, 54.92) --
	(153.08, 55.03) --
	(152.84, 55.14) --
	(152.60, 55.24) --
	(152.36, 55.35) --
	(152.12, 55.45) --
	(151.88, 55.54) --
	(151.64, 55.64) --
	(151.40, 55.73) --
	(151.16, 55.82) --
	(150.92, 55.91) --
	(150.68, 56.00) --
	(150.44, 56.08) --
	(150.20, 56.16) --
	(149.96, 56.25) --
	(149.72, 56.33) --
	(149.48, 56.41) --
	(149.24, 56.49) --
	(149.00, 56.57) --
	(148.76, 56.65) --
	(148.52, 56.73) --
	(148.28, 56.81) --
	(148.04, 56.90) --
	(147.80, 56.98) --
	(147.56, 57.08) --
	(147.32, 57.17) --
	(147.08, 57.27) --
	(146.84, 57.38) --
	(146.60, 57.48) --
	(146.36, 57.60) --
	(146.12, 57.71) --
	(145.88, 57.84) --
	(145.64, 57.96) --
	(145.40, 58.09) --
	(145.16, 58.22) --
	(144.92, 58.35) --
	(144.68, 58.48) --
	(144.44, 58.61) --
	(144.20, 58.74) --
	(143.96, 58.87) --
	(143.72, 58.99) --
	(143.48, 59.12) --
	(143.24, 59.24) --
	(143.00, 59.35) --
	(142.76, 59.47) --
	(142.52, 59.57) --
	(142.28, 59.68) --
	(142.04, 59.78) --
	(141.80, 59.87) --
	(141.56, 59.96) --
	(141.32, 60.04) --
	(141.08, 60.12) --
	(140.84, 60.20) --
	(140.60, 60.26) --
	(140.36, 60.32) --
	(140.12, 60.37) --
	(139.88, 60.42) --
	(139.64, 60.46) --
	(139.40, 60.49) --
	(139.16, 60.52) --
	(138.92, 60.54) --
	(138.68, 60.55) --
	(138.44, 60.55) --
	(138.20, 60.54) --
	(137.96, 60.53) --
	(137.72, 60.51) --
	(137.48, 60.48) --
	(137.24, 60.45) --
	(137.00, 60.41) --
	(136.76, 60.36) --
	(136.52, 60.31) --
	(136.28, 60.25) --
	(136.04, 60.18) --
	(135.80, 60.12) --
	(135.56, 60.04) --
	(135.32, 59.96) --
	(135.08, 59.88) --
	(134.84, 59.79) --
	(134.60, 59.70) --
	(134.36, 59.60) --
	(134.12, 59.49) --
	(133.88, 59.38) --
	(133.64, 59.26) --
	(133.40, 59.14) --
	(133.16, 59.01) --
	(132.92, 58.87) --
	(132.68, 58.73) --
	(132.44, 58.59) --
	(132.20, 58.43) --
	(131.96, 58.28) --
	(131.72, 58.12) --
	(131.48, 57.96) --
	(131.24, 57.79) --
	(131.00, 57.63) --
	(130.76, 57.46) --
	(130.52, 57.30) --
	(130.28, 57.13) --
	(130.04, 56.97) --
	(129.80, 56.80) --
	(129.56, 56.64) --
	(129.32, 56.48) --
	(129.08, 56.32) --
	(128.84, 56.16) --
	(128.60, 56.00) --
	(128.36, 55.84) --
	(128.12, 55.69) --
	(127.88, 55.53) --
	(127.64, 55.37) --
	(127.40, 55.22) --
	(127.16, 55.06) --
	(126.92, 54.90) --
	(126.68, 54.75) --
	(126.44, 54.59) --
	(126.20, 54.44) --
	(125.96, 54.28) --
	(125.72, 54.13) --
	(125.48, 53.98) --
	(125.24, 53.84) --
	(125.00, 53.70) --
	(124.76, 53.56) --
	(124.52, 53.43) --
	(124.28, 53.31) --
	(124.04, 53.19) --
	(123.80, 53.07) --
	(123.56, 52.97) --
	(123.32, 52.87) --
	(123.08, 52.77) --
	(122.84, 52.68) --
	(122.60, 52.59) --
	(122.36, 52.51) --
	(122.12, 52.43) --
	(121.88, 52.36) --
	(121.64, 52.28) --
	(121.40, 52.21) --
	(121.16, 52.14) --
	(120.92, 52.06) --
	(120.68, 51.99) --
	(120.44, 51.92) --
	(120.21, 51.85) --
	(119.97, 51.78) --
	(119.73, 51.71) --
	(119.49, 51.63) --
	(119.25, 51.56) --
	(119.01, 51.49) --
	(118.77, 51.42) --
	(118.53, 51.36) --
	(118.29, 51.29) --
	(118.05, 51.23) --
	(117.81, 51.17) --
	(117.57, 51.12) --
	(117.33, 51.07) --
	(117.09, 51.02) --
	(116.85, 50.98) --
	(116.61, 50.93) --
	(116.37, 50.90) --
	(116.13, 50.86) --
	(115.89, 50.83) --
	(115.65, 50.80) --
	(115.41, 50.78) --
	(115.17, 50.75) --
	(114.93, 50.73) --
	(114.69, 50.70) --
	(114.45, 50.68) --
	(114.21, 50.66) --
	(113.97, 50.64) --
	(113.73, 50.62) --
	(113.49, 50.60) --
	(113.25, 50.58) --
	(113.01, 50.56) --
	(112.77, 50.54) --
	(112.53, 50.52) --
	(112.29, 50.50) --
	(112.05, 50.48) --
	(111.81, 50.46) --
	(111.57, 50.44) --
	(111.33, 50.41) --
	(111.09, 50.39) --
	(110.85, 50.37) --
	(110.61, 50.35) --
	(110.37, 50.33) --
	(110.13, 50.30) --
	(109.89, 50.28) --
	(109.65, 50.26) --
	(109.41, 50.24) --
	(109.17, 50.22) --
	(108.93, 50.20) --
	(108.69, 50.18) --
	(108.45, 50.17) --
	(108.21, 50.15) --
	(107.97, 50.13) --
	(107.73, 50.12) --
	(107.49, 50.11) --
	(107.25, 50.09) --
	(107.01, 50.08) --
	(106.77, 50.07) --
	(106.53, 50.06) --
	(106.29, 50.05) --
	(106.05, 50.04) --
	(105.81, 50.03) --
	(105.57, 50.02) --
	(105.33, 50.01) --
	(105.33, 49.85) --
	cycle;
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 71.72, 37.19) rectangle (235.40,211.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77, 46.90) {Pooled estimate};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77, 68.13) {Study 1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77, 89.37) {Study 2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,110.60) {Study 3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,131.84) {Study 4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,153.07) {Study 5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,174.30) {Study 6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.77,195.54) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97, 49.93) --
	( 71.72, 49.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97, 71.16) --
	( 71.72, 71.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97, 92.40) --
	( 71.72, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,113.63) --
	( 71.72,113.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,134.87) --
	( 71.72,134.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,156.10) --
	( 71.72,156.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,177.34) --
	( 71.72,177.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.97,198.57) --
	( 71.72,198.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 74.56, 34.44) --
	( 74.56, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.56, 34.44) --
	(112.56, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.55, 34.44) --
	(150.55, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.55, 34.44) --
	(188.55, 37.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.55, 34.44) --
	(226.55, 37.19);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 74.56, 26.18) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.56, 26.18) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (150.55, 26.18) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (188.55, 26.18) {0.3};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (226.55, 26.18) {0.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 70.64, 11.26) {E};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 78.12, 11.26) {x};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 83.92, 11.26) {p};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 90.03, 11.26) {o};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 95.53, 11.26) {n};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (101.64, 11.26) {e};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (106.53, 11.26) {n};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (112.64, 11.26) {t};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (116.92, 11.26) {i};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (119.97, 11.26) {a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (125.47, 11.26) {l};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (128.52, 11.26) { };

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (134.02, 11.26) {g};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (139.52, 11.26) {r};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (143.83, 11.26) {o};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (149.33, 11.26) {w};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (157.27, 11.26) {t};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (161.55, 11.26) {h};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (167.66, 11.26) { };

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (173.16, 11.26) {r};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (177.46, 11.26) {a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (182.96, 11.26) {t};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (187.24, 11.26) {e};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (192.13, 11.26) { };

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.38] at (197.62, 11.26) {(};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (202.97, 11.26) {d};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (209.08, 11.26) {a};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (214.58, 11.26) {y};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (220.38, 11.26) {s};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.77] at (224.72, 15.76) {-};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.77] at (227.29, 15.76) {1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.38] at (231.14, 11.26) {)};
\end{scope}
\begin{scope}
\path[clip] (240.90,  0.00) rectangle (481.80,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (240.90,  0.00) rectangle (481.80,216.81);
\end{scope}
\begin{scope}
\path[clip] (312.62, 31.02) rectangle (476.30,211.31);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (312.62, 31.02) rectangle (476.30,211.31);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (337.45, 31.02) --
	(337.45,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (375.96, 31.02) --
	(375.96,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (414.47, 31.02) --
	(414.47,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (452.97, 31.02) --
	(452.97,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62, 44.21) --
	(476.30, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62, 66.20) --
	(476.30, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62, 88.18) --
	(476.30, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,110.17) --
	(476.30,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,132.16) --
	(476.30,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,154.14) --
	(476.30,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,176.13) --
	(476.30,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (312.62,198.12) --
	(476.30,198.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (318.20, 31.02) --
	(318.20,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.71, 31.02) --
	(356.71,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (395.21, 31.02) --
	(395.21,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.72, 31.02) --
	(433.72,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.23, 31.02) --
	(472.23,211.31);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (395.21, 66.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (364.41, 88.18) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (364.41,110.17) circle (  2.50);
\definecolor{drawColor}{RGB}{255,165,0}
\definecolor{fillColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (356.66,198.12) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (349.00,132.16) circle (  2.50);
\definecolor{drawColor}{RGB}{255,165,0}
\definecolor{fillColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (404.72,154.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (356.69,176.13) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) -- (395.21, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) -- (364.41, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) -- (364.41,110.17);
\definecolor{drawColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (320.06,198.12) -- (393.26,198.12);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) -- (349.00,132.16);
\definecolor{drawColor}{RGB}{255,165,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] (340.83,154.14) -- (468.86,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (349.40,176.13) -- (364.04,176.13);
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (327.85, 44.18) --
	(328.07, 44.18) --
	(328.28, 44.18) --
	(328.50, 44.18) --
	(328.72, 44.18) --
	(328.93, 44.18) --
	(329.15, 44.18) --
	(329.36, 44.17) --
	(329.58, 44.17) --
	(329.79, 44.17) --
	(330.01, 44.17) --
	(330.23, 44.16) --
	(330.44, 44.16) --
	(330.66, 44.16) --
	(330.87, 44.15) --
	(331.09, 44.15) --
	(331.30, 44.14) --
	(331.52, 44.14) --
	(331.74, 44.13) --
	(331.95, 44.13) --
	(332.17, 44.13) --
	(332.38, 44.12) --
	(332.60, 44.12) --
	(332.81, 44.12) --
	(333.03, 44.11) --
	(333.25, 44.11) --
	(333.46, 44.11) --
	(333.68, 44.10) --
	(333.89, 44.10) --
	(334.11, 44.10) --
	(334.32, 44.09) --
	(334.54, 44.09) --
	(334.76, 44.08) --
	(334.97, 44.08) --
	(335.19, 44.07) --
	(335.40, 44.07) --
	(335.62, 44.06) --
	(335.83, 44.06) --
	(336.05, 44.05) --
	(336.27, 44.05) --
	(336.48, 44.05) --
	(336.70, 44.05) --
	(336.91, 44.05) --
	(337.13, 44.05) --
	(337.34, 44.05) --
	(337.56, 44.05) --
	(337.78, 44.05) --
	(337.99, 44.05) --
	(338.21, 44.06) --
	(338.42, 44.06) --
	(338.64, 44.06) --
	(338.86, 44.07) --
	(339.07, 44.07) --
	(339.29, 44.07) --
	(339.50, 44.08) --
	(339.72, 44.08) --
	(339.93, 44.08) --
	(340.15, 44.09) --
	(340.37, 44.09) --
	(340.58, 44.09) --
	(340.80, 44.09) --
	(341.01, 44.08) --
	(341.23, 44.08) --
	(341.44, 44.07) --
	(341.66, 44.05) --
	(341.88, 44.04) --
	(342.09, 44.02) --
	(342.31, 43.99) --
	(342.52, 43.97) --
	(342.74, 43.94) --
	(342.95, 43.91) --
	(343.17, 43.88) --
	(343.39, 43.85) --
	(343.60, 43.81) --
	(343.82, 43.78) --
	(344.03, 43.75) --
	(344.25, 43.72) --
	(344.46, 43.68) --
	(344.68, 43.65) --
	(344.90, 43.63) --
	(345.11, 43.60) --
	(345.33, 43.57) --
	(345.54, 43.54) --
	(345.76, 43.52) --
	(345.97, 43.49) --
	(346.19, 43.46) --
	(346.41, 43.43) --
	(346.62, 43.40) --
	(346.84, 43.37) --
	(347.05, 43.34) --
	(347.27, 43.30) --
	(347.48, 43.26) --
	(347.70, 43.22) --
	(347.92, 43.17) --
	(348.13, 43.12) --
	(348.35, 43.07) --
	(348.56, 43.01) --
	(348.78, 42.95) --
	(348.99, 42.89) --
	(349.21, 42.83) --
	(349.43, 42.76) --
	(349.64, 42.70) --
	(349.86, 42.63) --
	(350.07, 42.56) --
	(350.29, 42.49) --
	(350.50, 42.42) --
	(350.72, 42.34) --
	(350.94, 42.27) --
	(351.15, 42.20) --
	(351.37, 42.12) --
	(351.58, 42.05) --
	(351.80, 41.98) --
	(352.02, 41.91) --
	(352.23, 41.84) --
	(352.45, 41.77) --
	(352.66, 41.71) --
	(352.88, 41.64) --
	(353.09, 41.57) --
	(353.31, 41.50) --
	(353.53, 41.42) --
	(353.74, 41.34) --
	(353.96, 41.25) --
	(354.17, 41.16) --
	(354.39, 41.05) --
	(354.60, 40.93) --
	(354.82, 40.80) --
	(355.04, 40.65) --
	(355.25, 40.49) --
	(355.47, 40.33) --
	(355.68, 40.15) --
	(355.90, 39.96) --
	(356.11, 39.77) --
	(356.33, 39.57) --
	(356.55, 39.37) --
	(356.76, 39.16) --
	(356.98, 38.96) --
	(357.19, 38.76) --
	(357.41, 38.56) --
	(357.62, 38.36) --
	(357.84, 38.17) --
	(358.06, 37.98) --
	(358.27, 37.79) --
	(358.49, 37.61) --
	(358.70, 37.43) --
	(358.92, 37.25) --
	(359.13, 37.08) --
	(359.35, 36.90) --
	(359.57, 36.72) --
	(359.78, 36.54) --
	(360.00, 36.36) --
	(360.21, 36.18) --
	(360.43, 35.99) --
	(360.64, 35.80) --
	(360.86, 35.62) --
	(361.08, 35.43) --
	(361.29, 35.25) --
	(361.51, 35.07) --
	(361.72, 34.89) --
	(361.94, 34.72) --
	(362.15, 34.55) --
	(362.37, 34.39) --
	(362.59, 34.23) --
	(362.80, 34.08) --
	(363.02, 33.95) --
	(363.23, 33.82) --
	(363.45, 33.70) --
	(363.66, 33.60) --
	(363.88, 33.51) --
	(364.10, 33.43) --
	(364.31, 33.36) --
	(364.53, 33.31) --
	(364.74, 33.27) --
	(364.96, 33.24) --
	(365.18, 33.22) --
	(365.39, 33.22) --
	(365.61, 33.22) --
	(365.82, 33.24) --
	(366.04, 33.27) --
	(366.25, 33.31) --
	(366.47, 33.36) --
	(366.69, 33.41) --
	(366.90, 33.48) --
	(367.12, 33.55) --
	(367.33, 33.63) --
	(367.55, 33.71) --
	(367.76, 33.80) --
	(367.98, 33.90) --
	(368.20, 33.99) --
	(368.41, 34.10) --
	(368.63, 34.20) --
	(368.84, 34.31) --
	(369.06, 34.42) --
	(369.27, 34.53) --
	(369.49, 34.64) --
	(369.71, 34.76) --
	(369.92, 34.88) --
	(370.14, 34.99) --
	(370.35, 35.11) --
	(370.57, 35.24) --
	(370.78, 35.36) --
	(371.00, 35.49) --
	(371.22, 35.61) --
	(371.43, 35.75) --
	(371.65, 35.88) --
	(371.86, 36.02) --
	(372.08, 36.17) --
	(372.29, 36.32) --
	(372.51, 36.47) --
	(372.73, 36.63) --
	(372.94, 36.79) --
	(373.16, 36.96) --
	(373.37, 37.12) --
	(373.59, 37.28) --
	(373.80, 37.44) --
	(374.02, 37.60) --
	(374.24, 37.76) --
	(374.45, 37.92) --
	(374.67, 38.07) --
	(374.88, 38.23) --
	(375.10, 38.38) --
	(375.31, 38.54) --
	(375.53, 38.71) --
	(375.75, 38.87) --
	(375.96, 39.04) --
	(376.18, 39.22) --
	(376.39, 39.39) --
	(376.61, 39.57) --
	(376.82, 39.75) --
	(377.04, 39.92) --
	(377.26, 40.09) --
	(377.47, 40.26) --
	(377.69, 40.42) --
	(377.90, 40.57) --
	(378.12, 40.71) --
	(378.33, 40.84) --
	(378.55, 40.96) --
	(378.77, 41.07) --
	(378.98, 41.17) --
	(379.20, 41.26) --
	(379.41, 41.34) --
	(379.63, 41.42) --
	(379.85, 41.48) --
	(380.06, 41.55) --
	(380.28, 41.61) --
	(380.49, 41.66) --
	(380.71, 41.72) --
	(380.92, 41.78) --
	(381.14, 41.83) --
	(381.36, 41.89) --
	(381.57, 41.95) --
	(381.79, 42.01) --
	(382.00, 42.07) --
	(382.22, 42.13) --
	(382.43, 42.19) --
	(382.65, 42.25) --
	(382.87, 42.30) --
	(383.08, 42.36) --
	(383.30, 42.42) --
	(383.51, 42.47) --
	(383.73, 42.52) --
	(383.94, 42.57) --
	(384.16, 42.62) --
	(384.38, 42.67) --
	(384.59, 42.72) --
	(384.81, 42.77) --
	(385.02, 42.82) --
	(385.24, 42.87) --
	(385.45, 42.92) --
	(385.67, 42.97) --
	(385.89, 43.02) --
	(386.10, 43.07) --
	(386.32, 43.12) --
	(386.53, 43.17) --
	(386.75, 43.21) --
	(386.96, 43.26) --
	(387.18, 43.30) --
	(387.40, 43.34) --
	(387.61, 43.37) --
	(387.83, 43.40) --
	(388.04, 43.43) --
	(388.26, 43.46) --
	(388.47, 43.48) --
	(388.69, 43.50) --
	(388.91, 43.52) --
	(389.12, 43.54) --
	(389.34, 43.56) --
	(389.55, 43.58) --
	(389.77, 43.60) --
	(389.98, 43.62) --
	(390.20, 43.64) --
	(390.42, 43.66) --
	(390.63, 43.68) --
	(390.85, 43.71) --
	(391.06, 43.73) --
	(391.28, 43.75) --
	(391.49, 43.77) --
	(391.71, 43.79) --
	(391.93, 43.80) --
	(392.14, 43.82) --
	(392.36, 43.83) --
	(392.57, 43.85) --
	(392.79, 43.86) --
	(393.01, 43.87) --
	(393.22, 43.87) --
	(393.44, 43.88) --
	(393.65, 43.89) --
	(393.87, 43.89) --
	(394.08, 43.89) --
	(394.30, 43.90) --
	(394.52, 43.90) --
	(394.73, 43.90) --
	(394.95, 43.91) --
	(395.16, 43.91) --
	(395.38, 43.92) --
	(395.59, 43.92) --
	(395.81, 43.93) --
	(396.03, 43.93) --
	(396.24, 43.94) --
	(396.46, 43.95) --
	(396.67, 43.95) --
	(396.89, 43.96) --
	(397.10, 43.97) --
	(397.32, 43.98) --
	(397.54, 43.99) --
	(397.75, 44.00) --
	(397.97, 44.01) --
	(398.18, 44.02) --
	(398.40, 44.02) --
	(398.61, 44.03) --
	(398.83, 44.04) --
	(399.05, 44.05) --
	(399.26, 44.05) --
	(399.48, 44.06) --
	(399.69, 44.06) --
	(399.91, 44.06) --
	(400.12, 44.06) --
	(400.34, 44.06) --
	(400.56, 44.06) --
	(400.77, 44.06) --
	(400.99, 44.06) --
	(401.20, 44.05) --
	(401.42, 44.05) --
	(401.63, 44.05) --
	(401.85, 44.04) --
	(402.07, 44.04) --
	(402.28, 44.04) --
	(402.50, 44.03) --
	(402.71, 44.03) --
	(402.93, 44.03) --
	(403.14, 44.04) --
	(403.36, 44.04) --
	(403.58, 44.05) --
	(403.79, 44.05) --
	(404.01, 44.06) --
	(404.22, 44.07) --
	(404.44, 44.07) --
	(404.65, 44.08) --
	(404.87, 44.09) --
	(405.09, 44.09) --
	(405.30, 44.10) --
	(405.52, 44.11) --
	(405.73, 44.11) --
	(405.95, 44.12) --
	(406.16, 44.12) --
	(406.38, 44.12) --
	(406.60, 44.13) --
	(406.81, 44.13) --
	(407.03, 44.13) --
	(407.24, 44.13) --
	(407.46, 44.14) --
	(407.68, 44.14) --
	(407.89, 44.14) --
	(408.11, 44.15) --
	(408.32, 44.15) --
	(408.54, 44.15) --
	(408.75, 44.16) --
	(408.97, 44.16) --
	(409.19, 44.16) --
	(409.40, 44.17) --
	(409.62, 44.17) --
	(409.83, 44.17) --
	(410.05, 44.18) --
	(410.26, 44.18) --
	(410.48, 44.19) --
	(410.70, 44.19) --
	(410.91, 44.19) --
	(411.13, 44.20) --
	(411.34, 44.20) --
	(411.56, 44.20) --
	(411.77, 44.20) --
	(411.99, 44.20) --
	(412.21, 44.21) --
	(412.42, 44.21) --
	(412.64, 44.21) --
	(412.85, 44.21) --
	(413.07, 44.21) --
	(413.28, 44.21) --
	(413.50, 44.21) --
	(413.72, 44.21) --
	(413.93, 44.21) --
	(414.15, 44.21) --
	(414.36, 44.21) --
	(414.58, 44.21) --
	(414.79, 44.21) --
	(415.01, 44.21) --
	(415.23, 44.21) --
	(415.44, 44.21) --
	(415.66, 44.21) --
	(415.87, 44.21) --
	(416.09, 44.21) --
	(416.30, 44.21) --
	(416.52, 44.21) --
	(416.74, 44.21) --
	(416.95, 44.21) --
	(417.17, 44.21) --
	(417.38, 44.21) --
	(417.60, 44.21) --
	(417.81, 44.21) --
	(418.03, 44.21) --
	(418.25, 44.21) --
	(418.46, 44.21) --
	(418.68, 44.21) --
	(418.89, 44.21) --
	(419.11, 44.21) --
	(419.32, 44.21) --
	(419.54, 44.21) --
	(419.76, 44.21) --
	(419.97, 44.21) --
	(420.19, 44.21) --
	(420.40, 44.21) --
	(420.62, 44.21) --
	(420.84, 44.21) --
	(421.05, 44.21) --
	(421.27, 44.21) --
	(421.48, 44.21) --
	(421.70, 44.21) --
	(421.91, 44.21) --
	(422.13, 44.21) --
	(422.35, 44.21) --
	(422.56, 44.21) --
	(422.78, 44.21) --
	(422.99, 44.21) --
	(423.21, 44.21) --
	(423.42, 44.21) --
	(423.64, 44.21) --
	(423.86, 44.21) --
	(424.07, 44.21) --
	(424.29, 44.21) --
	(424.50, 44.21) --
	(424.72, 44.21) --
	(424.93, 44.21) --
	(425.15, 44.21) --
	(425.37, 44.21) --
	(425.58, 44.21) --
	(425.80, 44.21) --
	(426.01, 44.21) --
	(426.23, 44.21) --
	(426.44, 44.20) --
	(426.66, 44.20) --
	(426.88, 44.20) --
	(427.09, 44.20) --
	(427.31, 44.20) --
	(427.52, 44.19) --
	(427.74, 44.19) --
	(427.95, 44.19) --
	(428.17, 44.19) --
	(428.39, 44.19) --
	(428.60, 44.19) --
	(428.82, 44.18) --
	(429.03, 44.18) --
	(429.25, 44.18) --
	(429.46, 44.18) --
	(429.68, 44.18) --
	(429.90, 44.19) --
	(430.11, 44.19) --
	(430.33, 44.19) --
	(430.54, 44.19) --
	(430.76, 44.19) --
	(430.97, 44.20) --
	(431.19, 44.20) --
	(431.41, 44.20) --
	(431.62, 44.20) --
	(431.84, 44.20) --
	(432.05, 44.21) --
	(432.27, 44.21) --
	(432.48, 44.21) --
	(432.70, 44.21) --
	(432.92, 44.21) --
	(433.13, 44.21) --
	(433.35, 44.21) --
	(433.56, 44.21) --
	(433.78, 44.21) --
	(433.99, 44.21) --
	(434.21, 44.21) --
	(434.43, 44.21) --
	(434.64, 44.21) --
	(434.86, 44.21) --
	(435.07, 44.21) --
	(435.29, 44.20) --
	(435.51, 44.20) --
	(435.72, 44.20) --
	(435.94, 44.20) --
	(436.15, 44.20) --
	(436.37, 44.20) --
	(436.58, 44.19) --
	(436.80, 44.19) --
	(437.02, 44.19) --
	(437.23, 44.19) --
	(437.45, 44.19) --
	(437.66, 44.18) --
	(437.88, 44.18) --
	(438.09, 44.18) --
	(438.09, 44.24) --
	(437.88, 44.24) --
	(437.66, 44.24) --
	(437.45, 44.24) --
	(437.23, 44.24) --
	(437.02, 44.23) --
	(436.80, 44.23) --
	(436.58, 44.23) --
	(436.37, 44.23) --
	(436.15, 44.22) --
	(435.94, 44.22) --
	(435.72, 44.22) --
	(435.51, 44.22) --
	(435.29, 44.22) --
	(435.07, 44.22) --
	(434.86, 44.22) --
	(434.64, 44.21) --
	(434.43, 44.21) --
	(434.21, 44.21) --
	(433.99, 44.21) --
	(433.78, 44.21) --
	(433.56, 44.21) --
	(433.35, 44.21) --
	(433.13, 44.21) --
	(432.92, 44.21) --
	(432.70, 44.21) --
	(432.48, 44.21) --
	(432.27, 44.22) --
	(432.05, 44.22) --
	(431.84, 44.22) --
	(431.62, 44.22) --
	(431.41, 44.22) --
	(431.19, 44.22) --
	(430.97, 44.23) --
	(430.76, 44.23) --
	(430.54, 44.23) --
	(430.33, 44.23) --
	(430.11, 44.23) --
	(429.90, 44.24) --
	(429.68, 44.24) --
	(429.46, 44.24) --
	(429.25, 44.24) --
	(429.03, 44.24) --
	(428.82, 44.24) --
	(428.60, 44.24) --
	(428.39, 44.24) --
	(428.17, 44.23) --
	(427.95, 44.23) --
	(427.74, 44.23) --
	(427.52, 44.23) --
	(427.31, 44.23) --
	(427.09, 44.22) --
	(426.88, 44.22) --
	(426.66, 44.22) --
	(426.44, 44.22) --
	(426.23, 44.22) --
	(426.01, 44.22) --
	(425.80, 44.21) --
	(425.58, 44.21) --
	(425.37, 44.21) --
	(425.15, 44.21) --
	(424.93, 44.21) --
	(424.72, 44.21) --
	(424.50, 44.21) --
	(424.29, 44.21) --
	(424.07, 44.21) --
	(423.86, 44.21) --
	(423.64, 44.21) --
	(423.42, 44.21) --
	(423.21, 44.21) --
	(422.99, 44.21) --
	(422.78, 44.21) --
	(422.56, 44.21) --
	(422.35, 44.21) --
	(422.13, 44.21) --
	(421.91, 44.21) --
	(421.70, 44.21) --
	(421.48, 44.21) --
	(421.27, 44.21) --
	(421.05, 44.21) --
	(420.84, 44.21) --
	(420.62, 44.21) --
	(420.40, 44.21) --
	(420.19, 44.21) --
	(419.97, 44.21) --
	(419.76, 44.21) --
	(419.54, 44.21) --
	(419.32, 44.21) --
	(419.11, 44.21) --
	(418.89, 44.21) --
	(418.68, 44.21) --
	(418.46, 44.21) --
	(418.25, 44.21) --
	(418.03, 44.21) --
	(417.81, 44.21) --
	(417.60, 44.21) --
	(417.38, 44.21) --
	(417.17, 44.21) --
	(416.95, 44.21) --
	(416.74, 44.21) --
	(416.52, 44.21) --
	(416.30, 44.21) --
	(416.09, 44.21) --
	(415.87, 44.21) --
	(415.66, 44.21) --
	(415.44, 44.21) --
	(415.23, 44.21) --
	(415.01, 44.21) --
	(414.79, 44.21) --
	(414.58, 44.21) --
	(414.36, 44.21) --
	(414.15, 44.21) --
	(413.93, 44.21) --
	(413.72, 44.21) --
	(413.50, 44.21) --
	(413.28, 44.21) --
	(413.07, 44.21) --
	(412.85, 44.21) --
	(412.64, 44.21) --
	(412.42, 44.22) --
	(412.21, 44.22) --
	(411.99, 44.22) --
	(411.77, 44.22) --
	(411.56, 44.22) --
	(411.34, 44.22) --
	(411.13, 44.23) --
	(410.91, 44.23) --
	(410.70, 44.23) --
	(410.48, 44.24) --
	(410.26, 44.24) --
	(410.05, 44.24) --
	(409.83, 44.25) --
	(409.62, 44.25) --
	(409.40, 44.26) --
	(409.19, 44.26) --
	(408.97, 44.26) --
	(408.75, 44.27) --
	(408.54, 44.27) --
	(408.32, 44.27) --
	(408.11, 44.28) --
	(407.89, 44.28) --
	(407.68, 44.28) --
	(407.46, 44.29) --
	(407.24, 44.29) --
	(407.03, 44.29) --
	(406.81, 44.29) --
	(406.60, 44.30) --
	(406.38, 44.30) --
	(406.16, 44.30) --
	(405.95, 44.31) --
	(405.73, 44.31) --
	(405.52, 44.32) --
	(405.30, 44.32) --
	(405.09, 44.33) --
	(404.87, 44.34) --
	(404.65, 44.34) --
	(404.44, 44.35) --
	(404.22, 44.36) --
	(404.01, 44.36) --
	(403.79, 44.37) --
	(403.58, 44.38) --
	(403.36, 44.38) --
	(403.14, 44.39) --
	(402.93, 44.39) --
	(402.71, 44.39) --
	(402.50, 44.39) --
	(402.28, 44.39) --
	(402.07, 44.38) --
	(401.85, 44.38) --
	(401.63, 44.38) --
	(401.42, 44.37) --
	(401.20, 44.37) --
	(400.99, 44.37) --
	(400.77, 44.36) --
	(400.56, 44.36) --
	(400.34, 44.36) --
	(400.12, 44.36) --
	(399.91, 44.36) --
	(399.69, 44.36) --
	(399.48, 44.37) --
	(399.26, 44.37) --
	(399.05, 44.38) --
	(398.83, 44.38) --
	(398.61, 44.39) --
	(398.40, 44.40) --
	(398.18, 44.41) --
	(397.97, 44.42) --
	(397.75, 44.43) --
	(397.54, 44.43) --
	(397.32, 44.44) --
	(397.10, 44.45) --
	(396.89, 44.46) --
	(396.67, 44.47) --
	(396.46, 44.48) --
	(396.24, 44.48) --
	(396.03, 44.49) --
	(395.81, 44.50) --
	(395.59, 44.50) --
	(395.38, 44.51) --
	(395.16, 44.51) --
	(394.95, 44.52) --
	(394.73, 44.52) --
	(394.52, 44.52) --
	(394.30, 44.53) --
	(394.08, 44.53) --
	(393.87, 44.53) --
	(393.65, 44.54) --
	(393.44, 44.54) --
	(393.22, 44.55) --
	(393.01, 44.56) --
	(392.79, 44.57) --
	(392.57, 44.58) --
	(392.36, 44.59) --
	(392.14, 44.60) --
	(391.93, 44.62) --
	(391.71, 44.64) --
	(391.49, 44.66) --
	(391.28, 44.68) --
	(391.06, 44.70) --
	(390.85, 44.72) --
	(390.63, 44.74) --
	(390.42, 44.76) --
	(390.20, 44.78) --
	(389.98, 44.80) --
	(389.77, 44.82) --
	(389.55, 44.84) --
	(389.34, 44.86) --
	(389.12, 44.88) --
	(388.91, 44.90) --
	(388.69, 44.92) --
	(388.47, 44.94) --
	(388.26, 44.96) --
	(388.04, 44.99) --
	(387.83, 45.02) --
	(387.61, 45.05) --
	(387.40, 45.09) --
	(387.18, 45.12) --
	(386.96, 45.16) --
	(386.75, 45.21) --
	(386.53, 45.25) --
	(386.32, 45.30) --
	(386.10, 45.35) --
	(385.89, 45.40) --
	(385.67, 45.45) --
	(385.45, 45.50) --
	(385.24, 45.55) --
	(385.02, 45.60) --
	(384.81, 45.65) --
	(384.59, 45.70) --
	(384.38, 45.75) --
	(384.16, 45.80) --
	(383.94, 45.85) --
	(383.73, 45.90) --
	(383.51, 45.95) --
	(383.30, 46.01) --
	(383.08, 46.06) --
	(382.87, 46.12) --
	(382.65, 46.18) --
	(382.43, 46.24) --
	(382.22, 46.30) --
	(382.00, 46.36) --
	(381.79, 46.42) --
	(381.57, 46.47) --
	(381.36, 46.53) --
	(381.14, 46.59) --
	(380.92, 46.65) --
	(380.71, 46.70) --
	(380.49, 46.76) --
	(380.28, 46.82) --
	(380.06, 46.87) --
	(379.85, 46.94) --
	(379.63, 47.01) --
	(379.41, 47.08) --
	(379.20, 47.16) --
	(378.98, 47.25) --
	(378.77, 47.35) --
	(378.55, 47.46) --
	(378.33, 47.58) --
	(378.12, 47.71) --
	(377.90, 47.85) --
	(377.69, 48.00) --
	(377.47, 48.16) --
	(377.26, 48.33) --
	(377.04, 48.50) --
	(376.82, 48.68) --
	(376.61, 48.85) --
	(376.39, 49.03) --
	(376.18, 49.21) --
	(375.96, 49.38) --
	(375.75, 49.55) --
	(375.53, 49.72) --
	(375.31, 49.88) --
	(375.10, 50.04) --
	(374.88, 50.20) --
	(374.67, 50.35) --
	(374.45, 50.51) --
	(374.24, 50.66) --
	(374.02, 50.82) --
	(373.80, 50.98) --
	(373.59, 51.14) --
	(373.37, 51.30) --
	(373.16, 51.47) --
	(372.94, 51.63) --
	(372.73, 51.79) --
	(372.51, 51.95) --
	(372.29, 52.10) --
	(372.08, 52.25) --
	(371.86, 52.40) --
	(371.65, 52.54) --
	(371.43, 52.68) --
	(371.22, 52.81) --
	(371.00, 52.94) --
	(370.78, 53.06) --
	(370.57, 53.19) --
	(370.35, 53.31) --
	(370.14, 53.43) --
	(369.92, 53.55) --
	(369.71, 53.66) --
	(369.49, 53.78) --
	(369.27, 53.89) --
	(369.06, 54.01) --
	(368.84, 54.12) --
	(368.63, 54.22) --
	(368.41, 54.33) --
	(368.20, 54.43) --
	(367.98, 54.53) --
	(367.76, 54.62) --
	(367.55, 54.71) --
	(367.33, 54.79) --
	(367.12, 54.87) --
	(366.90, 54.95) --
	(366.69, 55.01) --
	(366.47, 55.07) --
	(366.25, 55.11) --
	(366.04, 55.15) --
	(365.82, 55.18) --
	(365.61, 55.20) --
	(365.39, 55.20) --
	(365.18, 55.20) --
	(364.96, 55.18) --
	(364.74, 55.16) --
	(364.53, 55.11) --
	(364.31, 55.06) --
	(364.10, 54.99) --
	(363.88, 54.91) --
	(363.66, 54.82) --
	(363.45, 54.72) --
	(363.23, 54.60) --
	(363.02, 54.48) --
	(362.80, 54.34) --
	(362.59, 54.19) --
	(362.37, 54.04) --
	(362.15, 53.87) --
	(361.94, 53.71) --
	(361.72, 53.53) --
	(361.51, 53.35) --
	(361.29, 53.17) --
	(361.08, 52.99) --
	(360.86, 52.80) --
	(360.64, 52.62) --
	(360.43, 52.43) --
	(360.21, 52.25) --
	(360.00, 52.06) --
	(359.78, 51.88) --
	(359.57, 51.70) --
	(359.35, 51.52) --
	(359.13, 51.35) --
	(358.92, 51.17) --
	(358.70, 50.99) --
	(358.49, 50.81) --
	(358.27, 50.63) --
	(358.06, 50.44) --
	(357.84, 50.26) --
	(357.62, 50.06) --
	(357.41, 49.87) --
	(357.19, 49.67) --
	(356.98, 49.46) --
	(356.76, 49.26) --
	(356.55, 49.05) --
	(356.33, 48.85) --
	(356.11, 48.65) --
	(355.90, 48.46) --
	(355.68, 48.27) --
	(355.47, 48.10) --
	(355.25, 47.93) --
	(355.04, 47.77) --
	(354.82, 47.63) --
	(354.60, 47.49) --
	(354.39, 47.37) --
	(354.17, 47.27) --
	(353.96, 47.17) --
	(353.74, 47.08) --
	(353.53, 47.00) --
	(353.31, 46.92) --
	(353.09, 46.85) --
	(352.88, 46.78) --
	(352.66, 46.72) --
	(352.45, 46.65) --
	(352.23, 46.58) --
	(352.02, 46.51) --
	(351.80, 46.44) --
	(351.58, 46.37) --
	(351.37, 46.30) --
	(351.15, 46.23) --
	(350.94, 46.15) --
	(350.72, 46.08) --
	(350.50, 46.01) --
	(350.29, 45.93) --
	(350.07, 45.86) --
	(349.86, 45.79) --
	(349.64, 45.72) --
	(349.43, 45.66) --
	(349.21, 45.59) --
	(348.99, 45.53) --
	(348.78, 45.47) --
	(348.56, 45.41) --
	(348.35, 45.35) --
	(348.13, 45.30) --
	(347.92, 45.25) --
	(347.70, 45.21) --
	(347.48, 45.16) --
	(347.27, 45.12) --
	(347.05, 45.09) --
	(346.84, 45.05) --
	(346.62, 45.02) --
	(346.41, 44.99) --
	(346.19, 44.96) --
	(345.97, 44.93) --
	(345.76, 44.91) --
	(345.54, 44.88) --
	(345.33, 44.85) --
	(345.11, 44.83) --
	(344.90, 44.80) --
	(344.68, 44.77) --
	(344.46, 44.74) --
	(344.25, 44.71) --
	(344.03, 44.68) --
	(343.82, 44.64) --
	(343.60, 44.61) --
	(343.39, 44.58) --
	(343.17, 44.54) --
	(342.95, 44.51) --
	(342.74, 44.48) --
	(342.52, 44.45) --
	(342.31, 44.43) --
	(342.09, 44.41) --
	(341.88, 44.39) --
	(341.66, 44.37) --
	(341.44, 44.36) --
	(341.23, 44.35) --
	(341.01, 44.34) --
	(340.80, 44.34) --
	(340.58, 44.33) --
	(340.37, 44.33) --
	(340.15, 44.34) --
	(339.93, 44.34) --
	(339.72, 44.34) --
	(339.50, 44.34) --
	(339.29, 44.35) --
	(339.07, 44.35) --
	(338.86, 44.36) --
	(338.64, 44.36) --
	(338.42, 44.36) --
	(338.21, 44.37) --
	(337.99, 44.37) --
	(337.78, 44.37) --
	(337.56, 44.37) --
	(337.34, 44.38) --
	(337.13, 44.38) --
	(336.91, 44.38) --
	(336.70, 44.38) --
	(336.48, 44.38) --
	(336.27, 44.37) --
	(336.05, 44.37) --
	(335.83, 44.37) --
	(335.62, 44.36) --
	(335.40, 44.36) --
	(335.19, 44.35) --
	(334.97, 44.35) --
	(334.76, 44.34) --
	(334.54, 44.33) --
	(334.32, 44.33) --
	(334.11, 44.33) --
	(333.89, 44.32) --
	(333.68, 44.32) --
	(333.46, 44.31) --
	(333.25, 44.31) --
	(333.03, 44.31) --
	(332.81, 44.31) --
	(332.60, 44.30) --
	(332.38, 44.30) --
	(332.17, 44.30) --
	(331.95, 44.29) --
	(331.74, 44.29) --
	(331.52, 44.28) --
	(331.30, 44.28) --
	(331.09, 44.27) --
	(330.87, 44.27) --
	(330.66, 44.27) --
	(330.44, 44.26) --
	(330.23, 44.26) --
	(330.01, 44.26) --
	(329.79, 44.25) --
	(329.58, 44.25) --
	(329.36, 44.25) --
	(329.15, 44.25) --
	(328.93, 44.25) --
	(328.72, 44.25) --
	(328.50, 44.24) --
	(328.28, 44.24) --
	(328.07, 44.24) --
	(327.85, 44.24) --
	(327.85, 44.18) --
	cycle;
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (312.62, 31.02) rectangle (476.30,211.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67, 41.18) {Pooled estimate};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67, 63.17) {Study 1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67, 85.15) {Study 2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,107.14) {Study 3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,129.13) {Study 4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,151.11) {Study 5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,173.10) {Study 6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (307.67,195.09) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87, 44.21) --
	(312.62, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87, 66.20) --
	(312.62, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87, 88.18) --
	(312.62, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,110.17) --
	(312.62,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,132.16) --
	(312.62,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,154.14) --
	(312.62,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,176.13) --
	(312.62,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (309.87,198.12) --
	(312.62,198.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (318.20, 28.27) --
	(318.20, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.71, 28.27) --
	(356.71, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (395.21, 28.27) --
	(395.21, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.72, 28.27) --
	(433.72, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.23, 28.27) --
	(472.23, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.20, 20.01) {6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (356.71, 20.01) {8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (395.21, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (433.72, 20.01) {12};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (472.23, 20.01) {14};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (394.46,  7.82) {Mean generation interval (days)};
\end{scope}
\begin{scope}
\path[clip] (481.80,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (481.80,  0.00) rectangle (722.70,216.81);
\end{scope}
\begin{scope}
\path[clip] (553.52, 31.02) rectangle (717.20,211.31);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (553.52, 31.02) rectangle (717.20,211.31);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (579.31, 31.02) --
	(579.31,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (616.03, 31.02) --
	(616.03,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (652.74, 31.02) --
	(652.74,211.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (689.45, 31.02) --
	(689.45,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52, 44.21) --
	(717.20, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52, 66.20) --
	(717.20, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52, 88.18) --
	(717.20, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,110.17) --
	(717.20,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,132.16) --
	(717.20,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,154.14) --
	(717.20,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,176.13) --
	(717.20,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.52,198.12) --
	(717.20,198.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (560.96, 31.02) --
	(560.96,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.67, 31.02) --
	(597.67,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.38, 31.02) --
	(634.38,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (671.09, 31.02) --
	(671.09,211.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (707.80, 31.02) --
	(707.80,211.31);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (634.38, 66.20) circle (  2.50);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (597.67, 92.07) --
	(601.03, 86.24) --
	(594.31, 86.24) --
	(597.67, 92.07);

\path[fill=fillColor] (575.98,110.17) circle (  2.50);

\path[fill=fillColor] (560.96,198.12) circle (  2.50);

\path[fill=fillColor] (597.67,132.16) circle (  2.50);

\path[fill=fillColor] (597.67,154.14) circle (  2.50);

\path[fill=fillColor] (575.64,176.13) circle (  2.50);

\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) -- (634.38, 66.20);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) -- (597.67, 88.18);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) -- (575.98,110.17);

\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) -- (560.96,198.12);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) -- (597.67,132.16);

\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) -- (597.67,154.14);

\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) -- (575.64,176.13);
\definecolor{drawColor}{RGB}{255,0,0}
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (571.29, 44.01) --
	(571.56, 43.99) --
	(571.83, 43.96) --
	(572.10, 43.92) --
	(572.37, 43.89) --
	(572.65, 43.85) --
	(572.92, 43.80) --
	(573.19, 43.76) --
	(573.46, 43.71) --
	(573.73, 43.65) --
	(574.00, 43.59) --
	(574.27, 43.53) --
	(574.54, 43.47) --
	(574.81, 43.40) --
	(575.08, 43.32) --
	(575.36, 43.24) --
	(575.63, 43.16) --
	(575.90, 43.07) --
	(576.17, 42.98) --
	(576.44, 42.88) --
	(576.71, 42.78) --
	(576.98, 42.67) --
	(577.25, 42.56) --
	(577.52, 42.44) --
	(577.79, 42.31) --
	(578.07, 42.18) --
	(578.34, 42.03) --
	(578.61, 41.89) --
	(578.88, 41.73) --
	(579.15, 41.56) --
	(579.42, 41.39) --
	(579.69, 41.20) --
	(579.96, 41.01) --
	(580.23, 40.81) --
	(580.50, 40.60) --
	(580.77, 40.38) --
	(581.05, 40.15) --
	(581.32, 39.92) --
	(581.59, 39.68) --
	(581.86, 39.44) --
	(582.13, 39.19) --
	(582.40, 38.94) --
	(582.67, 38.68) --
	(582.94, 38.43) --
	(583.21, 38.17) --
	(583.48, 37.92) --
	(583.76, 37.66) --
	(584.03, 37.42) --
	(584.30, 37.17) --
	(584.57, 36.93) --
	(584.84, 36.70) --
	(585.11, 36.47) --
	(585.38, 36.25) --
	(585.65, 36.04) --
	(585.92, 35.84) --
	(586.19, 35.64) --
	(586.47, 35.45) --
	(586.74, 35.27) --
	(587.01, 35.09) --
	(587.28, 34.93) --
	(587.55, 34.77) --
	(587.82, 34.61) --
	(588.09, 34.46) --
	(588.36, 34.32) --
	(588.63, 34.19) --
	(588.90, 34.07) --
	(589.18, 33.95) --
	(589.45, 33.84) --
	(589.72, 33.74) --
	(589.99, 33.64) --
	(590.26, 33.56) --
	(590.53, 33.48) --
	(590.80, 33.41) --
	(591.07, 33.36) --
	(591.34, 33.31) --
	(591.61, 33.27) --
	(591.89, 33.24) --
	(592.16, 33.23) --
	(592.43, 33.22) --
	(592.70, 33.22) --
	(592.97, 33.23) --
	(593.24, 33.25) --
	(593.51, 33.28) --
	(593.78, 33.31) --
	(594.05, 33.36) --
	(594.32, 33.41) --
	(594.59, 33.47) --
	(594.87, 33.54) --
	(595.14, 33.61) --
	(595.41, 33.69) --
	(595.68, 33.78) --
	(595.95, 33.88) --
	(596.22, 33.98) --
	(596.49, 34.09) --
	(596.76, 34.20) --
	(597.03, 34.32) --
	(597.30, 34.45) --
	(597.58, 34.57) --
	(597.85, 34.70) --
	(598.12, 34.84) --
	(598.39, 34.97) --
	(598.66, 35.10) --
	(598.93, 35.24) --
	(599.20, 35.37) --
	(599.47, 35.49) --
	(599.74, 35.62) --
	(600.01, 35.74) --
	(600.29, 35.85) --
	(600.56, 35.96) --
	(600.83, 36.06) --
	(601.10, 36.16) --
	(601.37, 36.25) --
	(601.64, 36.34) --
	(601.91, 36.42) --
	(602.18, 36.50) --
	(602.45, 36.58) --
	(602.72, 36.66) --
	(603.00, 36.74) --
	(603.27, 36.81) --
	(603.54, 36.89) --
	(603.81, 36.97) --
	(604.08, 37.05) --
	(604.35, 37.13) --
	(604.62, 37.22) --
	(604.89, 37.30) --
	(605.16, 37.39) --
	(605.43, 37.49) --
	(605.70, 37.58) --
	(605.98, 37.67) --
	(606.25, 37.77) --
	(606.52, 37.87) --
	(606.79, 37.96) --
	(607.06, 38.05) --
	(607.33, 38.14) --
	(607.60, 38.23) --
	(607.87, 38.32) --
	(608.14, 38.40) --
	(608.41, 38.47) --
	(608.69, 38.55) --
	(608.96, 38.62) --
	(609.23, 38.68) --
	(609.50, 38.74) --
	(609.77, 38.80) --
	(610.04, 38.85) --
	(610.31, 38.90) --
	(610.58, 38.95) --
	(610.85, 39.00) --
	(611.12, 39.04) --
	(611.40, 39.09) --
	(611.67, 39.13) --
	(611.94, 39.17) --
	(612.21, 39.22) --
	(612.48, 39.27) --
	(612.75, 39.32) --
	(613.02, 39.37) --
	(613.29, 39.42) --
	(613.56, 39.48) --
	(613.83, 39.54) --
	(614.11, 39.60) --
	(614.38, 39.67) --
	(614.65, 39.74) --
	(614.92, 39.82) --
	(615.19, 39.90) --
	(615.46, 39.98) --
	(615.73, 40.07) --
	(616.00, 40.16) --
	(616.27, 40.25) --
	(616.54, 40.34) --
	(616.81, 40.43) --
	(617.09, 40.53) --
	(617.36, 40.62) --
	(617.63, 40.71) --
	(617.90, 40.80) --
	(618.17, 40.89) --
	(618.44, 40.98) --
	(618.71, 41.07) --
	(618.98, 41.15) --
	(619.25, 41.23) --
	(619.52, 41.31) --
	(619.80, 41.39) --
	(620.07, 41.46) --
	(620.34, 41.53) --
	(620.61, 41.60) --
	(620.88, 41.67) --
	(621.15, 41.73) --
	(621.42, 41.79) --
	(621.69, 41.85) --
	(621.96, 41.91) --
	(622.23, 41.97) --
	(622.51, 42.02) --
	(622.78, 42.08) --
	(623.05, 42.13) --
	(623.32, 42.17) --
	(623.59, 42.22) --
	(623.86, 42.27) --
	(624.13, 42.31) --
	(624.40, 42.35) --
	(624.67, 42.39) --
	(624.94, 42.42) --
	(625.22, 42.46) --
	(625.49, 42.49) --
	(625.76, 42.52) --
	(626.03, 42.54) --
	(626.30, 42.57) --
	(626.57, 42.59) --
	(626.84, 42.62) --
	(627.11, 42.64) --
	(627.38, 42.66) --
	(627.65, 42.68) --
	(627.92, 42.70) --
	(628.20, 42.71) --
	(628.47, 42.73) --
	(628.74, 42.74) --
	(629.01, 42.76) --
	(629.28, 42.77) --
	(629.55, 42.79) --
	(629.82, 42.80) --
	(630.09, 42.81) --
	(630.36, 42.83) --
	(630.63, 42.84) --
	(630.91, 42.85) --
	(631.18, 42.87) --
	(631.45, 42.88) --
	(631.72, 42.90) --
	(631.99, 42.92) --
	(632.26, 42.93) --
	(632.53, 42.95) --
	(632.80, 42.97) --
	(633.07, 42.99) --
	(633.34, 43.01) --
	(633.62, 43.04) --
	(633.89, 43.06) --
	(634.16, 43.08) --
	(634.43, 43.11) --
	(634.70, 43.13) --
	(634.97, 43.16) --
	(635.24, 43.18) --
	(635.51, 43.21) --
	(635.78, 43.23) --
	(636.05, 43.26) --
	(636.33, 43.28) --
	(636.60, 43.30) --
	(636.87, 43.32) --
	(637.14, 43.34) --
	(637.41, 43.36) --
	(637.68, 43.38) --
	(637.95, 43.39) --
	(638.22, 43.41) --
	(638.49, 43.42) --
	(638.76, 43.43) --
	(639.03, 43.44) --
	(639.31, 43.45) --
	(639.58, 43.46) --
	(639.85, 43.47) --
	(640.12, 43.48) --
	(640.39, 43.49) --
	(640.66, 43.50) --
	(640.93, 43.50) --
	(641.20, 43.51) --
	(641.47, 43.52) --
	(641.74, 43.53) --
	(642.02, 43.53) --
	(642.29, 43.54) --
	(642.56, 43.55) --
	(642.83, 43.56) --
	(643.10, 43.57) --
	(643.37, 43.58) --
	(643.64, 43.60) --
	(643.91, 43.61) --
	(644.18, 43.62) --
	(644.45, 43.64) --
	(644.73, 43.65) --
	(645.00, 43.67) --
	(645.27, 43.68) --
	(645.54, 43.70) --
	(645.81, 43.72) --
	(646.08, 43.73) --
	(646.35, 43.75) --
	(646.62, 43.76) --
	(646.89, 43.78) --
	(647.16, 43.80) --
	(647.44, 43.81) --
	(647.71, 43.83) --
	(647.98, 43.84) --
	(648.25, 43.86) --
	(648.52, 43.87) --
	(648.79, 43.89) --
	(649.06, 43.90) --
	(649.33, 43.91) --
	(649.60, 43.93) --
	(649.87, 43.94) --
	(650.15, 43.95) --
	(650.42, 43.96) --
	(650.69, 43.97) --
	(650.96, 43.98) --
	(651.23, 43.98) --
	(651.50, 43.99) --
	(651.77, 43.99) --
	(652.04, 44.00) --
	(652.31, 44.00) --
	(652.58, 44.01) --
	(652.85, 44.01) --
	(653.13, 44.01) --
	(653.40, 44.01) --
	(653.67, 44.01) --
	(653.94, 44.01) --
	(654.21, 44.01) --
	(654.48, 44.01) --
	(654.75, 44.01) --
	(655.02, 44.02) --
	(655.29, 44.02) --
	(655.56, 44.02) --
	(655.84, 44.02) --
	(656.11, 44.02) --
	(656.38, 44.03) --
	(656.65, 44.03) --
	(656.92, 44.03) --
	(657.19, 44.04) --
	(657.46, 44.04) --
	(657.73, 44.05) --
	(658.00, 44.05) --
	(658.27, 44.06) --
	(658.55, 44.07) --
	(658.82, 44.07) --
	(659.09, 44.08) --
	(659.36, 44.08) --
	(659.63, 44.09) --
	(659.90, 44.09) --
	(660.17, 44.10) --
	(660.44, 44.10) --
	(660.71, 44.11) --
	(660.98, 44.11) --
	(661.26, 44.12) --
	(661.53, 44.12) --
	(661.80, 44.13) --
	(662.07, 44.13) --
	(662.34, 44.13) --
	(662.61, 44.14) --
	(662.88, 44.14) --
	(663.15, 44.14) --
	(663.42, 44.15) --
	(663.69, 44.15) --
	(663.96, 44.15) --
	(664.24, 44.16) --
	(664.51, 44.16) --
	(664.78, 44.16) --
	(665.05, 44.16) --
	(665.32, 44.16) --
	(665.59, 44.16) --
	(665.86, 44.17) --
	(666.13, 44.17) --
	(666.40, 44.17) --
	(666.67, 44.17) --
	(666.95, 44.16) --
	(667.22, 44.16) --
	(667.49, 44.16) --
	(667.76, 44.16) --
	(668.03, 44.16) --
	(668.30, 44.16) --
	(668.57, 44.16) --
	(668.84, 44.15) --
	(669.11, 44.15) --
	(669.38, 44.15) --
	(669.66, 44.15) --
	(669.93, 44.15) --
	(670.20, 44.15) --
	(670.47, 44.15) --
	(670.74, 44.14) --
	(671.01, 44.14) --
	(671.28, 44.14) --
	(671.55, 44.14) --
	(671.82, 44.15) --
	(672.09, 44.15) --
	(672.37, 44.15) --
	(672.64, 44.15) --
	(672.91, 44.15) --
	(673.18, 44.15) --
	(673.45, 44.15) --
	(673.72, 44.15) --
	(673.99, 44.15) --
	(674.26, 44.15) --
	(674.53, 44.15) --
	(674.80, 44.15) --
	(675.07, 44.15) --
	(675.35, 44.15) --
	(675.62, 44.15) --
	(675.89, 44.15) --
	(676.16, 44.15) --
	(676.43, 44.15) --
	(676.70, 44.15) --
	(676.97, 44.15) --
	(677.24, 44.15) --
	(677.51, 44.15) --
	(677.78, 44.15) --
	(678.06, 44.15) --
	(678.33, 44.15) --
	(678.60, 44.15) --
	(678.87, 44.14) --
	(679.14, 44.14) --
	(679.41, 44.14) --
	(679.68, 44.14) --
	(679.95, 44.14) --
	(680.22, 44.14) --
	(680.49, 44.14) --
	(680.77, 44.14) --
	(681.04, 44.14) --
	(681.31, 44.14) --
	(681.58, 44.14) --
	(681.85, 44.14) --
	(682.12, 44.14) --
	(682.39, 44.14) --
	(682.66, 44.14) --
	(682.93, 44.14) --
	(683.20, 44.13) --
	(683.48, 44.13) --
	(683.75, 44.13) --
	(684.02, 44.13) --
	(684.29, 44.13) --
	(684.56, 44.13) --
	(684.83, 44.13) --
	(685.10, 44.13) --
	(685.37, 44.13) --
	(685.64, 44.13) --
	(685.91, 44.13) --
	(686.18, 44.13) --
	(686.46, 44.14) --
	(686.73, 44.14) --
	(687.00, 44.14) --
	(687.27, 44.15) --
	(687.54, 44.15) --
	(687.81, 44.15) --
	(688.08, 44.16) --
	(688.35, 44.16) --
	(688.62, 44.16) --
	(688.89, 44.17) --
	(689.17, 44.17) --
	(689.44, 44.17) --
	(689.71, 44.18) --
	(689.98, 44.18) --
	(690.25, 44.18) --
	(690.52, 44.18) --
	(690.79, 44.18) --
	(691.06, 44.19) --
	(691.33, 44.19) --
	(691.60, 44.19) --
	(691.88, 44.19) --
	(692.15, 44.19) --
	(692.42, 44.19) --
	(692.69, 44.19) --
	(692.96, 44.19) --
	(693.23, 44.19) --
	(693.50, 44.19) --
	(693.77, 44.19) --
	(694.04, 44.19) --
	(694.31, 44.19) --
	(694.59, 44.19) --
	(694.86, 44.19) --
	(695.13, 44.19) --
	(695.40, 44.19) --
	(695.67, 44.19) --
	(695.94, 44.19) --
	(696.21, 44.19) --
	(696.48, 44.19) --
	(696.75, 44.20) --
	(697.02, 44.20) --
	(697.29, 44.20) --
	(697.57, 44.20) --
	(697.84, 44.20) --
	(698.11, 44.20) --
	(698.38, 44.20) --
	(698.65, 44.20) --
	(698.92, 44.20) --
	(699.19, 44.20) --
	(699.46, 44.20) --
	(699.73, 44.20) --
	(700.00, 44.20) --
	(700.28, 44.20) --
	(700.55, 44.20) --
	(700.82, 44.20) --
	(701.09, 44.20) --
	(701.36, 44.20) --
	(701.63, 44.20) --
	(701.90, 44.20) --
	(702.17, 44.20) --
	(702.44, 44.19) --
	(702.71, 44.19) --
	(702.99, 44.19) --
	(703.26, 44.19) --
	(703.53, 44.19) --
	(703.80, 44.19) --
	(704.07, 44.19) --
	(704.34, 44.18) --
	(704.61, 44.18) --
	(704.88, 44.18) --
	(705.15, 44.18) --
	(705.42, 44.18) --
	(705.70, 44.18) --
	(705.97, 44.18) --
	(706.24, 44.18) --
	(706.51, 44.18) --
	(706.78, 44.18) --
	(707.05, 44.18) --
	(707.32, 44.18) --
	(707.59, 44.18) --
	(707.86, 44.18) --
	(708.13, 44.18) --
	(708.41, 44.18) --
	(708.68, 44.18) --
	(708.95, 44.18) --
	(709.22, 44.18) --
	(709.49, 44.18) --
	(709.76, 44.18) --
	(709.76, 44.24) --
	(709.49, 44.24) --
	(709.22, 44.24) --
	(708.95, 44.24) --
	(708.68, 44.24) --
	(708.41, 44.24) --
	(708.13, 44.24) --
	(707.86, 44.24) --
	(707.59, 44.24) --
	(707.32, 44.24) --
	(707.05, 44.24) --
	(706.78, 44.24) --
	(706.51, 44.24) --
	(706.24, 44.24) --
	(705.97, 44.24) --
	(705.70, 44.24) --
	(705.42, 44.24) --
	(705.15, 44.24) --
	(704.88, 44.24) --
	(704.61, 44.24) --
	(704.34, 44.24) --
	(704.07, 44.24) --
	(703.80, 44.24) --
	(703.53, 44.23) --
	(703.26, 44.23) --
	(702.99, 44.23) --
	(702.71, 44.23) --
	(702.44, 44.23) --
	(702.17, 44.23) --
	(701.90, 44.23) --
	(701.63, 44.22) --
	(701.36, 44.22) --
	(701.09, 44.22) --
	(700.82, 44.22) --
	(700.55, 44.22) --
	(700.28, 44.22) --
	(700.00, 44.22) --
	(699.73, 44.22) --
	(699.46, 44.22) --
	(699.19, 44.22) --
	(698.92, 44.22) --
	(698.65, 44.22) --
	(698.38, 44.22) --
	(698.11, 44.22) --
	(697.84, 44.22) --
	(697.57, 44.22) --
	(697.29, 44.22) --
	(697.02, 44.23) --
	(696.75, 44.23) --
	(696.48, 44.23) --
	(696.21, 44.23) --
	(695.94, 44.23) --
	(695.67, 44.23) --
	(695.40, 44.23) --
	(695.13, 44.23) --
	(694.86, 44.23) --
	(694.59, 44.24) --
	(694.31, 44.24) --
	(694.04, 44.24) --
	(693.77, 44.24) --
	(693.50, 44.24) --
	(693.23, 44.24) --
	(692.96, 44.24) --
	(692.69, 44.24) --
	(692.42, 44.24) --
	(692.15, 44.24) --
	(691.88, 44.24) --
	(691.60, 44.24) --
	(691.33, 44.24) --
	(691.06, 44.24) --
	(690.79, 44.24) --
	(690.52, 44.24) --
	(690.25, 44.24) --
	(689.98, 44.24) --
	(689.71, 44.25) --
	(689.44, 44.25) --
	(689.17, 44.25) --
	(688.89, 44.25) --
	(688.62, 44.26) --
	(688.35, 44.26) --
	(688.08, 44.27) --
	(687.81, 44.27) --
	(687.54, 44.27) --
	(687.27, 44.28) --
	(687.00, 44.28) --
	(686.73, 44.28) --
	(686.46, 44.29) --
	(686.18, 44.29) --
	(685.91, 44.29) --
	(685.64, 44.29) --
	(685.37, 44.29) --
	(685.10, 44.29) --
	(684.83, 44.29) --
	(684.56, 44.29) --
	(684.29, 44.29) --
	(684.02, 44.29) --
	(683.75, 44.29) --
	(683.48, 44.29) --
	(683.20, 44.29) --
	(682.93, 44.29) --
	(682.66, 44.29) --
	(682.39, 44.29) --
	(682.12, 44.28) --
	(681.85, 44.28) --
	(681.58, 44.28) --
	(681.31, 44.28) --
	(681.04, 44.28) --
	(680.77, 44.28) --
	(680.49, 44.28) --
	(680.22, 44.28) --
	(679.95, 44.28) --
	(679.68, 44.28) --
	(679.41, 44.28) --
	(679.14, 44.28) --
	(678.87, 44.28) --
	(678.60, 44.28) --
	(678.33, 44.28) --
	(678.06, 44.28) --
	(677.78, 44.27) --
	(677.51, 44.27) --
	(677.24, 44.27) --
	(676.97, 44.27) --
	(676.70, 44.27) --
	(676.43, 44.27) --
	(676.16, 44.27) --
	(675.89, 44.27) --
	(675.62, 44.27) --
	(675.35, 44.27) --
	(675.07, 44.27) --
	(674.80, 44.27) --
	(674.53, 44.27) --
	(674.26, 44.27) --
	(673.99, 44.27) --
	(673.72, 44.27) --
	(673.45, 44.27) --
	(673.18, 44.27) --
	(672.91, 44.27) --
	(672.64, 44.27) --
	(672.37, 44.28) --
	(672.09, 44.28) --
	(671.82, 44.28) --
	(671.55, 44.28) --
	(671.28, 44.28) --
	(671.01, 44.28) --
	(670.74, 44.28) --
	(670.47, 44.28) --
	(670.20, 44.28) --
	(669.93, 44.28) --
	(669.66, 44.27) --
	(669.38, 44.27) --
	(669.11, 44.27) --
	(668.84, 44.27) --
	(668.57, 44.27) --
	(668.30, 44.27) --
	(668.03, 44.26) --
	(667.76, 44.26) --
	(667.49, 44.26) --
	(667.22, 44.26) --
	(666.95, 44.26) --
	(666.67, 44.26) --
	(666.40, 44.26) --
	(666.13, 44.26) --
	(665.86, 44.26) --
	(665.59, 44.26) --
	(665.32, 44.26) --
	(665.05, 44.26) --
	(664.78, 44.26) --
	(664.51, 44.26) --
	(664.24, 44.27) --
	(663.96, 44.27) --
	(663.69, 44.27) --
	(663.42, 44.28) --
	(663.15, 44.28) --
	(662.88, 44.28) --
	(662.61, 44.29) --
	(662.34, 44.29) --
	(662.07, 44.29) --
	(661.80, 44.30) --
	(661.53, 44.30) --
	(661.26, 44.31) --
	(660.98, 44.31) --
	(660.71, 44.31) --
	(660.44, 44.32) --
	(660.17, 44.32) --
	(659.90, 44.33) --
	(659.63, 44.33) --
	(659.36, 44.34) --
	(659.09, 44.35) --
	(658.82, 44.35) --
	(658.55, 44.36) --
	(658.27, 44.36) --
	(658.00, 44.37) --
	(657.73, 44.37) --
	(657.46, 44.38) --
	(657.19, 44.38) --
	(656.92, 44.39) --
	(656.65, 44.39) --
	(656.38, 44.40) --
	(656.11, 44.40) --
	(655.84, 44.40) --
	(655.56, 44.40) --
	(655.29, 44.41) --
	(655.02, 44.41) --
	(654.75, 44.41) --
	(654.48, 44.41) --
	(654.21, 44.41) --
	(653.94, 44.41) --
	(653.67, 44.41) --
	(653.40, 44.41) --
	(653.13, 44.41) --
	(652.85, 44.41) --
	(652.58, 44.42) --
	(652.31, 44.42) --
	(652.04, 44.42) --
	(651.77, 44.43) --
	(651.50, 44.43) --
	(651.23, 44.44) --
	(650.96, 44.45) --
	(650.69, 44.46) --
	(650.42, 44.46) --
	(650.15, 44.47) --
	(649.87, 44.49) --
	(649.60, 44.50) --
	(649.33, 44.51) --
	(649.06, 44.52) --
	(648.79, 44.53) --
	(648.52, 44.55) --
	(648.25, 44.56) --
	(647.98, 44.58) --
	(647.71, 44.59) --
	(647.44, 44.61) --
	(647.16, 44.63) --
	(646.89, 44.64) --
	(646.62, 44.66) --
	(646.35, 44.67) --
	(646.08, 44.69) --
	(645.81, 44.71) --
	(645.54, 44.72) --
	(645.27, 44.74) --
	(645.00, 44.76) --
	(644.73, 44.77) --
	(644.45, 44.79) --
	(644.18, 44.80) --
	(643.91, 44.81) --
	(643.64, 44.83) --
	(643.37, 44.84) --
	(643.10, 44.85) --
	(642.83, 44.86) --
	(642.56, 44.87) --
	(642.29, 44.88) --
	(642.02, 44.89) --
	(641.74, 44.90) --
	(641.47, 44.90) --
	(641.20, 44.91) --
	(640.93, 44.92) --
	(640.66, 44.93) --
	(640.39, 44.93) --
	(640.12, 44.94) --
	(639.85, 44.95) --
	(639.58, 44.96) --
	(639.31, 44.97) --
	(639.03, 44.98) --
	(638.76, 44.99) --
	(638.49, 45.00) --
	(638.22, 45.01) --
	(637.95, 45.03) --
	(637.68, 45.04) --
	(637.41, 45.06) --
	(637.14, 45.08) --
	(636.87, 45.10) --
	(636.60, 45.12) --
	(636.33, 45.14) --
	(636.05, 45.16) --
	(635.78, 45.19) --
	(635.51, 45.21) --
	(635.24, 45.24) --
	(634.97, 45.26) --
	(634.70, 45.29) --
	(634.43, 45.31) --
	(634.16, 45.34) --
	(633.89, 45.36) --
	(633.62, 45.39) --
	(633.34, 45.41) --
	(633.07, 45.43) --
	(632.80, 45.45) --
	(632.53, 45.47) --
	(632.26, 45.49) --
	(631.99, 45.51) --
	(631.72, 45.52) --
	(631.45, 45.54) --
	(631.18, 45.55) --
	(630.91, 45.57) --
	(630.63, 45.58) --
	(630.36, 45.60) --
	(630.09, 45.61) --
	(629.82, 45.62) --
	(629.55, 45.64) --
	(629.28, 45.65) --
	(629.01, 45.66) --
	(628.74, 45.68) --
	(628.47, 45.69) --
	(628.20, 45.71) --
	(627.92, 45.73) --
	(627.65, 45.75) --
	(627.38, 45.76) --
	(627.11, 45.78) --
	(626.84, 45.81) --
	(626.57, 45.83) --
	(626.30, 45.85) --
	(626.03, 45.88) --
	(625.76, 45.91) --
	(625.49, 45.94) --
	(625.22, 45.97) --
	(624.94, 46.00) --
	(624.67, 46.04) --
	(624.40, 46.07) --
	(624.13, 46.12) --
	(623.86, 46.16) --
	(623.59, 46.20) --
	(623.32, 46.25) --
	(623.05, 46.30) --
	(622.78, 46.35) --
	(622.51, 46.40) --
	(622.23, 46.45) --
	(621.96, 46.51) --
	(621.69, 46.57) --
	(621.42, 46.63) --
	(621.15, 46.69) --
	(620.88, 46.76) --
	(620.61, 46.82) --
	(620.34, 46.89) --
	(620.07, 46.96) --
	(619.80, 47.03) --
	(619.52, 47.11) --
	(619.25, 47.19) --
	(618.98, 47.27) --
	(618.71, 47.35) --
	(618.44, 47.44) --
	(618.17, 47.53) --
	(617.90, 47.62) --
	(617.63, 47.71) --
	(617.36, 47.80) --
	(617.09, 47.90) --
	(616.81, 47.99) --
	(616.54, 48.08) --
	(616.27, 48.18) --
	(616.00, 48.27) --
	(615.73, 48.35) --
	(615.46, 48.44) --
	(615.19, 48.52) --
	(614.92, 48.60) --
	(614.65, 48.68) --
	(614.38, 48.75) --
	(614.11, 48.82) --
	(613.83, 48.88) --
	(613.56, 48.94) --
	(613.29, 49.00) --
	(613.02, 49.06) --
	(612.75, 49.11) --
	(612.48, 49.16) --
	(612.21, 49.20) --
	(611.94, 49.25) --
	(611.67, 49.29) --
	(611.40, 49.34) --
	(611.12, 49.38) --
	(610.85, 49.43) --
	(610.58, 49.47) --
	(610.31, 49.52) --
	(610.04, 49.57) --
	(609.77, 49.62) --
	(609.50, 49.68) --
	(609.23, 49.74) --
	(608.96, 49.81) --
	(608.69, 49.88) --
	(608.41, 49.95) --
	(608.14, 50.03) --
	(607.87, 50.11) --
	(607.60, 50.19) --
	(607.33, 50.28) --
	(607.06, 50.37) --
	(606.79, 50.46) --
	(606.52, 50.56) --
	(606.25, 50.65) --
	(605.98, 50.75) --
	(605.70, 50.84) --
	(605.43, 50.94) --
	(605.16, 51.03) --
	(604.89, 51.12) --
	(604.62, 51.21) --
	(604.35, 51.29) --
	(604.08, 51.38) --
	(603.81, 51.46) --
	(603.54, 51.53) --
	(603.27, 51.61) --
	(603.00, 51.69) --
	(602.72, 51.76) --
	(602.45, 51.84) --
	(602.18, 51.92) --
	(601.91, 52.00) --
	(601.64, 52.08) --
	(601.37, 52.17) --
	(601.10, 52.26) --
	(600.83, 52.36) --
	(600.56, 52.46) --
	(600.29, 52.57) --
	(600.01, 52.69) --
	(599.74, 52.81) --
	(599.47, 52.93) --
	(599.20, 53.06) --
	(598.93, 53.19) --
	(598.66, 53.32) --
	(598.39, 53.45) --
	(598.12, 53.59) --
	(597.85, 53.72) --
	(597.58, 53.85) --
	(597.30, 53.98) --
	(597.03, 54.10) --
	(596.76, 54.22) --
	(596.49, 54.33) --
	(596.22, 54.44) --
	(595.95, 54.55) --
	(595.68, 54.64) --
	(595.41, 54.73) --
	(595.14, 54.81) --
	(594.87, 54.89) --
	(594.59, 54.95) --
	(594.32, 55.01) --
	(594.05, 55.07) --
	(593.78, 55.11) --
	(593.51, 55.15) --
	(593.24, 55.17) --
	(592.97, 55.19) --
	(592.70, 55.20) --
	(592.43, 55.20) --
	(592.16, 55.20) --
	(591.89, 55.18) --
	(591.61, 55.15) --
	(591.34, 55.11) --
	(591.07, 55.07) --
	(590.80, 55.01) --
	(590.53, 54.94) --
	(590.26, 54.87) --
	(589.99, 54.78) --
	(589.72, 54.69) --
	(589.45, 54.58) --
	(589.18, 54.47) --
	(588.90, 54.36) --
	(588.63, 54.23) --
	(588.36, 54.10) --
	(588.09, 53.96) --
	(587.82, 53.81) --
	(587.55, 53.66) --
	(587.28, 53.50) --
	(587.01, 53.33) --
	(586.74, 53.16) --
	(586.47, 52.97) --
	(586.19, 52.78) --
	(585.92, 52.59) --
	(585.65, 52.38) --
	(585.38, 52.17) --
	(585.11, 51.95) --
	(584.84, 51.72) --
	(584.57, 51.49) --
	(584.30, 51.25) --
	(584.03, 51.01) --
	(583.76, 50.76) --
	(583.48, 50.51) --
	(583.21, 50.25) --
	(582.94, 50.00) --
	(582.67, 49.74) --
	(582.40, 49.49) --
	(582.13, 49.23) --
	(581.86, 48.98) --
	(581.59, 48.74) --
	(581.32, 48.50) --
	(581.05, 48.27) --
	(580.77, 48.04) --
	(580.50, 47.82) --
	(580.23, 47.62) --
	(579.96, 47.41) --
	(579.69, 47.22) --
	(579.42, 47.04) --
	(579.15, 46.86) --
	(578.88, 46.69) --
	(578.61, 46.54) --
	(578.34, 46.39) --
	(578.07, 46.25) --
	(577.79, 46.11) --
	(577.52, 45.98) --
	(577.25, 45.86) --
	(576.98, 45.75) --
	(576.71, 45.64) --
	(576.44, 45.54) --
	(576.17, 45.44) --
	(575.90, 45.35) --
	(575.63, 45.26) --
	(575.36, 45.18) --
	(575.08, 45.10) --
	(574.81, 45.03) --
	(574.54, 44.96) --
	(574.27, 44.89) --
	(574.00, 44.83) --
	(573.73, 44.77) --
	(573.46, 44.72) --
	(573.19, 44.67) --
	(572.92, 44.62) --
	(572.65, 44.58) --
	(572.37, 44.54) --
	(572.10, 44.50) --
	(571.83, 44.47) --
	(571.56, 44.44) --
	(571.29, 44.41) --
	(571.29, 44.01) --
	cycle;
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (553.52, 31.02) rectangle (717.20,211.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57, 41.18) {Pooled estimate};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57, 63.17) {Study 1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57, 85.15) {Study 2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,107.14) {Study 3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,129.13) {Study 4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,151.11) {Study 5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,173.10) {Study 6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (548.57,195.09) {Study 7};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77, 44.21) --
	(553.52, 44.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77, 66.20) --
	(553.52, 66.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77, 88.18) --
	(553.52, 88.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,110.17) --
	(553.52,110.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,132.16) --
	(553.52,132.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,154.14) --
	(553.52,154.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,176.13) --
	(553.52,176.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.77,198.12) --
	(553.52,198.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (560.96, 28.27) --
	(560.96, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.67, 28.27) --
	(597.67, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.38, 28.27) --
	(634.38, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (671.09, 28.27) --
	(671.09, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (707.80, 28.27) --
	(707.80, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (560.96, 20.01) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (597.67, 20.01) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (634.38, 20.01) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (671.09, 20.01) {1.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (707.80, 20.01) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (635.36,  7.82) {Generation-interval dispersion};
\end{scope}
\end{tikzpicture}

\end{document}

### 2020-04-28 figure2.tex
working on it...

% Created by tikzDevice version 0.12.3 on 2020-04-27 21:17:30
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 37.26, 33.48) rectangle (283.08,264.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 37.26, 33.48) rectangle (283.08,264.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
	(283.08, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 77.88) --
	(283.08, 77.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,122.28) --
	(283.08,122.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,166.68) --
	(283.08,166.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,211.08) --
	(283.08,211.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,255.48) --
	(283.08,255.48);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 65.62,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (112.90,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (160.17,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (207.44,122.88) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.72,119.58) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
	( 65.62,122.88);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
	( 65.62,122.88);

\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
	( 65.62,122.88);

\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90,234.50) --
	(112.90,234.50);

\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90,234.50) --
	(112.90, 55.99);

\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90, 55.99) --
	(112.90, 55.99);

\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17,153.48) --
	(160.17,153.48);

\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17,153.48) --
	(160.17, 99.27);

\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17, 99.27) --
	(160.17, 99.27);

\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44,164.15) --
	(207.44,164.15);

\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44,164.15) --
	(207.44, 68.61);

\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44, 68.61) --
	(207.44, 68.61);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72,260.46) --
	(254.72,260.46);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72,260.46) --
	(254.72, 41.21);

\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72, 41.21) --
	(254.72, 41.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
	( 37.26,264.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86, 30.17) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86, 74.57) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,118.97) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,163.37) {3.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,207.78) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,252.18) {4.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26, 33.48) --
	( 37.26, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26, 77.88) --
	( 37.26, 77.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,122.28) --
	( 37.26,122.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,166.68) --
	( 37.26,166.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,211.08) --
	( 37.26,211.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,255.48) --
	( 37.26,255.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
	(283.08, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 65.62, 30.48) --
	( 65.62, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.90, 30.48) --
	(112.90, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (160.17, 30.48) --
	(160.17, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.44, 30.48) --
	(207.44, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.72, 30.48) --
	(254.72, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 65.62, 21.46) {base};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (112.90, 21.46) {$\mu_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (160.17, 21.46) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (207.44, 21.46) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (254.72, 21.46) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (160.17,  8.33) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 14.26,148.92) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.44] at ( 37.26,273.16) {A. Baseline};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] (326.34, 33.48) rectangle (572.16,264.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (326.34, 33.48) rectangle (572.16,264.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 79.65) --
	(572.16, 79.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34,141.22) --
	(572.16,141.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34,202.79) --
	(572.16,202.79);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (354.70,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (401.98,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (449.25,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (496.52,142.26) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (543.80,140.42) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
	(354.70,142.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
	(354.70,142.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
	(354.70,142.26);

\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,187.46) --
	(401.98,187.46);

\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,187.46) --
	(401.98,111.70);

\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,111.70) --
	(401.98,111.70);

\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,195.31) --
	(449.25,195.31);

\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,195.31) --
	(449.25,101.35);

\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,101.35) --
	(449.25,101.35);

\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52,213.80) --
	(496.52,213.80);

\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52,213.80) --
	(496.52, 48.20);

\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52, 48.20) --
	(496.52, 48.20);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80,261.16) --
	(543.80,261.16);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80,261.16) --
	(543.80, 41.24);

\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80, 41.24) --
	(543.80, 41.24);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 33.48) --
	(326.34,264.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94, 76.35) {2.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94,137.92) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94,199.49) {3.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34, 79.65) --
	(326.34, 79.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34,141.22) --
	(326.34,141.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34,202.79) --
	(326.34,202.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 33.48) --
	(572.16, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.70, 30.48) --
	(354.70, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.98, 30.48) --
	(401.98, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (449.25, 30.48) --
	(449.25, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.52, 30.48) --
	(496.52, 33.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (543.80, 30.48) --
	(543.80, 33.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (354.70, 21.46) {base};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (401.98, 21.46) {$\hat{\mu}_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (449.25, 21.46) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (496.52, 21.46) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (543.80, 21.46) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (449.25,  8.33) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (303.34,148.92) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.44] at (326.34,273.16) {B. Reduced uncertainty in $r$};
\end{scope}
\end{tikzpicture}

\end{document}

### 2020-04-28 ncov.tex
working on it...

@@
 However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary significantly even though most analyses relied on reports of confirmed cases from China, particularly from the Hubei province.
-Understanding the differences between \Ro estimates is necessary as \Ro provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
-It can be difficult to compare a disparate set of estimates of \Ro when the estimation methods and their underlying assumptions vary widely.
+Compare a disparate set of estimates of \Ro is difficult when the estimation methods and their underlying assumptions vary widely.
+In some cases, similar methods can give different estimates; in other cases, different methods can give similar estimates.
+Understanding the differences between \Ro estimates is critical to controlling an epidemic as \Ro provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 
@@
 \hline
-Study 2 & Stochastic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020)  & 2.5 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 2 & Stochastic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020)  & 2.6 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
 \hline
@@
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-Typically, Bayesian multilevel models seek to leverage estimates across spatially and/or temporally varying outbreaks of a same disease.
 In this case, most estimates are based on reported number of cases from China, particularly from Wuhan province.
@@
 Thus, the pooled estimator should be interpreted with care.
-
-\rev{Authors state that “We note that this approach does not account for non- independence between the parameter estimates made by different modelers.” Authors should discuss this point more in depth.
-This seems to me an important limitation.
-With this respect I wonder what data the studies focus on.
-Are all analyzing the same epidemic (in China) considering same or similar data?
-Again, more details on these studies (in this case the location, period of the epidemic, and data source) should be provided.}
 
@@
 
+\subsection{Sensitivity analysis}
+
+In order to understand how uncertainties in each component ($r_i$, $\bar G_i$, and $\kappa_i$) affect the estimate of $(\Ro)_i$ from each study $i$, we replace $r_i$, $\bar G_i$, and $\kappa_i$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculate the basic reproductive number \Ro.
+We refer to the resulting estimates of \Ro as ``substitute'' estimates.
+For example, $r$-substitute estimate for Study $i$ is computed as:
+\begin{equation}
+\left(1 + \kappa_i \mu_r \bar{G}_i\right)^{1/\kappa_i},
+\end{equation}
+where $\kappa_i$ and $\bar{G}_i$ are taken from their corresponding parameter sets and $\mu_r$ is taken from the posterior distribution.
+This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
+We compare substitute estimates with the base estimates of \Ro (based on $r_i$, $\bar G_i$, and $\kappa_i$) as well as pooled estimates of \Ro (baed on $\mu_r$, $\mu_G$, and $\mu_\kappa$).
+
 \section{Results}
@@
 
-\rev{In figure 1, the posterior density distribution for the 3 parameters [exponential growth rate (r),
-Mean generation interval ($\bar G$), Generation interval dispersion (kappa))] must replace the
-segment representing the median and the CI. In figure 1, the legend must be self-sufficient. ‘See
-text’ it is not possible. A short explanation is necessary.}
-
-\rev{In the first paragraph of the Results page 7 the value of the median and the CI of the 3
-parameters must be presented.}
-
-\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model.
+\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model:
+$\mu_r = 0.17\,\textrm{days}^{-1}$ (95\% CI: 0.12--0.25 $\textrm{days}^{-1}$),
+$\mu_G = 8.51\,\textrm{days}$ (95\% CI: 7.60--9.63 days),
+and
+$\mu_\kappa = 0.50$ (95\% CI: 0.26--1.10).
 We find that there is a large uncertainty associated with the underlying parameters;
@@
 \textbf{Comparisons of the reported parameter values with our pooled estimates.}
-We inferred point estimates (black), uniform distributions (orange) or confidence/credible intervals (purple) for each parameter from each study, and combined them into pooled estimates (red; see text).
+We inferred point estimates (black), uniform distributions (orange) or confidence/credible intervals (purple) for each parameter from each study, and combined them into pooled estimates using a Bayesian multilevel model (red).
+Points represent medians calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$ (orange and purple).
+Error bars represent 95\% quantiles calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$.
+Red violin plots represent posterior distributions.
 Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval dispersion.
@@
 
-\rev{The estimation of R0 median and its credible interval “2.9 (95\% CI: 2.1-4.5)” must be presented
-in page 7 with the figure 2 in the second paragraph of Results and not at the end of the Results
-part (page 9 line 26).}
-
-\rev{The uncertainty related to kappa is not described in Figure 2. Even the authors mentioned in
-Page 7, Lines 44-50, the results should also be visualized in Figure 2 along with other
-parameters.}
-
 \rev{Page 7, Line 38. The authors might give some information about the incubation period of
@@
 3rd day of the city lockdown.}
+
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro. 
 For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26 (\fref{eff}A).
-Comparing the models that include only some sources of uncertainty to the ``all'' model, we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: the middle bar (``GI mean''), which lacks growth-rate uncertainty, is relatively narrow.
+Comparing the estimates that include only some sources of uncertainty to the ``all'' estimate ($\Ro = 3.0$; 95\% CI: 2.1--4.6), we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: two middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, is relatively narrow.
 In this case, propagating error from the mean generation interval has negligible effect compared to propagating the uncertainty in $r$.
-Uncertainty in the generation-interval dispersion $\kappa$ also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare ``growth rate + GI mean'' with ``all'').
+Uncertainty in the generation-interval dispersion $\kappa$ also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare $\mu_G$ with $\mu_\kappa$).
 For example, reducing $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
 
-As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important (\fref{eff}B).
+We further explore how the effects of uncertainties in generation-interval distributions change when there is less uncertainty associated with the exponential growth rate.
+This hypothetical example reflects recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
+To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (\fref{eff}B); 
+then $\hat{\mu}_r$ has the same median as $\mu_r$ but the associated 95\% CI is 4 times narrower (0.16--0.19 $\textrm{days}^{-1}$).
+As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important.
 Propagating error only from the growth rate gives very narrow credible intervals in this case. 
 Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow credible intervals.
-We expect this hypothetical example to better reflect more recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
-
-\rev{Please clearly explain the formula in the caption of Figure 2 (Page 8, line 33) in the Methods
-section.}
 
@@
 We also compare the estimates of \Ro across different studies by 
-replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro (\fref{R0}).
+replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro.
+We refer to these estimates as ``substitue estimates''.
 This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
+
+\fref{R0} compares the ``base'' estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 ``substitute'' estimates (3 parameter substitutions $\times$ 7 studies).
+
 We find that incorporating uncertainties one at a time increases the width of the credible intervals in all but 7 cases.

### 2020-04-28 MCMC_all.R

@@ -10,7 +10,7 @@ priorfun <- function(theta) {
     dgamma(theta[5], 2, 2/0.5, log=TRUE)
 }
 
-sample_data$`Majumder et al. (2020)`$kappa <- 0.02
+sample_data$`Majumder and Mandl (2020)`$kappa <- 0.02
 
 postfun <- function(theta) {
   if (any(theta < 0))

### 2020-04-28 compare_R0.R

@@ -2,8 +2,22 @@ library(ggplot2); theme_set(theme_bw())
 library(gridExtra)
 library(colorspace)
 library(dplyr)
-source("published_estimate.R")
 source("study_number.R")
+
+load("sample_data.rda")
+
+Rdata_adj <- sample_data %>%
+  bind_rows(.id="study") %>%
+  group_by(study) %>%
+  mutate(
+    R2=(1 + kappa * r * gbar)^(1/kappa),
+    R2=ifelse(kappa==0, exp(r * gbar), R2),
+  ) %>%
+  summarize(
+    est=median(R2),
+    lwr=quantile(R2, 0.025),
+    upr=quantile(R2, 0.975)
+  )
 
 load("R0_r.rda")
 load("R0_gbar.rda")
@@ -20,7 +34,7 @@ R0all <- Rdata_adj %>%
   ) %>%
   mutate(
     type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
-                labels=c("base", "growth rate", "mean generation interval", "generation-interval dispersion", "all"))
+                labels=c("base", "exponential growth rate", "mean generation interval", "generation-interval dispersion", "all"))
   ) %>%
   merge(study_number) %>%
   mutate(
@@ -42,7 +56,7 @@ R0all %>%
   mutate(
     width=upr-lwr
   ) %>%
-  merge(summarize(group_by(Rdata, study), bwidth=upr-lwr)) %>%
+  merge(summarize(group_by(Rdata_adj, study), bwidth=upr-lwr)) %>%
   mutate(
     ww=width>bwidth
   )
@@ -65,5 +79,7 @@ g1 <- ggplot(R0all) +
     legend.title = element_blank()
   )
 
-ggsave("compare_R0.pdf", g1, width=7, height=4)
-ggsave("compare_R0.png", g1, width=7, height=4)
+tikz(file = "compare_R0.tex", width = 8, height = 4, standAlone = T)
+g1
+dev.off()
+tools::texi2dvi('compare_R0.tex', pdf = T, clean = T)

### 2020-04-28 compare_assumption.R

@@ -2,7 +2,7 @@ library(coda)
 library(ggplot2); theme_set(theme_bw())
 library(dplyr)
 library(gridExtra)
-source("published_estimate.R")
+library(tikzDevice)
 source("study_number.R")
 
 load("sample_data.rda")
@@ -19,82 +19,111 @@ rdata2 <- sample_data %>%
   }) %>%
   bind_rows(.id="study")
 
+gdata2 <- sample_data %>%
+  lapply(function(x) {
+    data.frame(
+      est=median(x$gbar),
+      lwr=quantile(x$gbar, 0.025),
+      upr=quantile(x$gbar, 0.975)
+    )
+  }) %>%
+  bind_rows(.id="study")
+
+kappadata2 <- sample_data %>%
+  lapply(function(x) {
+    data.frame(
+      est=median(x$kappa),
+      lwr=quantile(x$kappa, 0.025),
+      upr=quantile(x$kappa, 0.975)
+    )
+  }) %>%
+  bind_rows(.id="study")
+
 avg_r <- data.frame(
-  est=median(unlist(unlist(MCMC_all[,1]))),
-  lwr=quantile(unlist(unlist(MCMC_all[,1])), 0.025),
-  upr=quantile(unlist(unlist(MCMC_all[,1])), 0.975),
+  est=unlist(unlist(MCMC_all[,1])),
   anon="Pooled estimate"
-)
+) %>%
+  mutate(
+    anon=factor(anon, levels=c("Pooled estimate", paste0("Study ", 7:1)))
+  )
+
+quantile(avg_r$est, c(0.025, 0.5, 0.975))
 
 avg_gen <- data.frame(
-  est=median(unlist(unlist(MCMC_all[,3]))),
-  lwr=quantile(unlist(unlist(MCMC_all[,3])), 0.025),
-  upr=quantile(unlist(unlist(MCMC_all[,3])), 0.975),
+  est=unlist(unlist(MCMC_all[,3])),
   anon="Pooled estimate"
-)
+) %>%
+  mutate(
+    anon=factor(anon, levels=c("Pooled estimate", paste0("Study ", 7:1)))
+  )
+
+quantile(avg_gen$est, c(0.025, 0.5, 0.975))
 
 avg_kappa <- data.frame(
-  est=median(unlist(unlist(MCMC_all[,5]))),
-  lwr=quantile(unlist(unlist(MCMC_all[,5])), 0.025),
[…]

### 2020-04-28 figure2.R

@@ -1,6 +1,7 @@
 library(coda)
-library(ggplot2); theme_set(theme_bw(base_size=16))
+library(ggplot2); theme_set(theme_bw(base_size=12))
 library(gridExtra)
+library(tikzDevice)
 
 load("sample_data.rda")
 load("MCMC_all.rda")
@@ -21,21 +22,21 @@ compfun <- function(x) {
     est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
     lwr=quantile((1+median(kappa)*median(gbar)*x)^(1/median(kappa)), 0.025),
     upr=quantile((1+median(kappa)*median(gbar)*x)^(1/median(kappa)), 0.975),
-    type="growth rate"
+    type="$\\mu_r$"
   )
   
   dd_gbar <- data.frame(
     est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
     lwr=quantile((1+median(kappa)*gbar*median(x))^(1/median(kappa)), 0.025),
     upr=quantile((1+median(kappa)*gbar*median(x))^(1/median(kappa)), 0.975),
-    type="GI mean"
+    type="$\\mu_G$"
   )
   
-  dd_both <- data.frame(
-    est=median((1+median(kappa)*gbar*x)^(1/median(kappa))),
-    lwr=quantile((1+median(kappa)*gbar*x)^(1/median(kappa)), 0.025),
-    upr=quantile((1+median(kappa)*gbar*x)^(1/median(kappa)), 0.975),
-    type="growth rate\n+\nGI mean"
+  dd_kappa <- data.frame(
+    est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
+    lwr=quantile((1+kappa*median(gbar)*median(x))^(1/kappa), 0.025),
+    upr=quantile((1+kappa*median(gbar)*median(x))^(1/kappa), 0.975),
+    type="$\\mu_\\kappa$"
   )
   
   dd_all <- data.frame(
@@ -45,17 +46,22 @@ compfun <- function(x) {
     type="all"
   )
   
-  dd_combine <- rbind(dd, dd_r, dd_gbar, dd_both, dd_all)
+  dd_combine <- rbind(dd, dd_r, dd_gbar, dd_kappa, dd_all)
   
   dd_combine
 }
 
 dd_combine <- compfun(r)
 
-dd_combine2 <- compfun((r + median(r)*3)/4)
[…]

### 2020-04-28 published_estimate.R

@@ -1,29 +1,0 @@
-Rdata <- data.frame(
-  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder and Mandl (2020)", "Liu et al. (2020)"),
-  est=c(NA, 2.5, 2.2, 3.8, 5.47, NA, 2.92),
-  lwr=c(1.5, 1.5, 1.4, 3.6, 4.16, 2, 2.28),
-  upr=c(3.5, 3.5, 3.8, 4, 7.1, 3.1, 3.67)
-)
-
-Rdata_adj <- data.frame(
-  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder and Mandl (2020)", "Liu et al. (2020)"),
-  est=c(NA, 2.5, 2.2, 3.8, 5.47, NA, 2.92),
-  lwr=c(qunif(0.025, 1.5, 3.6), 1.5, qgamma(0.025, 12, 12/2.2), 3.6, 4.16, exp(0.114 * qunif(0.025, 6, 10)), 2.28),
-  upr=c(qunif(0.975, 1.5, 3.6), 3.5, qgamma(0.975, 12, 12/2.2), 4, 7.1, exp(0.114 * qunif(0.975, 6, 10)), 3.67)
-)
-
-gdata <- data.frame(
-  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder and Mandl (2020)", "Liu et al. (2020)"),
-  est=c(10, 8.4, NA, 7.6, NA, NA, 8.4),
-  lwr=c(10, 8.4, 7, 7.6, 7.6, 6, 8.4),
-  upr=c(10, 8.4, 14, 7.6, 8.4, 10, 8.4)
-)
-
-kappadata <- data.frame(
-  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Riou and Althaus (2020)", "Read et al. (2020)",
-          "Zhao et al. (2020)", "Majumder and Mandl (2020)", "Liu et al. (2020)"),
-  est=c(1, NA, 0.5, 0.5, 0.2, 0, 0.2)
-)

### 2020-04-28 sample_author.R

@@ -9,24 +9,12 @@ sample_bedford <- function(n=1) {
 }
 
 sample_imai <- function(n=1) {
-  # pgamma(3.5, 28, 28/2.6)-pgamma(1.5, 28, 28/2.6)
+  # pgamma(3.5, 18, 18/2.6)-pgamma(1.5, 18, 18/2.6)
   
   dplyr::data_frame(
     gbar=8.4,
     kappa=0.5,
-    R=rgamma(n, 28, rate=28/2.6),
-    r=(R^kappa-1)/(kappa*gbar),
-    rho=r*gbar
-  )
-}
-
-sample_imai <- function(n=1) {
-  # pgamma(3.5, 28, 28/2.6)-pgamma(1.5, 28, 28/2.6)
-  
-  dplyr::data_frame(
-    gbar=8.4,
-    kappa=0.5,
-    R=rgamma(n, 28, rate=28/2.6),
+    R=rgamma(n, 18, rate=18/2.6),
     r=(R^kappa-1)/(kappa*gbar),
     rho=r*gbar
   )

### 2020-04-29 review/response.tex
push

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

\usepackage{color}

\usepackage{lmodern}
\usepackage{amssymb,amsmath}

\newcommand{\rR}{\mbox{$r$--$\cal R$}}
\newcommand{\RR}{\ensuremath{{\cal R}}}
\newcommand{\RRhat}{\ensuremath{{\hat \cal R}}}
\newcommand{\Rx}[1]{\ensuremath{{\cal R}_{#1}}} 
\newcommand{\Ro}{\Rx{0}}
\newcommand{\Reff}{\Rx{\mathit{eff}}}
\newcommand{\Tc}{\ensuremath{C}}

\newcommand{\rev}{\subsection*}
\newcommand{\revtext}{\textsf}
\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0em}

\newcommand{\comment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\dc}[1]{\comment{blue}{DC}{#1}}
\newcommand{\jsw}[1]{\comment{green}{JSW}{#1}}
\newcommand{\hotcomment}[1]{\comment{red}{HOT}{#1}}

\begin{document}

\noindent Dear Editor:

Thank you for the chance to revise and resubmit our manuscript. 
Below please find our responses to reviewers.

\rev{Reviewer \#1}

\revtext{
The manuscript presents a statistical framework for comparing different estimates of R0 decomposed in their building blocks (exponential growth, average generation interval, and dispersion in the generation interval). The analysis allows for pulled estimates of the different quantities, combining information from the different studies. In addition, authors provide a reflection on how the uncertainty in the different components is propagated in the R0 estimate and its confidence interval.}

\revtext{It seems to me that the paper is well written.}

Thank you!

\revtext{As a disclaimer, I must say that I am not sufficiently expert of the methodology used here to be fully confident in my judgment of the soundness and the completeness of the analysis. Maybe it is related to my limitation, but I find the main focus of the paper a little too technical for the Journal (e.g. see the last point below). Thus, I believe that authors could make an extra effort to convince a more general and interdisciplinary public on the importance of the point they are raising.}

We have made significant revisions to the manuscript to address this point.

\revtext{I am not sure whether authors incorporate well all assumptions of the studies they consider.}

During the exponential growth period, the basic reproduction number entirely depends on the exponential growth rate and the generation-interval distributions.
Therefore, all model assumptions can be boiled down to assumptions about the exponential growth rate and the shape of the generation-interval distribution.
So we are effectively considering all model assumptions.
We have tried to make this point clearer:

``Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic grows exponentially in the beginning.
While the IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth, Study 7 estimates $d=0$ across all parameters they consider, in which case the IDEA model becomes equivalent to the exponential growth model.
When the epidemic is growing exponentially, the estimated basic reproduction number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$:
\begin{equation}
1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
\label{eq:euler}
\end{equation}
In other words, all model assumptions come down to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence intervals of the model parameters will be necessarily narrow.
Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and shape of the generation-interval distribution to understand the estimates of the basic reproduction number.''

\revtext{In addition, it is not clear to me on what extent the assumptions made by authors themselves affect the results - e.g. the assumption of a gamma and uniform distribution for the estimates adopted to compute the growth rate.}

???

\revtext{Other hypotheses underlie these works other than average generation interval, its coefficient of variation and the exponential growth. For instance, what about the offspring distribution (Poisson vs. negative binomial), or the observation model? Authors should provide a synthesis of all assumptions of these studies}

See above.

\revtext{and discuss how well the gamma approximation framework embrace all cases.}

The gamma approximation framework assumes that the generation-interval distribution is gamma distributed.
Other assumptions mentioned above (e.g., offspring distributions or the observation model) do not affect the generation-interval distributions.
Moreover, this approximation has been widely used previously and has been tested across multiple diseases.
We have clarified the text and included appropriate citations:

``Here, we use the gamma approximation framework to the generation-interval distribution to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.''

\revtext{
Authors state that “We note that this approach does not account for non- independence between the parameter estimates made by different modelers.” Authors should discuss this point more in depth. This seems to me an important limitation. With this respect I wonder what data the studies focus on. Are all analyzing the same epidemic (in China) considering same or similar data? Again, more details on these studies (in this case the location, period of the epidemic, and data source) should be provided.}

We now provide a table that summarizes this information.

\revtext{
At the beginning of the Discussion section authors list a series of important issues and possible sources of errors that often limit the utility of early outbreak analyses (e.g. fitting the cumulative curve of cases, assuming a fixed generation interval when in fact it varies during the outbreak, changing reporting rate, etc.). Compared with these issues the main focus of the paper (not properly accounting for the uncertainty in the generation time and growth rate parameters) appears a small caveat. Maybe author could give a sense of the importance of the element they are focusing on with the respect to the other issues?}

\rev{Reviewer \#2}

We might have inadvertently confused reviewer 2.
Our method does not rely on any data or dynamical models but the reviewer asks us to compare our results with the reported number of cases (this is not possible as we do not make any predictions about the epidemic trajectory).
We use a multi-level model to calculate pooled estimates of parameters that have been estimated or assumed from other studies.
Then, we use the pooled estimates to calculate $\mathcal R_0$.
We have made significant improvement to our manuscript to make our methods clearer.

\revtext{The subject is interesting for the ongoing pandemic. The authors propose an estimation method
for the basic reproduction number and describe its uncertainty from the estimates made by
several early studies. The consensus epidemiological estimate could be beneficial for making
control strategies.}

Thank you!

\revtext{The authors propose a statistical method for an epidemiological situation. They combine basic
reproduction rate estimates and other estimates in the early stage of the outbreak and their
uncertainty to obtain a global estimate with its uncertainty. So, this subject is at the interface of
biomathematics and epidemiology of the infectious disease, which is the scope of the journal.}

\revtext{The interesting and primordial subject of the estimation of R0 and its uncertainty is studied using
the Bayesian inference tools. The authors do not develop enough the underlying hypotheses
and the method used. Many elements are not described, even in the appendix. This should be
described to estimate the validity of these underlying assumptions.}

\revtext{More specific points:}

\revtext{The authors must use the same variable names and their units throughout the documents: e.g.,
for kappa, which is a key parameter in this study. In the document, it appears in following
various terms that make the reader confused: GI variation, squared coefficient of variation,
generation-interval dispersion. Please always keep the same terms for the same variables. The
same remark is also applied to the mean generation interval.}

Done.

\revtext{The unit of time changes throughout the document. Choose the same time unit (e.g., days,
which is the most used and change the Equation 5, Page 6, which is in weeks).}

Done.

\revtext{As in the Bayesian inference, there are no Confidence Intervals like in frequentist inference, but
Credible Interval. The authors should replace Confidence Intervals with Credible Intervals.}

Done.

\revtext{Methods}

\revtext{The authors should clearly explain the hypotheses they have done in the Methods section (e.g.,
the source of the uncertainty).}

\revtext{The authors do not describe all the methods used in this article within the Methods section.
Some of them are in the figure captions, e.g., Figures 2 and 3. Also, the authors should describe
all the methods in this section. Using subheadings could help the reader to better understand
the methods, e.g., likelihood, prior distributions, sensitivity analysis, etc.}

\revtext{Please clearly explain the formula in the caption of Figure 2 (Page 8, line 33) in the Methods
section.}

\revtext{The Bayesian estimation framework is not clearly explained. Readers could easily get confused
about which variables serve as observed data, parameters, or hyperparameters. It is also not
very clear about the number of observed data, e.g., i = 7 means that there are only 7 data (1 per
study) or there are 700,000 simulated data (100,000 per study) in the model. We strongly
suggest that the authors should provide a directed acyclic graph (DAG) representing the entire
model, which is a common practice for the Bayesian estimation.}

\revtext{The authors consider that all the 7 studies have the same weight in the estimation of basic
reproduction weight even though their estimations and precision vary a lot. The authors might
explain why they make this hypothesis and why they did not explore a weighted estimation
based on the confidence of the estimation of R0.}

\revtext{In each chain, the authors use 500,000 sampling steps and remove every 1000th step.
Therefore, the posterior distributions of each parameter consist of 2,000 samples (500 samples
per chain with 4 independent chains). Is it sufficient for the effective sample size? Is it too much
for the burn-in steps of 500,000 and the thinning steps of 1,000?}

This may be an overkill but we wanted to make sure that the posterior samples are not autocorrelated.
This is definitely sufficient for the effective sample size to reliably calculate the 95\% confidence intervals, especially given that we're only averaging across previously estimated parameter values, rather than explicitly fitting a model to data.

\revtext{Results}

\revtext{In Figures 2 and 3, the authors replace the reported parameters (r, $\bar G$, and kappa) with the
pooled estimates mu ($\mu_r$, $\mu_G$, and $\mu_\kappa$). The authors do not give information about
the other sources of uncertainty linked with the variation of parameter sigma ($\sigma_r$, $\sigma_G$,
$\sigma_\kappa$).}

\revtext{The authors do not provide the validation on their estimates, e.g., a posterior predictive check
and comparisons with the reported cases during the study period.}

No...

\revtext{In figure 1, the posterior density distribution for the 3 parameters [exponential growth rate (r),
Mean generation interval ($\bar G$), Generation interval dispersion (kappa))] must replace the
segment representing the median and the CI. In figure 1, the legend must be self-sufficient. ‘See
text’ it is not possible. A short explanation is necessary.}

\revtext{In the first paragraph of the Results page 7 the value of the median and the CI of the 3
parameters must be presented.}

\revtext{The estimation of R0 median and its credible interval “2.9 (95\% CI: 2.1-4.5)” must be presented
in page 7 with the figure 2 in the second paragraph of Results and not at the end of the Results
part (page 9 line 26).}

\revtext{The uncertainty related to kappa is not described in Figure 2. Even the authors mentioned in
Page 7, Lines 44-50, the results should also be visualized in Figure 2 along with other
parameters.}

\revtext{Discussions}

\revtext{In the Results section, the authors occasionally discuss the results, i.e., Page 8, line 50 to Page
9, line 23. The opinions should be in the Discussion section.
In the Discussion section, the authors should discuss the quality of the reported estimates from
7 studies used in this subject. Some of them do not provide detailed methods and assumptions.
Also, most of them are preprint articles. It could be another important source of uncertainty on
R0.}

\revtext{Although the authors are in the case of the COVID-19 epidemic in Hubei before any
confinement, the authors do not mention in the introduction or in the discussion the other
sources of uncertainty linked to the data (such as the variations in diagnostic results related to
the day of the week, a saturation of diagnostic test capacity, transparency of data,
representativeness of samples and improvement of detection capacity as time goes by.}

\revtext{Page 4, Table 1. Please check Study 4, the article has been revised since 27 January 2020,
and their estimates also changed.}

\revtext{Page 6, Table 2. Please recheck the parameters of Study 5, which is inconsistent with the 90%
confidence interval in Table 1. The 90\% confidence interval for a Gamma distribution with mean
2.2 and shape 12 is 1.27 to 3.34, while the interval in Table 1 is 1.4 to 3.8.}

\revtext{Page 6, Line 29. Are the priors for $\mu_r$, $\mu_G$, and $\mu_\kappa$ weakly informative? Did the
authors check the sensibility with new priors?}

\revtext{Page 6 Lines 44-47. Please clarify the sentence: “Alternative choices of prior … can lead to
poor mixing”.}

\revtext{Page 7, Line 20. Figure 1 - Caption: Did the authors test the sensitivity of the value of kappa of
Study 2 (e.g., 0 or 1) on the pooled estimate?}

\revtext{Page 7, Line 38. The authors might give some information about the incubation period of
COVID-19. It could explain why using the data of 26 January is still acceptable even if it was the
3rd day of the city lockdown.}

\revtext{Page 8, Line 50. Please clarify the sentence: ”We find that incorporating uncertainties one at a
time increases the width of the confidence intervals in all but 7 cases”. It is not clear particularly
at “in all but 7 cases”.}

\rev{Reviewer \#3}

\revtext{
Here, Park and colleagues present a straight-forward, but important analysis of 7 early estimates of R0 for the ongoing covid outbreak. They stress that multiple forms of uncertainty have to be taken into account when calculating R0 and its confidence intervals, lest the intervals be too narrow and/or the estimates over-confident. The methods are sound, and the results compelling. I really don’t have any major comments, though I wonder if it would be better suited to another journal. I do apologize to the authors for the tardiness of my review.}

Thank you for the review. We have made significant revisions to our manuscript to make it approachable for a wider audience.

\end{document}

### 2020-04-29 ncov.tex
push

@@
 Specifically, 95\% quantiles for $\mu_r$, $\mu_G$, and $\mu_\kappa$ are 0.02--0.40 $\mathrm{days}^{-1}$, 0.8--19.5 days, and 0.1--1.4, respectively.
+Parameters that are outside these ranges are biologically unrealistic for SARS-CoV-2 outbreaks.
 Therefore, we do not expect our results to be sensitive to these priors.
@@
 This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
-We compare substitute estimates with the base estimates of \Ro (based on $r_i$, $\bar G_i$, and $\kappa_i$) as well as pooled estimates of \Ro (baed on $\mu_r$, $\mu_G$, and $\mu_\kappa$).
+We compare substitute estimates with the base estimates of \Ro (based on $r_i$, $\bar G_i$, and $\kappa_i$) as well as pooled estimates of \Ro (based on $\mu_r$, $\mu_G$, and $\mu_\kappa$).
 
@@
 
-\fref{assumption} compares the reported values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model:
+\fref{assumption} compares the estimated values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model:
 $\mu_r = 0.17\,\textrm{days}^{-1}$ (95\% CI: 0.12--0.25 $\textrm{days}^{-1}$),
@@
 Points represent medians calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$ (orange and purple).
-Error bars represent 95\% quantiles calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$.
-Red violin plots represent posterior distributions.
+Error bars represent 95\% equi-tailed quantiles calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$.
+Red violin plots represent distributions of 2000 posterior samples.
 Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval dispersion.

### 2020-05-03 ncov.tex
getting longer and longer!

@@
 Although the virus is likely to have originated from animal hosts \citep{andersen2020proximal}, the ability of SARS-CoV-2 to directly transmit between humans, particularly without symptoms, has posed a greater threat for its spread \citep{he2020temporal}.
-As of April 23, 2020, more than 2 million cases of the coronavirus disease 2019 (COVID-19) have been confirmed internationally \citep{who94}.
+As of March 2, 2020, more than 3 million cases of the coronavirus disease 2019 (COVID-19) have been confirmed internationally \citep{who103}.
 
@@
 However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary significantly even though most analyses relied on reports of confirmed cases from China, particularly from the Hubei province.
-Compare a disparate set of estimates of \Ro is difficult when the estimation methods and their underlying assumptions vary widely.
+Comparing a disparate set of estimates of \Ro can be particularly difficult when the estimation methods and their underlying assumptions vary widely.
 In some cases, similar methods can give different estimates; in other cases, different methods can give similar estimates.
-Understanding the differences between \Ro estimates is critical to controlling an epidemic as \Ro provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+Understanding the differences (or similarities) between \Ro estimates is critical to controlling an epidemic as \Ro provides information about the level of intervention required to prevent further transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 
-Here, we show that a wide range of approaches to estimating \Ro during the exponential growth period can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
 The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, characterizes the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
-therefore, estimates of \Ro from different models directly depend on their assumptions about the generation-interval distribution and the exponential growth rate.
+therefore, estimates of \Ro directly depend on their assumptions about the generation-interval distribution and the exponential growth rate.
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
@@
 \scriptsize
-\begin{tabular}{l|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{2.4cm}|p{2.7cm}|l}
- & Model & Data & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
+\begin{tabular}{l|p{2cm}|p{3.2cm}|p{2.8cm}|p{2.5cm}|p{2.4cm}|p{2.7cm}|p{2cm}}
+ & Model & Data (study period) & Data source & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
 \hline
-Study 1 & Deterministic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
+Study 1 & Deterministic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020) & Estimated by \cite{imaincov0} & 1.5--3.5 & 10 & 1 & \cite{bedfordncov} \\
 \hline
-Study 2 & Stochastic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020)  & 2.6 (1.5--3.5)$^\ast$ & 8.4 & unspecified$^\dagger$ & \cite{imaincov} \\
+Study 2 & Stochastic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020) & Estimated by \cite{imaincov0}  & 2.6 (1.5--3.5)$^\ast$ & 8.4 & Not reported$^\dagger$ & \cite{imaincov} \\
 \hline
-Study 3 & Poisson offspring distribution model & Confirmed cases from China and other countries\newline (Dec 29, 2019--Jan 23, 2020) & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
+Study 3 & Poisson offspring distribution model & Confirmed cases from China and other countries\newline (Dec 29, 2019--Jan 23, 2020) & Medical records and epidemiological investigations from Guangdong Province, China, and official websites of other regions in China & 2.92 (95\% CI: 2.28--3.67) & 8.4 & 0.2 & \cite{liuncov} \\
 \hline
-Study 4 & Metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model & Reported cases from China\newline (Jan 1--21, 2020) & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
+Study 4 & Metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model & Confirmed cases from China\newline (Jan 1--21, 2020) & Not reported & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
 \hline
-Study 5 & Stochastic branching process model & Estimated total number of cases in Wuhan City, China\newline (Jan 18, 2020) & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
+Study 5 & Stochastic branching process model & Total number of cases in Wuhan City, China\newline (Jan 18, 2020) & Estimated by \cite{imaincov0} & 2.2 (90\% CI: 1.4--3.8) & 7--14 & 0.5 & \cite{riouncov} \\
 \hline
-Study 6 & Exponential growth model & Reported cases from China\newline (Jan 1--21, 2020) & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
+Study 6 & Exponential growth model & Confirmed cases from China\newline (Jan 10--22, 2020) & Wuhan
+Municipal Health Commission, China and National Health Commission of China & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 & 0.2 & \cite{zhaoncov} \\
 \hline
-Study 7 & Incidence Decay and Exponential Adjustment (IDEA) model & Reported cases from Wuhan City, China\newline (Dec 8, 2019--Jan 26, 2020) & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
+Study 7 & Incidence Decay and Exponential Adjustment (IDEA) model & Reported cases from Wuhan City, China\newline (Dec 1, 2019--Jan 26, 2020) & World	Health Organization, National Health Commission of China, Wuhan Municipal	Health Commission, and \cite{huang2020clinical} & 2.0--3.1 & 6--10 & 0 & \cite{majumderncov} \\
 \hline
@@
 \textbf{Summary of the models used and reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
-}
 Model details, estimates of \Ro, and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
@@
 $^\ddagger$The authors presented \Ro estimates under different assumptions regarding the reporting rate; we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
+}
 \end{table}
@@
 We gathered information on estimates of \Ro and their model assumptions from 7 articles that were published online between January 23--26, 2020.
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on \url{nextstrain.org} \citep{bedfordncov}.
-Model details are summarized in Table 1.
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on \url{nextstrain.org} \citep{bedfordncov}.
+Model details and corresponding data are summarized in Table 1.
 
@@
 Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic grows exponentially in the beginning.
-While the IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth when $d \neq 0$ \citep{fisman2013idea}, Study 7 estimates $d=0$ across all parameters they consider.
-When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{wallinga2007generation}:
+The IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth when $d \neq 0$ \citep{fisman2013idea}, but Study 7 estimates $d=0$ across all parameters they consider.
+Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restricion that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
+given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delay of around 5 days \citep{sun2020early}, the majority of reported cases during study periods are likely to have been infected prior to the travel ban.
+
+When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population, via the Euler-Loka equation \citep{wallinga2007generation}:
 \begin{linenomath*}
@@
 
-\subsection{Pooled-estimates}
+\subsection{Pooled estimates}
 
-We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution.
-The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make:
+We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters estimates across different studies come from the same gamma distributions:
 \begin{linenomath*}
@@
 where $\mu_r$, $\mu_G$, $\mu_\kappa$ represent the pooled estimates, and $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations.
+The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make.
 We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
@@
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-In this case, most estimates are based on reported number of cases from China, particularly from Wuhan province.
+In this case, most estimates are primarily driven byreported cases from China, particularly from Wuhan province.
 Differences among estimates are likely to reflect differences in estimation methods and underlying assumptions, rather than epidemiological differences.
@@
 
+\subsection{Markov Chain Monte Carlo}
+
 We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
-Posterior samples are thinned every 1000 steps.
+Posterior samples are thinned every 1000 steps to remove autocorrelation among posterior samples.
 Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
@@
 95\% credible intervals (CI) are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
+
+\subsection{Comparing estimates of the basic reproduction number}
+
+In order to compare estimates of the basic reproduction number \Ro (and particularly their associated uncertainties) across different studies, we need a consistent measure of uncertainty.
+Instead of using reported uncertainty ranges from the original studies, we re-calculate the basic reproduction number from the parameter sets ($r_i$, $\bar G_i$, and $\kappa_i$) for each study using \eref{gamma} and calculate the median and 95\% quantiles.
+We refer to these estimates as the base estimtes.
+We note that the distribution of the basic reproduction number for each study corresponds to the assumed distributions in Table 2 for all studies except for Study 6.
+The assumed distribution in Study 6 does not account for the uncertainty in the meang generation interval $\bar G$, whereas the base estimates acconunt for this uncertainty.
+Furthermore, since the distribution in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles could differ slightly from the reported values.
+We compare the base estimates with \Ro estimates based on the pooled estimates ($\mu_r$, $\mu_G$, $\mu_\kappa$) using \eref{gamma}.
 
@@
 
-\rev{Page 7, Line 38. The authors might give some information about the incubation period of
-COVID-19. It could explain why using the data of 26 January is still acceptable even if it was the
-3rd day of the city lockdown.}
-
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro. 
-For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26 (\fref{eff}A).
+For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 26, 2020 (\fref{eff}A).
 Comparing the estimates that include only some sources of uncertainty to the ``all'' estimate ($\Ro = 3.0$; 95\% CI: 2.1--4.6), we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: two middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, is relatively narrow.
@@
 Uncertainty in the generation-interval dispersion $\kappa$ also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare $\mu_G$ with $\mu_\kappa$).
-For example, reducing $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential, therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
+For example, reducing $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential (see \eref{gamma}), therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
 
@@
 Propagating error only from the growth rate gives very narrow credible intervals in this case. 
-Likewise, propagating errors from the growth rate and the mean generation interval gives wider but still too narrow credible intervals.
+Propagating errors from the mean generation interval or generation-interval dispersion gives wider but still narrow credible intervals.
 
@@
   \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
-We compare estimates of \Ro under five scenarios that propagate different combinations of uncertainties (A) based on our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) and (B) assuming a 4-fold reduction in uncertainty of our pooled estimate of the exponential growth rate (using $(\mu_r + 3\times\mathrm{median}(\mu_r))/4$, instead).
-\textbf{base}: \Ro estimates based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-\textbf{growth rate}: \Ro estimates based on the the posterior distribution of $\mu_r$ while using median estimates of $\mu_G$ and $\mu_\kappa$.
-\textbf{GI mean}: \Ro estimates based on the the posterior distribution of $\mu_G$ while using median estimates of $\mu_r$ and $\mu_\kappa$.
-\textbf{growth rate + GI mean}: \Ro estimates based on the the joint posterior distributions of $\mu_r$ and $\mu_G$ while using a median estimate of $\mu_\kappa$.
-\textbf{all}: \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-Vertical lines represent the 95\% credible intervals.
+We compare estimates of \Ro under five scenarios that propagate different parameter uncertainties (A) based on our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) and (B) assuming a 4-fold reduction in uncertainty of our pooled estimate of the exponential growth rate (using $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$).
+Each uncertainty type represents \Ro estimates based the posterior distributions of one of three parameters ($\mu_r$, $\mu_G$, and $\mu_\kappa$) while using median estimates of two other parameters.
+The `base' type represents \Ro estimate based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+The `all' type represents \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+Points represent the median estimates.
+Vertical error bars represent the 95\% credible intervals.
 }
@@
 
-\rev{Page 8, Line 50. Please clarify the sentence: ”We find that incorporating uncertainties one at a
-time increases the width of the confidence intervals in all but 7 cases”. It is not clear particularly
-at “in all but 7 cases”.}
-We also compare the estimates of \Ro across different studies by 
-replacing their values of $r$, $\bar G$, and $\kappa$ with our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$, respectively) one at a time and recalculating the basic reproductive number \Ro.
-We refer to these estimates as ``substitue estimates''.
-This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
+We note that the pooled estimate of the basic reproduction number ($\Ro = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than
 
 \fref{R0} compares the ``base'' estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 ``substitute'' estimates (3 parameter substitutions $\times$ 7 studies).
+We find that all but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimtates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
+We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
+For example, the $r$-substitute estimate of \Ro for Study 7 is $\Ro = 3.9$ (95\% CI: 2.3--8.8), which is much wider than the uncertainty range they reported (2.0--3.1).
+This is consistent with our earlier results (\fref{eff}) that demonstrated the importance of accounting for uncertainty in the exponential growth rate $r$.
 
-We find that incorporating uncertainties one at a time increases the width of the credible intervals in all but 7 cases.
-We estimate narrower credible intervals for Study 3, Study 6, and Study 7 when we account for proper uncertainties in the generation-interval dispersion because they assume a narrow generation-interval distribution (compare ``base'' with ``generation-interval dispersion'');
-when higher values of $\kappa$ are used, their estimates of \Ro become less sensitive to the values of $r$ and $\bar G$, giving narrower credible intervals.
-We estimate narrower credible intervals for Study 5 and Study 7 when we account for proper uncertainties in the mean generation interval (compare ``base'' with ``mean generation interval'') because the range of uncertainty in the mean generation interval $\bar G$ they consider is much wider than the pooled range (\fref{assumption}).
-Substituting the reported $r$ or $\bar G$ from Study 1 with our pooled estimates give narrower credible intervals for similar reasons.
-
-We find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
-For example, recalculating \Ro for Study 7 by using our pooled estimate of $r$ gives $\Ro = 3.9$ (95\% CI: 2.3--8.6), which is much wider than the uncertainty range they reported (2.0--3.1).
-There are two explanations for this result.
-First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical mathematical effects on \Ro in our framework (\eref{gamma} in Methods),
-$r$ is more influential in this case because it is associated with more uncertainty (\fref{assumption}).
-Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$.
-One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive and thus tend to have particularly narrow credible intervals.
+% There are two explanations for this result.
+% First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical mathematical effects on \Ro in our framework (\eref{gamma} in Methods),
+% $r$ is more influential in this case because it is associated with more uncertainty (\fref{assumption}).
+% Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$.
+% One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
+% This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive and thus tend to have particularly narrow credible intervals.
 
@@
 
-Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate \Ro and compare it with the reported \Ro estimates.
-Our estimated \Ro from the pooled distribution has a median of 2.9 (95\% CI: 2.1--4.5).
-While the point estimate of \Ro is similar to other reported values from this date range, the credible intervals are wider than all but one study.
-This result does not imply that assumptions based on the pooled estimate are too weak;
-we believe that this credible interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
-In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
-Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
-We note in particular that, while the baseline example we used from Study 6 may appear to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
-Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
+%Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate \Ro and compare it with the reported \Ro estimates.
+%Our estimated \Ro from the pooled distribution has a median of 2.9 (95\% CI: 2.1--4.5).
+%While the point estimate of \Ro is similar to other reported values from this date range, the credible intervals are wider than all but one study.
+%This result does not imply that assumptions based on the pooled estimate are too weak;
+%we believe that this credible interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+%In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
+%Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
+%We note in particular that, while the baseline example we used from Study 6 may appear to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
+%Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
 

### 2020-05-03 compare_R0.R

@@ -49,7 +49,8 @@ R0all %>%
   merge(summarize(group_by(Rdata_adj, study), bwidth=upr-lwr)) %>%
   mutate(
     ww=width>bwidth
-  )
+  ) %>%
+  filter(!ww)
 
 R0all %>%
   filter(type=="all") %>%

### 2020-05-04 review/response.tex
almost done!!!

@@
 
-We might have inadvertently confused reviewer 2.
+We might have inadvertently confused reviewer 2 to some extent.
 Our method does not rely on any data or dynamical models but the reviewer asks us to compare our results with the reported number of cases (this is not possible as we do not make any predictions about the epidemic trajectory).
-We use a multi-level model to calculate pooled estimates of parameters that have been estimated or assumed from other studies.
+We use a multi-level model to calculate pooled estimates of parameters that have already been estimated or assumed from other studies.
 Then, we use the pooled estimates to calculate $\mathcal R_0$.
@@
 
+Done.
+
 \revtext{Please clearly explain the formula in the caption of Figure 2 (Page 8, line 33) in the Methods
 section.}
+
+Done.
 
@@
 
+Done.
+
 \revtext{The authors consider that all the 7 studies have the same weight in the estimation of basic
@@
 based on the confidence of the estimation of R0.}
+
+Done.
 
@@
 
+Done.
+
 \revtext{The authors do not provide the validation on their estimates, e.g., a posterior predictive check
@@
 
+Done.
+
 \revtext{In the first paragraph of the Results page 7 the value of the median and the CI of the 3
 parameters must be presented.}
+
+Done.
 
@@
 
+Done.
+
 \revtext{The uncertainty related to kappa is not described in Figure 2. Even the authors mentioned in
@@
 
+Done.
+
 \revtext{Discussions}
@@
 \revtext{In the Results section, the authors occasionally discuss the results, i.e., Page 8, line 50 to Page
-9, line 23. The opinions should be in the Discussion section.
-In the Discussion section, the authors should discuss the quality of the reported estimates from
+9, line 23. The opinions should be in the Discussion section.}
+
+Done.
+
+\revtext{In the Discussion section, the authors should discuss the quality of the reported estimates from
 7 studies used in this subject. Some of them do not provide detailed methods and assumptions.
@@
 R0.}
+
+Done.
 
@@
 
+Some of the issues raised here (e.g., day of the week effect and limited diagnostic resources) were already mentioned in the Discussion of the previous version of the manuscript.
+We now mention other points as well.
+
 \revtext{Page 4, Table 1. Please check Study 4, the article has been revised since 27 January 2020,
 and their estimates also changed.}
+
+Done.
 
@@
 
+Done.
+
 \revtext{Page 6, Line 29. Are the priors for $\mu_r$, $\mu_G$, and $\mu_\kappa$ weakly informative? Did the
 authors check the sensibility with new priors?}
+
+Done.
 
@@
 
+Done.
+
 \revtext{Page 7, Line 20. Figure 1 - Caption: Did the authors test the sensitivity of the value of kappa of
 Study 2 (e.g., 0 or 1) on the pooled estimate?}
+
+Done.
 
@@
 
+Done.
+
 \revtext{Page 8, Line 50. Please clarify the sentence: ”We find that incorporating uncertainties one at a
@@
 at “in all but 7 cases”.}
+
+Done.
 

### 2020-05-04 ncov.tex
almost done!!!

@@
 \newcommand{\Ro}{\ensuremath{{\mathcal R}_{0}}\xspace}
+\newcommand{\Rpool}{\ensuremath{{\mathcal R}_{\textrm{\tiny{pool}}}}\xspace}
 \newcommand{\RR}{\ensuremath{{\mathcal R}}}
@@
 We also use our framework to tease apart which assumptions of these different models led to their different estimates and the associated degrees of uncertainty.
-Despite the availability of more recent and/or updated estimates of \Ro, we restrict ourselves to the estimates above in order to focus on the resolution of uncertainty in the earliest stages of an epidemic.
 
@@
 
-\subsection{Comparing model assumptions}
+\subsection{Model assumptions}
 
@@
 
-%While some studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}, we strictly focus on estimates that were published between January 23--26, 2020.
-%The main purpose of this study is to illustrate that estimates of \Ro from different models can be understood by comparing their underlying assumptions about the exponential growth rates and generation-interval distributions, rather than to provide more precise or accurate estimate of \Ro.
+While some studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}, we strictly focus on estimates that were published between January 23--26, 2020 for two reasons:
+(i) to focus on the resolution of uncertainty in the earliest stages of an epidemic and
+(ii) to illustrate that estimates of \Ro from different models can be understood by comparing their underlying assumptions about the exponential growth rates and generation-interval distributions.
+We do not aim to provide more precise or accurate estimate of \Ro.
 
@@
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-In this case, most estimates are primarily driven byreported cases from China, particularly from Wuhan province.
+In this case, most estimates primarily depend on by reported cases from China, particularly from Hubei province.
 Differences among estimates are likely to reflect differences in estimation methods and underlying assumptions, rather than epidemiological differences.
-The pooled estimates can become sharper as we add more models even when the models no longer add more information about the epidemic.
+The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models no longer add more information about the epidemic.
 Thus, the pooled estimator should be interpreted with care.
@@
 Furthermore, since the distribution in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles could differ slightly from the reported values.
-We compare the base estimates with \Ro estimates based on the pooled estimates ($\mu_r$, $\mu_G$, $\mu_\kappa$) using \eref{gamma}.
+We compare the base estimates with pooled estimate the basic reproduction number \Rpool based on the pooled estimates of underying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
 
@@
 We find that there is a large uncertainty associated with the underlying parameters;
-many models rely on stronger assumptions that ignore these uncertainties.
-Surprisingly, no studies take into account how generation-interval dispersion affects their estimates of \Ro:
+most studies consider narrower ranges of uncertainties in these parameters.
+No studies take into account how generation-interval dispersion affects their estimates of \Ro:
 all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
-% Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
 
@@
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro. 
-For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 26, 2020 (\fref{eff}A).
-Comparing the estimates that include only some sources of uncertainty to the ``all'' estimate ($\Ro = 3.0$; 95\% CI: 2.1--4.6), we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: two middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, is relatively narrow.
+For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26, 2020 (\fref{eff}A).
+Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: two middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, is relatively narrow.
 In this case, propagating error from the mean generation interval has negligible effect compared to propagating the uncertainty in $r$.
@@
 For example, reducing $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential (see \eref{gamma}), therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
+We note that our estimate of \Rpool is relatively insensitive to our assumption of $\kappa=0.5$ for Study 2: assuming $\kappa=0.1$ gives $\Rpool = 3.0$ (95\% CI: 2.2--4.7), whereas assuming $\kappa=0.9$ gives $\Rpool = 2.9$ (95\% CI: 2.1--4.4).
 
@@
 
-We note that the pooled estimate of the basic reproduction number ($\Ro = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than
-
-\fref{R0} compares the ``base'' estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 ``substitute'' estimates (3 parameter substitutions $\times$ 7 studies).
-We find that all but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimtates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
-We also find that accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
+Finally, \fref{R0} compares the ``base'' estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
+All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimtates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
+Furthermore, accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
 For example, the $r$-substitute estimate of \Ro for Study 7 is $\Ro = 3.9$ (95\% CI: 2.3--8.8), which is much wider than the uncertainty range they reported (2.0--3.1).
 This is consistent with our earlier results (\fref{eff}) that demonstrated the importance of accounting for uncertainty in the exponential growth rate $r$.
-
-% There are two explanations for this result.
-% First, even though the exponential growth rate $r$ and the mean generation interval $\bar G$ have identical mathematical effects on \Ro in our framework (\eref{gamma} in Methods),
-% $r$ is more influential in this case because it is associated with more uncertainty (\fref{assumption}).
-% Second, assuming a fixed generation interval ($\kappa=0$) makes the estimate of \Ro too sensitive to $r$ and $\bar G$.
-% One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-% This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive and thus tend to have particularly narrow credible intervals.
+We also note that the pooled estimate of the basic reproduction number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates except for Study 2.
 
@@
 
-%Finally, we incorporate all uncertainties by using posterior samples for $\mu_r$, $\mu_G$, and $\mu_\kappa$ to recalculate \Ro and compare it with the reported \Ro estimates.
-%Our estimated \Ro from the pooled distribution has a median of 2.9 (95\% CI: 2.1--4.5).
-%While the point estimate of \Ro is similar to other reported values from this date range, the credible intervals are wider than all but one study.
-%This result does not imply that assumptions based on the pooled estimate are too weak;
-%we believe that this credible interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
-%In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
-%Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
-%We note in particular that, while the baseline example we used from Study 6 may appear to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
-%Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
-
 \section{Discussion}
 
-\rev{In the Results section, the authors occasionally discuss the results, i.e., Page 8, line 50 to Page
-9, line 23. The opinions should be in the Discussion section.
-In the Discussion section, the authors should discuss the quality of the reported estimates from
-7 studies used in this subject. Some of them do not provide detailed methods and assumptions.
-Also, most of them are preprint articles. It could be another important source of uncertainty on
-R0.}
+Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
+Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the SARS-CoV-2 outbreak.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals.
 
-\rev{Although the authors are in the case of the COVID-19 epidemic in Hubei before any
-confinement, the authors do not mention in the introduction or in the discussion the other
-sources of uncertainty linked to the data (such as the variations in diagnostic results related to
-the day of the week, a saturation of diagnostic test capacity, transparency of data,
-representativeness of samples and improvement of detection capacity as time goes by.}
+Our analysis shows that many early estimates of \Ro rely on strong assumptions.
+The lack of uncertainties in the generation-interval dispersion is particularly surprising because it determines the shape of the $r$-\Ro relationship (\fref{assumption});
+assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
+The lack of uncertainties in the generation-interval dispersion further explains the sensitivity of \Ro estimates to the exponential growth rate, particularly in Study 7 (\fref{R0}).
+Since Study 7 assumes a fixed generation interval ($\kappa=0$), they implicitly assume an exponential $r$--\Ro relationship, making their estimate too sensitive to $r$.
+One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
+This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive --- implicitly assuming a linear $r$--\Ro relationship --- and thus tend to have particularly narrow credible intervals.
 
-\rev{Page 4, Table 1. Please check Study 4, the article has been revised since 27 January 2020,
-and their estimates also changed.}
-
-\rev{Page 6, Table 2. Please recheck the parameters of Study 5, which is inconsistent with the 90%
-confidence interval in Table 1. The 90\% confidence interval for a Gamma distribution with mean
-2.2 and shape 12 is 1.27 to 3.34, while the interval in Table 1 is 1.4 to 3.8.}
-
-Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
-Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals.
-Our analysis of individual studies shows that many early estimates of \Ro rely on strong assumptions.
+As most studies rely on strong assumptions, the credible intervals associated with the base \Ro estimates are necessarily narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
+While the point estimate of \Rpool is similar to other reported values from this date range, the credible intervals are wider than all but one study.
+This result does not imply that assumptions based on the pooled estimate are too weak;
+we believe that this credible interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
+Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
+We note in particular that, while the baseline example we used from Study 6 may appear to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
+Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
 
@@
 This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
+Narrow uncertainties in the estimates of the exponential growth rate are likely to depend on this assumption.
 Naive fits to cumulative incidence data should therefore be avoided.
-\rev{
-At the beginning of the Discussion section authors list a series of important issues and possible sources of errors that often limit the utility of early outbreak analyses (e.g.
-fitting the cumulative curve of cases, assuming a fixed generation interval when in fact it varies during the outbreak, changing reporting rate, etc.).
-Compared with these issues the main focus of the paper (not properly accounting for the uncertainty in the generation time and growth rate parameters) appears a small caveat.
-Maybe author could give a sense of the importance of the element they are focusing on with the respect to the other issues?
-}
 
@@
 Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated credible intervals.
-Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
-Accounting for these factors is crucial for making accurate inferences.
+Other factors related to reporting --- including changes in case definition, saturation in diagnostic test capacity, transparency of data, and representativeness of samples --- will also affect the inference.
+Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects};
+since we are primarily focusing on the outbreak in Wuhan city before confinement, generation intervals are unlikely to vary significantly.
+All of these factors affect the estimation of the exponential growth rate, which, in turn, will affect the estimation of the basic reproduction number.
 
-Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26).
+Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26, 2020).
+Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates have not been finalized.
+For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020: $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
+we did not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
+Some studies also do not provide detailed description of their methods, data, and/or assumptions.
+The quality of these analyses adds further uncertainty to their results that are not captured by their uncertainty quantification (e.g., reported credible intervals) or by our analysis.
+While pre-prints can play an important role in disseminating information, they should be interpreted with care.
+
 During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.

### 2020-05-04 ncov.tex
maybe time to diff

@@
 While estimates of \Ro and associated confidence intervals for Study 6 in Table 1 represents estimates of \Ro based on $\bar G = 8\,\mathrm{days}$, we account for the uncertainty they consider for $\bar G$ in our formal analysis.
-\rev{Page 7, Line 20. Figure 1 - Caption: Did the authors test the sensitivity of the value of kappa of
-Study 2 (e.g., 0 or 1) on the pooled estimate?}
 
@@
 \section{Results}
-
-\rev{In Figures 2 and 3, the authors replace the reported parameters (r, $\bar G$, and kappa) with the
-pooled estimates mu ($\mu_r$, $\mu_G$, and $\mu_\kappa$). The authors do not give information about
-the other sources of uncertainty linked with the variation of parameter sigma ($\sigma_r$, $\sigma_G$,
-$\sigma_\kappa$).}
 

### 2020-05-04 ncov.tex
weird typos

@@
 The IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth when $d \neq 0$ \citep{fisman2013idea}, but Study 7 estimates $d=0$ across all parameters they consider.
-Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restricion that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
+Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
 given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delay of around 5 days \citep{sun2020early}, the majority of reported cases during study periods are likely to have been infected prior to the travel ban.
 
-When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population, via the Euler-Loka equation \citep{wallinga2007generation}:
+When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population, via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{linenomath*}
@@
 While most studies report confidence or credible intervals to quantify uncertainties associated with their estimates, some rely on different measures.
-In particular, study 2 reports a range of \Ro for the worst and best case scenarios, which correspond to the values of \Ro such that 95\% and 5\% of the simulated total number of cases by Janaury 18, 2020 is greater than or equal to 4000, respectively;
+In particular, study 2 reports a range of \Ro for the worst and best case scenarios, which correspond to the values of \Ro such that 95\% and 5\% of the simulated total number of cases by January 18, 2020 is greater than or equal to 4000, respectively;
 for brevity, we treat these intervals as a 90\% credible interval in our analysis.
@@
 As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that correspond to their model assumptions.
-Since the estimates of the basic reproductive number \Ro entirely depends on the exponential growth rate and the shape of generation-interval distributions, we can calulate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+Since the estimates of the basic reproductive number \Ro entirely depends on the exponential growth rate and the shape of generation-interval distributions, we can calculate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
@@
 Instead of using reported uncertainty ranges from the original studies, we re-calculate the basic reproduction number from the parameter sets ($r_i$, $\bar G_i$, and $\kappa_i$) for each study using \eref{gamma} and calculate the median and 95\% quantiles.
-We refer to these estimates as the base estimtes.
+We refer to these estimates as the base estimates.
 We note that the distribution of the basic reproduction number for each study corresponds to the assumed distributions in Table 2 for all studies except for Study 6.
-The assumed distribution in Study 6 does not account for the uncertainty in the meang generation interval $\bar G$, whereas the base estimates acconunt for this uncertainty.
+The assumed distribution in Study 6 does not account for the uncertainty in the mean generation interval $\bar G$, whereas the base estimates account for this uncertainty.
 Furthermore, since the distribution in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles could differ slightly from the reported values.
-We compare the base estimates with pooled estimate the basic reproduction number \Rpool based on the pooled estimates of underying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
+We compare the base estimates with pooled estimate the basic reproduction number \Rpool based on the pooled estimates of underlying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
 
@@
 Finally, \fref{R0} compares the ``base'' estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
-All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimtates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
+All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
 Furthermore, accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).

### 2020-05-04 ncov.tex
last paragraph!!!!????

@@
 
+\cite{flaxman2020estimating} recently estimated the basic reproduction number for SARS-CoV-2 outbreaks in 11 European countries to be around 3.87 (3.01--4.66), on average.
+While these estimates appear to be broadly consistent with to earlier estimates from China, comparing the exponential growth rate and the underlying generation-interval distributions suggest otherwise.
+They assumed a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
+based on these values, the exponential growth rate has to be much higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ compared to the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).
+We caution that naively comparing estimates of the basic reproduction number without accounting for differences in underlying assumptions can underestimate the differences in the estimates.
+
 In summary, we have provided a basis for comparing exponential-growth based estimates of \Ro and its associated uncertainty in terms of three components: the exponential growth rate, mean generation interval, and generation interval dispersion.

### 2020-05-04 figure2.tex
am i done?

@@
-% Created by tikzDevice version 0.12.3 on 2020-04-27 21:17:30
+% Created by tikzDevice version 0.12.3 on 2020-05-04 03:09:56
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,289.08);
+\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,289.08);
 \end{scope}
 \begin{scope}
-\path[clip] ( 37.26, 33.48) rectangle (283.08,264.36);
+\path[clip] ( 49.68, 44.63) rectangle (281.08,256.12);
 \definecolor{fillColor}{RGB}{255,255,255}
 
-\path[fill=fillColor] ( 37.26, 33.48) rectangle (283.08,264.36);
+\path[fill=fillColor] ( 49.68, 44.63) rectangle (281.08,256.12);
 \definecolor{drawColor}{gray}{0.92}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
-	(283.08, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
+	(281.08, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 77.88) --
-	(283.08, 77.88);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 85.31) --
+	(281.08, 85.31);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,122.28) --
-	(283.08,122.28);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,125.98) --
+	(281.08,125.98);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,166.68) --
-	(283.08,166.68);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,166.65) --
+	(281.08,166.65);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,211.08) --
-	(283.08,211.08);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,207.32) --
+	(281.08,207.32);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26,255.48) --
-	(283.08,255.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,247.99) --
+	(281.08,247.99);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 65.62,122.88) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.38,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (112.90,122.88) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (120.88,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (160.17,122.88) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (165.38,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (207.44,122.88) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (209.88,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.72,119.58) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.38,123.50) circle (  5.71);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
-	( 65.62,122.88);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
+	( 76.38,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
-	( 65.62,122.88);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
+	( 76.38,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 65.62,122.88) --
-	( 65.62,122.88);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
+	( 76.38,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90,234.50) --
-	(112.90,234.50);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,228.77) --
+	(120.88,228.77);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90,234.50) --
-	(112.90, 55.99);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,228.77) --
+	(120.88, 65.26);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (112.90, 55.99) --
-	(112.90, 55.99);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88, 65.26) --
+	(120.88, 65.26);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17,153.48) --
-	(160.17,153.48);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,154.56) --
+	(165.38,154.56);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17,153.48) --
-	(160.17, 99.27);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,154.56) --
+	(165.38,104.90);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (160.17, 99.27) --
-	(160.17, 99.27);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,104.90) --
+	(165.38,104.90);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44,164.15) --
-	(207.44,164.15);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,164.33) --
+	(209.88,164.33);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44,164.15) --
-	(207.44, 68.61);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,164.33) --
+	(209.88, 76.82);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (207.44, 68.61) --
-	(207.44, 68.61);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88, 76.82) --
+	(209.88, 76.82);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72,260.46) --
-	(254.72,260.46);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,252.55) --
+	(254.38,252.55);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72,260.46) --
-	(254.72, 41.21);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,252.55) --
+	(254.38, 51.71);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (254.72, 41.21) --
-	(254.72, 41.21);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38, 51.71) --
+	(254.38, 51.71);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
-	( 37.26,264.36);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
+	( 49.68,256.12);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86, 30.17) {2.0};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {2.0};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86, 74.57) {2.5};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 80.90) {2.5};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,118.97) {3.0};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,121.57) {3.0};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,163.37) {3.5};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,162.24) {3.5};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,207.78) {4.0};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,202.91) {4.0};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 31.86,252.18) {4.5};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,243.58) {4.5};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26, 33.48) --
-	( 37.26, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
+	( 49.68, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26, 77.88) --
-	( 37.26, 77.88);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 85.31) --
+	( 49.68, 85.31);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,122.28) --
-	( 37.26,122.28);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,125.98) --
+	( 49.68,125.98);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,166.68) --
-	( 37.26,166.68);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,166.65) --
+	( 49.68,166.65);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,211.08) --
-	( 37.26,211.08);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,207.32) --
+	( 49.68,207.32);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.26,255.48) --
-	( 37.26,255.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,247.99) --
+	( 49.68,247.99);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.26, 33.48) --
-	(283.08, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
+	(281.08, 44.63);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 65.62, 30.48) --
-	( 65.62, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 76.38, 40.63) --
+	( 76.38, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (112.90, 30.48) --
-	(112.90, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (120.88, 40.63) --
+	(120.88, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (160.17, 30.48) --
-	(160.17, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (165.38, 40.63) --
+	(165.38, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (207.44, 30.48) --
-	(207.44, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (209.88, 40.63) --
+	(209.88, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (254.72, 30.48) --
-	(254.72, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (254.38, 40.63) --
+	(254.38, 44.63);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 65.62, 21.46) {base};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 76.38, 28.62) {none};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (112.90, 21.46) {$\mu_r$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (120.88, 28.62) {$\mu_r$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (160.17, 21.46) {$\mu_G$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (165.38, 28.62) {$\mu_G$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (207.44, 21.46) {$\mu_\kappa$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (209.88, 28.62) {$\mu_\kappa$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (254.72, 21.46) {all};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (254.38, 28.62) {all};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (160.17,  8.33) {Uncertainty type};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (165.38, 11.11) {Uncertainty type};
 \end{scope}
@@
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 14.26,148.92) {Basic reproductive number};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,150.38) {Basic reproductive number};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.44] at ( 37.26,273.16) {A. Baseline};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at ( 49.68,267.86) {A. Baseline};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,289.08);
+\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,289.08);
 \end{scope}
 \begin{scope}
-\path[clip] (326.34, 33.48) rectangle (572.16,264.36);
+\path[clip] (338.76, 44.63) rectangle (570.16,256.12);
 \definecolor{fillColor}{RGB}{255,255,255}
 
-\path[fill=fillColor] (326.34, 33.48) rectangle (572.16,264.36);
+\path[fill=fillColor] (338.76, 44.63) rectangle (570.16,256.12);
 \definecolor{drawColor}{gray}{0.92}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 79.65) --
-	(572.16, 79.65);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 86.93) --
+	(570.16, 86.93);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34,141.22) --
-	(572.16,141.22);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,143.33) --
+	(570.16,143.33);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34,202.79) --
-	(572.16,202.79);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,199.73) --
+	(570.16,199.73);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (354.70,142.26) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (365.46,144.28) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (401.98,142.26) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (409.96,144.28) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (449.25,142.26) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (454.46,144.28) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (496.52,142.26) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (498.96,144.28) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (543.80,140.42) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (543.46,142.60) circle (  5.71);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
-	(354.70,142.26);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,144.28) --
+	(365.46,144.28);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
-	(354.70,142.26);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,144.28) --
+	(365.46,144.28);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (354.70,142.26) --
-	(354.70,142.26);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,144.28) --
+	(365.46,144.28);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,187.46) --
-	(401.98,187.46);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,185.68) --
+	(409.96,185.68);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,187.46) --
-	(401.98,111.70);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,185.68) --
+	(409.96,116.29);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (401.98,111.70) --
-	(401.98,111.70);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,116.29) --
+	(409.96,116.29);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,195.31) --
-	(449.25,195.31);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,192.88) --
+	(454.46,192.88);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,195.31) --
-	(449.25,101.35);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,192.88) --
+	(454.46,106.80);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (449.25,101.35) --
-	(449.25,101.35);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,106.80) --
+	(454.46,106.80);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52,213.80) --
-	(496.52,213.80);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,209.81) --
+	(498.96,209.81);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52,213.80) --
-	(496.52, 48.20);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,209.81) --
+	(498.96, 58.12);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (496.52, 48.20) --
-	(496.52, 48.20);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96, 58.12) --
+	(498.96, 58.12);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80,261.16) --
-	(543.80,261.16);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,253.19) --
+	(543.46,253.19);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80,261.16) --
-	(543.80, 41.24);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,253.19) --
+	(543.46, 51.75);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (543.80, 41.24) --
-	(543.80, 41.24);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46, 51.75) --
+	(543.46, 51.75);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 33.48) --
-	(326.34,264.36);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 44.63) --
+	(338.76,256.12);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94, 76.35) {2.6};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56, 82.52) {2.6};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94,137.92) {3.0};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,138.92) {3.0};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at (320.94,199.49) {3.4};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,195.32) {3.4};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34, 79.65) --
-	(326.34, 79.65);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76, 86.93) --
+	(338.76, 86.93);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34,141.22) --
-	(326.34,141.22);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,143.33) --
+	(338.76,143.33);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (323.34,202.79) --
-	(326.34,202.79);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,199.73) --
+	(338.76,199.73);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (326.34, 33.48) --
-	(572.16, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 44.63) --
+	(570.16, 44.63);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (354.70, 30.48) --
-	(354.70, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (365.46, 40.63) --
+	(365.46, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (401.98, 30.48) --
-	(401.98, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (409.96, 40.63) --
+	(409.96, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (449.25, 30.48) --
-	(449.25, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (454.46, 40.63) --
+	(454.46, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (496.52, 30.48) --
-	(496.52, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (498.96, 40.63) --
+	(498.96, 44.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (543.80, 30.48) --
-	(543.80, 33.48);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (543.46, 40.63) --
+	(543.46, 44.63);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (354.70, 21.46) {base};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (365.46, 28.62) {none};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (401.98, 21.46) {$\hat{\mu}_r$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (409.96, 28.62) {$\hat{\mu}_r$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (449.25, 21.46) {$\mu_G$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (454.46, 28.62) {$\mu_G$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (496.52, 21.46) {$\mu_\kappa$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (498.96, 28.62) {$\mu_\kappa$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (543.80, 21.46) {all};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (543.46, 28.62) {all};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (449.25,  8.33) {Uncertainty type};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (454.46, 11.11) {Uncertainty type};
 \end{scope}
@@
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (303.34,148.92) {Basic reproductive number};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (308.10,150.38) {Basic reproductive number};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.44] at (326.34,273.16) {B. Reduced uncertainty in $r$};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at (338.76,267.86) {B. Reduced uncertainty in $r$};
 \end{scope}

### 2020-05-04 ncov.tex
am i done?

@@
 \section*{Abstract}
-A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat. 
+A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat in December 2019. 
 As the epidemic progresses, disease modelers continue to focus on estimating the basic reproductive number \Ro\ --- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
-The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
+The modeling approaches and resulting estimates of \Ro during the beginning of the outbreak vary widely, despite relying on similar data sources.
 Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
@@
 
-Since December 2019, a novel coronavirus (SARS-CoV-2) has been spreading in many parts of the world \citep{pneumonia}.
+Since December 2019, a novel coronavirus (SARS-CoV-2) has been spreading globally \citep{pneumonia}.
 Although the virus is likely to have originated from animal hosts \citep{andersen2020proximal}, the ability of SARS-CoV-2 to directly transmit between humans, particularly without symptoms, has posed a greater threat for its spread \citep{he2020temporal}.
-As of March 2, 2020, more than 3 million cases of the coronavirus disease 2019 (COVID-19) have been confirmed internationally \citep{who103}.
+As of May 2, 2020, more than 3 million cases of the coronavirus disease 2019 (COVID-19) have been confirmed internationally \citep{who103}.
 
@@
 We present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
-Previous studies have directly calculated the average of reported \Ro values but such methods undermine differences in underlying model assumptions and statistical methods \citep{majumder2020early, liu2020reproductive}.
+Previous studies have directly calculated the average of reported \Ro values  \citep{majumder2020early, liu2020reproductive} but such methods can undermine differences in underlying model assumptions and statistical methods.
 Instead, we decompose estimated \Ro into three key quantities ($r$, $\bar G$, and $\kappa$) and calculate the average (pooled estimates) of these key parameters.
@@
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
-We also use our framework to tease apart which assumptions of these different models led to their different estimates and the associated degrees of uncertainty.
+% We also use our framework to tease apart which assumptions of these different models led to their different estimates and the associated degrees of uncertainty.
 
@@
 \hline
-Study 4 & Metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model & Confirmed cases from China\newline (Jan 1--21, 2020) & Not reported & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
+Study 4 & Metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model & Confirmed cases from China and other countries\newline (Jan 1--21, 2020) & Not reported & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
 \hline
@@
 \caption{
-\textbf{Summary of the models used and reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
+\textbf{Summary of the models, analyzed data, reported estimates of the basic reproductive number, and the assumptions about the generation-interval distributions.}
 Model details, estimates of \Ro, and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
@@
 In other words, all model assumptions come down to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence intervals associated with the exponential growth rates or parameters of the generation-interval distributions will be necessarily narrow.
+For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will be necessarily narrow.
 Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.
 
-As most studies do not report their estimates of the exponential growth rate, we summarize model outcomes using reported (either estimated or assumed) values of the basic reproductive number \Ro, mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation (Table 1);
-we re-estimate the corresponding exponential growth rates later.
+As most studies do not report their estimates of the exponential growth rate, we first summarize model outcomes using reported (either estimated or assumed) values of the basic reproductive number \Ro, mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation (Table 1);
+we re-estimate the corresponding exponential growth rates from these values later.
 Study 2 only reports their assumptions about the mean generation interval; for brevity, we assume $\kappa = 0.5$ in our analysis.
@@
 we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
-While estimates of \Ro and associated confidence intervals for Study 6 in Table 1 represents estimates of \Ro based on $\bar G = 8\,\mathrm{days}$, we account for the uncertainty they consider for $\bar G$ in our formal analysis.
+While estimates of \Ro and the associated confidence intervals for Study 6 in Table 1 represents estimates of \Ro based on $\bar G = 8\,\mathrm{days}$, we account for the uncertainty they consider for $\bar G$ in our formal analysis.
 
@@
 
-While some studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}, we strictly focus on estimates that were published between January 23--26, 2020 for two reasons:
-(i) to focus on the resolution of uncertainty in the earliest stages of an epidemic and
-(ii) to illustrate that estimates of \Ro from different models can be understood by comparing their underlying assumptions about the exponential growth rates and generation-interval distributions.
-We do not aim to provide more precise or accurate estimate of \Ro.
+Some studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+As our primary focus is on the resolution of uncertainty in the available information during the earliest stages of an epidemic, rather than to provide more precise or accurate estimate of \Ro, we strictly focus on estimates that were published between January 23--26, 2020.
 
@@
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
-We used gamma distributions to model values reported with confidence/credible intervals and uniform distributions to model values reported with ranges;
-when confidence/credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimate value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence/credible limits is equal to the reported confidence/credible level.
+We used gamma distributions to model values reported with confidence or credible intervals and uniform distributions to model values reported with ranges;
+when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
 This probability does not need to be equi-tailed.
@@
 
-For each study $i$, we construct a family of parameter sets by drawing $10^5$ random samples from the probability distributions (Table 2) that represent the estimates of $(\Ro)_{i,m}$ and the assumed values of $\bar G_{i,m}$ and $\kappa_{i,m}$ and calculate the exponential growth rate $r_{i,m}$ via the inverse of \eref{gamma}:
+For each study $i$, we construct a family of parameter sets by drawing $10^5$ random samples from the corresponding probability distributions (Table 2) that represent the estimates of $(\Ro)_{i,m}$ and the assumed values of $\bar G_{i,m}$ and $\kappa_{i,m}$ and calculate the exponential growth rate $r_{i,m}$ via the inverse of \eref{gamma}:
 \begin{linenomath*}
@@
 This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.
-Uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.
-Then, the set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ defines an empirical parameter distribution for each study without having to re-fit each model.
+We note that uncertainties in the probability distributions that we calculate for the estimated exponential growth rates reflects model assumptions, statistical methods, as well as quality of the data that each study relies on.
+Then, the the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ defines an empirical parameter distribution for each study $i$ without having to re-fit each model.
 
 For most cases, we model $(\Ro)_i$, $\bar G_i$, and $\kappa_i$ using independent distributions (as specified in Table 2) to calculate $r_i$.
-In practice, this assumption does not affect our analysis because we are re-calculating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
+This assumption does not affect our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
 in other words, we still obtain the same estimates and the associated uncertainty of $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.
-For study 6, we fix $\bar G=8\,\textrm{days}$ and use the gamma distribution (Table 2) that corresponds to $\mathcal R_0 = 5.47$ (95\% CI: 4.16--7.10) during the re-estimation step for $r$ to remain consistent with the original study that assumed $\bar G=8\,\textrm{days}$ to obtain this particular estimate;
-the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like) days.
+For study 6, we fix $\bar G=8\,\textrm{days}$ and use the gamma distribution (Table 2) that corresponds to $\mathcal R_0 = 5.47$ (95\% CI: 4.16--7.10) during the re-estimation step for $r$ to remain consistent with the original study that assumed $\bar G=8\,\textrm{days}$ to obtain this particular estimate.
 We account for uncertainties in $\bar G$ for Study 6 (Table 1) in all other steps in order to properly incorporate parameter uncertainties in the estimate of \Ro.
@@
 \scriptsize
-\begin{tabular}{l|p{4.5cm}|p{2.5cm}|p{2.7cm}}
+\begin{tabular}{l|p{5cm}|p{2.5cm}|p{2.7cm}}
  & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ \\
@@
 \hline
-Study 2 & $\gammdist(\mathrm{mean}=2.6, \alpha=28)$ & 8.4 & 0.5 \\
+Study 2 & $\gammdist(\mathrm{mean}=2.6, \mathrm{shape}=18)$ & 8.4 & 0.5 \\
 \hline
-Study 3 & $\gammdist(\mathrm{mean}=2.92, \alpha=67)$ & 8.4 & 0.2 \\
+Study 3 & $\gammdist(\mathrm{mean}=2.92, \mathrm{shape}=67)$ & 8.4 & 0.2 \\
 \hline
-Study 4 & $\gammdist(\mathrm{mean}=3.8, \alpha=1400)$ & 7.6 & 0.5 \\
+Study 4 & $\gammdist(\mathrm{mean}=3.8, \mathrm{shape}=1400)$ & 7.6 & 0.5 \\
 \hline
-Study 5 & $\gammdist(\mathrm{mean}=2.2, \alpha=12)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
+Study 5 & $\gammdist(\mathrm{mean}=2.2, \mathrm{shape}=12)$ & $\mathrm{Uniform}(7, 14)$ & 0.5\\
 \hline
-Study 6 & $\gammdist(\mathrm{mean}=5.47, \alpha=54)$ & $\mathrm{Uniform}(7.6, 8.4)^\ast$ & 0.2\\
+Study 6 & $\gammdist(\mathrm{mean}=5.47, \mathrm{shape}=54)$ & $\mathrm{Uniform}(7.6, 8.4)^\ast$ & 0.2\\
 \hline
@@
 We use these probability distributions to obtain a probability distribution for the exponential growth rate $r$.
-The gamma distribution is parameterized by its mean and shape $\alpha$.
+The gamma distribution is parameterized by its mean and shape.
 Constant values are fixed according to Table 1.
-$^\ast$We do not account for this uncertainty during our recalculation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$; the original article reports three $\mathcal R_0$ (and 95\% CIs) estimates using three different values of $\bar G$: 7.6 (MERS-like), 8 (average), and 8.4 (SARS-like).
+$^\ast$We do not account for this uncertainty during our re-estimation of the exponential growth rate $r$ because the reported estimate of $\mathcal R_0$ and its uncertainty assumes $\bar G = 8$.
 We still account for this uncertainty in our pooled estimates ($\mu_G$).
-$^\dagger$Study 7 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
-Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
+$^\dagger$Instead of modeling \Ro with a probability distribution and re-estimating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
 }
@@
 The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make.
-We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
+We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
 Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
@@
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-In this case, most estimates primarily depend on by reported cases from China, particularly from Hubei province.
-Differences among estimates are likely to reflect differences in estimation methods and underlying assumptions, rather than epidemiological differences.
-The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models no longer add more information about the epidemic.
+In this case, most estimates primarily depend on reported cases from China, particularly from Hubei Province.
+Differences among estimates are likely to driven by differences in estimation methods and underlying assumptions, rather than epidemiological differences.
+The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
+Since SARS-CoV-2 had primarily spread in Hubei Province, China during this period, it is not possible to include independent sources of data from other countries.
 Thus, the pooled estimator should be interpreted with care.
@@
 We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
-Posterior samples are thinned every 1000 steps to remove autocorrelation among posterior samples.
+Posterior samples are thinned every 1000 steps to remove autocorrelations among posterior samples.
 Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
@@
 
-\subsection{Comparing estimates of the basic reproduction number}
+\subsection{Comparing estimates of the basic reproductive number}
 
-In order to compare estimates of the basic reproduction number \Ro (and particularly their associated uncertainties) across different studies, we need a consistent measure of uncertainty.
-Instead of using reported uncertainty ranges from the original studies, we re-calculate the basic reproduction number from the parameter sets ($r_i$, $\bar G_i$, and $\kappa_i$) for each study using \eref{gamma} and calculate the median and 95\% quantiles.
+In order to compare estimates of the basic reproductive number \Ro (and particularly their associated uncertainties) across different studies, we need a consistent measure of uncertainty.
+Instead of using reported uncertainty ranges from the original studies, we re-calculate the basic reproductive number from the parameter sets ($r_i$, $\bar G_i$, and $\kappa_i$) for each study using \eref{gamma} and calculate the median and 95\% quantiles.
 We refer to these estimates as the base estimates.
-We note that the distribution of the basic reproduction number for each study corresponds to the assumed distributions in Table 2 for all studies except for Study 6.
-The assumed distribution in Study 6 does not account for the uncertainty in the mean generation interval $\bar G$, whereas the base estimates account for this uncertainty.
-Furthermore, since the distribution in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles could differ slightly from the reported values.
-We compare the base estimates with pooled estimate the basic reproduction number \Rpool based on the pooled estimates of underlying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
+We note that the distribution of the basic reproductive number for each study corresponds to the assumed distributions in Table 2 for all studies except for Study 6.
+The assumed distribution in Study 6 in Table 2 does not account for the uncertainty in the mean generation interval $\bar G$, whereas the base estimates account for this uncertainty.
+Furthermore, since the distribution in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles slightly differ from the reported values.
+We compare the base estimates with pooled estimate the basic reproductive number \Rpool based on the pooled estimates of underlying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
 
@@
 This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
-We compare substitute estimates with the base estimates of \Ro (based on $r_i$, $\bar G_i$, and $\kappa_i$) as well as pooled estimates of \Ro (based on $\mu_r$, $\mu_G$, and $\mu_\kappa$).
+We compare substitute estimates with the base estimates of \Ro (based on $r_i$, $\bar G_i$, and $\kappa_i$).
 
@@
 Each uncertainty type represents \Ro estimates based the posterior distributions of one of three parameters ($\mu_r$, $\mu_G$, and $\mu_\kappa$) while using median estimates of two other parameters.
-The `base' type represents \Ro estimate based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
-The `all' type represents \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+The `none' type represents \Ro estimate based on the median estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$.
+The `all' type represents \Ro estimates based on the joint posterior distributions of  $\mu_r$, $\mu_G$, and $\mu_\kappa$ (also corresponds to \Rpool).
 Points represent the median estimates.
@@
 
-Finally, \fref{R0} compares the ``base'' estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
+Finally, \fref{R0} compares the base estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
 All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
@@
 This is consistent with our earlier results (\fref{eff}) that demonstrated the importance of accounting for uncertainty in the exponential growth rate $r$.
-We also note that the pooled estimate of the basic reproduction number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates except for Study 2.
+We also note that the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates except for Study 2.
 
@@
 \textbf{Sensitivity of the reported \Ro estimates with respect to our pooled estimates of the underlying parameters.}
-We replace the reported parameter values (growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro (\textbf{growth rate}, \textbf{mean generation interval}, and \textbf{generation-interval dispersion}).
-The pooled estimate of \Ro is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$ (\textbf{all});
+We replace the reported parameter values (growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro; we refer to these estimates as substitute estimates.
+The pooled estimate represents \Rpool, which is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$;
 this corresponds to replacing all reported parameter values with our pooled estimates, which gives identical results across all studies.
-Horizontal dashed lines represent the 95\% credible intervals of our pooled estimate of \Ro.
-The reported \Ro estimates (\textbf{base}) have been adjusted to show the approximate 95\% credible interval using the probability distributions that we defined if they had relied on different measures for parameter uncertainties.
+Points represent the medians of our base, substitute, and pooled estimates.
+Horizontal dashed lines represent the 95\% credible intervals of our base, substitute, and pooled estimates.
 }
@@
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
+However, comparing disparate estimates of \Ro can be difficult when they rely on different methods and assumptions.
 Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the SARS-CoV-2 outbreak.
@@
 
-As most studies rely on strong assumptions, the credible intervals associated with the base \Ro estimates are necessarily narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
+As most studies rely on strong assumptions, the credible intervals associated with the base estimates of \Ro are necessarily narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
 While the point estimate of \Rpool is similar to other reported values from this date range, the credible intervals are wider than all but one study.
@@
 since we are primarily focusing on the outbreak in Wuhan city before confinement, generation intervals are unlikely to vary significantly.
-All of these factors affect the estimation of the exponential growth rate, which, in turn, will affect the estimation of the basic reproduction number.
+All of these factors, including fitting to cumulative curves or ignoring process errors, affect the estimation of the exponential growth rate (as well as the associated uncertainties), which in turn affects the estimation of the basic reproductive number.
 
@@
 
-\cite{flaxman2020estimating} recently estimated the basic reproduction number for SARS-CoV-2 outbreaks in 11 European countries to be around 3.87 (3.01--4.66), on average.
+\cite{flaxman2020estimating} recently estimated the basic reproductive number for SARS-CoV-2 outbreaks in 11 European countries to be around 3.87 (3.01--4.66), on average.
 While these estimates appear to be broadly consistent with to earlier estimates from China, comparing the exponential growth rate and the underlying generation-interval distributions suggest otherwise.
@@
 based on these values, the exponential growth rate has to be much higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ compared to the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).
-We caution that naively comparing estimates of the basic reproduction number without accounting for differences in underlying assumptions can underestimate the differences in the estimates.
+We caution that naively comparing estimates of the basic reproductive number without accounting for differences in underlying assumptions can underestimate the differences in the estimates.
 

### 2020-05-04 MCMC_all_kappa_01.R

@@ -1,0 +1,69 @@
+library(MCMCpack)
+load("sample_data.rda")
+
+nsample <- 100000
+
+priorfun <- function(theta) {
+  sum(dnorm(theta[-c(1, 3, 5)], 0, 10, log=TRUE)) +
+    dgamma(theta[1], 2, 2*7, log=TRUE) +
+    dgamma(theta[3], 2, 2/7, log=TRUE) +
+    dgamma(theta[5], 2, 2/0.5, log=TRUE)
+}
+
+sample_data$`Majumder and Mandl (2020)`$kappa <- 0.02
+sample_data$`Imai et al. (2020)`$kappa <- 0.1
+
+postfun <- function(theta) {
+  if (any(theta < 0))
+    return(-Inf)
+  
+  ## theta[2], theta[4], theta[6] are sd
+  mean_r <- theta[1]
+  shape_r <- mean_r^2/theta[2]^2
+  rate_r <- shape_r/mean_r
+  
+  mean_gbar <- theta[3]
+  shape_gbar <- mean_gbar^2/theta[4]^2
+  rate_gbar <- shape_gbar/mean_gbar
+  
+  mean_kappa <- theta[5]
+  shape_kappa <- mean_kappa^2/theta[6]^2
+  rate_kappa <- shape_kappa/mean_kappa
+  
+  ss <- sample(nsample, 1)
+  
+  r <- sapply(sample_data, function(x) x$r[ss])
+  gbar <- sapply(sample_data, function(x) x$gbar[ss])
+  kappa <- sapply(sample_data, function(x) x$kappa[ss])
+  
+  ll <- sum(dgamma(r, shape=shape_r, rate=rate_r, log=TRUE)) +
+    sum(dgamma(gbar, shape=shape_gbar, rate=rate_gbar, log=TRUE)) +
+    sum(dgamma(kappa, shape=shape_kappa, rate=rate_kappa, log=TRUE)) +
+    priorfun(theta)
+  
+  if (is.na(ll) || is.nan(ll)) {
+    return(-Inf)
+  }
+  
+  ll
+}
+
+V <- matrix(0, 6, 6)
+diag(V) <- c(2e-3, 0.004, 0.3, 1, 0.1, 0.1)/10
+
+reslist <- vector('list', 4)
+
+for (i in 1:4) {
+  reslist[[i]] <- MCMCmetrop1R(postfun,
+                               theta.init=c(0.2, 0.1, 8, 1, 0.5, 0.5),
+                               thin=1000,
+                               burnin=500000,
[…]

### 2020-05-04 MCMC_all_kappa_09.R

@@ -1,0 +1,69 @@
+library(MCMCpack)
+load("sample_data.rda")
+
+nsample <- 100000
+
+priorfun <- function(theta) {
+  sum(dnorm(theta[-c(1, 3, 5)], 0, 10, log=TRUE)) +
+    dgamma(theta[1], 2, 2*7, log=TRUE) +
+    dgamma(theta[3], 2, 2/7, log=TRUE) +
+    dgamma(theta[5], 2, 2/0.5, log=TRUE)
+}
+
+sample_data$`Majumder and Mandl (2020)`$kappa <- 0.02
+sample_data$`Imai et al. (2020)`$kappa <- 0.9
+
+postfun <- function(theta) {
+  if (any(theta < 0))
+    return(-Inf)
+  
+  ## theta[2], theta[4], theta[6] are sd
+  mean_r <- theta[1]
+  shape_r <- mean_r^2/theta[2]^2
+  rate_r <- shape_r/mean_r
+  
+  mean_gbar <- theta[3]
+  shape_gbar <- mean_gbar^2/theta[4]^2
+  rate_gbar <- shape_gbar/mean_gbar
+  
+  mean_kappa <- theta[5]
+  shape_kappa <- mean_kappa^2/theta[6]^2
+  rate_kappa <- shape_kappa/mean_kappa
+  
+  ss <- sample(nsample, 1)
+  
+  r <- sapply(sample_data, function(x) x$r[ss])
+  gbar <- sapply(sample_data, function(x) x$gbar[ss])
+  kappa <- sapply(sample_data, function(x) x$kappa[ss])
+  
+  ll <- sum(dgamma(r, shape=shape_r, rate=rate_r, log=TRUE)) +
+    sum(dgamma(gbar, shape=shape_gbar, rate=rate_gbar, log=TRUE)) +
+    sum(dgamma(kappa, shape=shape_kappa, rate=rate_kappa, log=TRUE)) +
+    priorfun(theta)
+  
+  if (is.na(ll) || is.nan(ll)) {
+    return(-Inf)
+  }
+  
+  ll
+}
+
+V <- matrix(0, 6, 6)
+diag(V) <- c(2e-3, 0.01, 0.3, 1.5, 0.1, 0.1)/10
+
+reslist <- vector('list', 4)
+
+for (i in 1:4) {
+  reslist[[i]] <- MCMCmetrop1R(postfun,
+                               theta.init=c(0.2, 0.1, 8, 1, 0.5, 0.5),
[…]

### 2020-05-04 R0_all_kappa_01.R

@@ -1,0 +1,25 @@
+library(coda)
+library(dplyr)
+
+load("sample_data.rda")
+load("MCMC_all_kappa_01.rda")
+
+r <- unlist(unlist(MCMC_all_kappa_01[,1]))
+gbar <- unlist(unlist(MCMC_all_kappa_01[,3]))
+kappa <- unlist(unlist(MCMC_all_kappa_01[,5]))
+
+R0_all <- lapply(sample_data, function(x) {
+  R <- (1+kappa*gbar*r)^(1/kappa)
+  
+  data.frame(
+    est=median(R),
+    lwr=quantile(R, 0.025),
+    upr=quantile(R, 0.975)
+  )
+}) %>%
+  bind_rows(.id="study") %>%
+  mutate(
+    type="all"
+  )
+
+R0_all

### 2020-05-04 R0_all_kappa_09.R

@@ -1,0 +1,25 @@
+library(coda)
+library(dplyr)
+
+load("sample_data.rda")
+load("MCMC_all_kappa_09.rda")
+
+r <- unlist(unlist(MCMC_all_kappa_09[,1]))
+gbar <- unlist(unlist(MCMC_all_kappa_09[,3]))
+kappa <- unlist(unlist(MCMC_all_kappa_09[,5]))
+
+R0_all <- lapply(sample_data, function(x) {
+  R <- (1+kappa*gbar*r)^(1/kappa)
+  
+  data.frame(
+    est=median(R),
+    lwr=quantile(R, 0.025),
+    upr=quantile(R, 0.975)
+  )
+}) %>%
+  bind_rows(.id="study") %>%
+  mutate(
+    type="all"
+  )
+
+R0_all

### 2020-05-04 figure2.R

@@ -1,5 +1,5 @@
 library(coda)
-library(ggplot2); theme_set(theme_bw(base_size=12))
+library(ggplot2); theme_set(theme_bw(base_size=16))
 library(gridExtra)
 library(tikzDevice)
 
@@ -15,7 +15,7 @@ compfun <- function(x) {
     est=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
     lwr=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
     upr=(1+median(kappa)*median(gbar)*median(x))^(1/median(kappa)),
-    type="base"
+    type="none"
   )
   
   dd_r <- data.frame(

### 2020-05-04 review/response.tex
done

@@
 
+\usepackage{xspace}
+
 \usepackage{lmodern}
 \usepackage{amssymb,amsmath}
+
+\usepackage[pdfencoding=auto, psdextra]{hyperref}
+
+\usepackage{natbib}
+\bibliographystyle{chicago}
 
@@
 \newcommand{\Rx}[1]{\ensuremath{{\cal R}_{#1}}} 
-\newcommand{\Ro}{\Rx{0}}
+\newcommand{\Ro}{\ensuremath{{\mathcal R}_{0}}\xspace}
+\newcommand{\Rpool}{\ensuremath{{\mathcal R}_{\textrm{\tiny{pool}}}}\xspace}
 \newcommand{\Reff}{\Rx{\mathit{eff}}}
@@
 
-We have made significant revisions to the manuscript to address this point.
+We have made significant revisions to our manuscript to make it approachable for a wider audience.
 
@@
 So we are effectively considering all model assumptions.
-We have tried to make this point clearer:
+We have tried to make this point clearer in the main text:
 
 ``Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic grows exponentially in the beginning.
-While the IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth, Study 7 estimates $d=0$ across all parameters they consider, in which case the IDEA model becomes equivalent to the exponential growth model.
-When the epidemic is growing exponentially, the estimated basic reproduction number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$:
+The IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth when $d \neq 0$ \citep{fisman2013idea}, but Study 7 estimates $d=0$ across all parameters they consider.
+Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
+given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delay of around 5 days \citep{sun2020early}, the majority of reported cases during study periods are likely to have been infected prior to the travel ban.
+
+When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population, via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{equation}
-1/\Ro = \int \exp(-r\tau) g(\tau) \, d\tau.
+\frac{1}{\Ro} = \int \exp(-r\tau) g(\tau) \, d\tau.
 \label{eq:euler}
@@
 In other words, all model assumptions come down to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence intervals of the model parameters will be necessarily narrow.
-Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and shape of the generation-interval distribution to understand the estimates of the basic reproduction number.''
+For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will be necessarily narrow.
+Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.''
 
@@
 
-???
+Since we are interested in the uncertainty, expressed by probability distributions, the overall probabilities are not affected by the assumptions of gamma or uniform distributions.
+We have added following sentences in the main text:
+
+``This assumption does not affect our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
+in other words, we still obtain the same estimates and the associated uncertainty of $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.''
 
@@
 
-See above.
+See above. We have also added a table to provide a better overview of the models and data used in each study.
 
@@
 
-``Here, we use the gamma approximation framework to the generation-interval distribution to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+``Here, we use the gamma approximation framework to the generation-interval distribution \citep{nishiura2009transmission, mcbryde2009early, roberts2011early, trichereau2012estimation, nishiura2015theoretical, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
 The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.''
@@
 
-We now provide a table that summarizes this information.
+We now provide a table that summarizes this information. We have also added the following text:
+
+``We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
+In this case, most estimates primarily depend on reported cases from China, particularly from Hubei Province.
+Differences among estimates are likely to driven by differences in estimation methods and underlying assumptions, rather than epidemiological differences.
+The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
+Since SARS-CoV-2 had primarily spread in Hubei Province, China during this period, it is not possible to include independent sources of data from other countries.
+Thus, the pooled estimator should be interpreted with care.''
 
@@
 Maybe author could give a sense of the importance of the element they are focusing on with the respect to the other issues?}
+
+All of these issues raised here directly affect estimates of growth rate $r$, which in turn affect estimates of the basic reproduction number \Ro.
+Therefore, comparing the growth rate with generation-interval distribution parameters allow us to take these assumptions into account implicitly. We have made this clearer in the main text.
 
@@
 
+We have made significant changes to our Methods section.
+
 \revtext{The authors do not describe all the methods used in this article within the Methods section.
@@
 
-Done.
+We have removed the formula from the caption. However, we decided to include the formula in the Results section as we think it provides better flow:
+
+``We further explore how the effects of uncertainties in generation-interval distributions change when there is less uncertainty associated with the exponential growth rate.
+This hypothetical example reflects recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
+To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$; 
+then $\hat{\mu}_r$ has the same median as $\mu_r$ but the associated 95\% CI is 4 times narrower (0.16--0.19 $\textrm{days}^{-1}$).''
 
@@
 
-Done.
+We have made significant revisions to the Methods section. 
+In particular, we made it very clear that we are simply averaging across the reported estimates and are not relying on any data or dynamical model.
+Therefore, we do not think that a directed acyclic graph is necessary with the current text, particularly given that we are simply averaging across reported values.
 
@@
 
+``Here, we assume that all 7 studies are equally weighted.
+During the initial phase of an outbreak of a novel virus such as SARS-CoV-2, uncertainties in parameter estimates are primarily driven by the underlying assumptions; narrow confidence intervals from a study are likely to reflect their strong assumptions rather than its precision of inference.
+Moreover, we cannot assess which assumptions are more realistic during the initial phase of an outbreak a novel virus due to the scarcity of information.
+Therefore, it is difficult to come up with a sensible way to weigh different estimates, and we do not explore weighted estimation.''
+
 \revtext{In each chain, the authors use 500,000 sampling steps and remove every 1000th step.
@@
 This may be an overkill but we wanted to make sure that the posterior samples are not autocorrelated.
-This is definitely sufficient for the effective sample size to reliably calculate the 95\% confidence intervals, especially given that we're only averaging across previously estimated parameter values, rather than explicitly fitting a model to data.
+On the other hand, this is definitely sufficient for the effective sample size to reliably calculate the 95\% confidence intervals, especially given that we're only averaging across previously estimated parameter values, rather than explicitly fitting a model to data.
 
@@
 
-Done.
+Variation parameters do not affect the basic reproduction number.
 
@@
 
-No...
+We do not rely on data or fit a model to data. Therefore, it is not possible to compare with the reported cases.
 
@@
 
-Done.
+We have changed the figure caption as follows:
+
+``We inferred point estimates (black), uniform distributions (orange) or confidence/credible intervals (purple) for each parameter from each study, and combined them into pooled estimates using a Bayesian multilevel model (red).
+Points represent medians calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$ (orange and purple).
+Error bars represent 95\% equi-tailed quantiles calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$.
+Red violin plots represent distributions of 2000 posterior samples.
+Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval dispersion.''
 
@@
 
-Done.
+We have removed most, if not all, interpretation of the results or opinions based on the results from the Results section.
 
@@
 
-Done.
+This is a very good point. We have added the following paragraph to the main text:
+
+``Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26, 2020).
+Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates have not been finalized.
+For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020: $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
+we did not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
+Some studies also do not provide detailed description of their methods, data, and/or assumptions.
+The quality of these analyses adds further uncertainty to their results that are not captured by their uncertainty quantification (e.g., reported credible intervals) or by our analysis.
+While pre-prints can play an important role in disseminating information, they should be interpreted with care.''
 
@@
 Some of the issues raised here (e.g., day of the week effect and limited diagnostic resources) were already mentioned in the Discussion of the previous version of the manuscript.
-We now mention other points as well.
+We now mention other points as well:
+
+``There are other important phenomena not covered by our simple framework.
+Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
+For emerging pathogens such as SARS-CoV-2, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
+for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
+Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{tariq2019assessing}.
+Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated credible intervals.
+Other factors related to reporting --- including changes in case definition, saturation in diagnostic test capacity, transparency of data, and representativeness of samples --- will also affect the inference.
+Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects};
+since we are primarily focusing on the outbreak in Wuhan city before confinement, generation intervals are unlikely to vary significantly.
+All of these factors, including fitting to cumulative curves or ignoring process errors, affect the estimation of the exponential growth rate (as well as the associated uncertainties), which in turn affects the estimation of the basic reproductive number.''
 
@@
 
-Done.
+We explain why we do not consider revised estimates in several places. See above in the discussion about pre-prints.
 
@@
 
-Done.
+We match probability, not necessarily equi-tailed quantiles. A Gamma distribution distribution with mean
+2.2 and shape 12 has a 90\% probability of containing a value between 1.27 and 3.34. We make this clear in our main text now:
+
+``We used gamma distributions to model values reported with confidence or credible intervals and uniform distributions to model values reported with ranges;
+when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
+This probability does not need to be equi-tailed.
+For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
+we model this estimate as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).''
 
@@
 
-Done.
+We believe that these priors are sufficiently wide and therefore did not try different priors. We provide a better justification in the main text:
+
+``These priors are chosen such that their 95\% quantiles are sufficiently wider than biologically realistic parameter ranges.
+Specifically, 95\% quantiles for $\mu_r$, $\mu_G$, and $\mu_\kappa$ are 0.02--0.40 $\mathrm{days}^{-1}$, 0.8--19.5 days, and 0.1--1.4, respectively.
+Parameters that are outside these ranges are biologically unrealistic for SARS-CoV-2 outbreaks.
+Therefore, we do not expect our results to be sensitive to these priors.''
 
@@
 
+``Weak priors (e.g. half-Cauchy(0,5)) can lead to inefficient sampling and poor convergence.''
+
 \revtext{Page 7, Line 20. Figure 1 - Caption: Did the authors test the sensitivity of the value of kappa of
@@
 
-Done.
+We now check the sensitivity of the value of kappa of Study 2 on the pooled estimate of the basic reproduction number.
+
+``We note that our estimate of \Rpool is relatively insensitive to our assumption of $\kappa=0.5$ for Study 2: assuming $\kappa=0.1$ gives $\Rpool = 3.0$ (95\% CI: 2.2--4.7), whereas assuming $\kappa=0.9$ gives $\Rpool = 2.9$ (95\% CI: 2.1--4.4).''
 
@@
 
+``Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
+given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delay of around 5 days \citep{sun2020early}, the majority of reported cases during study periods are likely to have been infected prior to the travel ban.''
+
 \revtext{Page 8, Line 50. Please clarify the sentence: ”We find that incorporating uncertainties one at a
@@
 
-Done.
+We have changed this explanation (and added Sensitivity Analysis section in the Methods):
+
+``Finally, Figure 3 compares the base estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
+All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.''
 
@@
 
+\pagebreak
+
+\bibliography{../ncov_abbr.bib}
+
 \end{document}

### 2020-05-06 review/response.tex
Saving response so I can move it (sorry)

@@
 
-During the exponential growth period, the basic reproduction number entirely depends on the exponential growth rate and the generation-interval distributions.
-Therefore, all model assumptions can be boiled down to assumptions about the exponential growth rate and the shape of the generation-interval distribution.
-So we are effectively considering all model assumptions.
+This is a key point. During the exponential growth period, the basic reproduction number entirely depends on the exponential growth rate and the generation-interval distributions.
+Therefore, the effect of model assumptions on the inferred reproductive number can in fact be considered via their effects on exponential growth rate and the shape of the generation-interval distribution.
 We have tried to make this point clearer in the main text:
@@
 
-Since we are interested in the uncertainty, expressed by probability distributions, the overall probabilities are not affected by the assumptions of gamma or uniform distributions.
+Since we are interested in the uncertainty, expressed by probability distributions, our analysis is consistent with the uncertainty reported by the authors.
 We have added following sentences in the main text:
@@
 
-See above. We have also added a table to provide a better overview of the models and data used in each study.
+See above.
+These assumptions will be reflected in the uncertainty distributions reported by the authors, which we then incorporate into our analysis.
+We have added a new table to provide a better overview of the models and data used in each study.
 
 \revtext{and discuss how well the gamma approximation framework embrace all cases.}
+
+\jd{I think all of the papers we reviewed actually used exponential, fixed or gamma, no? Should we say that?}
 
@@
 
-We might have inadvertently confused reviewer 2 to some extent.
-Our method does not rely on any data or dynamical models but the reviewer asks us to compare our results with the reported number of cases (this is not possible as we do not make any predictions about the epidemic trajectory).
-We use a multi-level model to calculate pooled estimates of parameters that have already been estimated or assumed from other studies.
+Our exposition may not have been clear enough for reviewer 2.
+Our method does not directly rely on data, and does not depend on a specific dynamical model, but the reviewer asks us to compare our results with the reported number of cases.
+We use a multi-level model to calculate pooled estimates of parameters that have already been estimated or assumed from other studies, based on fits to case reports.
 Then, we use the pooled estimates to calculate $\mathcal R_0$.

### 2020-05-06 ncov.tex
done
Saving response so I can move it (sorry)

@@
 We note that uncertainties in the probability distributions that we calculate for the estimated exponential growth rates reflects model assumptions, statistical methods, as well as quality of the data that each study relies on.
-Then, the the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ defines an empirical parameter distribution for each study $i$ without having to re-fit each model.
-
-For most cases, we model $(\Ro)_i$, $\bar G_i$, and $\kappa_i$ using independent distributions (as specified in Table 2) to calculate $r_i$.
 This assumption does not affect our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
@@
 The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make.
+Here, we do so by averaging across reported values, without explicitly re-fitting their models.
 We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
@@
 The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
-Since SARS-CoV-2 had primarily spread in Hubei Province, China during this period, it is not possible to include independent sources of data from other countries.
+Since SARS-CoV-2 had primarily spread in Hubei Province, China during this period, it was not possible to include independent sources of data from other countries.
 Thus, the pooled estimator should be interpreted with care.

### 2020-05-06 response.tex
Out of time

--- 2020-05-06 review/response.tex
@@
 
-We have removed the formula from the caption. However, we decided to include the formula in the Results section as we think it provides better flow:
+We have moved the formula to the Results section to provide a better context for improved explanation:
 
@@
 We have made significant revisions to the Methods section. 
-In particular, we made it very clear that we are simply averaging across the reported estimates and are not relying on any data or dynamical model.
-Therefore, we do not think that a directed acyclic graph is necessary with the current text, particularly given that we are simply averaging across reported values.
+In particular, we made it very clear that we are simply using Bayesian estimation to average across the reported estimates and are not relying on any data or dynamical model.
+We hope that the new text is sufficiently clear; a directed acyclic graph to explain that we are doing Bayesian averaging seems like overkill.
 
@@
 based on the confidence of the estimation of R0.}
+
+\jd{I'm confused. Isn't the reason to use Bayesian averaging is that it does weight implicitly by certainty?}
 
@@
 for the burn-in steps of 500,000 and the thinning steps of 1,000?}
+
+\jd{Do we examine effective sample size? Can we report them?}
 
@@
 
-\bibliography{../ncov_abbr.bib}
+%% \bibliography{../ncov_abbr.bib} Sorry!
+\bibliography{ncov_abbr.bib}
 

### 2020-05-06 response.tex
Done

@@
 
-Variation parameters do not affect the basic reproduction number.
+We have sampled appropriately from the Bayesian posteriors. We talk about pooled estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$ because those are used directly to calculate the reproduction number.
+
+\jd{Is there anything we can clarify in the MS?}
 
@@
 Red violin plots represent distributions of 2000 posterior samples.
-Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval dispersion.''
+Open triangle: we assumed $\kappa=0.5$ for Study 2, which does not report generation-interval assumptions.''
 
@@
 
-We have removed most, if not all, interpretation of the results or opinions based on the results from the Results section.
+We have removed these examples, and other interpretations and opinions, from the Results section.
 
@@
 
-We explain why we do not consider revised estimates in several places. See above in the discussion about pre-prints.
+We now clearly explain why we do not consider revised estimates. See above in the discussion about pre-prints.
 
-\revtext{Page 6, Table 2. Please recheck the parameters of Study 5, which is inconsistent with the 90%
+\revtext{Page 6, Table 2. Please recheck the parameters of Study 5, which is inconsistent with the 90\%
 confidence interval in Table 1. The 90\% confidence interval for a Gamma distribution with mean

### 2020-05-06 response.tex
tweaks

@@
 
-\jd{I think all of the papers we reviewed actually used exponential, fixed or gamma, no? Should we say that?}
-
 The gamma approximation framework assumes that the generation-interval distribution is gamma distributed.
@@
 ``Here, we use the gamma approximation framework to the generation-interval distribution \citep{nishiura2009transmission, mcbryde2009early, roberts2011early, trichereau2012estimation, nishiura2015theoretical, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
-The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.''
+The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.
+Study 1, 5, 6, and 7 also used a gamma distribution (including delta and exponential distributions) to model the generation-interval distribution for SARS-CoV-2.''
 
@@
 
-\jd{I'm confused. Isn't the reason to use Bayesian averaging is that it does weight implicitly by certainty?}
-
-Done.
-
-``Here, we assume that all 7 studies are equally weighted.
-During the initial phase of an outbreak of a novel virus such as SARS-CoV-2, uncertainties in parameter estimates are primarily driven by the underlying assumptions; narrow confidence intervals from a study are likely to reflect their strong assumptions rather than its precision of inference.
-Moreover, we cannot assess which assumptions are more realistic during the initial phase of an outbreak a novel virus due to the scarcity of information.
-Therefore, it is difficult to come up with a sensible way to weigh different estimates, and we do not explore weighted estimation.''
+We do not make that hypothesis. Bayesian multilevel models implicitly weight parameters by certainty.
 
@@
 
-\jd{Do we examine effective sample size? Can we report them?}
-
 This may be an overkill but we wanted to make sure that the posterior samples are not autocorrelated.
 On the other hand, this is definitely sufficient for the effective sample size to reliably calculate the 95\% confidence intervals, especially given that we're only averaging across previously estimated parameter values, rather than explicitly fitting a model to data.
+We now ensure that effective sample sizes are greater than at least 1,000 for all parameters.
 
@@
 
-We have sampled appropriately from the Bayesian posteriors. We talk about pooled estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$ because those are used directly to calculate the reproduction number.
-
-\jd{Is there anything we can clarify in the MS?}
+We have sampled appropriately from the Bayesian posteriors.
+We talk about pooled estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$ because those are used directly to calculate the reproduction number; $\sigma_r$, $\sigma_G$, and
+$\sigma_\kappa$ do not affect the calculation of the reproduction number.
 

### 2020-05-06 response.tex
minor

@@
 
-We do not make that hypothesis. Bayesian multilevel models implicitly weight parameters by certainty.
+We used Bayesian multilevel models precisely for this reason (to weight the studies appropriately based on reported certainty.)
 
@@
 
-We have sampled appropriately from the Bayesian posteriors.
-We talk about pooled estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$ because those are used directly to calculate the reproduction number; $\sigma_r$, $\sigma_G$, and
-$\sigma_\kappa$ do not affect the calculation of the reproduction number.
+We have sampled appropriately from the Bayesian posteriors. We talk about pooled estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$ because those are used directly to calculate the reproduction number.
 

### 2020-05-07 propagate_pix.tex
Make talk pictures downstream of figure2

% Created by tikzDevice version 0.12.3 on 2020-05-07 10:03:52
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (505.89,505.89);
\end{scope}
\begin{scope}
\path[clip] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(497.89, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,127.00) --
	(497.89,127.00);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,209.36) --
	(497.89,209.36);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,291.73) --
	(497.89,291.73);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,374.10) --
	(497.89,374.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,456.46) --
	(497.89,456.46);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.59,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,210.47) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,204.36) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
	(101.40,210.47);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,417.54) --
	(187.59,417.54);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,417.54) --
	(187.59, 86.40);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59, 86.40) --
	(187.59, 86.40);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
	(273.79,267.25);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
	(273.79,166.69);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,166.69) --
	(273.79,166.69);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
	(359.98,287.03);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
	(359.98,109.81);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,109.81) --
	(359.98,109.81);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,465.70) --
	(446.17,465.70);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,465.70) --
	(446.17, 58.97);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17, 58.97) --
	(446.17, 58.97);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	( 49.68,472.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {2.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,122.59) {2.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,204.96) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,287.32) {3.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,369.69) {4.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,452.05) {4.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,127.00) --
	( 49.68,127.00);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,209.36) --
	( 49.68,209.36);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,291.73) --
	( 49.68,291.73);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,374.10) --
	( 49.68,374.10);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,456.46) --
	( 49.68,456.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(497.89, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (101.40, 40.63) --
	(101.40, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (187.59, 40.63) --
	(187.59, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (273.79, 40.63) --
	(273.79, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (359.98, 40.63) --
	(359.98, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (446.17, 40.63) --
	(446.17, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (101.40, 28.62) {none};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (187.59, 28.62) {$\mu_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (273.79, 28.62) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (359.98, 28.62) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (446.17, 28.62) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (273.79, 11.11) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,258.78) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at ( 49.68,484.67) {A. Baseline};
\end{scope}
\end{tikzpicture}
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (505.89,505.89);
\end{scope}
\begin{scope}
\path[clip] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.68, 44.63) rectangle (497.89,472.93);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,130.29) --
	(497.89,130.29);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,244.51) --
	(497.89,244.51);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,358.72) --
	(497.89,358.72);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,246.43) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.59,246.43) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,246.43) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,246.43) circle (  5.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,243.03) circle (  5.71);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,246.43) --
	(101.40,246.43);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,246.43) --
	(101.40,246.43);

\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,246.43) --
	(101.40,246.43);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,330.28) --
	(187.59,330.28);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,330.28) --
	(187.59,189.75);

\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,189.75) --
	(187.59,189.75);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,344.85) --
	(273.79,344.85);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,344.85) --
	(273.79,170.54);

\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,170.54) --
	(273.79,170.54);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,379.13) --
	(359.98,379.13);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,379.13) --
	(359.98, 71.95);

\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98, 71.95) --
	(359.98, 71.95);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,466.99) --
	(446.17,466.99);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,466.99) --
	(446.17, 59.04);

\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17, 59.04) --
	(446.17, 59.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	( 49.68,472.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,125.89) {2.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,240.10) {3.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,354.31) {3.4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,130.29) --
	( 49.68,130.29);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,244.51) --
	( 49.68,244.51);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,358.72) --
	( 49.68,358.72);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
	(497.89, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (101.40, 40.63) --
	(101.40, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (187.59, 40.63) --
	(187.59, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (273.79, 40.63) --
	(273.79, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (359.98, 40.63) --
	(359.98, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (446.17, 40.63) --
	(446.17, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (101.40, 28.62) {none};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (187.59, 28.62) {$\hat{\mu}_r$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (273.79, 28.62) {$\mu_G$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (359.98, 28.62) {$\mu_\kappa$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (446.17, 28.62) {all};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (273.79, 11.11) {Uncertainty type};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,258.78) {Basic reproductive number};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.92] at ( 49.68,484.67) {B. Reduced uncertainty in $r$};
\end{scope}
\end{tikzpicture}

\end{document}

### 2020-05-07 figure2.R

@@ -1,4 +1,5 @@
 library(coda)
+library(dplyr)
 library(ggplot2); theme_set(theme_bw(base_size=16))
 library(gridExtra)
 library(tikzDevice)

### 2020-05-07 ncov.tex

@@
 The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.
+Study 1, 5, 6, and 7 also used a gamma distribution (including delta and exponential distributions) to model the generation-interval distribution for SARS-CoV-2.
 Assuming that generation intervals follow a gamma distribution with mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation of a gamma distribution, we have
@@
 
-Here, we assume that all 7 studies are equally weighted.
-During the initial phase of an outbreak of a novel virus such as SARS-CoV-2, uncertainties in parameter estimates are primarily driven by the underlying assumptions; narrow confidence intervals from a study are likely to reflect their strong assumptions rather than its precision of inference.
-Moreover, we cannot assess which assumptions are more realistic during the initial phase of an outbreak a novel virus due to the scarcity of information.
-Therefore, it is difficult to come up with a sensible way to weigh different estimates, and we do not explore weighted estimation.
-
 We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
@@
 Posterior samples are thinned every 1000 steps to remove autocorrelations among posterior samples.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 and the effective sample size is greater than 1000 for all hyperparameters \citep{gelman1992inference};
 trace plots and marginal posterior distribution plots are presented in Appendix.

### 2020-05-07 compare_assumption.tex
done

@@
-% Created by tikzDevice version 0.12.3 on 2020-04-27 16:28:44
+% Created by tikzDevice version 0.12.3 on 2020-05-07 17:10:16
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 95.51, 71.16) -- (167.58, 71.16);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (167.58, 71.16) --
+	(167.58, 71.16);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (167.58, 71.16) --
+	( 95.51, 71.16);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 95.51, 71.16) --
+	( 95.51, 71.16);
 \definecolor{drawColor}{RGB}{160,32,240}
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 96.55, 92.40) -- (163.42, 92.40);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (114.86,113.63) -- (141.83,113.63);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (163.42, 92.40) --
+	(163.42, 92.40);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (163.42, 92.40) --
+	( 96.55, 92.40);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 96.55, 92.40) --
+	( 96.55, 92.40);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (141.83,113.63) --
+	(141.83,113.63);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (141.83,113.63) --
+	(114.86,113.63);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (114.86,113.63) --
+	(114.86,113.63);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) -- (117.88,198.57);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) --
+	(117.88,198.57);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) --
+	(117.88,198.57);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (117.88,198.57) --
+	(117.88,198.57);
 \definecolor{drawColor}{RGB}{160,32,240}
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (164.38,134.87) -- (174.59,134.87);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 79.16,156.10) -- (149.69,156.10);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (152.24,177.34) -- (187.73,177.34);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (174.59,134.87) --
+	(174.59,134.87);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (174.59,134.87) --
+	(164.38,134.87);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (164.38,134.87) --
+	(164.38,134.87);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (149.69,156.10) --
+	(149.69,156.10);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (149.69,156.10) --
+	( 79.16,156.10);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 79.16,156.10) --
+	( 79.16,156.10);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (187.73,177.34) --
+	(187.73,177.34);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (187.73,177.34) --
+	(152.24,177.34);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (152.24,177.34) --
+	(152.24,177.34);
 \definecolor{drawColor}{RGB}{255,0,0}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (105.33, 49.85) --
-	(105.57, 49.84) --
-	(105.81, 49.83) --
-	(106.05, 49.82) --
-	(106.29, 49.81) --
-	(106.53, 49.80) --
-	(106.77, 49.79) --
-	(107.01, 49.78) --
-	(107.25, 49.77) --
-	(107.49, 49.76) --
-	(107.73, 49.74) --
-	(107.97, 49.73) --
-	(108.21, 49.71) --
-	(108.45, 49.69) --
-	(108.69, 49.68) --
-	(108.93, 49.66) --
-	(109.17, 49.64) --
-	(109.41, 49.62) --
-	(109.65, 49.60) --
-	(109.89, 49.58) --
-	(110.13, 49.56) --
-	(110.37, 49.54) --
-	(110.61, 49.51) --
-	(110.85, 49.49) --
-	(111.09, 49.47) --
-	(111.33, 49.45) --
-	(111.57, 49.43) --
-	(111.81, 49.40) --
-	(112.05, 49.38) --
-	(112.29, 49.36) --
-	(112.53, 49.34) --
-	(112.77, 49.32) --
-	(113.01, 49.30) --
-	(113.25, 49.28) --
-	(113.49, 49.26) --
-	(113.73, 49.24) --
-	(113.97, 49.22) --
-	(114.21, 49.20) --
-	(114.45, 49.18) --
-	(114.69, 49.16) --
-	(114.93, 49.13) --
-	(115.17, 49.11) --
-	(115.41, 49.09) --
-	(115.65, 49.06) --
-	(115.89, 49.03) --
-	(116.13, 49.00) --
-	(116.37, 48.96) --
-	(116.61, 48.93) --
-	(116.85, 48.89) --
-	(117.09, 48.84) --
-	(117.33, 48.79) --
-	(117.57, 48.74) --
-	(117.81, 48.69) --
-	(118.05, 48.63) --
-	(118.29, 48.57) --
-	(118.53, 48.50) --
-	(118.77, 48.44) --
-	(119.01, 48.37) --
-	(119.25, 48.30) --
-	(119.49, 48.23) --
-	(119.73, 48.16) --
-	(119.97, 48.08) --
-	(120.21, 48.01) --
-	(120.44, 47.94) --
-	(120.68, 47.87) --
-	(120.92, 47.80) --
-	(121.16, 47.72) --
-	(121.40, 47.65) --
-	(121.64, 47.58) --
-	(121.88, 47.51) --
-	(122.12, 47.43) --
-	(122.36, 47.35) --
-	(122.60, 47.27) --
-	(122.84, 47.18) --
-	(123.08, 47.09) --
-	(123.32, 47.00) --
-	(123.56, 46.89) --
-	(123.80, 46.79) --
-	(124.04, 46.67) --
-	(124.28, 46.55) --
-	(124.52, 46.43) --
-	(124.76, 46.30) --
-	(125.00, 46.16) --
-	(125.24, 46.02) --
-	(125.48, 45.88) --
-	(125.72, 45.73) --
-	(125.96, 45.58) --
-	(126.20, 45.42) --
-	(126.44, 45.27) --
-	(126.68, 45.11) --
-	(126.92, 44.96) --
-	(127.16, 44.80) --
-	(127.40, 44.64) --
-	(127.64, 44.49) --
-	(127.88, 44.33) --
-	(128.12, 44.17) --
-	(128.36, 44.02) --
-	(128.60, 43.86) --
-	(128.84, 43.70) --
-	(129.08, 43.54) --
-	(129.32, 43.38) --
-	(129.56, 43.22) --
-	(129.80, 43.06) --
-	(130.04, 42.89) --
-	(130.28, 42.73) --
-	(130.52, 42.56) --
-	(130.76, 42.40) --
-	(131.00, 42.23) --
-	(131.24, 42.07) --
-	(131.48, 41.90) --
-	(131.72, 41.74) --
-	(131.96, 41.58) --
-	(132.20, 41.43) --
-	(132.44, 41.27) --
-	(132.68, 41.13) --
-	(132.92, 40.99) --
-	(133.16, 40.85) --
-	(133.40, 40.72) --
-	(133.64, 40.60) --
-	(133.88, 40.48) --
-	(134.12, 40.37) --
-	(134.36, 40.26) --
-	(134.60, 40.16) --
-	(134.84, 40.07) --
-	(135.08, 39.98) --
-	(135.32, 39.90) --
-	(135.56, 39.82) --
-	(135.80, 39.75) --
-	(136.04, 39.68) --
-	(136.28, 39.61) --
-	(136.52, 39.55) --
-	(136.76, 39.50) --
-	(137.00, 39.45) --
-	(137.24, 39.41) --
-	(137.48, 39.38) --
-	(137.72, 39.35) --
-	(137.96, 39.33) --
-	(138.20, 39.32) --
-	(138.44, 39.31) --
-	(138.68, 39.32) --
-	(138.92, 39.32) --
-	(139.16, 39.34) --
-	(139.40, 39.37) --
-	(139.64, 39.40) --
-	(139.88, 39.44) --
-	(140.12, 39.49) --
-	(140.36, 39.54) --
-	(140.60, 39.60) --
-	(140.84, 39.67) --
-	(141.08, 39.74) --
-	(141.32, 39.82) --
-	(141.56, 39.90) --
-	(141.80, 39.99) --
-	(142.04, 40.08) --
-	(142.28, 40.18) --
-	(142.52, 40.29) --
-	(142.76, 40.40) --
-	(143.00, 40.51) --
-	(143.24, 40.62) --
-	(143.48, 40.75) --
-	(143.72, 40.87) --
-	(143.96, 40.99) --
-	(144.20, 41.12) --
-	(144.44, 41.25) --
-	(144.68, 41.38) --
-	(144.92, 41.51) --
-	(145.16, 41.65) --
-	(145.40, 41.77) --
-	(145.64, 41.90) --
-	(145.88, 42.03) --
-	(146.12, 42.15) --
-	(146.36, 42.26) --
-	(146.60, 42.38) --
-	(146.84, 42.49) --
-	(147.08, 42.59) --
-	(147.32, 42.69) --
-	(147.56, 42.79) --
-	(147.80, 42.88) --
-	(148.04, 42.97) --
-	(148.28, 43.05) --
-	(148.52, 43.13) --
-	(148.76, 43.21) --
-	(149.00, 43.29) --
-	(149.24, 43.37) --
-	(149.48, 43.45) --
-	(149.72, 43.53) --
-	(149.96, 43.61) --
-	(150.20, 43.70) --
-	(150.44, 43.78) --
-	(150.68, 43.86) --
-	(150.92, 43.95) --
-	(151.16, 44.04) --
-	(151.40, 44.13) --
-	(151.64, 44.22) --
-	(151.88, 44.32) --
-	(152.12, 44.41) --
-	(152.36, 44.52) --
-	(152.60, 44.62) --
-	(152.84, 44.72) --
-	(153.08, 44.83) --
-	(153.32, 44.94) --
-	(153.56, 45.06) --
-	(153.80, 45.18) --
-	(154.04, 45.30) --
-	(154.28, 45.42) --
-	(154.52, 45.54) --
-	(154.76, 45.67) --
-	(155.00, 45.80) --
-	(155.24, 45.92) --
-	(155.48, 46.05) --
-	(155.72, 46.18) --
-	(155.96, 46.30) --
-	(156.20, 46.42) --
-	(156.44, 46.53) --
-	(156.68, 46.65) --
-	(156.92, 46.75) --
-	(157.16, 46.85) --
-	(157.40, 46.95) --
-	(157.64, 47.04) --
-	(157.88, 47.12) --
-	(158.12, 47.20) --
-	(158.36, 47.27) --
-	(158.60, 47.34) --
-	(158.84, 47.40) --
-	(159.08, 47.45) --
-	(159.32, 47.50) --
-	(159.56, 47.55) --
-	(159.80, 47.59) --
-	(160.04, 47.62) --
-	(160.28, 47.66) --
-	(160.52, 47.69) --
-	(160.76, 47.71) --
-	(161.00, 47.74) --
-	(161.24, 47.76) --
-	(161.48, 47.78) --
-	(161.72, 47.81) --
-	(161.96, 47.83) --
-	(162.20, 47.85) --
-	(162.44, 47.88) --
-	(162.68, 47.91) --
-	(162.92, 47.94) --
-	(163.16, 47.97) --
-	(163.40, 48.01) --
-	(163.64, 48.05) --
-	(163.88, 48.09) --
-	(164.12, 48.13) --
-	(164.36, 48.18) --
-	(164.60, 48.22) --
-	(164.84, 48.27) --
-	(165.08, 48.32) --
-	(165.32, 48.36) --
-	(165.56, 48.41) --
-	(165.80, 48.46) --
-	(166.04, 48.50) --
-	(166.28, 48.54) --
-	(166.52, 48.59) --
-	(166.76, 48.63) --
-	(167.00, 48.66) --
-	(167.24, 48.70) --
-	(167.48, 48.74) --
-	(167.72, 48.77) --
-	(167.96, 48.81) --
-	(168.20, 48.84) --
-	(168.44, 48.88) --
-	(168.68, 48.91) --
-	(168.92, 48.95) --
-	(169.16, 48.98) --
-	(169.40, 49.01) --
-	(169.64, 49.05) --
-	(169.88, 49.08) --
-	(170.12, 49.11) --
-	(170.36, 49.14) --
-	(170.60, 49.17) --
-	(170.84, 49.20) --
-	(171.08, 49.22) --
-	(171.32, 49.25) --
-	(171.56, 49.27) --
-	(171.80, 49.29) --
-	(172.04, 49.31) --
-	(172.28, 49.33) --
-	(172.52, 49.34) --
-	(172.76, 49.36) --
-	(173.00, 49.37) --
-	(173.24, 49.38) --
-	(173.48, 49.39) --
-	(173.72, 49.41) --
-	(173.96, 49.42) --
-	(174.20, 49.43) --
-	(174.44, 49.44) --
-	(174.68, 49.45) --
-	(174.92, 49.46) --
-	(175.16, 49.47) --
-	(175.40, 49.48) --
-	(175.64, 49.49) --
-	(175.88, 49.50) --
-	(176.12, 49.51) --
-	(176.36, 49.52) --
-	(176.60, 49.53) --
-	(176.84, 49.53) --
-	(177.08, 49.54) --
-	(177.32, 49.55) --
-	(177.56, 49.55) --
-	(177.80, 49.56) --
-	(178.04, 49.56) --
-	(178.28, 49.57) --
-	(178.52, 49.58) --
-	(178.76, 49.58) --
-	(179.00, 49.59) --
-	(179.24, 49.59) --
-	(179.48, 49.60) --
-	(179.72, 49.60) --
-	(179.96, 49.61) --
-	(180.20, 49.62) --
-	(180.44, 49.63) --
-	(180.68, 49.64) --
-	(180.92, 49.65) --
-	(181.16, 49.66) --
-	(181.40, 49.67) --
-	(181.64, 49.68) --
-	(181.88, 49.69) --
-	(182.12, 49.70) --
-	(182.36, 49.72) --
-	(182.60, 49.73) --
-	(182.84, 49.74) --
-	(183.08, 49.75) --
-	(183.32, 49.76) --
-	(183.56, 49.77) --
-	(183.80, 49.78) --
-	(184.04, 49.79) --
-	(184.28, 49.80) --
-	(184.52, 49.81) --
-	(184.76, 49.81) --
-	(185.00, 49.82) --
-	(185.24, 49.82) --
-	(185.48, 49.83) --
-	(185.72, 49.83) --
-	(185.96, 49.83) --
-	(186.20, 49.83) --
-	(186.44, 49.83) --
-	(186.68, 49.83) --
-	(186.92, 49.83) --
-	(187.16, 49.83) --
-	(187.40, 49.83) --
-	(187.64, 49.83) --
-	(187.88, 49.82) --
-	(188.12, 49.82) --
-	(188.36, 49.82) --
-	(188.60, 49.82) --
-	(188.84, 49.81) --
-	(189.08, 49.81) --
-	(189.32, 49.81) --
-	(189.56, 49.81) --
-	(189.80, 49.81) --
-	(190.04, 49.81) --
-	(190.28, 49.81) --
-	(190.52, 49.81) --
-	(190.76, 49.81) --
-	(191.00, 49.81) --
-	(191.24, 49.81) --
-	(191.48, 49.82) --
-	(191.72, 49.82) --
-	(191.96, 49.82) --
-	(192.20, 49.83) --
-	(192.44, 49.83) --
-	(192.68, 49.84) --
-	(192.92, 49.84) --
-	(193.16, 49.85) --
-	(193.40, 49.86) --
-	(193.64, 49.86) --
-	(193.88, 49.87) --
-	(194.12, 49.87) --
-	(194.36, 49.87) --
-	(194.60, 49.88) --
-	(194.84, 49.88) --
-	(195.08, 49.88) --
-	(195.32, 49.88) --
-	(195.56, 49.88) --
-	(195.80, 49.88) --
-	(196.04, 49.88) --
-	(196.28, 49.88) --
-	(196.52, 49.88) --
-	(196.76, 49.88) --
-	(197.00, 49.88) --
-	(197.24, 49.88) --
-	(197.48, 49.88) --
-	(197.72, 49.88) --
-	(197.96, 49.88) --
-	(198.20, 49.88) --
-	(198.44, 49.88) --
-	(198.68, 49.88) --
-	(198.92, 49.88) --
-	(199.16, 49.88) --
-	(199.40, 49.89) --
-	(199.64, 49.89) --
-	(199.88, 49.89) --
-	(200.12, 49.89) --
-	(200.36, 49.90) --
-	(200.60, 49.90) --
-	(200.84, 49.90) --
-	(201.08, 49.91) --
-	(201.32, 49.91) --
-	(201.56, 49.91) --
-	(201.80, 49.91) --
-	(202.04, 49.92) --
-	(202.28, 49.92) --
-	(202.52, 49.92) --
-	(202.76, 49.92) --
-	(203.00, 49.92) --
+\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (105.33, 49.93) --
+	(105.57, 49.93) --
+	(105.81, 49.93) --
+	(106.05, 49.93) --
+	(106.29, 49.93) --
+	(106.53, 49.93) --
+	(106.77, 49.93) --
+	(107.01, 49.93) --
+	(107.25, 49.93) --
+	(107.49, 49.93) --
+	(107.73, 49.93) --
+	(107.97, 49.93) --
+	(108.21, 49.93) --
+	(108.45, 49.93) --
+	(108.69, 49.93) --
+	(108.93, 49.93) --
+	(109.17, 49.93) --
+	(109.41, 49.93) --
+	(109.65, 49.93) --
+	(109.89, 49.93) --
+	(110.13, 49.93) --
+	(110.37, 49.93) --
+	(110.61, 49.93) --
+	(110.85, 49.93) --
+	(111.09, 49.93) --
+	(111.33, 49.93) --
+	(111.57, 49.93) --
+	(111.81, 49.93) --
+	(112.05, 49.93) --
+	(112.29, 49.93) --
+	(112.53, 49.93) --
+	(112.77, 49.93) --
+	(113.01, 49.93) --
+	(113.25, 49.93) --
+	(113.49, 49.93) --
+	(113.73, 49.93) --
+	(113.97, 49.93) --
+	(114.21, 49.93) --
+	(114.45, 49.93) --
+	(114.69, 49.93) --
+	(114.93, 49.93) --
+	(115.17, 49.93) --
+	(115.41, 49.93) --
+	(115.65, 49.93) --
+	(115.89, 49.93) --
+	(116.13, 49.93) --
+	(116.37, 49.93) --
+	(116.61, 49.93) --
+	(116.85, 49.93) --
+	(117.09, 49.93) --
+	(117.33, 49.93) --
+	(117.57, 49.93) --
+	(117.81, 49.93) --
+	(118.05, 49.93) --
+	(118.29, 49.93) --
+	(118.53, 49.93) --
+	(118.77, 49.93) --
+	(119.01, 49.93) --
+	(119.25, 49.93) --
+	(119.49, 49.93) --
+	(119.73, 49.93) --
+	(119.97, 49.93) --
+	(120.21, 49.93) --
+	(120.44, 49.93) --
+	(120.68, 49.93) --
+	(120.92, 49.93) --
+	(121.16, 49.93) --
+	(121.40, 49.93) --
+	(121.64, 49.93) --
+	(121.88, 49.93) --
+	(122.12, 49.93) --
+	(122.36, 49.93) --
+	(122.60, 49.93) --
+	(122.84, 49.93) --
+	(123.08, 49.93) --
+	(123.32, 49.93) --
+	(123.56, 49.93) --
+	(123.80, 49.93) --
+	(124.04, 49.93) --
+	(124.28, 49.93) --
+	(124.52, 49.93) --
+	(124.76, 49.93) --
+	(125.00, 49.93) --
+	(125.24, 49.93) --
+	(125.48, 49.93) --
+	(125.72, 49.93) --
+	(125.96, 49.93) --
+	(126.20, 49.93) --
+	(126.44, 49.93) --
+	(126.68, 49.93) --
+	(126.92, 49.93) --
+	(127.16, 49.93) --
+	(127.40, 49.93) --
+	(127.64, 49.93) --
+	(127.88, 49.93) --
+	(128.12, 49.93) --
+	(128.36, 49.93) --
+	(128.60, 49.93) --
+	(128.84, 49.93) --
+	(129.08, 49.93) --
+	(129.32, 49.93) --
+	(129.56, 49.93) --
+	(129.80, 49.93) --
+	(130.04, 49.93) --
+	(130.28, 49.93) --
+	(130.52, 49.93) --
+	(130.76, 49.93) --
+	(131.00, 49.93) --
+	(131.24, 49.93) --
+	(131.48, 49.93) --
+	(131.72, 49.93) --
+	(131.96, 49.93) --
+	(132.20, 49.93) --
+	(132.44, 49.93) --
+	(132.68, 49.93) --
+	(132.92, 49.93) --
+	(133.16, 49.93) --
+	(133.40, 49.93) --
+	(133.64, 49.93) --
+	(133.88, 49.93) --
+	(134.12, 49.93) --
+	(134.36, 49.93) --
+	(134.60, 49.93) --
+	(134.84, 49.93) --
+	(135.08, 49.93) --
+	(135.32, 49.93) --
+	(135.56, 49.93) --
+	(135.80, 49.93) --
+	(136.04, 49.93) --
+	(136.28, 49.93) --
+	(136.52, 49.93) --
+	(136.76, 49.93) --
+	(137.00, 49.93) --
+	(137.24, 49.93) --
+	(137.48, 49.93) --
+	(137.72, 49.93) --
+	(137.96, 49.93) --
+	(138.20, 49.93) --
+	(138.44, 49.93) --
+	(138.68, 49.93) --
+	(138.92, 49.93) --
+	(139.16, 49.93) --
+	(139.40, 49.93) --
+	(139.64, 49.93) --
+	(139.88, 49.93) --
+	(140.12, 49.93) --
+	(140.36, 49.93) --
+	(140.60, 49.93) --
+	(140.84, 49.93) --
+	(141.08, 49.93) --
+	(141.32, 49.93) --
+	(141.56, 49.93) --
+	(141.80, 49.93) --
+	(142.04, 49.93) --
+	(142.28, 49.93) --
+	(142.52, 49.93) --
+	(142.76, 49.93) --
+	(143.00, 49.93) --
+	(143.24, 49.93) --
+	(143.48, 49.93) --
+	(143.72, 49.93) --
+	(143.96, 49.93) --
+	(144.20, 49.93) --
+	(144.44, 49.93) --
+	(144.68, 49.93) --
+	(144.92, 49.93) --
+	(145.16, 49.93) --
+	(145.40, 49.93) --
+	(145.64, 49.93) --
+	(145.88, 49.93) --
+	(146.12, 49.93) --
+	(146.36, 49.93) --
+	(146.60, 49.93) --
+	(146.84, 49.93) --
+	(147.08, 49.93) --
+	(147.32, 49.93) --
+	(147.56, 49.93) --
+	(147.80, 49.93) --
+	(148.04, 49.93) --
+	(148.28, 49.93) --
+	(148.52, 49.93) --
+	(148.76, 49.93) --
+	(149.00, 49.93) --
+	(149.24, 49.93) --
+	(149.48, 49.93) --
+	(149.72, 49.93) --
+	(149.96, 49.93) --
+	(150.20, 49.93) --
+	(150.44, 49.93) --
+	(150.68, 49.93) --
+	(150.92, 49.93) --
+	(151.16, 49.93) --
+	(151.40, 49.93) --
+	(151.64, 49.93) --
+	(151.88, 49.93) --
+	(152.12, 49.93) --
+	(152.36, 49.93) --
+	(152.60, 49.93) --
+	(152.84, 49.93) --
+	(153.08, 49.93) --
+	(153.32, 49.93) --
+	(153.56, 49.93) --
+	(153.80, 49.93) --
+	(154.04, 49.93) --
+	(154.28, 49.93) --
+	(154.52, 49.93) --
+	(154.76, 49.93) --
+	(155.00, 49.93) --
+	(155.24, 49.93) --
+	(155.48, 49.93) --
+	(155.72, 49.93) --
+	(155.96, 49.93) --
+	(156.20, 49.93) --
+	(156.44, 49.93) --
+	(156.68, 49.93) --
+	(156.92, 49.93) --
+	(157.16, 49.93) --
+	(157.40, 49.93) --
+	(157.64, 49.93) --
+	(157.88, 49.93) --
+	(158.12, 49.93) --
+	(158.36, 49.93) --
+	(158.60, 49.93) --
+	(158.84, 49.93) --
+	(159.08, 49.93) --
+	(159.32, 49.93) --
+	(159.56, 49.93) --
+	(159.80, 49.93) --
+	(160.04, 49.93) --
+	(160.28, 49.93) --
+	(160.52, 49.93) --
+	(160.76, 49.93) --
+	(161.00, 49.93) --
+	(161.24, 49.93) --
+	(161.48, 49.93) --
+	(161.72, 49.93) --
+	(161.96, 49.93) --
+	(162.20, 49.93) --
+	(162.44, 49.93) --
+	(162.68, 49.93) --
+	(162.92, 49.93) --
+	(163.16, 49.93) --
+	(163.40, 49.93) --
+	(163.64, 49.93) --
+	(163.88, 49.93) --
+	(164.12, 49.93) --
+	(164.36, 49.93) --
+	(164.60, 49.93) --
+	(164.84, 49.93) --
+	(165.08, 49.93) --
+	(165.32, 49.93) --
+	(165.56, 49.93) --
+	(165.80, 49.93) --
+	(166.04, 49.93) --
+	(166.28, 49.93) --
+	(166.52, 49.93) --
+	(166.76, 49.93) --
+	(167.00, 49.93) --
+	(167.24, 49.93) --
+	(167.48, 49.93) --
+	(167.72, 49.93) --
+	(167.96, 49.93) --
+	(168.20, 49.93) --
+	(168.44, 49.93) --
+	(168.68, 49.93) --
+	(168.92, 49.93) --
+	(169.16, 49.93) --
+	(169.40, 49.93) --
+	(169.64, 49.93) --
+	(169.88, 49.93) --
+	(170.12, 49.93) --
+	(170.36, 49.93) --
+	(170.60, 49.93) --
+	(170.84, 49.93) --
+	(171.08, 49.93) --
+	(171.32, 49.93) --
+	(171.56, 49.93) --
+	(171.80, 49.93) --
+	(172.04, 49.93) --
+	(172.28, 49.93) --
+	(172.52, 49.93) --
+	(172.76, 49.93) --
+	(173.00, 49.93) --
+	(173.24, 49.93) --
+	(173.48, 49.93) --
+	(173.72, 49.93) --
+	(173.96, 49.93) --
+	(174.20, 49.93) --
+	(174.44, 49.93) --
+	(174.68, 49.93) --
+	(174.92, 49.93) --
+	(175.16, 49.93) --
+	(175.40, 49.93) --
+	(175.64, 49.93) --
+	(175.88, 49.93) --
+	(176.12, 49.93) --
+	(176.36, 49.93) --
+	(176.60, 49.93) --
+	(176.84, 49.93) --
+	(177.08, 49.93) --
+	(177.32, 49.93) --
+	(177.56, 49.93) --
+	(177.80, 49.93) --
+	(178.04, 49.93) --
+	(178.28, 49.93) --
+	(178.52, 49.93) --
+	(178.76, 49.93) --
+	(179.00, 49.93) --
+	(179.24, 49.93) --
+	(179.48, 49.93) --
+	(179.72, 49.93) --
+	(179.96, 49.93) --
+	(180.20, 49.93) --
+	(180.44, 49.93) --
+	(180.68, 49.93) --
+	(180.92, 49.93) --
+	(181.16, 49.93) --
+	(181.40, 49.93) --
+	(181.64, 49.93) --
+	(181.88, 49.93) --
+	(182.12, 49.93) --
+	(182.36, 49.93) --
+	(182.60, 49.93) --
+	(182.84, 49.93) --
+	(183.08, 49.93) --
+	(183.32, 49.93) --
+	(183.56, 49.93) --
+	(183.80, 49.93) --
+	(184.04, 49.93) --
+	(184.28, 49.93) --
+	(184.52, 49.93) --
+	(184.76, 49.93) --
+	(185.00, 49.93) --
+	(185.24, 49.93) --
+	(185.48, 49.93) --
+	(185.72, 49.93) --
+	(185.96, 49.93) --
+	(186.20, 49.93) --
+	(186.44, 49.93) --
+	(186.68, 49.93) --
+	(186.92, 49.93) --
+	(187.16, 49.93) --
+	(187.40, 49.93) --
+	(187.64, 49.93) --
+	(187.88, 49.93) --
+	(188.12, 49.93) --
+	(188.36, 49.93) --
+	(188.60, 49.93) --
+	(188.84, 49.93) --
+	(189.08, 49.93) --
+	(189.32, 49.93) --
+	(189.56, 49.93) --
+	(189.80, 49.93) --
+	(190.04, 49.93) --
+	(190.28, 49.93) --
+	(190.52, 49.93) --
+	(190.76, 49.93) --
+	(191.00, 49.93) --
+	(191.24, 49.93) --
+	(191.48, 49.93) --
+	(191.72, 49.93) --
+	(191.96, 49.93) --
+	(192.20, 49.93) --
+	(192.44, 49.93) --
+	(192.68, 49.93) --
+	(192.92, 49.93) --
+	(193.16, 49.93) --
+	(193.40, 49.93) --
+	(193.64, 49.93) --
+	(193.88, 49.93) --
+	(194.12, 49.93) --
+	(194.36, 49.93) --
+	(194.60, 49.93) --
+	(194.84, 49.93) --
+	(195.08, 49.93) --
+	(195.32, 49.93) --
+	(195.56, 49.93) --
+	(195.80, 49.93) --
+	(196.04, 49.93) --
+	(196.28, 49.93) --
+	(196.52, 49.93) --
+	(196.76, 49.93) --
+	(197.00, 49.93) --
+	(197.24, 49.93) --
+	(197.48, 49.93) --
+	(197.72, 49.93) --
+	(197.96, 49.93) --
+	(198.20, 49.93) --
+	(198.44, 49.93) --
+	(198.68, 49.93) --
+	(198.92, 49.93) --
+	(199.16, 49.93) --
+	(199.40, 49.93) --
+	(199.64, 49.93) --
+	(199.88, 49.93) --
+	(200.12, 49.93) --
+	(200.36, 49.93) --
+	(200.60, 49.93) --
+	(200.84, 49.93) --
+	(201.08, 49.93) --
+	(201.32, 49.93) --
+	(201.56, 49.93) --
+	(201.80, 49.93) --
+	(202.04, 49.93) --
+	(202.28, 49.93) --
+	(202.52, 49.93) --
+	(202.76, 49.93) --
+	(203.00, 49.93) --
 	(203.24, 49.93) --
@@
 	(209.48, 49.93) --
-	(209.72, 49.92) --
-	(209.96, 49.92) --
-	(210.20, 49.92) --
-	(210.44, 49.92) --
-	(210.68, 49.92) --
-	(210.92, 49.92) --
-	(211.16, 49.92) --
-	(211.40, 49.91) --
-	(211.64, 49.91) --
-	(211.88, 49.91) --
-	(212.12, 49.91) --
-	(212.36, 49.91) --
-	(212.60, 49.91) --
-	(212.84, 49.91) --
-	(213.08, 49.91) --
-	(213.32, 49.91) --
-	(213.56, 49.90) --
-	(213.80, 49.90) --
-	(214.04, 49.90) --
-	(214.28, 49.91) --
-	(214.52, 49.91) --
-	(214.76, 49.91) --
-	(215.00, 49.91) --
-	(215.24, 49.91) --
-	(215.48, 49.91) --
-	(215.72, 49.91) --
-	(215.96, 49.91) --
-	(216.20, 49.92) --
-	(216.44, 49.92) --
-	(216.68, 49.92) --
-	(216.92, 49.92) --
-	(217.16, 49.92) --
-	(217.40, 49.92) --
-	(217.64, 49.92) --
+	(209.72, 49.93) --
+	(209.96, 49.93) --
+	(210.20, 49.93) --
+	(210.44, 49.93) --
+	(210.68, 49.93) --
+	(210.92, 49.93) --
+	(211.16, 49.93) --
+	(211.40, 49.93) --
+	(211.64, 49.93) --
+	(211.88, 49.93) --
+	(212.12, 49.93) --
+	(212.36, 49.93) --
+	(212.60, 49.93) --
+	(212.84, 49.93) --
+	(213.08, 49.93) --
+	(213.32, 49.93) --
+	(213.56, 49.93) --
+	(213.80, 49.93) --
+	(214.04, 49.93) --
+	(214.28, 49.93) --
+	(214.52, 49.93) --
+	(214.76, 49.93) --
+	(215.00, 49.93) --
+	(215.24, 49.93) --
+	(215.48, 49.93) --
+	(215.72, 49.93) --
+	(215.96, 49.93) --
+	(216.20, 49.93) --
+	(216.44, 49.93) --
+	(216.68, 49.93) --
+	(216.92, 49.93) --
+	(217.16, 49.93) --
+	(217.40, 49.93) --
+	(217.64, 49.93) --
 	(217.88, 49.93) --
@@
 	(223.88, 49.93) --
-	(224.12, 49.92) --
-	(224.36, 49.92) --
-	(224.60, 49.92) --
-	(224.84, 49.92) --
-	(225.08, 49.92) --
-	(225.32, 49.92) --
-	(225.56, 49.92) --
-	(225.80, 49.91) --
-	(226.04, 49.91) --
-	(226.28, 49.91) --
-	(226.52, 49.91) --
-	(226.76, 49.91) --
-	(227.00, 49.91) --
-	(227.24, 49.91) --
-	(227.48, 49.91) --
-	(227.72, 49.90) --
-	(227.96, 49.90) --
+	(224.12, 49.93) --
+	(224.36, 49.93) --
+	(224.60, 49.93) --
+	(224.84, 49.93) --
+	(225.08, 49.93) --
+	(225.32, 49.93) --
+	(225.56, 49.93) --
+	(225.80, 49.93) --
+	(226.04, 49.93) --
+	(226.28, 49.93) --
+	(226.52, 49.93) --
+	(226.76, 49.93) --
+	(227.00, 49.93) --
+	(227.24, 49.93) --
+	(227.48, 49.93) --
+	(227.72, 49.93) --
+	(227.96, 49.93) --
 	(227.96, 49.96) --
@@
 	(105.33, 50.01) --
-	(105.33, 49.85) --
+	(105.33, 49.93) --
 	cycle;
@@
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) -- (395.21, 66.20);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) -- (364.41, 88.18);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) -- (364.41,110.17);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) --
+	(395.21, 66.20);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) --
+	(395.21, 66.20);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (395.21, 66.20) --
+	(395.21, 66.20);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) --
+	(364.41, 88.18);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) --
+	(364.41, 88.18);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41, 88.18) --
+	(364.41, 88.18);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) --
+	(364.41,110.17);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) --
+	(364.41,110.17);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.41,110.17) --
+	(364.41,110.17);
 \definecolor{drawColor}{RGB}{255,165,0}
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (320.06,198.12) -- (393.26,198.12);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (393.26,198.12) --
+	(393.26,198.12);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (393.26,198.12) --
+	(320.06,198.12);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (320.06,198.12) --
+	(320.06,198.12);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) -- (349.00,132.16);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) --
+	(349.00,132.16);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) --
+	(349.00,132.16);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (349.00,132.16) --
+	(349.00,132.16);
 \definecolor{drawColor}{RGB}{255,165,0}
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (340.83,154.14) -- (468.86,154.14);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (349.40,176.13) -- (364.04,176.13);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (468.86,154.14) --
+	(468.86,154.14);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (468.86,154.14) --
+	(340.83,154.14);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (340.83,154.14) --
+	(340.83,154.14);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.04,176.13) --
+	(364.04,176.13);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (364.04,176.13) --
+	(349.40,176.13);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (349.40,176.13) --
+	(349.40,176.13);
 \definecolor{drawColor}{RGB}{255,0,0}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (327.85, 44.18) --
-	(328.07, 44.18) --
-	(328.28, 44.18) --
-	(328.50, 44.18) --
-	(328.72, 44.18) --
-	(328.93, 44.18) --
-	(329.15, 44.18) --
-	(329.36, 44.17) --
-	(329.58, 44.17) --
-	(329.79, 44.17) --
-	(330.01, 44.17) --
-	(330.23, 44.16) --
-	(330.44, 44.16) --
-	(330.66, 44.16) --
-	(330.87, 44.15) --
-	(331.09, 44.15) --
-	(331.30, 44.14) --
-	(331.52, 44.14) --
-	(331.74, 44.13) --
-	(331.95, 44.13) --
-	(332.17, 44.13) --
-	(332.38, 44.12) --
-	(332.60, 44.12) --
-	(332.81, 44.12) --
-	(333.03, 44.11) --
-	(333.25, 44.11) --
-	(333.46, 44.11) --
-	(333.68, 44.10) --
-	(333.89, 44.10) --
-	(334.11, 44.10) --
-	(334.32, 44.09) --
-	(334.54, 44.09) --
-	(334.76, 44.08) --
-	(334.97, 44.08) --
-	(335.19, 44.07) --
-	(335.40, 44.07) --
-	(335.62, 44.06) --
-	(335.83, 44.06) --
-	(336.05, 44.05) --
-	(336.27, 44.05) --
-	(336.48, 44.05) --
-	(336.70, 44.05) --
-	(336.91, 44.05) --
-	(337.13, 44.05) --
-	(337.34, 44.05) --
-	(337.56, 44.05) --
-	(337.78, 44.05) --
-	(337.99, 44.05) --
-	(338.21, 44.06) --
-	(338.42, 44.06) --
-	(338.64, 44.06) --
-	(338.86, 44.07) --
-	(339.07, 44.07) --
-	(339.29, 44.07) --
-	(339.50, 44.08) --
-	(339.72, 44.08) --
-	(339.93, 44.08) --
-	(340.15, 44.09) --
-	(340.37, 44.09) --
-	(340.58, 44.09) --
-	(340.80, 44.09) --
-	(341.01, 44.08) --
-	(341.23, 44.08) --
-	(341.44, 44.07) --
-	(341.66, 44.05) --
-	(341.88, 44.04) --
-	(342.09, 44.02) --
-	(342.31, 43.99) --
-	(342.52, 43.97) --
-	(342.74, 43.94) --
-	(342.95, 43.91) --
-	(343.17, 43.88) --
-	(343.39, 43.85) --
-	(343.60, 43.81) --
-	(343.82, 43.78) --
-	(344.03, 43.75) --
-	(344.25, 43.72) --
-	(344.46, 43.68) --
-	(344.68, 43.65) --
-	(344.90, 43.63) --
-	(345.11, 43.60) --
-	(345.33, 43.57) --
-	(345.54, 43.54) --
-	(345.76, 43.52) --
-	(345.97, 43.49) --
-	(346.19, 43.46) --
-	(346.41, 43.43) --
-	(346.62, 43.40) --
-	(346.84, 43.37) --
-	(347.05, 43.34) --
-	(347.27, 43.30) --
-	(347.48, 43.26) --
-	(347.70, 43.22) --
-	(347.92, 43.17) --
-	(348.13, 43.12) --
-	(348.35, 43.07) --
-	(348.56, 43.01) --
-	(348.78, 42.95) --
-	(348.99, 42.89) --
-	(349.21, 42.83) --
-	(349.43, 42.76) --
-	(349.64, 42.70) --
-	(349.86, 42.63) --
-	(350.07, 42.56) --
-	(350.29, 42.49) --
-	(350.50, 42.42) --
-	(350.72, 42.34) --
-	(350.94, 42.27) --
-	(351.15, 42.20) --
-	(351.37, 42.12) --
-	(351.58, 42.05) --
-	(351.80, 41.98) --
-	(352.02, 41.91) --
-	(352.23, 41.84) --
-	(352.45, 41.77) --
-	(352.66, 41.71) --
-	(352.88, 41.64) --
-	(353.09, 41.57) --
-	(353.31, 41.50) --
-	(353.53, 41.42) --
-	(353.74, 41.34) --
-	(353.96, 41.25) --
-	(354.17, 41.16) --
-	(354.39, 41.05) --
-	(354.60, 40.93) --
-	(354.82, 40.80) --
-	(355.04, 40.65) --
-	(355.25, 40.49) --
-	(355.47, 40.33) --
-	(355.68, 40.15) --
-	(355.90, 39.96) --
-	(356.11, 39.77) --
-	(356.33, 39.57) --
-	(356.55, 39.37) --
-	(356.76, 39.16) --
-	(356.98, 38.96) --
-	(357.19, 38.76) --
-	(357.41, 38.56) --
-	(357.62, 38.36) --
-	(357.84, 38.17) --
-	(358.06, 37.98) --
-	(358.27, 37.79) --
-	(358.49, 37.61) --
-	(358.70, 37.43) --
-	(358.92, 37.25) --
-	(359.13, 37.08) --
-	(359.35, 36.90) --
-	(359.57, 36.72) --
-	(359.78, 36.54) --
-	(360.00, 36.36) --
-	(360.21, 36.18) --
-	(360.43, 35.99) --
-	(360.64, 35.80) --
-	(360.86, 35.62) --
-	(361.08, 35.43) --
-	(361.29, 35.25) --
-	(361.51, 35.07) --
-	(361.72, 34.89) --
-	(361.94, 34.72) --
-	(362.15, 34.55) --
-	(362.37, 34.39) --
-	(362.59, 34.23) --
-	(362.80, 34.08) --
-	(363.02, 33.95) --
-	(363.23, 33.82) --
-	(363.45, 33.70) --
-	(363.66, 33.60) --
-	(363.88, 33.51) --
-	(364.10, 33.43) --
-	(364.31, 33.36) --
-	(364.53, 33.31) --
-	(364.74, 33.27) --
-	(364.96, 33.24) --
-	(365.18, 33.22) --
-	(365.39, 33.22) --
-	(365.61, 33.22) --
-	(365.82, 33.24) --
-	(366.04, 33.27) --
-	(366.25, 33.31) --
-	(366.47, 33.36) --
-	(366.69, 33.41) --
-	(366.90, 33.48) --
-	(367.12, 33.55) --
-	(367.33, 33.63) --
-	(367.55, 33.71) --
-	(367.76, 33.80) --
-	(367.98, 33.90) --
-	(368.20, 33.99) --
-	(368.41, 34.10) --
-	(368.63, 34.20) --
-	(368.84, 34.31) --
-	(369.06, 34.42) --
-	(369.27, 34.53) --
-	(369.49, 34.64) --
-	(369.71, 34.76) --
-	(369.92, 34.88) --
-	(370.14, 34.99) --
-	(370.35, 35.11) --
-	(370.57, 35.24) --
-	(370.78, 35.36) --
-	(371.00, 35.49) --
-	(371.22, 35.61) --
-	(371.43, 35.75) --
-	(371.65, 35.88) --
-	(371.86, 36.02) --
-	(372.08, 36.17) --
-	(372.29, 36.32) --
-	(372.51, 36.47) --
-	(372.73, 36.63) --
-	(372.94, 36.79) --
-	(373.16, 36.96) --
-	(373.37, 37.12) --
-	(373.59, 37.28) --
-	(373.80, 37.44) --
-	(374.02, 37.60) --
-	(374.24, 37.76) --
-	(374.45, 37.92) --
-	(374.67, 38.07) --
-	(374.88, 38.23) --
-	(375.10, 38.38) --
-	(375.31, 38.54) --
-	(375.53, 38.71) --
-	(375.75, 38.87) --
-	(375.96, 39.04) --
-	(376.18, 39.22) --
-	(376.39, 39.39) --
-	(376.61, 39.57) --
-	(376.82, 39.75) --
-	(377.04, 39.92) --
-	(377.26, 40.09) --
-	(377.47, 40.26) --
-	(377.69, 40.42) --
-	(377.90, 40.57) --
-	(378.12, 40.71) --
-	(378.33, 40.84) --
-	(378.55, 40.96) --
-	(378.77, 41.07) --
-	(378.98, 41.17) --
-	(379.20, 41.26) --
-	(379.41, 41.34) --
-	(379.63, 41.42) --
-	(379.85, 41.48) --
-	(380.06, 41.55) --
-	(380.28, 41.61) --
-	(380.49, 41.66) --
-	(380.71, 41.72) --
-	(380.92, 41.78) --
-	(381.14, 41.83) --
-	(381.36, 41.89) --
-	(381.57, 41.95) --
-	(381.79, 42.01) --
-	(382.00, 42.07) --
-	(382.22, 42.13) --
-	(382.43, 42.19) --
-	(382.65, 42.25) --
-	(382.87, 42.30) --
-	(383.08, 42.36) --
-	(383.30, 42.42) --
-	(383.51, 42.47) --
-	(383.73, 42.52) --
-	(383.94, 42.57) --
-	(384.16, 42.62) --
-	(384.38, 42.67) --
-	(384.59, 42.72) --
-	(384.81, 42.77) --
-	(385.02, 42.82) --
-	(385.24, 42.87) --
-	(385.45, 42.92) --
-	(385.67, 42.97) --
-	(385.89, 43.02) --
-	(386.10, 43.07) --
-	(386.32, 43.12) --
-	(386.53, 43.17) --
-	(386.75, 43.21) --
-	(386.96, 43.26) --
-	(387.18, 43.30) --
-	(387.40, 43.34) --
-	(387.61, 43.37) --
-	(387.83, 43.40) --
-	(388.04, 43.43) --
-	(388.26, 43.46) --
-	(388.47, 43.48) --
-	(388.69, 43.50) --
-	(388.91, 43.52) --
-	(389.12, 43.54) --
-	(389.34, 43.56) --
-	(389.55, 43.58) --
-	(389.77, 43.60) --
-	(389.98, 43.62) --
-	(390.20, 43.64) --
-	(390.42, 43.66) --
-	(390.63, 43.68) --
-	(390.85, 43.71) --
-	(391.06, 43.73) --
-	(391.28, 43.75) --
-	(391.49, 43.77) --
-	(391.71, 43.79) --
-	(391.93, 43.80) --
-	(392.14, 43.82) --
-	(392.36, 43.83) --
-	(392.57, 43.85) --
-	(392.79, 43.86) --
-	(393.01, 43.87) --
-	(393.22, 43.87) --
-	(393.44, 43.88) --
-	(393.65, 43.89) --
-	(393.87, 43.89) --
-	(394.08, 43.89) --
-	(394.30, 43.90) --
-	(394.52, 43.90) --
-	(394.73, 43.90) --
-	(394.95, 43.91) --
-	(395.16, 43.91) --
-	(395.38, 43.92) --
-	(395.59, 43.92) --
-	(395.81, 43.93) --
-	(396.03, 43.93) --
-	(396.24, 43.94) --
-	(396.46, 43.95) --
-	(396.67, 43.95) --
-	(396.89, 43.96) --
-	(397.10, 43.97) --
-	(397.32, 43.98) --
-	(397.54, 43.99) --
-	(397.75, 44.00) --
-	(397.97, 44.01) --
-	(398.18, 44.02) --
-	(398.40, 44.02) --
-	(398.61, 44.03) --
-	(398.83, 44.04) --
-	(399.05, 44.05) --
-	(399.26, 44.05) --
-	(399.48, 44.06) --
-	(399.69, 44.06) --
-	(399.91, 44.06) --
-	(400.12, 44.06) --
-	(400.34, 44.06) --
-	(400.56, 44.06) --
-	(400.77, 44.06) --
-	(400.99, 44.06) --
-	(401.20, 44.05) --
-	(401.42, 44.05) --
-	(401.63, 44.05) --
-	(401.85, 44.04) --
-	(402.07, 44.04) --
-	(402.28, 44.04) --
-	(402.50, 44.03) --
-	(402.71, 44.03) --
-	(402.93, 44.03) --
-	(403.14, 44.04) --
-	(403.36, 44.04) --
-	(403.58, 44.05) --
-	(403.79, 44.05) --
-	(404.01, 44.06) --
-	(404.22, 44.07) --
-	(404.44, 44.07) --
-	(404.65, 44.08) --
-	(404.87, 44.09) --
-	(405.09, 44.09) --
-	(405.30, 44.10) --
-	(405.52, 44.11) --
-	(405.73, 44.11) --
-	(405.95, 44.12) --
-	(406.16, 44.12) --
-	(406.38, 44.12) --
-	(406.60, 44.13) --
-	(406.81, 44.13) --
-	(407.03, 44.13) --
-	(407.24, 44.13) --
-	(407.46, 44.14) --
-	(407.68, 44.14) --
-	(407.89, 44.14) --
-	(408.11, 44.15) --
-	(408.32, 44.15) --
-	(408.54, 44.15) --
-	(408.75, 44.16) --
-	(408.97, 44.16) --
-	(409.19, 44.16) --
-	(409.40, 44.17) --
-	(409.62, 44.17) --
-	(409.83, 44.17) --
-	(410.05, 44.18) --
-	(410.26, 44.18) --
-	(410.48, 44.19) --
-	(410.70, 44.19) --
-	(410.91, 44.19) --
-	(411.13, 44.20) --
-	(411.34, 44.20) --
-	(411.56, 44.20) --
-	(411.77, 44.20) --
-	(411.99, 44.20) --
+\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (327.85, 44.21) --
+	(328.07, 44.21) --
+	(328.28, 44.21) --
+	(328.50, 44.21) --
+	(328.72, 44.21) --
+	(328.93, 44.21) --
+	(329.15, 44.21) --
+	(329.36, 44.21) --
+	(329.58, 44.21) --
+	(329.79, 44.21) --
+	(330.01, 44.21) --
+	(330.23, 44.21) --
+	(330.44, 44.21) --
+	(330.66, 44.21) --
+	(330.87, 44.21) --
+	(331.09, 44.21) --
+	(331.30, 44.21) --
+	(331.52, 44.21) --
+	(331.74, 44.21) --
+	(331.95, 44.21) --
+	(332.17, 44.21) --
+	(332.38, 44.21) --
+	(332.60, 44.21) --
+	(332.81, 44.21) --
+	(333.03, 44.21) --
+	(333.25, 44.21) --
+	(333.46, 44.21) --
+	(333.68, 44.21) --
+	(333.89, 44.21) --
+	(334.11, 44.21) --
+	(334.32, 44.21) --
+	(334.54, 44.21) --
+	(334.76, 44.21) --
+	(334.97, 44.21) --
+	(335.19, 44.21) --
+	(335.40, 44.21) --
+	(335.62, 44.21) --
+	(335.83, 44.21) --
+	(336.05, 44.21) --
+	(336.27, 44.21) --
+	(336.48, 44.21) --
+	(336.70, 44.21) --
+	(336.91, 44.21) --
+	(337.13, 44.21) --
+	(337.34, 44.21) --
+	(337.56, 44.21) --
+	(337.78, 44.21) --
+	(337.99, 44.21) --
+	(338.21, 44.21) --
+	(338.42, 44.21) --
+	(338.64, 44.21) --
+	(338.86, 44.21) --
+	(339.07, 44.21) --
+	(339.29, 44.21) --
+	(339.50, 44.21) --
+	(339.72, 44.21) --
+	(339.93, 44.21) --
+	(340.15, 44.21) --
+	(340.37, 44.21) --
+	(340.58, 44.21) --
+	(340.80, 44.21) --
+	(341.01, 44.21) --
+	(341.23, 44.21) --
+	(341.44, 44.21) --
+	(341.66, 44.21) --
+	(341.88, 44.21) --
+	(342.09, 44.21) --
+	(342.31, 44.21) --
+	(342.52, 44.21) --
+	(342.74, 44.21) --
+	(342.95, 44.21) --
+	(343.17, 44.21) --
+	(343.39, 44.21) --
+	(343.60, 44.21) --
+	(343.82, 44.21) --
+	(344.03, 44.21) --
+	(344.25, 44.21) --
+	(344.46, 44.21) --
+	(344.68, 44.21) --
+	(344.90, 44.21) --
+	(345.11, 44.21) --
+	(345.33, 44.21) --
+	(345.54, 44.21) --
+	(345.76, 44.21) --
+	(345.97, 44.21) --
+	(346.19, 44.21) --
+	(346.41, 44.21) --
+	(346.62, 44.21) --
+	(346.84, 44.21) --
+	(347.05, 44.21) --
+	(347.27, 44.21) --
+	(347.48, 44.21) --
+	(347.70, 44.21) --
+	(347.92, 44.21) --
+	(348.13, 44.21) --
+	(348.35, 44.21) --
+	(348.56, 44.21) --
+	(348.78, 44.21) --
+	(348.99, 44.21) --
+	(349.21, 44.21) --
+	(349.43, 44.21) --
+	(349.64, 44.21) --
+	(349.86, 44.21) --
+	(350.07, 44.21) --
+	(350.29, 44.21) --
+	(350.50, 44.21) --
+	(350.72, 44.21) --
+	(350.94, 44.21) --
+	(351.15, 44.21) --
+	(351.37, 44.21) --
+	(351.58, 44.21) --
+	(351.80, 44.21) --
+	(352.02, 44.21) --
+	(352.23, 44.21) --
+	(352.45, 44.21) --
+	(352.66, 44.21) --
+	(352.88, 44.21) --
+	(353.09, 44.21) --
+	(353.31, 44.21) --
+	(353.53, 44.21) --
+	(353.74, 44.21) --
+	(353.96, 44.21) --
+	(354.17, 44.21) --
+	(354.39, 44.21) --
+	(354.60, 44.21) --
+	(354.82, 44.21) --
+	(355.04, 44.21) --
+	(355.25, 44.21) --
+	(355.47, 44.21) --
+	(355.68, 44.21) --
+	(355.90, 44.21) --
+	(356.11, 44.21) --
+	(356.33, 44.21) --
+	(356.55, 44.21) --
+	(356.76, 44.21) --
+	(356.98, 44.21) --
+	(357.19, 44.21) --
+	(357.41, 44.21) --
+	(357.62, 44.21) --
+	(357.84, 44.21) --
+	(358.06, 44.21) --
+	(358.27, 44.21) --
+	(358.49, 44.21) --
+	(358.70, 44.21) --
+	(358.92, 44.21) --
+	(359.13, 44.21) --
+	(359.35, 44.21) --
+	(359.57, 44.21) --
+	(359.78, 44.21) --
+	(360.00, 44.21) --
+	(360.21, 44.21) --
+	(360.43, 44.21) --
+	(360.64, 44.21) --
+	(360.86, 44.21) --
+	(361.08, 44.21) --
+	(361.29, 44.21) --
+	(361.51, 44.21) --
+	(361.72, 44.21) --
+	(361.94, 44.21) --
+	(362.15, 44.21) --
+	(362.37, 44.21) --
+	(362.59, 44.21) --
+	(362.80, 44.21) --
+	(363.02, 44.21) --
+	(363.23, 44.21) --
+	(363.45, 44.21) --
+	(363.66, 44.21) --
+	(363.88, 44.21) --
+	(364.10, 44.21) --
+	(364.31, 44.21) --
+	(364.53, 44.21) --
+	(364.74, 44.21) --
+	(364.96, 44.21) --
+	(365.18, 44.21) --
+	(365.39, 44.21) --
+	(365.61, 44.21) --
+	(365.82, 44.21) --
+	(366.04, 44.21) --
+	(366.25, 44.21) --
+	(366.47, 44.21) --
+	(366.69, 44.21) --
+	(366.90, 44.21) --
+	(367.12, 44.21) --
+	(367.33, 44.21) --
+	(367.55, 44.21) --
+	(367.76, 44.21) --
+	(367.98, 44.21) --
+	(368.20, 44.21) --
+	(368.41, 44.21) --
+	(368.63, 44.21) --
+	(368.84, 44.21) --
+	(369.06, 44.21) --
+	(369.27, 44.21) --
+	(369.49, 44.21) --
+	(369.71, 44.21) --
+	(369.92, 44.21) --
+	(370.14, 44.21) --
+	(370.35, 44.21) --
+	(370.57, 44.21) --
+	(370.78, 44.21) --
+	(371.00, 44.21) --
+	(371.22, 44.21) --
+	(371.43, 44.21) --
+	(371.65, 44.21) --
+	(371.86, 44.21) --
+	(372.08, 44.21) --
+	(372.29, 44.21) --
+	(372.51, 44.21) --
+	(372.73, 44.21) --
+	(372.94, 44.21) --
+	(373.16, 44.21) --
+	(373.37, 44.21) --
+	(373.59, 44.21) --
+	(373.80, 44.21) --
+	(374.02, 44.21) --
+	(374.24, 44.21) --
+	(374.45, 44.21) --
+	(374.67, 44.21) --
+	(374.88, 44.21) --
+	(375.10, 44.21) --
+	(375.31, 44.21) --
+	(375.53, 44.21) --
+	(375.75, 44.21) --
+	(375.96, 44.21) --
+	(376.18, 44.21) --
+	(376.39, 44.21) --
+	(376.61, 44.21) --
+	(376.82, 44.21) --
+	(377.04, 44.21) --
+	(377.26, 44.21) --
+	(377.47, 44.21) --
+	(377.69, 44.21) --
+	(377.90, 44.21) --
+	(378.12, 44.21) --
+	(378.33, 44.21) --
+	(378.55, 44.21) --
+	(378.77, 44.21) --
+	(378.98, 44.21) --
+	(379.20, 44.21) --
+	(379.41, 44.21) --
+	(379.63, 44.21) --
+	(379.85, 44.21) --
+	(380.06, 44.21) --
+	(380.28, 44.21) --
+	(380.49, 44.21) --
+	(380.71, 44.21) --
+	(380.92, 44.21) --
+	(381.14, 44.21) --
+	(381.36, 44.21) --
+	(381.57, 44.21) --
+	(381.79, 44.21) --
+	(382.00, 44.21) --
+	(382.22, 44.21) --
+	(382.43, 44.21) --
+	(382.65, 44.21) --
+	(382.87, 44.21) --
+	(383.08, 44.21) --
+	(383.30, 44.21) --
+	(383.51, 44.21) --
+	(383.73, 44.21) --
+	(383.94, 44.21) --
+	(384.16, 44.21) --
+	(384.38, 44.21) --
+	(384.59, 44.21) --
+	(384.81, 44.21) --
+	(385.02, 44.21) --
+	(385.24, 44.21) --
+	(385.45, 44.21) --
+	(385.67, 44.21) --
+	(385.89, 44.21) --
+	(386.10, 44.21) --
+	(386.32, 44.21) --
+	(386.53, 44.21) --
+	(386.75, 44.21) --
+	(386.96, 44.21) --
+	(387.18, 44.21) --
+	(387.40, 44.21) --
+	(387.61, 44.21) --
+	(387.83, 44.21) --
+	(388.04, 44.21) --
+	(388.26, 44.21) --
+	(388.47, 44.21) --
+	(388.69, 44.21) --
+	(388.91, 44.21) --
+	(389.12, 44.21) --
+	(389.34, 44.21) --
+	(389.55, 44.21) --
+	(389.77, 44.21) --
+	(389.98, 44.21) --
+	(390.20, 44.21) --
+	(390.42, 44.21) --
+	(390.63, 44.21) --
+	(390.85, 44.21) --
+	(391.06, 44.21) --
+	(391.28, 44.21) --
+	(391.49, 44.21) --
+	(391.71, 44.21) --
+	(391.93, 44.21) --
+	(392.14, 44.21) --
+	(392.36, 44.21) --
+	(392.57, 44.21) --
+	(392.79, 44.21) --
+	(393.01, 44.21) --
+	(393.22, 44.21) --
+	(393.44, 44.21) --
+	(393.65, 44.21) --
+	(393.87, 44.21) --
+	(394.08, 44.21) --
+	(394.30, 44.21) --
+	(394.52, 44.21) --
+	(394.73, 44.21) --
+	(394.95, 44.21) --
+	(395.16, 44.21) --
+	(395.38, 44.21) --
+	(395.59, 44.21) --
+	(395.81, 44.21) --
+	(396.03, 44.21) --
+	(396.24, 44.21) --
+	(396.46, 44.21) --
+	(396.67, 44.21) --
+	(396.89, 44.21) --
+	(397.10, 44.21) --
+	(397.32, 44.21) --
+	(397.54, 44.21) --
+	(397.75, 44.21) --
+	(397.97, 44.21) --
+	(398.18, 44.21) --
+	(398.40, 44.21) --
+	(398.61, 44.21) --
+	(398.83, 44.21) --
+	(399.05, 44.21) --
+	(399.26, 44.21) --
+	(399.48, 44.21) --
+	(399.69, 44.21) --
+	(399.91, 44.21) --
+	(400.12, 44.21) --
+	(400.34, 44.21) --
+	(400.56, 44.21) --
+	(400.77, 44.21) --
+	(400.99, 44.21) --
+	(401.20, 44.21) --
+	(401.42, 44.21) --
+	(401.63, 44.21) --
+	(401.85, 44.21) --
+	(402.07, 44.21) --
+	(402.28, 44.21) --
+	(402.50, 44.21) --
+	(402.71, 44.21) --
+	(402.93, 44.21) --
+	(403.14, 44.21) --
+	(403.36, 44.21) --
+	(403.58, 44.21) --
+	(403.79, 44.21) --
+	(404.01, 44.21) --
+	(404.22, 44.21) --
+	(404.44, 44.21) --
+	(404.65, 44.21) --
+	(404.87, 44.21) --
+	(405.09, 44.21) --
+	(405.30, 44.21) --
+	(405.52, 44.21) --
+	(405.73, 44.21) --
+	(405.95, 44.21) --
+	(406.16, 44.21) --
+	(406.38, 44.21) --
+	(406.60, 44.21) --
+	(406.81, 44.21) --
+	(407.03, 44.21) --
+	(407.24, 44.21) --
+	(407.46, 44.21) --
+	(407.68, 44.21) --
+	(407.89, 44.21) --
+	(408.11, 44.21) --
+	(408.32, 44.21) --
+	(408.54, 44.21) --
+	(408.75, 44.21) --
+	(408.97, 44.21) --
+	(409.19, 44.21) --
+	(409.40, 44.21) --
+	(409.62, 44.21) --
+	(409.83, 44.21) --
+	(410.05, 44.21) --
+	(410.26, 44.21) --
+	(410.48, 44.21) --
+	(410.70, 44.21) --
+	(410.91, 44.21) --
+	(411.13, 44.21) --
+	(411.34, 44.21) --
+	(411.56, 44.21) --
+	(411.77, 44.21) --
+	(411.99, 44.21) --
 	(412.21, 44.21) --
@@
 	(426.23, 44.21) --
-	(426.44, 44.20) --
-	(426.66, 44.20) --
-	(426.88, 44.20) --
-	(427.09, 44.20) --
-	(427.31, 44.20) --
-	(427.52, 44.19) --
-	(427.74, 44.19) --
-	(427.95, 44.19) --
-	(428.17, 44.19) --
-	(428.39, 44.19) --
-	(428.60, 44.19) --
-	(428.82, 44.18) --
-	(429.03, 44.18) --
-	(429.25, 44.18) --
-	(429.46, 44.18) --
-	(429.68, 44.18) --
-	(429.90, 44.19) --
-	(430.11, 44.19) --
-	(430.33, 44.19) --
-	(430.54, 44.19) --
-	(430.76, 44.19) --
-	(430.97, 44.20) --
-	(431.19, 44.20) --
-	(431.41, 44.20) --
-	(431.62, 44.20) --
-	(431.84, 44.20) --
+	(426.44, 44.21) --
+	(426.66, 44.21) --
+	(426.88, 44.21) --
+	(427.09, 44.21) --
+	(427.31, 44.21) --
+	(427.52, 44.21) --
+	(427.74, 44.21) --
+	(427.95, 44.21) --
+	(428.17, 44.21) --
+	(428.39, 44.21) --
+	(428.60, 44.21) --
+	(428.82, 44.21) --
+	(429.03, 44.21) --
+	(429.25, 44.21) --
+	(429.46, 44.21) --
+	(429.68, 44.21) --
+	(429.90, 44.21) --
+	(430.11, 44.21) --
+	(430.33, 44.21) --
+	(430.54, 44.21) --
+	(430.76, 44.21) --
+	(430.97, 44.21) --
+	(431.19, 44.21) --
+	(431.41, 44.21) --
+	(431.62, 44.21) --
+	(431.84, 44.21) --
 	(432.05, 44.21) --
@@
 	(435.07, 44.21) --
-	(435.29, 44.20) --
-	(435.51, 44.20) --
-	(435.72, 44.20) --
-	(435.94, 44.20) --
-	(436.15, 44.20) --
-	(436.37, 44.20) --
-	(436.58, 44.19) --
-	(436.80, 44.19) --
-	(437.02, 44.19) --
-	(437.23, 44.19) --
-	(437.45, 44.19) --
-	(437.66, 44.18) --
-	(437.88, 44.18) --
-	(438.09, 44.18) --
+	(435.29, 44.21) --
+	(435.51, 44.21) --
+	(435.72, 44.21) --
+	(435.94, 44.21) --
+	(436.15, 44.21) --
+	(436.37, 44.21) --
+	(436.58, 44.21) --
+	(436.80, 44.21) --
+	(437.02, 44.21) --
+	(437.23, 44.21) --
+	(437.45, 44.21) --
+	(437.66, 44.21) --
+	(437.88, 44.21) --
+	(438.09, 44.21) --
 	(438.09, 44.24) --
@@
 	(327.85, 44.24) --
-	(327.85, 44.18) --
+	(327.85, 44.21) --
 	cycle;
@@
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) -- (634.38, 66.20);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) -- (597.67, 88.18);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) -- (575.98,110.17);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) -- (560.96,198.12);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) -- (597.67,132.16);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) -- (597.67,154.14);
-
-\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) -- (575.64,176.13);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) --
+	(634.38, 66.20);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) --
+	(634.38, 66.20);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (634.38, 66.20) --
+	(634.38, 66.20);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) --
+	(597.67, 88.18);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) --
+	(597.67, 88.18);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67, 88.18) --
+	(597.67, 88.18);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) --
+	(575.98,110.17);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) --
+	(575.98,110.17);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (575.98,110.17) --
+	(575.98,110.17);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) --
+	(560.96,198.12);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) --
+	(560.96,198.12);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (560.96,198.12) --
+	(560.96,198.12);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) --
+	(597.67,132.16);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) --
+	(597.67,132.16);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,132.16) --
+	(597.67,132.16);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) --
+	(597.67,154.14);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) --
+	(597.67,154.14);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (597.67,154.14) --
+	(597.67,154.14);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) --
+	(575.64,176.13);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) --
+	(575.64,176.13);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (575.64,176.13) --
+	(575.64,176.13);
 \definecolor{drawColor}{RGB}{255,0,0}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (571.29, 44.01) --
-	(571.56, 43.99) --
-	(571.83, 43.96) --
-	(572.10, 43.92) --
-	(572.37, 43.89) --
-	(572.65, 43.85) --
-	(572.92, 43.80) --
-	(573.19, 43.76) --
-	(573.46, 43.71) --
-	(573.73, 43.65) --
-	(574.00, 43.59) --
-	(574.27, 43.53) --
-	(574.54, 43.47) --
-	(574.81, 43.40) --
-	(575.08, 43.32) --
-	(575.36, 43.24) --
-	(575.63, 43.16) --
-	(575.90, 43.07) --
-	(576.17, 42.98) --
-	(576.44, 42.88) --
-	(576.71, 42.78) --
-	(576.98, 42.67) --
-	(577.25, 42.56) --
-	(577.52, 42.44) --
-	(577.79, 42.31) --
-	(578.07, 42.18) --
-	(578.34, 42.03) --
-	(578.61, 41.89) --
-	(578.88, 41.73) --
-	(579.15, 41.56) --
-	(579.42, 41.39) --
-	(579.69, 41.20) --
-	(579.96, 41.01) --
-	(580.23, 40.81) --
-	(580.50, 40.60) --
-	(580.77, 40.38) --
-	(581.05, 40.15) --
-	(581.32, 39.92) --
-	(581.59, 39.68) --
-	(581.86, 39.44) --
-	(582.13, 39.19) --
-	(582.40, 38.94) --
-	(582.67, 38.68) --
-	(582.94, 38.43) --
-	(583.21, 38.17) --
-	(583.48, 37.92) --
-	(583.76, 37.66) --
-	(584.03, 37.42) --
-	(584.30, 37.17) --
-	(584.57, 36.93) --
-	(584.84, 36.70) --
-	(585.11, 36.47) --
-	(585.38, 36.25) --
-	(585.65, 36.04) --
-	(585.92, 35.84) --
-	(586.19, 35.64) --
-	(586.47, 35.45) --
-	(586.74, 35.27) --
-	(587.01, 35.09) --
-	(587.28, 34.93) --
-	(587.55, 34.77) --
-	(587.82, 34.61) --
-	(588.09, 34.46) --
-	(588.36, 34.32) --
-	(588.63, 34.19) --
-	(588.90, 34.07) --
-	(589.18, 33.95) --
-	(589.45, 33.84) --
-	(589.72, 33.74) --
-	(589.99, 33.64) --
-	(590.26, 33.56) --
-	(590.53, 33.48) --
-	(590.80, 33.41) --
-	(591.07, 33.36) --
-	(591.34, 33.31) --
-	(591.61, 33.27) --
-	(591.89, 33.24) --
-	(592.16, 33.23) --
-	(592.43, 33.22) --
-	(592.70, 33.22) --
-	(592.97, 33.23) --
-	(593.24, 33.25) --
-	(593.51, 33.28) --
-	(593.78, 33.31) --
-	(594.05, 33.36) --
-	(594.32, 33.41) --
-	(594.59, 33.47) --
-	(594.87, 33.54) --
-	(595.14, 33.61) --
-	(595.41, 33.69) --
-	(595.68, 33.78) --
-	(595.95, 33.88) --
-	(596.22, 33.98) --
-	(596.49, 34.09) --
-	(596.76, 34.20) --
-	(597.03, 34.32) --
-	(597.30, 34.45) --
-	(597.58, 34.57) --
-	(597.85, 34.70) --
-	(598.12, 34.84) --
-	(598.39, 34.97) --
-	(598.66, 35.10) --
-	(598.93, 35.24) --
-	(599.20, 35.37) --
-	(599.47, 35.49) --
-	(599.74, 35.62) --
-	(600.01, 35.74) --
-	(600.29, 35.85) --
-	(600.56, 35.96) --
-	(600.83, 36.06) --
-	(601.10, 36.16) --
-	(601.37, 36.25) --
-	(601.64, 36.34) --
-	(601.91, 36.42) --
-	(602.18, 36.50) --
-	(602.45, 36.58) --
-	(602.72, 36.66) --
-	(603.00, 36.74) --
-	(603.27, 36.81) --
-	(603.54, 36.89) --
-	(603.81, 36.97) --
-	(604.08, 37.05) --
-	(604.35, 37.13) --
-	(604.62, 37.22) --
-	(604.89, 37.30) --
-	(605.16, 37.39) --
-	(605.43, 37.49) --
-	(605.70, 37.58) --
-	(605.98, 37.67) --
-	(606.25, 37.77) --
-	(606.52, 37.87) --
-	(606.79, 37.96) --
-	(607.06, 38.05) --
-	(607.33, 38.14) --
-	(607.60, 38.23) --
-	(607.87, 38.32) --
-	(608.14, 38.40) --
-	(608.41, 38.47) --
-	(608.69, 38.55) --
-	(608.96, 38.62) --
-	(609.23, 38.68) --
-	(609.50, 38.74) --
-	(609.77, 38.80) --
-	(610.04, 38.85) --
-	(610.31, 38.90) --
-	(610.58, 38.95) --
-	(610.85, 39.00) --
-	(611.12, 39.04) --
-	(611.40, 39.09) --
-	(611.67, 39.13) --
-	(611.94, 39.17) --
-	(612.21, 39.22) --
-	(612.48, 39.27) --
-	(612.75, 39.32) --
-	(613.02, 39.37) --
-	(613.29, 39.42) --
-	(613.56, 39.48) --
-	(613.83, 39.54) --
-	(614.11, 39.60) --
-	(614.38, 39.67) --
-	(614.65, 39.74) --
-	(614.92, 39.82) --
-	(615.19, 39.90) --
-	(615.46, 39.98) --
-	(615.73, 40.07) --
-	(616.00, 40.16) --
-	(616.27, 40.25) --
-	(616.54, 40.34) --
-	(616.81, 40.43) --
-	(617.09, 40.53) --
-	(617.36, 40.62) --
-	(617.63, 40.71) --
-	(617.90, 40.80) --
-	(618.17, 40.89) --
-	(618.44, 40.98) --
-	(618.71, 41.07) --
-	(618.98, 41.15) --
-	(619.25, 41.23) --
-	(619.52, 41.31) --
-	(619.80, 41.39) --
-	(620.07, 41.46) --
-	(620.34, 41.53) --
-	(620.61, 41.60) --
-	(620.88, 41.67) --
-	(621.15, 41.73) --
-	(621.42, 41.79) --
-	(621.69, 41.85) --
-	(621.96, 41.91) --
-	(622.23, 41.97) --
-	(622.51, 42.02) --
-	(622.78, 42.08) --
-	(623.05, 42.13) --
-	(623.32, 42.17) --
-	(623.59, 42.22) --
-	(623.86, 42.27) --
-	(624.13, 42.31) --
-	(624.40, 42.35) --
-	(624.67, 42.39) --
-	(624.94, 42.42) --
-	(625.22, 42.46) --
-	(625.49, 42.49) --
-	(625.76, 42.52) --
-	(626.03, 42.54) --
-	(626.30, 42.57) --
-	(626.57, 42.59) --
-	(626.84, 42.62) --
-	(627.11, 42.64) --
-	(627.38, 42.66) --
-	(627.65, 42.68) --
-	(627.92, 42.70) --
-	(628.20, 42.71) --
-	(628.47, 42.73) --
-	(628.74, 42.74) --
-	(629.01, 42.76) --
-	(629.28, 42.77) --
-	(629.55, 42.79) --
-	(629.82, 42.80) --
-	(630.09, 42.81) --
-	(630.36, 42.83) --
-	(630.63, 42.84) --
-	(630.91, 42.85) --
-	(631.18, 42.87) --
-	(631.45, 42.88) --
-	(631.72, 42.90) --
-	(631.99, 42.92) --
-	(632.26, 42.93) --
-	(632.53, 42.95) --
-	(632.80, 42.97) --
-	(633.07, 42.99) --
-	(633.34, 43.01) --
-	(633.62, 43.04) --
-	(633.89, 43.06) --
-	(634.16, 43.08) --
-	(634.43, 43.11) --
-	(634.70, 43.13) --
-	(634.97, 43.16) --
-	(635.24, 43.18) --
-	(635.51, 43.21) --
-	(635.78, 43.23) --
-	(636.05, 43.26) --
-	(636.33, 43.28) --
-	(636.60, 43.30) --
-	(636.87, 43.32) --
-	(637.14, 43.34) --
-	(637.41, 43.36) --
-	(637.68, 43.38) --
-	(637.95, 43.39) --
-	(638.22, 43.41) --
-	(638.49, 43.42) --
-	(638.76, 43.43) --
-	(639.03, 43.44) --
-	(639.31, 43.45) --
-	(639.58, 43.46) --
-	(639.85, 43.47) --
-	(640.12, 43.48) --
-	(640.39, 43.49) --
-	(640.66, 43.50) --
-	(640.93, 43.50) --
-	(641.20, 43.51) --
-	(641.47, 43.52) --
-	(641.74, 43.53) --
-	(642.02, 43.53) --
-	(642.29, 43.54) --
-	(642.56, 43.55) --
-	(642.83, 43.56) --
-	(643.10, 43.57) --
-	(643.37, 43.58) --
-	(643.64, 43.60) --
-	(643.91, 43.61) --
-	(644.18, 43.62) --
-	(644.45, 43.64) --
-	(644.73, 43.65) --
-	(645.00, 43.67) --
-	(645.27, 43.68) --
-	(645.54, 43.70) --
-	(645.81, 43.72) --
-	(646.08, 43.73) --
-	(646.35, 43.75) --
-	(646.62, 43.76) --
-	(646.89, 43.78) --
-	(647.16, 43.80) --
-	(647.44, 43.81) --
-	(647.71, 43.83) --
-	(647.98, 43.84) --
-	(648.25, 43.86) --
-	(648.52, 43.87) --
-	(648.79, 43.89) --
-	(649.06, 43.90) --
-	(649.33, 43.91) --
-	(649.60, 43.93) --
-	(649.87, 43.94) --
-	(650.15, 43.95) --
-	(650.42, 43.96) --
-	(650.69, 43.97) --
-	(650.96, 43.98) --
-	(651.23, 43.98) --
-	(651.50, 43.99) --
-	(651.77, 43.99) --
-	(652.04, 44.00) --
-	(652.31, 44.00) --
-	(652.58, 44.01) --
-	(652.85, 44.01) --
-	(653.13, 44.01) --
-	(653.40, 44.01) --
-	(653.67, 44.01) --
-	(653.94, 44.01) --
-	(654.21, 44.01) --
-	(654.48, 44.01) --
-	(654.75, 44.01) --
-	(655.02, 44.02) --
-	(655.29, 44.02) --
-	(655.56, 44.02) --
-	(655.84, 44.02) --
-	(656.11, 44.02) --
-	(656.38, 44.03) --
-	(656.65, 44.03) --
-	(656.92, 44.03) --
-	(657.19, 44.04) --
-	(657.46, 44.04) --
-	(657.73, 44.05) --
-	(658.00, 44.05) --
-	(658.27, 44.06) --
-	(658.55, 44.07) --
-	(658.82, 44.07) --
-	(659.09, 44.08) --
-	(659.36, 44.08) --
-	(659.63, 44.09) --
-	(659.90, 44.09) --
-	(660.17, 44.10) --
-	(660.44, 44.10) --
-	(660.71, 44.11) --
-	(660.98, 44.11) --
-	(661.26, 44.12) --
-	(661.53, 44.12) --
-	(661.80, 44.13) --
-	(662.07, 44.13) --
-	(662.34, 44.13) --
-	(662.61, 44.14) --
-	(662.88, 44.14) --
-	(663.15, 44.14) --
-	(663.42, 44.15) --
-	(663.69, 44.15) --
-	(663.96, 44.15) --
-	(664.24, 44.16) --
-	(664.51, 44.16) --
-	(664.78, 44.16) --
-	(665.05, 44.16) --
-	(665.32, 44.16) --
-	(665.59, 44.16) --
-	(665.86, 44.17) --
-	(666.13, 44.17) --
-	(666.40, 44.17) --
-	(666.67, 44.17) --
-	(666.95, 44.16) --
-	(667.22, 44.16) --
-	(667.49, 44.16) --
-	(667.76, 44.16) --
-	(668.03, 44.16) --
-	(668.30, 44.16) --
-	(668.57, 44.16) --
-	(668.84, 44.15) --
-	(669.11, 44.15) --
-	(669.38, 44.15) --
-	(669.66, 44.15) --
-	(669.93, 44.15) --
-	(670.20, 44.15) --
-	(670.47, 44.15) --
-	(670.74, 44.14) --
-	(671.01, 44.14) --
-	(671.28, 44.14) --
-	(671.55, 44.14) --
-	(671.82, 44.15) --
-	(672.09, 44.15) --
-	(672.37, 44.15) --
-	(672.64, 44.15) --
-	(672.91, 44.15) --
-	(673.18, 44.15) --
-	(673.45, 44.15) --
-	(673.72, 44.15) --
-	(673.99, 44.15) --
-	(674.26, 44.15) --
-	(674.53, 44.15) --
-	(674.80, 44.15) --
-	(675.07, 44.15) --
-	(675.35, 44.15) --
-	(675.62, 44.15) --
-	(675.89, 44.15) --
-	(676.16, 44.15) --
-	(676.43, 44.15) --
-	(676.70, 44.15) --
-	(676.97, 44.15) --
-	(677.24, 44.15) --
-	(677.51, 44.15) --
-	(677.78, 44.15) --
-	(678.06, 44.15) --
-	(678.33, 44.15) --
-	(678.60, 44.15) --
-	(678.87, 44.14) --
-	(679.14, 44.14) --
-	(679.41, 44.14) --
-	(679.68, 44.14) --
-	(679.95, 44.14) --
-	(680.22, 44.14) --
-	(680.49, 44.14) --
-	(680.77, 44.14) --
-	(681.04, 44.14) --
-	(681.31, 44.14) --
-	(681.58, 44.14) --
-	(681.85, 44.14) --
-	(682.12, 44.14) --
-	(682.39, 44.14) --
-	(682.66, 44.14) --
-	(682.93, 44.14) --
-	(683.20, 44.13) --
-	(683.48, 44.13) --
-	(683.75, 44.13) --
-	(684.02, 44.13) --
-	(684.29, 44.13) --
-	(684.56, 44.13) --
-	(684.83, 44.13) --
-	(685.10, 44.13) --
-	(685.37, 44.13) --
-	(685.64, 44.13) --
-	(685.91, 44.13) --
-	(686.18, 44.13) --
-	(686.46, 44.14) --
-	(686.73, 44.14) --
-	(687.00, 44.14) --
-	(687.27, 44.15) --
-	(687.54, 44.15) --
-	(687.81, 44.15) --
-	(688.08, 44.16) --
-	(688.35, 44.16) --
-	(688.62, 44.16) --
-	(688.89, 44.17) --
-	(689.17, 44.17) --
-	(689.44, 44.17) --
-	(689.71, 44.18) --
-	(689.98, 44.18) --
-	(690.25, 44.18) --
-	(690.52, 44.18) --
-	(690.79, 44.18) --
-	(691.06, 44.19) --
-	(691.33, 44.19) --
-	(691.60, 44.19) --
-	(691.88, 44.19) --
-	(692.15, 44.19) --
-	(692.42, 44.19) --
-	(692.69, 44.19) --
-	(692.96, 44.19) --
-	(693.23, 44.19) --
-	(693.50, 44.19) --
-	(693.77, 44.19) --
-	(694.04, 44.19) --
-	(694.31, 44.19) --
-	(694.59, 44.19) --
-	(694.86, 44.19) --
-	(695.13, 44.19) --
-	(695.40, 44.19) --
-	(695.67, 44.19) --
-	(695.94, 44.19) --
-	(696.21, 44.19) --
-	(696.48, 44.19) --
-	(696.75, 44.20) --
-	(697.02, 44.20) --
-	(697.29, 44.20) --
-	(697.57, 44.20) --
-	(697.84, 44.20) --
-	(698.11, 44.20) --
-	(698.38, 44.20) --
-	(698.65, 44.20) --
-	(698.92, 44.20) --
-	(699.19, 44.20) --
-	(699.46, 44.20) --
-	(699.73, 44.20) --
-	(700.00, 44.20) --
-	(700.28, 44.20) --
-	(700.55, 44.20) --
-	(700.82, 44.20) --
-	(701.09, 44.20) --
-	(701.36, 44.20) --
-	(701.63, 44.20) --
-	(701.90, 44.20) --
-	(702.17, 44.20) --
-	(702.44, 44.19) --
-	(702.71, 44.19) --
-	(702.99, 44.19) --
-	(703.26, 44.19) --
-	(703.53, 44.19) --
-	(703.80, 44.19) --
-	(704.07, 44.19) --
-	(704.34, 44.18) --
-	(704.61, 44.18) --
-	(704.88, 44.18) --
-	(705.15, 44.18) --
-	(705.42, 44.18) --
-	(705.70, 44.18) --
-	(705.97, 44.18) --
-	(706.24, 44.18) --
-	(706.51, 44.18) --
-	(706.78, 44.18) --
-	(707.05, 44.18) --
-	(707.32, 44.18) --
-	(707.59, 44.18) --
-	(707.86, 44.18) --
-	(708.13, 44.18) --
-	(708.41, 44.18) --
-	(708.68, 44.18) --
-	(708.95, 44.18) --
-	(709.22, 44.18) --
-	(709.49, 44.18) --
-	(709.76, 44.18) --
+\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.20] (571.29, 44.21) --
+	(571.56, 44.21) --
+	(571.83, 44.21) --
+	(572.10, 44.21) --
+	(572.37, 44.21) --
+	(572.65, 44.21) --
+	(572.92, 44.21) --
+	(573.19, 44.21) --
+	(573.46, 44.21) --
+	(573.73, 44.21) --
+	(574.00, 44.21) --
+	(574.27, 44.21) --
+	(574.54, 44.21) --
+	(574.81, 44.21) --
+	(575.08, 44.21) --
+	(575.36, 44.21) --
+	(575.63, 44.21) --
+	(575.90, 44.21) --
+	(576.17, 44.21) --
+	(576.44, 44.21) --
+	(576.71, 44.21) --
+	(576.98, 44.21) --
+	(577.25, 44.21) --
+	(577.52, 44.21) --
+	(577.79, 44.21) --
+	(578.07, 44.21) --
+	(578.34, 44.21) --
+	(578.61, 44.21) --
+	(578.88, 44.21) --
+	(579.15, 44.21) --
+	(579.42, 44.21) --
+	(579.69, 44.21) --
+	(579.96, 44.21) --
+	(580.23, 44.21) --
+	(580.50, 44.21) --
+	(580.77, 44.21) --
+	(581.05, 44.21) --
+	(581.32, 44.21) --
+	(581.59, 44.21) --
+	(581.86, 44.21) --
+	(582.13, 44.21) --
+	(582.40, 44.21) --
+	(582.67, 44.21) --
+	(582.94, 44.21) --
+	(583.21, 44.21) --
+	(583.48, 44.21) --
+	(583.76, 44.21) --
+	(584.03, 44.21) --
+	(584.30, 44.21) --
+	(584.57, 44.21) --
+	(584.84, 44.21) --
+	(585.11, 44.21) --
+	(585.38, 44.21) --
+	(585.65, 44.21) --
+	(585.92, 44.21) --
+	(586.19, 44.21) --
+	(586.47, 44.21) --
+	(586.74, 44.21) --
+	(587.01, 44.21) --
+	(587.28, 44.21) --
+	(587.55, 44.21) --
+	(587.82, 44.21) --
+	(588.09, 44.21) --
+	(588.36, 44.21) --
+	(588.63, 44.21) --
+	(588.90, 44.21) --
+	(589.18, 44.21) --
+	(589.45, 44.21) --
+	(589.72, 44.21) --
+	(589.99, 44.21) --
+	(590.26, 44.21) --
+	(590.53, 44.21) --
+	(590.80, 44.21) --
+	(591.07, 44.21) --
+	(591.34, 44.21) --
+	(591.61, 44.21) --
+	(591.89, 44.21) --
+	(592.16, 44.21) --
+	(592.43, 44.21) --
+	(592.70, 44.21) --
+	(592.97, 44.21) --
+	(593.24, 44.21) --
+	(593.51, 44.21) --
+	(593.78, 44.21) --
+	(594.05, 44.21) --
+	(594.32, 44.21) --
+	(594.59, 44.21) --
+	(594.87, 44.21) --
+	(595.14, 44.21) --
+	(595.41, 44.21) --
+	(595.68, 44.21) --
+	(595.95, 44.21) --
+	(596.22, 44.21) --
+	(596.49, 44.21) --
+	(596.76, 44.21) --
+	(597.03, 44.21) --
+	(597.30, 44.21) --
+	(597.58, 44.21) --
+	(597.85, 44.21) --
+	(598.12, 44.21) --
+	(598.39, 44.21) --
+	(598.66, 44.21) --
+	(598.93, 44.21) --
+	(599.20, 44.21) --
+	(599.47, 44.21) --
+	(599.74, 44.21) --
+	(600.01, 44.21) --
+	(600.29, 44.21) --
+	(600.56, 44.21) --
+	(600.83, 44.21) --
+	(601.10, 44.21) --
+	(601.37, 44.21) --
+	(601.64, 44.21) --
+	(601.91, 44.21) --
+	(602.18, 44.21) --
+	(602.45, 44.21) --
+	(602.72, 44.21) --
+	(603.00, 44.21) --
+	(603.27, 44.21) --
+	(603.54, 44.21) --
+	(603.81, 44.21) --
+	(604.08, 44.21) --
+	(604.35, 44.21) --
+	(604.62, 44.21) --
+	(604.89, 44.21) --
+	(605.16, 44.21) --
+	(605.43, 44.21) --
+	(605.70, 44.21) --
+	(605.98, 44.21) --
+	(606.25, 44.21) --
+	(606.52, 44.21) --
+	(606.79, 44.21) --
+	(607.06, 44.21) --
+	(607.33, 44.21) --
+	(607.60, 44.21) --
+	(607.87, 44.21) --
+	(608.14, 44.21) --
+	(608.41, 44.21) --
+	(608.69, 44.21) --
+	(608.96, 44.21) --
+	(609.23, 44.21) --
+	(609.50, 44.21) --
+	(609.77, 44.21) --
+	(610.04, 44.21) --
+	(610.31, 44.21) --
+	(610.58, 44.21) --
+	(610.85, 44.21) --
+	(611.12, 44.21) --
+	(611.40, 44.21) --
+	(611.67, 44.21) --
+	(611.94, 44.21) --
+	(612.21, 44.21) --
+	(612.48, 44.21) --
+	(612.75, 44.21) --
+	(613.02, 44.21) --
+	(613.29, 44.21) --
+	(613.56, 44.21) --
+	(613.83, 44.21) --
+	(614.11, 44.21) --
+	(614.38, 44.21) --
+	(614.65, 44.21) --
+	(614.92, 44.21) --
+	(615.19, 44.21) --
+	(615.46, 44.21) --
+	(615.73, 44.21) --
+	(616.00, 44.21) --
+	(616.27, 44.21) --
+	(616.54, 44.21) --
+	(616.81, 44.21) --
+	(617.09, 44.21) --
+	(617.36, 44.21) --
+	(617.63, 44.21) --
+	(617.90, 44.21) --
+	(618.17, 44.21) --
+	(618.44, 44.21) --
+	(618.71, 44.21) --
+	(618.98, 44.21) --
+	(619.25, 44.21) --
+	(619.52, 44.21) --
+	(619.80, 44.21) --
+	(620.07, 44.21) --
+	(620.34, 44.21) --
+	(620.61, 44.21) --
+	(620.88, 44.21) --
+	(621.15, 44.21) --
+	(621.42, 44.21) --
+	(621.69, 44.21) --
+	(621.96, 44.21) --
+	(622.23, 44.21) --
+	(622.51, 44.21) --
+	(622.78, 44.21) --
+	(623.05, 44.21) --
+	(623.32, 44.21) --
+	(623.59, 44.21) --
+	(623.86, 44.21) --
+	(624.13, 44.21) --
+	(624.40, 44.21) --
+	(624.67, 44.21) --
+	(624.94, 44.21) --
+	(625.22, 44.21) --
+	(625.49, 44.21) --
+	(625.76, 44.21) --
+	(626.03, 44.21) --
+	(626.30, 44.21) --
+	(626.57, 44.21) --
+	(626.84, 44.21) --
+	(627.11, 44.21) --
+	(627.38, 44.21) --
+	(627.65, 44.21) --
+	(627.92, 44.21) --
+	(628.20, 44.21) --
+	(628.47, 44.21) --
+	(628.74, 44.21) --
+	(629.01, 44.21) --
+	(629.28, 44.21) --
+	(629.55, 44.21) --
+	(629.82, 44.21) --
+	(630.09, 44.21) --
+	(630.36, 44.21) --
+	(630.63, 44.21) --
+	(630.91, 44.21) --
+	(631.18, 44.21) --
+	(631.45, 44.21) --
+	(631.72, 44.21) --
+	(631.99, 44.21) --
+	(632.26, 44.21) --
+	(632.53, 44.21) --
+	(632.80, 44.21) --
+	(633.07, 44.21) --
+	(633.34, 44.21) --
+	(633.62, 44.21) --
+	(633.89, 44.21) --
+	(634.16, 44.21) --
+	(634.43, 44.21) --
+	(634.70, 44.21) --
+	(634.97, 44.21) --
+	(635.24, 44.21) --
+	(635.51, 44.21) --
+	(635.78, 44.21) --
+	(636.05, 44.21) --
+	(636.33, 44.21) --
+	(636.60, 44.21) --
+	(636.87, 44.21) --
+	(637.14, 44.21) --
+	(637.41, 44.21) --
+	(637.68, 44.21) --
+	(637.95, 44.21) --
+	(638.22, 44.21) --
+	(638.49, 44.21) --
+	(638.76, 44.21) --
+	(639.03, 44.21) --
+	(639.31, 44.21) --
+	(639.58, 44.21) --
+	(639.85, 44.21) --
+	(640.12, 44.21) --
+	(640.39, 44.21) --
+	(640.66, 44.21) --
+	(640.93, 44.21) --
+	(641.20, 44.21) --
+	(641.47, 44.21) --
+	(641.74, 44.21) --
+	(642.02, 44.21) --
+	(642.29, 44.21) --
+	(642.56, 44.21) --
+	(642.83, 44.21) --
+	(643.10, 44.21) --
+	(643.37, 44.21) --
+	(643.64, 44.21) --
+	(643.91, 44.21) --
+	(644.18, 44.21) --
+	(644.45, 44.21) --
+	(644.73, 44.21) --
+	(645.00, 44.21) --
+	(645.27, 44.21) --
+	(645.54, 44.21) --
+	(645.81, 44.21) --
+	(646.08, 44.21) --
+	(646.35, 44.21) --
+	(646.62, 44.21) --
+	(646.89, 44.21) --
+	(647.16, 44.21) --
+	(647.44, 44.21) --
+	(647.71, 44.21) --
+	(647.98, 44.21) --
+	(648.25, 44.21) --
+	(648.52, 44.21) --
+	(648.79, 44.21) --
+	(649.06, 44.21) --
+	(649.33, 44.21) --
+	(649.60, 44.21) --
+	(649.87, 44.21) --
+	(650.15, 44.21) --
+	(650.42, 44.21) --
+	(650.69, 44.21) --
+	(650.96, 44.21) --
+	(651.23, 44.21) --
+	(651.50, 44.21) --
+	(651.77, 44.21) --
+	(652.04, 44.21) --
+	(652.31, 44.21) --
+	(652.58, 44.21) --
+	(652.85, 44.21) --
+	(653.13, 44.21) --
+	(653.40, 44.21) --
+	(653.67, 44.21) --
+	(653.94, 44.21) --
+	(654.21, 44.21) --
+	(654.48, 44.21) --
+	(654.75, 44.21) --
+	(655.02, 44.21) --
+	(655.29, 44.21) --
+	(655.56, 44.21) --
+	(655.84, 44.21) --
+	(656.11, 44.21) --
+	(656.38, 44.21) --
+	(656.65, 44.21) --
+	(656.92, 44.21) --
+	(657.19, 44.21) --
+	(657.46, 44.21) --
+	(657.73, 44.21) --
+	(658.00, 44.21) --
+	(658.27, 44.21) --
+	(658.55, 44.21) --
+	(658.82, 44.21) --
+	(659.09, 44.21) --
+	(659.36, 44.21) --
+	(659.63, 44.21) --
+	(659.90, 44.21) --
+	(660.17, 44.21) --
+	(660.44, 44.21) --
+	(660.71, 44.21) --
+	(660.98, 44.21) --
+	(661.26, 44.21) --
+	(661.53, 44.21) --
+	(661.80, 44.21) --
+	(662.07, 44.21) --
+	(662.34, 44.21) --
+	(662.61, 44.21) --
+	(662.88, 44.21) --
+	(663.15, 44.21) --
+	(663.42, 44.21) --
+	(663.69, 44.21) --
+	(663.96, 44.21) --
+	(664.24, 44.21) --
+	(664.51, 44.21) --
+	(664.78, 44.21) --
+	(665.05, 44.21) --
+	(665.32, 44.21) --
+	(665.59, 44.21) --
+	(665.86, 44.21) --
+	(666.13, 44.21) --
+	(666.40, 44.21) --
+	(666.67, 44.21) --
+	(666.95, 44.21) --
+	(667.22, 44.21) --
+	(667.49, 44.21) --
+	(667.76, 44.21) --
+	(668.03, 44.21) --
+	(668.30, 44.21) --
+	(668.57, 44.21) --
+	(668.84, 44.21) --
+	(669.11, 44.21) --
+	(669.38, 44.21) --
+	(669.66, 44.21) --
+	(669.93, 44.21) --
+	(670.20, 44.21) --
+	(670.47, 44.21) --
+	(670.74, 44.21) --
+	(671.01, 44.21) --
+	(671.28, 44.21) --
+	(671.55, 44.21) --
+	(671.82, 44.21) --
+	(672.09, 44.21) --
+	(672.37, 44.21) --
+	(672.64, 44.21) --
+	(672.91, 44.21) --
+	(673.18, 44.21) --
+	(673.45, 44.21) --
+	(673.72, 44.21) --
+	(673.99, 44.21) --
+	(674.26, 44.21) --
+	(674.53, 44.21) --
+	(674.80, 44.21) --
+	(675.07, 44.21) --
+	(675.35, 44.21) --
+	(675.62, 44.21) --
+	(675.89, 44.21) --
+	(676.16, 44.21) --
+	(676.43, 44.21) --
+	(676.70, 44.21) --
+	(676.97, 44.21) --
+	(677.24, 44.21) --
+	(677.51, 44.21) --
+	(677.78, 44.21) --
+	(678.06, 44.21) --
+	(678.33, 44.21) --
+	(678.60, 44.21) --
+	(678.87, 44.21) --
+	(679.14, 44.21) --
+	(679.41, 44.21) --
+	(679.68, 44.21) --
+	(679.95, 44.21) --
+	(680.22, 44.21) --
+	(680.49, 44.21) --
+	(680.77, 44.21) --
+	(681.04, 44.21) --
+	(681.31, 44.21) --
+	(681.58, 44.21) --
+	(681.85, 44.21) --
+	(682.12, 44.21) --
+	(682.39, 44.21) --
+	(682.66, 44.21) --
+	(682.93, 44.21) --
+	(683.20, 44.21) --
+	(683.48, 44.21) --
+	(683.75, 44.21) --
+	(684.02, 44.21) --
+	(684.29, 44.21) --
+	(684.56, 44.21) --
+	(684.83, 44.21) --
+	(685.10, 44.21) --
+	(685.37, 44.21) --
+	(685.64, 44.21) --
+	(685.91, 44.21) --
+	(686.18, 44.21) --
+	(686.46, 44.21) --
+	(686.73, 44.21) --
+	(687.00, 44.21) --
+	(687.27, 44.21) --
+	(687.54, 44.21) --
+	(687.81, 44.21) --
+	(688.08, 44.21) --
+	(688.35, 44.21) --
+	(688.62, 44.21) --
+	(688.89, 44.21) --
+	(689.17, 44.21) --
+	(689.44, 44.21) --
+	(689.71, 44.21) --
+	(689.98, 44.21) --
+	(690.25, 44.21) --
+	(690.52, 44.21) --
+	(690.79, 44.21) --
+	(691.06, 44.21) --
+	(691.33, 44.21) --
+	(691.60, 44.21) --
+	(691.88, 44.21) --
+	(692.15, 44.21) --
+	(692.42, 44.21) --
+	(692.69, 44.21) --
+	(692.96, 44.21) --
+	(693.23, 44.21) --
+	(693.50, 44.21) --
+	(693.77, 44.21) --
+	(694.04, 44.21) --
+	(694.31, 44.21) --
+	(694.59, 44.21) --
+	(694.86, 44.21) --
+	(695.13, 44.21) --
+	(695.40, 44.21) --
+	(695.67, 44.21) --
+	(695.94, 44.21) --
+	(696.21, 44.21) --
+	(696.48, 44.21) --
+	(696.75, 44.21) --
+	(697.02, 44.21) --
+	(697.29, 44.21) --
+	(697.57, 44.21) --
+	(697.84, 44.21) --
+	(698.11, 44.21) --
+	(698.38, 44.21) --
+	(698.65, 44.21) --
+	(698.92, 44.21) --
+	(699.19, 44.21) --
+	(699.46, 44.21) --
+	(699.73, 44.21) --
+	(700.00, 44.21) --
+	(700.28, 44.21) --
+	(700.55, 44.21) --
+	(700.82, 44.21) --
+	(701.09, 44.21) --
+	(701.36, 44.21) --
+	(701.63, 44.21) --
+	(701.90, 44.21) --
+	(702.17, 44.21) --
+	(702.44, 44.21) --
+	(702.71, 44.21) --
+	(702.99, 44.21) --
+	(703.26, 44.21) --
+	(703.53, 44.21) --
+	(703.80, 44.21) --
+	(704.07, 44.21) --
+	(704.34, 44.21) --
+	(704.61, 44.21) --
+	(704.88, 44.21) --
+	(705.15, 44.21) --
+	(705.42, 44.21) --
+	(705.70, 44.21) --
+	(705.97, 44.21) --
+	(706.24, 44.21) --
+	(706.51, 44.21) --
+	(706.78, 44.21) --
+	(707.05, 44.21) --
+	(707.32, 44.21) --
+	(707.59, 44.21) --
+	(707.86, 44.21) --
+	(708.13, 44.21) --
+	(708.41, 44.21) --
+	(708.68, 44.21) --
+	(708.95, 44.21) --
+	(709.22, 44.21) --
+	(709.49, 44.21) --
+	(709.76, 44.21) --
 	(709.76, 44.24) --
@@
 	(571.29, 44.41) --
-	(571.29, 44.01) --
+	(571.29, 44.21) --
 	cycle;

### 2020-05-07 figure2.tex
Make talk pictures downstream of figure2
done

@@
-% Created by tikzDevice version 0.12.3 on 2020-05-04 03:09:56
+% Created by tikzDevice version 0.12.3 on 2020-05-07 17:13:04
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 86.93) --
-	(570.16, 86.93);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 44.63) --
+	(570.16, 44.63);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,143.33) --
-	(570.16,143.33);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76, 85.31) --
+	(570.16, 85.31);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,199.73) --
-	(570.16,199.73);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,125.98) --
+	(570.16,125.98);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,166.65) --
+	(570.16,166.65);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,207.32) --
+	(570.16,207.32);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (338.76,247.99) --
+	(570.16,247.99);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (365.46,144.28) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (365.46,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (409.96,144.28) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (409.96,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (454.46,144.28) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (454.46,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (498.96,144.28) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (498.96,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (543.46,142.60) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (543.46,125.55) circle (  5.71);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,144.28) --
-	(365.46,144.28);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,126.52) --
+	(365.46,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,144.28) --
-	(365.46,144.28);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,126.52) --
+	(365.46,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,144.28) --
-	(365.46,144.28);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (365.46,126.52) --
+	(365.46,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,185.68) --
-	(409.96,185.68);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,150.41) --
+	(409.96,150.41);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,185.68) --
-	(409.96,116.29);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,150.41) --
+	(409.96,110.38);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,116.29) --
-	(409.96,116.29);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (409.96,110.38) --
+	(409.96,110.38);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,192.88) --
-	(454.46,192.88);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,154.56) --
+	(454.46,154.56);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,192.88) --
-	(454.46,106.80);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,154.56) --
+	(454.46,104.90);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,106.80) --
-	(454.46,106.80);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (454.46,104.90) --
+	(454.46,104.90);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,209.81) --
-	(498.96,209.81);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,164.33) --
+	(498.96,164.33);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,209.81) --
-	(498.96, 58.12);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96,164.33) --
+	(498.96, 76.82);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96, 58.12) --
-	(498.96, 58.12);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (498.96, 76.82) --
+	(498.96, 76.82);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,253.19) --
-	(543.46,253.19);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,189.36) --
+	(543.46,189.36);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,253.19) --
-	(543.46, 51.75);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46,189.36) --
+	(543.46, 73.14);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46, 51.75) --
-	(543.46, 51.75);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (543.46, 73.14) --
+	(543.46, 73.14);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56, 82.52) {2.6};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56, 40.23) {2.0};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,138.92) {3.0};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56, 80.90) {2.5};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,195.32) {3.4};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,121.57) {3.0};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,162.24) {3.5};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,202.91) {4.0};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (331.56,243.58) {4.5};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76, 86.93) --
-	(338.76, 86.93);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76, 44.63) --
+	(338.76, 44.63);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,143.33) --
-	(338.76,143.33);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76, 85.31) --
+	(338.76, 85.31);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,199.73) --
-	(338.76,199.73);
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,125.98) --
+	(338.76,125.98);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,166.65) --
+	(338.76,166.65);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,207.32) --
+	(338.76,207.32);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] (334.76,247.99) --
+	(338.76,247.99);
 \end{scope}

### 2020-05-07 figure2.R

@@ -79,7 +79,7 @@ g2 <- ggplot(dd_combine2) +
   geom_errorbar(aes(type, min=lwr, max=upr), width=0, lwd=2) +
   xlab("Uncertainty type") +
   ggtitle("B. Reduced uncertainty in $r$") +
-  scale_y_continuous("Basic reproductive number", limits=c(2.3, 3.8), expand=c(0, 0)) +
+  scale_y_continuous("Basic reproductive number", limits=c(2, 4.6), expand=c(0, 0)) +
   theme(
     panel.grid.major.x = element_blank(),
     panel.grid.minor.x = element_blank(),

### 2020-05-07 compare_R0.tex

@@
-% Created by tikzDevice version 0.12.3 on 2020-04-27 21:16:13
+% Created by tikzDevice version 0.12.3 on 2020-05-07 17:27:07
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25, 77.19) rectangle ( 69.18, 81.12);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 75.59) rectangle ( 70.79, 82.72);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25, 77.19) -- ( 69.18, 81.12);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 75.59) -- ( 70.79, 82.72);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.25, 81.12) -- ( 69.18, 77.19);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 82.72) -- ( 70.79, 75.59);
 \definecolor{drawColor}{RGB}{145,131,230}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.42, 78.87) -- (148.97, 78.87);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (141.15, 78.87) -- (151.24, 78.87);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.19, 76.10) -- (146.19, 81.65);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.19, 73.83) -- (146.19, 83.92);
 \definecolor{fillColor}{RGB}{0,173,154}
 
-\path[fill=fillColor] (135.92, 58.38) --
-	(139.84, 58.38) --
-	(139.84, 62.30) --
-	(135.92, 62.30) --
+\path[fill=fillColor] (134.31, 56.77) --
+	(141.45, 56.77) --
+	(141.45, 63.91) --
+	(134.31, 63.91) --
 	cycle;
@@
 
-\path[fill=fillColor] (129.57, 75.55) --
-	(132.21, 70.98) --
-	(126.92, 70.98) --
+\path[fill=fillColor] (129.57, 78.05) --
+	(134.37, 69.73) --
+	(124.76, 69.73) --
 	cycle;
@@
 
-\path[fill=fillColor] (121.25, 66.37) circle (  1.96);
+\path[fill=fillColor] (121.25, 66.37) circle (  3.57);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.92, 66.49) -- (215.47, 66.49);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.65, 66.49) -- (217.74, 66.49);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.70, 63.71) -- (212.70, 69.26);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.70, 61.44) -- (212.70, 71.54);
 \definecolor{fillColor}{RGB}{0,173,154}
 
-\path[fill=fillColor] (202.42, 66.60) --
-	(206.35, 66.60) --
-	(206.35, 70.53) --
-	(202.42, 70.53) --
+\path[fill=fillColor] (200.82, 65.00) --
+	(207.95, 65.00) --
+	(207.95, 72.13) --
+	(200.82, 72.13) --
 	cycle;
@@
 
-\path[fill=fillColor] (196.07, 82.29) --
-	(198.71, 77.71) --
-	(193.43, 77.71) --
+\path[fill=fillColor] (196.07, 84.79) --
+	(200.88, 76.46) --
+	(191.27, 76.46) --
 	cycle;
@@
 
-\path[fill=fillColor] (187.76, 67.91) circle (  1.96);
+\path[fill=fillColor] (187.76, 67.91) circle (  3.57);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.43, 66.66) -- (281.98, 66.66);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.16, 66.66) -- (284.25, 66.66);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.20, 63.89) -- (279.20, 69.44);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.20, 61.62) -- (279.20, 71.71);
 \definecolor{fillColor}{RGB}{0,173,154}
 
-\path[fill=fillColor] (268.93, 76.35) --
-	(272.85, 76.35) --
-	(272.85, 80.27) --
-	(268.93, 80.27) --
+\path[fill=fillColor] (267.32, 74.74) --
+	(274.46, 74.74) --
+	(274.46, 81.88) --
+	(267.32, 81.88) --
 	cycle;
@@
 
-\path[fill=fillColor] (262.58, 98.08) --
-	(265.22, 93.51) --
-	(259.93, 93.51) --
+\path[fill=fillColor] (262.58,100.58) --
+	(267.38, 92.26) --
+	(257.77, 92.26) --
 	cycle;
@@
 
-\path[fill=fillColor] (254.26, 77.41) circle (  1.96);
+\path[fill=fillColor] (254.26, 77.41) circle (  3.57);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.93,101.40) -- (348.48,101.40);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.66,101.40) -- (350.76,101.40);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.71, 98.63) -- (345.71,104.18);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.71, 96.36) -- (345.71,106.45);
 \definecolor{fillColor}{RGB}{0,173,154}
 
-\path[fill=fillColor] (335.43,112.03) --
-	(339.36,112.03) --
-	(339.36,115.96) --
-	(335.43,115.96) --
+\path[fill=fillColor] (333.83,110.42) --
+	(340.96,110.42) --
+	(340.96,117.56) --
+	(333.83,117.56) --
 	cycle;
@@
 
-\path[fill=fillColor] (329.08, 75.97) --
-	(331.73, 71.40) --
-	(326.44, 71.40) --
+\path[fill=fillColor] (329.08, 78.47) --
+	(333.89, 70.15) --
+	(324.28, 70.15) --
 	cycle;
@@
 
-\path[fill=fillColor] (320.77,101.64) circle (  1.96);
+\path[fill=fillColor] (320.77,101.64) circle (  3.57);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.44, 55.76) -- (414.99, 55.76);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (407.17, 55.76) -- (417.26, 55.76);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.22, 52.99) -- (412.22, 58.54);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.22, 50.72) -- (412.22, 60.81);
 \definecolor{fillColor}{RGB}{0,173,154}
 
-\path[fill=fillColor] (401.94, 47.95) --
-	(405.86, 47.95) --
-	(405.86, 51.88) --
-	(401.94, 51.88) --
+\path[fill=fillColor] (400.33, 46.35) --
+	(407.47, 46.35) --
+	(407.47, 53.48) --
+	(400.33, 53.48) --
 	cycle;
@@
 
-\path[fill=fillColor] (395.59, 99.37) --
-	(398.23, 94.79) --
-	(392.95, 94.79) --
+\path[fill=fillColor] (395.59,101.86) --
+	(400.39, 93.54) --
+	(390.78, 93.54) --
 	cycle;
@@
 
-\path[fill=fillColor] (387.28, 56.59) circle (  1.96);
+\path[fill=fillColor] (387.28, 56.59) circle (  3.57);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.95,105.92) -- (481.50,105.92);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.67,105.92) -- (483.77,105.92);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.72,103.15) -- (478.72,108.70);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.72,100.87) -- (478.72,110.97);
 \definecolor{fillColor}{RGB}{0,173,154}
 
-\path[fill=fillColor] (468.45,157.79) --
-	(472.37,157.79) --
-	(472.37,161.71) --
-	(468.45,161.71) --
+\path[fill=fillColor] (466.84,156.18) --
+	(473.98,156.18) --
+	(473.98,163.32) --
+	(466.84,163.32) --
 	cycle;
@@
 
-\path[fill=fillColor] (462.09, 93.47) --
-	(464.74, 88.89) --
-	(459.45, 88.89) --
+\path[fill=fillColor] (462.09, 95.96) --
+	(466.90, 87.64) --
+	(457.29, 87.64) --
 	cycle;
@@
 
-\path[fill=fillColor] (453.78,145.76) circle (  1.96);
+\path[fill=fillColor] (453.78,145.76) circle (  3.57);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.45, 55.35) -- (548.00, 55.35);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (540.18, 55.35) -- (550.27, 55.35);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.23, 52.58) -- (545.23, 58.13);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.23, 50.30) -- (545.23, 60.40);
 \definecolor{fillColor}{RGB}{0,173,154}
 
-\path[fill=fillColor] (534.95, 68.19) --
-	(538.88, 68.19) --
-	(538.88, 72.11) --
-	(534.95, 72.11) --
+\path[fill=fillColor] (533.34, 66.58) --
+	(540.48, 66.58) --
+	(540.48, 73.72) --
+	(533.34, 73.72) --
 	cycle;
@@
 
-\path[fill=fillColor] (528.60,108.72) --
-	(531.24,104.14) --
-	(525.96,104.14) --
+\path[fill=fillColor] (528.60,111.22) --
+	(533.41,102.89) --
+	(523.79,102.89) --
 	cycle;
@@
 
-\path[fill=fillColor] (520.29, 66.12) circle (  1.96);
+\path[fill=fillColor] (520.29, 66.12) circle (  3.57);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22,122.16) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,122.16) --
 	( 67.22,122.16);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22,122.16) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,122.16) --
 	( 67.22, 55.23);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 55.23) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22, 55.23) --
 	( 67.22, 55.23);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (146.19,148.57) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19,148.57) --
 	(146.19,148.57);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (146.19,148.57) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19,148.57) --
 	(146.19, 42.52);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (146.19, 42.52) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19, 42.52) --
 	(146.19, 42.52);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (137.88, 83.60) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 83.60) --
 	(137.88, 83.60);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (137.88, 83.60) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 83.60) --
 	(137.88, 38.44);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (137.88, 38.44) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 38.44) --
 	(137.88, 38.44);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (129.57, 93.55) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 93.55) --
 	(129.57, 93.55);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (129.57, 93.55) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 93.55) --
 	(129.57, 57.68);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (129.57, 57.68) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 57.68) --
 	(129.57, 57.68);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (121.25, 92.13) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 92.13) --
 	(121.25, 92.13);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (121.25, 92.13) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 92.13) --
 	(121.25, 40.71);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (121.25, 40.71) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 40.71) --
 	(121.25, 40.71);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (212.70,109.43) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70,109.43) --
 	(212.70,109.43);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (212.70,109.43) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70,109.43) --
 	(212.70, 40.25);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (212.70, 40.25) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70, 40.25) --
 	(212.70, 40.25);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (204.38,108.47) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38,108.47) --
 	(204.38,108.47);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (204.38,108.47) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38,108.47) --
 	(204.38, 40.60);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (204.38, 40.60) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38, 40.60) --
 	(204.38, 40.60);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (196.07,112.61) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07,112.61) --
 	(196.07,112.61);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (196.07,112.61) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07,112.61) --
 	(196.07, 59.21);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (196.07, 59.21) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07, 59.21) --
 	(196.07, 59.21);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (187.76,105.16) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76,105.16) --
 	(187.76,105.16);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (187.76,105.16) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76,105.16) --
 	(187.76, 40.54);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (187.76, 40.54) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76, 40.54) --
 	(187.76, 40.54);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (279.20, 84.95) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 84.95) --
 	(279.20, 84.95);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (279.20, 84.95) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 84.95) --
 	(279.20, 52.73);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (279.20, 52.73) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 52.73) --
 	(279.20, 52.73);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (270.89,103.19) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89,103.19) --
 	(270.89,103.19);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (270.89,103.19) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89,103.19) --
 	(270.89, 59.79);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (270.89, 59.79) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89, 59.79) --
 	(270.89, 59.79);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (262.58,154.24) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58,154.24) --
 	(262.58,154.24);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (262.58,154.24) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58,154.24) --
 	(262.58, 65.38);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (262.58, 65.38) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58, 65.38) --
 	(262.58, 65.38);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (254.26, 97.96) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 97.96) --
 	(254.26, 97.96);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (254.26, 97.96) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 97.96) --
 	(254.26, 60.10);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (254.26, 60.10) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 60.10) --
 	(254.26, 60.10);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (345.71,126.97) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71,126.97) --
 	(345.71,126.97);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (345.71,126.97) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71,126.97) --
 	(345.71, 74.24);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (345.71, 74.24) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71, 74.24) --
 	(345.71, 74.24);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (337.40,131.59) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,131.59) --
 	(337.40,131.59);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (337.40,131.59) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,131.59) --
 	(337.40,100.29);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (337.40,100.29) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,100.29) --
 	(337.40,100.29);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (329.08,101.76) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08,101.76) --
 	(329.08,101.76);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (329.08,101.76) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08,101.76) --
 	(329.08, 55.45);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (329.08, 55.45) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08, 55.45) --
 	(329.08, 55.45);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (320.77,107.13) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77,107.13) --
 	(320.77,107.13);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (320.77,107.13) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77,107.13) --
 	(320.77, 96.34);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (320.77, 96.34) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77, 96.34) --
 	(320.77, 96.34);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (412.22, 98.29) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 98.29) --
 	(412.22, 98.29);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (412.22, 98.29) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 98.29) --
 	(412.22, 29.35);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (412.22, 29.35) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 29.35) --
 	(412.22, 29.35);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (403.90, 91.28) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 91.28) --
 	(403.90, 91.28);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (403.90, 91.28) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 91.28) --
 	(403.90, 28.63);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (403.90, 28.63) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 28.63) --
 	(403.90, 28.63);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (395.59,165.53) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59,165.53) --
 	(395.59,165.53);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (395.59,165.53) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59,165.53) --
 	(395.59, 61.03);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (395.59, 61.03) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59, 61.03) --
 	(395.59, 61.03);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (387.28, 96.20) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 96.20) --
 	(387.28, 96.20);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (387.28, 96.20) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 96.20) --
 	(387.28, 29.36);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (387.28, 29.36) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 29.36) --
 	(387.28, 29.36);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (478.72,148.26) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72,148.26) --
 	(478.72,148.26);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (478.72,148.26) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72,148.26) --
 	(478.72, 74.69);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (478.72, 74.69) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72, 74.69) --
 	(478.72, 74.69);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,223.85) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,223.85) --
 	(470.41,223.85);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,223.85) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,223.85) --
 	(470.41,114.90);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,114.90) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,114.90) --
 	(470.41,114.90);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (462.09,146.48) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09,146.48) --
 	(462.09,146.48);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (462.09,146.48) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09,146.48) --
 	(462.09, 62.15);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (462.09, 62.15) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09, 62.15) --
 	(462.09, 62.15);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (453.78,191.49) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,191.49) --
 	(453.78,191.49);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (453.78,191.49) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,191.49) --
 	(453.78,109.22);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (453.78,109.22) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,109.22) --
 	(453.78,109.22);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (545.23, 67.69) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 67.69) --
 	(545.23, 67.69);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (545.23, 67.69) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 67.69) --
 	(545.23, 46.39);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (545.23, 46.39) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 46.39) --
 	(545.23, 46.39);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (536.91, 79.88) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 79.88) --
 	(536.91, 79.88);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (536.91, 79.88) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 79.88) --
 	(536.91, 63.12);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (536.91, 63.12) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 63.12) --
 	(536.91, 63.12);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (528.60,236.72) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60,236.72) --
 	(528.60,236.72);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (528.60,236.72) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60,236.72) --
 	(528.60, 61.88);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (528.60, 61.88) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60, 61.88) --
 	(528.60, 61.88);
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (520.29, 82.44) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 82.44) --
 	(520.29, 82.44);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (520.29, 82.44) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 82.44) --
 	(520.29, 52.97);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (520.29, 52.97) --
+\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 52.97) --
 	(520.29, 52.97);
@@
 
-\path[fill=fillColor] ( 85.79,270.85) circle (  1.96);
+\path[fill=fillColor] ( 85.79,270.85) circle (  3.57);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.01,270.85) -- ( 91.57,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 80.01,270.85) -- ( 91.57,270.85);
 \end{scope}
@@
 
-\path[fill=fillColor] (127.91,273.90) --
-	(130.56,269.33) --
-	(125.27,269.33) --
+\path[fill=fillColor] (127.91,276.40) --
+	(132.72,268.08) --
+	(123.11,268.08) --
 	cycle;
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (122.13,270.85) -- (133.69,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (122.13,270.85) -- (133.69,270.85);
 \end{scope}
@@
 
-\path[fill=fillColor] (243.34,268.89) --
-	(247.27,268.89) --
-	(247.27,272.82) --
-	(243.34,272.82) --
+\path[fill=fillColor] (241.74,267.28) --
+	(248.87,267.28) --
+	(248.87,274.42) --
+	(241.74,274.42) --
 	cycle;
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (239.52,270.85) -- (251.09,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (239.52,270.85) -- (251.09,270.85);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (363.34,270.85) -- (368.89,270.85);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.07,270.85) -- (371.17,270.85);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.12,268.08) -- (366.12,273.63);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.12,265.81) -- (366.12,275.90);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (360.34,270.85) -- (371.90,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (360.34,270.85) -- (371.90,270.85);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.93,268.89) rectangle (506.86,272.82);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,267.28) rectangle (508.46,274.42);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.93,268.89) -- (506.86,272.82);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,267.28) -- (508.46,274.42);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.93,272.82) -- (506.86,268.89);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,274.42) -- (508.46,267.28);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] (499.11,270.85) -- (510.68,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (499.11,270.85) -- (510.68,270.85);
 \end{scope}

### 2020-05-07 ncov.tex

@@
 	\textbf\newline{
-		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (SARS-CoV-2) outbreak
+		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the SARS-CoV-2 outbreak
 	}
@@
 The modeling approaches and resulting estimates of \Ro during the beginning of the outbreak vary widely, despite relying on similar data sources.
-Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate, the mean generation interval, and the generation-interval dispersion.
 We then apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak.
@@
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
-Ignoring these uncertainties inevitably leads to strong conclusions.
+Ignoring these uncertainties inevitably leads to overly confident conclusions.
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
@@
 \begin{tabular}{l|p{2cm}|p{3.2cm}|p{2.8cm}|p{2.5cm}|p{2.4cm}|p{2.7cm}|p{2cm}}
- & Model & Data (study period) & Data source & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & \\
+ & Model & Data (study period) & Data source & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & Reference \\
 \hline
@@
 
-When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population, via the Euler-Lotka equation \citep{wallinga2007generation}:
+When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{linenomath*}
@@
 In this case, most estimates primarily depend on reported cases from China, particularly from Hubei Province.
-Differences among estimates are likely to driven by differences in estimation methods and underlying assumptions, rather than epidemiological differences.
+Differences among estimates are likely to be driven by differences in estimation methods and underlying assumptions, rather than epidemiological differences.
 The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
@@
 Error bars represent 95\% equi-tailed quantiles calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$.
-Red violin plots represent distributions of 2000 posterior samples.
-Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval dispersion.
+Red density plots represent distributions of 2000 posterior samples.
+Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval assumptions.
 }
@@
 The lack of uncertainties in the generation-interval dispersion is particularly surprising because it determines the shape of the $r$-\Ro relationship (\fref{assumption});
-assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
+assuming fixed parameter values can lead to overly confident conclusions \citep{elderd2006uncertainty}.
 The lack of uncertainties in the generation-interval dispersion further explains the sensitivity of \Ro estimates to the exponential growth rate, particularly in Study 7 (\fref{R0}).
@@
 
-As most studies rely on strong assumptions, the credible intervals associated with the base estimates of \Ro are necessarily narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
+As most studies rely on overly confident assumptions, the credible intervals associated with the base estimates of \Ro are necessarily narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
 While the point estimate of \Rpool is similar to other reported values from this date range, the credible intervals are wider than all but one study.

### 2020-05-07 propagate_pix.tex

@@
-% Created by tikzDevice version 0.12.3 on 2020-05-07 10:03:52
+% Created by tikzDevice version 0.12.3 on 2020-05-07 18:04:06
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,130.29) --
-	(497.89,130.29);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 44.63) --
+	(497.89, 44.63);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,244.51) --
-	(497.89,244.51);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,127.00) --
+	(497.89,127.00);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,358.72) --
-	(497.89,358.72);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,209.36) --
+	(497.89,209.36);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,291.73) --
+	(497.89,291.73);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,374.10) --
+	(497.89,374.10);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68,456.46) --
+	(497.89,456.46);
 \definecolor{drawColor}{RGB}{0,0,0}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,246.43) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,210.47) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.59,246.43) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.59,210.47) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,246.43) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,210.47) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,246.43) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,210.47) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,243.03) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,208.51) circle (  5.71);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,246.43) --
-	(101.40,246.43);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
+	(101.40,210.47);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,246.43) --
-	(101.40,246.43);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
+	(101.40,210.47);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,246.43) --
-	(101.40,246.43);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
+	(101.40,210.47);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,330.28) --
-	(187.59,330.28);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,258.85) --
+	(187.59,258.85);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,330.28) --
-	(187.59,189.75);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,258.85) --
+	(187.59,177.77);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,189.75) --
-	(187.59,189.75);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,177.77) --
+	(187.59,177.77);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,344.85) --
-	(273.79,344.85);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
+	(273.79,267.25);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,344.85) --
-	(273.79,170.54);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
+	(273.79,166.69);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,170.54) --
-	(273.79,170.54);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,166.69) --
+	(273.79,166.69);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,379.13) --
-	(359.98,379.13);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
+	(359.98,287.03);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,379.13) --
-	(359.98, 71.95);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
+	(359.98,109.81);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98, 71.95) --
-	(359.98, 71.95);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,109.81) --
+	(359.98,109.81);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,466.99) --
-	(446.17,466.99);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,337.72) --
+	(446.17,337.72);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,466.99) --
-	(446.17, 59.04);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,337.72) --
+	(446.17,102.36);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17, 59.04) --
-	(446.17, 59.04);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,102.36) --
+	(446.17,102.36);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,125.89) {2.6};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {2.0};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,240.10) {3.0};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,122.59) {2.5};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,354.31) {3.4};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,204.96) {3.0};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,287.32) {3.5};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,369.69) {4.0};
+
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,452.05) {4.5};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,130.29) --
-	( 49.68,130.29);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
+	( 49.68, 44.63);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,244.51) --
-	( 49.68,244.51);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,127.00) --
+	( 49.68,127.00);
 
-\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,358.72) --
-	( 49.68,358.72);
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,209.36) --
+	( 49.68,209.36);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,291.73) --
+	( 49.68,291.73);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,374.10) --
+	( 49.68,374.10);
+
+\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,456.46) --
+	( 49.68,456.46);
 \end{scope}

### 2020-05-07 ncov.tex
One comment and some minor changes.

@@
 
+\jd{This \P\ seems awkward to me. I'd like to shorten it. The stuff about why $\kappa$ can matter seems badly out of place.}
 \fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro. 
 For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26, 2020 (\fref{eff}A).
-Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: two middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, is relatively narrow.
+Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are quite narrow.
 In this case, propagating error from the mean generation interval has negligible effect compared to propagating the uncertainty in $r$.
@@
 \cite{flaxman2020estimating} recently estimated the basic reproductive number for SARS-CoV-2 outbreaks in 11 European countries to be around 3.87 (3.01--4.66), on average.
-While these estimates appear to be broadly consistent with to earlier estimates from China, comparing the exponential growth rate and the underlying generation-interval distributions suggest otherwise.
-They assumed a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
+While these estimates appear to be broadly consistent with earlier estimates from China, comparing the exponential growth rate and the underlying generation-interval distributions suggest otherwise.
+The later paper assumed a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
 based on these values, the exponential growth rate has to be much higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ compared to the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).

### 2020-05-08 ncov.tex
tomorrow

--- 2020-05-07 ncov.tex
@@
 This is consistent with our earlier results (\fref{eff}) that demonstrated the importance of accounting for uncertainty in the exponential growth rate $r$.
-We also note that the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates except for Study 2.
+We also note that the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates except for Study 6.
 

### 2020-05-08 compare_R0.R

@@ -52,15 +52,15 @@ R0all %>%
   ) %>%
   filter(!ww)
 
-R0all %>%
-  filter(type=="all") %>%
+R0_all %>%
   mutate(
     width=upr-lwr
   ) %>%
-  merge(summarize(group_by(Rdata_adj, study), bwidth=upr-lwr)) %>%
+  merge(summarize(group_by(Rdata_adj, study), bwidth=upr-lwr), all=TRUE) %>%
   mutate(
     ww=width>bwidth
-  )
+  ) %>%
+  merge(study_number)
 
 all_col <- "#000000" ## black
 g1 <- ggplot(R0all) +

### 2020-05-09 ncov.tex
bmb suggestions!

@@
 \newcommand*{\ie}{i.e.\@\xspace}
+
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 \section*{Abstract}
-A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat in December 2019. 
+A novel coronavirus (SARS-CoV-2) emerged as a global threat in December 2019. 
 As the epidemic progresses, disease modelers continue to focus on estimating the basic reproductive number \Ro\ --- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
@@
 Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate, the mean generation interval, and the generation-interval dispersion.
-We then apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak.
-We show that many early \Ro estimates are overly confident.
+We apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak, showing that many \Ro estimates are overly confident.
 Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.
@@
 
-As SARS-CoV-2 began to spread in other parts of China, outside Hubei province, as well as in other countries, many analyses of the outbreak were published as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}).
-These analyses particularly focused on estimating the basic reproductive number \Ro\ --- the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition} --- in order to assess the pandemic potential of SARS-CoV-2.
-Such rapid dissemination of the early analyses played an important role in shaping the response to the outbreak \citep{majumder2020early}.
+As SARS-CoV-2 began to spread in parts of China outside Hubei province, as well as in other countries, many analyses of the outbreak were published as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}).
+These analyses focused on estimating the basic reproductive number \Ro\ --- the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition} --- in order to assess the pandemic potential of SARS-CoV-2.
+Rapid dissemination of these early analyses played an important role in shaping the response to the outbreak \citep{majumder2020early}.
 
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary significantly even though most analyses relied on reports of confirmed cases from China, particularly from the Hubei province.
-Comparing a disparate set of estimates of \Ro can be particularly difficult when the estimation methods and their underlying assumptions vary widely.
+However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary considerably even though most analyses relied on similar data --- reports of confirmed cases from China, particularly from the Hubei province.
+Comparing a disparate set of estimates of \Ro can be difficult when the estimation methods and their underlying assumptions vary widely.
 In some cases, similar methods can give different estimates; in other cases, different methods can give similar estimates.
-Understanding the differences (or similarities) between \Ro estimates is critical to controlling an epidemic as \Ro provides information about the level of intervention required to prevent further transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
+Understanding the differences between \Ro estimates is critical to controlling an epidemic as \Ro provides information about the level of intervention required to prevent further transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 
@@
 The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, characterizes the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
-therefore, estimates of \Ro directly depend on their assumptions about the generation-interval distribution and the exponential growth rate.
+therefore, estimates of \Ro depend directly on their assumptions about the generation-interval distribution and the exponential growth rate.
 Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
-Ignoring these uncertainties inevitably leads to overly confident conclusions.
+Ignoring these uncertainties leads to overly confident conclusions.
 We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
@@
 We present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
-Previous studies have directly calculated the average of reported \Ro values  \citep{majumder2020early, liu2020reproductive} but such methods can undermine differences in underlying model assumptions and statistical methods.
+Previous studies have directly calculated the average of reported \Ro values  \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
 Instead, we decompose estimated \Ro into three key quantities ($r$, $\bar G$, and $\kappa$) and calculate the average (pooled estimates) of these key parameters.
-Calculating \Ro based on the pooled estimates allow us to appropriately average across the uncertainties present in modeling approaches and their underlying assumptions.
+Calculating \Ro based on the pooled estimates allows us to average appropriately across the uncertainties present in modeling approaches and in their underlying assumptions.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
-% We also use our framework to tease apart which assumptions of these different models led to their different estimates and the associated degrees of uncertainty.
 
@@
 \scriptsize
-\begin{tabular}{l|p{2cm}|p{3.2cm}|p{2.8cm}|p{2.5cm}|p{2.4cm}|p{2.7cm}|p{2cm}}
+\begin{tabular}{l|L{2cm}|L{3.2cm}|L{2.8cm}|L{2.5cm}|L{2.4cm}|L{2.7cm}|L{2cm}}
  & Model & Data (study period) & Data source & Basic reproductive\newline number \Ro & Mean generation\newline interval $\bar G$ (days) & Generation-interval\newline dispersion $\kappa$ & Reference \\
@@
 Model details, estimates of \Ro, and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
-Generation-interval dispersion represent the squared coefficient of variations in generation intervals.
+Generation-interval dispersiosn represent the squared coefficients of variation in generation intervals.
 $^\ast$These intervals reflect \Ro values for best and worst scenarios. We treat these intervals as a 90\% credible interval in our analysis.
@@
 
-We gathered information on estimates of \Ro and their model assumptions from 7 articles that were published online between January 23--26, 2020.
-Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on \url{nextstrain.org} \citep{bedfordncov}.
-Model details and corresponding data are summarized in Table 1.
+We gathered information on estimates of \Ro and their model assumptions from seven articles that were published online between January 23--26, 2020.
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on \url{nextstrain.org} \citep{bedfordncov} (Table 1).
 
@@
 
-Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic grows exponentially in the beginning.
-The IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth when $d \neq 0$ \citep{fisman2013idea}, but Study 7 estimates $d=0$ across all parameters they consider.
-Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
-given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delay of around 5 days \citep{sun2020early}, the majority of reported cases during study periods are likely to have been infected prior to the travel ban.
+Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic initially grows exponentially.
+The IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from exponential growth when $d \neq 0$ \citep{fisman2013idea}, but Study 7 estimates $d=0$ across all parameters they consider.
+Even though some studies consider reported cases up to January 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
+given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delays of around 5 days \citep{sun2020early}, the majority of reported cases during the study periods are likely to have been infected prior to the travel ban.
 
-When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
+When the epidemic is growing exponentially, the estimated basic reproductive number depends entirely on the exponential growth rate $r$ and the intrinsic generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{linenomath*}
@@
 \end{linenomath*}
-In other words, all model assumptions come down to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will be necessarily narrow.
+In other words, all model assumptions reduce to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+If a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
 Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.
 
-As most studies do not report their estimates of the exponential growth rate, we first summarize model outcomes using reported (either estimated or assumed) values of the basic reproductive number \Ro, mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation (Table 1);
+As most studies do not report their estimates of the exponential growth rate, we first summarize model outcomes using reported (either estimated or assumed) values of the basic reproductive number \Ro, mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation (Table 1) ---
 we re-estimate the corresponding exponential growth rates from these values later.
-Study 2 only reports their assumptions about the mean generation interval; for brevity, we assume $\kappa = 0.5$ in our analysis.
+Study 2 only reports their assumptions about the mean generation interval; for simplicity, we assume $\kappa = 0.5$ in our analysis.
 Study 6 presents \Ro estimates under 12 different scenarios regarding reporting rates (0, 0.5, 1 or 2 fold increase in reporting rate) and the shapes of the generation-interval distributions (MERS-like, SARS-like, and average);
 we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
-While estimates of \Ro and the associated confidence intervals for Study 6 in Table 1 represents estimates of \Ro based on $\bar G = 8\,\mathrm{days}$, we account for the uncertainty they consider for $\bar G$ in our formal analysis.
+While estimates of \Ro and the associated confidence intervals for Study 6 in Table 1 are based on $\bar G = 8\,\mathrm{days}$, we account for the uncertainty they consider for $\bar G$ in our formal analysis.
 
-While most studies report confidence or credible intervals to quantify uncertainties associated with their estimates, some rely on different measures.
-In particular, study 2 reports a range of \Ro for the worst and best case scenarios, which correspond to the values of \Ro such that 95\% and 5\% of the simulated total number of cases by January 18, 2020 is greater than or equal to 4000, respectively;
-for brevity, we treat these intervals as a 90\% credible interval in our analysis.
+While most studies report confidence or credible intervals to quantify uncertainties associated with their estimates, some use different measures.
+In particular, Study 2 reports a range of \Ro for the worst and best case scenarios, which correspond to the values of \Ro such that 95\% and 5\% of the simulated total number of cases by January 18, 2020 are greater than or equal to 4000, respectively;
+for simplicity, we treat these intervals as a 90\% credible interval in our analysis.
 Uncertainty ranges reported by Study 1 and Study 7 are assumed to be uniform ranges.
 
-Some studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
-As our primary focus is on the resolution of uncertainty in the available information during the earliest stages of an epidemic, rather than to provide more precise or accurate estimate of \Ro, we strictly focus on estimates that were published between January 23--26, 2020.
+Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+As our primary focus is on the resolution of uncertainty in the available information during the earliest stages of an epidemic, rather than to provide more precise or accurate estimates of \Ro, we focus strictly on estimates that were published between January 23--26, 2020.
 
@@
 Here, we use the gamma approximation framework to the generation-interval distribution \citep{nishiura2009transmission, mcbryde2009early, roberts2011early, trichereau2012estimation, nishiura2015theoretical, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
-The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.
-Study 1, 5, 6, and 7 also used a gamma distribution (including delta and exponential distributions) to model the generation-interval distribution for SARS-CoV-2.
-Assuming that generation intervals follow a gamma distribution with mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation of a gamma distribution, we have
+The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, measles, and rabies \citep{park2019practical}.
+Studies 1, 5, 6, and 7 also used a gamma distribution (including the special cases of Dirac delta and exponential distributions) to model the generation-interval distribution for SARS-CoV-2.
+Assuming that generation intervals follow a gamma distribution with mean generation interval $\bar G$ and generation-interval dispersion $\kappa$, represented by the squared coefficient of variation of a gamma distribution, we have \citep{park2019practical}:
 \begin{linenomath*}
@@
 
-As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that correspond to their model assumptions.
-Since the estimates of the basic reproductive number \Ro entirely depends on the exponential growth rate and the shape of generation-interval distributions, we can calculate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that corresponds to their model assumptions.
+Since the estimates of the basic reproductive number \Ro depend entirely on the exponential growth rate and the shape of generation-interval distributions, we can calculate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
@@
 
-For each study $i$, we construct a family of parameter sets by drawing $10^5$ random samples from the corresponding probability distributions (Table 2) that represent the estimates of $(\Ro)_{i,m}$ and the assumed values of $\bar G_{i,m}$ and $\kappa_{i,m}$ and calculate the exponential growth rate $r_{i,m}$ via the inverse of \eref{gamma}:
+For each study $i$, we construct a family of parameter sets by drawing $10^5$ random samples from the corresponding probability distributions (Table 2) that represent the estimates of $(\Ro)_{i,m}$ and the assumed values of $\bar G_{i,m}$ and $\kappa_{i,m}$ and calculate the exponential growth rate $r_{i,m}$ by inverting \eref{gamma}:
 \begin{linenomath*}
@@
 where $m=1,\dots,10^5$.
-This allows us to approximate the probability distributions of the estimated exponential growth rates by each study.
-We note that uncertainties in the probability distributions that we calculate for the estimated exponential growth rates reflects model assumptions, statistical methods, as well as quality of the data that each study relies on.
-This assumption does not affect our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
-in other words, we still obtain the same estimates and the associated uncertainty of $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.
-For study 6, we fix $\bar G=8\,\textrm{days}$ and use the gamma distribution (Table 2) that corresponds to $\mathcal R_0 = 5.47$ (95\% CI: 4.16--7.10) during the re-estimation step for $r$ to remain consistent with the original study that assumed $\bar G=8\,\textrm{days}$ to obtain this particular estimate.
+This allows us to approximate the probability distributions of the exponential growth rates estimated by each study.
+Uncertainties in the probability distributions that we calculate for the estimated exponential growth rates reflect model assumptions, statistical methods, as well as quality of the data that each study relies on.
+This approach of re-estimating the exponential growth rate does not affect our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
+in other words, we still obtain the same estimate and the same associated uncertainty of $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.
+For study 6, we fix $\bar G=8\,\textrm{days}$ and use the gamma distribution (Table 2) that corresponds to $\mathcal R_0 = 5.47$ (95\% CI: 4.16--7.10) during the re-estimation step for $r$ to remain consistent with the original study, which assumed $\bar G=8\,\textrm{days}$ for this particular estimate.
 We account for uncertainties in $\bar G$ for Study 6 (Table 1) in all other steps in order to properly incorporate parameter uncertainties in the estimate of \Ro.
 Study 7 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
-Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$ in the original article.
+Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which accounts all uncertainty in the reported \Ro when combined with the considered range of $\bar G$ in the original article.
 
@@
 We still account for this uncertainty in our pooled estimates ($\mu_G$).
-$^\dagger$Instead of modeling \Ro with a probability distribution and re-estimating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
+$^\dagger$Instead of modeling \Ro with a probability distribution and re-estimating $r$, we use $r=0.114\,\mathrm{days}^{-1}$ (see text).
 }
@@
 
-We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters estimates across different studies come from the same gamma distributions:
+We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameter estimates across different studies are all drawn from the same gamma distributions:
 \begin{linenomath*}
@@
 where $\mu_r$, $\mu_G$, $\mu_\kappa$ represent the pooled estimates, and $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations.
-The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make.
+The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches while accounting for the uncertainties in their assumptions.
 Here, we do so by averaging across reported values, without explicitly re-fitting their models.
-We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
-Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
+We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
+Since the gamma distribution does not allow $\kappa=0$ (this corresponds to a Dirac delta generation-interval distribution), we substitute $\kappa=0.02$ for Study 7.
 
-We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
+Our approach does not account for non-independence between the parameter estimates made by different modelers.
 In this case, most estimates primarily depend on reported cases from China, particularly from Hubei Province.
-Differences among estimates are likely to be driven by differences in estimation methods and underlying assumptions, rather than epidemiological differences.
-The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
+Differences among estimates are primarily driven by differences in estimation methods and underlying assumptions, rather than by epidemiological differences.
+The pooled estimates can become sharper (i.e., have narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
 Since SARS-CoV-2 had primarily spread in Hubei Province, China during this period, it was not possible to include independent sources of data from other countries.
@@
 \end{linenomath*}
-These priors are chosen such that their 95\% quantiles are sufficiently wider than biologically realistic parameter ranges.
-Specifically, 95\% quantiles for $\mu_r$, $\mu_G$, and $\mu_\kappa$ are 0.02--0.40 $\mathrm{days}^{-1}$, 0.8--19.5 days, and 0.1--1.4, respectively.
+These priors are chosen such that their 95\% quantile ranges are sufficiently wider than biologically realistic parameter ranges.
+Specifically, 95\% quantile ranges for $\mu_r$, $\mu_G$, and $\mu_\kappa$ are 0.02--0.40 $\mathrm{days}^{-1}$, 0.8--19.5 days, and 0.1--1.4, respectively;
+95\% prior quantile range for \Ro then corresponds to 1.05--12.00.
 Parameters that are outside these ranges are biologically unrealistic for SARS-CoV-2 outbreaks.
@@
 We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
-We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
-This phenomenon is a known problem \citep{gelman2006prior}.
+We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances --- a known problem \citep{gelman2006prior}.
 Alternative choices of prior for the between-study shape parameters are also suboptimal. 
-Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$ assumes \textit{a priori} that between-study variance is large (and therefore do not pool different estimates sufficiently).
-Weak priors (e.g. half-Cauchy(0,5)) can lead to inefficient sampling and poor convergence.
+Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$ assumes \textit{a priori} that between-study variance is large (and therefore does not pool different estimates sufficiently).
+Overly weak priors (e.g. half-Cauchy(0,5)) leads to inefficient sampling and poor convergence.
 
@@
 
-We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
+We ran 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
 Posterior samples are thinned every 1000 steps to remove autocorrelations among posterior samples.
@@
 trace plots and marginal posterior distribution plots are presented in Appendix.
-95\% credible intervals (CI) are calculated by taking 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
+95\% credible intervals (CI) are calculated by computing 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
 
@@
 We refer to these estimates as the base estimates.
-We note that the distribution of the basic reproductive number for each study corresponds to the assumed distributions in Table 2 for all studies except for Study 6.
-The assumed distribution in Study 6 in Table 2 does not account for the uncertainty in the mean generation interval $\bar G$, whereas the base estimates account for this uncertainty.
-Furthermore, since the distribution in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles slightly differ from the reported values.
-We compare the base estimates with pooled estimate the basic reproductive number \Rpool based on the pooled estimates of underlying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
+The distribution of the basic reproductive number for each study corresponds to the assumed distributions in Table 2 for all studies except for Study 6.
+The assumed distribution in Study 6 in Table 2 neglects uncertainty in the mean generation interval $\bar G$, whereas the base estimates account for this uncertainty.
+Furthermore, since the distributions in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles differ slightly from the reported values.
+We compare the base estimates with a pooled estimate the basic reproductive number (\Rpool) based on the pooled estimates of underlying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
 
@@
 We refer to the resulting estimates of \Ro as ``substitute'' estimates.
-For example, $r$-substitute estimate for Study $i$ is computed as:
+For example, the $r$-substitute estimate for Study $i$ is computed as:
 \begin{equation}
@@
 \end{equation}
-where $\kappa_i$ and $\bar{G}_i$ are taken from their corresponding parameter sets and $\mu_r$ is taken from the posterior distribution.
+where $\kappa_i$ and $\bar{G}_i$ are taken from their corresponding parameter sets and $\mu_r$ is drawn from the posterior distribution.
 This procedure allows us to assess the sensitivity of the estimates of \Ro across appropriate ranges of uncertainties.
@@
 $\mu_\kappa = 0.50$ (95\% CI: 0.26--1.10).
-We find that there is a large uncertainty associated with the underlying parameters;
+Despite the large uncertainty associated with the underlying parameters,
 most studies consider narrower ranges of uncertainties in these parameters.
-No studies take into account how generation-interval dispersion affects their estimates of \Ro:
+No studies take into account how uncertainty in the generation-interval dispersion affects their estimates of \Ro:
 all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
@@
 
-\jd{This \P\ seems awkward to me. I'd like to shorten it. The stuff about why $\kappa$ can matter seems badly out of place.}
-\fref{eff} shows how propagating uncertainty in different combinations would affect estimates and CIs for \Ro. 
+\fref{eff} shows how propagating uncertainty in underlying parameters affect estimates and CIs for \Ro. 
 For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26, 2020 (\fref{eff}A).
-Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (which all but one of the studies reviewed did) is absolutely crucial: both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are quite narrow.
-In this case, propagating error from the mean generation interval has negligible effect compared to propagating the uncertainty in $r$.
-Uncertainty in the generation-interval dispersion $\kappa$ also has important effects as it determines the functional form of the relationship between $r$ and \Ro (compare $\mu_G$ with $\mu_\kappa$).
-For example, reducing $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro\ relationship from linear to exponential (see \eref{gamma}), therefore increasing the sensitivity of \Ro estimates to $r$ and $\bar G$.
-We note that our estimate of \Rpool is relatively insensitive to our assumption of $\kappa=0.5$ for Study 2: assuming $\kappa=0.1$ gives $\Rpool = 3.0$ (95\% CI: 2.2--4.7), whereas assuming $\kappa=0.9$ gives $\Rpool = 2.9$ (95\% CI: 2.1--4.4).
+Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (as done by all but one studies reviewed) is absolutely crucial: both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are overly narrow.
+In this case, propagating error from the mean generation interval has a negligible effect compared to propagating the uncertainty in $r$.
+Uncertainty in the generation-interval dispersion $\kappa$ also has important effects (compare $\mu_G$ credible intervals with $\mu_\kappa$ credible intervals in \fref{eff}A).
+However, our estimate of \Rpool is relatively insensitive to our assumption of $\kappa=0.5$ for Study 2: assuming $\kappa=0.1$ gives $\Rpool = 3.0$ (95\% CI: 2.2--4.7), whereas assuming $\kappa=0.9$ gives $\Rpool = 2.9$ (95\% CI: 2.1--4.4).
 
-We further explore how the effects of uncertainties in generation-interval distributions change when there is less uncertainty associated with the exponential growth rate.
-This hypothetical example reflects recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
+We further explore how the effects of uncertainties in generation-interval distributions change when the exponential growth rate is more certain.
+This hypothetical example reflects recent scenarios, as increased data availability allows researchers to estimate $r$ with more certainty.
 To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (\fref{eff}B); 
@@
 Propagating error only from the growth rate gives very narrow credible intervals in this case. 
-Propagating errors from the mean generation interval or generation-interval dispersion gives wider but still narrow credible intervals.
+Propagating errors from the mean generation interval or generation-interval dispersion gives more realistic but still narrow credible intervals.
 
@@
 Finally, \fref{R0} compares the base estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
-All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
-Furthermore, accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
-For example, the $r$-substitute estimate of \Ro for Study 7 is $\Ro = 3.9$ (95\% CI: 2.3--8.8), which is much wider than the uncertainty range they reported (2.0--3.1).
+All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- the cases with more certain substitute estimates are the $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
+Accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
+For example, the $r$-substitute estimate of \Ro for Study 7 is $\Ro = 3.9$ (95\% CI: 2.3--8.8), which is much wider than the uncertainty range reported by the authors (2.0--3.1).
 This is consistent with our earlier results (\fref{eff}) that demonstrated the importance of accounting for uncertainty in the exponential growth rate $r$.
-We also note that the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates except for Study 6.
+We also note that the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates in all cases except for Study 6.
 
@@
 \textbf{Sensitivity of the reported \Ro estimates with respect to our pooled estimates of the underlying parameters.}
-We replace the reported parameter values (growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro; we refer to these estimates as substitute estimates.
+We calculate substitute estimates by replacing the reported parameter values (growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$) with our corresponding pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) one at a time and recalculate \Ro.
 The pooled estimate represents \Rpool, which is calculated from the joint posterior distribution of $\mu_r$, $\mu_G$, and $\mu_\kappa$;
@@
 Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the SARS-CoV-2 outbreak.
-Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the degree of dispersion in the generation intervals.
 
 Our analysis shows that many early estimates of \Ro rely on strong assumptions.
-The lack of uncertainties in the generation-interval dispersion is particularly surprising because it determines the shape of the $r$-\Ro relationship (\fref{assumption});
-assuming fixed parameter values can lead to overly confident conclusions \citep{elderd2006uncertainty}.
-The lack of uncertainties in the generation-interval dispersion further explains the sensitivity of \Ro estimates to the exponential growth rate, particularly in Study 7 (\fref{R0}).
+The neglect of uncertainties in the generation-interval dispersion is particularly important because it determines the shape of the $r$--\Ro relationship (\fref{assumption}):
+reducing $\kappa$ from 1 (assuming exponentially distributed generation intervals) to 0 (assuming fixed generation intervals) changes the $r$--\Ro relationship from linear to exponential (see \eref{gamma}).
+Assuming fixed parameter values here will lead to overly confident conclusions \citep{elderd2006uncertainty}.
+The lack of uncertainty in the generation-interval dispersion further explains the sensitivity of \Ro estimates to the exponential growth rate, particularly in Study 7 (\fref{R0}).
 Since Study 7 assumes a fixed generation interval ($\kappa=0$), they implicitly assume an exponential $r$--\Ro relationship, making their estimate too sensitive to $r$.
-One exception is Study 1: we find this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$.
-This is because Study 1 assumes an exponentially distributed generation interval ($\kappa=1$): estimates that rely on this assumption make \Ro relatively insensitive --- implicitly assuming a linear $r$--\Ro relationship --- and thus tend to have particularly narrow credible intervals.
+One exception is Study 1: this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$,
+because the study assumes an exponentially distributed generation interval ($\kappa=1$). 
+Estimates that rely on this assumption make \Ro relatively insensitive --- implicitly assuming a linear $r$--\Ro relationship.
 
@@
 While the point estimate of \Rpool is similar to other reported values from this date range, the credible intervals are wider than all but one study.
-This result does not imply that assumptions based on the pooled estimate are too weak;
-we believe that this credible interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
-In fact, because the pooled estimate does not account for overlap in data sources used by the models, we feel that it is more likely to be over-confident than under-confident.
-Our median estimate averages over the various studies, and therefore particular studies have higher or lower median estimates.
-We note in particular that, while the baseline example we used from Study 6 may appear to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
-Here, our focus is on estimating uncertainty, not on identifying potential explanations for these discrepancies.
+This result does not mean that assumptions underlying the pooled estimate are too weak;
+rather, this credible interval more accurately reflects the level of uncertainties present in the information that was available when these models were fitted.
+In fact, because the pooled estimate does not account for overlap in data sources used by the models, it is more likely to be over-confident than under-confident.
+Because our median estimate averages over the various studies, particular studies have higher or lower median estimates.
+In particular, while the baseline example we used from Study 6 may appear to be an outlier, the authors of this study also explore different scenarios involving changes in reporting rate over time, under which their estimates of \Ro are similar to other reported estimates.
 
-Of the seven studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
-Narrow uncertainties in the estimates of the exponential growth rate are likely to depend on this assumption.
-Naive fits to cumulative incidence data should therefore be avoided.
+Of the seven studies that we reviewed, two of them directly fit their models to the cumulative number of confirmed cases.
+This approach is appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence neglects autocorrelation between successive counts of cumulative cases. 
+As a result, this approach both biases parameter estimates and gives verly narrow confidence intervals \citep{ma2014estimating, king2015avoidable}.
+Narrow uncertainties in the estimates of the exponential growth rate are probably driven by this approach.
 
-Many sources of noise affect real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects disease transmission); and observation noise (randomness underlying how many of the true cases are reported).
+Many sources of noise affect real-world incidence data, including both dynamical, or ``process'', noise (randomness that directly or indirectly affects the actual number of cases occurring); and observation noise (randomness underlying how many of these cases are reported).
 Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
-This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \citep{ma2014estimating}.
-Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \citep{king2015avoidable,taylor2016stochasticity}.
+Neglecting one or the other is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \citep{ma2014estimating}.
+Modelers should, however, be aware that oversimplifying the error model can give overly narrow confidence intervals \citep{king2015avoidable,taylor2016stochasticity}.
 
-There are other important phenomena not covered by our simple framework.
-Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
+Our simple framework neglects some other important phenomena.
+Examples that seem relevant to this outbreak include: changing reporting rates; reporting delays (including the effects of weekends and holidays); and changing generation intervals.
 For emerging pathogens such as SARS-CoV-2, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
@@
 Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated credible intervals.
-Other factors related to reporting --- including changes in case definition, saturation in diagnostic test capacity, transparency of data, and representativeness of samples --- will also affect the inference.
-Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects};
-since we are primarily focusing on the outbreak in Wuhan city before confinement, generation intervals are unlikely to vary significantly.
+Other factors related to reporting --- including changes in case definition, saturation in diagnostic test capacity, transparency of data, and representativeness of samples --- will also affect estimation and inference.
+Finally, generation intervals can become shorter throughout an epidemic as intervention strategies such as isolation of detected cases can reduce the infectious period \citep{hethcote2002effects};
+since we are primarily focusing on the outbreak in Hubei Province before confinement, generation intervals are unlikely to vary significantly.
 All of these factors, including fitting to cumulative curves or ignoring process errors, affect the estimation of the exponential growth rate (as well as the associated uncertainties), which in turn affects the estimation of the basic reproductive number.
@@
 Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates have not been finalized.
-For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020: $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
+For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020 to $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
 we did not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
-Some studies also do not provide detailed description of their methods, data, and/or assumptions.
-The quality of these analyses adds further uncertainty to their results that are not captured by their uncertainty quantification (e.g., reported credible intervals) or by our analysis.
-While pre-prints can play an important role in disseminating information, they should be interpreted with care.
+Some studies also lack detailed description of their methods, data, and/or assumptions.
+The variation in quality of these analyses adds further uncertainty to their results that is not captured by their uncertainty quantification (e.g., reported credible intervals) or by our analysis.
+While pre-prints can play an important role in disseminating information, they should always be interpreted with care.
 
 During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
-However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
-Although our analysis only reflects a snapshot of a fast-moving epidemic, we expect certain lessons to hold: credible intervals must combine different sources of uncertainty. 
-In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate (and other epidemiological parameters) will become more precise; thus the risk of over-confidence when uncertainty about the generation-interval distribution is neglected will become greater.
+However, as the number of susceptible individuals decreases or behaviour changes in response to perception of the epidemic, the growth rate will decrease: estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
+Although our analysis strictly to the earliest stages of an epidemic, we expect certain lessons to hold more generally: credible intervals must combine as many sources of uncertainty as possible. 
+In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate (and other epidemiological parameters) will generally become more precise; thus the risk of over-confidence (when uncertainty about the generation-interval distribution is neglected) will become greater.
 
@@
 Model estimates during an outbreak, even in pre-prints, should include code links and complete explanations.
-We suggest using methods based on open-source tools allow for maximal reproducibility.
+Methods based on open-source tools allow for maximal reproducibility.
 
@@
 The later paper assumed a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
-based on these values, the exponential growth rate has to be much higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ compared to the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).
-We caution that naively comparing estimates of the basic reproductive number without accounting for differences in underlying assumptions can underestimate the differences in the estimates.
+based on these values, the exponential growth rate has to be considerably higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ than the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).
+Naively comparing estimates of the basic reproductive number without accounting for differences in underlying assumptions can underestimate the differences in the estimates.
 
-In summary, we have provided a basis for comparing exponential-growth based estimates of \Ro and its associated uncertainty in terms of three components: the exponential growth rate, mean generation interval, and generation interval dispersion.
+We have provided a basis for comparing exponential-growth based estimates of \Ro and its associated uncertainty in terms of three components: the exponential growth rate, mean generation interval, and generation interval dispersion.
 We hope this framework will help researchers understand and reconcile disparate estimates of disease transmission early in an epidemic.

### 2020-05-09 ncov.tex
tiny corrections

@@
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary considerably even though most analyses relied on similar data --- reports of confirmed cases from China, particularly from the Hubei province.
+However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary considerably even though most analyses relied on similar data --- reports of confirmed cases from China, particularly from the Wuhan City.
 Comparing a disparate set of estimates of \Ro can be difficult when the estimation methods and their underlying assumptions vary widely.
@@
 
-We present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+We present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak in Wuhan City, China \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 Previous studies have directly calculated the average of reported \Ro values  \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
@@
 
-We gathered information on estimates of \Ro and their model assumptions from seven articles that were published online between January 23--26, 2020.
+We gathered information on estimates of \Ro for the SARS-CoV-2 outbreak in Wuhan City, China and their model assumptions from seven articles that were published online between January 23--26, 2020.
 Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the website of Imperial College London \citep{imaincov}; and one report was posted on \url{nextstrain.org} \citep{bedfordncov} (Table 1).
@@
 Study 2 only reports their assumptions about the mean generation interval; for simplicity, we assume $\kappa = 0.5$ in our analysis.
-Study 6 presents \Ro estimates under 12 different scenarios regarding reporting rates (0, 0.5, 1 or 2 fold increase in reporting rate) and the shapes of the generation-interval distributions (MERS-like, SARS-like, and average);
+Study 6 presents \Ro estimates under 12 different scenarios regarding reporting rates (0, 0.5, 1 or 2 fold increase in reporting rate) and the shapes of the generation-interval distributions based on previous coronavirus outbreaks (Middle East Respiratory Syndrome, MERS; Severe Acute Respiratory Syndrome, SARS; and their average) ---
 we use their baseline scenario in our analysis to remain consistent with other studies, which do not account for changes in the reporting rate.
@@
 Our approach does not account for non-independence between the parameter estimates made by different modelers.
-In this case, most estimates primarily depend on reported cases from China, particularly from Hubei Province.
+In this case, most estimates primarily depend on reported cases from China, particularly from Wuhan City.
 Differences among estimates are primarily driven by differences in estimation methods and underlying assumptions, rather than by epidemiological differences.
 The pooled estimates can become sharper (i.e., have narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
-Since SARS-CoV-2 had primarily spread in Hubei Province, China during this period, it was not possible to include independent sources of data from other countries.
+Since SARS-CoV-2 had primarily spread in Wuhan City, China during this period, it was not possible to include independent sources of data from other countries.
 Thus, the pooled estimator should be interpreted with care.
@@
 Finally, generation intervals can become shorter throughout an epidemic as intervention strategies such as isolation of detected cases can reduce the infectious period \citep{hethcote2002effects};
-since we are primarily focusing on the outbreak in Hubei Province before confinement, generation intervals are unlikely to vary significantly.
+since we are primarily focusing on the outbreak in Wuhan City before confinement, generation intervals are unlikely to vary significantly.
 All of these factors, including fitting to cumulative curves or ignoring process errors, affect the estimation of the exponential growth rate (as well as the associated uncertainties), which in turn affects the estimation of the basic reproductive number.

### 2020-05-09 ncov.tex
super minor tweaks

@@
 therefore, estimates of \Ro depend directly on their assumptions about the generation-interval distribution and the exponential growth rate.
-Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
+Early in an epidemic, information is scarce and there is uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
 Ignoring these uncertainties leads to overly confident conclusions.
-We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
+% We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
 
-We present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies and apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak in Wuhan City, China \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+To formalize the estimation of uncertainty at the onset of an outbreak, we present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies. 
+We apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak in Wuhan City, China \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 Previous studies have directly calculated the average of reported \Ro values  \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
@@
 Model details, estimates of \Ro, and their assumptions about the shape of the generation interval distributions were collected from 7 studies.
-Generation-interval dispersiosn represent the squared coefficients of variation in generation intervals.
+Generation-interval dispersion represent the squared coefficients of variation in generation intervals.
 $^\ast$These intervals reflect \Ro values for best and worst scenarios. We treat these intervals as a 90\% credible interval in our analysis.
@@
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
-We used gamma distributions to model values reported with confidence or credible intervals and uniform distributions to model values reported with ranges;
+We used gamma distributions to model values reported with confidence or credible intervals (CI) and uniform distributions to model values reported with ranges;
 when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
@@
 Alternative choices of prior for the between-study shape parameters are also suboptimal. 
-Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$ assumes \textit{a priori} that between-study variance is large (and therefore does not pool different estimates sufficiently).
+Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$) assumes \textit{a priori} that between-study variance is large (and therefore does not pool different estimates sufficiently).
 Overly weak priors (e.g. half-Cauchy(0,5)) leads to inefficient sampling and poor convergence.
@@
 For illustrative purposes, we use our pooled estimates, which may represent a reasonable proxy for the state of knowledge as of January 23--26, 2020 (\fref{eff}A).
-Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (as done by all but one studies reviewed) is absolutely crucial: both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are overly narrow.
+Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (as done by all but one studies reviewed) is absolutely crucial: uncertainty in the pooled estimates for both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are overly narrow.
 In this case, propagating error from the mean generation interval has a negligible effect compared to propagating the uncertainty in $r$.
@@
 We further explore how the effects of uncertainties in generation-interval distributions change when the exponential growth rate is more certain.
-This hypothetical example reflects recent scenarios, as increased data availability allows researchers to estimate $r$ with more certainty.
+This hypothetical example reflects scenarios, in which increased data availability allows researchers to estimate $r$ with more certainty.
 To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (\fref{eff}B); 
@@
 
-Our analysis shows that many early estimates of \Ro rely on strong assumptions.
+Our analysis shows that many early estimates of \Ro rely on overly confident assumptions.
 The neglect of uncertainties in the generation-interval dispersion is particularly important because it determines the shape of the $r$--\Ro relationship (\fref{assumption}):
@@
 
-As most studies rely on overly confident assumptions, the credible intervals associated with the base estimates of \Ro are necessarily narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
+As most studies rely on overly confident assumptions, the credible intervals associated with the base estimates of \Ro should tend to be narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
 While the point estimate of \Rpool is similar to other reported values from this date range, the credible intervals are wider than all but one study.
@@
 This approach is appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence neglects autocorrelation between successive counts of cumulative cases. 
-As a result, this approach both biases parameter estimates and gives verly narrow confidence intervals \citep{ma2014estimating, king2015avoidable}.
+As a result, this approach both biases parameter estimates and gives very narrow confidence intervals \citep{ma2014estimating, king2015avoidable}.
 Narrow uncertainties in the estimates of the exponential growth rate are probably driven by this approach.
@@
 Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26, 2020).
-Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates have not been finalized.
+Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates were not necessarily finalized.
 For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020 to $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
@@
 During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
-However, as the number of susceptible individuals decreases or behaviour changes in response to perception of the epidemic, the growth rate will decrease: estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
+However, as the number of susceptible individuals decreases or behavior changes in response to perception of the epidemic, the growth rate will decrease: estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
 Although our analysis strictly to the earliest stages of an epidemic, we expect certain lessons to hold more generally: credible intervals must combine as many sources of uncertainty as possible. 
@@
 
-\cite{flaxman2020estimating} recently estimated the basic reproductive number for SARS-CoV-2 outbreaks in 11 European countries to be around 3.87 (3.01--4.66), on average.
+Despite our focus on estimating \Ro at the onset of an outbreak, many of the issues persist now. 
+For example, \cite{flaxman2020estimating} recently estimated the basic reproductive number for SARS-CoV-2 outbreaks in 11 European countries to be around 3.87 (3.01--4.66), on average.
 While these estimates appear to be broadly consistent with earlier estimates from China, comparing the exponential growth rate and the underlying generation-interval distributions suggest otherwise.
@@
 based on these values, the exponential growth rate has to be considerably higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ than the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).
-Naively comparing estimates of the basic reproductive number without accounting for differences in underlying assumptions can underestimate the differences in the estimates.
+Naively comparing estimates of the basic reproductive number without accounting for differences in underlying assumptions can lead to over-interpretation of apparent differences in the estimates.
 
@@
 
-SWP and JD developed the statistical framework. 
+SWP and JD developed the statistical framework, with contributions from all authors. 
 SWP reviewed the published literature.

### 2020-05-09 response.tex
update

@@
 Thank you for the chance to revise and resubmit our manuscript. 
+We have re-written a considerable amount of the manuscript in order to convince a more general readership (suggested by reviewer 1) and to provide a clearer explanation of the methods (suggested by reviewer 2).
+A large proportion of the Results section has been moved to the Discussion section, as suggested by reviewer 2.
+We have also included a new paragraph in the Discussion section to compare early estimates with more recent estimates of the basic reproduction number.
+The main figures and the overall conclusions remain unchanged.
 Below please find our responses to reviewers.
@@
 
-``Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic grows exponentially in the beginning.
-The IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from the exponential growth when $d \neq 0$ \citep{fisman2013idea}, but Study 7 estimates $d=0$ across all parameters they consider.
-Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
-given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delay of around 5 days \citep{sun2020early}, the majority of reported cases during study periods are likely to have been infected prior to the travel ban.
+``Despite a wide range of models considered across Study 1--7, all of them assume that the epidemic initially grows exponentially.
+The IDEA model (used in Study 7) includes a discount parameter $d$ that allows the model to deviate from exponential growth when $d \neq 0$ \citep{fisman2013idea}, but Study 7 estimates $d=0$ across all parameters they consider.
+Even though some studies consider reported cases up to January 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
+given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delays of around 5 days \citep{sun2020early}, the majority of reported cases during the study periods are likely to have been infected prior to the travel ban.
 
-When the epidemic is growing exponentially, the estimated basic reproductive number entirely depends on the exponential growth rate $r$ and the \emph{intrinsic} generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population, via the Euler-Lotka equation \citep{wallinga2007generation}:
+When the epidemic is growing exponentially, the estimated basic reproductive number depends entirely on the exponential growth rate $r$ and the intrinsic generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{equation}
@@
 \end{equation}
-In other words, all model assumptions come down to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-For example, if a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will be necessarily narrow.
+In other words, all model assumptions reduce to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+If a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
 Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.''
@@
 
-``This assumption does not affect our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
-in other words, we still obtain the same estimates and the associated uncertainty of $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.''
+``This approach of re-estimating the exponential growth rate does not affect the uncertainty captured by our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
+in other words, we still obtain the same degree of associated uncertainty in $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.''
 
@@
 ``Here, we use the gamma approximation framework to the generation-interval distribution \citep{nishiura2009transmission, mcbryde2009early, roberts2011early, trichereau2012estimation, nishiura2015theoretical, park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
-The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, influenza, measles, and rabies.
-Study 1, 5, 6, and 7 also used a gamma distribution (including delta and exponential distributions) to model the generation-interval distribution for SARS-CoV-2.''
+The gamma distribution provides a reasonable approximation for generation-interval distributions of many diseases, including Ebola, measles, and rabies \citep{park2019practical}.
+Studies 1, 5, 6, and 7 also used a gamma distribution (including the special cases of Dirac delta and exponential distributions) to model the generation-interval distribution for SARS-CoV-2.''
 
@@
 
-``We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
-In this case, most estimates primarily depend on reported cases from China, particularly from Hubei Province.
-Differences among estimates are likely to driven by differences in estimation methods and underlying assumptions, rather than epidemiological differences.
-The pooled estimates can become sharper (i.e., give narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
-Since SARS-CoV-2 had primarily spread in Hubei Province, China during this period, it is not possible to include independent sources of data from other countries.
+``Our approach does not account for non-independence between the parameter estimates made by different modelers.
+In this case, most estimates primarily depend on reported cases from China, particularly from Wuhan City.
+Differences among estimates are primarily driven by differences in estimation methods and underlying assumptions, rather than by epidemiological differences.
+The pooled estimates can become sharper (i.e., have narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
+Since SARS-CoV-2 had primarily spread in Wuhan City, China during this period, it was not possible to include independent sources of data from other countries.
 Thus, the pooled estimator should be interpreted with care.''
@@
 \rev{Reviewer \#2}
-
-Our exposition may not have been clear enough for reviewer 2.
-Our method does not directly rely on data, and does not depend on a specific dynamical model, but the reviewer asks us to compare our results with the reported number of cases.
-We use a multi-level model to calculate pooled estimates of parameters that have already been estimated or assumed from other studies, based on fits to case reports.
-Then, we use the pooled estimates to calculate $\mathcal R_0$.
-We have made significant improvement to our manuscript to make our methods clearer.
 
@@
 
-Done.
+We refer to uncertainty intervals that we estimate as credible intervals as suggested.
+We still use confidence intervals to refer to uncertainty intervals reported in the original study if it relies on frequentist inference.
 
@@
 
-We have made significant changes to our Methods section.
+We have tried to clarify assumptions of our Bayesian multilevel model as much as possible. 
+In particular, we made it clear that we do not rely on time series data or fit a model to time series data; 
+we take reported estimates and average them.
+We also explain potential sources of uncertainty in the reported estimates.
 
@@
 
-Done.
+We now explain all methods, including those that were described in figure and table captions, within the Methods section.
+We left simplified versions of the descriptions in captions to assist readers in interpreting the tables and figures.
+We have also added subheadings in the Methods section.
 
@@
 
-``We further explore how the effects of uncertainties in generation-interval distributions change when there is less uncertainty associated with the exponential growth rate.
-This hypothetical example reflects recent scenarios, as increased data availability will allow researchers to estimate $r$ with more certainty.
-To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$; 
-then $\hat{\mu}_r$ has the same median as $\mu_r$ but the associated 95\% CI is 4 times narrower (0.16--0.19 $\textrm{days}^{-1}$).''
+``We further explore how the effects of uncertainties in generation-interval distributions change when the exponential growth rate is more certain.
+This hypothetical example reflects scenarios, in which increased data availability allows researchers to estimate $r$ with more certainty.
+To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (\fref{eff}B); 
+then $\hat{\mu}_r$ has the same median as $\mu_r$ but the associated 95\% CI is 4 times narrower (0.16--0.19 $\textrm{days}^{-1}$).
+As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important.
+Propagating error only from the growth rate gives very narrow credible intervals in this case. 
+Propagating errors from the mean generation interval or generation-interval dispersion gives more realistic but still narrow credible intervals.''
 
@@
 We have made significant revisions to the Methods section. 
-In particular, we made it very clear that we are simply using Bayesian estimation to average across the reported estimates and are not relying on any data or dynamical model.
+In particular, we made it very clear that we are simply using Bayesian estimation to average across the reported estimates and are not relying on any time series data or dynamical model.
 We hope that the new text is sufficiently clear; a directed acyclic graph to explain that we are doing Bayesian averaging seems like overkill.
@@
 
-We do not rely on data or fit a model to data. Therefore, it is not possible to compare with the reported cases.
+We do not rely on time series data or fit a model to time series data. Therefore, it is not possible to compare with the reported cases.
 
@@
 Error bars represent 95\% equi-tailed quantiles calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$.
-Red violin plots represent distributions of 2000 posterior samples.
-Open triangle: we assumed $\kappa=0.5$ for Study 2, which does not report generation-interval assumptions.''
+Red density plots represent distributions of 2000 posterior samples.
+Open triangle: we assumed $\kappa=0.5$ for Study 2 which does not report generation-interval assumptions.''
 
@@
 
-Done.
+Done:
+
+``Figure 1 compares the estimated values of the exponential growth rate $r$, mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ from different studies with the pooled estimates that we calculate from our multilevel model:
+$\mu_r = 0.17\,\textrm{days}^{-1}$ (95\% CI: 0.12--0.25 $\textrm{days}^{-1}$),
+$\mu_G = 8.51\,\textrm{days}$ (95\% CI: 7.60--9.63 days),
+and
+$\mu_\kappa = 0.50$ (95\% CI: 0.26--1.10).''
 
@@
 
-Done.
+Done:
+
+``Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (as done by all but one studies reviewed) is absolutely crucial: uncertainty in the pooled estimates for both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are overly narrow.''
 
@@
 
-Done.
+We now incorporate the uncertainty related to kappa in Figure 2.
 
@@
 ``Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26, 2020).
-Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates have not been finalized.
-For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020: $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
+Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates were not necessarily finalized.
+For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020 to $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
 we did not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
-Some studies also do not provide detailed description of their methods, data, and/or assumptions.
-The quality of these analyses adds further uncertainty to their results that are not captured by their uncertainty quantification (e.g., reported credible intervals) or by our analysis.
-While pre-prints can play an important role in disseminating information, they should be interpreted with care.''
+Some studies also lack detailed description of their methods, data, and/or assumptions.
+The variation in quality of these analyses adds further uncertainty to their results that is not captured by their uncertainty quantification (e.g., reported credible intervals) or by our analysis.''
 
@@
 
-``There are other important phenomena not covered by our simple framework.
-Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
+``Our simple framework neglects some other important phenomena.
+Examples that seem relevant to this outbreak include: changing reporting rates; reporting delays (including the effects of weekends and holidays); and changing generation intervals.
 For emerging pathogens such as SARS-CoV-2, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
@@
 Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated credible intervals.
-Other factors related to reporting --- including changes in case definition, saturation in diagnostic test capacity, transparency of data, and representativeness of samples --- will also affect the inference.
-Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects};
-since we are primarily focusing on the outbreak in Wuhan city before confinement, generation intervals are unlikely to vary significantly.
+Other factors related to reporting --- including changes in case definition, saturation in diagnostic test capacity, transparency of data, and representativeness of samples --- will also affect estimation and inference.
+Finally, generation intervals can become shorter throughout an epidemic as intervention strategies such as isolation of detected cases can reduce the infectious period \citep{hethcote2002effects};
+since we are primarily focusing on the outbreak in Wuhan City before confinement, generation intervals are unlikely to vary significantly.
 All of these factors, including fitting to cumulative curves or ignoring process errors, affect the estimation of the exponential growth rate (as well as the associated uncertainties), which in turn affects the estimation of the basic reproductive number.''
@@
 
-``We used gamma distributions to model values reported with confidence or credible intervals and uniform distributions to model values reported with ranges;
+``We used gamma distributions to model values reported with confidence or credible intervals (CI) and uniform distributions to model values reported with ranges;
 when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
@@
 
-We believe that these priors are sufficiently wide and therefore did not try different priors. We provide a better justification in the main text:
+We believe that these priors are sufficiently wide and therefore did not try different priors.
+We also perform a prior predictive simulations on \Ro based on $\mu_r$, $\mu_G$, and $\mu_\kappa$ to provide further justification.
+We provide a better justification in the main text:
 
-``These priors are chosen such that their 95\% quantiles are sufficiently wider than biologically realistic parameter ranges.
-Specifically, 95\% quantiles for $\mu_r$, $\mu_G$, and $\mu_\kappa$ are 0.02--0.40 $\mathrm{days}^{-1}$, 0.8--19.5 days, and 0.1--1.4, respectively.
+``These priors are chosen such that their 95\% quantile ranges are sufficiently wider than biologically realistic parameter ranges.
+Specifically, 95\% quantile ranges for $\mu_r$, $\mu_G$, and $\mu_\kappa$ are 0.02--0.40 $\mathrm{days}^{-1}$, 0.8--19.5 days, and 0.1--1.4, respectively;
+95\% prior quantile range for \Ro then corresponds to 1.05--12.00.
 Parameters that are outside these ranges are biologically unrealistic for SARS-CoV-2 outbreaks.
@@
 
-``Weak priors (e.g. half-Cauchy(0,5)) can lead to inefficient sampling and poor convergence.''
+``Alternative choices of prior for the between-study shape parameters are also suboptimal. 
+Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$) assumes \textit{a priori} that between-study variance is large (and therefore does not pool different estimates sufficiently).
+Overly weak priors (e.g. half-Cauchy(0,5)) leads to inefficient sampling and poor convergence.''
 
@@
 
-``We note that our estimate of \Rpool is relatively insensitive to our assumption of $\kappa=0.5$ for Study 2: assuming $\kappa=0.1$ gives $\Rpool = 3.0$ (95\% CI: 2.2--4.7), whereas assuming $\kappa=0.9$ gives $\Rpool = 2.9$ (95\% CI: 2.1--4.4).''
+``However, our estimate of \Rpool is relatively insensitive to our assumption of $\kappa=0.5$ for Study 2: assuming $\kappa=0.1$ gives $\Rpool = 3.0$ (95\% CI: 2.2--4.7), whereas assuming $\kappa=0.9$ gives $\Rpool = 2.9$ (95\% CI: 2.1--4.4).''
 
@@
 
-``Even though some studies consider reported cases up to Janaury 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
-given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delay of around 5 days \citep{sun2020early}, the majority of reported cases during study periods are likely to have been infected prior to the travel ban.''
+``Even though some studies consider reported cases up to January 26, 2020 --- three days after the travel restriction that took place on January 23, 2020 \citep{Tianeabb6105} --- the exponential growth assumption can still describe the number of reported cases reasonably well;
+given the incubation period of around 5 days \citep{lauer2020incubation} as well as reporting delays of around 5 days \citep{sun2020early}, the majority of reported cases during the study periods are likely to have been infected prior to the travel ban.''
 
@@
 ``Finally, Figure 3 compares the base estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
-All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- namely, $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.''
+All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- the cases with more certain substitute estimates are the $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.''
 
@@
 
-Thank you for the review. We have made significant revisions to our manuscript to make it approachable for a wider audience.
+As stated in the journal scope, JRSI publishes research applying mathematics and physics to the biological and medical sciences.
+We assume that the ``and-s'' and both sides implies that any combination is sufficiently in scope, i.e., math+bio, math+medical sciences, physics+bio, physics+medical sciences, or a combination.
+This paper applies mathematical methods to pressing epidemiological questions of broad relevance, both within the field of epidemiology and outside given the impacts of $\Ro$ estimation on forecasting and control efforts.
 
@@
 
-%% \bibliography{../ncov_abbr.bib} Sorry!
 \bibliography{ncov_abbr.bib}

### 2020-05-09 ncov.tex
need to sleep
update

@@
 Uncertainties in the probability distributions that we calculate for the estimated exponential growth rates reflect model assumptions, statistical methods, as well as quality of the data that each study relies on.
-This approach of re-estimating the exponential growth rate does not affect our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
-in other words, we still obtain the same estimate and the same associated uncertainty of $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.
+This approach of re-estimating the exponential growth rate does not affect the uncertainty captured by our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
+in other words, we still obtain the same degree of associated uncertainty in $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.
 For study 6, we fix $\bar G=8\,\textrm{days}$ and use the gamma distribution (Table 2) that corresponds to $\mathcal R_0 = 5.47$ (95\% CI: 4.16--7.10) during the re-estimation step for $r$ to remain consistent with the original study, which assumed $\bar G=8\,\textrm{days}$ for this particular estimate.
@@
 The variation in quality of these analyses adds further uncertainty to their results that is not captured by their uncertainty quantification (e.g., reported credible intervals) or by our analysis.
-While pre-prints can play an important role in disseminating information, they should always be interpreted with care.
 
@@
 
-We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical frameworks for inferring generation-interval distributions from such data \citep{britton2019estimation}.
-A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
+We strongly emphasize the value of attention to accurate characterization of the transmission chains via both contact tracing and improved statistical frameworks for inferring generation-interval distributions from such data \citep{britton2019estimation}.
+A combined effort between public-health workers and modelers in this direction will be crucial both for predicting the course of an epidemic and for controlling it.
 We also emphasize the value of transparency from modelers.
 Model estimates during an outbreak, even in pre-prints, should include code links and complete explanations.
-Methods based on open-source tools allow for maximal reproducibility.
+Methods based on open-source tools allow for maximal reproducibility \citep{barton2020call}.
 

### 2020-05-10 response.tex
done?

@@
 A large proportion of the Results section has been moved to the Discussion section, as suggested by reviewer 2.
-We have also included a new paragraph in the Discussion section to compare early estimates with more recent estimates of the basic reproduction number.
-The main figures and the overall conclusions remain unchanged.
+We have also included a new paragraph in the Discussion section to compare early estimates with more recent estimates of the basic reproductive number.
+We had previous interpreted the uncertainty range reported by Study 2 as a 95\% credible interval; we now interpret it as 90\% credible interval as it better suits their definition.
+Otherwise, the main figures and the overall conclusions remain unchanged.
 Below please find our responses to reviewers.
@@
 
-This is a key point. During the exponential growth period, the basic reproduction number entirely depends on the exponential growth rate and the generation-interval distributions.
+This is a key point. During the exponential growth period, the basic reproductive number is determined by the exponential growth rate and the generation-interval distributions.
 Therefore, the effect of model assumptions on the inferred reproductive number can in fact be considered via their effects on exponential growth rate and the shape of the generation-interval distribution.
@@
 
-When the epidemic is growing exponentially, the estimated basic reproductive number depends entirely on the exponential growth rate $r$ and the intrinsic generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
+When the epidemic is growing exponentially, the estimated basic reproductive number is determined by the exponential growth rate $r$ and the intrinsic generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{equation}
@@
 
-All of these issues raised here directly affect estimates of growth rate $r$, which in turn affect estimates of the basic reproduction number \Ro.
+All of these issues raised here directly affect estimates of growth rate $r$, which in turn affect estimates of the basic reproductive number \Ro.
 Therefore, comparing the growth rate with generation-interval distribution parameters allow us to take these assumptions into account implicitly. We have made this clearer in the main text.
@@
 On the other hand, this is definitely sufficient for the effective sample size to reliably calculate the 95\% confidence intervals, especially given that we're only averaging across previously estimated parameter values, rather than explicitly fitting a model to data.
-We now ensure that effective sample sizes are greater than at least 1,000 for all parameters.
+We now ensure that effective sample sizes are greater than at least 1,000 for all hyperparameters.
 
@@
 
-We have sampled appropriately from the Bayesian posteriors. We talk about pooled estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$ because those are used directly to calculate the reproduction number.
+We have sampled appropriately from the Bayesian posteriors. We talk about pooled estimates of $\mu_r$, $\mu_G$, and $\mu_\kappa$ because those are used directly to calculate the reproductive number.
 
@@
 
-We now check the sensitivity of the value of kappa of Study 2 on the pooled estimate of the basic reproduction number.
+We now check the sensitivity of the value of kappa of Study 2 on the pooled estimate of the basic reproductive number.
 

### 2020-05-10 ncov.tex
done?

@@
 
-As SARS-CoV-2 began to spread in parts of China outside Hubei province, as well as in other countries, many analyses of the outbreak were published as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}).
+As SARS-CoV-2 began to spread in parts of China outside Hubei province, as well as in other countries, many analyses of the outbreak were published as pre-prints \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov} and in peer-reviewed journals \citep{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}.
 These analyses focused on estimating the basic reproductive number \Ro\ --- the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition} --- in order to assess the pandemic potential of SARS-CoV-2.
@@
 
-When the epidemic is growing exponentially, the estimated basic reproductive number depends entirely on the exponential growth rate $r$ and the intrinsic generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
+When the epidemic is growing exponentially, the estimated basic reproductive number is determined by the exponential growth rate $r$ and the intrinsic generation-interval distribution $g(\tau)$, which describes the infection time of secondary cases caused by a primary case in a fully susceptible population \citep{champredon2015intrinsic}, via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{linenomath*}
@@
 As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that corresponds to their model assumptions.
-Since the estimates of the basic reproductive number \Ro depend entirely on the exponential growth rate and the shape of generation-interval distributions, we can calculate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+Since the estimates of the basic reproductive number \Ro depend is determined by the exponential growth rate and the shape of generation-interval distributions, we can calculate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
@@
 We account for uncertainties in $\bar G$ for Study 6 (Table 1) in all other steps in order to properly incorporate parameter uncertainties in the estimate of \Ro.
-Study 7 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+Study 7 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which accounts all uncertainty in the reported \Ro when combined with the considered range of $\bar G$ in the original article.
@@
 
-We use weakly informative priors on hyperparameters:
+We use weakly informative priors on hyperparameters ($\mu_r, \mu_G, \mu_\kappa, \sigma_r, \sigma_G, \sigma_\kappa$):
 \begin{linenomath*}
@@
 Posterior samples are thinned every 1000 steps to remove autocorrelations among posterior samples.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 and the effective sample size is greater than 1000 for all hyperparameters \citep{gelman1992inference};
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01  \citep{gelman1992inference} and the effective sample size is greater than 1000 for all hyperparameters ($\mu_r, \mu_G, \mu_\kappa, \sigma_r, \sigma_G, \sigma_\kappa$);
 trace plots and marginal posterior distribution plots are presented in Appendix.
-95\% credible intervals (CI) are calculated by computing 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each parameter.
+95\% credible intervals (CI) are calculated by computing 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each hyperparameter.
 
@@
 Points represent the medians of our base, substitute, and pooled estimates.
-Horizontal dashed lines represent the 95\% credible intervals of our base, substitute, and pooled estimates.
+Vertical error bars represent the 95\% credible intervals of our base, substitute, and pooled estimates.
+Horizontal dashed lines represent the 95\% credible intervals of our pooled estimates.
 }
@@
 
-Of the seven studies that we reviewed, two of them directly fit their models to the cumulative number of confirmed cases.
+Of the seven studies that we reviewed, at least one of them directly fit their models to the cumulative number of confirmed cases.
 This approach is appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence neglects autocorrelation between successive counts of cumulative cases. 

### 2020-05-10 ncov.tex
done?

@@
 
-We ran 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps.
+We ran 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps using Metropolis-Hastings algorithm.
+Proposal distributions are modeled using independent normal distributions.
+Initial values and variances of the proposal distributions are chosen by trial-and-error to ensure reasonable acceptance rate (around 10\%) and convergence within 1,000,000 steps.
 Posterior samples are thinned every 1000 steps to remove autocorrelations among posterior samples.

### 2020-05-10 ncov.tex
done?

@@
 Proposal distributions are modeled using independent normal distributions.
-Initial values and variances of the proposal distributions are chosen by trial-and-error to ensure reasonable acceptance rate (around 10\%) and convergence within 1,000,000 steps.
+Initial values and variances of the proposal distributions are chosen by trial-and-error to ensure a reasonable acceptance rate (around 10\%) and convergence within 1,000,000 steps.
 Posterior samples are thinned every 1000 steps to remove autocorrelations among posterior samples.

### 2020-05-10 ncov.tex
the

@@
 
-We ran 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps using Metropolis-Hastings algorithm.
+We run 4 independent Markov Chain Monte Carlo chains each consisting of 500,000 burnin steps and 500,000 sampling steps using the Metropolis-Hastings algorithm.
 Proposal distributions are modeled using independent normal distributions.

### 2020-05-10 ncov.tex
tense

@@
 Ignoring these uncertainties leads to overly confident conclusions.
-% We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
 
@@
 We apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak in Wuhan City, China \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
-Previous studies have directly calculated the average of reported \Ro values  \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
+Previous studies directly calculate the average of reported \Ro values \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
 Instead, we decompose estimated \Ro into three key quantities ($r$, $\bar G$, and $\kappa$) and calculate the average (pooled estimates) of these key parameters.
@@
 \end{linenomath*}
-In other words, all model assumptions reduce to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-If a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
 Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.
+All model assumptions reduce to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+For example, if a model relies on strong assumptions about the underlying observation (how new cases are reported) or process (how new cases are generated) model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
 
@@
 \end{linenomath*}
-This equation demonstrates that a generation-interval distribution that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$) will give a higher estimate of \Ro for the same value of $r$ \citep{wallinga2007generation}.
+This equation demonstrates that a generation-interval distribution that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$) gives a higher estimate of \Ro for the same value of $r$ \citep{wallinga2007generation}.
 
@@
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
-We used gamma distributions to model values reported with confidence or credible intervals (CI) and uniform distributions to model values reported with ranges;
+We use gamma distributions to model values reported with confidence or credible intervals (CI) and uniform distributions to model values reported with ranges;
 when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
@@
 Here, we do so by averaging across reported values, without explicitly re-fitting their models.
-We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
+We use an Markov Chain Monte Carlo approach (cf.~S\ref{section:MCMC)} and account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
 Since the gamma distribution does not allow $\kappa=0$ (this corresponds to a Dirac delta generation-interval distribution), we substitute $\kappa=0.02$ for Study 7.
@@
 The pooled estimates can become sharper (i.e., have narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
-Since SARS-CoV-2 had primarily spread in Wuhan City, China during this period, it was not possible to include independent sources of data from other countries.
+Since SARS-CoV-2 spread primarily in Wuhan City, China, during this period, it is not possible to include independent sources of data from other countries.
 Thus, the pooled estimator should be interpreted with care.
@@
 
-We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
-We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances --- a known problem \citep{gelman2006prior}.
+We follow recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances --- a known problem \citep{gelman2006prior}.
 Alternative choices of prior for the between-study shape parameters are also suboptimal. 
@@
 \subsection{Markov Chain Monte Carlo}
+\label{section:MCMC}
 
@@
 Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01  \citep{gelman1992inference} and the effective sample size is greater than 1000 for all hyperparameters ($\mu_r, \mu_G, \mu_\kappa, \sigma_r, \sigma_G, \sigma_\kappa$);
-trace plots and marginal posterior distribution plots are presented in Appendix.
+trace plots and marginal posterior distribution plots are presented in the Appendix.
 95\% credible intervals (CI) are calculated by computing 2.5\% and 97.5\% quantiles from the marginal posterior distribution for each hyperparameter.
@@
 
-Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26, 2020).
-Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates were not necessarily finalized.
+Here, we focus on the estimates of \Ro that are published within a very short time frame (January 23--26, 2020).
+Since the estimates are published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates are not necessarily finalized.
 For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020 to $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
-we did not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
+we do not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
 Some studies also lack detailed description of their methods, data, and/or assumptions.
@@
 We strongly emphasize the value of attention to accurate characterization of the transmission chains via both contact tracing and improved statistical frameworks for inferring generation-interval distributions from such data \citep{britton2019estimation}.
-A combined effort between public-health workers and modelers in this direction will be crucial both for predicting the course of an epidemic and for controlling it.
+A combined effort between public-health workers and modelers in this direction is crucial both for predicting the course of an epidemic and for controlling it.
 We also emphasize the value of transparency from modelers.
@@
 While these estimates appear to be broadly consistent with earlier estimates from China, comparing the exponential growth rate and the underlying generation-interval distributions suggest otherwise.
-The later paper assumed a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
+The later paper assume a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
 based on these values, the exponential growth rate has to be considerably higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ than the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).

### 2020-05-10 posterior.R

@@ -7,7 +7,7 @@ library(gridExtra)
 
 load("MCMC_all.rda")
 
-parnames <- c("mu[r]~(days^{-1})", "sigma[r]", "mu[G]~(days)", "sigma[G]", "mu[kappa]", "sigma[kappa]")
+parnames <- c("mu[r]~(days^{-1})", "sigma[r]~(days^{-1})", "mu[G]~(days)", "sigma[G]~(days)", "mu[kappa]", "sigma[kappa]")
 
 gelman.diag(MCMC_all)
 
@@ -108,4 +108,4 @@ gpost <- ggplot(MCMC_summ) +
     strip.background = element_blank()
   )
 
-ggsave("posterior_dist.pdf", gpost, width=8, height=8)
+ggsave("posterior_dist.pdf", gpost, width=6, height=6)

### 2020-05-10 ncov.tex
done for now.

@@
 Here, we do so by averaging across reported values, without explicitly re-fitting their models.
-We use an Markov Chain Monte Carlo approach (cf.~S\ref{section:MCMC)} and account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
+We use a Markov Chain Monte Carlo approach (cf.~S\ref{section:MCMC}) and account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
 Since the gamma distribution does not allow $\kappa=0$ (this corresponds to a Dirac delta generation-interval distribution), we substitute $\kappa=0.02$ for Study 7.

### 2020-05-10 ncov.tex
tweak

@@
 However, as the number of susceptible individuals decreases or behavior changes in response to perception of the epidemic, the growth rate will decrease: estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
-Although our analysis strictly to the earliest stages of an epidemic, we expect certain lessons to hold more generally: credible intervals must combine as many sources of uncertainty as possible. 
+Although our analysis applies strictly to the earliest stages of an epidemic, we expect certain lessons to hold more generally: credible intervals must combine as many sources of uncertainty as possible.
 In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate (and other epidemiological parameters) will generally become more precise; thus the risk of over-confidence (when uncertainty about the generation-interval distribution is neglected) will become greater.
+Incorporating estimates of the dynamics of susceptibility (e.g., using properly calibrated serological studies \citep{metcalf2016use}) is also important for characterizing transmission as the outbreak progresses.
 

### 2020-05-12 ncov.tex
super-minor edits here and there

@@
 	\textbf\newline{
-		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the SARS-CoV-2 outbreak
+		Reconciling early-outbreak estimates of the basic reproductive number and its uncertainty: framework and applications to the novel coronavirus (SARS-CoV-2) outbreak
 	}
@@
 Although the virus is likely to have originated from animal hosts \citep{andersen2020proximal}, the ability of SARS-CoV-2 to directly transmit between humans, particularly without symptoms, has posed a greater threat for its spread \citep{he2020temporal}.
-As of May 2, 2020, more than 3 million cases of the coronavirus disease 2019 (COVID-19) have been confirmed internationally \citep{who103}.
+As of May 11, 2020, more than 4 million cases of the coronavirus disease 2019 (COVID-19) have been confirmed internationally \citep{who112}.
 
@@
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary considerably even though most analyses relied on similar data --- reports of confirmed cases from China, particularly from the Wuhan City.
+However, the estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) vary considerably even though most analyses rely on similar data --- reports of confirmed cases from China, particularly from Wuhan City.
 Comparing a disparate set of estimates of \Ro can be difficult when the estimation methods and their underlying assumptions vary widely.
@@
 To formalize the estimation of uncertainty at the onset of an outbreak, we present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies. 
-We apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak in Wuhan City, China \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
-Previous studies directly calculate the average of reported \Ro values \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
+We apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimate \Ro for the SARS-CoV-2 outbreak in Wuhan City, China \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+Previous studies have directly calculated the average of reported \Ro values \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
 Instead, we decompose estimated \Ro into three key quantities ($r$, $\bar G$, and $\kappa$) and calculate the average (pooled estimates) of these key parameters.
@@
 \hline
-Study 4 & Metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model & Confirmed cases from China and other countries\newline (Jan 1--21, 2020) & Not reported & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
+Study 4 & Deterministic Metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model & Confirmed cases from China and other countries\newline (Jan 1--21, 2020) & Not reported & 3.8 (95\% CI: 3.6--4.0) & 7.6 & 0.5 & \cite{readncov} \\
 \hline
@@
 Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.
-All model assumptions reduce to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+All model assumptions reduce to properties of the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
 For example, if a model relies on strong assumptions about the underlying observation (how new cases are reported) or process (how new cases are generated) model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
@@
 As most studies do not report their estimates of the exponential growth rate, we first re-estimate the exponential growth rate that corresponds to their model assumptions.
-Since the estimates of the basic reproductive number \Ro depend is determined by the exponential growth rate and the shape of generation-interval distributions, we can calculate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+Since the estimates of the basic reproductive number \Ro is determined by the exponential growth rate and the shape of generation-interval distributions, we can calculate the exponential growth rate from the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
@@
 when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
-This probability does not need to be equi-tailed.
+This probability does not need to depend on equi-tailed quantiles.
 For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
@@
 in other words, we still obtain the same degree of associated uncertainty in $(\Ro)_i$ if we calculate it from $r_i$, $\bar G_i$, and $\kappa_i$.
+
 For study 6, we fix $\bar G=8\,\textrm{days}$ and use the gamma distribution (Table 2) that corresponds to $\mathcal R_0 = 5.47$ (95\% CI: 4.16--7.10) during the re-estimation step for $r$ to remain consistent with the original study, which assumed $\bar G=8\,\textrm{days}$ for this particular estimate.
@@
 Here, we do so by averaging across reported values, without explicitly re-fitting their models.
-We use a Markov Chain Monte Carlo approach (cf.~S\ref{section:MCMC}) and account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
+We use a Markov Chain Monte Carlo approach (cf.~Section \ref{section:MCMC}) and account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
 Since the gamma distribution does not allow $\kappa=0$ (this corresponds to a Dirac delta generation-interval distribution), we substitute $\kappa=0.02$ for Study 7.
@@
 In order to compare estimates of the basic reproductive number \Ro (and particularly their associated uncertainties) across different studies, we need a consistent measure of uncertainty.
-Instead of using reported uncertainty ranges from the original studies, we re-calculate the basic reproductive number from the parameter sets ($r_i$, $\bar G_i$, and $\kappa_i$) for each study using \eref{gamma} and calculate the median and 95\% quantiles.
+Instead of using reported uncertainty ranges from the original studies, we re-calculate the basic reproductive number from the parameter sets ($r_i$, $\bar G_i$, and $\kappa_i$) for each study using \eref{gamma} and calculate the median and 95\% equi-tailed quantile.
 We refer to these estimates as the base estimates.
@@
 As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important.
-Propagating error only from the growth rate gives very narrow credible intervals in this case. 
-Propagating errors from the mean generation interval or generation-interval dispersion gives more realistic but still narrow credible intervals.
+Propagating error only from the growth rate ($\hat{\mu}_r$ in \fref{eff}B) gives very narrow credible intervals in this case. 
+Propagating errors from the mean generation interval ($\mu_G$ in \fref{eff}B) or generation-interval dispersion ($\mu_\kappa$ in \fref{eff}B) gives more realistic but still narrow credible intervals.
 
@@
   \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
-We compare estimates of \Ro under five scenarios that propagate different parameter uncertainties (A) based on our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) and (B) assuming a 4-fold reduction in uncertainty of our pooled estimate of the exponential growth rate (using $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$).
+We compare estimates of \Ro under nine scenarios that propagate different parameter uncertainties (A) based on our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) and (B) assuming a 4-fold reduction in uncertainty of our pooled estimate of the exponential growth rate (using $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$).
 Each uncertainty type represents \Ro estimates based the posterior distributions of one of three parameters ($\mu_r$, $\mu_G$, and $\mu_\kappa$) while using median estimates of two other parameters.
@@
 However, comparing disparate estimates of \Ro can be difficult when they rely on different methods and assumptions.
-Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the SARS-CoV-2 outbreak.
+Here, we use a gamma approximation framework \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the SARS-CoV-2 outbreak.
 Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the degree of dispersion in the generation intervals.
@@
 Assuming fixed parameter values here will lead to overly confident conclusions \citep{elderd2006uncertainty}.
+
 The lack of uncertainty in the generation-interval dispersion further explains the sensitivity of \Ro estimates to the exponential growth rate, particularly in Study 7 (\fref{R0}).
 Since Study 7 assumes a fixed generation interval ($\kappa=0$), they implicitly assume an exponential $r$--\Ro relationship, making their estimate too sensitive to $r$.
+Similarly, the credible intervals associated with the base estimates of Studies 3 ($\kappa=0.2$), 6  ($\kappa=0.2$), and 7 ($\kappa=0$) are wider than the credible intervals associated with their corresponding $\kappa$-substitute estimates, which rely on wider generation-interval distributions ($\mu_\kappa=0.50$; 95\% CI: 0.26--1.10) and, therefore, are less sensitive to uncertainties in $r$ and $\bar G$.
 One exception is Study 1: this estimate of \Ro is most sensitive to generation-interval dispersion $\kappa$,
 because the study assumes an exponentially distributed generation interval ($\kappa=1$). 
-Estimates that rely on this assumption make \Ro relatively insensitive --- implicitly assuming a linear $r$--\Ro relationship.
+Estimates that rely on this assumption implicitly assume a linear $r$--\Ro relationship.
 
 As most studies rely on overly confident assumptions, the credible intervals associated with the base estimates of \Ro should tend to be narrower than the credible intervals of the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6).
-While the point estimate of \Rpool is similar to other reported values from this date range, the credible intervals are wider than all but one study.
+While the point estimate of \Rpool is similar to other reported values from this date range, its credible interval is wider than the credible intervals of the base estimates of all but one study.
 This result does not mean that assumptions underlying the pooled estimate are too weak;
@@
 
-Of the seven studies that we reviewed, at least one of them directly fit their models to the cumulative number of confirmed cases.
+Of the seven studies that we review, at least one of them directly fit their models to the cumulative number of confirmed cases.
 This approach is appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence neglects autocorrelation between successive counts of cumulative cases. 
-As a result, this approach both biases parameter estimates and gives very narrow confidence intervals \citep{ma2014estimating, king2015avoidable}.
+As a result, this approach both biases parameter estimates and gives overly narrow credible intervals \citep{ma2014estimating, king2015avoidable}.
 Narrow uncertainties in the estimates of the exponential growth rate are probably driven by this approach.
@@
 Neglecting one or the other is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \citep{ma2014estimating}.
-Modelers should, however, be aware that oversimplifying the error model can give overly narrow confidence intervals \citep{king2015avoidable,taylor2016stochasticity}.
+Modelers should, however, be aware that oversimplifying the error model can give overly narrow credible intervals \citep{king2015avoidable,taylor2016stochasticity}.
 
@@
 Here, we focus on the estimates of \Ro that are published within a very short time frame (January 23--26, 2020).
-Since the estimates are published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates are not necessarily finalized.
+Since these estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates were not necessarily finalized.
 For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020 to $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
@@
 While these estimates appear to be broadly consistent with earlier estimates from China, comparing the exponential growth rate and the underlying generation-interval distributions suggest otherwise.
-The later paper assume a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
+The later paper assumes a shorter mean generation interval ($\bar G = 6.5\,\textrm{days}$) but similar generation-interval dispersion ($\kappa = 0.38$);
 based on these values, the exponential growth rate has to be considerably higher ($r = 0.27\,\textrm{days}^{-1}$) to obtain $\Ro = 3.87$ than the exponential growth rate observed in China ($\mu_r = 0.17\,\textrm{days}^{-1}$; 95\% CI: 0.12--0.25 $\textrm{days}^{-1}$).

### 2020-05-12 response.tex
ready?

@@
 A large proportion of the Results section has been moved to the Discussion section, as suggested by reviewer 2.
-We have also included a new paragraph in the Discussion section to compare early estimates with more recent estimates of the basic reproductive number.
+We have also included a new paragraph in the Discussion section to compare early estimates  of the basic reproductive number with more recent estimates.
 We had previous interpreted the uncertainty range reported by Study 2 as a 95\% credible interval; we now interpret it as 90\% credible interval as it better suits their definition.
@@
 \end{equation}
-In other words, all model assumptions reduce to assumptions about the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-If a model relies on strong assumptions about the underlying observation or process model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
-Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.''
+Therefore, it is sufficient to consider the estimates and assumptions about the exponential growth rates and the shapes of the generation-interval distributions to understand disparate estimates of the basic reproductive number.
+All model assumptions reduce to properties of the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
+For example, if a model relies on overly confident assumptions about the underlying observation (how new cases are reported) or process (how new cases are generated) model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.''
 
@@
 The pooled estimates can become sharper (i.e., have narrower credible intervals) as we add more models even when the models or the data no longer add more information about the epidemic.
-Since SARS-CoV-2 had primarily spread in Wuhan City, China during this period, it was not possible to include independent sources of data from other countries.
+Since SARS-CoV-2 spread primarily in Wuhan City, China, during this period, it is not possible to include independent sources of data from other countries.
 Thus, the pooled estimator should be interpreted with care.''
@@
 This hypothetical example reflects scenarios, in which increased data availability allows researchers to estimate $r$ with more certainty.
-To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (\fref{eff}B); 
-then $\hat{\mu}_r$ has the same median as $\mu_r$ but the associated 95\% CI is 4 times narrower (0.16--0.19 $\textrm{days}^{-1}$).
-As uncertainty associated with the exponential growth rate decreases, accounting for uncertainties in generation intervals becomes even more important.
-Propagating error only from the growth rate gives very narrow credible intervals in this case. 
-Propagating errors from the mean generation interval or generation-interval dispersion gives more realistic but still narrow credible intervals.''
+To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (Fig.~2B); 
+then $\hat{\mu}_r$ has the same median as $\mu_r$ but the associated 95\% CI is 4 times narrower (0.16--0.19 $\textrm{days}^{-1}$).''
 
@@
 
-We used Bayesian multilevel models precisely for this reason (to weight the studies appropriately based on reported certainty.)
+We used Bayesian multilevel models precisely for this reason (to weight the studies appropriately based on reported certainty).
 
@@
 This may be an overkill but we wanted to make sure that the posterior samples are not autocorrelated.
-On the other hand, this is definitely sufficient for the effective sample size to reliably calculate the 95\% confidence intervals, especially given that we're only averaging across previously estimated parameter values, rather than explicitly fitting a model to data.
+On the other hand, this is definitely sufficient for the effective sample size to reliably calculate the 95\% confidence intervals, especially given that we are simply averaging across previously estimated parameter values, rather than explicitly fitting a model to data.
 We now ensure that effective sample sizes are greater than at least 1,000 for all hyperparameters.
@@
 
-``Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in \fref{eff}), we see that propagating error from the growth rate (as done by all but one studies reviewed) is absolutely crucial: uncertainty in the pooled estimates for both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are overly narrow.''
+``Comparing the estimates that include only some sources of uncertainty to the pooled estimate ($\Rpool = 3.0$; 95\% CI: 2.1--4.6; see `all' in Fig.~2), we see that propagating error from the growth rate (as done by all but one studies reviewed) is absolutely crucial: uncertainty in the pooled estimates for both middle bars ($\mu_G$ and $\mu_\kappa$), which lack growth-rate uncertainty, are overly narrow.''
 
@@
 
-``Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26, 2020).
-Since the estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates were not necessarily finalized.
+``Here, we focus on the estimates of \Ro that are published within a very short time frame (January 23--26, 2020).
+Since these estimates were published as pre-prints, rather than in peer-reviewed journals, the quality of the analyses as well as the resulting estimates were not necessarily finalized.
 For example, Study 4 initially estimated $\Ro = 3.8$ (95\% CI: 3.6--4.0; \cite{readncov}) but revised their estimate on January 28, 2020 to $\Ro = 3.11$ (95\% CI: 2.39--4.13; \cite{readncov2});
-we did not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
+we do not include their revised estimates in our analysis in order to focus on available information at the very beginning of the outbreak.
 Some studies also lack detailed description of their methods, data, and/or assumptions.
@@
 
-We now clearly explain why we do not consider revised estimates. See above in the discussion about pre-prints.
+We now clearly explain why we do not consider revised estimates. See above in the discussion about pre-prints. We have also added the following text to the Methods section:
+
+``Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
+As our primary focus is on the resolution of uncertainty in the available information during the earliest stages of an epidemic, rather than to provide more precise or accurate estimates of \Ro, we focus strictly on estimates that were published between January 23--26, 2020.''
 
@@
 We match probability, not necessarily equi-tailed quantiles. A Gamma distribution distribution with mean
-2.2 and shape 12 has a 90\% probability of containing a value between 1.27 and 3.34. We make this clear in our main text now:
+2.2 and shape 12 has a 90\% probability of containing a value between 1.4 and 3.8. We make this clear in our main text now:
 
-``We used gamma distributions to model values reported with confidence or credible intervals (CI) and uniform distributions to model values reported with ranges;
+``We use gamma distributions to model values reported with confidence or credible intervals (CI) and uniform distributions to model values reported with ranges;
 when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
-This probability does not need to be equi-tailed.
+This probability does not need to be calculated from equi-tailed quantiles.
 For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
@@
 Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$) assumes \textit{a priori} that between-study variance is large (and therefore does not pool different estimates sufficiently).
-Overly weak priors (e.g. half-Cauchy(0,5)) leads to inefficient sampling and poor convergence.''
+Overly weak priors (e.g. half-Cauchy(0,5)) lead to inefficient sampling and poor convergence.''
 

### 2020-05-12 ncov.tex
ready?

@@
 The modeling approaches and resulting estimates of \Ro during the beginning of the outbreak vary widely, despite relying on similar data sources.
-Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate, the mean generation interval, and the generation-interval dispersion.
+Here, we present a statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate, the mean generation interval, and the generation-interval dispersion.
 We apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak, showing that many \Ro estimates are overly confident.
@@
 All model assumptions reduce to properties of the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-For example, if a model relies on strong assumptions about the underlying observation (how new cases are reported) or process (how new cases are generated) model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
+For example, if a model relies on overly confident assumptions about the underlying observation (how new cases are reported) or process (how new cases are generated) model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
 
@@
 when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
-This probability does not need to depend on equi-tailed quantiles.
+This probability does not need to be calculated from equi-tailed quantiles.
 For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
@@
 Imposing strong priors (e.g. half-$t(\mu=0,\sigma=1,\nu=4)$) assumes \textit{a priori} that between-study variance is large (and therefore does not pool different estimates sufficiently).
-Overly weak priors (e.g. half-Cauchy(0,5)) leads to inefficient sampling and poor convergence.
+Overly weak priors (e.g. half-Cauchy(0,5)) lead to inefficient sampling and poor convergence.
 
@@
 Vertical error bars represent the 95\% credible intervals of our base, substitute, and pooled estimates.
-Horizontal dashed lines represent the 95\% credible intervals of our pooled estimates.
+Horizontal dashed lines represent the 95\% credible intervals of our pooled estimate.
 }

### 2020-05-12 response.tex
bib

@@
 We have also included a new paragraph in the Discussion section to compare early estimates  of the basic reproductive number with more recent estimates.
-We had previous interpreted the uncertainty range reported by Study 2 as a 95\% credible interval; we now interpret it as 90\% credible interval as it better suits their definition.
+We had previously interpreted the uncertainty range reported by Study 2 as a 95\% credible interval; we now interpret it as 90\% credible interval as it better suits their definition.
 Otherwise, the main figures and the overall conclusions remain unchanged.

### 2020-06-21 backup/ncov_b1.tex
done

--- 2020-02-07 ncov_euro.tex
@@
 
-\usepackage[numbers,sort&compress]{natbib}
-\bibliographystyle{unsrtnat}
+\usepackage{natbib}
+\bibliographystyle{chicago}
 \date{\today}
@@
 \section*{Abstract}
-
-\subsection*{Background}
-
-A novel coronavirus (2019-nCoV) has recently emerged as a global threat.
-As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro, the average number of secondary cases caused by a primary case in an otherwise susceptible population.
+A novel coronavirus (SARS-CoV-2) has recently emerged as a global threat. 
+As the epidemic progresses, many disease modelers have focused on estimating the basic reproductive number \Ro -- the average number of secondary cases caused by a primary case in an otherwise susceptible population.
 The modeling approaches and resulting estimates of \Ro vary widely, despite relying on similar data sources.
-
-\subsection*{Aim}
-
-We aimed to develop a framework for comparing and combining different estimates of \Ro across a wide range of models.
-
-\subsection*{Methods}
-
-We reviewed 7 model-based analyses of the 2019-nCoV outbreak that were published online between January 23--26, 2020.
-We decompose their \Ro estimates into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.
-
-\subsection*{Results}
-
-We find that most early estimates of \Ro rely on strong assumptions, especially about the generation-interval dispersion.
-Estimates that rely on narrow generation-interval distributions are overly sensitive to estimates of the exponential growth rate.
-
-\subsection*{Conclusion}
-
+Here, we present a novel statistical framework for comparing and combining different estimates of \Ro across a wide range of models by decomposing the basic reproductive number into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
+We then apply our framework to early estimates of \Ro for the SARS-CoV-2 outbreak.
+We show that many early \Ro estimates are overly confident.
 Our results emphasize the importance of propagating uncertainties in all components of \Ro, including the shape of the generation-interval distribution, in efforts to estimate \Ro at the outset of an epidemic.
-
-\pagebreak
 
@@
 
-Basic reproductive number, 2019-nCoV, novel coronavirus, Bayesian multilevel model
-
-\section*{Funding}
-
-BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC).
-ML was supported by Canadian Institutes of Health Research (CIHR).
-The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
-
-\section*{Declaration of interests}
-
-We declare no competing interests.
-
-\section*{Acknowledgements}
-
-We thank Daihai He for providing helpful comments on the manuscript.
-
-\section*{Contribution}
-
-SWP and JD developed the statistical framework. 
-SWP reviewed the published literature.
-SWP performed the analysis. 
-SWP, BMB, and JD created the figures. 
-SWP and JD wrote the first draft.
-All authors contributed to the writing and approval of the final report.
-
-\section*{Data availability}
-
-\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
-
-\pagebreak
+SARS-CoV-2, COVID-19, novel coronavirus, basic reproductive number, generation interval, Bayesian multilevel model
 
@@
 
-Since December 2019 \cite{pneumonia}, a novel coronavirus (2019-nCoV) has been spreading in China and other parts of the world.
-Although the virus is believed to have originated from animal reservoirs, the ability of 2019-nCoV to directly transmit between humans has posed a greater threat for its spread \cite{chan2020familial}.
-As of February 3, 2020, the World Health Organization (WHO) has confirmed 17391 cases, including 153 confirmed cases in 23 different countries, outside China \cite{who02report}.
-% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \cite{whoemer}.
+Since December 2019, a novel coronavirus (SARS-CoV-2) has been spreading in China and other parts of the world \citep{pneumonia}.
+Although the virus is believed to have originated from animal reservoirs \citep{cdcncov}, the ability of SARS-CoV-2 ability to directly transmit between humans has posed a greater threat for its spread \citep{huang2020clinical,who26report}.
+As of February 27, 2020, the World Health Organization (WHO) has confirmed 82,294 cases of the coronavirus disease (COVID-19), including 3,664 confirmed cases in 46 different countries, outside China \citep{who38report}.
+% Also on January 30, 2020, WHO declared the outbreak a Public Health Emergency of International Concern \citep{whoemer}.
 
-As the disease continues to spread, many researchers have published their preliminary analyses of the outbreak as pre-prints \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov, riouncov}, focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \cite{anderson1991infectious}).
-Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission, and about the final size of the outbreak \cite{anderson1991infectious}.
+As the disease continues to spread, many researchers have already published their analyses of the outbreak as pre-prints (e.g., \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, zhaoncov}) and in peer-reviewed journals (e.g., \cite{li2020early, riou2020pattern, wu2020nowcasting, zhao2020preliminary}), focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
+Estimates of the basic reproductive number are of interest during an outbreak because they provide information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the potential final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
 We commend these researchers for their timely contribution and those who made the data publicly available.
-However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups when the estimation methods and their underlying assumptions vary widely.
+However, it can be difficult to compare a disparate set of estimates of \Ro from different research groups (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely.
 
 Here, we show that a wide range of approaches to estimating \Ro can be understood and compared in terms of estimates of three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
-The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \cite{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \cite{wallinga2007generation};
+The generation interval, defined as the interval between the time when an individual becomes infected and the time when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical};
 therefore, estimates of \Ro from different models directly depend on their implicit assumptions about the generation-interval distribution and the exponential growth rate.
-Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding disease transmission.
-We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \cite{elderd2006uncertainty}.
+Early in an epidemic, information is scarce and there is inevitably a great deal of uncertainty surrounding both case reports (affecting the estimates of the exponential growth rate) and contact tracing (affecting the estimates of the generation-interval distribution).
+We suggest that disease modelers should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty in \Ro should propagate error from all three sources \citep{elderd2006uncertainty}.
 
-We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the 2019-nCoV outbreak \cite{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
-We use our framework to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
-the pooled estimates reflect the uncertainties present in modeling approaches.
+We compare seven disparate models published online between January 23--26, 2020 that estimated \Ro for the SARS-CoV-2 outbreak \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
+We use a Bayesian multilevel model to construct pooled estimates for the three key quantities: $r$, $\bar G$, and $\kappa$;
+the pooled estimates reflect the uncertainties present in modeling approaches and their underlying assumptions.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. 
@@
 We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
-Five studies \cite{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \cite{imaincov}; and one report was posted on nextstrain.org \cite{bedfordncov}.
-Their modeling approaches vary widely: a branching process model \cite{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \cite{readncov}, an exponential growth model \cite{zhaoncov}, a Poisson offspring distribution model \cite{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \cite{majumderncov}.
-Four studies estimated \Ro by directly fitting their models to incidence data \cite{readncov, zhaoncov, liuncov, majumderncov}.
-The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \cite{imaincov0})
-Some of these studies have now been published in peer-reviewed journals \cite{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \cite{readncov2}.
+Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
+Their modeling approaches vary widely: a branching process model \citep{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \citep{readncov}, an exponential growth model \citep{zhaoncov}, a Poisson offspring distribution model \citep{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \citep{majumderncov}.
+Four studies estimated \Ro by directly fitting their models to incidence data \citep{readncov, zhaoncov, liuncov, majumderncov}.
+The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0})
+Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.
 
@@
 Early in an outbreak, \Ro is difficult to estimate directly;
-instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \cite{ma2014estimating}.
-Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \cite{champredon2015intrinsic}, the basic reproductive
-number can be estimated via the Euler-Lotka equation \cite{wallinga2007generation}:
+instead, \Ro is often inferred from the exponential growth rate $r$, which can be estimated reliably from incidence data \citep{ma2014estimating}.
+Given an estimate of the exponential growth rate $r$ and an \emph{intrinsic} generation-interval distribution $g(\tau)$ \citep{champredon2015intrinsic}, the basic reproductive
+number can be estimated via the Euler-Lotka equation \citep{wallinga2007generation}:
 \begin{linenomath*}
@@
 
-Here, we use the gamma approximation framework \cite{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
-Assuming that generation intervals follow a gamma distribution
-with the mean $\bar G$ and the squared coefficient of variation $\kappa$, 
-we have
+Here, we use the gamma approximation framework \citep{park2019practical} to (i) characterize the amount of uncertainty present in the exponential growth rates and the shape of the generation-interval distribution and (ii) assess the degree to which these uncertainties affect the estimate of \Ro.
+Assuming that generation intervals follow a gamma distribution with the mean $\bar G$ and the squared coefficient of variation $\kappa$, we have
 \begin{linenomath*}
@@
 \end{linenomath*}
-This equation demonstrates that a generation-interval distribution
-that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$)
-will give a higher estimate of \Ro for the same value of $r$.
+This equation demonstrates that a generation-interval distribution that has a larger mean (higher $\bar{G}$) or is less variable (lower $\kappa$) will give a higher estimate of \Ro for the same value of $r$ \citep{wallinga2007generation}.
 
@@
 We still account for this uncertainty in our pooled estimates ($\mu_G$).
-$^\dagger$Study 6 uses the IDEA model \cite{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
+$^\dagger$Study 6 uses the IDEA model \citep{fisman2013idea}, through which the authors effectively fit an exponential curve to the cumulative number of confirmed cases without propagating any statistical uncertainty.
 Instead of modeling \Ro with a probability distribution and recalculating $r$, we use $r=0.114\,\mathrm{days}^{-1}$, which explains all uncertainty in the reported \Ro, when combined with the considered range of $\bar G$.
@@
 \end{linenomath*}
-We followed recommendations outlined in Gelman \textit{et al.} \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
+We followed recommendations outlined in \cite{gelman2006prior}, parameterizing the top-level gamma distributions in terms of their means and standard deviations and imposing weakly informative prior distributions on between-study standard deviations, i.e., $\textrm{half-normal}(0, 10)$.
 We had initially used gamma priors with small shape parameters ($< 1$) on between-study shape parameters ($=\mu^2/\sigma^2$) but found this put too much prior probability on large between-study variances.
-This phenomenon is a known problem \cite{gelman2006prior}.
+This phenomenon is a known problem \citep{gelman2006prior}.
 Alternative choices of prior for the between-study shape parameters are also suboptimal: imposing strong priors (e.g.
@@
 Posterior samples are thinned every 1000 steps.
-Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \cite{gelman1992inference};
+Convergence is assessed by ensuring that the Gelman-Rubin statistic is below 1.01 for all hyperparameters \citep{gelman1992inference};
 trace plots and marginal posterior distribution plots are presented in Appendix.
@@
 all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
-% Assuming fixed parameter values can lead to overly strong conclusions \cite{elderd2006uncertainty}.
+Assuming fixed parameter values can lead to overly strong conclusions \citep{elderd2006uncertainty}.
 
@@
 Estimating the basic reproductive number \Ro is crucial for predicting the course of an outbreak and planning intervention strategies.
-Here, we use a gamma approximation \cite{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
-Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals:
-our analysis of individual studies shows that assuming too narrow a generation-interval distribution can make the estimate of \Ro overly sensitive to the estimates of the exponential growth rate $r$.
+Here, we use a gamma approximation \citep{park2019practical} to decompose \Ro estimates into three key quantities ($r$, $\bar G$, and $\kappa$) and apply a multilevel Bayesian framework to compare estimates of \Ro for the novel coronavirus outbreak.
+Our results demonstrate the importance of accounting for uncertainties associated with the underlying generation-interval distributions, including uncertainties in the amount of dispersion in the generation intervals.
+Our analysis of individual studies shows that many early estimates of \Ro rely on strong assumptions.
 
 Of the seven studies that we reviewed, two of them directly fit their models to cumulative number of confirmed cases.
-This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \cite{ma2014estimating, king2015avoidable}.
+This approach can be appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence instead of raw incidence can both bias parameters and give overly narrow confidence intervals, if the resulting non-independent error structures are not taken into account \citep{ma2014estimating, king2015avoidable}.
 Naive fits to cumulative incidence data should therefore be avoided.
@@
 Disease modelers face the choice of incorporating one or both of these in their data-fitting and modeling steps. 
-This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \cite{ma2014estimating}.
-Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \cite{king2015avoidable,taylor2016stochasticity}.
+This is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \citep{ma2014estimating}.
+Modelers should however be aware of the possibility that ignoring one kind of error can give overly narrow confidence intervals \citep{king2015avoidable,taylor2016stochasticity}.
 
@@
 Examples that seem relevant to this outbreak include: changing reporting rates, reporting delays (including the effects of weekends and holidays), and changing generation intervals.
-For emerging pathogens such as 2019-nCoV, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
-for example, Zhao \textit{et al.} \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \cite{who17protocol}, and January 20.
-Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \cite{tariq2019assessing}.
-% Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
-Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \cite{hethcote2002effects}.
+For emerging pathogens such as SARS-CoV-2, there may be an early period of time when the reporting rate is very low due to limited awareness or diagnostic resources;
+for example, \cite{zhaoncov} (Study 6) demonstrated that estimates of \Ro can change from 5.47 (95\% CI: 4.16--7.10) to 3.30 (95\% CI: 2.73--3.96) when they assume 2-fold changes in the reporting rate between January 17, when the official diagnostic guidelines were released \citep{who17protocol}, and January 20.
+Delays between key epidemiological timings (e.g., infection, symptom onset, and detection) can also shift the shape of an observed epidemic curve and, therefore, affect parameter estimates as well as predictions of the course of an outbreak \citep{tariq2019assessing}.
+Even though a constant delay between infection and detection may not affect the estimate of the growth rate, it can still affect the associated confidence intervals.
+Finally, generation intervals can become shorter throughout an epidemic as intervention strategies, such as quarantine, can reduce the infectious period \citep{hethcote2002effects}.
 Accounting for these factors is crucial for making accurate inferences.
@@
 Here, we focused on the estimates of \Ro that were published within a very short time frame (January 23--26).
-During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \cite{anderson1991infectious}.
+During early phases of an outbreak, it is reasonable to assume that the epidemic grows exponentially \citep{anderson1991infectious}.
 However, as the number of susceptible individuals decreases, the epidemic will saturate, and estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
@@
 
-We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical frameworks for inferring generation-interval distributions from such data \cite{britton2019estimation}.
+We strongly emphasize the value of attention to accurate characterization of the transmission chains via contact tracing and better statistical frameworks for inferring generation-interval distributions from such data \citep{britton2019estimation}.
 A combined effort between public-health workers and modelers in this direction will be crucial for predicting the course of an epidemic and controlling it.
@@
 We hope this framework will help researchers understand and reconcile disparate estimates of disease transmission early in an epidemic.
+
+\pagebreak
+
+\section*{Funding}
+
+BMB and DJDE were supported by Natural Sciences and Engineering Research Council (NSERC).
+ML was supported by Canadian Institutes of Health Research (CIHR).
+The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
+\section*{Competing interests}
+
+We declare no competing interests.
+
+\section*{Acknowledgements}
+
+We thank Daihai He for providing helpful comments on the manuscript.
+
+\section*{Contribution}
+
+SWP and JD developed the statistical framework. 
+SWP reviewed the published literature.
+SWP performed the analysis. 
+SWP, BMB, and JD created the figures. 
+SWP and JD wrote the first draft.
+All authors contributed to the writing and approval of the final report.
+
+\section*{Data availability}
+
+\texttt{R} code is available in GitHub (\url{https://github.com/parksw3/nCoV_framework}).
 

### 2020-06-21 backup/ncov_b2.tex
done

--- 2020-05-12 ncov.tex
@@
 
-\bibliography{ncov_abbr}
+\begin{thebibliography}{}
+
+\bibitem[\protect\citeauthoryear{Andersen, Rambaut, Lipkin, Holmes, and
+  Garry}{Andersen et~al.}{2020}]{andersen2020proximal}
+Andersen, K.~G., A.~Rambaut, W.~I. Lipkin, E.~C. Holmes, and R.~F. Garry
+  (2020).
+\newblock The proximal origin of {SARS-CoV-2}.
+\newblock {\em Nat Med\/}~{\em 26\/}(4), 450--452.
+
+\bibitem[\protect\citeauthoryear{Anderson and May}{Anderson and
+  May}{1991}]{anderson1991infectious}
+Anderson, R.~M. and R.~M. May (1991).
+\newblock {\em Infectious diseases of humans: dynamics and control}.
+\newblock Oxford University Press.
+
+\bibitem[\protect\citeauthoryear{Barton, Alberti, Ames, Atkinson, Bales, Burke,
+  Chen, Diallo, Earn, Fath, et~al.}{Barton et~al.}{2020}]{barton2020call}
+Barton, C.~M., M.~Alberti, D.~Ames, J.-A. Atkinson, J.~Bales, E.~Burke,
+  M.~Chen, S.~Y. Diallo, D.~J. Earn, B.~Fath, et~al. (2020).
+\newblock Call for transparency of {COVID-19} models.
+\newblock {\em Science\/}~{\em 368\/}(6490), 482--483.
+
+\bibitem[\protect\citeauthoryear{Bedford, Neher, Hadfield, Hodcroft, Ilcisin,
+  and M{\"u}ller}{Bedford et~al.}{2020}]{bedfordncov}
+Bedford, T., R.~Neher, J.~Hadfield, E.~Hodcroft, M.~Ilcisin, and N.~M{\"u}ller
+  (2020).
+\newblock {Genomic analysis of nCoV spread. Situation report 2020-01-23.}
+\newblock \url{https://nextstrain.org/narratives/ncov/sit-rep/2020-01-23}.
+  Accessed 24, January, 2020.
+
+\bibitem[\protect\citeauthoryear{Britton and Scalia~Tomba}{Britton and
+  Scalia~Tomba}{2019}]{britton2019estimation}
+Britton, T. and G.~Scalia~Tomba (2019).
+\newblock Estimation in emerging epidemics: Biases and remedies.
+\newblock {\em J R Soc Interface\/}~{\em 16\/}(150), 20180670.
+
+\bibitem[\protect\citeauthoryear{Champredon and Dushoff}{Champredon and
+  Dushoff}{2015}]{champredon2015intrinsic}
+Champredon, D. and J.~Dushoff (2015).
+\newblock Intrinsic and realized generation intervals in infectious-disease
+  transmission.
+\newblock {\em Proc R Soc Lond B Biol Sci\/}~{\em 282\/}(1821), 20152026.
+
+\bibitem[\protect\citeauthoryear{Diekmann, Heesterbeek, and Metz}{Diekmann
+  et~al.}{1990}]{diekmann1990definition}
+Diekmann, O., J.~A.~P. Heesterbeek, and J.~A. Metz (1990).
+\newblock On the definition and the computation of the basic reproduction ratio
+  $\mathcal{R}_0$ in models for infectious diseases in heterogeneous
+  populations.
+\newblock {\em J Math Biol\/}~{\em 28\/}(4), 365--382.
+
+\bibitem[\protect\citeauthoryear{Elderd, Dukic, and Dwyer}{Elderd
+  et~al.}{2006}]{elderd2006uncertainty}
+Elderd, B.~D., V.~M. Dukic, and G.~Dwyer (2006, October).
+\newblock Uncertainty in predictions of disease spread and public health
+  responses to bioterrorism and emerging diseases.
+\newblock {\em Proc Natl Acad Sci USA\/}~{\em 103\/}(42), 15693 --15697.
+
+\bibitem[\protect\citeauthoryear{Fisman, Hauck, Tuite, and Greer}{Fisman
+  et~al.}{2013}]{fisman2013idea}
+Fisman, D.~N., T.~S. Hauck, A.~R. Tuite, and A.~L. Greer (2013).
+\newblock {An IDEA for short term outbreak projection: nearcasting using the
+  basic reproduction number}.
+\newblock {\em PloS One\/}~{\em 8\/}(12).
+
+\bibitem[\protect\citeauthoryear{Flaxman, Mishra, Gandy, Unwin, Coupland,
+  Mellan, Zhu, Berah, Eaton, Guzman, Schmit, Callizo, Team, Whittaker,
+  Winskill, Xi, Ghani, Donnelly, Riley, Okell, Vollmer, Ferguson, and
+  Bhatt}{Flaxman et~al.}{2020}]{flaxman2020estimating}
+Flaxman, S., S.~Mishra, A.~Gandy, H.~J.~T. Unwin, H.~Coupland, T.~A. Mellan,
+  H.~Zhu, T.~Berah, J.~W. Eaton, P.~N.~P. Guzman, N.~Schmit, L.~Callizo, I.~C.
+  C.-.~R. Team, C.~Whittaker, P.~Winskill, X.~Xi, A.~Ghani, C.~A. Donnelly,
+  S.~Riley, L.~C. Okell, M.~A.~C. Vollmer, N.~M. Ferguson, and S.~Bhatt (2020).
+\newblock Estimating the number of infections and the impact of
+  non-pharmaceutical interventions on covid-19 in 11 european countries.
+\newblock
+  \url{https://www.imperial.ac.uk/mrc-global-infectious-disease-analysis/covid-19/report-13-europe-npi-impact/}.
+  Accessed May 4, 2020.
+
+\bibitem[\protect\citeauthoryear{Gelman et~al.}{Gelman
+  et~al.}{2006}]{gelman2006prior}
+Gelman, A. et~al. (2006).
+\newblock {Prior distributions for variance parameters in hierarchical models
+  (comment on article by Browne and Draper)}.
+\newblock {\em Bayesian analysis\/}~{\em 1\/}(3), 515--534.
+
+\bibitem[\protect\citeauthoryear{Gelman, Rubin, et~al.}{Gelman
+  et~al.}{1992}]{gelman1992inference}
+Gelman, A., D.~B. Rubin, et~al. (1992).
+\newblock Inference from iterative simulation using multiple sequences.
+\newblock {\em Stat Sci\/}~{\em 7\/}(4), 457--472.
+
+\bibitem[\protect\citeauthoryear{He, Lau, Wu, Deng, Wang, Hao, Lau, Wong, Guan,
+  Tan, et~al.}{He et~al.}{2020}]{he2020temporal}
+He, X., E.~H. Lau, P.~Wu, X.~Deng, J.~Wang, X.~Hao, Y.~C. Lau, J.~Y. Wong,
+  Y.~Guan, X.~Tan, et~al. (2020).
+\newblock Temporal dynamics in viral shedding and transmissibility of
+  {COVID-19}.
+\newblock {\em Nat Med\/}, 1--4.
+
+\bibitem[\protect\citeauthoryear{Hethcote, Zhien, and Shengbing}{Hethcote
+  et~al.}{2002}]{hethcote2002effects}
+Hethcote, H., M.~Zhien, and L.~Shengbing (2002).
+\newblock Effects of quarantine in six endemic models for infectious diseases.
+\newblock {\em Math Biosci\/}~{\em 180\/}(1-2), 141--160.
+
+\bibitem[\protect\citeauthoryear{Huang, Wang, Li, Ren, Zhao, Hu, Zhang, Fan,
+  Xu, Gu, et~al.}{Huang et~al.}{2020}]{huang2020clinical}
+Huang, C., Y.~Wang, X.~Li, L.~Ren, J.~Zhao, Y.~Hu, L.~Zhang, G.~Fan, J.~Xu,
+  X.~Gu, et~al. (2020).
+\newblock {Clinical features of patients infected with 2019 novel coronavirus
+  in Wuhan, China}.
+\newblock {\em Lancet\/}~{\em 395\/}(10223), 497--506.
+
+\bibitem[\protect\citeauthoryear{Imai, Cori, Dorigatti, Baguelin, Donelly,
+  Riley, and Ferguson}{Imai et~al.}{2020}]{imaincov}
+Imai, N., A.~Cori, I.~Dorigatti, M.~Baguelin, C.~A. Donelly, S.~Riley, and
+  N.~M. Ferguson (2020).
+\newblock {Report 3: Transmissibility of 2019-nCoV}.
+\newblock
+  \url{https://www.imperial.ac.uk/media/imperial-college/medicine/sph/ide/gida-fellowships/Imperial-2019-nCoV-transmissibility.pdf}.
+  Accessed 26, January, 2020.
+
+\bibitem[\protect\citeauthoryear{Imai, Dorigatti, Cori, Donelly, Riley, and
+  Ferguson}{Imai et~al.}{2020}]{imaincov0}
+Imai, N., I.~Dorigatti, A.~Cori, C.~A. Donelly, S.~Riley, and N.~M. Ferguson
+  (2020).
+\newblock {Report 2: Estimating the potential total number of novel Coronavirus
+  cases in Wuhan City, China}.
+\newblock
+  \url{https://www.imperial.ac.uk/media/imperial-college/medicine/sph/ide/gida-fellowships/2019-nCoV-outbreak-report-22-01-2020.pdf}.
+  Accessed 3, February, 2020.
+
+\bibitem[\protect\citeauthoryear{King, Domenech~de Cell{\`e}s, Magpantay, and
+  Rohani}{King et~al.}{2015}]{king2015avoidable}
+King, A.~A., M.~Domenech~de Cell{\`e}s, F.~M. Magpantay, and P.~Rohani (2015).
+\newblock Avoidable errors in the modelling of outbreaks of emerging pathogens,
+  with special reference to {Ebola}.
+\newblock {\em Proc R Soc Lond B Biol Sci\/}~{\em 282\/}(1806), 20150347.
+
+\bibitem[\protect\citeauthoryear{Lauer, Grantz, Bi, Jones, Zheng, Meredith,
+  Azman, Reich, and Lessler}{Lauer et~al.}{2020}]{lauer2020incubation}
+Lauer, S.~A., K.~H. Grantz, Q.~Bi, F.~K. Jones, Q.~Zheng, H.~R. Meredith, A.~S.
+  Azman, N.~G. Reich, and J.~Lessler (2020).
+\newblock {The incubation period of coronavirus disease 2019 (COVID-19) from
+  publicly reported confirmed cases: estimation and application}.
+\newblock {\em Ann Intern Med\/}.
+
+\bibitem[\protect\citeauthoryear{Li, Guan, Wu, Wang, Zhou, Tong, Ren, Leung,
+  Lau, Wong, et~al.}{Li et~al.}{2020}]{li2020early}
+Li, Q., X.~Guan, P.~Wu, X.~Wang, L.~Zhou, Y.~Tong, R.~Ren, K.~S. Leung, E.~H.
+  Lau, J.~Y. Wong, et~al. (2020).
+\newblock Early transmission dynamics in {Wuhan, China}, of novel
+  coronavirus--infected pneumonia.
+\newblock {\em N Engl J Med\/}.
+
+\bibitem[\protect\citeauthoryear{Liu, Hu, Kang, Lin, Zhong, Xiao, He, Song,
+  Huang, Rong, Deng, Zeng, Tan, Zeng, Zhu, Li, Wan, Lu, Deng, He, and Ma}{Liu
+  et~al.}{2020}]{liuncov}
+Liu, T., J.~Hu, M.~Kang, L.~Lin, H.~Zhong, J.~Xiao, G.~He, T.~Song, Q.~Huang,
+  Z.~Rong, A.~Deng, W.~Zeng, X.~Tan, S.~Zeng, Z.~Zhu, J.~Li, D.~Wan, J.~Lu,
+  H.~Deng, J.~He, and W.~Ma (2020).
+\newblock {Transmission dynamics of 2019 novel coronavirus (2019-nCoV)}.
+\newblock \url{https://www.biorxiv.org/content/10.1101/2020.01.25.919787v1}.
+  Accessed 27, January, 2020.
+
+\bibitem[\protect\citeauthoryear{Liu, Gayle, Wilder-Smith, and Rockl{\"o}v}{Liu
+  et~al.}{2020}]{liu2020reproductive}
+Liu, Y., A.~A. Gayle, A.~Wilder-Smith, and J.~Rockl{\"o}v (2020).
+\newblock {The reproductive number of COVID-19 is higher compared to SARS
+  coronavirus}.
+\newblock {\em J Travel Med\/}.
+
+\bibitem[\protect\citeauthoryear{Ma, Dushoff, Bolker, and Earn}{Ma
+  et~al.}{2014}]{ma2014estimating}
+Ma, J., J.~Dushoff, B.~M. Bolker, and D.~J. Earn (2014).
+\newblock Estimating initial epidemic growth rates.
+\newblock {\em Bull Math Biol\/}~{\em 76\/}(1), 245--260.
+
+\bibitem[\protect\citeauthoryear{Ma and Earn}{Ma and
+  Earn}{2006}]{ma2006generality}
+Ma, J. and D.~J. Earn (2006).
+\newblock Generality of the final size formula for an epidemic of a newly
+  invading infectious disease.
+\newblock {\em Bull Math Biol\/}~{\em 68\/}(3), 679--702.
+
+\bibitem[\protect\citeauthoryear{Majumder and Mandl}{Majumder and
+  Mandl}{2020a}]{majumderncov}
+Majumder, M. and K.~D. Mandl (2020a).
+\newblock Early transmissibility assessment of a novel coronavirus in {Wuhan,
+  China}.
+\newblock \url{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3524675}.
+  Accessed 27, January, 2020.
+
+\bibitem[\protect\citeauthoryear{Majumder and Mandl}{Majumder and
+  Mandl}{2020b}]{majumder2020early}
+Majumder, M.~S. and K.~D. Mandl (2020b).
+\newblock Early in the epidemic: impact of preprints on global discourse about
+  {COVID-19} transmissibility.
+\newblock {\em Lancet Glob Health\/}.
+
+\bibitem[\protect\citeauthoryear{McBryde, Bergeri, van Gemert, Rotty, Headley,
+  Simpson, Lester, Hellard, and Fielding}{McBryde
+  et~al.}{2009}]{mcbryde2009early}
+McBryde, E., I.~Bergeri, C.~van Gemert, J.~Rotty, E.~Headley, K.~Simpson,
+  R.~Lester, M.~Hellard, and J.~E. Fielding (2009).
+\newblock {Early transmission characteristics of influenza A (H1N1) v in
+  Australia: Victorian state, 16 May--3 June 2009}.
+\newblock {\em Euro Surveill\/}~{\em 14\/}(42), 19363.
+
+\bibitem[\protect\citeauthoryear{Metcalf, Farrar, Cutts, Basta, Graham,
+  Lessler, Ferguson, Burke, and Grenfell}{Metcalf
+  et~al.}{2016}]{metcalf2016use}
+Metcalf, C. J.~E., J.~Farrar, F.~T. Cutts, N.~E. Basta, A.~L. Graham,
+  J.~Lessler, N.~M. Ferguson, D.~S. Burke, and B.~T. Grenfell (2016).
+\newblock Use of serological surveys to generate key insights into the changing
+  global landscape of infectious disease.
+\newblock {\em Lancet\/}~{\em 388\/}(10045), 728--730.
+
+\bibitem[\protect\citeauthoryear{Nishiura, Castillo-Chavez, Safan, and
+  Chowell}{Nishiura et~al.}{2009}]{nishiura2009transmission}
+Nishiura, H., C.~Castillo-Chavez, M.~Safan, and G.~Chowell (2009).
+\newblock {Transmission potential of the new influenza A (H1N1) virus and its
+  age-specificity in Japan}.
+\newblock {\em Euro Surveill\/}~{\em 14\/}(22), 19227.
+
+\bibitem[\protect\citeauthoryear{Nishiura and Chowell}{Nishiura and
+  Chowell}{2015}]{nishiura2015theoretical}
+Nishiura, H. and G.~Chowell (2015).
+\newblock {Theoretical perspectives on the infectiousness of Ebola virus
+  disease}.
+\newblock {\em Theor Biol Med Model\/}~{\em 12\/}(1), 1.
+
+\bibitem[\protect\citeauthoryear{Park, Champredon, Weitz, and Dushoff}{Park
+  et~al.}{2019}]{park2019practical}
+Park, S.~W., D.~Champredon, J.~S. Weitz, and J.~Dushoff (2019).
+\newblock A practical generation-interval-based approach to inferring the
+  strength of epidemics from their speed.
+\newblock {\em Epidemics\/}~{\em 27}, 12--18.
+
+\bibitem[\protect\citeauthoryear{Read, Bridgen, Cummings, Ho, and Jewell}{Read
+  et~al.}{2020a}]{readncov}
+Read, J.~M., J.~R. Bridgen, D.~A. Cummings, A.~Ho, and C.~P. Jewell (2020a).
+\newblock {Novel coronavirus 2019-nCoV: early estimation of epidemiological
+  parameters and epidemic predictions}.
+\newblock \url{https://www.medrxiv.org/content/10.1101/2020.01.23.20018549v1}.
+  Accessed 26, January, 2020.
+
+\bibitem[\protect\citeauthoryear{Read, Bridgen, Cummings, Ho, and Jewell}{Read
+  et~al.}{2020b}]{readncov2}
+Read, J.~M., J.~R. Bridgen, D.~A. Cummings, A.~Ho, and C.~P. Jewell (2020b).
+\newblock {Novel coronavirus 2019-nCoV: early estimation of epidemiological
+  parameters and epidemic predictions}.
+\newblock \url{https://www.medrxiv.org/content/10.1101/2020.01.23.20018549v2}.
+  Accessed 5, February, 2020.
+
+\bibitem[\protect\citeauthoryear{Riou and Althaus}{Riou and
+  Althaus}{2020a}]{riouncov}
+Riou, J. and C.~L. Althaus (2020a).
+\newblock {Pattern of early human-to-human transmission of wuhan 2019-nCoV}.
+\newblock \url{https://www.biorxiv.org/content/10.1101/2020.01.23.917351v1}.
+  Accessed 26, January, 2020.
+
+\bibitem[\protect\citeauthoryear{Riou and Althaus}{Riou and
+  Althaus}{2020b}]{riou2020pattern}
+Riou, J. and C.~L. Althaus (2020b).
+\newblock {Pattern of early human-to-human transmission of Wuhan 2019 novel
+  coronavirus (2019-nCoV), December 2019 to January 2020}.
+\newblock {\em Euro Surveill\/}~{\em 25\/}(4), 2000058.
+
+\bibitem[\protect\citeauthoryear{Roberts and Heesterbeek}{Roberts and
+  Heesterbeek}{2007}]{roberts2007model}
+Roberts, M. and J.~Heesterbeek (2007).
+\newblock Model-consistent estimation of the basic reproduction number from the
+  incidence of an emerging infection.
+\newblock {\em J Math Biol\/}~{\em 55\/}(5-6), 803.
+
+\bibitem[\protect\citeauthoryear{Roberts and Nishiura}{Roberts and
+  Nishiura}{2011}]{roberts2011early}
+Roberts, M.~G. and H.~Nishiura (2011).
+\newblock {Early estimation of the reproduction number in the presence of
+  imported cases: pandemic influenza H1N1-2009 in New Zealand}.
+\newblock {\em PLoS One\/}~{\em 6\/}(5).
+
+\bibitem[\protect\citeauthoryear{Sun, Chen, and Viboud}{Sun
+  et~al.}{2020}]{sun2020early}
+Sun, K., J.~Chen, and C.~Viboud (2020).
+\newblock Early epidemiological analysis of the coronavirus disease 2019
+  outbreak based on crowdsourced data: a population-level observational study.
+\newblock {\em Lancet Dig Health\/}.
+
+\bibitem[\protect\citeauthoryear{Svensson}{Svensson}{2007}]{svensson2007note}
+Svensson, {\AA}. (2007).
+\newblock A note on generation times in epidemic models.
+\newblock {\em Math Biosci\/}~{\em 208\/}(1), 300--311.
+
+\bibitem[\protect\citeauthoryear{Tariq, Roosa, Mizumoto, and Chowell}{Tariq
+  et~al.}{2019}]{tariq2019assessing}
+Tariq, A., K.~Roosa, K.~Mizumoto, and G.~Chowell (2019).
+\newblock {Assessing reporting delays and the effective reproduction number:
+  The Ebola epidemic in DRC, May 2018--January 2019}.
+\newblock {\em Epidemics\/}~{\em 26}, 128--133.
+
+\bibitem[\protect\citeauthoryear{Taylor, Dushoff, and Weitz}{Taylor
+  et~al.}{2016}]{taylor2016stochasticity}
+Taylor, B.~P., J.~Dushoff, and J.~S. Weitz (2016).
+\newblock Stochasticity and the limits to confidence when estimating
+  $\mathcal{R}_0$ of {Ebola} and other emerging infectious diseases.
+\newblock {\em J Theor Biol\/}~{\em 408}, 145--154.
+
+\bibitem[\protect\citeauthoryear{Tian, Liu, Li, Wu, Chen, Kraemer, Li, Cai, Xu,
+  Yang, Wang, Yang, Cui, Song, Zheng, Wang, Bjornstad, Yang, Grenfell, Pybus,
+  and Dye}{Tian et~al.}{2020}]{Tianeabb6105}
+Tian, H., Y.~Liu, Y.~Li, C.-H. Wu, B.~Chen, M.~U.~G. Kraemer, B.~Li, J.~Cai,
+  B.~Xu, Q.~Yang, B.~Wang, P.~Yang, Y.~Cui, Y.~Song, P.~Zheng, Q.~Wang, O.~N.
+  Bjornstad, R.~Yang, B.~T. Grenfell, O.~G. Pybus, and C.~Dye (2020).
+\newblock An investigation of transmission control measures during the first 50
+  days of the covid-19 epidemic in china.
+\newblock {\em Science\/}.
+
+\bibitem[\protect\citeauthoryear{Trichereau, Verret, Mayet, Manet, Decam,
+  Meynard, Deparis, and Migliani}{Trichereau
+  et~al.}{2012}]{trichereau2012estimation}
+Trichereau, J., C.~Verret, A.~Mayet, G.~Manet, C.~Decam, J.-B. Meynard,
+  X.~Deparis, and R.~Migliani (2012).
+\newblock {Estimation of the reproductive number for A (H1N1) pdm09 influenza
+  among the French armed forces, September 2009--March 2010}.
+\newblock {\em J Infect\/}~{\em 64\/}(6), 628--630.
+
+\bibitem[\protect\citeauthoryear{Wallinga and Lipsitch}{Wallinga and
+  Lipsitch}{2007}]{wallinga2007generation}
+Wallinga, J. and M.~Lipsitch (2007).
+\newblock How generation intervals shape the relationship between growth rates
+  and reproductive numbers.
+\newblock {\em Proc R Soc Lond B Biol Sci\/}~{\em 274\/}(1609), 599--604.
+
+\bibitem[\protect\citeauthoryear{Wearing, Rohani, and Keeling}{Wearing
+  et~al.}{2005}]{wearing2005appropriate}
+Wearing, H.~J., P.~Rohani, and M.~J. Keeling (2005).
+\newblock Appropriate models for the management of infectious diseases.
+\newblock {\em PLoS Med\/}~{\em 2\/}(7).
+
+\bibitem[\protect\citeauthoryear{{World Health Organization}}{{World Health
+  Organization}}{2020a}]{who112}
+{World Health Organization} (2020a).
+\newblock {Coronavirus disease 2019 (COVID-19) Situation Report - 112}.
+\newblock
+  \url{https://www.who.int/docs/default-source/coronaviruse/situation-reports/20200511-covid-19-sitrep-112.pdf?sfvrsn=813f2669_2}.
+  Accessed May 11, 2020.
+
+\bibitem[\protect\citeauthoryear{{World Health Organization}}{{World Health
+  Organization}}{2020b}]{who17protocol}
+{World Health Organization} (2020b).
+\newblock {Laboratory testing for 2019 novel coronavirus (2019-nCoV) in
+  suspected human cases}.
+\newblock
+  \url{https://www.who.int/publications-detail/laboratory-testing-for-2019-novel-coronavirus-in-suspected-human-cases-20200117}.
+  Accessed February 4, 2020.
+
+\bibitem[\protect\citeauthoryear{{World Health Organization}}{{World Health
+  Organization}}{2020c}]{pneumonia}
+{World Health Organization} (2020c).
+\newblock Pneumonia of unknown cause -- {China}.
+\newblock
+  \url{https://www.who.int/csr/don/05-january-2020-pneumonia-of-unkown-cause-china/en/}.
+  Accessed January 30, 2020.
+
+\bibitem[\protect\citeauthoryear{Wu, Leung, and Leung}{Wu
+  et~al.}{2020}]{wu2020nowcasting}
+Wu, J.~T., K.~Leung, and G.~M. Leung (2020).
+\newblock {Nowcasting and forecasting the potential domestic and international
+  spread of the 2019-nCoV outbreak originating in Wuhan, China: a modelling
+  study}.
+\newblock {\em Lancet\/}.
+
+\bibitem[\protect\citeauthoryear{Zhao, Lin, Ran, Musa, Yang, Wang, Lou, Gao,
+  Yang, He, et~al.}{Zhao et~al.}{2020}]{zhao2020preliminary}
+Zhao, S., Q.~Lin, J.~Ran, S.~S. Musa, G.~Yang, W.~Wang, Y.~Lou, D.~Gao,
+  L.~Yang, D.~He, et~al. (2020).
+\newblock {Preliminary estimation of the basic reproduction number of novel
+  coronavirus (2019-nCoV) in China, from 2019 to 2020: A data-driven analysis
+  in the early phase of the outbreak}.
+\newblock {\em Int J Infect Dis\/}.
+
+\bibitem[\protect\citeauthoryear{Zhao, Ran, Musa, Yang, Lou, Gao, Yang, and
+  He}{Zhao et~al.}{2020}]{zhaoncov}
+Zhao, S., J.~Ran, S.~S. Musa, G.~Yang, Y.~Lou, D.~Gao, L.~Yang, and D.~He
+  (2020).
+\newblock {Preliminary estimation of the basic reproduction number of novel
+  coronavirus (2019-nCoV) in China, from 2019 to 2020: A data-driven analysis
+  in the early phase of the outbreak}.
+\newblock \url{https://www.biorxiv.org/content/10.1101/2020.01.23.916395v1}.
+  Accessed 26, January, 2020.
+
+\end{thebibliography}
 

### 2020-06-21 compare_R0.tex
done

@@
-% Created by tikzDevice version 0.12.3 on 2020-05-07 17:27:07
+% Created by tikzDevice version 0.12.3 on 2020-06-20 18:53:21
 % !TEX encoding = UTF-8 Unicode
@@
 \begin{scope}
-\path[clip] ( 27.31, 18.22) rectangle (572.66,247.13);
+\path[clip] ( 27.31, 18.22) rectangle (572.66,232.67);
 \definecolor{fillColor}{RGB}{255,255,255}
 
-\path[fill=fillColor] ( 27.31, 18.22) rectangle (572.66,247.13);
+\path[fill=fillColor] ( 27.31, 18.22) rectangle (572.66,232.67);
 \definecolor{drawColor}{gray}{0.92}
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 25.76) --
-	(572.66, 25.76);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 25.29) --
+	(572.66, 25.29);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 79.98) --
-	(572.66, 79.98);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31, 76.08) --
+	(572.66, 76.08);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,134.20) --
-	(572.66,134.20);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,126.87) --
+	(572.66,126.87);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,188.41) --
-	(572.66,188.41);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,177.67) --
+	(572.66,177.67);
 
-\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,242.63) --
-	(572.66,242.63);
+\path[draw=drawColor,line width= 0.3pt,line join=round] ( 27.31,228.46) --
+	(572.66,228.46);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 52.87) --
-	(572.66, 52.87);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 50.68) --
+	(572.66, 50.68);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,107.09) --
-	(572.66,107.09);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,101.48) --
+	(572.66,101.48);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,161.31) --
-	(572.66,161.31);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,152.27) --
+	(572.66,152.27);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,215.52) --
-	(572.66,215.52);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,203.06) --
+	(572.66,203.06);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.22, 18.22) --
-	( 67.22,247.13);
+	( 67.22,232.67);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] (133.72, 18.22) --
-	(133.72,247.13);
+	(133.72,232.67);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] (200.23, 18.22) --
-	(200.23,247.13);
+	(200.23,232.67);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] (266.73, 18.22) --
-	(266.73,247.13);
+	(266.73,232.67);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] (333.24, 18.22) --
-	(333.24,247.13);
+	(333.24,232.67);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] (399.75, 18.22) --
-	(399.75,247.13);
+	(399.75,232.67);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] (466.25, 18.22) --
-	(466.25,247.13);
+	(466.25,232.67);
 
 \path[draw=drawColor,line width= 0.6pt,line join=round] (532.76, 18.22) --
-	(532.76,247.13);
+	(532.76,232.67);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 55.23) -- (572.66, 55.23);
+\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 52.89) -- (572.66, 52.89);
 
-\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,122.16) -- (572.66,122.16);
-\definecolor{drawColor}{RGB}{0,0,0}
+\path[draw=drawColor,draw opacity=0.50,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,115.60) -- (572.66,115.60);
+\definecolor{drawColor}{RGB}{182,117,224}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 75.59) rectangle ( 70.79, 82.72);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 71.74) -- ( 70.79, 78.88);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 75.59) -- ( 70.79, 82.72);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 78.88) -- ( 70.79, 71.74);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.65, 82.72) -- ( 70.79, 75.59);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.17, 75.31) -- ( 72.26, 75.31);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (141.15, 78.87) -- (151.24, 78.87);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 67.22, 70.26) -- ( 67.22, 80.35);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.19, 73.83) -- (146.19, 83.92);
-\definecolor{fillColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.45, 71.47) rectangle (150.59, 78.61);
 
-\path[fill=fillColor] (134.31, 56.77) --
-	(141.45, 56.77) --
-	(141.45, 63.91) --
-	(134.31, 63.91) --
-	cycle;
-\definecolor{fillColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.45, 71.47) -- (150.59, 78.61);
 
-\path[fill=fillColor] (129.57, 78.05) --
-	(134.37, 69.73) --
-	(124.76, 69.73) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.45, 78.61) -- (150.59, 71.47);
+\definecolor{drawColor}{RGB}{0,170,90}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.33, 57.68) -- (145.42, 57.68);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.37, 52.63) -- (140.37, 62.73);
+\definecolor{fillColor}{RGB}{170,144,0}
+
+\path[fill=fillColor] (130.15, 65.51) --
+	(137.29, 65.51) --
+	(137.29, 72.64) --
+	(130.15, 72.64) --
 	cycle;
@@
 
-\path[fill=fillColor] (121.25, 66.37) circle (  3.57);
+\path[fill=fillColor] (127.07, 68.88) --
+	(131.88, 60.55) --
+	(122.27, 60.55) --
+	cycle;
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.65, 66.49) -- (217.74, 66.49);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.96, 59.87) rectangle (217.10, 67.01);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.70, 61.44) -- (212.70, 71.54);
-\definecolor{fillColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.96, 59.87) -- (217.10, 67.01);
 
-\path[fill=fillColor] (200.82, 65.00) --
-	(207.95, 65.00) --
-	(207.95, 72.13) --
-	(200.82, 72.13) --
-	cycle;
-\definecolor{fillColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.96, 67.01) -- (217.10, 59.87);
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[fill=fillColor] (196.07, 84.79) --
-	(200.88, 76.46) --
-	(191.27, 76.46) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.83, 65.39) -- (211.93, 65.39);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.88, 60.34) -- (206.88, 70.43);
+\definecolor{fillColor}{RGB}{170,144,0}
+
+\path[fill=fillColor] (196.66, 71.82) --
+	(203.80, 71.82) --
+	(203.80, 78.95) --
+	(196.66, 78.95) --
 	cycle;
@@
 
-\path[fill=fillColor] (187.76, 67.91) circle (  3.57);
+\path[fill=fillColor] (193.58, 70.32) --
+	(198.38, 62.00) --
+	(188.77, 62.00) --
+	cycle;
+\definecolor{fillColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (274.16, 66.66) -- (284.25, 66.66);
+\path[fill=fillColor] (186.93, 65.92) circle (  3.57);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.20, 61.62) -- (279.20, 71.71);
-\definecolor{fillColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.47, 60.04) rectangle (283.60, 67.17);
 
-\path[fill=fillColor] (267.32, 74.74) --
-	(274.46, 74.74) --
-	(274.46, 81.88) --
-	(267.32, 81.88) --
-	cycle;
-\definecolor{fillColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.47, 60.04) -- (283.60, 67.17);
 
-\path[fill=fillColor] (262.58,100.58) --
-	(267.38, 92.26) --
-	(257.77, 92.26) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.47, 67.17) -- (283.60, 60.04);
+\definecolor{drawColor}{RGB}{0,170,90}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.34, 74.52) -- (278.43, 74.52);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.38, 69.47) -- (273.38, 79.56);
+\definecolor{fillColor}{RGB}{170,144,0}
+
+\path[fill=fillColor] (263.17, 86.61) --
+	(270.30, 86.61) --
+	(270.30, 93.75) --
+	(263.17, 93.75) --
 	cycle;
@@
 
-\path[fill=fillColor] (254.26, 77.41) circle (  3.57);
+\path[fill=fillColor] (260.08, 79.22) --
+	(264.89, 70.90) --
+	(255.28, 70.90) --
+	cycle;
+\definecolor{fillColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.66,101.40) -- (350.76,101.40);
+\path[fill=fillColor] (253.43, 74.05) circle (  3.57);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.71, 96.36) -- (345.71,106.45);
-\definecolor{fillColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.97, 92.58) rectangle (350.11, 99.72);
 
-\path[fill=fillColor] (333.83,110.42) --
-	(340.96,110.42) --
-	(340.96,117.56) --
-	(333.83,117.56) --
-	cycle;
-\definecolor{fillColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.97, 92.58) -- (350.11, 99.72);
 
-\path[fill=fillColor] (329.08, 78.47) --
-	(333.89, 70.15) --
-	(324.28, 70.15) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.97, 99.72) -- (350.11, 92.58);
+\definecolor{drawColor}{RGB}{0,170,90}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.84,107.95) -- (344.94,107.95);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.89,102.90) -- (339.89,112.99);
+\definecolor{fillColor}{RGB}{170,144,0}
+
+\path[fill=fillColor] (329.67, 65.90) --
+	(336.81, 65.90) --
+	(336.81, 73.04) --
+	(329.67, 73.04) --
 	cycle;
@@
 
-\path[fill=fillColor] (320.77,101.64) circle (  3.57);
+\path[fill=fillColor] (326.59,101.92) --
+	(331.40, 93.60) --
+	(321.78, 93.60) --
+	cycle;
+\definecolor{fillColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (407.17, 55.76) -- (417.26, 55.76);
+\path[fill=fillColor] (319.94, 96.40) circle (  3.57);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.22, 50.72) -- (412.22, 60.81);
-\definecolor{fillColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.48, 49.82) rectangle (416.61, 56.96);
 
-\path[fill=fillColor] (400.33, 46.35) --
-	(407.47, 46.35) --
-	(407.47, 53.48) --
-	(400.33, 53.48) --
-	cycle;
-\definecolor{fillColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.48, 49.82) -- (416.61, 56.96);
 
-\path[fill=fillColor] (395.59,101.86) --
-	(400.39, 93.54) --
-	(390.78, 93.54) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (409.48, 56.96) -- (416.61, 49.82);
+\definecolor{drawColor}{RGB}{0,170,90}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.35, 47.91) -- (411.44, 47.91);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.40, 42.87) -- (406.40, 52.96);
+\definecolor{fillColor}{RGB}{170,144,0}
+
+\path[fill=fillColor] (396.18, 87.82) --
+	(403.31, 87.82) --
+	(403.31, 94.95) --
+	(396.18, 94.95) --
 	cycle;
@@
 
-\path[fill=fillColor] (387.28, 56.59) circle (  3.57);
+\path[fill=fillColor] (393.09, 59.72) --
+	(397.90, 51.39) --
+	(388.29, 51.39) --
+	cycle;
+\definecolor{fillColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (473.67,105.92) -- (483.77,105.92);
+\path[fill=fillColor] (386.44, 55.76) circle (  3.57);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.72,100.87) -- (478.72,110.97);
-\definecolor{fillColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.98, 96.81) rectangle (483.12,103.95);
 
-\path[fill=fillColor] (466.84,156.18) --
-	(473.98,156.18) --
-	(473.98,163.32) --
-	(466.84,163.32) --
-	cycle;
-\definecolor{fillColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.98, 96.81) -- (483.12,103.95);
 
-\path[fill=fillColor] (462.09, 95.96) --
-	(466.90, 87.64) --
-	(457.29, 87.64) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.98,103.95) -- (483.12, 96.81);
+\definecolor{drawColor}{RGB}{0,170,90}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.85,150.81) -- (477.95,150.81);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.90,145.76) -- (472.90,155.86);
+\definecolor{fillColor}{RGB}{170,144,0}
+
+\path[fill=fillColor] (462.68, 82.29) --
+	(469.82, 82.29) --
+	(469.82, 89.42) --
+	(462.68, 89.42) --
 	cycle;
@@
 
-\path[fill=fillColor] (453.78,145.76) circle (  3.57);
+\path[fill=fillColor] (459.60,143.26) --
+	(464.41,134.93) --
+	(454.79,134.93) --
+	cycle;
+\definecolor{fillColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (540.18, 55.35) -- (550.27, 55.35);
+\path[fill=fillColor] (452.95,138.81) circle (  3.57);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.23, 50.30) -- (545.23, 60.40);
-\definecolor{fillColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.49, 49.44) rectangle (549.63, 56.57);
 
-\path[fill=fillColor] (533.34, 66.58) --
-	(540.48, 66.58) --
-	(540.48, 73.72) --
-	(533.34, 73.72) --
-	cycle;
-\definecolor{fillColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.49, 49.44) -- (549.63, 56.57);
 
-\path[fill=fillColor] (528.60,111.22) --
-	(533.41,102.89) --
-	(523.79,102.89) --
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (542.49, 56.57) -- (549.63, 49.44);
+\definecolor{drawColor}{RGB}{0,170,90}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.36, 66.87) -- (544.45, 66.87);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.41, 61.82) -- (539.41, 71.92);
+\definecolor{fillColor}{RGB}{170,144,0}
+
+\path[fill=fillColor] (529.19, 96.58) --
+	(536.33, 96.58) --
+	(536.33,103.71) --
+	(529.19,103.71) --
 	cycle;
@@
 
-\path[fill=fillColor] (520.29, 66.12) circle (  3.57);
+\path[fill=fillColor] (526.11, 68.64) --
+	(530.91, 60.32) --
+	(521.30, 60.32) --
+	cycle;
+\definecolor{drawColor}{RGB}{182,117,224}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,115.60) --
+	( 67.22,115.60);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,115.60) --
+	( 67.22, 52.89);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22, 52.89) --
+	( 67.22, 52.89);
+\definecolor{drawColor}{RGB}{0,166,202}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (147.02,140.34) --
+	(147.02,140.34);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (147.02,140.34) --
+	(147.02, 40.98);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (147.02, 40.98) --
+	(147.02, 40.98);
+\definecolor{drawColor}{RGB}{0,170,90}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (140.37, 79.47) --
+	(140.37, 79.47);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (140.37, 79.47) --
+	(140.37, 37.16);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (140.37, 37.16) --
+	(140.37, 37.16);
+\definecolor{drawColor}{RGB}{170,144,0}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (133.72, 88.80) --
+	(133.72, 88.80);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (133.72, 88.80) --
+	(133.72, 55.19);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (133.72, 55.19) --
+	(133.72, 55.19);
+\definecolor{drawColor}{RGB}{225,106,134}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (127.07, 87.46) --
+	(127.07, 87.46);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (127.07, 87.46) --
+	(127.07, 39.29);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (127.07, 39.29) --
+	(127.07, 39.29);
 \definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,122.16) --
-	( 67.22,122.16);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (120.42, 88.78) --
+	(120.42, 88.78);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22,122.16) --
-	( 67.22, 55.23);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (120.42, 88.78) --
+	(120.42, 37.98);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 67.22, 55.23) --
-	( 67.22, 55.23);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (120.42, 37.98) --
+	(120.42, 37.98);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19,148.57) --
-	(146.19,148.57);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (213.53,103.67) --
+	(213.53,103.67);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19,148.57) --
-	(146.19, 42.52);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (213.53,103.67) --
+	(213.53, 38.86);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (146.19, 42.52) --
-	(146.19, 42.52);
-\definecolor{drawColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (213.53, 38.86) --
+	(213.53, 38.86);
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 83.60) --
-	(137.88, 83.60);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (206.88,102.77) --
+	(206.88,102.77);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 83.60) --
-	(137.88, 38.44);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (206.88,102.77) --
+	(206.88, 39.19);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (137.88, 38.44) --
-	(137.88, 38.44);
-\definecolor{drawColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (206.88, 39.19) --
+	(206.88, 39.19);
+\definecolor{drawColor}{RGB}{170,144,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 93.55) --
-	(129.57, 93.55);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (200.23,106.65) --
+	(200.23,106.65);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 93.55) --
-	(129.57, 57.68);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (200.23,106.65) --
+	(200.23, 56.62);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (129.57, 57.68) --
-	(129.57, 57.68);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (200.23, 56.62) --
+	(200.23, 56.62);
 \definecolor{drawColor}{RGB}{225,106,134}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 92.13) --
-	(121.25, 92.13);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (193.58, 99.67) --
+	(193.58, 99.67);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 92.13) --
-	(121.25, 40.71);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (193.58, 99.67) --
+	(193.58, 39.13);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (121.25, 40.71) --
-	(121.25, 40.71);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (193.58, 39.13) --
+	(193.58, 39.13);
+\definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70,109.43) --
-	(212.70,109.43);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (186.93, 88.78) --
+	(186.93, 88.78);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70,109.43) --
-	(212.70, 40.25);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (186.93, 88.78) --
+	(186.93, 37.98);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (212.70, 40.25) --
-	(212.70, 40.25);
-\definecolor{drawColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (186.93, 37.98) --
+	(186.93, 37.98);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38,108.47) --
-	(204.38,108.47);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (280.03, 80.74) --
+	(280.03, 80.74);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38,108.47) --
-	(204.38, 40.60);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (280.03, 80.74) --
+	(280.03, 50.55);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (204.38, 40.60) --
-	(204.38, 40.60);
-\definecolor{drawColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (280.03, 50.55) --
+	(280.03, 50.55);
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07,112.61) --
-	(196.07,112.61);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (273.38, 97.82) --
+	(273.38, 97.82);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07,112.61) --
-	(196.07, 59.21);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (273.38, 97.82) --
+	(273.38, 57.16);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (196.07, 59.21) --
-	(196.07, 59.21);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (273.38, 57.16) --
+	(273.38, 57.16);
+\definecolor{drawColor}{RGB}{170,144,0}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (266.73,145.65) --
+	(266.73,145.65);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (266.73,145.65) --
+	(266.73, 62.40);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (266.73, 62.40) --
+	(266.73, 62.40);
 \definecolor{drawColor}{RGB}{225,106,134}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76,105.16) --
-	(187.76,105.16);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (260.08, 92.92) --
+	(260.08, 92.92);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76,105.16) --
-	(187.76, 40.54);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (260.08, 92.92) --
+	(260.08, 57.46);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (187.76, 40.54) --
-	(187.76, 40.54);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (260.08, 57.46) --
+	(260.08, 57.46);
+\definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 84.95) --
-	(279.20, 84.95);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (253.43, 93.10) --
+	(253.43, 93.10);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 84.95) --
-	(279.20, 52.73);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (253.43, 93.10) --
+	(253.43, 57.79);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (279.20, 52.73) --
-	(279.20, 52.73);
-\definecolor{drawColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (253.43, 57.79) --
+	(253.43, 57.79);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89,103.19) --
-	(270.89,103.19);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (346.54,120.10) --
+	(346.54,120.10);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89,103.19) --
-	(270.89, 59.79);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (346.54,120.10) --
+	(346.54, 70.70);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (270.89, 59.79) --
-	(270.89, 59.79);
-\definecolor{drawColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (346.54, 70.70) --
+	(346.54, 70.70);
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58,154.24) --
-	(262.58,154.24);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (339.89,124.43) --
+	(339.89,124.43);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58,154.24) --
-	(262.58, 65.38);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (339.89,124.43) --
+	(339.89, 95.11);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (262.58, 65.38) --
-	(262.58, 65.38);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (339.89, 95.11) --
+	(339.89, 95.11);
+\definecolor{drawColor}{RGB}{170,144,0}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (333.24, 96.49) --
+	(333.24, 96.49);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (333.24, 96.49) --
+	(333.24, 53.10);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (333.24, 53.10) --
+	(333.24, 53.10);
 \definecolor{drawColor}{RGB}{225,106,134}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 97.96) --
-	(254.26, 97.96);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (326.59,101.51) --
+	(326.59,101.51);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 97.96) --
-	(254.26, 60.10);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (326.59,101.51) --
+	(326.59, 91.41);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (254.26, 60.10) --
-	(254.26, 60.10);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (326.59, 91.41) --
+	(326.59, 91.41);
+\definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71,126.97) --
-	(345.71,126.97);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (319.94,101.48) --
+	(319.94,101.48);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71,126.97) --
-	(345.71, 74.24);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (319.94,101.48) --
+	(319.94, 91.32);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (345.71, 74.24) --
-	(345.71, 74.24);
-\definecolor{drawColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (319.94, 91.32) --
+	(319.94, 91.32);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,131.59) --
-	(337.40,131.59);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (413.05, 93.24) --
+	(413.05, 93.24);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,131.59) --
-	(337.40,100.29);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (413.05, 93.24) --
+	(413.05, 28.65);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (337.40,100.29) --
-	(337.40,100.29);
-\definecolor{drawColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (413.05, 28.65) --
+	(413.05, 28.65);
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08,101.76) --
-	(329.08,101.76);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (406.40, 86.66) --
+	(406.40, 86.66);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08,101.76) --
-	(329.08, 55.45);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (406.40, 86.66) --
+	(406.40, 27.97);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (329.08, 55.45) --
-	(329.08, 55.45);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (406.40, 27.97) --
+	(406.40, 27.97);
+\definecolor{drawColor}{RGB}{170,144,0}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (399.75,156.23) --
+	(399.75,156.23);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (399.75,156.23) --
+	(399.75, 58.32);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (399.75, 58.32) --
+	(399.75, 58.32);
 \definecolor{drawColor}{RGB}{225,106,134}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77,107.13) --
-	(320.77,107.13);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (393.09, 91.27) --
+	(393.09, 91.27);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77,107.13) --
-	(320.77, 96.34);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (393.09, 91.27) --
+	(393.09, 28.65);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (320.77, 96.34) --
-	(320.77, 96.34);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (393.09, 28.65) --
+	(393.09, 28.65);
+\definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 98.29) --
-	(412.22, 98.29);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (386.44, 96.40) --
+	(386.44, 96.40);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 98.29) --
-	(412.22, 29.35);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (386.44, 96.40) --
+	(386.44, 35.44);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (412.22, 29.35) --
-	(412.22, 29.35);
-\definecolor{drawColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (386.44, 35.44) --
+	(386.44, 35.44);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 91.28) --
-	(403.90, 91.28);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (479.55,140.05) --
+	(479.55,140.05);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 91.28) --
-	(403.90, 28.63);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (479.55,140.05) --
+	(479.55, 71.12);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (403.90, 28.63) --
-	(403.90, 28.63);
-\definecolor{drawColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (479.55, 71.12) --
+	(479.55, 71.12);
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59,165.53) --
-	(395.59,165.53);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (472.90,210.87) --
+	(472.90,210.87);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59,165.53) --
-	(395.59, 61.03);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (472.90,210.87) --
+	(472.90,108.80);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (395.59, 61.03) --
-	(395.59, 61.03);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (472.90,108.80) --
+	(472.90,108.80);
+\definecolor{drawColor}{RGB}{170,144,0}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (466.25,138.38) --
+	(466.25,138.38);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (466.25,138.38) --
+	(466.25, 59.37);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (466.25, 59.37) --
+	(466.25, 59.37);
 \definecolor{drawColor}{RGB}{225,106,134}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 96.20) --
-	(387.28, 96.20);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (459.60,180.55) --
+	(459.60,180.55);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 96.20) --
-	(387.28, 29.36);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (459.60,180.55) --
+	(459.60,103.48);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (387.28, 29.36) --
-	(387.28, 29.36);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (459.60,103.48) --
+	(459.60,103.48);
+\definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72,148.26) --
-	(478.72,148.26);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (452.95,180.21) --
+	(452.95,180.21);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72,148.26) --
-	(478.72, 74.69);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (452.95,180.21) --
+	(452.95,105.54);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (478.72, 74.69) --
-	(478.72, 74.69);
-\definecolor{drawColor}{RGB}{0,173,154}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (452.95,105.54) --
+	(452.95,105.54);
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,223.85) --
-	(470.41,223.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (546.06, 64.57) --
+	(546.06, 64.57);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,223.85) --
-	(470.41,114.90);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (546.06, 64.57) --
+	(546.06, 44.61);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (470.41,114.90) --
-	(470.41,114.90);
-\definecolor{drawColor}{RGB}{144,152,0}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (546.06, 44.61) --
+	(546.06, 44.61);
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09,146.48) --
-	(462.09,146.48);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (539.41, 75.99) --
+	(539.41, 75.99);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09,146.48) --
-	(462.09, 62.15);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (539.41, 75.99) --
+	(539.41, 60.28);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (462.09, 62.15) --
-	(462.09, 62.15);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (539.41, 60.28) --
+	(539.41, 60.28);
+\definecolor{drawColor}{RGB}{170,144,0}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (532.76,222.92) --
+	(532.76,222.92);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (532.76,222.92) --
+	(532.76, 59.12);
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (532.76, 59.12) --
+	(532.76, 59.12);
 \definecolor{drawColor}{RGB}{225,106,134}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,191.49) --
-	(453.78,191.49);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (526.11, 78.38) --
+	(526.11, 78.38);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,191.49) --
-	(453.78,109.22);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (526.11, 78.38) --
+	(526.11, 50.78);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (453.78,109.22) --
-	(453.78,109.22);
-\definecolor{drawColor}{RGB}{145,131,230}
+\path[draw=drawColor,line width= 1.1pt,line join=round] (526.11, 50.78) --
+	(526.11, 50.78);
+\definecolor{drawColor}{RGB}{0,0,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 67.69) --
-	(545.23, 67.69);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (519.46, 78.62) --
+	(519.46, 78.62);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 67.69) --
-	(545.23, 46.39);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (519.46, 78.62) --
+	(519.46, 50.68);
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (545.23, 46.39) --
-	(545.23, 46.39);
-\definecolor{drawColor}{RGB}{0,173,154}
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 79.88) --
-	(536.91, 79.88);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 79.88) --
-	(536.91, 63.12);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (536.91, 63.12) --
-	(536.91, 63.12);
-\definecolor{drawColor}{RGB}{144,152,0}
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60,236.72) --
-	(528.60,236.72);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60,236.72) --
-	(528.60, 61.88);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (528.60, 61.88) --
-	(528.60, 61.88);
-\definecolor{drawColor}{RGB}{225,106,134}
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 82.44) --
-	(520.29, 82.44);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 82.44) --
-	(520.29, 52.97);
-
-\path[draw=drawColor,line width= 1.1pt,line join=round] (520.29, 52.97) --
-	(520.29, 52.97);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (519.46, 50.68) --
+	(519.46, 50.68);
 \definecolor{drawColor}{gray}{0.20}
 
-\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 18.22) rectangle (572.66,247.13);
+\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 18.22) rectangle (572.66,232.67);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 49.84) {2};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 47.65) {2};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,104.06) {4};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 98.45) {4};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,158.27) {6};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,149.24) {6};
 
-\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,212.49) {8};
+\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,200.03) {8};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 52.87) --
-	( 27.31, 52.87);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 50.68) --
+	( 27.31, 50.68);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,107.09) --
-	( 27.31,107.09);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,101.48) --
+	( 27.31,101.48);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,161.31) --
-	( 27.31,161.31);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,152.27) --
+	( 27.31,152.27);
 
-\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,215.52) --
-	( 27.31,215.52);
+\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,203.06) --
+	( 27.31,203.06);
 \end{scope}
@@
 
-\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,132.67) {Basic reproductive number};
+\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,125.45) {Basic reproductive number};
 \end{scope}
@@
 
-\path[fill=fillColor] ( 67.57,258.13) rectangle (532.41,283.58);
+\path[fill=fillColor] (135.69,243.67) rectangle (464.28,283.58);
 \end{scope}
@@
 
-\path[fill=fillColor] ( 78.57,263.63) rectangle ( 93.02,278.08);
+\path[fill=fillColor] (146.69,263.63) rectangle (161.14,278.08);
+\end{scope}
+\begin{scope}
+\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
+\definecolor{fillColor}{RGB}{0,0,0}
+
+\path[fill=fillColor] (153.92,270.85) circle (  3.57);
+\end{scope}
+\begin{scope}
+\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
+\definecolor{drawColor}{RGB}{0,0,0}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (148.13,270.85) -- (159.70,270.85);
+\end{scope}
+\begin{scope}
+\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
+\definecolor{fillColor}{RGB}{255,255,255}
+
+\path[fill=fillColor] (146.69,249.17) rectangle (161.14,263.63);
 \end{scope}
@@
 
-\path[fill=fillColor] ( 85.79,270.85) circle (  3.57);
+\path[fill=fillColor] (153.92,261.95) --
+	(158.72,253.62) --
+	(149.11,253.62) --
+	cycle;
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] ( 80.01,270.85) -- ( 91.57,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (148.13,256.40) -- (159.70,256.40);
 \end{scope}
@@
 
-\path[fill=fillColor] (120.69,263.63) rectangle (135.14,278.08);
+\path[fill=fillColor] (204.70,263.63) rectangle (219.15,278.08);
 \end{scope}
@@
 \path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
-\definecolor{fillColor}{RGB}{144,152,0}
+\definecolor{fillColor}{RGB}{170,144,0}
 
-\path[fill=fillColor] (127.91,276.40) --
-	(132.72,268.08) --
-	(123.11,268.08) --
+\path[fill=fillColor] (208.35,267.28) --
+	(215.49,267.28) --
+	(215.49,274.42) --
+	(208.35,274.42) --
 	cycle;
@@
 \path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
-\definecolor{drawColor}{RGB}{144,152,0}
+\definecolor{drawColor}{RGB}{170,144,0}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (122.13,270.85) -- (133.69,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (206.14,270.85) -- (217.70,270.85);
 \end{scope}
@@
 
-\path[fill=fillColor] (238.08,263.63) rectangle (252.53,278.08);
+\path[fill=fillColor] (204.70,249.17) rectangle (219.15,263.63);
 \end{scope}
@@
 \path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
-\definecolor{fillColor}{RGB}{0,173,154}
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[fill=fillColor] (241.74,267.28) --
-	(248.87,267.28) --
-	(248.87,274.42) --
-	(241.74,274.42) --
-	cycle;
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.88,256.40) -- (216.97,256.40);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.92,251.35) -- (211.92,261.45);
 \end{scope}
@@
 \path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
-\definecolor{drawColor}{RGB}{0,173,154}
+\definecolor{drawColor}{RGB}{0,170,90}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (239.52,270.85) -- (251.09,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (206.14,256.40) -- (217.70,256.40);
 \end{scope}
@@
 
-\path[fill=fillColor] (358.89,263.63) rectangle (373.35,278.08);
+\path[fill=fillColor] (325.51,263.63) rectangle (339.96,278.08);
 \end{scope}
@@
 \path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
-\definecolor{drawColor}{RGB}{145,131,230}
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.07,270.85) -- (371.17,270.85);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,267.28) rectangle (336.30,274.42);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.12,265.81) -- (366.12,275.90);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,267.28) -- (336.30,274.42);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,274.42) -- (336.30,267.28);
 \end{scope}
@@
 \path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
-\definecolor{drawColor}{RGB}{145,131,230}
+\definecolor{drawColor}{RGB}{0,166,202}
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (360.34,270.85) -- (371.90,270.85);
+\path[draw=drawColor,line width= 1.1pt,line join=round] (326.95,270.85) -- (338.52,270.85);
 \end{scope}
@@
 
-\path[fill=fillColor] (497.67,263.63) rectangle (512.12,278.08);
+\path[fill=fillColor] (325.51,249.17) rectangle (339.96,263.63);
+\end{scope}
+\begin{scope}
+\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
+\definecolor{drawColor}{RGB}{182,117,224}
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,252.83) -- (336.30,259.97);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.17,259.97) -- (336.30,252.83);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (327.69,256.40) -- (337.78,256.40);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (332.74,251.35) -- (332.74,261.45);
+\end{scope}
+\begin{scope}
+\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
+\definecolor{drawColor}{RGB}{182,117,224}
+
+\path[draw=drawColor,line width= 1.1pt,line join=round] (326.95,256.40) -- (338.52,256.40);
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,267.28) rectangle (508.46,274.42);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,267.28) -- (508.46,274.42);
-
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.33,274.42) -- (508.46,267.28);
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.64,267.82) {reported};
 \end{scope}
@@
 
-\path[draw=drawColor,line width= 1.1pt,line join=round] (499.11,270.85) -- (510.68,270.85);
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.64,253.37) {base};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 98.52,267.82) {base};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (224.65,267.82) {exponential growth rate};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (140.64,267.82) {exponential growth rate};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (224.65,253.37) {mean generation interval};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (258.03,267.82) {mean generation interval};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.46,267.82) {generation-interval dispersion};
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (378.85,267.82) {generation-interval dispersion};
-\end{scope}
-\begin{scope}
-\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
-\definecolor{drawColor}{RGB}{0,0,0}
-
-\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (517.62,267.82) {all};
+\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.46,253.37) {all};
 \end{scope}

### 2020-06-21 response2.tex
done

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}

\usepackage{color}

\usepackage{xspace}

\usepackage{lmodern}
\usepackage{amssymb,amsmath}

\usepackage[pdfencoding=auto, psdextra]{hyperref}

\usepackage{natbib}
\bibliographystyle{chicago}

\newcommand{\rR}{\mbox{$r$--$\cal R$}}
\newcommand{\RR}{\ensuremath{{\cal R}}}
\newcommand{\RRhat}{\ensuremath{{\hat \cal R}}}
\newcommand{\Rx}[1]{\ensuremath{{\cal R}_{#1}}} 
\newcommand{\Ro}{\ensuremath{{\mathcal R}_{0}}\xspace}
\newcommand{\Rpool}{\ensuremath{{\mathcal R}_{\textrm{\tiny{pool}}}}\xspace}
\newcommand{\Reff}{\Rx{\mathit{eff}}}
\newcommand{\Tc}{\ensuremath{C}}

\newcommand{\rev}{\subsection*}
\newcommand{\revtext}{\textsf}
\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0em}

\newcommand{\comment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\jd}[1]{\comment{cyan}{JD}{#1}}
\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\dc}[1]{\comment{blue}{DC}{#1}}
\newcommand{\jsw}[1]{\comment{green}{JSW}{#1}}
\newcommand{\hotcomment}[1]{\comment{red}{HOT}{#1}}

\begin{document}

\noindent Dear Editor:

Thank you for the chance to revise and resubmit our manuscript.
Below please find our responses to reviewers.

\rev{Reviewer \#1}

\revtext{Authors have properly addressed my comments and I believe that the manuscript is now suitable for publication}

Thank you.

\rev{Reviewer \#2}

\revtext{
The authors have made a good effort of explanation in the writing of the material and methods
part. They did not wish to carry out a DAG. Indeed, the Bayesian estimation part is relatively
simple and does not necessarily require a DAG. However, the whole process is complex and
to facilitate the understanding of the readers a conceptual framework of the whole study is
more than necessary.}

We have added the following text in the last paragraph of the introduction to provide a conceptual overview of the framework:

``Instead, we model the estimate of \Ro (as well as the associated generation-interval parameters, $\bar G$, and $\kappa$) from each study with probability distributions that account for the uncertainty in the estimates;
this allows us to re-estimate the corresponding distributions of the exponential growth rates $r$.
We then use a Bayesian multi-level model to calculate the average (pooled estimates) of three key quantities ($r$, $\bar G$, and $\kappa$).
The resulting pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) are used to calculate the pooled estimate of the basic reproduction number, \Rpool.
Calculating \Rpool based on the pooled estimates allows us to average appropriately across the uncertainties present in modeling approaches and in their underlying assumptions.''

\revtext{There is still a confusion between credible intervals and confidence
intervals. When the authors use a frequentist approach, they estimate a confidence interval;
while the authors use a bayesian approach they estimate a credible interval. So it is study,
they estimate credible intervals but they estimated intervals they used from the literature can
be both confidence and credible intervals. Therefore, it is case, use the terms
confidence/credible intervals.}

We tried to be as careful as possible in the uses of confidence and credible intervals.
When we refer to based estimates, however, we still use credible intervals because the base estimates are our approximations of the reported estimates (taken from each study); we have tried to make this distinction clear in the text.

\revtext{In this work, all the studies have all the same weight 1/7 when they calculated pooled estimates
cf. Equation 4. They did not give more weight to the studies, which were the more likely}

We now provide a better justification:

``Although this approach assumes that all studies are equally weighted, the overall pooled estimate will still be weighted by the certainty of the reported estimates (e.g., $r_i$ will be sampled from a narrow distribution and therefore have stronger influence on $\mu_r$ if the reported confidence/credible interval on $r_i$ is narrow).''

\revtext{More specific comments}

\revtext{The authors do not yet give information about the other sources of uncertainty linked
with the variation of parameter sigma ($\sigma_r$, $\sigma_G$, $\sigma_\kappa$).}

We have added the following sentences:

``The estimates of the between-study standard deviations further suggests that there is a large variability in the underlying parameters among the seven studies, particularly in $r$ and $\kappa$:
$\sigma_r = 0.07\,\textrm{days}^{-1}$ (95\% CI: 0.04--0.19 $\textrm{days}^{-1}$),
$\sigma_G = 1.02\,\textrm{days}$ (95\% CI: 0.54--2.50 days),
and
$\sigma_\kappa = 0.51$ (95\% CI: 0.24--1.52).
This variability is likely driven by the differences in modeling approaches and assumptions.''

\revtext{Keep the same order for exemple always $r$, $\bar G$ and $\kappa$.}

Done.

\revtext{
The new figure 2 is nice. Change the caption of the figure 2 “Effects of the uncertainties of $r$, $\bar G$, and $\kappa$ on the estimates of $\Ro$” by “Effects of growth rate ($r$), Mean generation interval ($\bar G$), and Generation-interval dispersion ($\kappa$) on the estimates of $\Ro$”. The caption must be explicit by itself.
}

Done.

\revtext{
The new figure 3 is nice but please represent on figure 3 the value of the R0
and its IC published in the 7 studies (Cf table1) for example in black.
}

Done.

\revtext{
Indicate the number of posterior samples used (2000? or 1000?).
}

Done.

\revtext{
When the authors speak about “credible intervals” from the literature, they must
replace every time by “confidence/credible” because it depends on the methods
of previous studies.
}

Done.

\revtext{
The mutation of the virus or the presence of a new clade of the virus having
different R0 can also affect estimation and inference.
}

We have added the following sentence in the discussion:

``Emergence of a new strain (or a clade) with different transmissibility can also affect disease dynamics as well as inference, although it is less likely to be of a concern in the beginning of an epidemic.''

\pagebreak

\bibliography{ncov_abbr.bib}

\end{document}

### 2020-06-21 compare_R0.R

@@ -3,6 +3,7 @@ library(gridExtra)
 library(colorspace)
 library(dplyr)
 source("study_number.R")
+source("study_estimate.R")
 
 load("sample_data.rda")
 
@@ -30,19 +31,20 @@ R0all <- Rdata_adj %>%
   ) %>%
   bind_rows(
     R0_r, R0_gbar, R0_kappa,
-    mutate(filter(R0_all, study=="Imai et al. (2020)"))
+    mutate(filter(R0_all, study=="Imai et al. (2020)")),
+    mutate(study_estimate, type="reported")
   ) %>%
   mutate(
-    type=factor(type, levels=c("base", "r", "gbar", "kappa", "all"),
-                labels=c("base", "exponential growth rate", "mean generation interval", "generation-interval dispersion", "all"))
+    type=factor(type, levels=c("reported", "base", "r", "gbar", "kappa", "all"),
+                labels=c("reported", "base", "exponential growth rate", "mean generation interval", "generation-interval dispersion", "all"))
   ) %>%
   merge(study_number) %>%
   mutate(
     anon=ifelse(type=="all", "Pooled estimate", as.character(anon))
   )
-  
+
 R0all %>%
-  filter(type!="all", type!="base") %>%
+  filter(type!="all", type!="base", type!="reported") %>%
   mutate(
     width=upr-lwr
   ) %>%
@@ -73,7 +75,7 @@ g1 <- ggplot(R0all) +
                 width=0, lwd=1) +
   scale_y_continuous("Basic reproductive number", breaks=c(2, 4, 6, 8, 10)) +
   ## scale_colour_discrete_qualitative() +
-  scale_colour_manual(values=c(colorspace::qualitative_hcl(4),"#000000")) +
+  scale_colour_manual(values=c("#000000", colorspace::qualitative_hcl(5))) +
   theme(
     axis.title.x = element_blank(),
     legend.position = "top",

### 2020-06-21 prior_pred.R

@@ -1,0 +1,11 @@
+set.seed(101)
+
+nsim <- 100000
+
+r <- rgamma(nsim, 2, 2*7)
+gbar <- rgamma(nsim, 2, 2/7) 
+kappa <- rgamma(nsim, 2, 2/0.5)
+
+R0 <- (1 + kappa * r * gbar)^(1/kappa)
+
+quantile(R0, c(0.025, 0.975))

### 2020-06-21 study_estimate.R

@@ -1,0 +1,7 @@
+study_estimate <- data.frame(
+  study=c("Bedford et al. (2020)", "Imai et al. (2020)", "Liu et al. (2020)", "Majumder and Mandl (2020)", "Read et al. (2020)",
+          "Riou and Althaus (2020)", "Zhao et al. (2020)"),
+  est=c(NA, 2.6, 2.92, NA, 3.8, 2.2, 5.47),
+  lwr=c(1.5, 1.5, 2.28, 2.0, 3.6, 1.4, 4.16),
+  upr=c(3.5, 3.5, 3.67, 3.1, 4.0, 3.8, 7.10)
+)

### 2020-06-21 summarize_MCMC.R

@@ -1,0 +1,17 @@
+library(coda)
+library(dplyr)
+library(tidyr)
+
+load("MCMC_all.rda")
+
+MCMC_all %>%
+  lapply(as.data.frame) %>%
+  bind_rows %>%
+  setNames(c("mu_r", "sigma_r", "mu_G", "sigma_G", "mu_kappa", "sigma_kappa")) %>%
+  gather(key, value) %>%
+  group_by(key) %>%
+  summarize(
+    est=median(value),
+    lwr=quantile(value, 0.025),
+    upr=quantile(value, 0.975)
+  )

### 2020-06-24 response2.tex
Suggestions for response letter (including some that need to move to MS).

@@
 \usepackage[pdfencoding=auto, psdextra]{hyperref}
-
-\usepackage{natbib}
-\bibliographystyle{chicago}
 
@@
 this allows us to re-estimate the corresponding distributions of the exponential growth rates $r$.
-We then use a Bayesian multi-level model to calculate the average (pooled estimates) of three key quantities ($r$, $\bar G$, and $\kappa$).
+We then use a Bayesian multi-level model to estimate the three  key quantities ($r$, $\bar G$, and $\kappa$).
 The resulting pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) are used to calculate the pooled estimate of the basic reproduction number, \Rpool.
-Calculating \Rpool based on the pooled estimates allows us to average appropriately across the uncertainties present in modeling approaches and in their underlying assumptions.''
+Using pooled estimates allows us to average appropriately across the uncertainties present in modeling approaches and in their underlying assumptions.''
 
@@
 We tried to be as careful as possible in the uses of confidence and credible intervals.
-When we refer to based estimates, however, we still use credible intervals because the base estimates are our approximations of the reported estimates (taken from each study); we have tried to make this distinction clear in the text.
+We still describe our base estimates with credible intervals because the base estimates are our Bayesian approximations based on the reported estimates taken from each study; we have tried to make this distinction clear in the text.
 
@@
 
-``Although this approach assumes that all studies are equally weighted, the overall pooled estimate will still be weighted by the certainty of the reported estimates (e.g., $r_i$ will be sampled from a narrow distribution and therefore have stronger influence on $\mu_r$ if the reported confidence/credible interval on $r_i$ is narrow).''
+``Although this approach nominally treats all studies are equally weighted, the overall pooled estimate will still be weighted by the certainty of the reported estimates (e.g., $r_i$ will be sampled from a narrow distribution and therefore have stronger influence on $\mu_r$ if the reported confidence/credible interval on $r_i$ is narrow).''
 
@@
 
-``Emergence of a new strain (or a clade) with different transmissibility can also affect disease dynamics as well as inference, although it is less likely to be of a concern in the beginning of an epidemic.''
-
-\pagebreak
-
-\bibliography{ncov_abbr.bib}
+``Emergence of a new strain (or a clade) with different transmissibility could also affect disease dynamics, and complicate inference; this study does not address this possibility.''
 

### 2020-06-25 ncov.tex
diff

@@
 
-To formalize the estimation of uncertainty at the onset of an outbreak, we present a statistical framework for averaging across estimates of basic reproductive number \Ro from multiple studies. 
+To formalize the estimation of uncertainty at the onset of an outbreak, we present a statistical framework for averaging across estimates of the basic reproductive number \Ro from multiple studies. 
 We apply the method to seven disparate models published online as pre-prints between January 23--26, 2020 that estimate \Ro for the SARS-CoV-2 outbreak in Wuhan City, China \citep{bedfordncov, imaincov, liuncov, majumderncov, readncov, riouncov, zhaoncov}.
 Previous studies have directly calculated the average of reported \Ro values \citep{majumder2020early, liu2020reproductive} but such methods mask differences in underlying model assumptions and statistical methods.
-Instead, we decompose estimated \Ro into three key quantities ($r$, $\bar G$, and $\kappa$) and calculate the average (pooled estimates) of these key parameters.
-Calculating \Ro based on the pooled estimates allows us to average appropriately across the uncertainties present in modeling approaches and in their underlying assumptions.
+Instead, we model the estimate of \Ro (as well as the associated generation-interval parameters, $\bar G$, and $\kappa$) from each study with probability distributions that account for the uncertainty in the estimates;
+this allows us to re-estimate the corresponding distributions of the exponential growth rates $r$.
+We then use a Bayesian multi-level model to average the three key quantities ($r$, $\bar G$, and $\kappa$).
+The resulting pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) are used to calculate the pooled estimate of the basic reproduction number, \Rpool.
+Using pooled estimates allows us to average appropriately across the uncertainties present in modeling approaches and in their underlying assumptions.
 We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval.
@@
 Generation-interval dispersion represent the squared coefficients of variation in generation intervals.
-$^\ast$These intervals reflect \Ro values for best and worst scenarios. We treat these intervals as a 90\% credible interval in our analysis.
+$^\ast$These intervals reflect \Ro values for best and worst scenarios. We treat these intervals as a 90\% confidence/credible interval in our analysis.
 $^\dagger$We assume $\kappa = 0.5$ in our analysis.
@@
 All model assumptions reduce to properties of the exponential growth rate $r$ and the shape of the generation-interval distribution $g(\tau)$.
-For example, if a model relies on overly confident assumptions about the underlying observation (how new cases are reported) or process (how new cases are generated) model, the estimated confidence or credible intervals associated with the exponential growth rates or parameters of the generation-interval distributions will necessarily be narrow.
+For example, if a model relies on overly confident assumptions about the underlying observation (how new cases are reported) or process (how new cases are generated) model, the estimated confidence/credible intervals  associated with the exponential growth rates or parameters of the generation-interval distributions (from each study) will necessarily be narrow.
 
@@
 
-While most studies report confidence or credible intervals to quantify uncertainties associated with their estimates, some use different measures.
+While most studies report confidence/credible intervals to quantify uncertainties associated with their estimates, some use different measures.
 In particular, Study 2 reports a range of \Ro for the worst and best case scenarios, which correspond to the values of \Ro such that 95\% and 5\% of the simulated total number of cases by January 18, 2020 are greater than or equal to 4000, respectively;
-for simplicity, we treat these intervals as a 90\% credible interval in our analysis.
+for simplicity, we treat these intervals as a 90\% confidence/credible interval in our analysis.
 Uncertainty ranges reported by Study 1 and Study 7 are assumed to be uniform ranges.
@@
 First, to account for uncertainties in these parameters, we model reported values of the basic reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$ with appropriate probability distributions.
-We use gamma distributions to model values reported with confidence or credible intervals (CI) and uniform distributions to model values reported with ranges;
-when confidence or credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence or credible limits is equal to the reported confidence or credible level.
+We use gamma distributions to model values reported with confidence/credible intervals (CI) and uniform distributions to model values reported with ranges;
+when confidence/credible intervals are reported, we parameterize the gamma distribution such that (i) its mean matches the estimated value and (ii) the probability that a random variable following the specified gamma distribution falls between the lower and upper confidence/credible limits is equal to the reported confidence/credible level.
 This probability does not need to be calculated from equi-tailed quantiles.
@@
 This allows us to approximate the probability distributions of the exponential growth rates estimated by each study.
-Uncertainties in the probability distributions that we calculate for the estimated exponential growth rates reflect model assumptions, statistical methods, as well as quality of the data that each study relies on.
+Uncertainties in the probability distributions that we calculate for the estimated exponential growth rates reflect model assumptions, statistical methods, as well as the quality of the data that each study relies on.
 This approach of re-estimating the exponential growth rate does not affect the uncertainty captured by our analysis because we are re-estimating the probability distribution of $r_i$ that is consistent with the reported values of $(\Ro)_i$, $\bar G_i$, and $\kappa_i$;
@@
 Here, we do so by averaging across reported values, without explicitly re-fitting their models.
-We use a Markov Chain Monte Carlo approach (cf.~Section \ref{section:MCMC}) and account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(\bar{G}_{i,m}, \kappa_{i,m}, r_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
+We use a Markov Chain Monte Carlo approach (cf.~Section \ref{section:MCMC}) and account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and correlation among them), by drawing a random set from the family of parameter sets $(r_{i,m}, \bar{G}_{i,m}, \kappa_{i,m})$ for each study $i$ at each Metropolis-Hastings step.
 Since the gamma distribution does not allow $\kappa=0$ (this corresponds to a Dirac delta generation-interval distribution), we substitute $\kappa=0.02$ for Study 7.
+Although this approach nominally treats all studies are equally weighted, the overall pooled estimate will still be weighted by the certainty of the reported estimates (e.g., $r_i$ will be sampled from a narrow distribution and therefore have stronger influence on $\mu_r$ if the reported confidence/credible interval on $r_i$ is narrow).
 
@@
 Since SARS-CoV-2 spread primarily in Wuhan City, China, during this period, it is not possible to include independent sources of data from other countries.
-Thus, the pooled estimator should be interpreted with care.
+Thus, the pooled estimates should be interpreted with care.
 
@@
 The assumed distribution in Study 6 in Table 2 neglects uncertainty in the mean generation interval $\bar G$, whereas the base estimates account for this uncertainty.
-Furthermore, since the distributions in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles differ slightly from the reported values.
+Furthermore, since the distributions in Table 2 are constructed by matching the mean and the probabilities associated with the reported uncertainty ranges, the exact values of the base estimates and their 95\% quantiles differ slightly from the reported values in Table 1.
 We compare the base estimates with a pooled estimate the basic reproductive number (\Rpool) based on the pooled estimates of underlying parameters (by substituting $\mu_r$, $\mu_G$, $\mu_\kappa$ in \eref{gamma}).
@@
 all studies assumed fixed values for $\kappa$, ranging from 0 to 1.
+The estimates of the between-study standard deviations further suggests that there is a large variability in the underlying parameters among the seven studies, particularly in $r$ and $\kappa$:
+$\sigma_r = 0.07\,\textrm{days}^{-1}$ (95\% CI: 0.04--0.19 $\textrm{days}^{-1}$),
+$\sigma_G = 1.02\,\textrm{days}$ (95\% CI: 0.54--2.50 days),
+and
+$\sigma_\kappa = 0.51$ (95\% CI: 0.24--1.52).
+This variability is likely driven by the differences in modeling approaches and assumptions.
 
@@
 We inferred point estimates (black), uniform distributions (orange) or confidence/credible intervals (purple) for each parameter from each study, and combined them into pooled estimates using a Bayesian multilevel model (red).
-Points represent medians calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$ (orange and purple).
-Error bars represent 95\% equi-tailed quantiles calculated from the parameter set $(\bar{G}_{i}, \kappa_{i}, r_{i})$ for each study $i$.
+Points represent medians calculated from the parameter set $(r_{i}, \bar{G}_{i}, \kappa_{i})$ for each study $i$ (orange and purple).
+Error bars represent 95\% equi-tailed quantiles calculated from the parameter set $(r_{i}, \bar{G}_{i}, \kappa_{i})$ for each study $i$.
 Red density plots represent distributions of 2000 posterior samples.
@@
 This hypothetical example reflects scenarios, in which increased data availability allows researchers to estimate $r$ with more certainty.
-To simulate estimates of the exponential growth rate with stronger confidence, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (\fref{eff}B); 
+To simulate estimates of the exponential growth rate with narrower uncertainty, we use $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$ (\fref{eff}B); 
 then $\hat{\mu}_r$ has the same median as $\mu_r$ but the associated 95\% CI is 4 times narrower (0.16--0.19 $\textrm{days}^{-1}$).
@@
 \caption{
-  \textbf{Effects of $r$, $\bar G$, and $\kappa$ on the estimates of \Ro.}
+  \textbf{Effects of the exponential growth rate $r$, mean generation interval $\bar G$, and generation-interval dispersion $\kappa$ on the estimates of the basic reproduction number \Ro.}
 We compare estimates of \Ro under nine scenarios that propagate different parameter uncertainties (A) based on our pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) and (B) assuming a 4-fold reduction in uncertainty of our pooled estimate of the exponential growth rate (using $\hat{\mu}_r = (\mu_r + 3\times\mathrm{median}(\mu_r))/4$ instead of $\mu_r$).
@@
 
-Finally, \fref{R0} compares the base estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) and 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
+Finally, \fref{R0} compares the reported estimates (Table 1) with the base estimates (based on $r_i$, $\bar G_i$, and $\kappa_i$ for each study $i$) as well as 21 substitute estimates (3 parameter substitutions $\times$ 7 studies).
+The base estimates, which are probability-based approximations of the reported estimates, are broadly consistent with the reported estimates.
 All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- the cases with more certain substitute estimates are the $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
@@
 This is consistent with our earlier results (\fref{eff}) that demonstrated the importance of accounting for uncertainty in the exponential growth rate $r$.
-We also note that the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates in all cases except for Study 6.
+In addition, the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates in all cases except for Study 6.
 
@@
 this corresponds to replacing all reported parameter values with our pooled estimates, which gives identical results across all studies.
-Points represent the medians of our base, substitute, and pooled estimates.
+The reported estimates refer to estimates listed in Table 1.
+Points represent the medians of the reported, base, substitute, and pooled estimates.
 Vertical error bars represent the 95\% credible intervals of our base, substitute, and pooled estimates.
 Horizontal dashed lines represent the 95\% credible intervals of our pooled estimate.
+2000 posterior samples are used for the analysis.
 }
@@
 This approach is appealing because of its simplicity and apparent robustness, but fitting a model to cumulative incidence neglects autocorrelation between successive counts of cumulative cases. 
-As a result, this approach both biases parameter estimates and gives overly narrow credible intervals \citep{ma2014estimating, king2015avoidable}.
+As a result, this approach both biases parameter estimates and gives overly narrow confidence/credible intervals \citep{ma2014estimating, king2015avoidable}.
 Narrow uncertainties in the estimates of the exponential growth rate are probably driven by this approach.
@@
 Neglecting one or the other is not always a serious problem, particularly if the goal is inferring parameters rather than directly making forecasts \citep{ma2014estimating}.
-Modelers should, however, be aware that oversimplifying the error model can give overly narrow credible intervals \citep{king2015avoidable,taylor2016stochasticity}.
+Modelers should, however, be aware that oversimplifying the error model can give overly narrow confidence/credible intervals \citep{king2015avoidable,taylor2016stochasticity}.
 
@@
 All of these factors, including fitting to cumulative curves or ignoring process errors, affect the estimation of the exponential growth rate (as well as the associated uncertainties), which in turn affects the estimation of the basic reproductive number.
+Emergence of a new strain (or a clade) with different transmissibility could also affect disease dynamics, and complicate inference; this study does not address this possibility.
 
@@
 However, as the number of susceptible individuals decreases or behavior changes in response to perception of the epidemic, the growth rate will decrease: estimates of $r$ used for \Ro should account for the possibility that $r$ is decreasing through time.
-Although our analysis applies strictly to the earliest stages of an epidemic, we expect certain lessons to hold more generally: credible intervals must combine as many sources of uncertainty as possible.
+Although our analysis applies strictly to the earliest stages of an epidemic, we expect certain lessons to hold more generally: confidence/credible intervals must combine as many sources of uncertainty as possible.
 In fact, as epidemics progress and more data becomes available, it is likely that inferences about exponential growth rate (and other epidemiological parameters) will generally become more precise; thus the risk of over-confidence (when uncertainty about the generation-interval distribution is neglected) will become greater.
@@
 SWP reviewed the published literature.
-SWP performed the analysis.
+SWP performed the analysis, with contributions from all authors.
 SWP, BMB, and JD created the figures. 

### 2020-06-25 ncov.tex
A few changes and comments

@@
 The base estimates, which are probability-based approximations of the reported estimates, are broadly consistent with the reported estimates.
-All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- the cases with more certain substitute estimates are the $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
+\jd{Only 7 exceptions are listed.} All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- the cases with more certain substitute estimates are the $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
 Accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
@@
 This is consistent with our earlier results (\fref{eff}) that demonstrated the importance of accounting for uncertainty in the exponential growth rate $r$.
-In addition, the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates in all cases except for Study 6.
+In addition, the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates for all studies except for Study 6.
+\jd{I feel like this is something I should have noticed sooner.
+It makes most sense to plot and compare these values on a log scale, which would change this answer (6 would be smaller now, but I don't know about some of the others).
+I'm willing to let this go, but just want to feel annoyed at myself for a moment.}
 
@@
 Points represent the medians of the reported, base, substitute, and pooled estimates.
-Vertical error bars represent the 95\% credible intervals of our base, substitute, and pooled estimates.
+Vertical error bars represent the 95\% credible intervals of our base, substitute, and pooled estimates (based on 2000 posterior samples).
 Horizontal dashed lines represent the 95\% credible intervals of our pooled estimate.
-2000 posterior samples are used for the analysis.
+
 }
@@
 All of these factors, including fitting to cumulative curves or ignoring process errors, affect the estimation of the exponential growth rate (as well as the associated uncertainties), which in turn affects the estimation of the basic reproductive number.
-Emergence of a new strain (or a clade) with different transmissibility could also affect disease dynamics, and complicate inference; this study does not address this possibility.
+Emergence of a new strain with different transmissibility could also affect disease dynamics, and complicate inference; this study does not address this possibility.
 

### 2020-06-25 ncov.tex
done

@@
 The base estimates, which are probability-based approximations of the reported estimates, are broadly consistent with the reported estimates.
-\jd{Only 7 exceptions are listed.} All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- the cases with more certain substitute estimates are the $\bar G$-substitute estimates for Study 1 and 7, $r$-substitute estimates for Study 1 and 2, and  $\kappa$-substitute estimates for Study 3, 6, and 7.
+All but 8 substitute estimates have wider credible intervals compared to their corresponding base estimates --- the cases with more certain substitute estimates are the $\bar G$-substitute estimates for Studies 1, 5, and 7, $r$-substitute estimates for Studies 1 and 2, and  $\kappa$-substitute estimates for Studies 3, 6, and 7.
 Accounting for uncertainties in the estimate of $r$ has the largest effect on the estimates of \Ro\ in most cases (\fref{R0}).
@@
 In addition, the pooled estimate of the basic reproductive number ($\Rpool = 3.0$; 95\% CI: 2.1--4.6) has wider credible intervals than the base estimates for all studies except for Study 6.
-\jd{I feel like this is something I should have noticed sooner.
-It makes most sense to plot and compare these values on a log scale, which would change this answer (6 would be smaller now, but I don't know about some of the others).
-I'm willing to let this go, but just want to feel annoyed at myself for a moment.}
 
@@
 Horizontal dashed lines represent the 95\% credible intervals of our pooled estimate.
-
 }

### 2020-06-25 response2.tex
diff
done

@@
 
-Thank you.
+Thank you for your review.
 
@@
 this allows us to re-estimate the corresponding distributions of the exponential growth rates $r$.
-We then use a Bayesian multi-level model to estimate the three  key quantities ($r$, $\bar G$, and $\kappa$).
+We then use a Bayesian multi-level model to average the three key quantities ($r$, $\bar G$, and $\kappa$).
 The resulting pooled estimates ($\mu_r$, $\mu_G$, and $\mu_\kappa$) are used to calculate the pooled estimate of the basic reproduction number, \Rpool.
@@
 
-``Emergence of a new strain (or a clade) with different transmissibility could also affect disease dynamics, and complicate inference; this study does not address this possibility.''
+``Emergence of a new strain with different transmissibility could also affect disease dynamics, and complicate inference; this study does not address this possibility.''
 

### 2021-09-10 figure2.tex
home from trip
Installing on new alishan

@@
-% Created by tikzDevice version 0.12.3 on 2020-05-07 17:13:04
+% Created by tikzDevice version 0.12.3.1 on 2021-09-10 14:27:21
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.38,126.52) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.38,126.52) circle (  5.71);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (165.38,126.52) circle (  5.71);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (165.38,126.52) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.38,126.52) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (209.88,126.52) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (209.88,123.50) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.38,123.50) circle (  5.71);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,126.52) --
+	(254.38,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
-	( 76.38,126.52);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,126.52) --
+	(254.38,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
-	( 76.38,126.52);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,126.52) --
+	(254.38,126.52);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,126.52) --
-	( 76.38,126.52);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,228.77) --
+	(165.38,228.77);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,228.77) --
-	(120.88,228.77);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,228.77) --
+	(165.38, 65.26);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,228.77) --
-	(120.88, 65.26);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38, 65.26) --
+	(165.38, 65.26);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88, 65.26) --
-	(120.88, 65.26);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,154.56) --
+	(120.88,154.56);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,154.56) --
-	(165.38,154.56);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,154.56) --
+	(120.88,104.90);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,154.56) --
-	(165.38,104.90);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (120.88,104.90) --
+	(120.88,104.90);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (165.38,104.90) --
-	(165.38,104.90);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,164.33) --
+	( 76.38,164.33);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,164.33) --
-	(209.88,164.33);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38,164.33) --
+	( 76.38, 76.82);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,164.33) --
-	(209.88, 76.82);
+\path[draw=drawColor,line width= 2.3pt,line join=round] ( 76.38, 76.82) --
+	( 76.38, 76.82);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88, 76.82) --
-	(209.88, 76.82);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,252.55) --
+	(209.88,252.55);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,252.55) --
-	(254.38,252.55);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88,252.55) --
+	(209.88, 51.71);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38,252.55) --
-	(254.38, 51.71);
-
-\path[draw=drawColor,line width= 2.3pt,line join=round] (254.38, 51.71) --
-	(254.38, 51.71);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (209.88, 51.71) --
+	(209.88, 51.71);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 76.38, 28.62) {none};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 76.38, 28.62) {$\mu_\kappa$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (120.88, 28.62) {$\mu_r$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (120.88, 28.62) {$\mu_G$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (165.38, 28.62) {$\mu_G$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (165.38, 28.62) {$\mu_r$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (209.88, 28.62) {$\mu_\kappa$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (209.88, 28.62) {all};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (254.38, 28.62) {all};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (254.38, 28.62) {none};
 \end{scope}

### 2021-09-10 propagate_pix.tex
home from trip
good night
Installing on new alishan

@@
-% Created by tikzDevice version 0.12.3 on 2020-05-07 18:04:06
+% Created by tikzDevice version 0.12.3.1 on 2021-09-10 14:27:30
 % !TEX encoding = UTF-8 Unicode
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,210.47) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,210.47) circle (  5.71);
+
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,210.47) circle (  5.71);
 
@@
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.79,210.47) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.40,210.47) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,210.47) circle (  5.71);
+\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.98,204.36) circle (  5.71);
 
-\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (446.17,204.36) circle (  5.71);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,210.47) --
+	(446.17,210.47);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
-	(101.40,210.47);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,210.47) --
+	(446.17,210.47);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
-	(101.40,210.47);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,210.47) --
+	(446.17,210.47);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,210.47) --
-	(101.40,210.47);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,417.54) --
+	(273.79,417.54);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,417.54) --
-	(187.59,417.54);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,417.54) --
+	(273.79, 86.40);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,417.54) --
-	(187.59, 86.40);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79, 86.40) --
+	(273.79, 86.40);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59, 86.40) --
-	(187.59, 86.40);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,267.25) --
+	(187.59,267.25);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
-	(273.79,267.25);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,267.25) --
+	(187.59,166.69);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,267.25) --
-	(273.79,166.69);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (187.59,166.69) --
+	(187.59,166.69);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (273.79,166.69) --
-	(273.79,166.69);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,287.03) --
+	(101.40,287.03);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
-	(359.98,287.03);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,287.03) --
+	(101.40,109.81);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,287.03) --
-	(359.98,109.81);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (101.40,109.81) --
+	(101.40,109.81);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,109.81) --
-	(359.98,109.81);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,465.70) --
+	(359.98,465.70);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,465.70) --
-	(446.17,465.70);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98,465.70) --
+	(359.98, 58.97);
 
-\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17,465.70) --
-	(446.17, 58.97);
-
-\path[draw=drawColor,line width= 2.3pt,line join=round] (446.17, 58.97) --
-	(446.17, 58.97);
+\path[draw=drawColor,line width= 2.3pt,line join=round] (359.98, 58.97) --
+	(359.98, 58.97);
 \end{scope}
@@
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (101.40, 28.62) {none};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (101.40, 28.62) {$\mu_\kappa$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (187.59, 28.62) {$\mu_r$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (187.59, 28.62) {$\mu_G$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (273.79, 28.62) {$\mu_G$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (273.79, 28.62) {$\mu_r$};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (359.98, 28.62) {$\mu_\kappa$};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (359.98, 28.62) {all};
 
-\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (446.17, 28.62) {all};
+\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (446.17, 28.62) {none};
 \end{scope}
